### 2023-04-11 writeups/results.qmd
quarto model summary

---
title: "Results"
format:
  html:
    theme: flatly
    toc: true
---

```{r}
#| echo: false
############################ Parameters to set ############################
model_name <- "sv_user_guide_reparameterised_ksc_priors_ksc_phi_0.97779_sig_0.1585_beta_0.64733_"
pp_flag <- "posterior" # prior = no likelihood estimation, posterior = likelihood estimation

# Data location
data_loc <- "simulated"
data_type <- "ksc"
file_name <- "phi_0.97779_sig_0.1585_beta_0.64733"
############################ Parameters to set ############################
```

```{r}
#| echo: false
#| warnings: false
#| messages: false
#| output: false
library(tidyverse)
library(shinystan)
library(posterior)

source(here::here("R", "plots.R"))
source(here::here("R", "model_eval.R"))

path <- here::here("output", model_name)
rds_path <- list.files(path = here::here("output", model_name), full.names = TRUE, pattern = "*.RDS")
csv_path <- list.files(path = here::here("output", model_name), full.names = TRUE, pattern = "*.csv")

# Load model results
model_fit <- readRDS(rds_path[grep(paste(pp_flag, "fit", sep = ""), rds_path)])

# Load training data
data <- read.csv(here::here("data", data_loc, data_type, paste(file_name, ".csv", sep = "")))
dependent_variable <- "yobs" # log squared returns
returns <- data[complete.cases(data[dependent_variable]), dependent_variable]

# Data for predictive checks
y_rep_df_sample <- mcmc_output_df(model_obj = model_fit, variable = "y_rep", sample_n = 10)
log_y_squared_avg <- average_over_draws(model_obj = model_fit, variable = "log_y_squared")
auto_corr <- mcmc_output_df(model_obj = model_fit, variable = "log_y_squared_autocorr", sample_n = NULL)
kurtosis <- mcmc_output_df(model_obj = model_fit, variable = "y_rep_kurtosis", sample_n = NULL)
skewness <- mcmc_output_df(model_obj = model_fit, variable = "y_rep_skewness", sample_n = NULL)
true_returns_df <- returns_df(returns)
true_log_sq_returns <- log_squared_returns(returns)
```

# Model summary

## Model code

```{r}
#| echo: false
cat(model_fit$code(), sep = "\n")
```

## Predictive checks

### y_t
```{r}
#| echo: false
plot_predictive_check(
  dataframe = y_rep_df_sample,
  x_axis = time,
  y_axis = y_rep,
  groups = mcmc_draw,
  prior_post = pp_flag,
  save = FALSE,
  path = NULL,
  true_data = true_returns_df
)
```

### Distribution of log(y*^2)
```{r}
#| echo: false
plot_log_y_sqd_hist(
  mcmc_data = log_y_squared_avg,
  true_data = true_log_sq_returns,
  mcmc_x_axis = average,
  true_x_axis = log_y_squared,
  prior_post = pp_flag,
  save = FALSE,
  path = fit_location
)
```

```{r}
#| echo: false
plot_log_y_sqd_kde(
  mcmc_data = log_y_squared_avg,
  true_data = true_log_sq_returns,
  mcmc_x_axis = average,
  true_x_axis = log_y_squared,
  prior_post = pp_flag,
  save = FALSE,
  path = fit_location
)
```

### Autocorrelation log(Y^2)

```{r}
#| echo: false
plot_hist(
  data = auto_corr,
  x_axis = log_y_squared_autocorr,
  variable_name = "log(Y^2) autocorrelation",
  prior_post = pp_flag,
  save = FALSE,
  path = fit_location
)
```

### Kurtosis y_t

```{r}
#| echo: false
plot_hist(
  data = kurtosis,
  x_axis = y_rep_kurtosis,
  variable_name = "y_t kurtosis",
  prior_post = pp_flag,
  save = FALSE,
  path = fit_location
)
```

### Skewness y_t
```{r}
#| echo: false
plot_hist(
  data = skewness,
  x_axis = y_rep_skewness,
  variable_name = "y_t skewness",
  prior_post = pp_flag,
  save = FALSE,
  path = fit_location
)
```

### Checking phi and p (beta prior)

```{r}
#| echo: false
plot_hist(
  data = mcmc_output_df(model_obj = model_fit, variable = "phi"),
  x_axis = phi,
  variable_name = "phi",
  prior_post = pp_flag,
  save = FALSE,
  path = fit_location
)
```

```{r}
#| echo: false
plot_hist(
  data = mcmc_output_df(model_obj = model_fit, variable = "p"),
  x_axis = p,
  variable_name = "p",
  prior_post = pp_flag,
  save = FALSE,
  path = fit_location
)

```

### 2023-04-16 scripts/run_model.r
phi plots

@@ -48,7 +48,7 @@ for (prior in c(1, 0)) {
     chains = 4,
     parallel_chains = 4,
     refresh = 500,
-    adapt_delta = 0.9
+    adapt_delta = 0.8
   )
 
   # Save model
@@ -148,3 +148,21 @@ for (prior in c(1, 0)) {
     path = fit_location
   )
 }
+
+# plot_hist(
+#     data = mcmc_output_df(model_obj = model_fit, variable = "phi"),
+#     x_axis = phi,
+#     variable_name = "phi",
+#     prior_post = 'prior',
+#     save = FALSE,
+#     path = fit_location
+#   )
+
+# plot_hist(
+#     data = mcmc_output_df(model_obj = model_fit, variable = "p"),
+#     x_axis = p,
+#     variable_name = "p",
+#     prior_post = 'prior',
+#     save = FALSE,
+#     path = fit_location
+#   )

### 2023-04-16 config.r
added config and make files

@@ -1,0 +1,29 @@
+config = TRUE
+
+# Create Yahoo Data
+
+# Create Simulated Data
+
+# Model training
+
+## Model metadata
+model_name <- "sv_user_guide_reparameterised_ksc_priors"
+unique_identifier <- "adapt_delta_0.9"
+dependent_variable <- "yobs"
+
+## Data location
+data_loc <- "simulated"
+data_type <- "ksc"
+data_file_name <- "phi_0.97779_sig_0.1585_beta_0.64733"
+
+## Stan sampling
+seed <- 123
+chains <- 4
+parallel_chains <- 4
+refresh <- 500
+adapt_delta <- 0.9
+save_warmup <- TRUE
+
+# Evaluation
+output_name <- paste(model_name, data_type, data_file_name, unique_identifier, sep = "_")
+pp_flag <- "posterior"

### 2023-04-16 writeups/results.qmd
put code at bottom
added summary
updated config options

@@
 #| echo: false
+source(here::here("config.r"))
+
 ############################ Parameters to set ############################
-model_name <- "sv_user_guide_reparameterised_ksc_priors_ksc_phi_0.97779_sig_0.1585_beta_0.64733_"
+if (!(config)) {
+output_name <- "sv_user_guide_reparameterised_ksc_priors_ksc_phi_0.97779_sig_0.1585_beta_0.64733_"
 pp_flag <- "posterior" # prior = no likelihood estimation, posterior = likelihood estimation
@@
 data_type <- "ksc"
-file_name <- "phi_0.97779_sig_0.1585_beta_0.64733"
+data_file_name <- "phi_0.97779_sig_0.1585_beta_0.64733"
+}
 ############################ Parameters to set ############################
@@
 
-path <- here::here("output", model_name)
-rds_path <- list.files(path = here::here("output", model_name), full.names = TRUE, pattern = "*.RDS")
-csv_path <- list.files(path = here::here("output", model_name), full.names = TRUE, pattern = "*.csv")
+path <- here::here("output", output_name)
+rds_path <- list.files(path = here::here("output", output_name), full.names = TRUE, pattern = "*.RDS")
+csv_path <- list.files(path = here::here("output", output_name), full.names = TRUE, pattern = "*.csv")
 
@@
 # Load training data
-data <- read.csv(here::here("data", data_loc, data_type, paste(file_name, ".csv", sep = "")))
+data <- read.csv(here::here("data", data_loc, data_type, paste(data_file_name, ".csv", sep = "")))
 dependent_variable <- "yobs" # log squared returns
@@
 
+<!-- ```{r}
+model_fit$summary()
+``` -->
+
 # Model summary
-
-## Model code
 
@@
 #| echo: false
-cat(model_fit$code(), sep = "\n")
+#| output: 'asis'
+paste(pp_flag, output_name, sep = " ")
 ```
@@
 ```
+
+# Model code
+
+```{r}
+#| echo: false
+cat(model_fit$code(), sep = "\n")
+```
+
+# Model sampling time
+
+```{r}
+# Time to sample
+model_fit$time()
+```
+
+# Model console messages
+
+```{r}
+# Console messages
+model_fit$output()
+```

### 2023-04-16 scripts/eval.r

@@ -2,24 +2,27 @@ library(tidyverse)
 library(shinystan)
 library(posterior)
 
+source(here::here("config.r"))
 source(here::here("R", "plots.R"))
 source(here::here("R", "model_eval.R"))
 
 ############################ Parameters to set ############################
 
-model_name <- "sv_user_guide_reparameterised_ksc_priors_ksc_phi_0.97779_sig_0.1585_beta_0.64733_beta_0.5_0.5"
-pp_flag <- "posterior" # prior = no likelihood estimation, posterior = likelihood estimation
+if (!(config)) {
+  output_name <- "sv_user_guide_reparameterised_ksc_priors_ksc_phi_0.97779_sig_0.1585_beta_0.64733_adapt_delta_0.8"
+  pp_flag <- "posterior" # prior = no likelihood estimation, posterior = likelihood estimation
 
 
-# Data location
-data_loc <- "simulated"
-data_type <- "ksc"
-file_name <- "phi_0.97779_sig_0.1585_beta_0.64733"
+  # Data location
+  data_loc <- "simulated"
+  data_type <- "ksc"
+  data_file_name <- "phi_0.97779_sig_0.1585_beta_0.64733"
+}
 ############################ Parameters to set ############################
 
-path <- here::here("output", model_name)
-rds_path <- list.files(path = here::here("output", model_name), full.names = TRUE, pattern = "*.RDS")
-csv_path <- list.files(path = here::here("output", model_name), full.names = TRUE, pattern = "*.csv")
+path <- here::here("output", output_name)
+rds_path <- list.files(path = here::here("output", output_name), full.names = TRUE, pattern = "*.RDS")
+csv_path <- list.files(path = here::here("output", output_name), full.names = TRUE, pattern = "*.csv")
 
 model_fit <- readRDS(rds_path[grep(paste(pp_flag, "fit", sep = ""), rds_path)])
 
@@ -28,6 +31,20 @@ model_fit <- readRDS(rds_path[grep(paste(pp_flag, "fit", sep = ""), rds_path)])
 
 # Contains ess and rhats (new)
 model_fit$summary()
+
+# Warm up samples
+model_fit$draws(variable = "sigma",
+  inc_warmup = TRUE,
+  format = "draws_df") %>%
+  pivot_wider(id_cols = ".iteration",
+    names_from = ".chain",
+    values_from = "sigma") %>% View # %>% filter(.iteration >= 1000)
+
+# Time to sample
[…]

### 2023-04-16 scripts/run_model.r

@@ -1,20 +1,33 @@
 library(tidyverse)
 library(cmdstanr)
+source(here::here("config.r"))
 source(here::here("R", "plots.R"))
 source(here::here("R", "model_eval.R"))
 
 set.seed(321321)
 
 ############################ Parameters to set ############################
-# Model description
-model_name <- "sv_user_guide_reparameterised_ksc_priors"
-dependent_variable <- "yobs"
-unique_identifier <- ""
 
-# Data location
-data_loc <- "simulated"
-data_type <- "ksc"
-file_name <- "phi_0.97779_sig_0.1585_beta_0.64733"
+if (!(config)) {
+  # Model description
+  model_name <- NULL
+  dependent_variable <- NULL
+  unique_identifier <- NULL
+
+  # Data location
+  data_loc <- NULL
+  data_type <- NULL
+  data_file_name <- NULL
+
+  # Stan paramaters
+  seed = 123
+  chains = 4
+  parallel_chains = 4
+  refresh = 500
+  adapt_delta = 0.8
+  save_warmup <- TRUE
+}
+
 ############################ Parameters to set ############################
 
 executables_path <- here::here("model_executables")
@@ -37,22 +50,23 @@ for (prior in c(1, 0)) {
   mod <- cmdstan_model(file, include_paths = here::here("models", "functions"), dir = executables_path)
 
   # Get data
-  data <- read.csv(here::here("data", data_loc, data_type, paste(file_name, ".csv", sep = "")))
+  data <- read.csv(here::here("data", data_loc, data_type, paste(data_file_name, ".csv", sep = "")))
   returns <- data[complete.cases(data[dependent_variable]), dependent_variable]
 
   # Fit model
   data_list <- list(T = length(returns), y = returns, sample_prior = sample_prior)
   model_fit <- mod$sample(
     data = data_list,
-    seed = 123,
-    chains = 4,
-    parallel_chains = 4,
-    refresh = 500,
-    adapt_delta = 0.8
+    seed = seed,
+    chains = chains,
+    parallel_chains = parallel_chains,
+    refresh = refresh,
+    adapt_delta = adapt_delta,
+    save_warmup = save_warmup
   )
 
   # Save model
-  fit_location <- paste(model_name, "_", data_type, "_", file_name, "_", unique_identifier, sep = "") # "_", format(Sys.time(), "%Y%m%d%H%M%S"), sep = "")
[…]

### 2023-04-16 config.r
updated data generation config and make files

@@ -1,11 +1,18 @@
 config = TRUE
 
 # Create Yahoo Data
+stock_index <- "GSPC"
+start_date <- "2019-01-04"
+end_date <- "2023-02-01"
 
 # Create Simulated Data
+set.seed(323651)
+size <- 1000
+phi <- 0.97779
+sig <- 0.15850
+beta <- 0.64733
 
 # Model training
-
 ## Model metadata
 model_name <- "sv_user_guide_reparameterised_ksc_priors"
 unique_identifier <- "adapt_delta_0.9"

### 2023-04-16 scripts/create_yahoo_data.r

@@ -1,10 +1,14 @@
 # Load data functions
 source(here::here("R", "data.R"))
+source(here::here("config.r"))
 
 # (Download) and load data
-stock_index <- "GSPC"
-start_date <- "2019-01-04"
-end_date <- "2023-02-01"
+if (!(config)) {
+    stock_index <- "GSPC"
+    start_date <- "2019-01-04"
+    end_date <- "2023-02-01"
+}
+
 raw_data_loc <- fin_data_location(stock_index, "raw", start_date, end_date)
 preprocessed_data_loc <- fin_data_location(stock_index, "preprocessed", start_date, end_date)
 

### 2023-08-18 scripts/sim_ksc_results.r
diagnostics complete

@@ -2,11 +2,14 @@ library(jsonlite)
 library(dplyr)
 library(tidyr)
 library(arrow)
+library(posterior)
+library(future.apply)
+plan(multisession)
 
 config <- jsonlite::read_json("configs/ksc.json")
+tmp_data_location <- paste("simulation_output/", config$simulation_name, "/tmp", sep ="") 
 
-data_location <- paste("simulation_output/", config$simulation_name, "/tmp", sep ="") 
-file_metadata <- as.data.frame(list.files(data_location))
+file_metadata <- as.data.frame(list.files(tmp_data_location))
 names(file_metadata) <- c("files")
 
 file_metadata <- separate_wider_delim(file_metadata, 
@@ -17,11 +20,92 @@ file_metadata <- separate_wider_delim(file_metadata,
     separate_wider_delim(., cols = mcmc_seed, delim = ".", names = c("mcmc_seed", "parquet")) %>% 
     select(-parquet)
 
-file_metadata[file_metadata["dataset"] == 1, ][["files"]]
+simulation_results <- function(file_number){
+    results <- list()
+    results[["all_chains"]] <- list()
+    results[["one_chain"]] <- list()
 
-interation_list <- list()
-iter_files <- file_metadata[file_metadata["dataset"] == 1, ][["files"]]
-iter_files_path <- paste(data_location, "/", iter_files, sep = "")
-interation_list <- lapply(iter_files_path, arrow::read_parquet)
+    iteration_list <- list()
+    iter_files <- file_metadata[file_metadata["dataset"] == file_number, ][["files"]]
+    iter_files_path <- paste(tmp_data_location, "/", iter_files, sep = "")
+    iteration_list <- lapply(iter_files_path, arrow::read_parquet, as_data_frame=TRUE)
+    parameter_names <- names(iteration_list[[1]])
 
-arrow::read_parquet(file = data_location)
+    combine_chains <- function(parameter, list_obj){
+        parameter_df = as.data.frame(cbind(
+        list_obj[[1]][parameter],
+        list_obj[[2]][parameter],
+        list_obj[[3]][parameter],
+        list_obj[[4]][parameter]
+            )
+        )
+
+        names(parameter_df) <- c("chain_1", "chain_2", "chain_3", "chain_4")
+
+        return(parameter_df)
+    }
+
+
+    all_chains = sapply(parameter_names, combine_chains, list_obj = iteration_list, simplify=FALSE)
[…]

### 2023-08-19 sim_ksc.py
async parallelisation

@@ -1,0 +1,126 @@
+from __future__ import division
+import os
+import json
+import random
+
+import numpy as np
+import pandas as pd
+import pyarrow as pa
+import pyarrow.parquet as pq
+# from joblib import Parallel, delayed
+import multiprocessing as mp
+np.set_printoptions(suppress=True, precision=5)
+
+from src.gaussianMixture import *
+
+nprocs = 25
+
+# Load config and create output directories
+config_path = "configs/ksc.json"
+data_path = "data/simulated/sbc"
+
+with open(config_path) as f:
+    config = json.load(f)
+
+# Create output directories
+output_dir = "simulation_output/" + config["simulation_name"] 
+os.mkdir(output_dir)
+os.mkdir(output_dir + "/output")
+os.mkdir(output_dir + "/tmp")
+os.mkdir(output_dir + "/std_out")
+
+# Save config
+with open(output_dir + "/config.json", "w") as outfile:
+    json.dump(config, outfile, indent=4)
+
+# Set seed and configure seed, data location and chains
+random.seed(config["seed"])
+
+sample_vect = random.sample(list(range(1, int(1e6))), config["n_iterations"] * config["chains"])
+data_file_indexes = [i for i in range(1, config["n_iterations"]+1) for _ in range(config["chains"])]
+seed_data_files = list(zip(sample_vect, data_file_indexes))
+
+
+def simulation(seed_data):
+    random.seed(seed_data[0])
+    data_file = data_path + "/" + config["data_location"] + f"/{seed_data[1]}.json"
+
+    with open(data_file) as f:
+        prior_params = json.load(f)
+
+    endog = pd.DataFrame(dict(y_obs = prior_params["y_sim"]))["y_obs"]
+
+    # Setup the model and simulation smoother
+    mod = TVLLDT(endog)
+    mod.set_smoother_output(0, smoother_state=True) # Calculate and return the smoothed states. Other options like cov matrices available
+    sim = mod.simulation_smoother()
+
+    # Storage for traces
+    trace_smoothed = np.zeros((config["mcmc_samples"] + 1, mod.nobs))
+    trace_states = np.zeros((config["mcmc_samples"] + 1, mod.nobs))
+    trace_mixing = np.zeros((config["mcmc_samples"] + 1, mod.nobs), int)
+    trace_mu = np.zeros((config["mcmc_samples"] + 1, 1))
[…]

### 2023-08-19 scripts/sbc_data_gen.r
ncp dgf

@@ -1,12 +1,9 @@
-model_name <- "sbc_data_gen_sv_ncp_ksc_priors"
+model_name <- "sbc_data_gen_sv_cp_ksc_priors_1000_datasets"
 n_time_points <- 1000
 n_datasets <- 1000
 data_path <- here::here("data", "simulated", "sbc", model_name)
 seed <- 543
-
-executables_path <- here::here("models/executables")
-if (!dir.exists(executables_path)) dir.create(executables_path)
-if (!dir.exists(data_path)) dir.create(data_path, recursive = TRUE)
+parameterisation <- "cp"
 
 gen_ncp_sv_dataset <- function(iter, seed_list, T){
     set.seed(seed_list[iter])
@@ -46,9 +43,56 @@ gen_ncp_sv_dataset <- function(iter, seed_list, T){
     
 }
 
+gen_cp_sv_dataset <- function(iter, seed_list, T){
+    set.seed(seed_list[iter])
+    sim_results <- list()
+
+    # Priors
+    mu <- rnorm(1, 0, sqrt(10))
+    sigma_sqd <- 1 / rgamma(1, 5/2, (0.01*5)/2)
+    sigma <- sqrt(sigma_sqd)
+    p <- rbeta(1, 20, 1.5)
+    phi <- 2*p - 1
+
+    # Generate log volatilities
+    h <- rep(0, T)
+    h[1] <- rnorm(1, mean=mu, sd=sigma/sqrt(1-phi^2))
+    for(i in 2:T){
+        h[i] <- rnorm(1, mean=(mu+phi*(h[(i-1)]-mu)), sd=sigma)
+      }
+
+    # Generate data from prior
+    y_sim <- exp(h/2)*rnorm(T, 0, 1)
+
+    sim_results[["y_sim"]] <- y_sim
+    sim_results[["sigma_sqd"]] <- sigma_sqd
+    sim_results[["sigma"]] <- sigma
+    sim_results[["mu"]] <- mu
+    sim_results[["phi"]] <- phi
+    
+    for(i in 1:length(h)){
+        sim_results[[paste("h[", i, "]", sep = "")]] <- h[i]
+    }
+
+    saveRDS(sim_results, file = paste(data_path, paste(iter, "RDS", sep = "."), sep = "/"))
+    
+}
+
+executables_path <- here::here("models/executables")
+if (!dir.exists(executables_path)) dir.create(executables_path)
[…]

### 2023-08-19 scripts/sim_ksc_results.r
revert to old tidyr funcs

@@ -12,12 +12,12 @@ tmp_data_location <- paste("simulation_output/", config$simulation_name, "/tmp",
 file_metadata <- as.data.frame(list.files(tmp_data_location))
 names(file_metadata) <- c("files")
 
-file_metadata <- separate_wider_delim(file_metadata, 
-                     cols = files, 
-                     delim = "_", 
-                     names = c("dataset", "mcmc_seed"), 
-                     cols_remove = FALSE) %>% 
-    separate_wider_delim(., cols = mcmc_seed, delim = ".", names = c("mcmc_seed", "parquet")) %>% 
+file_metadata <- tidyr::separate(file_metadata, 
+                     col = files, 
+                     sep = "_", 
+                     into = c("dataset", "mcmc_seed"),
+                     remove = FALSE) %>% 
+    tidyr::separate(., col = mcmc_seed, sep = "\\.", into = c("mcmc_seed", "parquet"), remove = TRUE) %>% 
     select(-parquet)
 
 simulation_results <- function(file_number){

### 2023-08-19 jobs/ksc.sh
reparameterised

@@ -1,0 +1,12 @@
+#!/bin/bash
+#SBATCH --job-name=sbc_cp_ksc_model_r3
+#SBATCH --time=06:00:00
+#SBATCH --mem-per-cpu=4096
+#SBATCH --array=1
+#SBATCH --ntasks=1
+#SBATCH --cpus-per-task=25
+#SBATCH --ntasks-per-socket=1
+
+source .env/bin/activate
+python sim_ksc.py
+R --vanilla < scripts/sim_ksc_results.r

### 2023-08-19 analysis/gauss_mix/gma.py
ksc scripts

@@ -1,0 +1,142 @@
+from __future__ import division
+
+import numpy as np
+import pandas as pd
+import statsmodels.api as sm
+
+from statsmodels.tsa.statespace.tools import (
+    constrain_stationary_univariate, unconstrain_stationary_univariate)
+
+from scipy.stats import norm, invgamma
+from scipy.special import logsumexp
+from scipy.signal import lfilter
+
+ksc_params = np.array([[0.04395, 2.77786,   0.16735],
+                       [0.24566, 1.79518,   0.34023],
+                       [0.34001, 0.61942,  0.64009],
+                       [0.25750, -1.08819,  1.26261],
+                       [0.10556, -3.97281,  2.61369],
+                       [0.00002, -8.56686,  5.17950],
+                       [0.00730, -10.12999, 5.79596]])
+
+
+
+class TVLLDT(sm.tsa.statespace.MLEModel):
+    """
+    Time-varying local linear deterministic trend
+    Mathematical form of class attributes:
+    https://www.statsmodels.org/dev/generated/statsmodels.tsa.statespace.representation.Representation.html#statsmodels.tsa.statespace.representation.Representation-attributes
+    """
+    def __init__(self, endog, offset=0.001):
+        # Convert to log squares, with offset
+        endog = np.log(endog**2 + offset)
+
+        # Initialize base model
+        super(TVLLDT, self).__init__(endog, k_states=1, k_posdef=1,
+                                     initialization='stationary')
+
+        # Setup time-varying arrays for observation equation
+        self['obs_intercept'] = np.zeros((1, self.nobs)) # mod.ssm.obs_intercept
+        self['obs_cov'] = np.zeros((1, 1, self.nobs)) # mod.ssm.obs_cov
+
+        # Setup fixed components of state space matrices
+        self['design', 0, 0] = 1 # mod.ssm.design. 
+        self['selection', 0, 0] = 1 # mod.ssm.selection
+
+    def update_mixing(self, indicators):
+        # z_t | s_t ~ N(m_i - 1.27036, v_i^2)
+        # See equation (10), p. 371
+        # Indicators is a vector
[…]

### 2023-08-19 analysis/gauss_mix/run_gma.py

@@ -1,0 +1,79 @@
+#Multiple chains
+#Seeds + datasets
+from __future__ import division
+
+import numpy as np
+import pandas as pd
+import pyarrow as pa
+import pyarrow.parquet as pq
+np.set_printoptions(suppress=True, precision=5)
+
+from gma import *
+
+np.random.seed(1234)
+
+df = pd.read_csv("../../data/simulated/ksc/phi_0.97779_sig_0.1585_beta_0.64733.csv")
+endog = df['yobs']
+
+# Setup the model and simulation smoother
+mod = TVLLDT(endog)
+mod.set_smoother_output(0, smoother_state=True) # Calculate and return the smoothed states. Other options like cov matrices available
+sim = mod.simulation_smoother()
+
+# Simulation parameters
+n_iterations = 1000
+burn = 100
+thin = 10
+
+# Storage for traces
+trace_smoothed = np.zeros((n_iterations + 1, mod.nobs))
+trace_states = np.zeros((n_iterations + 1, mod.nobs))
+trace_mixing = np.zeros((n_iterations + 1, mod.nobs), int)
+trace_mu = np.zeros((n_iterations + 1, 1))
+trace_phi = np.zeros((n_iterations + 1, 1))
+trace_sigma2 = np.zeros((n_iterations + 1, 1))
+
+# Initial values (p. 367)
+trace_mixing[0] = 0 # Initialised all with the 4th gaussian. So all draws from a specific normal distribution. Not specified in the paper how this is done
+trace_mu[0] = 0
+trace_phi[0] = 0.95
+trace_sigma2[0] = 0.02 #0.5 in the code. says 0.02 in the paper
+
+# Iterations
+for s in range(1, n_iterations + 1):
+    # Update the parameters of the model
+    mod.update_mixing(trace_mixing[s-1])
+    params = np.r_[trace_mu[s-1], trace_phi[s-1], trace_sigma2[s-1]]
+    mod.update(params, transformed=True)
+    
+    # Simulation smoothing
+    sim.simulate()
+    states = sim.simulated_state
+    trace_states[s] = states[0]
+
+    # Draw mixing indicators
+    trace_mixing[s] = draw_mixing(mod, states)
+    
+    # Draw parameters
+    trace_phi[s] = draw_posterior_phi(mod, states, trace_phi[s-1], trace_mu[s-1], trace_sigma2[s-1])
[…]

### 2023-08-19 scripts/_sim_ksc.py

@@ -1,0 +1,81 @@
+from __future__ import division
+import json
+
+import numpy as np
+import pandas as pd
+np.set_printoptions(suppress=True, precision=5)
+
+from src.gaussianMixture import *
+
+config_path = "configs/ksc.json"
+
+with open(config_path) as f:
+    config = json.load(f)
+
+np.random.seed(config["seed"])
+
+df = pd.read_csv("../../data/simulated/ksc/phi_0.97779_sig_0.1585_beta_0.64733.csv")
+endog = df['yobs']
+
+# Setup the model and simulation smoother
+mod = TVLLDT(endog)
+mod.set_smoother_output(0, smoother_state=True) # Calculate and return the smoothed states. Other options like cov matrices available
+sim = mod.simulation_smoother()
+
+# Storage for traces
+trace_smoothed = np.zeros((n_iterations + 1, mod.nobs))
+trace_states = np.zeros((n_iterations + 1, mod.nobs))
+trace_mixing = np.zeros((n_iterations + 1, mod.nobs), int)
+trace_mu = np.zeros((n_iterations + 1, 1))
+trace_phi = np.zeros((n_iterations + 1, 1))
+trace_sigma2 = np.zeros((n_iterations + 1, 1))
+
+# Initial values (p. 367)
+trace_mixing[0] = 0 # Initialised all with the 4th gaussian. So all draws from a specific normal distribution. Not specified in the paper how this is done
+trace_mu[0] = 0
+trace_phi[0] = 0.95
+trace_sigma2[0] = 0.02 #0.5 in the code. says 0.02 in the paper
+
+# Iterations
+for s in range(1, n_iterations + 1):
+    # Update the parameters of the model
+    mod.update_mixing(trace_mixing[s-1])
+    params = np.r_[trace_mu[s-1], trace_phi[s-1], trace_sigma2[s-1]]
+    mod.update(params, transformed=True)
+    
+    # Simulation smoothing
+    sim.simulate()
+    states = sim.simulated_state
+    trace_states[s] = states[0]
+
+    # Draw mixing indicators
+    trace_mixing[s] = draw_mixing(mod, states)
+    
+    # Draw parameters
+    trace_phi[s] = draw_posterior_phi(mod, states, trace_phi[s-1], trace_mu[s-1], trace_sigma2[s-1])
[…]

### 2023-08-19 scripts/sim_ksc.py

@@ -1,103 +1,0 @@
-from __future__ import division
-import os
-import json
-import random
-
-import numpy as np
-import pandas as pd
-import pyarrow as pa
-import pyarrow.parquet as pq
-from joblib import Parallel, delayed
-np.set_printoptions(suppress=True, precision=5)
-
-from src.gaussianMixture import *
-
-# Load config and create output directories
-config_path = "configs/ksc.json"
-data_path = "data/simulated/sbc"
-
-with open(config_path) as f:
-    config = json.load(f)
-
-# Create output directories
-output_dir = "simulation_output/" + config["simulation_name"] 
-os.mkdir(output_dir)
-os.mkdir(output_dir + "/output")
-os.mkdir(output_dir + "/tmp")
-os.mkdir(output_dir + "/std_out")
-
-# Save config
-with open(output_dir + "/config.json", "w") as outfile:
-    json.dump(config, outfile, indent=4)
-
-# Set seed and configure seed, data location and chains
-random.seed(config["seed"])
-
-sample_vect = random.sample(list(range(1, int(1e6))), config["n_iterations"] * config["chains"])
-data_file_indexes = [i for i in range(1, config["n_iterations"]+1) for _ in range(config["chains"])]
-seed_data_files = list(zip(sample_vect, data_file_indexes))
-
-
-def simulation(seed_data):
-    random.seed(seed_data[0])
-    data_file = data_path + "/" + config["data_location"] + f"/{seed_data[1]}.json"
-
-    with open(data_file) as f:
-        prior_params = json.load(f)
-
-    endog = pd.DataFrame(dict(y_obs = prior_params["y_sim"]))["y_obs"]
-
-    # Setup the model and simulation smoother
-    mod = TVLLDT(endog)
-    mod.set_smoother_output(0, smoother_state=True) # Calculate and return the smoothed states. Other options like cov matrices available
-    sim = mod.simulation_smoother()
-
-    # Storage for traces
-    trace_smoothed = np.zeros((config["mcmc_samples"] + 1, mod.nobs))
-    trace_states = np.zeros((config["mcmc_samples"] + 1, mod.nobs))
-    trace_mixing = np.zeros((config["mcmc_samples"] + 1, mod.nobs), int)
-    trace_mu = np.zeros((config["mcmc_samples"] + 1, 1))
[…]

### 2023-08-19 scripts/test_py.py

@@ -1,0 +1,50 @@
+# from joblib import Parallel, delayed
+# from multiprocessing import Pool
+import multiprocessing as mp
+
+print("Start")
+
+# def power_n_list(x_list, n):
+#     return [x ** n for x in x_list]
+
+def slice_data(data, nprocs):
+    aver, res = divmod(len(data), nprocs)
+    nums = []
+    for proc in range(nprocs):
+        if proc < res:
+            nums.append(aver + 1)
+        else:
+            nums.append(aver)
+    count = 0
+    slices = []
+    for proc in range(nprocs):
+        slices.append(data[count: count+nums[proc]])
+        count += nums[proc]
+    return slices
+
+def test(v):
+    return(v+1)
+
+
+nprocs = 2
+pool = mp.Pool(processes=nprocs)
+
+inp_lists = slice_data(range(10), nprocs)
+multi_result = [pool.apply_async(test, inp) for inp in inp_lists]
+
+# result = [x for p in multi_result for x in p.get()]
+print(multi_result)
+
+# def test(v):
+#     print(v)
+
+# Parallel(n_jobs=2, backend="multiprocessing")(delayed(test)(i) for i in [1,2,3,4])
+
+# with Pool(2) as p:
+    # p.map(test, [1, 2, 3,4,5,6])
+    # p.starmap(test, [1, 2, 3,4,5,6])
+
+print("Finish")
+
+
+

### 2023-08-21 writeups/research_proposal/propsal.qmd
init

---
title: "Comparison of MCMC algorithms in Stochastic Volatility Models"
format: pdf
---

# Background and Motivation
Maybe focus on the MCMC as opposed to SV upfront.

## SV
Why model volatility. Quantifying risks, tail probabilities, worst case scenarios. These are all inputs in the pricing of assets and derivatives. (pg92 FE slides)

SV models is one way to model volatility. Unlike the classic Black Scholes option pricing formula, it assumes that the underlying variance of an asset's return is not constant. That is, the variance of the underlying asset return is treated as a latent random variable. This deals with some of the empirical short comings of the BS model such as the volatility smile and skew in the realised variance in financial assets.

SV models are typically expressed as non linear Gaussian state space models. This can make it difficult to estimate using classical methods. The likelihood is unavailable in closed form and there are at least as many variables as data points. There are however, a variety of Frequentist and Bayesian strategies for estimating SV models. Frequentist approaches include quasi likelihood estimation and GMM which rely on asymptotic sampling distributions. Bayesian approaches rely on computational Markov Chain Monte Carlo techniques to sample from the target joint posterior of these high dimensional state space models.

This research will focus on Bayesian estimation strategies for Stochastic Volatility models. Particular attention will be given to the computational tools and approaches used to sample from high dimensional posterior distributions. Early approaches for estimating these models relied on classic Gibbs and Metropolis Hastings algorithms to apply the MCMC routines. Since then, there have been many developments in the availability of new kinds of MCMC samplers.

Hamiltonian Monte Carlo is a . It has been popularised through its implementation in a variety of open source programming languages. It presents a new way to sample from increasingly complicated generative models .

However, as the developments of these tools enable the estimation of more complicated models (with relative ease and accessibility through the development of open source software), so do the risk of mistakes and the need for procedures diagnose problems or mistakes when implementing these methods. How do we know that these algorithms are doing what we expect them to be doing, and how can we effectively differentiate the strengths and weaknesses of different tools.

How do we validate the computation and implementation of our sampling strategies and algorithms in increasingly complex models? And how do these form part of the computational modelling workflow?

<!-- ## HMC
- There have been many developemnts over the past few decades on estimating Bayesian models through Markov Chain Monte Carlo methods.

The original BUGs software implemented Gibbs and Metropolis Hastings MCMC algorithms for estimating this model.

In particular, the implementation of Hamiltonian Monte Carlo, a variant of the Metropolis Hastings algorithm, represents a recent advancement MCMC.

- The development of HMC is another advancement in estimating complicated, high dimensional models. 
- Development of diagnostics for different types of HMC and tools for evaluating the convergence and efficiency of the MCMC sampler.

- SV models are a primary candidate for using MCMC methods? Difficult to

## SBC -->

## Research Goal
The objective of this research is to evaluate the performance of MCMC algorithms for estimating SV models through a simulation study. Simulations are useful since there is control over the data generating process. This will help us understand the behaviour of MCMC and evaluate its performance against known parameters. Particular attention will be given to the calibration of the MCMC algorithms in the stochastic volatility context. Further diagnostics to evaluate performance will be explored to determine the efficiency of the MCMC sampler.

The first step is to understand what "performance" of MCMC algorithms means and what exactly we are measuring. At a high level, the objective of MCMC is to sample from the target posterior distribution and to provide reliable inference.  Convergence to the target posterior can be measured through efficiency measures such as rhat and effective sample size.

- Explcitly outline what is compared. KSC and HMC algorithms and estimation strategies.

<!-- 
Frequentist methods

- Variety of estimation strategies

- There are a variety of ways of estimating this model using a Bayesian frameworks.

Over the past few decades there have been many developments in computing hardware -->

RESEARCH GOAL: To compare the performance of different MCMC algorithms and estimation strategies when fitting stochastic volatility models. The study will use a discrete formulation is described

# Literature Review

## KSC
- KSC (1998) compare the performance of likelihood based inference and Bayesian inference at estimating the fit of these models. 
- Specifically KSC propose a method of sampling using an offset mixture normal approximation with a reweighting procedure to sample from the joint posterior of the SV model. This is estimated using conjugate priors as well as MH.
- The use of simulation and sampling are used as part of model diagnostics and evaluation. This is compared with other estimation strategies (MLE) and volatility models such as autogregressive conditional heteroskedasticity (ARCH).

## Stan HMC + NUTs
There have been many developments in MCMC algorithms since the introduction of classical methods such as GIBBS and MH. Hamiltoninn Monte Carlo (algorithm introduced when, widely available since 2010...?) is a relatively new im.

- Since then, there have been many developments in MCMC algorithms. In particular, HMC provides a flexible alternative to fitting and sampling from complicated high dimensional Bayesian models.
- It's key innovation is the use of gradients - enabling the flexible specification of any likelihood function. 
- Constraints- can only sample from continuous 
- HMC also provides new diagnostics for evaluating the effectiveness of the sampling procedure and how well the sampler is targeting the joint dist.

- How do we compare models? SBC. ESS in the tails, rhats.

- Given this context, my research looks to compare the efficacy of KSCs proposed MCMC sampling strategy with Hamiltonina Monte Carlo (NUTs).

## Simulation based calibration
- 2 papers

# 3) Methodology
The aim of this research is to use Simulation Based Calibration (SBC) to validate the inference of posterior samples generated by different MCMC algorithms in the context of Stochastic Volatility models. This will help identify any bias in computation and inconsistencies in model implementation and estimation strategies. Comparison of the SBC results between different computational approaches may help identify any strengths or shortcomings as well as any errors in model specification.

## 3.1) Stochastic Volatility Model
The univariate stochastic volatility model described by Kim, Shepherd and Chib (1998), now denoted as KSC, models the variance as a random variable following some latent stochastic process. $y_t$ is the mean corrected returns of some asset for equally spaced intervals t. $\beta$ is a constant scaling factor which is also defined as $exp(\mu / 2)$ representing instantaneous volatility. $h_t$ is log volatility, where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the autoregressive parameter $\phi$ such that $|\phi|<1$. This autoregressive parameter represents the persistence or "stickiness" of log volatility and the dispersion parameter $\sigma_{\eta}$ is its variance. $\epsilon_t$ and $\eta_t$ are standard normal white noise shocks.

$$
\begin{aligned}
y_t =& \space \beta exp(h_t/2) \epsilon_t \\
h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
h_1 \sim& \space normal\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
\end{aligned}
$$

$$
\begin{aligned}
\epsilon_t \sim& \space normal(0,1) \\
\delta_t \sim& \space normal(0,1)
\end{aligned}
$$

Setting $\beta=1$, the model can be expressed more succinctly as:

$$
\begin{aligned}
y_t \sim& \space normal(0, exp(h_t/2)) \\ 
h_1 \sim& \space normal \left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
h_{t+1} \sim& \space normal(\mu +\phi(h_t - \mu) , \sigma_{\eta}^2), \space\space t\neq 1\\ 
\end{aligned}
$$

With priors for the static parameters as defined in KSC:

$$
\begin{aligned}
\mu \sim& \space normal(0, 10^2) \\
\sigma_{\eta}^2 \sim& \space IG(5/2, (0.01\times 5) / 2) \\
\phi^{\ast} \sim& \space beta(20, 1.5) \\
\phi &=  2\phi^{\ast} - 1
\end{aligned}
$$

The prior on $\phi$ is a "stretched" beta distribution. This is a beta distribution (as defined on the parameter $\phi^*$) which has been transformed to have support (-1, 1).

## 3.2) Estimation Strategy

### KSC
KSC's model uses a simulation smoother[^1] (de Jong and Shephard, 1995) to estimate from the latent states. This requires the state space model to be linear and conditionally Gaussian.

[^1]: In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.

Since the model is not linear, a transformation is applied by squaring and taking the log of $y_t$.

$$
\begin{aligned}
y_t^{*} &= log(y_t^2) \\ 
&= log((\epsilon_t exp(h_t/2))^2) \\
&=  log(exp(h_t)) + log(\epsilon_t^2) \\
&= h_t + log(\epsilon_t^2)  \\
&= h_t + z_t \\
\end{aligned}
$$

Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and varaince 4.93. However, it is not simple to sample from this parameterisation of the model. KSC use a gaussian mixture model to **approximate** the first 4 moments of the log chi squared distribution. This is defined by:

$$
\begin{aligned}
f(z_t) = \sum_{i=1}^{K} q_if_N(z_i|m_i-1.2704, \nu_i^2)
\end{aligned}
$$

Where K is the mixture of normal densities $f_N$, component probabilities $q_i$, mean $m_i-1.2704$ and variance $\nu_i^2$. These parameters were seleted using moment matching where they found 7 normal densities with varying mean and variance parameters best approximated the log chi squared moments. These parameters and weights can be found in Appendix A.

The model can be sampled via the simulation smoother since the model is now linear and conditionally gaussian. The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions where as $\phi$ is sampled via a metropolis hastings accept/reject procedure. The details can be be around in Appendix B.

### Hamiltonian Monte Carlo (No-U-Turn Sampler)
Stan's default Hamiltonian Monte Carlo algorithm, the No-U-Turn Sampler, allows for direct sampling of the spcified stochastic volatility model. The HMC algorithm and more broadly the Stan programming language allows for sampling of the generative model and can flexibly handle complicated likelihood functions.

HMC will use the same priors are specified in the KSC model, however, the HMC algorithm does not require the priors to be conjugate (nor does the algorithm sample from conjugate or conditionally posterior distributions).

## Simulation Based Calibration (SBC)
<!-- SBC validates posterior samples and inferences generated by MCMC algorithms. It identifies inconsistencies and bias in computation and estimation of a generative model. This procedure  -->

SBC is conducted by comparing the distribution of rank statistics to the discrete uniform distribution which arises from a correctly calibrated algorithm. The procedure starts by generating datasets implied by the prior distribution and data generating process and calculating rank statistics from the resulting posterior samples. Starting with the parameters $\theta$ and data $y$, draw a sample from the prior distribution:

$$
\begin{aligned}
\theta^{sim} \sim \pi(\theta)
\end{aligned}
$$

Generate a dataset given by the prior draws and data generating process p.

$$
\begin{aligned}
y^{sim} \sim p(y|\theta^{sim})
\end{aligned}
$$

Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (i.e estimate the model on the simulated data using either KSC's strategy or HMC):
$$
\begin{aligned}
\{\theta_1,\dots , \theta_{L}\} \sim p(\theta | y^{sim})
\end{aligned}
$$

A key result from this procedure is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples $\theta^{sim}$. This is implied by the following expression:

$$
\begin{aligned}
\pi(\theta) &= \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})dy^{sim} d\theta^{sim} \\
&= \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) dy^{sim} d\theta^{sim}
\end{aligned}
$$

That is, the posterior averaged over the joint distribution follows the same distribution as the prior. The procedure of generating posterior samples implicitly performs this integral. The expression on the right hand side is proportional to the prior density and the simulation if done correctly should produce samples from the prior distribution. Therefore, any deviation of the posterior samples from the prior distribution suggests an error - either in computation, code or overall analysis.

Since the distribution of the posterior samples follows the prior distribution, the rank statistic for a given parameter and simultion follows a discrete uniform distribution[^2].

[^2]: Proof of this result in Talts et al.

$$
\begin{aligned}
r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim}) = \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}]
\end{aligned}
$$

This completes one iteration of SBC. Multiple iterations are run and the rank statistics are calculated for each parameter. The resulting rank statistics are compared to the discrete uniform distribution to determine if any problematic features or errors exist.

If the computation is well calibrated and the samples follow a discrete uniform distribution, then the posterior (credible) intervals will have sufficient coverage. That is, for any percentage interval selected (for example 90%) then there is 90\% chance that any $\theta^{sim}$ falls within this interval. In other words, a Bayesian analysis and computation is well calibrated if 90\% of constructed intervals contain the "true" parameter 90\% of the time.

<!-- ## Performance diagnostics

The goal of this simulation is to determine the "calibration" of MCMC when estimating SV models.

ADEMP:
- Aims
- Data generating mechanism
    + Write out mathematical form of the model here. Discrete time. 
- Estimand /target of analysis (maybe after explaining SBC?)
- Methods
    + SBC. Evaluating "calibration" of MCMC. What does calibration mean?
- Performance measures
    + Rhats, ESS, histogram

- Convergence diagnostic onto target joint posterior distribution? rhats/ess. This assess "efficiency" of MCMC.
- Tail ESS calculates ESS in the tail quantiles. Such  -->

# Results
SBC for the HMC algorithm and the KSC Gaussian approxmiation are conducted with simulated dataset sizes of 1000 observations and 1000 simulation interations. The preliminary results focus on a subset of parameters since there are parameters for each latent state (i.e one for each data observation) plus the static parameters.

Figure 1 shows the the distribution of rank statistics for the HMC algorithm, with the horizontal black line at 50 representing the discrete uniform distribution. Phi and sigma_sqd look relatively uniform. However mu and h.995 have some lumpiness which may suggest a lack of calibration.

![Distribution of parameter ranks for HMC algorithm](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/static_state_hist.png){width=75%}

One of the drackbacks of SBC is that any deviation from uniform from a calibrated analysis maybe due to finite samples. Increasing the number of iterations from 1000 to 5000 shows more consistent uniform behaviour from all parameters (Figure 2). This suggests that the HMC algorithm and overall analysis is well calibrated.

![Distribution of parameter ranks for HMC algorithm (5000 iterations)](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r6_5000_iterations/static_state_hist.png){width=75%}

Figure 3 displays the results from KSC's Gaussian approximation. Compared to HMC, the distribution of ranks deviate heavily from the expected discrete uniform distribution. The lack of calibration relative to the HMC algorithm is expected. The Gaussian Mixture Approximiation only approximates the log chi squared distribution. The data generating process is from the specified Stochastic Volatility model. We expect that since the HMC strategy samples directly from the true data generating process that it would have more success at

![Distribution of parameter ranks for HMC algorithm (5000 iterations)](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png){width=75%}

# Conclusion and future research considerations

## Plan
The KSC model also also apply a Metropois Hastings correction for the fact that the sampling strategy approximates the error distribution.

Extensions to the model table:

    + Stan (HMC), KSC (GMA), KSC (GMA + MH correction)
    + CP, NCP

Parameterisation

## Additional considerations
- Out of sample prediction on real dataset

# References

# Appendix A: Mixture Gaussian weights
Weights

# Appendix B: Conjugate posterior distributions

### 2023-08-21 writeups/research_proposal/propsal.qmd
fixed figures

@@
 title: "Comparison of MCMC algorithms in Stochastic Volatility Models"
-format: pdf
+format:
+  pdf:
+    fig-pos: H
 ---
@@
 # Results
-SBC for the HMC algorithm and the KSC Gaussian approxmiation are conducted with simulated dataset sizes of 1000 observations and 1000 simulation interations.
+SBC for the HMC algorithm and the KSC Gaussian approxmiation are conducted with simulated dataset sizes of 1000 observations, 1000 simulation interations with post warmup/burnin samples of 999 (which gives 1000 possible rank statistics).
+The histogram is drawn with 20 bins and 50 observations in each bin which is the expected count for a discrete uniform distribution.
 The preliminary results focus on a subset of parameters since there are parameters for each latent state (i.e one for each data observation) plus the static parameters.
@@
 
-![Distribution of parameter ranks for HMC algorithm](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/static_state_hist.png){width=75%}
+![Distribution of parameter ranks for HMC algorithm (1000 iterations)](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/static_state_hist.png){width=75%}
 
@@
 The Gaussian Mixture Approximiation only approximates the log chi squared distribution.
-The data generating process is from the specified Stochastic Volatility model.
-We expect that since the HMC strategy samples directly from the true data generating process that it would have more success at
+The HMC strategy samples directly from the same likelihood as the data generating process.
 
-![Distribution of parameter ranks for HMC algorithm (5000 iterations)](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png){width=75%}
+![Distribution of parameter ranks for KSC sampling strategy (1000 iterations)](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png){width=75%}
 
-# Conclusion and future research considerations
+An alternative to checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each bin.
+Let $b_j$ be the number of counts and $e_j$ the expected count in bin $j$.
+Then the chi squared statistic is given by:
 
-## Plan
-The KSC model also also apply a Metropois Hastings correction for the fact that the sampling strategy approximates the error distribution.
+$$
+\begin{aligned}
+\chi^2 = \sum_{j=1}^J \frac{(b_{j} - e_{j})^2}{e_j}
+\end{aligned}
+$$
 
-Extensions to the model table:
+These chi squared statistics can be considered independent draws from a chi squared distribution with 19 degrees of freedom (since there are 20 bins).
+Visualising this should give a chi squared distribution (if performing a hypothesis test, this would be the sampling distribution under a null of uniformity).
 
-    + Stan (HMC), KSC (GMA), KSC (GMA + MH correction)
-    + CP, NCP
+Figure 4 and 5 are the chi squared statistic histograms results of the HMC algorithm and KSC algorithm respectively.
+The vertical black line marks the 95th quantile which represents the critical value for hypothesis testing.
+The HMC chi squared statistics look like they are sampled from the correct distribution whereas KSC are inconsistent with the expected distribution.
 
-Parameterisation
+![Chi squared statistics for HMC](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/chi_sq_hist.png){width=70%}
 
-## Additional considerations
-- Out of sample prediction on real dataset
+![Chi squared statistics for KSC](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/chi_sq_hist.png){width=70%}
+
+# Conclusion and next steps
+KSC also also apply a Metropolis Hastings reweighting to correct the approximation of the error distribution.
+According to the authors, the correction ensures that samples are from the exact posterior density.
+The next step is to perform this correction and evaluate the SBC results.
+
+Another factor that may affect sampling performance is the parameterisation of the model.
+This can cause significant problems to samplers in models with increasing complexity.
+A classic example of this is Neal's funnel which arises from Hierarchical models - causing bias and performance issues with the HMC sampler.
+The solution is to rewrite the model with non centered parameterisation which significantly improves the performance of the sampler.
+Another step in this research is to consider the results of SBC under different model parameterisations.
+The models and parametiersations for this research are summarised in the below table:
+
+![Model matrix](model_table.png)
 

### 2023-08-21 writeups/research_proposal/propsal.qmd
spacing

@@
     fig-pos: H
+    include-in-header:
+        text: |
+          \usepackage{lipsum}
+          \usepackage{setspace}
+          \onehalfspacing
+          \linespread{1.5}
+fontsize: 12pt
+geometry: margin=1in
 ---
 
-# Background and Motivation
+<!-- # Background and Motivation
 Maybe focus on the MCMC as opposed to SV upfront.
@@
 
-How do we validate the computation and implementation of our sampling strategies and algorithms in increasingly complex models? And how do these form part of the computational modelling workflow?
+How do we validate the computation and implementation of our sampling strategies and algorithms in increasingly complex models?
+And how do these form part of the computational modelling workflow?
+-->
 
@@
 
-# Literature Review
+# Background and Literature Review
 
-## KSC
-- KSC (1998) compare the performance of likelihood based inference and Bayesian inference at estimating the fit of these models. 
-- Specifically KSC propose a method of sampling using an offset mixture normal approximation with a reweighting procedure to sample from the joint posterior of the SV model.
-This is estimated using conjugate priors as well as MH.
-- The use of simulation and sampling are used as part of model diagnostics and evaluation.
-This is compared with other estimation strategies (MLE) and volatility models such as autogregressive conditional heteroskedasticity (ARCH).
+Kim Sherpherd and Chib (1998) estimate the discrete time Stochastic Volatility model by proposing a simulation strategy using Bayesian methods.
+Specifically, KSC propose a method of sampling using an offset mixture normal approximation with a reweighting procedure to sample from the posterior of the SV model parameters.
+By linearing the model and using a mixture of Gaussian's, they were able to apply a Kalman Filter to estimate the latent volatilities of the state space models.
+The use of simulation and sampling are used as part of model diagnostics and evaluation.
+This is compared with other estimation strategies and volatility models such as autogregressive conditional heteroskedasticity (ARCH).
+Their specificaiton of the Stochastic Volatility model and estimation approaches provide the context for comparing the performance of different MCMC algorithms.
 
-## Stan HMC + NUTs
-There have been many developments in MCMC algorithms since the introduction of classical methods such as GIBBS and MH.
+The simulations run by KSC rely on sampling directly from conjugate priors and classic Markov Chain Monte Carlo algorithms, Gibbs and Metropolis Hastings.
+Since the paper was written, new and more powerful MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
+Hamiltonian Monte Carlo, through open source programming languages such as Stan and PyMC, have become
+
+There have been many developments in MCMC algorithms since the introduction of classical methods such as Gibbs and Metropolis Hastings.
+
 Hamiltoninn Monte Carlo (algorithm introduced when, widely available since 2010...?) is a relatively new im.
@@
 
-## Simulation Based Calibration (SBC)
+## 3.3 Simulation Based Calibration (SBC)
 <!-- SBC validates posterior samples and inferences generated by MCMC algorithms. It identifies inconsistencies and bias in computation and estimation of a generative model. This procedure  -->

### 2023-08-22 writeups/research_proposal/propsal.qmd
first draft

@@
 ## Research Goal
-The objective of this research is to evaluate the performance of MCMC algorithms for estimating SV models through a simulation study.
+The objective of this research is to validate and compare the comutational methods used to estimate Bayesian Stochastic Volatiity models.
+Specifically, Simulation Based Calibration will be used to identify bias and errors arising from the results of computational algorithms.
+The source of such errors include, but are not limited to, different estimation strategies, problems with the sampler, problematic model specifications and coding mistakes.
+This study will compare the estimation strategies detailed in the Stochastic Volatility literature and the application of Hamiltonina Monte Carlo for estimating these class of models.
+Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches and provide tools for evaluating other types models and algorithms.
+
+<!-- The objective of this research is to evaluate the performance of MCMC algorithms for estimating SV models through a simulation study.
 Simulations are useful since there is control over the data generating process.
@@
 
-- Explcitly outline what is compared. KSC and HMC algorithms and estimation strategies.
+- Explcitly outline what is compared. KSC and HMC algorithms and estimation strategies.  -->
 
@@
 Over the past few decades there have been many developments in computing hardware -->
-
-RESEARCH GOAL: To compare the performance of different MCMC algorithms and estimation strategies when fitting stochastic volatility models. The study will use a discrete formulation is described
+<!-- 
+RESEARCH GOAL: To compare the performance of different MCMC algorithms and estimation strategies when fitting stochastic volatility models. The study will use a discrete formulation is described  -->
 
@@
 
-Kim Sherpherd and Chib (1998) estimate the discrete time Stochastic Volatility model by proposing a simulation strategy using Bayesian methods.
-Specifically, KSC propose a method of sampling using an offset mixture normal approximation with a reweighting procedure to sample from the posterior of the SV model parameters.
+Kim, Sherpherd and Chib (1998) estimate the discrete time Stochastic Volatility model by proposing a simulation strategy using Bayesian methods.
+Specifically, they propose a method of sampling using an offset mixture normal approximation with a reweighting procedure to sample from the posterior of the SV model parameters.
 By linearing the model and using a mixture of Gaussian's, they were able to apply a Kalman Filter to estimate the latent volatilities of the state space models.
@@
 
-The simulations run by KSC rely on sampling directly from conjugate priors and classic Markov Chain Monte Carlo algorithms, Gibbs and Metropolis Hastings.
-Since the paper was written, new and more powerful MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
-Hamiltonian Monte Carlo, through open source programming languages such as Stan and PyMC, have become
+The simulations run by Kim, Sherpherd and Chib rely on sampling directly from conjugate priors and use of classic Markov Chain Monte Carlo algorithms, Gibbs and Metropolis Hastings.
+Since the paper was written, new MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
+Specifically, Hamiltonian Monte Carlo is an a MCMC algorithm which has been become widely available for efficiently sampling from sophisticated models.
+Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo was developed in the physics literature (Duane et al., 1987) before being applied in the statistics literature by Radford Neal through his works in Bayesean Neural Networks (Nael, 1995) and statistical computing (Neal 2011).
+The algorithm has since become widely available through open source developement projects such as Stan (Stan Development Team, 2017) and PyMC (2023).
+
+New developments in MCMC algorithms and improved computing resources have enabled estimation of richer and more complicated models.
+However, as the development of tools to enable us to fit more complicated models, so does the need to validate the output of these tools.
+Cook, Gelman and Rubin (2006) introduce simulation techniques to evaluate the "correctness" of the software used to sample Bayesian models.
+Their contribution used CDF quantiles from the posterior samples will be uniformly distributed.
+Talts et al.
+build upon this work by developing Simulation Based Calibration to provide more robust validation of output from a MCMC sampler through the use of rank statistics.
+
+ <!-- through the use of the CDF quantiles approximated from the posterior distribution. -->
+
+<!-- , through open source programming languages such as Stan and PyMC, have become
+
+As new and complicated
 
@@
 
-- Given this context, my research looks to compare the efficacy of KSCs proposed MCMC sampling strategy with Hamiltonina Monte Carlo (NUTs).
-
-## Simulation based calibration
-- 2 papers
+- Given this context, my research looks to compare the efficacy of KSCs proposed MCMC sampling strategy with Hamiltonina Monte Carlo (NUTs).  -->
 

### 2023-08-23 writeups/research_proposal/propsal.qmd
first draft edited

@@
 The objective of this research is to validate and compare the comutational methods used to estimate Bayesian Stochastic Volatiity models.
-Specifically, Simulation Based Calibration will be used to identify bias and errors arising from the results of computational algorithms.
-The source of such errors include, but are not limited to, different estimation strategies, problems with the sampler, problematic model specifications and coding mistakes.
-This study will compare the estimation strategies detailed in the Stochastic Volatility literature and the application of Hamiltonina Monte Carlo for estimating these class of models.
-Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches and provide tools for evaluating other types models and algorithms.
+Specifically, Simulation Based Calibration will be used to identify biases and errors arising from the results of Markov Chain Monte Carlo (MCMC) algorithms.
+The source of any miscalibration include, but are not limited to, estimation strategies, choice of sampler, model specifications and coding mistakes.
+This study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
+Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches.
+
+<!-- and provide tools for evaluating other types models and algorithms.  -->
 
@@
 
-Kim, Sherpherd and Chib (1998) estimate the discrete time Stochastic Volatility model by proposing a simulation strategy using Bayesian methods.
-Specifically, they propose a method of sampling using an offset mixture normal approximation with a reweighting procedure to sample from the posterior of the SV model parameters.
-By linearing the model and using a mixture of Gaussian's, they were able to apply a Kalman Filter to estimate the latent volatilities of the state space models.
+Kim, Sherpherd and Chib (1998) estimate the discrete time stochastic volatility model by proposing a simulation strategy using Bayesian methods.
+Specifically, they propose a method of sampling using an offset mixture normal approximation with a reweighting procedure to sample from the posterior of the stochastic volatility model parameters.
+By linearing the model and using a mixture of Gaussian's, they were able to apply a Kalman Filter to estimate the posterior density over the latent volatilities and static parameters.
 The use of simulation and sampling are used as part of model diagnostics and evaluation.
 This is compared with other estimation strategies and volatility models such as autogregressive conditional heteroskedasticity (ARCH).
-Their specificaiton of the Stochastic Volatility model and estimation approaches provide the context for comparing the performance of different MCMC algorithms.
+Their specificaiton of the stochastic volatility model and estimation approaches provide the context for comparing the performance of different MCMC algorithms.
 
-The simulations run by Kim, Sherpherd and Chib rely on sampling directly from conjugate priors and use of classic Markov Chain Monte Carlo algorithms, Gibbs and Metropolis Hastings.
+The simulations run by Kim, Sherpherd and Chib rely on sampling directly from conjugate priors and use of classic MCMC algorithms, Gibbs and Metropolis Hastings.
 Since the paper was written, new MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
@@
 Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo was developed in the physics literature (Duane et al., 1987) before being applied in the statistics literature by Radford Neal through his works in Bayesean Neural Networks (Nael, 1995) and statistical computing (Neal 2011).
-The algorithm has since become widely available through open source developement projects such as Stan (Stan Development Team, 2017) and PyMC (2023).
+The algorithm has since become widely available through open source developement projects such as Stan (Stan Development Team, 2023) and PyMC (2023).
+
+The key innovation of Hamiltonian Monte Carlo is using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
+Unlike Random Walk Metropolis Hastings, it takes advantage of the geometry of the posterior to determine its next step.
+A comprehensive explanation of the sampler is beyond the scope of this research and can be found in the above references.
 
@@
 Cook, Gelman and Rubin (2006) introduce simulation techniques to evaluate the "correctness" of the software used to sample Bayesian models.
-Their contribution used CDF quantiles from the posterior samples will be uniformly distributed.
+Their contribution identified that cumulative density quantiles from the posterior samples will be uniformly distributed.
 Talts et al.
@@
 ## 3.1) Stochastic Volatility Model
-The univariate stochastic volatility model described by Kim, Shepherd and Chib (1998), now denoted as KSC, models the variance as a random variable following some latent stochastic process.
+The univariate stochastic volatility model described by Kim, Shepherd and Chib (1998), models the variance as a random variable following some latent stochastic process.
 $y_t$ is the mean corrected returns of some asset for equally spaced intervals t.
@@
 
-With priors for the static parameters as defined in KSC:
+Priors for the static parameters are defined below with $\mu$ and $\sigma^2$ being conjugate priors:
 
@@
 
-### KSC
-KSC's model uses a simulation smoother[^1] (de Jong and Shephard, 1995) to estimate from the latent states. This requires the state space model to be linear and conditionally Gaussian.
+### Gaussian Mixture Approximation
+Kim Shepherd and Chib estimate the stochastic volatility model using a Kalman Filter and smoother[^1] (de Jong and Shephard, 1995) to compute the posterior distribution over the latent states.
+This requires the state space model to be linear and conditionally Gaussian.
 
@@
 However, it is not simple to sample from this parameterisation of the model.
-KSC use a gaussian mixture model to **approximate** the first 4 moments of the log chi squared distribution.
+Kim, Sherpherd and Chib use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution.
 This is defined by:
@@
 ### Hamiltonian Monte Carlo (No-U-Turn Sampler)
-Stan's default Hamiltonian Monte Carlo algorithm, the No-U-Turn Sampler, allows for direct sampling of the spcified stochastic volatility model.
-The HMC algorithm and more broadly the Stan programming language allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
-
-HMC will use the same priors are specified in the KSC model, however, the HMC algorithm does not require the priors to be conjugate (nor does the algorithm sample from conjugate or conditionally posterior distributions).
+The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
+Stan's default algorithm, the No-U-Turn Sampler (Hoffman & Gelman 2014), allows for direct sampling of the spcified stochastic volatility model.
+Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
+This approach will use the same priors are specified in the Gaussian mixture approximation.
+However, Hamiltonian does not sample directly from the analytical form of conjugate or conditionally conjugate distribution, so different priors can be chosen if necessary.
 
@@
 
-SBC is conducted by comparing the distribution of rank statistics to the discrete uniform distribution which arises from a correctly calibrated algorithm.
-The procedure starts by generating datasets implied by the prior distribution and data generating process and calculating rank statistics from the resulting posterior samples.
-Starting with the parameters $\theta$ and data $y$, draw a sample from the prior distribution:
-
+SBC is conducted by comparing the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.
+The procedure starts by taking draws from the prior distribution and creating datasets implied by each draw.
+Rank statistics are then calculated on the posterior samples estimated from the simulated data.
+Let $\theta$ be a parameter and $y$ represent the dataset.
+Start with draw a sample from the prior distribution:
 $$
@@
 
-Generate a dataset given by the prior draws and data generating process p.
+Generate a dataset given by the prior draws.
 
@@
 \begin{aligned}
-y^{sim} \sim p(y|\theta^{sim})
+y^{sim} \sim \pi (y|\theta^{sim})
 \end{aligned}
@@
 
-Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (i.e estimate the model on the simulated data using either KSC's strategy or HMC):
+Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (Hamiltonian Monte Carlo or Gaussian apprimation in our case).
+
 $$
 \begin{aligned}
-\{\theta_1,\dots , \theta_{L}\} \sim p(\theta | y^{sim})
+\{\theta_1,\dots , \theta_{L}\} \sim \pi (\theta | y^{sim})
 \end{aligned}
@@
 That is, the posterior averaged over the joint distribution follows the same distribution as the prior.
-The procedure of generating posterior samples implicitly performs this integral.
-The expression on the right hand side is proportional to the prior density and the simulation if done correctly should produce samples from the prior distribution.
+The procedure of generating posterior samples implicitly performs this integral since the expression on the right of the integral is proportional to the prior density.
 Therefore, any deviation of the posterior samples from the prior distribution suggests an error - either in computation, code or overall analysis.
+
+<!-- The simulation if done correctly should produce samples from the prior distribution.  -->
 
@@
 
-[^2]: Proof of this result in Talts et al.
+[^2]: Proof of this result in Talts et al 2018.
 
@@
 This completes one iteration of SBC.
-Multiple iterations are run and the rank statistics are calculated for each parameter.
+To complete the algorithm, multiple iterations are run and the rank statistics are calculated for each parameter.
 The resulting rank statistics are compared to the discrete uniform distribution to determine if any problematic features or errors exist.
@@
 # Results
-SBC for the HMC algorithm and the KSC Gaussian approxmiation are conducted with simulated dataset sizes of 1000 observations, 1000 simulation interations with post warmup/burnin samples of 999 (which gives 1000 possible rank statistics).
+SBC for Hamiltonian Monte Carlo and the Gaussian approxmiation are performed with simulated dataset sizes of 1000 simulation interations, 1000 observations, and post warmup/burnin samples of 999 (which gives 1000 possible rank statistics).
 The histogram is drawn with 20 bins and 50 observations in each bin which is the expected count for a discrete uniform distribution.
-The preliminary results focus on a subset of parameters since there are parameters for each latent state (i.e one for each data observation) plus the static parameters.
+This algorithm is summarised below:
 
-Figure 1 shows the the distribution of rank statistics for the HMC algorithm, with the horizontal black line at 50 representing the discrete uniform distribution.
+| **1) for n in 1000 iterations:**
+|   2) Draw from prior: $\theta^{sim}\sim\pi (\theta)$
+|   3) Simulate dataset with 1000 observations: $y^{sim} \sim p(y|\theta^{sim})$
+|   4) Draw 999 posterior samples (post warmup) $\{\theta_1,\dots , \theta_{L}\} \sim p(\theta | y^{sim})$
+|   5) Compute rank statistics $r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim})$
+
+Figure 1 shows the the distribution of rank statistics for Hamiltonian Monte Carlo, with the horizontal black line at 50 representing the discrete uniform distribution.
+The preliminary figures focus on the static parameters and arbitrarily chosen states (since there is one state parameter per data observation).
 Phi and sigma_sqd look relatively uniform.
@@
 One of the drackbacks of SBC is that any deviation from uniform from a calibrated analysis maybe due to finite samples.
-Increasing the number of iterations from 1000 to 5000 shows more consistent uniform behaviour from all parameters (Figure 2).
-This suggests that the HMC algorithm and overall analysis is well calibrated.
+Figure 2 shows more consistent uniform bahviour from all parameters after increasing the number of iterations from 1000 to 5000.
+This suggests that Hamiltonian Monte Carlo and overall analysis is well calibrated.
 
@@
 
-Figure 3 displays the results from KSC's Gaussian approximation.
-Compared to HMC, the distribution of ranks deviate heavily from the expected discrete uniform distribution.
-The lack of calibration relative to the HMC algorithm is expected.
-The Gaussian Mixture Approximiation only approximates the log chi squared distribution.
-The HMC strategy samples directly from the same likelihood as the data generating process.
+Figure 3 shows results for the Gaussian approximation.
+The distribution of the rank statistics deviate heavily from the expected discrete uniform distribution, with various peaks at each end of the histogram as well as peaks in the middle.
+This lack of calibration is expected.
+The Gaussian approximiation only approximates the log chi squared distribution, where as Hamiltonian Monte Carlo samples directly from the same likelihood as the data generating process.
 
@@
 
-Figure 4 and 5 are the chi squared statistic histograms results of the HMC algorithm and KSC algorithm respectively.
+Figure 4 are the chi squared statistic histograms results of Hamiltonian Monte Carlo and Gaussian approximation respectively.
 The vertical black line marks the 95th quantile which represents the critical value for hypothesis testing.
-The HMC chi squared statistics look like they are sampled from the correct distribution whereas KSC are inconsistent with the expected distribution.
+The HMC chi squared statistics look like they are sampled from the correct distribution, with 6.2\% of the samples greater than the 95\% critical value (doing this calculation for 5000 iterations leads to 4.8\% of samples in the critical area).
+Figure 5 shows the majority of chi squared statistics for Gaussian approximation within the critical area, suggesting the estimation strategy is not well calibrated.
 
@@
 # Conclusion and next steps
-KSC also also apply a Metropolis Hastings reweighting to correct the approximation of the error distribution.
+There are two key extensions to the simulation results worth exploring.
+
+**1) Metropolis Hastings reweighting for KSC model**
+
+Kim, Sherpherd and Chib apply a Metropolis Hastings reweighting to the posterior samples to correct for the approximation of the error distribution.
 According to the authors, the correction ensures that samples are from the exact posterior density.
 The next step is to perform this correction and evaluate the SBC results.
+I'd expect the Gaussian mixture approximation diagnostics to improve after this correction.
 
-Another factor that may affect sampling performance is the parameterisation of the model.
-This can cause significant problems to samplers in models with increasing complexity.
-A classic example of this is Neal's funnel which arises from Hierarchical models - causing bias and performance issues with the HMC sampler.
-The solution is to rewrite the model with non centered parameterisation which significantly improves the performance of the sampler.
-Another step in this research is to consider the results of SBC under different model parameterisations.
+**2) Centered and non centered parameterisation**
+
+The parameterisation of a model can affect the performance of a sampler.
+This can cause significant problems in getting unbiased MCMC samples for complex models.
+An example of this is Neal's funnel (Neal 2003) which arises in hierarchical models which causes performance issues with the Hamiltonian Monte Carlo sampler.
+Rewriting the model with non centered parameterisation significantly improves the performance of MCMC.
+Another consideration for this research is to consider both centered and non centered parametersations of the stochastic volatility model to see if there are any improvements to the simulation diagnostics.
+
 The models and parametiersations for this research are summarised in the below table:
@@
 
+## Plan
+Week 7: Reflect and integrate feedback from presentation into plan for the rest of the semester
+
+Week 8: Finish off any simulation or modelling not yet complete (listed in above matrix)
+
+Week 9: Write up any other results and have a first draft ready for feedback
+
+Week 10-11: Action any feedback, prepare for final submission and presentation
+
+Week 12: Submit
+
 # References
 
+Abril-Pla O, Andreani V, Carroll C, Dong L, Fonnesbeck CJ, Kochurov M, Kumar R, Lao J, Luhmann CC, Martin OA, Osthege M, Vieira R, Wiecki T, Zinkov R.
+(2023) PyMC: a modern, and comprehensive probabilistic programming framework in Python.
+
+Duane, Simon; Kennedy, Anthony D.; Pendleton, Brian J.; Roweth, Duncan (1987). "Hybrid Monte Carlo". Physics Letters B. 195 (2): 216–222.
+
+Hoffman, M. D., & Gelman, A. (2014). The No-U-Turn sampler: adaptively setting path lengths in Hamiltonian Monte Carlo. J. Mach. Learn. Res., 15(1), 1593-1623.
+
+Kim, Sangjoon, Neil Shephard, and Siddhartha Chib. 1998. “Stochastic Volatility: Likelihood Inference and Comparison with ARCH Models.” Review of Economic Studies 65: 361–93.
+
+Neal, R.M. (1995) Bayesian Learning for Neural Networks. Ph.D. Thesis, Graduate Department of Computer Science, University of Toronto, Toronto.
+
+Neal, R. M. (2003). Slice Sampling. The Annals of Statistics, 31(3), 705–741. http://www.jstor.org/stable/3448413
+
+Neal, R.M. (2011) MCMC Using Hamiltonian Dynamics. In Handbook of Markov Chain Mone Carlo, CRC Press, New York
+
+Stan Development Team. 2023. Stan Modeling Language Users Guide and Reference Manual, 2.32. https://mc-stan.org
+
+Talts, S., Betancourt, M., Simpson, D., Vehtari, A., & Gelman, A. (2018). Validating Bayesian inference algorithms with simulation-based calibration. arXiv preprint arXiv:1804.06788.
+
 # Appendix A: Mixture Gaussian weights
-Weights
 

### 2023-08-24 writeups/research_proposal/propsal.qmd
typos and independence of chi

@@
 ## Research Goal
-The objective of this research is to validate and compare the comutational methods used to estimate Bayesian Stochastic Volatiity models.
+The objective of this research is to validate and compare the computational methods used to estimate Bayesian stochastic volatility models.
 Specifically, Simulation Based Calibration will be used to identify biases and errors arising from the results of Markov Chain Monte Carlo (MCMC) algorithms.
@@
 
+<!-- The use of simulation and sampling are used as part of model diagnostics and evaluation. -->
+
 # Background and Literature Review
@@
 Kim, Sherpherd and Chib (1998) estimate the discrete time stochastic volatility model by proposing a simulation strategy using Bayesian methods.
-Specifically, they propose a method of sampling using an offset mixture normal approximation with a reweighting procedure to sample from the posterior of the stochastic volatility model parameters.
-By linearing the model and using a mixture of Gaussian's, they were able to apply a Kalman Filter to estimate the posterior density over the latent volatilities and static parameters.
-The use of simulation and sampling are used as part of model diagnostics and evaluation.
-This is compared with other estimation strategies and volatility models such as autogregressive conditional heteroskedasticity (ARCH).
-Their specificaiton of the stochastic volatility model and estimation approaches provide the context for comparing the performance of different MCMC algorithms.
+Specifically, they propose a method of sampling by using an offset mixture normal approximation with a reweighting procedure to sample from the posterior.
+By linearing the model and using a mixture of Gaussians, they were able to apply a Kalman Filter to estimate the posterior density over the latent volatilities and static parameters.
+This is compared with other estimation strategies and volatility models such as autoregressive conditional heteroskedasticity (ARCH).
+Their specification of the stochastic volatility model and estimation approaches provide the context for comparing the performance of different MCMC algorithms.
 
-The simulations run by Kim, Sherpherd and Chib rely on sampling directly from conjugate priors and use of classic MCMC algorithms, Gibbs and Metropolis Hastings.
+This strategy relies on sampling directly from conjugate posterior distributions and the use of classic MCMC algorithms, Gibbs and Random Walk Metropolis Hastings.
 Since the paper was written, new MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
-Specifically, Hamiltonian Monte Carlo is an a MCMC algorithm which has been become widely available for efficiently sampling from sophisticated models.
-Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo was developed in the physics literature (Duane et al., 1987) before being applied in the statistics literature by Radford Neal through his works in Bayesean Neural Networks (Nael, 1995) and statistical computing (Neal 2011).
-The algorithm has since become widely available through open source developement projects such as Stan (Stan Development Team, 2023) and PyMC (2023).
+Specifically, Hamiltonian Monte Carlo is a MCMC algorithm which has become widely available for efficiently sampling from sophisticated models.
+Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature (Duane et al.
+1987) before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks (Nael, 1995) and statistical computing (Neal 2011).
+The algorithm has since become widely available through open source development projects such as Stan (Stan Development Team, 2023) and PyMC (2023).
 
@@
 New developments in MCMC algorithms and improved computing resources have enabled estimation of richer and more complicated models.
-However, as the development of tools to enable us to fit more complicated models, so does the need to validate the output of these tools.
+However, as the development of tools enables us to fit more complicated models, so does the need to validate the output of these tools.
 Cook, Gelman and Rubin (2006) introduce simulation techniques to evaluate the "correctness" of the software used to sample Bayesian models.
-Their contribution identified that cumulative density quantiles from the posterior samples will be uniformly distributed.
-Talts et al.
+Their contribution identified that cumulative density quantiles from the posterior samples can be tested for uniformity.
+Talts et al (2018).
 build upon this work by developing Simulation Based Calibration to provide more robust validation of output from a MCMC sampler through the use of rank statistics.
@@
 ## 3.1) Stochastic Volatility Model
-The univariate stochastic volatility model described by Kim, Shepherd and Chib (1998), models the variance as a random variable following some latent stochastic process.
+The univariate stochastic volatility model described by Kim, Shephard and Chib (1998) models the variance as a random variable following some latent stochastic process.
 $y_t$ is the mean corrected returns of some asset for equally spaced intervals t.
@@
 
-Priors for the static parameters are defined below with $\mu$ and $\sigma^2$ being conjugate priors:
+Priors for the static parameters are defined below with conjugate priors on $\mu$ and $\sigma^2$:
 
@@
 ### Gaussian Mixture Approximation
-Kim Shepherd and Chib estimate the stochastic volatility model using a Kalman Filter and smoother[^1] (de Jong and Shephard, 1995) to compute the posterior distribution over the latent states.
+Kim, Shephard and Chib estimate the stochastic volatility model using a Kalman Filter and smoother[^1] (de Jong and Shephard, 1995) to compute the posterior distribution over the latent states.
 This requires the state space model to be linear and conditionally Gaussian.
@@
 
-Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and varaince 4.93.
+Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
 However, it is not simple to sample from this parameterisation of the model.
@@
 Where K is the mixture of normal densities $f_N$, component probabilities $q_i$, mean $m_i-1.2704$ and variance $\nu_i^2$.
-These parameters were seleted using moment matching where they found 7 normal densities with varying mean and variance parameters best approximated the log chi squared moments.
+These parameters were selected using moment matching where they found 7 normal densities with varying mean and variance parameters best approximated the log chi squared moments.
 These parameters and weights can be found in Appendix A.
@@
 The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
-Stan's default algorithm, the No-U-Turn Sampler (Hoffman & Gelman 2014), allows for direct sampling of the spcified stochastic volatility model.
+Stan's default algorithm, the No-U-Turn Sampler (Hoffman & Gelman 2014), allows for direct sampling of the specified stochastic volatility model.
 Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
-This approach will use the same priors are specified in the Gaussian mixture approximation.
+This approach will use the same priors as specified in the Gaussian mixture approximation.
 However, Hamiltonian does not sample directly from the analytical form of conjugate or conditionally conjugate distribution, so different priors can be chosen if necessary.
@@
 
-Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (Hamiltonian Monte Carlo or Gaussian apprimation in our case).
+Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (Hamiltonian Monte Carlo or Gaussian approximation in our case).
 
@@
 
-Since the distribution of the posterior samples follows the prior distribution, the rank statistic for a given parameter and simultion follows a discrete uniform distribution[^2].
+Since the distribution of the posterior samples follows the prior distribution, the rank statistic for a given parameter and simulation follows a discrete uniform distribution[^2].
 
@@
 # Results
-SBC for Hamiltonian Monte Carlo and the Gaussian approxmiation are performed with simulated dataset sizes of 1000 simulation interations, 1000 observations, and post warmup/burnin samples of 999 (which gives 1000 possible rank statistics).
+SBC for Hamiltonian Monte Carlo and the Gaussian approximation are performed with simulated dataset sizes of 1000 simulation iterations, 1000 observations, and post warmup/burnin samples of 999 (which gives 1000 possible rank statistics).
 The histogram is drawn with 20 bins and 50 observations in each bin which is the expected count for a discrete uniform distribution.
@@
 
-Figure 1 shows the the distribution of rank statistics for Hamiltonian Monte Carlo, with the horizontal black line at 50 representing the discrete uniform distribution.
+Figure 1 shows the distribution of rank statistics for Hamiltonian Monte Carlo, with the horizontal black line at 50 representing the discrete uniform distribution.
 The preliminary figures focus on the static parameters and arbitrarily chosen states (since there is one state parameter per data observation).
 Phi and sigma_sqd look relatively uniform.
-However mu and h.995 have some lumpiness which may suggest a lack of calibration.
+However mu and h.995 (the 995th state parameter) have some lumpiness which may suggest a lack of calibration.
 
@@
 
-One of the drackbacks of SBC is that any deviation from uniform from a calibrated analysis maybe due to finite samples.
-Figure 2 shows more consistent uniform bahviour from all parameters after increasing the number of iterations from 1000 to 5000.
+One of the drawbacks of SBC is that any deviation from uniform from a calibrated analysis maybe due to finite samples.
+Figure 2 shows more consistent uniform behaviour from all parameters after increasing the number of iterations from 1000 to 5000.
 This suggests that Hamiltonian Monte Carlo and overall analysis is well calibrated.
@@
 Figure 3 shows results for the Gaussian approximation.
-The distribution of the rank statistics deviate heavily from the expected discrete uniform distribution, with various peaks at each end of the histogram as well as peaks in the middle.
+The distribution of the rank statistics deviates heavily from the expected discrete uniform distribution, with various peaks at each end of the histogram as well as peaks in the middle.
 This lack of calibration is expected.
-The Gaussian approximiation only approximates the log chi squared distribution, where as Hamiltonian Monte Carlo samples directly from the same likelihood as the data generating process.
+The Gaussian approximation only approximates the log chi squared distribution, whereas Hamiltonian Monte Carlo samples directly from the same likelihood as the data generating process.
+
+This result may also be due to the small number of posterior samples drawn from the model.
+The MCMC sampler may not have converged yet which may explain the deviations from uniformity.
+A follow up step is to compare the effective sample sizes (ESS) of the two methods and to increase the length of the MCMC chain to ensure the ESS is approximately the same before comparing histograms.
 
@@
 
-These chi squared statistics can be considered independent draws from a chi squared distribution with 19 degrees of freedom (since there are 20 bins).
+These chi squared statistics can be considered draws from a chi squared distribution with 19 degrees of freedom (since there are 20 bins).
 Visualising this should give a chi squared distribution (if performing a hypothesis test, this would be the sampling distribution under a null of uniformity).
+One caveat about this is the chi squared statistics calculated in this way may not be independent.
+That is, each iteration of SBC may be independent however, the calculation of the rank statistics for a given iteration may be dependent between some parameters (since each state parameter is a function of the previous state parameter).
 
@@
 
-The models and parametiersations for this research are summarised in the below table:
+The models and parameterisations for this research are summarised in the below table:
 
@@
 
-Week 8: Finish off any simulation or modelling not yet complete (listed in above matrix)
+Week 8: Finish off any simulation or modeling not yet complete (listed in above matrix)
 

### 2023-08-25 analysis/sbc/sbc_results.r
save plots

@@ -2,11 +2,12 @@ library(dplyr)
 library(ggplot2)
 library(future.apply)
 library(posterior)
+library(car)
 source("R/sbc.r")
 plan(multicore, workers = 9)
 
 # Get all parameter data
-path <- "simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r5_1000_iterations"
+path <- "simulation_output/sbc_cp_ksc_model_cp_dgf_r1"
 rds_list <- list.files(path = paste(path, "output", sep = "/"), pattern = "*.RDS")
 results <- lapply(paste("output", rds_list, sep = "/"), get_ranks, model_path = path)
 df <- as.data.frame(do.call("rbind", results))
@@ -28,18 +29,25 @@ state_parameters <- names(rank_bins)[!(names(rank_bins) %in% static_parameters)]
 additional_parameters <- c("h.1.", "h.10.", "h.100.", "h.500.", "h.1000.", "h.995.")
 
 # Facet plots (hist of params)
-facet_hist(data = rank_bins, variables = static_parameters, expected_bin_count = expected_count, nbins = n_bins)
-facet_hist(data = rank_bins, variables = append(static_parameters, additional_parameters), expected_bin_count = expected_count, nbins = n_bins)
+static_hist <- facet_hist(data = rank_bins, variables = static_parameters, expected_bin_count = expected_count, nbins = n_bins)
+ggsave(paste(path, "/static_hist.png", sep =""), static_hist, bg = "white")
+
+static_state_hist <- facet_hist(data = rank_bins, variables = append(static_parameters, additional_parameters), expected_bin_count = expected_count, nbins = n_bins)
+ggsave(paste(path, "/static_state_hist.png", sep =""), static_state_hist, bg = "white")
 
 # Chi squared state parameters (dot plots)
-dot_plots(data = rank_stats,
+chi_sq_dot_plots <- dot_plots(data = rank_stats,
     variables = state_parameters,
     plot_title = "Chi squared estimates for state rank statistics")
+ggsave(paste(path, "/chi_sq_dot_plots.png", sep =""), chi_sq_dot_plots, bg = "white")
 
 # Chi squared state parameters (hist)
-chi_sq_hist(data = rank_stats,
+chi_sq_histogram <- chi_sq_hist(data = rank_stats,
     variables = state_parameters,
[…]

### 2023-08-25 writeups/research_proposal/propsal.qmd
typos

@@
 # 3) Methodology
-The aim of this research is to use Simulation Based Calibration (SBC) to validate the inference of posterior samples generated by different MCMC algorithms in the context of Stochastic Volatility models.
+The aim of this research is to use Simulation Based Calibration (SBC) to validate the inference of posterior samples generated by different MCMC algorithms in the context of stochastic volatility models.
 This will help identify any bias in computation and inconsistencies in model implementation and estimation strategies.
@@
 This approach will use the same priors as specified in the Gaussian mixture approximation.
-However, Hamiltonian does not sample directly from the analytical form of conjugate or conditionally conjugate distribution, so different priors can be chosen if necessary.
+However, Hamiltonian Monte Carlo does not sample directly from the analytical form of conjugate or conditionally conjugate distribution, so different priors can be chosen if necessary.
 
@@
 
-| **1) for n in 1000 iterations:**
+| **1) for sim in 1000 iterations:**
 |   2) Draw from prior: $\theta^{sim}\sim\pi (\theta)$
@@
 
-Figure 4 are the chi squared statistic histograms results of Hamiltonian Monte Carlo and Gaussian approximation respectively.
+Figure 4 and 5 are the chi squared statistic histograms results of Hamiltonian Monte Carlo and Gaussian approximation respectively.
 The vertical black line marks the 95th quantile which represents the critical value for hypothesis testing.
-The HMC chi squared statistics look like they are sampled from the correct distribution, with 6.2\% of the samples greater than the 95\% critical value (doing this calculation for 5000 iterations leads to 4.8\% of samples in the critical area).
+The Hamiltonian Monte Carlo chi squared statistics look like they are sampled from the correct distribution, with 6.2\% of the samples greater than the 95\% critical value (doing this calculation for 5000 iterations leads to 4.8\% of samples in the critical area).
 Figure 5 shows the majority of chi squared statistics for Gaussian approximation within the critical area, suggesting the estimation strategy is not well calibrated.
 
-![Chi squared statistics for HMC](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/chi_sq_hist.png){width=70%}
+![Chi squared statistics for Hamiltonian Monte Carlo](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/chi_sq_hist.png){width=70%}
 
-![Chi squared statistics for KSC](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/chi_sq_hist.png){width=70%}
+![Chi squared statistics for Gaussian approximation](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/chi_sq_hist.png){width=70%}
 

### 2023-08-25 sim_ksc.py
parametiersation

@@ -7,7 +7,6 @@ import numpy as np
 import pandas as pd
 import pyarrow as pa
 import pyarrow.parquet as pq
-# from joblib import Parallel, delayed
 import multiprocessing as mp
 np.set_printoptions(suppress=True, precision=5)
 
@@ -72,9 +71,14 @@ def simulation(seed_data):
     # Iterations
     for s in range(1, config["mcmc_samples"] + 1):
         # Update the parameters of the model
-        mod.update_mixing(trace_mixing[s-1])
         params = np.r_[trace_mu[s-1], trace_phi[s-1], trace_sigma2[s-1]]
-        mod.update(params, transformed=True)
+        mod.update_mixing(indicators = trace_mixing[s-1], 
+                          params = params,
+                          parameterisation = config["parameterisation"])
+                          
+        mod.update(params = params, 
+                   parameterisation = config["parameterisation"],
+                   transformed=True)
 
         # Simulation smoothing
         sim.simulate()

### 2023-08-25 src/gaussianMixture.py

@@ -43,17 +43,29 @@ class TVLLDT(sm.tsa.statespace.MLEModel):
         self['design', 0, 0] = 1 # mod.ssm.design. 
         self['selection', 0, 0] = 1 # mod.ssm.selection
 
-    def update_mixing(self, indicators):
+    def update_mixing(self, indicators, params, parameterisation):
         # z_t | s_t ~ N(m_i - 1.27036, v_i^2)
         # See equation (10), p. 371
         # Indicators is a vector
-        self['obs_intercept', 0] = ksc_params[indicators, 1] - 1.27036
+        if parameterisation == "centered":
+            self['obs_intercept', 0] = ksc_params[indicators, 1] - 1.27036
+        elif parameterisation == "noncentered":
+            self['obs_intercept', 0] = ksc_params[indicators, 1] - 1.27036 + params[0] * (1 - params[1])
+        else:
+            raise ValueError("Incorrect parameterisation")
+
         self['obs_cov', 0, 0] = ksc_params[indicators, 2]
 
-    def update(self, params, **kwargs):
+    def update(self, params, parameterisation, **kwargs):
         params = super(TVLLDT, self).update(params, **kwargs)
-
-        self['state_intercept', 0, 0] = params[0] * (1 - params[1])
+        
+        if parameterisation == "centered":
+            self['state_intercept', 0, 0] = params[0] * (1 - params[1])
+        elif parameterisation == "noncentered":
+            self['state_intercept', 0, 0] = 0
+        else:
+            raise ValueError("Incorrect parameterisation")
+        
         self['transition', 0, 0] = params[1]
         self['state_cov', 0, 0] = params[2]
 

### 2023-08-28 writeups/proposal_presentation/proposal_presentation_sv.qmd
first draft

---
title: "Comparison of MCMC algorithms in Stochastic Volatility Models"
format: 
    beamer:
        incremental: true
        header-includes: |
            \setbeamertemplate{itemize items}[circle]
author: "By Benjamin Wee"
institute: "Supervised by Catherine Forbes and Lauren Kennedy"
---

## Background
- Hello

## Research Goal

Validate the computational methods used to estimate Bayesian stochastic volatility (SV) models

. . .

- 3 key components of research
    + Kim, Sherphard and Chib (1998) SV model
    + Hamiltonian Monte Carlo
    + Simulation Based Calibration to validate MCMC computation

## Stochastic Volatility (SV) model
KSC (1998) estimate a univariate discrete time SV model which models the variance as a latent stochastic process.

. . .

$$
\begin{aligned}
y_t =& \space \beta exp(h_t/2) \epsilon_t \\
h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
h_1 \sim& \space normal\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
\end{aligned}
$$

. . .

:::{.nonincremental}
- $y_t$ is mean corrected returns for equally spaced intervals $t$

- $\mu$ is the "average" log volatility

- $\beta = exp(\mu / 2)$ constant scaling factor (set to 1)

- $h_t$ is log volatility where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the autoregressive parameter $\phi$ such that $|\phi|<1$.

:::

## Kim Sherphard Chib (1998) Estimation strategy
:::{.nonincremental}
- Estimation of states using a Kalman filter, however, this requires the model to be linear and Gaussian.

. . .

- Log and square both sides to linearise the model

. . .

$$
\begin{aligned}
y_t^{*} &= log(y_t^2) \\ 
&= log((\epsilon_t exp(h_t/2))^2) \\
&=  log(exp(h_t)) + log(\epsilon_t^2) \\
&= h_t + log(\epsilon_t^2)  \\
&= h_t + z_t \\
\end{aligned}
$$

Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.

:::

## Kim Sherphard Chib (1998) Estimation strategy
Model is now linear but not Gaussian. KSC use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution through moment matching. This is defined by:

$$
\begin{aligned}
f(z_t) = \sum_{i=1}^{K} q_if_N(z_i|m_i-1.2704, \nu_i^2)
\end{aligned}
$$

K is the mixture of 7 normal densities $f_N$, component probabilities $q_i$, mean $m_i-1.2704$ and variance $\nu_i^2$.

. . .

The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions where as $\phi$ is sampled via a metropolis hastings accept/reject procedure.

## Hamiltonian Monte Carlo (HMC)
- Developed in the physics literature (Duane et al. 1987) before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks (Nael, 1995) and statistical computing (Neal 2011).

- Key innovation: using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore. It takes advantage of the geometry of the posterior to determine its next step.

- The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study. Stan's default algorithm, the No-U-Turn Sampler (Hoffman & Gelman 2014), allows for direct sampling of the specified stochastic volatility model.

## Simulation Based Calibration (SBC)
<!-- Probably need to talk about what calibration means-->
- The procedure starts by taking draws from the prior distribution and creating datasets implied by each draw.

- Rank statistics are then calculated on the posterior samples estimated from the simulated data.

- SBC compares the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.

- If the computation is well calibrated and the samples follow a discrete uniform distribution, then the posterior (credible) intervals will have sufficient coverage. That is, for any percentage interval selected (for example 90%) then there is 90\% chance that any $\theta^{sim}$ falls within this interval.

## Simulation Based Calibration (SBC)
Let $\theta$ be a parameter and $y$ represent the dataset. Start with draw a sample from the prior distribution:
$$
\begin{aligned}
\theta^{sim} \sim \pi(\theta)
\end{aligned}
$$

. . .

Generate a dataset given by the prior draws.

$$
\begin{aligned}
y^{sim} \sim \pi (y|\theta^{sim})
\end{aligned}
$$

. . .

Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (Hamiltonian Monte Carlo or Gaussian approximation in our case).

$$
\begin{aligned}
\{\theta_1,\dots , \theta_{L}\} \sim \pi (\theta | y^{sim})
\end{aligned}
$$

## Simulation Based Calibration (SBC)
A key result from this procedure is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples $\theta^{sim}$. This is implied by the following expression (and proven in Talts et al. (2018)):

$$
\begin{aligned}
\pi(\theta) &= \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})dy^{sim} d\theta^{sim} \\
&= \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) dy^{sim} d\theta^{sim}
\end{aligned}
$$

. . .

That is, the posterior averaged over the joint distribution follows the same distribution as the prior. Therefore, any deviation of the posterior samples from the prior distribution suggests an error - either in computation, code or overall analysis.

<!-- Algorithm, computation, model strategy on average fails to estimate the parameters of a known data generating process. Whether or not the model is fit for a particular use case is not answered here. It's whether the model can capture the _assumed_ structure of the proposed DGP (we can't tell if this DGP is "correct" in the real world) -->

## HMC Results

![1000 Iterations of SBC (HMC)](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/static_state_hist.png){width=80%}

## HMC Results

![5000 Iterations of SBC (HMC)](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r6_5000_iterations/static_state_hist.png){width=75%}

## Gaussian mixture Results

![1000 Iterations of SBC (KSC)](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png){width=75%}

## Next steps
- Consider different model parameterisation
    + Performance of a sampler may be sensititive to the shape of the posterior density

- Consider KSC's correction using importance sampling weights (MH correction step)
    + Reweights samples from approximate distribution such that reweighted draws come from the correct distribution

### 2023-08-29 writeups/proposal_presentation/proposal_presentation_sv.qmd
background

@@
 ## Background
-- Hello
+- Variety of Bayesian estimation strategies for different types of Stochastic Volatility models using classical Markov Chain Monte Carlo (MCMC) algorithms
+
+- Since then, more advanced algorithms such as Hamiltoninan Monte Carlo has enabled the estimation of richer and more complicated models
+
+- However, as the developments of these tools enable the estimation of more complicated models so do the risk of mistakes and the need to validate the output of our algorithms
+
+- How do we ensure the validity of inferences from our algorithms across different estimation strategies?
+
+<!-- What is the fundamental difference between convergence diagnostics, bayes factors with SBC? Check the fit of the model?-->
 
@@
 
-- 3 key components of research
+- 3 key components
     + Kim, Sherphard and Chib (1998) SV model
@@
 h_1 \sim& \space normal\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
+\epsilon_t \sim &N(0,1) \space\space\space \eta_t \sim N(0,1)
 \end{aligned}
@@
 
-- $h_t$ is log volatility where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the autoregressive parameter $\phi$ such that $|\phi|<1$.
+- $h_t$ is log volatility where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ is a stationary process governed by the autoregressive parameter $\phi$ where $|\phi|<1$.
 
@@
 ## Next steps
-- Consider different model parameterisation
+
+- Increasing number of posterior draws for KSC model
+    + MCMC may not have converged onto target distribution
+
+- Different model parameterisation
     + Performance of a sampler may be sensititive to the shape of the posterior density
 
-- Consider KSC's correction using importance sampling weights (MH correction step)
+- KSC's correction using importance sampling weights (MH correction step)
     + Reweights samples from approximate distribution such that reweighted draws come from the correct distribution

### 2023-08-30 writeups/proposal_presentation/proposal_presentation_sv.qmd
reformat

@@
 ## Background
-- Variety of Bayesian estimation strategies for different types of Stochastic Volatility models using classical Markov Chain Monte Carlo (MCMC) algorithms
+- Stochastic volatility models the variance of a financial instrument as a random variable
+
+- Variety of Bayesian estimation strategies using classic Markov Chain Monte Carlo (MCMC) algorithms such as Gibbs and Metropolis Hastings
 
@@
 
-- However, as the developments of these tools enable the estimation of more complicated models so do the risk of mistakes and the need to validate the output of our algorithms
-
-- How do we ensure the validity of inferences from our algorithms across different estimation strategies?
+- As the developments of these tools enable the estimation of more complicated models so do the risk of mistakes and the need to validate the output of our algorithms.
+How do we ensure the validity of our inferences across different estimation strategies?
 
@@
 
-:::{.nonincremental}
+<!-- :::{.nonincremental} -->
 - $y_t$ is mean corrected returns for equally spaced intervals $t$
-
-- $\mu$ is the "average" log volatility
 
@@
 
-:::
+- $\mu$ is the "average" log volatility
+
+<!-- ::: -->
 
@@
 <!-- Probably need to talk about what calibration means-->
-- The procedure starts by taking draws from the prior distribution and creating datasets implied by each draw.
+<!-- - The procedure starts by taking draws from the prior distribution and creating datasets implied by each draw.
 
@@
 
-- SBC compares the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.
+If B computation is correct it can capture the known parameers.-->
+:::{.nonincremental}
+SBC simulates data with known parameters and checks if on average, the interval of an estimated model contains the true parameter
 
-- If the computation is well calibrated and the samples follow a discrete uniform distribution, then the posterior (credible) intervals will have sufficient coverage.
-That is, for any percentage interval selected (for example 90%) then there is 90\% chance that any $\theta^{sim}$ falls within this interval.
+1) Take a draw from the joint prior $\theta^{sim}$
+2) Generate a dataset conditional on the prior draw
+3) Calculate a rank statistics
+4)  Compare the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.
+
+:::
+
+. . .
+
+If Bayesian computation is well calibrated, then the posterior (credible) intervals will have sufficient coverage.
+That is, for any percentage interval selected (for example 90%) then 90\% of these constructed intervals will contain $\theta^{sim}$.
 
@@
 ## Simulation Based Calibration (SBC)
-A key result from this procedure is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples $\theta^{sim}$.
-This is implied by the following expression (and proven in Talts et al.
-(2018)):
+A key result from this procedure is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples.
 
@@
 \begin{aligned}
-\pi(\theta) &= \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})dy^{sim} d\theta^{sim} \\
-&= \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) dy^{sim} d\theta^{sim}
+\pi(\theta) &= \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})dy^{sim} d\theta^{sim} 
+\end{aligned}
+$$
+<!-- &= \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) dy^{sim} d\theta^{sim}
+ -->
+
+That is, the posterior averaged over the joint distribution follows the same distribution as the prior.
+The rank statistic for a given parameter and simulation follows a discrete uniform distribution (proven in Talts et al.
+(2018)).
+
+$$
+\begin{aligned}
+r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim}) = \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}]
 \end{aligned}
@@
 
-. . .
-
-That is, the posterior averaged over the joint distribution follows the same distribution as the prior.
 Therefore, any deviation of the posterior samples from the prior distribution suggests an error - either in computation, code or overall analysis.
@@
 
-## HMC Results
+## HMC Results (1000 iterations)
 
-![1000 Iterations of SBC (HMC)](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/static_state_hist.png){width=80%}
+![](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/static_state_hist.png){fig-align="center" height=90%}
 
-## HMC Results
+## HMC Results (5000 iterations)
 
-![5000 Iterations of SBC (HMC)](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r6_5000_iterations/static_state_hist.png){width=75%}
+![](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r6_5000_iterations/static_state_hist.png){fig-align="center" height=90%}
 
-## Gaussian mixture Results
+## Gaussian mixture Results (1000 iterations)
 
-![1000 Iterations of SBC (KSC)](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png){width=75%}
+![](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png){fig-align="center" height=90%}
 
 ## Next steps
-
+:::{.nonincremental}
 - Increasing number of posterior draws for KSC model
@@
     + Reweights samples from approximate distribution such that reweighted draws come from the correct distribution
+
+:::

### 2023-09-01 writeups/proposal_presentation/proposal_presentation_sv.qmd
Feedback: results, sbc, hmc

@@
 
-- Since then, more advanced algorithms such as Hamiltoninan Monte Carlo has enabled the estimation of richer and more complicated models
+- Since then, more advanced algorithms such as Hamiltonian Monte Carlo has enabled the estimation of richer and more complicated models
 
@@
 \begin{aligned}
-y_t =& \space \beta exp(h_t/2) \epsilon_t \\
+y_t =& \space \beta e^{h_t/2} \epsilon_t \\
 h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
@@
 
-## Kim Sherphard Chib (1998) Estimation strategy
+#  Estimation strategy
+
+## Kim Sherphard Chib (1998)
 :::{.nonincremental}
@@
 ## Hamiltonian Monte Carlo (HMC)
-- Developed in the physics literature (Duane et al.
-1987) before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks (Nael, 1995) and statistical computing (Neal 2011).
+- Key innovations:
 
-- Key innovation: using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
-It takes advantage of the geometry of the posterior to determine its next step.
+    + Uses the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
 
-- The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
-Stan's default algorithm, the No-U-Turn Sampler (Hoffman & Gelman 2014), allows for direct sampling of the specified stochastic volatility model.
+    + Runs physics simulation - imagine the parameter is a marble rolling in random directions inside a bowl
 
-## Simulation Based Calibration (SBC)
-<!-- Probably need to talk about what calibration means-->
-<!-- - The procedure starts by taking draws from the prior distribution and creating datasets implied by each draw.
+    + Allows for direct sampling of the specified stochastic volatility model
 
-- Rank statistics are then calculated on the posterior samples estimated from the simulated data.
+- The Stan programming language's implementation of Hamiltonian Monte Carlo, the No-U-Turn Sampler (Hoffman & Gelman 2014), will be compared with KSC's strategy.
 
-If B computation is correct it can capture the known parameers.-->
-:::{.nonincremental}
-SBC simulates data with known parameters and checks if on average, the interval of an estimated model contains the true parameter
-
-1) Take a draw from the joint prior $\theta^{sim}$
-2) Generate a dataset conditional on the prior draw
-3) Calculate a rank statistics
-4)  Compare the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.
-
-:::
-
-. . .
-
-If Bayesian computation is well calibrated, then the posterior (credible) intervals will have sufficient coverage.
-That is, for any percentage interval selected (for example 90%) then 90\% of these constructed intervals will contain $\theta^{sim}$.
+# Simulation Based Calibration (SBC)
 
@@
 ## Simulation Based Calibration (SBC)
-A key result from this procedure is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples.
-
-$$
-\begin{aligned}
-\pi(\theta) &= \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})dy^{sim} d\theta^{sim} 
-\end{aligned}
-$$
-<!-- &= \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) dy^{sim} d\theta^{sim}
- -->
-
-That is, the posterior averaged over the joint distribution follows the same distribution as the prior.
-The rank statistic for a given parameter and simulation follows a discrete uniform distribution (proven in Talts et al.
-(2018)).
+The rank statistic for a given parameter and simulation follows a discrete uniform distribution (proven in Talts et al. (2018)).
 
@@
 
+If Bayesian computation is well calibrated, then the posterior (credible) intervals will have sufficient coverage.
+That is, for any percentage interval selected (for example 90%) then 90\% of these constructed intervals will contain $\theta^{sim}$.
+
 <!-- Algorithm, computation, model strategy on average fails to estimate the parameters of a known data generating process.
@@
 It's whether the model can capture the _assumed_ structure of the proposed DGP (we can't tell if this DGP is "correct" in the real world) -->
+
+# Results
 
@@
 
-- KSC's correction using importance sampling weights (MH correction step)
-    + Reweights samples from approximate distribution such that reweighted draws come from the correct distribution
+- KSC's correction using importance sampling weights
+    + Reweight samples from Gaussian mixture so that the new draws come from the correct distribution
 
 :::
+
+# Appendix
+
+## Simulation Based Calibration (SBC)
+A key result from this procedure is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples.
+
+$$
+\begin{aligned}
+\pi(\theta) &= \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})dy^{sim} d\theta^{sim}  \\
+&= \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) dy^{sim} d\theta^{sim}
+\end{aligned}
+$$
+
+That is, the posterior averaged over the joint distribution follows the same distribution as the prior.
+The rank statistic for a given parameter and simulation follows a discrete uniform distribution (proven in Talts et al.
+(2018)).

### 2023-09-01 writeups/proposal_presentation/sandp500.r
init finance plot

@@ -1,0 +1,3 @@
+library(ggplot2)
+df <- read.csv("data/preprocessed/GSPC/19900101_20230201.csv")
+ggplot(df, aes(x=as.Date(Date), y = log_return)) + geom_line()

### 2023-09-04 R/sbc.r
updated font size

@@ -30,9 +30,11 @@ facet_hist <- function(data, variables, nbins, expected_bin_count) {
         geom_histogram(aes(rank), bins = nbins, fill = "light blue", alpha = 0.7) +
         scale_x_continuous(labels = function(x) x * expected_bin_count) +
         facet_wrap(~variable) +
-        theme_minimal() +
+        theme_minimal(base_size = 22) +
         geom_hline(yintercept = expected_bin_count, size = 0.5, alpha = 0.3) +
-        labs(title = "Distribution of Rank Statistics")
+        labs(title = "Distribution of Rank Statistics",
+                y = "Count",
+                x = "Ranks")
 }
 
 dot_plots <- function(data, variables, plot_title) {
@@ -45,7 +47,7 @@ dot_plots <- function(data, variables, plot_title) {
         geom_segment(aes(x = parameter, xend = parameter, y = 0, yend = chisq_stat), color = "grey") +
         geom_point(aes(x = parameter, y = chisq_stat), size = 3, colour = "light blue") +
         coord_flip() +
-        theme_minimal() +
+        theme_minimal(base_size = 22) +
         theme(axis.title.y = element_blank(),
             axis.text.y = element_blank(),
             axis.ticks.y = element_blank()) +
@@ -62,7 +64,7 @@ chi_sq_hist <- function(data, variables, plot_title) {
         mutate(parameter = factor(parameter, unique(parameter))) %>%
         ggplot(.) +
         geom_histogram(aes(x = chisq_stat), fill = 'blue', alpha = 0.4) +
-        theme_minimal() +
+        theme_minimal(base_size = 22) +
         theme(axis.title.y = element_blank(),
             axis.text.y = element_blank(),
             axis.ticks.y = element_blank()) +

### 2023-09-04 writeups/proposal_presentation/proposal_presentation_sv.qmd
updated based on feedback

@@
 
-## Background
-- Stochastic volatility models the variance of a financial instrument as a random variable
+# What is Stochastic Volatility?
 
-- Variety of Bayesian estimation strategies using classic Markov Chain Monte Carlo (MCMC) algorithms such as Gibbs and Metropolis Hastings
+## Log returns S&P 500
+![]("returns.png")
 
-- Since then, more advanced algorithms such as Hamiltonian Monte Carlo has enabled the estimation of richer and more complicated models
+## Stochastic Volatility (SV) model
 
-- As the developments of these tools enable the estimation of more complicated models so do the risk of mistakes and the need to validate the output of our algorithms.
-How do we ensure the validity of our inferences across different estimation strategies?
+$$
+\begin{aligned}
+y_t =& \space \beta e^{h_t/2} \epsilon_t \\
+h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
+h_1 \sim&  N\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
+\epsilon_t \sim& N(0,1) \quad \eta_t \sim N(0,1) \\
+\end{aligned}
+$$
 
-<!-- What is the fundamental difference between convergence diagnostics, bayes factors with SBC? Check the fit of the model?-->
+- $y_t$ is mean corrected returns for equally spaced intervals $t$
+
+- $\beta = e^{\mu / 2}$ constant scaling factor (set to 1)
+
+- $h_t$ is log volatility where $h_1$ is a draw from a stationary distribution
+
+- State equation $h_{t+1}$ is a stationary process where $\mu$ is the "average" log volatility and determined by the autoregressive parameter $\phi$ where $|\phi|<1$.
 
@@
 
-- 3 key components
-    + Kim, Sherphard and Chib (1998) SV model
-    + Hamiltonian Monte Carlo
-    + Simulation Based Calibration to validate MCMC computation
+- Compare the Kim, Shephard and Chib (1998) estimation strategy with Hamiltonian Monte Carlo
+
+- Simulation Based Calibration to validate MCMC computation
+
+# Estimation strategy
 
@@
 KSC (1998) estimate a univariate discrete time SV model which models the variance as a latent stochastic process.
-
-. . .
 
@@
 h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
-h_1 \sim& \space normal\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
-\epsilon_t \sim &N(0,1) \space\space\space \eta_t \sim N(0,1)
+h_1 \sim& N\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
+\epsilon_t \sim& N(0,1) \quad \eta_t \sim N(0,1) \\
 \end{aligned}
@@
 
-<!-- :::{.nonincremental} -->
-- $y_t$ is mean corrected returns for equally spaced intervals $t$
+With priors:
+$$
+\begin{aligned}
+\mu \sim& \space N(0, 10^2) \\
+\sigma_{\eta}^2 \sim& \space IG(5/2, (0.01\times 5) / 2) \\
+\phi^{\ast} \sim& \space beta(20, 1.5) \\
+\phi =& 2\phi^{\ast} - 1
+\end{aligned}
+$$
 
-- $\beta = exp(\mu / 2)$ constant scaling factor (set to 1)
-
-- $h_t$ is log volatility where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ is a stationary process governed by the autoregressive parameter $\phi$ where $|\phi|<1$.
-
-- $\mu$ is the "average" log volatility
-
-<!-- ::: -->
-
-#  Estimation strategy
-
-## Kim Sherphard Chib (1998)
+## Kim Shephard Chib (1998)
 :::{.nonincremental}
-- Estimation of states using a Kalman filter, however, this requires the model to be linear and Gaussian.
+- Estimation of states using a standard Kalman filter, however, this requires the model to be linear and Gaussian.
 
@@
 
-## Kim Sherphard Chib (1998) Estimation strategy
+## Kim Shephard Chib (1998)
 Model is now linear but not Gaussian. KSC use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution through moment matching. This is defined by:
@@
 
-    + Runs physics simulation - imagine the parameter is a marble rolling in random directions inside a bowl
+    + Imagine the parameter is a marble rolling in random directions inside a bowl (negative log transformation of joint posterior)
 
@@
 
-Therefore, any deviation of the posterior samples from the prior distribution suggests an error - either in computation, code or overall analysis.
+. . .
+
+Therefore, if the distribution of rank statistics deviates from a discrete uniform distribution, there may be evidence that our analysis or computation is not returning the correct posteriors (not calibrated).
+
+. . .
 

### 2023-09-04 writeups/research_proposal/propsal.qmd
updated based on feedback

@@
 title: "Comparison of MCMC algorithms in Stochastic Volatility Models"
+author: 
+    - "By Benjamin Wee" 
+    - "Supervised by Catherine Forbes and Lauren Kennedy"
+# abstract: "The objective of this research is to validate and compare the computational methods used to estimate Bayesian stochastic volatility models.
+Specifically, Simulation Based Calibration will be used to identify biases and errors arising from the results of Markov Chain Monte Carlo (MCMC) algorithms.
+The source of any miscalibration include, but are not limited to, estimation strategies, choice of sampler, model specifications and coding mistakes.
+This study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
+Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches."
 format:
@@
 geometry: margin=1in
+mathspec: true
 ---
@@
 
-## Research Goal
+## 1. Research Goal
 The objective of this research is to validate and compare the computational methods used to estimate Bayesian stochastic volatility models.
@@
 
-# Background and Literature Review
+# 2. Background and Literature Review
 
@@
 
-# 3) Methodology
+# 3. Methodology
 The aim of this research is to use Simulation Based Calibration (SBC) to validate the inference of posterior samples generated by different MCMC algorithms in the context of stochastic volatility models.
@@
 
-## 3.1) Stochastic Volatility Model
+## 3.1 Stochastic Volatility Model
 The univariate stochastic volatility model described by Kim, Shephard and Chib (1998) models the variance as a random variable following some latent stochastic process.
@@
 
-## 3.2) Estimation Strategy
+## 3.2 Estimation Strategy
 
@@
 
-# Results
+# 4. Results
 SBC for Hamiltonian Monte Carlo and the Gaussian approximation are performed with simulated dataset sizes of 1000 simulation iterations, 1000 observations, and post warmup/burnin samples of 999 (which gives 1000 possible rank statistics).
@@
 
-# Conclusion and next steps
+# 5. Conclusion and next steps
 There are two key extensions to the simulation results worth exploring.
@@
 
+{{< pagebreak >}}
+
 # References
@@
 
+{{< pagebreak >}}
+
 # Appendix A: Mixture Gaussian weights
 
-# Appendix B: Conjugate posterior distributions
+| $\omega$ | $Pr(\omega = i)$ | $m_i$ | $\nu^2_i$ |
+|------|------|------|------|
+| 1    | 0.00730    | -10.12999   | 5.79596 | 
+| 2    | 0.10556    | -3.97281    | 2.61369 | 
+| 3    | 0.00002    | -8.56686    | 5.17950 | 
+| 4    | 0.04395    | 2.77786    | 0.16735 | 
+| 5    | 0.34001    | 0.61942    | 0.64009 | 
+| 6    | 0.24566    | 1.79518    | 0.34023 | 
+| 7    | 0.25750    | -1.08819    | 1.26261 |
+
+{{< pagebreak >}}
+
+# Appendix B: Sampling from mixture of Gaussians
+
+## Conjugate posterior distributions
+
+**Sampling** $\boldsymbol{\sigma_{\eta}^2}$
+
+Inverse gamma conjugate posterior distribution:
+
+$$
+\sigma^2_{\eta} | y,h,\phi,\mu \sim IG \Bigl\{\frac{n+\sigma_r}{2}, \frac{0.05 + (h_1 - \mu)^2 (1 - \phi^2) + \sum_{t=1}^{n-1}((h_{t+1} - \mu) - \phi(h_t - \mu))^2}{2}\Bigr\}
+$$
+
+**Sampling** $\boldsymbol{\mu}$
+
+Gaussian conjugate posterior distribution:
+
+$$
+\mu | h,\phi,\sigma^2_{\eta}  \sim N(\hat{\mu}, \sigma^2_{\mu})
+$$
+
+Where
+
+$$
+\begin{aligned}
+\hat{\mu} &= \sigma^2_{\mu} \Bigl\{\frac{(1-\phi^2)}{\sigma_{\eta}^2}h_1 +\frac{(1-\phi^2)}{\sigma_{\eta}^2} \sum_{t=1}^{n-1} (h_{t+1} - \phi h_t)\Bigr\} \\
+\sigma^2_{\mu} &= \sigma^2_{\eta} \{(n-1)(1-\phi)^2 + (1-\phi^2)\}^{-1}
+\end{aligned}
+$$
+
+**Sampling** $\boldsymbol{\phi}$
+
+Metropolis Hastings accept/reject procedure:
+
+1) Generate proposal $\phi^\ast$ from $N(\hat{\phi}, V_{\phi})$ where $\hat{\phi} = \frac{\sum_{t=1}^{n-1} (h_{t+1} - \mu)(h_t - \mu)}{\sum_{t=1}^{n-1} (h_t - \mu)^2}$ and $V_{\phi} = \sigma^2_{\eta} \{\sum_{t=1}^{n-1} (h_t - \mu)^2\}^{-1}$
+
+2) Accept proposal as $\phi^{(i)}$ with probability $e^{\{g(\phi^\ast) - g(\phi^{(i-1)}\}}$ such that $g(\phi) = log (\pi (\phi)) - \frac {(h_t - \mu)^2 (1-\phi^2)}{2 \sigma_{\eta}^2} + \frac{1}{2} log (1-\phi^2)$

### 2023-09-05 R/sbc.r
updated diagnostics

@@ -100,17 +100,26 @@ load_parameters <- function(rds_path) {
 
 get_rhat_basic <- function(rds_path, model_path) {
     data = readRDS(here::here(paste(model_path, rds_path, sep = "/")))
-    return(data$all_chains$rhat_basic)
+    diagnostics = do.call("rbind", data$all_chains$diagnostics)
+    results <- as.data.frame(cbind(rownames(diagnostics), diagnostics$rhat_basic))
+    names(results) <- c("parameters", "rhat_basic")
+    return(results)
 }
 
 get_rhat <- function(rds_path, model_path) {
     data = readRDS(here::here(paste(model_path, rds_path, sep = "/")))
-    return(data$all_chains$rhat)
+    diagnostics = do.call("rbind", data$all_chains$diagnostics)
+    results <- as.data.frame(cbind(rownames(diagnostics), diagnostics$rhat))
+    names(results) <- c("parameters", "rhat")
+    return(results)
 }
 
 get_ess <- function(rds_path, model_path, ess_type, chains) {
     data = readRDS(here::here(paste(model_path, rds_path, sep = "/")))
-    return(data[[chains]][[ess_type]])
+    diagnostics = data[[chains]][["diagnostics"]]
+    results <- as.data.frame(cbind(rownames(diagnostics), diagnostics[ess_type]))
+    names(results) <- c("parameters", ess_type)
+    return(results)
 }
 
 get_time <- function(rds_path, model_path) {

### 2023-09-07 writeups/research_proposal/propsal.qmd
restructured proposal

@@
 
-## 1. Research Goal
-The objective of this research is to validate and compare the computational methods used to estimate Bayesian stochastic volatility models.
-Specifically, Simulation Based Calibration will be used to identify biases and errors arising from the results of Markov Chain Monte Carlo (MCMC) algorithms.
-The source of any miscalibration include, but are not limited to, estimation strategies, choice of sampler, model specifications and coding mistakes.
-This study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
-Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches.
-
 <!-- and provide tools for evaluating other types models and algorithms.  -->
@@
 
+## 1. Research Goal
+
+The objective of this research is to compare the computational methods used to estimate Bayesian stochastic volatility models.
+A simulation study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo and their ability to sample from the model's posterior distribution.
+Specifically, Simulation Based Calibration (SBC) is used to check whether these sampling strategies are returning efficient and well calibrated posterior estimates.
+Key metrics of interest are the effective sample size to check the efficiency of the algorithm and tests of uniformity to assess the calibration of the posteriors.
+This will determine which algorithm is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
+
+<!-- Stochastic volatility models are used to explicitly model the variance of returns in financial instruments and can be estimated using frequentist or Bayesian frameworks.  -->
+
+<!-- 
+The objective of this research is to validate the computational methods used to estimate Bayesian stochastic volatility models.
+This study will assess the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
+A simulation study will compare the ability of these methods to sample from the model's posterior distribution.
+Specifically, Simulation Based Calibration (SBC) is used to check whether these sampling strategies are returning efficient and well calibrated posterior estimates.
+Key metrics to check these features are the effective sample size to check efficiency of the algorithm and tests of uniformity to determine calibration of the posteriors.
+This will determine which algorithm is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
+-->
+
+<!-- Simulation Based Calibration will be used to identify biases and errors arising from the results of Markov Chain Monte Carlo (MCMC) algorithms.
+The source of any miscalibration include, but are not limited to, estimation strategies, choice of sampler, model specifications and coding mistakes.
+This study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
+Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches.
+-->
+
+<!-- The objective of this research is to validate and compare the computational methods used to estimate Bayesian stochastic volatility models.
+Specifically, Simulation Based Calibration will be used to identify biases and errors arising from the results of Markov Chain Monte Carlo (MCMC) algorithms.
+The source of any miscalibration include, but are not limited to, estimation strategies, choice of sampler, model specifications and coding mistakes.
+This study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
+Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches.
+-->
+
+<!-- Specifically, they propose a method of sampling by using an offset mixture normal approximation with a reweighting procedure to sample from the posterior distribution.
+By linearing the model and using a mixture of Gaussians, they were able to apply a Kalman Filter to estimate the posterior density over the latent volatilities and static parameters.
+-->
+
+<!-- The univariate stochastic volatility model described in their paper outlines the state space representation of the the model which is described below.
+ -->
 # 2. Background and Literature Review
 
-Kim, Sherpherd and Chib (1998) estimate the discrete time stochastic volatility model by proposing a simulation strategy using Bayesian methods.
-Specifically, they propose a method of sampling by using an offset mixture normal approximation with a reweighting procedure to sample from the posterior.
-By linearing the model and using a mixture of Gaussians, they were able to apply a Kalman Filter to estimate the posterior density over the latent volatilities and static parameters.
-This is compared with other estimation strategies and volatility models such as autoregressive conditional heteroskedasticity (ARCH).
-Their specification of the stochastic volatility model and estimation approaches provide the context for comparing the performance of different MCMC algorithms.
+## 2.1 Stochastic Volatility
+Stochastic volatility models are used in financial econometrics to model the variance or volatility of returns in a financial instrument.
+The return of a financial asset often exhibits pattterns in the variance (heteroskedasticity) which can be modelled to improve prediction and estimates of risk or tail events.
+Unlike other volatility models such as ARCH or GARCH, these stochastic volatility models treat the variance as a random variable.
+Furthermore, these models have a state space representation since the variance is treated as a latent stochastic process.
+State space models can be difficult to estimate since there is a latent state for every data observation in the time series, which means there are more parameters than data points.
+These models can be estimated using both frequent tools but this research will focus on Bayesian estimation strategies.
 
-This strategy relies on sampling directly from conjugate posterior distributions and the use of classic MCMC algorithms, Gibbs and Random Walk Metropolis Hastings.
+Kim, Sherpherd and Chib (1998) estimate a discrete time, univariate stochastic volatility model using Bayesian methods which is outlined below.
+$y_t$ is the mean corrected returns of some asset for equally spaced intervals t.
+$\beta$ is a constant scaling factor which is also defined as $exp(\mu / 2)$ representing instantaneous volatility.
+$h_t$ is log volatility, where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the autoregressive parameter $\phi$ such that $|\phi|<1$.
+This autoregressive parameter represents the persistence or "stickiness" of log volatility and the dispersion parameter $\sigma_{\eta}$ is its variance.
+$\epsilon_t$ and $\eta_t$ are standard normal white noise shocks.
+
+$$
+\begin{aligned}
+y_t =& \space \beta exp(h_t/2) \epsilon_t \\
+h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
+h_1 \sim& \space normal\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
+\end{aligned}
+$$
+
+$$
+\begin{aligned}
+\epsilon_t \sim& \space normal(0,1) \\
+\delta_t \sim& \space normal(0,1)
+\end{aligned}
+$$
+
+Setting $\beta=1$, the model can be expressed more succinctly as:
+
+$$
+\begin{aligned}
+y_t \sim& \space normal(0, exp(h_t/2)) \\ 
+h_1 \sim& \space normal \left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
+h_{t+1} \sim& \space normal(\mu +\phi(h_t - \mu) , \sigma_{\eta}^2), \space\space t\neq 1\\ 
+\end{aligned}
+$$
+
+Priors for the static parameters are defined below with conjugate priors on $\mu$ and $\sigma^2$:
+
+$$
+\begin{aligned}
+\mu \sim& \space normal(0, 10^2) \\
+\sigma_{\eta}^2 \sim& \space IG(5/2, (0.01\times 5) / 2) \\
+\phi^{\ast} \sim& \space beta(20, 1.5) \\
+\phi &=  2\phi^{\ast} - 1
+\end{aligned}
+$$
+
+The prior on $\phi$ is a "stretched" beta distribution. This is a beta distribution (as defined on the parameter $\phi^*$) which has been transformed to have support (-1, 1).
+
+### Estimation strategy
+Kim, Shephard and Chib's sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother[^1] (de Jong and Shephard, 1995).
+
+[^1]: In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.
+
+The standard Kalman Filter and smoother is used to compute the posterior distribution over the latent states.
+This requires the state and measurement equations to be linear and conditionally Gaussian.
+Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring and taking the log of $y_t$.
+
+$$
+\begin{aligned}
+y_t^{*} &= log(y_t^2) \\ 
+&= log((\epsilon_t exp(h_t/2))^2) \\
+&=  log(exp(h_t)) + log(\epsilon_t^2) \\
+&= h_t + log(\epsilon_t^2)  \\
+&= h_t + z_t \\
+\end{aligned}
+$$
+
+Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
+However, it is not simple to sample from this parameterisation of the model.
+Kim, Sherpherd and Chib use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution.
+This is defined by:
+
+$$
+\begin{aligned}
+f(z_t) = \sum_{i=1}^{K} q_if_N(z_i|m_i-1.2704, \nu_i^2)
+\end{aligned}
+$$
+
+Where K is the mixture of normal densities $f_N$, component probabilities $q_i$, mean $m_i-1.2704$ and variance $\nu_i^2$.
+These parameters were selected using moment matching where they found 7 normal densities with varying mean and variance parameters best approximated the log chi squared moments.
+These parameters and weights can be found in Appendix A.
+
+The model can be sampled via the simulation smoother since the model is now linear and conditionally gaussian.
+The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions where as $\phi$ is sampled via a metropolis hastings accept/reject procedure.
+The details can be be around in Appendix B.
+
+## 2.2 Hamiltoninan Monte Carlo
 Since the paper was written, new MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
@@
 
-New developments in MCMC algorithms and improved computing resources have enabled estimation of richer and more complicated models.
+The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
+Stan's default algorithm, the No-U-Turn Sampler (Hoffman & Gelman 2014), allows for direct sampling of the specified stochastic volatility model.
+Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
+This approach will use the same priors as specified in the Gaussian mixture approximation.
+However, Hamiltonian Monte Carlo does not sample directly from the analytical form of conjugate or conditionally conjugate distribution, so different priors can be chosen if necessary.
+
+## 2.3 Research contribution
+The goal of this research is to determine which algorithm is better at estimating stochastic volatility models.
+There are clear differences between Kim Sherphard and Chib's estimation strategy and Hamiltoninan Monte Carlo.
+My main contribution is to compare these algorithms and determine which of these sampling strategies is better based on whether the algorithm returns calibrated posterior estimates and the efficiency of the MCMC sampler.
+
+<!-- New developments in MCMC algorithms and improved computing resources have enabled estimation of richer and more complicated models.
 However, as the development of tools enables us to fit more complicated models, so does the need to validate the output of these tools.
@@
 build upon this work by developing Simulation Based Calibration to provide more robust validation of output from a MCMC sampler through the use of rank statistics.
+-->
 
@@
 # 3. Methodology
+
+## 3.1 Design
+Simulation Based Calibration (SBC) is a simulation technique that will help check the calibration of posterior estimates generated by different MCMC algorithms.
+SBC is conducted by comparing the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.
+The procedure starts by taking draws from the prior distribution and creating datasets implied by each draw.
+Rank statistics are then calculated on the posterior samples estimated from the simulated data.
+Let $\theta$ be a parameter and $y$ represent the dataset.
+Start with draw a sample from the prior distribution:
+$$
+\begin{aligned}
+\theta^{sim} \sim \pi(\theta)
+\end{aligned}
+$$
+
+Generate a dataset given by the prior draws.
+
+$$
+\begin{aligned}
+y^{sim} \sim \pi (y|\theta^{sim})
+\end{aligned}
+$$
+
+Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (Hamiltonian Monte Carlo or Gaussian approximation in our case).
+
+$$
+\begin{aligned}
+\{\theta_1,\dots , \theta_{L}\} \sim \pi (\theta | y^{sim})
+\end{aligned}
+$$
+
+A key result from this procedure is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples $\theta^{sim}$.
+This is implied by the following expression:
+
+$$
+\begin{aligned}
+\pi(\theta) &= \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})dy^{sim} d\theta^{sim} \\
+&= \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) dy^{sim} d\theta^{sim}
+\end{aligned}
+$$
+
+That is, the posterior averaged over the joint distribution follows the same distribution as the prior.
+The procedure of generating posterior samples implicitly performs this integral since the expression on the right of the integral is proportional to the prior density.
+Therefore, any deviation of the posterior samples from the prior distribution suggests an error - either in computation, code or overall analysis.
+
+<!-- The simulation if done correctly should produce samples from the prior distribution.  -->
+
+Since the distribution of the posterior samples follows the prior distribution, the rank statistic for a given parameter and simulation follows a discrete uniform distribution[^2].
+
+[^2]: Proof of this result in Talts et al 2018.
+
+$$
+\begin{aligned}
+r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim}) = \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}]
+\end{aligned}
+$$
+
+This completes one iteration of SBC.
+To complete the algorithm, multiple iterations are run and the rank statistics are calculated for each parameter.
+The resulting rank statistics are compared to the discrete uniform distribution to determine if any problematic features or errors exist.
+
+If the computation is well calibrated and the samples follow a discrete uniform distribution, then the posterior (credible) intervals will have sufficient coverage.
+That is, for any percentage interval selected (for example 90%) then there is 90\% chance that any $\theta^{sim}$ falls within this interval.
+In other words, a Bayesian analysis and computation is well calibrated if the 90\% interval contains the true parameter in 90\% interations of the SBC procedure.
+
+## 3.2 Implementation
+SBC for Hamiltonian Monte Carlo and the Gaussian approximation are performed with simulated dataset sizes of 1000 simulation iterations, 1000 observations, and post warmup/burnin samples of 999 (which gives 1000 possible rank statistics).
+This algorithm is summarised below:
+
+| **1) for sim in 1000 iterations:**
+|   2) Draw from prior: $\theta^{sim}\sim\pi (\theta)$
+|   3) Simulate dataset with 1000 observations: $y^{sim} \sim p(y|\theta^{sim})$
+|   4) Draw 999 posterior samples (post warmup) $\{\theta_1,\dots , \theta_{L}\} \sim p(\theta | y^{sim})$
+|   5) Compute rank statistics $r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim})$
+
+## 3.3 Metrics
+The key metrics and diagnostics to compare the performance of these samplers is the effective sample size, histograms of rank statistics and chi-squared test statistics.
+
+Effecitve sample size (ESS) measures the efficiency of the MCMC sampler.
+It calculates the number of (effectively) independent draws from the posterior draws generated by a Markov chain.
+A poor ESS can arise from high autocorrelation in the Markov chain which leads to highly dependent samples.
+If it requires long MCMC chains to get to a desired ESS, then this is considered inefficient since it takes more resources to get a representative sample of the target distribution.
+
+Histograms of rank statistics are a visual diagnostic to determine if the distribution of ranks follows a discrete uniform distribution.
+As described, a calibrated algorithm implies rank statistics which are randomly drawn from a uniform.
+A weakness of this approach is there are at least as many parameters as there data points.
+Since there are 1000 observations per dataset, this means there are 1003 parameters to estimate.
+
+An alternative to checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each bin.
+Let $b_j$ be the number of counts and $e_j$ the expected count in bin $j$.
+Then the chi squared statistic is given by:
+
+$$
+\begin{aligned}
+\chi^2 = \sum_{j=1}^J \frac{(b_{j} - e_{j})^2}{e_j}
+\end{aligned}
+$$
+
+The chi squared statistics provide a more succinct summary of the overall performance of the algorithm.
+
+<!-- 
 The aim of this research is to use Simulation Based Calibration (SBC) to validate the inference of posterior samples generated by different MCMC algorithms in the context of stochastic volatility models.
@@
 Comparison of the SBC results between different computational approaches may help identify any strengths or shortcomings as well as any errors in model specification.
+-->
 
-## 3.1 Stochastic Volatility Model
+<!-- ## 3.1 Stochastic Volatility Model
 The univariate stochastic volatility model described by Kim, Shephard and Chib (1998) models the variance as a random variable following some latent stochastic process.
@@
 However, Hamiltonian Monte Carlo does not sample directly from the analytical form of conjugate or conditionally conjugate distribution, so different priors can be chosen if necessary.
+ -->
 
-## 3.3 Simulation Based Calibration (SBC)
 <!-- SBC validates posterior samples and inferences generated by MCMC algorithms. It identifies inconsistencies and bias in computation and estimation of a generative model. This procedure  -->
-
-SBC is conducted by comparing the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.
-The procedure starts by taking draws from the prior distribution and creating datasets implied by each draw.
-Rank statistics are then calculated on the posterior samples estimated from the simulated data.
-Let $\theta$ be a parameter and $y$ represent the dataset.
-Start with draw a sample from the prior distribution:
-$$
-\begin{aligned}
-\theta^{sim} \sim \pi(\theta)
-\end{aligned}
-$$
-
-Generate a dataset given by the prior draws.
-
-$$
-\begin{aligned}
-y^{sim} \sim \pi (y|\theta^{sim})
-\end{aligned}
-$$
-
-Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (Hamiltonian Monte Carlo or Gaussian approximation in our case).
-
-$$
-\begin{aligned}
-\{\theta_1,\dots , \theta_{L}\} \sim \pi (\theta | y^{sim})
-\end{aligned}
-$$
-
-A key result from this procedure is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples $\theta^{sim}$.
-This is implied by the following expression:
-
-$$
-\begin{aligned}
-\pi(\theta) &= \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})dy^{sim} d\theta^{sim} \\
-&= \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) dy^{sim} d\theta^{sim}
-\end{aligned}
-$$
-
-That is, the posterior averaged over the joint distribution follows the same distribution as the prior.
-The procedure of generating posterior samples implicitly performs this integral since the expression on the right of the integral is proportional to the prior density.
-Therefore, any deviation of the posterior samples from the prior distribution suggests an error - either in computation, code or overall analysis.
-
-<!-- The simulation if done correctly should produce samples from the prior distribution.  -->
-
-Since the distribution of the posterior samples follows the prior distribution, the rank statistic for a given parameter and simulation follows a discrete uniform distribution[^2].
-
-[^2]: Proof of this result in Talts et al 2018.
-
-$$
-\begin{aligned}
-r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim}) = \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}]
-\end{aligned}
-$$
-
-This completes one iteration of SBC.
-To complete the algorithm, multiple iterations are run and the rank statistics are calculated for each parameter.
-The resulting rank statistics are compared to the discrete uniform distribution to determine if any problematic features or errors exist.
-
-If the computation is well calibrated and the samples follow a discrete uniform distribution, then the posterior (credible) intervals will have sufficient coverage.
-That is, for any percentage interval selected (for example 90%) then there is 90\% chance that any $\theta^{sim}$ falls within this interval.
-In other words, a Bayesian analysis and computation is well calibrated if 90\% of constructed intervals contain the "true" parameter 90\% of the time.
 
@@
 
-# 4. Results
-SBC for Hamiltonian Monte Carlo and the Gaussian approximation are performed with simulated dataset sizes of 1000 simulation iterations, 1000 observations, and post warmup/burnin samples of 999 (which gives 1000 possible rank statistics).
-The histogram is drawn with 20 bins and 50 observations in each bin which is the expected count for a discrete uniform distribution.
-This algorithm is summarised below:
-
-| **1) for sim in 1000 iterations:**
-|   2) Draw from prior: $\theta^{sim}\sim\pi (\theta)$
-|   3) Simulate dataset with 1000 observations: $y^{sim} \sim p(y|\theta^{sim})$
-|   4) Draw 999 posterior samples (post warmup) $\{\theta_1,\dots , \theta_{L}\} \sim p(\theta | y^{sim})$
-|   5) Compute rank statistics $r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim})$
-
+# 4. Preliminary Results
 Figure 1 shows the distribution of rank statistics for Hamiltonian Monte Carlo, with the horizontal black line at 50 representing the discrete uniform distribution.
@@
 
-An alternative to checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each bin.
+<!-- An alternative to checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each bin.
 Let $b_j$ be the number of counts and $e_j$ the expected count in bin $j$.
@@
 
-![Chi squared statistics for Gaussian approximation](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/chi_sq_hist.png){width=70%}
+![Chi squared statistics for Gaussian approximation](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/chi_sq_hist.png){width=70%} -->
 

### 2023-09-09 writeups/proposal_presentation/sandp500.r
presentation plot

@@ -1,3 +1,16 @@
 library(ggplot2)
 df <- read.csv("data/preprocessed/GSPC/19900101_20230201.csv")
-ggplot(df, aes(x=as.Date(Date), y = log_return)) + geom_line()
+
+ts <- ggplot(df[!is.na(df['log_return']), ], aes(x=as.Date(Date), y = log_return)) + 
+    geom_line(colour = 'blue') +
+    theme_minimal() +
+    labs(x = "Date",
+         y = "Log Return") + 
+    theme(axis.title.x = element_text(size=20),
+          axis.title.y = element_text(size=20), #angle = 0, vjust = 0.5),
+          axis.text.x = element_text(size=16),
+          axis.text.y = element_text(size=16))
+
+ggsave(paste(here::here(), "/writeups/proposal_presentation/returns.png", sep =""), ts, bg = "white")
+
+# plot(rnorm(2e4),type="l")

### 2023-09-09 sim_ksc.py
fixed burn in

@@ -97,7 +97,7 @@ def simulation(seed_data):
     draws = np.concatenate((trace_mu, trace_phi, trace_sigma2, trace_states), axis = 1)
     static_names = ['mu', 'phi', 'sigma2']
     state_names = [f"h[{state}]" for state in np.arange(1, trace_states.shape[1]+1)]
-    samples = pd.DataFrame(draws[config["burn"]:draws.shape[0]])
+    samples = pd.DataFrame(draws[config["burn"]+1:draws.shape[0]])
     samples.columns = static_names + state_names
 
     # Save as parquet

### 2023-09-09 analysis/gauss_mix/gma.py
Reweighting functions

@@ -1,7 +1,7 @@
 from __future__ import division
 
 import numpy as np
-import pandas as pd
+import scipy as sp
 import statsmodels.api as sm
 
 from statsmodels.tsa.statespace.tools import (
@@ -140,3 +140,31 @@ def draw_mixing(mod, states):
     sample = np.argmax(tmp, axis=1)
 
     return sample
+
+def mixture_loglike_v(data,states, burn):
+    state_val = states[burn+1:states.shape[0]]
+    norm_list = [ksc_params[i,0] * norm.pdf(x=data, loc=state_val+ksc_params[i,1]- 1.2704, scale=ksc_params[i,2]**0.5) for i in range(0,7)]
+    log_kn = np.log(sum(norm_list))
+    sum_log_kn = np.sum(log_kn, axis = 1)
+    return sum_log_kn
+
+def logchi2_loglike(x):
+    return np.sum(np.log(sp.stats.chi2.pdf(np.exp(x), df = 1) * np.exp(x)))
+
+def logchi2_loglike_v(data, states, burn):
+    index = np.arange(burn+1, states.shape[0])
+    sum_log_fn = [logchi2_loglike(data - states[i]) for i in index]
+    return sum_log_fn
+
+def importance_weights(data, states, burn):
+    log_k = mixture_loglike_v(data = data, 
+            states = states, 
+            burn = burn)
+
+    log_f = logchi2_loglike_v(data = data, 
+            states = states, 
+            burn = burn)
+
+    w = log_f - log_k
+    c = np.exp(w) / np.sum(np.exp(w))
+    return c

### 2023-09-09 analysis/gauss_mix/run_gma.py

@@ -1,19 +1,25 @@
 #Multiple chains
 #Seeds + datasets
 from __future__ import division
+import math
 
 import numpy as np
 import pandas as pd
+import matplotlib.pyplot as plt
 import pyarrow as pa
 import pyarrow.parquet as pq
+import scipy as sp
 np.set_printoptions(suppress=True, precision=5)
 
 from gma import *
 
 np.random.seed(1234)
 
-df = pd.read_csv("../../data/simulated/ksc/phi_0.97779_sig_0.1585_beta_0.64733.csv")
-endog = df['yobs']
+# df = pd.read_csv("../../data/simulated/ksc/phi_0.97779_sig_0.1585_beta_0.64733.csv")
+# endog = df['yobs']
+ex = pd.read_excel("../../data/raw/xrates.xls", engine = "xlrd")
+dta = np.log(ex).diff().iloc[1:]
+endog = (dta['USXUK'] - dta['USXUK'].mean()) * 100
 
 # Setup the model and simulation smoother
 mod = TVLLDT(endog)
@@ -21,8 +27,8 @@ mod.set_smoother_output(0, smoother_state=True) # Calculate and return the smoot
 sim = mod.simulation_smoother()
 
 # Simulation parameters
-n_iterations = 1000
-burn = 100
+n_iterations = 19999
+burn = 10000
 thin = 10
 
 # Storage for traces
@@ -60,10 +66,78 @@ for s in range(1, n_iterations + 1):
     trace_mu[s] = draw_posterior_mu(mod, states, trace_phi[s-1], trace_sigma2[s-1])
 
 # Create dataframe
-draws = np.concatenate((trace_mu, trace_phi, trace_sigma2, trace_states), axis = 1)
+burn_draws = np.concatenate((trace_mu, trace_phi, trace_sigma2, trace_states), axis = 1)
+draws = burn_draws[burn:burn_draws.shape[0]]
 static_names = ['mu', 'phi', 'sigma2']
 state_names = [f"h[{state}]" for state in np.arange(1, trace_states.shape[1]+1)]
-samples = pd.DataFrame(draws[burn:draws.shape[0]])
+
+# Weights
+y_star = np.log(endog**2 + 0.001)
+
+def f_weights(data, states, burn):
[…]

### 2023-09-09 sim_ksc.py

@@ -94,11 +94,19 @@ def simulation(seed_data):
         trace_mu[s] = draw_posterior_mu(mod, states, trace_phi[s-1], trace_sigma2[s-1])
 
     # Create dataframe
-    draws = np.concatenate((trace_mu, trace_phi, trace_sigma2, trace_states), axis = 1)
+    y_star = np.log(endog**2 + 0.001)
+    weights = importance_weights(data = y_star, 
+          states = trace_states, 
+          burn = config["burn"])
+
+    burn_draws = np.concatenate((trace_mu, trace_phi, trace_sigma2, trace_states), axis = 1)
+    param_draws = burn_draws[config["burn"]+1:burn_draws.shape[0]]
+    draws = np.concatenate((param_draws, weights.reshape(-1,1)), axis = 1)
+    samples = pd.DataFrame(draws)
+    
     static_names = ['mu', 'phi', 'sigma2']
     state_names = [f"h[{state}]" for state in np.arange(1, trace_states.shape[1]+1)]
-    samples = pd.DataFrame(draws[config["burn"]+1:draws.shape[0]])
-    samples.columns = static_names + state_names
+    samples.columns = static_names + state_names + ['weights']
 
     # Save as parquet
     table = pa.Table.from_pandas(samples)

### 2023-09-09 src/gaussianMixture.py

@@ -1,7 +1,7 @@
 from __future__ import division
 
 import numpy as np
-import pandas as pd
+import scipy as sp
 import statsmodels.api as sm
 
 from statsmodels.tsa.statespace.tools import (
@@ -152,3 +152,31 @@ def draw_mixing(mod, states):
     sample = np.argmax(tmp, axis=1)
 
     return sample
+
+def mixture_loglike_v(data,states, burn):
+    state_val = states[burn+1:states.shape[0]]
+    norm_list = [ksc_params[i,0] * norm.pdf(x=data, loc=state_val+ksc_params[i,1]- 1.2704, scale=ksc_params[i,2]**0.5) for i in range(0,7)]
+    log_kn = np.log(sum(norm_list))
+    sum_log_kn = np.sum(log_kn, axis = 1)
+    return sum_log_kn
+
+def logchi2_loglike(x):
+    return np.sum(np.log(sp.stats.chi2.pdf(np.exp(x), df = 1) * np.exp(x)))
+
+def logchi2_loglike_v(data, states, burn):
+    index = np.arange(burn+1, states.shape[0])
+    sum_log_fn = [logchi2_loglike(data - states[i]) for i in index]
+    return sum_log_fn
+
+def importance_weights(data, states, burn):
+    log_k = mixture_loglike_v(data = data, 
+            states = states, 
+            burn = burn)
+
+    log_f = logchi2_loglike_v(data = data, 
+            states = states, 
+            burn = burn)
+
+    w = log_f - log_k
+    c = np.exp(w) / np.sum(np.exp(w))
+    return c

### 2023-09-10 R/ksc_results.r
refactored and added weights

@@ -1,0 +1,56 @@
+combine_chains <- function(parameter, list_obj){
+        parameter_df = as.data.frame(cbind(
+        list_obj[[1]][parameter],
+        list_obj[[2]][parameter],
+        list_obj[[3]][parameter],
+        list_obj[[4]][parameter]
+            )
+        )
+
+        names(parameter_df) <- c("chain_1", "chain_2", "chain_3", "chain_4")
+
+        return(parameter_df)
+    }
+
+ diagnostics <- function(dataframe){
+        rhat <- posterior::rhat(as.matrix(dataframe))
+        rhat_basic <- posterior::rhat_basic(as.matrix(dataframe))
+        ess_bulk <- posterior::ess_bulk(as.matrix(dataframe))
+        ess_tail <- posterior::ess_tail(as.matrix(dataframe))
+        ess_basic <- posterior::ess_basic(as.matrix(dataframe))
+
+        results = as.data.frame(list(
+                           rhat=rhat, 
+                           rhat_basic=rhat_basic, 
+                           ess_bulk=ess_bulk,
+                           ess_tail=ess_tail,
+                           ess_basic=ess_basic))
+
+        return(results)
+    }
+
+    one_chain <- function(df){
+        return(df$chain_1)
+    }
+
+# Rank statistics
+rank_stats <- function(parameter, draws, prior_parameters) {
+        sum(draws[[parameter]] < prior_parameters[[parameter]])
+    }
+    
+weighted_ranks <- function(parameter, draws, prior_parameters, posterior_weights, one_chain){
+    unweighted_bool <- as.data.frame(draws[[parameter]] < prior_parameters[[parameter]] )
+    unweighted <- as.data.frame(sapply(unweighted_bool, as.numeric)) %>% 
+    tibble::rowid_to_column("index") %>% 
+    pivot_longer(cols = -c(index), names_to = "chain", values_to = "ranks") 
+    
+    weighted = posterior_weights %>% 
+        inner_join(unweighted, by = c("chain", "index")) %>% 
+        mutate(weighted_rank = weights * ranks)
+    if(one_chain){
+        result <- sum(weighted[weighted["chain"] == "chain_1", ]$weighted_rank)
+    } else{
+        result <- sum(weighted$weighted_rank)
+    }
+    return(result)
+}

### 2023-09-10 scripts/sim_ksc_results.r

@@ -6,8 +6,10 @@ library(posterior)
 library(future.apply)
 plan(multisession)
 
+source("R/ksc_results.r")
+
 config <- jsonlite::read_json("configs/ksc.json")
-tmp_data_location <- paste("simulation_output/", config$simulation_name, "/tmp", sep ="") 
+tmp_data_location <- "data/preprocessed/ksc_sbc" #paste("data/preprocessed/", config$simulation_name, "/tmp", sep ="") 
 
 file_metadata <- as.data.frame(list.files(tmp_data_location))
 names(file_metadata) <- c("files")
@@ -20,6 +22,9 @@ file_metadata <- tidyr::separate(file_metadata,
     tidyr::separate(., col = mcmc_seed, sep = "\\.", into = c("mcmc_seed", "parquet"), remove = TRUE) %>% 
     select(-parquet)
 
+## DElETE
+# file_number = "1"
+
 simulation_results <- function(file_number){
     results <- list()
     results[["all_chains"]] <- list()
@@ -31,50 +36,27 @@ simulation_results <- function(file_number){
     iteration_list <- lapply(iter_files_path, arrow::read_parquet, as_data_frame=TRUE)
     parameter_names <- names(iteration_list[[1]])
 
-    combine_chains <- function(parameter, list_obj){
-        parameter_df = as.data.frame(cbind(
-        list_obj[[1]][parameter],
-        list_obj[[2]][parameter],
-        list_obj[[3]][parameter],
-        list_obj[[4]][parameter]
-            )
-        )
-
-        names(parameter_df) <- c("chain_1", "chain_2", "chain_3", "chain_4")
-
-        return(parameter_df)
-    }
-
-
     all_chains = sapply(parameter_names, combine_chains, list_obj = iteration_list, simplify=FALSE)
     names(all_chains)[names(all_chains) == "sigma2"] <- "sigma_sqd"
+    
+    # ###REMOVE    
+    # weights <- as.data.frame(cbind(rep(1,9999),rep(1,9999),rep(1,9999),rep(1,9999)))
+    # names(weights) <- c("chain_1", "chain_2", "chain_3", "chain_4")
+    # all_chains[["weights"]] <- weights
+    # ###REMOVE
 
-    one_chain <- function(df){
-        return(df$chain_1)
-    }
+    post_weights <- all_chains[['weights']] %>% 
+            tibble::rowid_to_column("index") %>% 
+            pivot_longer(cols = -c(index),  names_to = "chain", values_to = "weights")
 
[…]

### 2023-09-10 scripts/sim_ksc_results.r
added back tmp data

@@ -9,7 +9,7 @@ plan(multisession)
 source("R/ksc_results.r")
 
 config <- jsonlite::read_json("configs/ksc.json")
-tmp_data_location <- "data/preprocessed/ksc_sbc" #paste("data/preprocessed/", config$simulation_name, "/tmp", sep ="") 
+tmp_data_location <- paste("data/preprocessed/", config$simulation_name, "/tmp", sep ="") 
 
 file_metadata <- as.data.frame(list.files(tmp_data_location))
 names(file_metadata) <- c("files")

### 2023-09-10 scripts/sim_ksc_results.r
typo

@@ -9,7 +9,7 @@ plan(multisession)
 source("R/ksc_results.r")
 
 config <- jsonlite::read_json("configs/ksc.json")
-tmp_data_location <- paste("data/preprocessed/", config$simulation_name, "/tmp", sep ="") 
+tmp_data_location <- paste("simulation_output/", config$simulation_name, "/tmp", sep ="") 
 
 file_metadata <- as.data.frame(list.files(tmp_data_location))
 names(file_metadata) <- c("files")

### 2023-09-13 R/sbc.r
fixed bug different dataset structure

@@ -100,16 +100,26 @@ load_parameters <- function(rds_path) {
 
 get_rhat_basic <- function(rds_path, model_path) {
     data = readRDS(here::here(paste(model_path, rds_path, sep = "/")))
-    diagnostics = do.call("rbind", data$all_chains$diagnostics)
-    results <- as.data.frame(cbind(rownames(diagnostics), diagnostics$rhat_basic))
+    if(is.null(dim(data$all_chains$diagnostics))){
+        diagnostics = do.call("rbind", data$all_chains$diagnostics)
+        results <- as.data.frame(cbind(rownames(diagnostics), diagnostics$rhat_basic))
+    } else{
+        diagnostics = data$all_chains$diagnostics
+        results <- as.data.frame(cbind(diagnostics$variable, diagnostics$rhat_basic))
+    }
     names(results) <- c("parameters", "rhat_basic")
     return(results)
 }
 
 get_rhat <- function(rds_path, model_path) {
     data = readRDS(here::here(paste(model_path, rds_path, sep = "/")))
-    diagnostics = do.call("rbind", data$all_chains$diagnostics)
-    results <- as.data.frame(cbind(rownames(diagnostics), diagnostics$rhat))
+    if(is.null(dim(data$all_chains$diagnostics))){
+        diagnostics = do.call("rbind", data$all_chains$diagnostics)
+        results <- as.data.frame(cbind(rownames(diagnostics), diagnostics$rhat))
+    } else{
+        diagnostics = data$all_chains$diagnostics
+        results <- as.data.frame(cbind(diagnostics$variable, diagnostics$rhat))
+    }
     names(results) <- c("parameters", "rhat")
     return(results)
 }

### 2023-09-13 analysis/sbc/sbc_results.r
weighted ranks

@@ -7,9 +7,13 @@ source("R/sbc.r")
 plan(multicore, workers = 9)
 
 # Get all parameter data
-path <- "simulation_output/sbc_cp_ksc_model_cp_dgf_10kmcmc_r1"
+path <- "simulation_output/ksc/ncp/sbc_ncp_ksc_model_ncp_dgf_r1"
 rds_list <- list.files(path = paste(path, "output", sep = "/"), pattern = "*.RDS")
 rds_path <- paste("output", rds_list, sep = "/")
+
+# Rank type
+# rank_type = "weighted_ranks"
+rank_type = "agg_ranks"
 
 # Parameters
 n_iterations <- length(rds_list)
@@ -18,10 +22,15 @@ n_bins <- 20 # Bins
 expected_count <- n_iterations / n_bins # Chi sqd stat
 
 # Load ranks
-weighted_results <- lapply(rds_path, get_ranks, model_path = path, rank_type = "weighted_ranks")
-results <- lapply(rds_path, get_ranks, model_path = path, rank_type = "agg_ranks")
-df <- as.data.frame(do.call("rbind", results))
-wdf <- as.data.frame(do.call("rbind", weighted_results)) * posterior_samples
+results <- lapply(rds_path, get_ranks, model_path = path, rank_type = rank_type)
+
+if (rank_type == "agg_ranks"){
+    df <- as.data.frame(do.call("rbind", results))
+} else if(rank_type == "weighted_ranks"){
+    df <- as.data.frame(do.call("rbind", results)) * (posterior_samples+1)
+} else {
+    print("error")
+}
 
 # Create bins and chi sq stats
 rank_bins <- as.data.frame(lapply(df, binning, posterior_draws = posterior_samples, bins = n_bins))
@@ -41,7 +50,7 @@ static_hist <- facet_hist(data = rank_bins,
                           rank_scales = (posterior_samples+1)
                           )
 
-ggsave(paste(path, "/static_hist.png", sep =""), static_hist, bg = "white")
+ggsave(paste(path, "/static_hist_", rank_type, ".png", sep =""), static_hist, bg = "white", width = 11)
 
 static_state_hist <- facet_hist(data = rank_bins, 
                                 variables = append(static_parameters, additional_parameters), 
@@ -49,20 +58,20 @@ static_state_hist <- facet_hist(data = rank_bins,
[…]

### 2023-09-13 src/gaussianMixture.py
updated initialisation.

@@ -33,7 +33,7 @@ class TVLLDT(sm.tsa.statespace.MLEModel):
 
         # Initialize base model
         super(TVLLDT, self).__init__(endog, k_states=1, k_posdef=1,
-                                     initialization='stationary')
+                                     initialization='known', initial_state =[0], initial_state_cov=np.array([[0.02/(1-0.95**2)]]))
 
         # Setup time-varying arrays for observation equation
         self['obs_intercept'] = np.zeros((1, self.nobs)) # mod.ssm.obs_intercept

### 2023-09-15 R/ksc_results.r
updated to handle only one chain

@@ -1,4 +1,6 @@
 combine_chains <- function(parameter, list_obj){
+
+        if(length(list_obj)==4){
         parameter_df = as.data.frame(cbind(
         list_obj[[1]][parameter],
         list_obj[[2]][parameter],
@@ -6,9 +8,13 @@ combine_chains <- function(parameter, list_obj){
         list_obj[[4]][parameter]
             )
         )
-
         names(parameter_df) <- c("chain_1", "chain_2", "chain_3", "chain_4")
-
+        } else if (length(list_obj)==1) {
+        parameter_df = as.data.frame(list_obj[[1]][parameter])
+        names(parameter_df) <- c("chain_1")
+        } else {
+            print("error")
+        }        
         return(parameter_df)
     }
 
@@ -29,7 +35,7 @@ combine_chains <- function(parameter, list_obj){
         return(results)
     }
 
-    one_chain <- function(df){
+    one_chain_f <- function(df){
         return(df$chain_1)
     }
 

### 2023-09-15 scripts/sim_ksc_results.r

@@ -22,9 +22,6 @@ file_metadata <- tidyr::separate(file_metadata,
     tidyr::separate(., col = mcmc_seed, sep = "\\.", into = c("mcmc_seed", "parquet"), remove = TRUE) %>% 
     select(-parquet)
 
-## DElETE
-# file_number = "1"
-
 simulation_results <- function(file_number){
     results <- list()
     results[["all_chains"]] <- list()
@@ -38,12 +35,6 @@ simulation_results <- function(file_number){
 
     all_chains = sapply(parameter_names, combine_chains, list_obj = iteration_list, simplify=FALSE)
     names(all_chains)[names(all_chains) == "sigma2"] <- "sigma_sqd"
-    
-    # ###REMOVE    
-    # weights <- as.data.frame(cbind(rep(1,9999),rep(1,9999),rep(1,9999),rep(1,9999)))
-    # names(weights) <- c("chain_1", "chain_2", "chain_3", "chain_4")
-    # all_chains[["weights"]] <- weights
-    # ###REMOVE
 
     post_weights <- all_chains[['weights']] %>% 
             tibble::rowid_to_column("index") %>% 
@@ -51,10 +42,14 @@ simulation_results <- function(file_number){
 
     all_chains <- within(all_chains, rm(weights))
     
-    one_chain = sapply(all_chains, one_chain, simplify=FALSE)
+    one_chain = sapply(all_chains, one_chain_f, simplify=FALSE)
 
     # All chain diagnostics
+    if(length(iteration_list)>1){
     all_chains_diagnostics <- sapply(all_chains, diagnostics, simplify=FALSE)
+    } else{
+    all_chains_diagnostics <- "Only one chain"
+    }
 
     # One chain diagnostics
     ess_basic <- sapply(one_chain, posterior::ess_basic, USE.NAMES = TRUE)
@@ -68,9 +63,11 @@ simulation_results <- function(file_number){
     data_location <- here::here("data/simulated/sbc", config$data_location, paste(file_number, "json", sep = "."))
     prior_params <- jsonlite::read_json(data_location)
 
+    if(length(iteration_list)>1){
     results[["all_chains"]][["agg_ranks"]] <- sapply(names(all_chains), rank_stats, draws = all_chains, prior_parameters = prior_params, USE.NAMES = TRUE)
[…]

### 2023-09-15 scripts/sim_ksc_results.r
ess weights

@@ -40,6 +40,10 @@ simulation_results <- function(file_number){
             tibble::rowid_to_column("index") %>% 
             pivot_longer(cols = -c(index),  names_to = "chain", values_to = "weights")
 
+    post_weights_ess <- post_weights %>% 
+                            group_by(chain) %>% 
+                            summarise(ess_weights=1/sum(weights^2))
+
     all_chains <- within(all_chains, rm(weights))
     
     one_chain = sapply(all_chains, one_chain_f, simplify=FALSE)
@@ -75,6 +79,7 @@ simulation_results <- function(file_number){
     results[["one_chain"]][["diagnostics"]] <- one_chain_diagnostic
     results[["config"]] <- config
     results[["seed_files"]] <- file_metadata[file_metadata["dataset"] == file_number, ]
+    results[["weights_ess"]] <- post_weights_ess
 
     saveRDS(results,
             file = here::here("simulation_output",

### 2023-09-16 configs/sbc_sim.r
latest configs

@@ -1,21 +1,21 @@
-simulation_name <- "sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r2"
+simulation_name <- "sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r9_5000_iterations"
 simulation_description = "
-SBC with non centered model and ksc priors with pre-simulated datasets run: 2.0
+SBC with noncentered model and ksc priors and noncentered dataset with pre-simulated datasets run: 9.0
 "
 
 # Simulation iterations
-sim_iter <- 1000
+sim_iter <- 5000
 
 # Model metadata
 model_name <- "sv_ncp_ksc_sbc" # Name of the .stan file in the models directory
 
 # Data location
-data_location <- "sbc_data_gen_sv_ncp_ksc_priors"
+data_location <- "sbc_data_gen_sv_ncp_ksc_priors_5000_datasets"
 dependent_variable <- "y_sim"
 
 ## Stan parameters
-chains <- 1
-parallel_chains <- 1
+chains <- 4
+parallel_chains <- 4
 refresh <- 500
 adapt_delta <- 0.999
 save_warmup <- FALSE

### 2023-09-16 jobs/ksc.sh

@@ -1,12 +1,12 @@
 #!/bin/bash
-#SBATCH --job-name=sbc_cp_ksc_model_r3
-#SBATCH --time=06:00:00
+#SBATCH --job-name=sbc_cp_ksc_model_cp_dgf_10kmcmc_3000_niter_r1
+#SBATCH --time=12:00:00
 #SBATCH --mem-per-cpu=4096
 #SBATCH --array=1
 #SBATCH --ntasks=1
-#SBATCH --cpus-per-task=25
+#SBATCH --cpus-per-task=40
 #SBATCH --ntasks-per-socket=1
 
-source .env/bin/activate
-python sim_ksc.py
+# source .env/bin/activate
+# python sim_ksc.py
 R --vanilla < scripts/sim_ksc_results.r

### 2023-09-16 scripts/convert_rds_json.r

@@ -10,7 +10,7 @@ get_data_name <- function(data_name){
     return(unlist(strsplit(data_name, split='.', fixed=TRUE))[1])
 }
 
-data_path <- "data/simulated/sbc/sbc_data_gen_sv_ncp_ksc_priors"
+data_path <- "data/simulated/sbc/sbc_data_gen_sv_cp_ksc_priors_5000_datasets"
 data_list <- list.files(path = data_path, pattern = "*.RDS")
 file_names <- unlist(lapply(data_list, get_data_name))
 file_names <- file_names[file_names != "metadata"]

### 2023-09-16 scripts/sbc_data_gen.r

@@ -1,6 +1,6 @@
-model_name <- "sbc_data_gen_sv_cp_ksc_priors_1000_datasets"
+model_name <- "sbc_data_gen_sv_cp_ksc_priors_5000_datasets"
 n_time_points <- 1000
-n_datasets <- 1000
+n_datasets <- 5000
 data_path <- here::here("data", "simulated", "sbc", model_name)
 seed <- 543
 parameterisation <- "cp"

### 2023-09-16 analysis/gauss_mix/gma.py
sandbox

@@ -32,8 +32,8 @@ class TVLLDT(sm.tsa.statespace.MLEModel):
         endog = np.log(endog**2 + offset)
 
         # Initialize base model
-        super(TVLLDT, self).__init__(endog, k_states=1, k_posdef=1,
-                                     initialization='stationary')
+        super(TVLLDT, self).__init__(endog, k_states=1,
+                                     initialization='known', initial_state =[0], initial_state_cov=np.array([[0.02/(1-0.95**2)]]))
 
         # Setup time-varying arrays for observation equation
         self['obs_intercept'] = np.zeros((1, self.nobs)) # mod.ssm.obs_intercept

### 2023-09-16 analysis/gauss_mix/run_gma.py

@@ -8,12 +8,12 @@ import pandas as pd
 import matplotlib.pyplot as plt
 import pyarrow as pa
 import pyarrow.parquet as pq
-import scipy as sp
 np.set_printoptions(suppress=True, precision=5)
 
 from gma import *
 
-np.random.seed(1234)
+seed=2413
+np.random.seed(seed)
 
 # df = pd.read_csv("../../data/simulated/ksc/phi_0.97779_sig_0.1585_beta_0.64733.csv")
 # endog = df['yobs']
@@ -67,10 +67,56 @@ for s in range(1, n_iterations + 1):
 
 # Create dataframe
 burn_draws = np.concatenate((trace_mu, trace_phi, trace_sigma2, trace_states), axis = 1)
-draws = burn_draws[burn:burn_draws.shape[0]]
+draws = burn_draws[burn+1:burn_draws.shape[0]]
 static_names = ['mu', 'phi', 'sigma2']
 state_names = [f"h[{state}]" for state in np.arange(1, trace_states.shape[1]+1)]
 
+y_star = np.log(endog**2 + 0.001)
+weights = importance_weights(data = y_star, 
+          states = trace_states, 
+          burn = burn)
+
+weights_ess = 1/sum(weights**2)
+
+burn_draws = np.concatenate((trace_mu, trace_phi, trace_sigma2, trace_states), axis = 1)
+param_draws = burn_draws[burn+1:burn_draws.shape[0]]
+draws = np.concatenate((param_draws, weights.reshape(-1,1)), axis = 1)
+samples = pd.DataFrame(draws)
+    
+static_names = ['mu', 'phi', 'sigma2']
+state_names = [f"h[{state}]" for state in np.arange(1, trace_states.shape[1]+1)]
+samples.columns = static_names + state_names + ['weights']
+
+table = pa.Table.from_pandas(samples)
+pq.write_table(table, f'data/{seed}.parquet')
+
+mean_phi = np.mean(trace_phi[burn::thin])
+mean_sigma = np.mean(trace_sigma2[burn::thin]**0.5)
+mean_beta = np.mean(np.exp(trace_mu[burn::thin] / 2))
+
+
+fig, axes = plt.subplots(1, 3, figsize=(15, 4))
+
+axes[0].hist(trace_phi[burn::thin])
+ylim = (0, 140)
[…]

### 2023-09-16 analysis/gauss_mix/sim_ksc_results.r

@@ -1,0 +1,91 @@
+library(jsonlite)
+library(dplyr)
+library(tidyr)
+library(arrow)
+library(posterior)
+library(future.apply)
+plan(multisession)
+
+source("R/ksc_results.r")
+
+config <- jsonlite::read_json("configs/ksc.json")
+tmp_data_location <- paste("simulation_output/", config$simulation_name, "/tmp", sep ="") 
+
+file_metadata <- as.data.frame(list.files("analysis/gauss_mix/data"))
+names(file_metadata) <- c("files")
+
+file_metadata <- tidyr::separate(file_metadata, 
+                     col = files, 
+                     sep = "_", 
+                     into = c("dataset", "mcmc_seed"),
+                     remove = FALSE) %>% 
+    tidyr::separate(., col = mcmc_seed, sep = "\\.", into = c("mcmc_seed", "parquet"), remove = TRUE) %>% 
+    select(-parquet)
+
+    results <- list()
+    results[["all_chains"]] <- list()
+    results[["one_chain"]] <- list()
+
+    iteration_list <- list()
+    iter_files <- file_metadata[["files"]]
+    iter_files_path <- paste("analysis/gauss_mix/data", "/", iter_files, sep = "")
+    iteration_list <- lapply(iter_files_path, arrow::read_parquet, as_data_frame=TRUE)
+    parameter_names <- names(iteration_list[[1]])
+
+    all_chains = sapply(parameter_names, combine_chains, list_obj = iteration_list, simplify=FALSE)
+    names(all_chains)[names(all_chains) == "sigma2"] <- "sigma_sqd"
+
+    post_weights <- all_chains[['weights']] %>% 
+            tibble::rowid_to_column("index") %>% 
+            pivot_longer(cols = -c(index),  names_to = "chain", values_to = "weights")
+
+    all_chains <- within(all_chains, rm(weights))
+    
+    one_chain = sapply(all_chains, one_chain, simplify=FALSE)
+
+    # All chain diagnostics
+    all_chains_diagnostics <- sapply(all_chains, diagnostics, simplify=FALSE)
+
+    # One chain diagnostics
[…]

### 2023-09-16 writeups/week8_update/week8.qmd
docs

---
title: " "
format:
  pdf:
    fig-pos: H
    include-in-header:
        text: |
          \usepackage{lipsum}
          \usepackage{setspace}
          \onehalfspacing
          \linespread{1.5}
fontsize: 12pt
geometry: margin=1in
mathspec: true
---

## Rank formula

$$
\begin{aligned}
r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim}) = \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}]
\end{aligned}
$$

## Reweighting

Let $f(\theta, h_t | y_t)$ be the posterior of the true stochastic volatility model, $g(\theta, h_t | y^{\ast}_t)$ be the linearised model with log chi squared standard errors, and $k(\theta, h_t | y^{\ast}_t)$ be the Gaussian approximation model. KSC define their log weights as:

$$
\begin{aligned}
w(\theta, h) = log\: f(\theta, h | y) - log\:  k(\theta, h_t | y^{\ast}_t) = const + log\: f(y|h) - log\: k(y^{\ast} | h)
\end{aligned}
$$

However, I think the weights should be with respect to the log chi squared distribtion. So I define log weights as:

$$
\begin{aligned}
w(\theta, h) = log\: g(\theta, h | y^{\ast}_t) - log\:  k(\theta, h_t | y^{\ast}_t)
\end{aligned}
$$

Take the exponential and normalise the weights for the $l^{th}$ posterior draw:

$$
\begin{aligned}
c^l = \frac{exp(w(\theta, h)_l)}{\sum_i exp(w(\theta, h)_i)}
\end{aligned}
$$

# Reweighted Rank formula
Note that multiplying the L (the number of posterior draws) brings it back to the correct scale (otherwise all the ranks are between 0 and 1).
$$
\begin{aligned}
r &= L \times \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}]\times c^l   \\
&= L \times \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}]\times \frac{exp(w(\theta, h)_l)}{\sum_i exp(w(\theta, h)_i)}
\end{aligned}
$$

# Visualising histograms of log weights

KSC use the SV model with normal errors as the target "true" distribution instead of the linearised log chi squared model when calculating weights. If you create a histogram using their formulation for log weights, you get:

![](../reweight_samples/incorrect.png)

This is way off from zero. However, if you use the log chi squared density  as the "true" density we are trying to approximate, we get the correct plot which they have in their paper. These weights are defined as: $w(\theta, h) = log\: g(\theta, h_t | y^{\ast}_t) - log\:  k(\theta, h_t | y^{\ast}_t)$

![](../reweight_samples/correct.png)

This is much better. This is also why I needed help getting the density of the log chi squared distribution in order to calculate this. And this also makes much more sense - the whole point of the Gaussian mixture was to approximate the lineared log chi squared model _not_ the SV model with normal errors.

Taking the delta:

![](delta_histograms.png)

- count	9999
- mean -1004.297735
- std	0.120312
- min	-1004.943507
- 25%	-1004.375330
- 50%	-1004.289475
- 75%	-1004.214196
- max	-1003.923458

### 2023-09-16 writeups/research_proposal/propsal.qmd

@@
 author: 
-    - "By Benjamin Wee" 
+    - "By Benjamin Wee (32921020)" 
     - "Supervised by Catherine Forbes and Lauren Kennedy"
@@
 Stochastic volatility models are used in financial econometrics to model the variance or volatility of returns in a financial instrument.
-The return of a financial asset often exhibits pattterns in the variance (heteroskedasticity) which can be modelled to improve prediction and estimates of risk or tail events.
-Unlike other volatility models such as ARCH or GARCH, these stochastic volatility models treat the variance as a random variable.
-Furthermore, these models have a state space representation since the variance is treated as a latent stochastic process.
-State space models can be difficult to estimate since there is a latent state for every data observation in the time series, which means there are more parameters than data points.
-These models can be estimated using both frequent tools but this research will focus on Bayesian estimation strategies.
+The return of a financial asset often exhibits patterns in the variance which can be modeled to improve prediction and estimates of risk or tail events.
+Unlike other volatility models such as ARCH or GARCH, stochastic volatility models treat the variance as a random variable.
+Furthermore, these models have a state space representation since the variance is expressed as a latent stochastic process.
+State space models can be difficult to estimate since there is a latent state for every data observation in the time series, which means there are often more parameters than data points.
+These models can be estimated using frequent frameworks but this research will focus on Bayesian estimation strategies.
 
-Kim, Sherpherd and Chib (1998) estimate a discrete time, univariate stochastic volatility model using Bayesian methods which is outlined below.
+Kim, Shephard and Chib (1998) estimate a discrete time, univariate stochastic volatility model using Bayesian methods which is outlined below.
 $y_t$ is the mean corrected returns of some asset for equally spaced intervals t.
@@
 $h_t$ is log volatility, where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the autoregressive parameter $\phi$ such that $|\phi|<1$.
-This autoregressive parameter represents the persistence or "stickiness" of log volatility and the dispersion parameter $\sigma_{\eta}$ is its variance.
-$\epsilon_t$ and $\eta_t$ are standard normal white noise shocks.
+This autoregressive parameter represents the persistence or "stickiness" of log volatility and the dispersion parameter $\sigma_{\eta}$ is the constant variance of the states.
+$\epsilon_t$ and $\eta_t$ are standard normal white noise shocks and are uncorrelated with each other.
 
@@
 ### Estimation strategy
-Kim, Shephard and Chib's sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother[^1] (de Jong and Shephard, 1995).
+Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother[^1] (de Jong and Shephard, 1995).
 
@@
 Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
-However, it is not simple to sample from this parameterisation of the model.
-Kim, Sherpherd and Chib use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution.
+The relationship between $y_t$ and $h_t$ is now linear; however, the error is not Gaussian.
+Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution.
 This is defined by:
@@
 
-The model can be sampled via the simulation smoother since the model is now linear and conditionally gaussian.
-The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions where as $\phi$ is sampled via a metropolis hastings accept/reject procedure.
-The details can be be around in Appendix B.
+The model can be sampled via the Kalman Filter and simulation smoother since the model is now linear and conditionally gaussian.
+The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions whereas $\phi$ is sampled via a Metropolis Hastings accept/reject procedure.
+The details can be around in Appendix B.
 
@@
 Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature (Duane et al.
-1987) before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks (Nael, 1995) and statistical computing (Neal 2011).
+1987) before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks (Neal, 1995) and statistical computing (Neal 2011).
 The algorithm has since become widely available through open source development projects such as Stan (Stan Development Team, 2023) and PyMC (2023).
@@
 The key innovation of Hamiltonian Monte Carlo is using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
-Unlike Random Walk Metropolis Hastings, it takes advantage of the geometry of the posterior to determine its next step.
+Unlike Random Walk Metropolis Hastings, it takes advantage of the geometry of the posterior to determine its next proposal step.
 A comprehensive explanation of the sampler is beyond the scope of this research and can be found in the above references.
@@
 Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
-This approach will use the same priors as specified in the Gaussian mixture approximation.
-However, Hamiltonian Monte Carlo does not sample directly from the analytical form of conjugate or conditionally conjugate distribution, so different priors can be chosen if necessary.
+This approach will also use the same priors as specified in the Gaussian mixture approximation.
 
 ## 2.3 Research contribution
-The goal of this research is to determine which algorithm is better at estimating stochastic volatility models.
-There are clear differences between Kim Sherphard and Chib's estimation strategy and Hamiltoninan Monte Carlo.
-My main contribution is to compare these algorithms and determine which of these sampling strategies is better based on whether the algorithm returns calibrated posterior estimates and the efficiency of the MCMC sampler.
+The goal of this research is to determine which algorithm is better at estimating the discrete time stochastic volatility model (conditional on the same priors).
+My main contribution is to design a simulation study that compares these algorithms and determines which sampling strategy is better based on the calibration of posterior estimates and the efficiency of the MCMC sampler.
 
@@
 ## 3.1 Design
-Simulation Based Calibration (SBC) is a simulation technique that will help check the calibration of posterior estimates generated by different MCMC algorithms.
+Simulation Based Calibration (SBC) is a technique that checks the calibration of posterior estimates generated by MCMC algorithms.
 SBC is conducted by comparing the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.
 The procedure starts by taking draws from the prior distribution and creating datasets implied by each draw.
-Rank statistics are then calculated on the posterior samples estimated from the simulated data.
-Let $\theta$ be a parameter and $y$ represent the dataset.
-Start with draw a sample from the prior distribution:
+Rank statistics are then calculated on the posterior samples conditional on the simulated data.
+
+To illustrate this procedure, let $\theta$ be a parameter and $y$ represent the dataset. Start with a single draw from the prior distribution:
 $$
@@
 
-Generate a dataset given by the prior draws.
+Generate a dataset given by the prior draw.
 
@@
 
-Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (Hamiltonian Monte Carlo or Gaussian approximation in our case).
+Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (Hamiltonian Monte Carlo or KSC) conditional on this dataset.
 
@@
 
-A key result from this procedure is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples $\theta^{sim}$.
-This is implied by the following expression:
+A key result is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples $\theta^{sim}$. This is implied by the following expression:
 
@@
 The procedure of generating posterior samples implicitly performs this integral since the expression on the right of the integral is proportional to the prior density.
-Therefore, any deviation of the posterior samples from the prior distribution suggests an error - either in computation, code or overall analysis.
+Therefore, any deviation of the posterior samples from the prior distribution suggests that the sampling methodology is not producing the correct posteriors.
+
+<!-- n error - either in computation, code or overall analysis.  -->
 
@@
 
-Since the distribution of the posterior samples follows the prior distribution, the rank statistic for a given parameter and simulation follows a discrete uniform distribution[^2].
+If the posterior samples follows the prior distribution, the rank statistic for a given parameter follows a discrete uniform distribution[^2]. The rank statistic is defined as:
 
-[^2]: Proof of this result in Talts et al 2018.
+[^2]: Proof of this result in Talts et al (2018).
 
@@
 To complete the algorithm, multiple iterations are run and the rank statistics are calculated for each parameter.
-The resulting rank statistics are compared to the discrete uniform distribution to determine if any problematic features or errors exist.
+The resulting rank statistics are compared to the discrete uniform distribution to determine if any problematic features exist.
 
-If the computation is well calibrated and the samples follow a discrete uniform distribution, then the posterior (credible) intervals will have sufficient coverage.
-That is, for any percentage interval selected (for example 90%) then there is 90\% chance that any $\theta^{sim}$ falls within this interval.
-In other words, a Bayesian analysis and computation is well calibrated if the 90\% interval contains the true parameter in 90\% interations of the SBC procedure.
+If the computation is well calibrated and the rank statistics follow a discrete uniform distribution, then the posterior credible intervals have sufficient coverage.
+That is, one way to describe calibration is: for any percentage interval selected over the posterior samples (for example 90%) then there is a 90\% chance that $\theta^{sim}$ falls within this interval.
+Another way of saying this is a Bayesian analysis is well calibrated if a 90\% credible interval contains the true parameter in 90\% of the SBC iterations.
 
 ## 3.2 Implementation
-SBC for Hamiltonian Monte Carlo and the Gaussian approximation are performed with simulated dataset sizes of 1000 simulation iterations, 1000 observations, and post warmup/burnin samples of 999 (which gives 1000 possible rank statistics).
+SBC for Hamiltonian Monte Carlo and the Gaussian approximation are performed with 1000 simulation iterations, 1000 observations in each dataset, and post warmup/burnin samples of 999 (which gives 1000 possible rank statistics).
 This algorithm is summarised below:
@@
 ## 3.3 Metrics
-The key metrics and diagnostics to compare the performance of these samplers is the effective sample size, histograms of rank statistics and chi-squared test statistics.
+The key metrics and diagnostics to compare the performance of these methods are the effective sample size (ESS), rank statistics, and chi-squared test statistics.
 
-Effecitve sample size (ESS) measures the efficiency of the MCMC sampler.
+ESS measures the efficiency of the MCMC sampler.
 It calculates the number of (effectively) independent draws from the posterior draws generated by a Markov chain.
 A poor ESS can arise from high autocorrelation in the Markov chain which leads to highly dependent samples.
-If it requires long MCMC chains to get to a desired ESS, then this is considered inefficient since it takes more resources to get a representative sample of the target distribution.
+An efficient MCMC algorithm takes less resources (for example, time and number of draws) to get a representative sample of the target distribution.
+If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), then we may conclude that this method is a more efficient sampler (conditional on the model).
 
-Histograms of rank statistics are a visual diagnostic to determine if the distribution of ranks follows a discrete uniform distribution.
-As described, a calibrated algorithm implies rank statistics which are randomly drawn from a uniform.
-A weakness of this approach is there are at least as many parameters as there data points.
-Since there are 1000 observations per dataset, this means there are 1003 parameters to estimate.
+Rank statistics as described in the SBC section are used to evaluate the calibration of a posterior.
+Histograms will be used to evaluate the distribution of rank statistics.
+If a posterior is well calibrated then it is expected that the histogram is uniform.
 
-An alternative to checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each bin.
+A drawback of this approach is there are more parameters than data points in this model.
+An alternative to visually checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each histogram bin.
 Let $b_j$ be the number of counts and $e_j$ the expected count in bin $j$.
@@
 
-The chi squared statistics provide a more succinct summary of the overall performance of the algorithm.
+Chi squared statistics can be used to test for uniformity and thus calibration of all the parameters in the model.
 
@@
 Figure 1 shows the distribution of rank statistics for Hamiltonian Monte Carlo, with the horizontal black line at 50 representing the discrete uniform distribution.
-The preliminary figures focus on the static parameters and arbitrarily chosen states (since there is one state parameter per data observation).
+The preliminary figures focus on the static parameters and arbitrarily chosen states.
 Phi and sigma_sqd look relatively uniform.
@@
 
-One of the drawbacks of SBC is that any deviation from uniform from a calibrated analysis maybe due to finite samples.
+A drawback of SBC is that any deviation in uniformity from a calibrated analysis may be due to noisy estimates from small samples.
 Figure 2 shows more consistent uniform behaviour from all parameters after increasing the number of iterations from 1000 to 5000.
-This suggests that Hamiltonian Monte Carlo and overall analysis is well calibrated.
+This suggests that Hamiltonian Monte Carlo and the overall analysis is well calibrated.
 
@@
 The distribution of the rank statistics deviates heavily from the expected discrete uniform distribution, with various peaks at each end of the histogram as well as peaks in the middle.
-This lack of calibration is expected.
-The Gaussian approximation only approximates the log chi squared distribution, whereas Hamiltonian Monte Carlo samples directly from the same likelihood as the data generating process.
+A reason for this miscalibration may be due to the Gaussian mixture only approximating the log chi squared distribution.
+Hamiltonian Monte Carlo on the other hand, samples directly from the target model which more closely represents the data generating process.
 
-This result may also be due to the small number of posterior samples drawn from the model.
+This result may also be due to the relatively small number of posterior samples drawn from the model.
 The MCMC sampler may not have converged yet which may explain the deviations from uniformity.
-A follow up step is to compare the effective sample sizes (ESS) of the two methods and to increase the length of the MCMC chain to ensure the ESS is approximately the same before comparing histograms.
+A follow up step is to compare the ESS of the two methods and to increase the length of the MCMC chain to ensure the ESS is approximately the same before comparing histograms.
 
-![Distribution of parameter ranks for KSC sampling strategy (1000 iterations)](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png){width=75%}
+![Distribution of parameter ranks for Gaussian mixture model (1000 iterations)](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png){width=75%}
 
@@
 
-Kim, Sherpherd and Chib apply a Metropolis Hastings reweighting to the posterior samples to correct for the approximation of the error distribution.
+Kim, Shephard and Chib apply importance sampling weights to correct for the approximation error of the Gaussian mixture model.
 According to the authors, the correction ensures that samples are from the exact posterior density.
@@
 This can cause significant problems in getting unbiased MCMC samples for complex models.
-An example of this is Neal's funnel (Neal 2003) which arises in hierarchical models which causes performance issues with the Hamiltonian Monte Carlo sampler.
+An example of this is Neal's funnel (Neal 2003) which arises in hierarchical models which causes performance issues in the Hamiltonian Monte Carlo sampler.
 Rewriting the model with non centered parameterisation significantly improves the performance of MCMC.
-Another consideration for this research is to consider both centered and non centered parametersations of the stochastic volatility model to see if there are any improvements to the simulation diagnostics.
+Another consideration for this research is to consider both centered and non centered parameterisations of the stochastic volatility model to see if there are any improvements to the simulation diagnostics.
 
@@
 
-![Model matrix](model_table.png)
+![Model matrix](model_table.jpg)
 
@@
 
-Week 8: Finish off any simulation or modeling not yet complete (listed in above matrix)
+Week 8: Finish off any simulation or modelling not yet complete (listed in above matrix)
 
-Week 9: Write up any other results and have a first draft ready for feedback
+Week 9: Write up results and have a first draft ready for feedback
 
@@
 2) Accept proposal as $\phi^{(i)}$ with probability $e^{\{g(\phi^\ast) - g(\phi^{(i-1)}\}}$ such that $g(\phi) = log (\pi (\phi)) - \frac {(h_t - \mu)^2 (1-\phi^2)}{2 \sigma_{\eta}^2} + \frac{1}{2} log (1-\phi^2)$
+
+**Sampling mixture density:**
+
+Rewrite mixture density with respect to a indicator variable $s_t$
+
+$$
+\begin{aligned}
+&z_t | s_t = i \sim N(m_i - 1.2704, \nu^2) \\
+&Pr(s_t = i) = q_i
+\end{aligned}
+$$
+
+Sample $s_t$ from probability mass function:
+
+$$
+\begin{aligned}
+Pr(s_t = i | y_t^{\ast}, h_t) \propto q_i f_N(y_t^{\ast} | h_t + m_t - 1.2704, \nu^2)
+\end{aligned}
+$$
+
+**Sampling steps**
+
+1. Set initial values $s$, $\phi$, $\sigma^2_{\eta}$ and $\mu$
+
+2. Sample $h$ from $h|y^{\ast}, s, \phi, \sigma^2_{\eta}, \mu$
+
+3. Sample s from $s|y^{\ast}, h$
+
+4. Sample $\mu|y^{\ast}, s, \phi, \sigma^2_{\eta}, h$ according to conjugate posterior distribution
+
+4. Sample $\sigma_{\eta}^2|y^{\ast}, s, \phi, \mu, h$ according to conjugate posterior distribution
+
+4. Sample $\phi|y^{\ast}, s, \mu, \sigma^2_{\eta}, h$ using metropolis hastings

### 2023-09-16 manuscript/main.tex
init

--- 2023-09-16 writeups/research_proposal/propsal.qmd
@@
----
-title: "Comparison of MCMC algorithms in Stochastic Volatility Models"
-author: 
-    - "By Benjamin Wee (32921020)" 
-    - "Supervised by Catherine Forbes and Lauren Kennedy"
-# abstract: "The objective of this research is to validate and compare the computational methods used to estimate Bayesian stochastic volatility models.
-Specifically, Simulation Based Calibration will be used to identify biases and errors arising from the results of Markov Chain Monte Carlo (MCMC) algorithms.
-The source of any miscalibration include, but are not limited to, estimation strategies, choice of sampler, model specifications and coding mistakes.
-This study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
-Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches."
-format:
-  pdf:
-    fig-pos: H
-    include-in-header:
-        text: |
-          \usepackage{lipsum}
-          \usepackage{setspace}
-          \onehalfspacing
+\documentclass[12pt, a4paper]{article}
+\usepackage{graphicx} % Required for inserting images
+\usepackage[parfill]{parskip} % Removes indent from new paragraphs
+\usepackage{amsmath, amssymb, amsfonts}
+\usepackage[
+top    = 1.0in,
+bottom = 1.2in,
+left   = 1.0in,
+right  = 1.0in]{geometry}
 \linespread{1.5}
-fontsize: 12pt
-geometry: margin=1in
-mathspec: true
----
 
-<!-- # Background and Motivation
-Maybe focus on the MCMC as opposed to SV upfront.
+\title{Comparison of MCMC algorithms in Stochastic Volatility Models}
+\author{Benjamin Wee}
+\date{October 2023}
 
-## SV
-Why model volatility. Quantifying risks, tail probabilities, worst case scenarios. These are all inputs in the pricing of assets and derivatives. (pg92 FE slides)
+\begin{document}
 
-SV models is one way to model volatility.
-Unlike the classic Black Scholes option pricing formula, it assumes that the underlying variance of an asset's return is not constant.
-That is, the variance of the underlying asset return is treated as a latent random variable.
-This deals with some of the empirical short comings of the BS model such as the volatility smile and skew in the realised variance in financial assets.
+\maketitle
 
-SV models are typically expressed as non linear Gaussian state space models.
-This can make it difficult to estimate using classical methods.
-The likelihood is unavailable in closed form and there are at least as many variables as data points.
-There are however, a variety of Frequentist and Bayesian strategies for estimating SV models.
-Frequentist approaches include quasi likelihood estimation and GMM which rely on asymptotic sampling distributions.
-Bayesian approaches rely on computational Markov Chain Monte Carlo techniques to sample from the target joint posterior of these high dimensional state space models.
-
-This research will focus on Bayesian estimation strategies for Stochastic Volatility models.
-Particular attention will be given to the computational tools and approaches used to sample from high dimensional posterior distributions.
-Early approaches for estimating these models relied on classic Gibbs and Metropolis Hastings algorithms to apply the MCMC routines.
-Since then, there have been many developments in the availability of new kinds of MCMC samplers.
-
-Hamiltonian Monte Carlo is a .
-It has been popularised through its implementation in a variety of open source programming languages.
-It presents a new way to sample from increasingly complicated generative models .
-
-However, as the developments of these tools enable the estimation of more complicated models (with relative ease and accessibility through the development of open source software), so do the risk of mistakes and the need for procedures diagnose problems or mistakes when implementing these methods.
-How do we know that these algorithms are doing what we expect them to be doing, and how can we effectively differentiate the strengths and weaknesses of different tools.
-
-How do we validate the computation and implementation of our sampling strategies and algorithms in increasingly complex models?
-And how do these form part of the computational modelling workflow?
--->
-
-<!-- ## HMC
-- There have been many developemnts over the past few decades on estimating Bayesian models through Markov Chain Monte Carlo methods.
-
-The original BUGs software implemented Gibbs and Metropolis Hastings MCMC algorithms for estimating this model.
-
-In particular, the implementation of Hamiltonian Monte Carlo, a variant of the Metropolis Hastings algorithm, represents a recent advancement MCMC.
-
-- The development of HMC is another advancement in estimating complicated, high dimensional models. 
-- Development of diagnostics for different types of HMC and tools for evaluating the convergence and efficiency of the MCMC sampler.
-
-- SV models are a primary candidate for using MCMC methods? Difficult to
-
-## SBC -->
-
-<!-- and provide tools for evaluating other types models and algorithms.  -->
-
-<!-- The objective of this research is to evaluate the performance of MCMC algorithms for estimating SV models through a simulation study.
-Simulations are useful since there is control over the data generating process.
-This will help us understand the behaviour of MCMC and evaluate its performance against known parameters.
-Particular attention will be given to the calibration of the MCMC algorithms in the stochastic volatility context.
-Further diagnostics to evaluate performance will be explored to determine the efficiency of the MCMC sampler.
-
-The first step is to understand what "performance" of MCMC algorithms means and what exactly we are measuring.
-At a high level, the objective of MCMC is to sample from the target posterior distribution and to provide reliable inference.
-Convergence to the target posterior can be measured through efficiency measures such as rhat and effective sample size.
-
-- Explcitly outline what is compared. KSC and HMC algorithms and estimation strategies.  -->
-
-<!-- 
-Frequentist methods
-
-- Variety of estimation strategies
-
-- There are a variety of ways of estimating this model using a Bayesian frameworks.
-
-Over the past few decades there have been many developments in computing hardware -->
-<!-- 
-RESEARCH GOAL: To compare the performance of different MCMC algorithms and estimation strategies when fitting stochastic volatility models. The study will use a discrete formulation is described  -->
-
-<!-- The use of simulation and sampling are used as part of model diagnostics and evaluation. -->
-
-## 1. Research Goal
-
+\begin{abstract}
     The objective of this research is to compare the computational methods used to estimate Bayesian stochastic volatility models.
@@
 This will determine which algorithm is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
+\end{abstract}
 
-<!-- Stochastic volatility models are used to explicitly model the variance of returns in financial instruments and can be estimated using frequentist or Bayesian frameworks.  -->
+\section{Introduction}
 
-<!-- 
-The objective of this research is to validate the computational methods used to estimate Bayesian stochastic volatility models.
-This study will assess the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
-A simulation study will compare the ability of these methods to sample from the model's posterior distribution.
-Specifically, Simulation Based Calibration (SBC) is used to check whether these sampling strategies are returning efficient and well calibrated posterior estimates.
-Key metrics to check these features are the effective sample size to check efficiency of the algorithm and tests of uniformity to determine calibration of the posteriors.
-This will determine which algorithm is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
--->
-
-<!-- Simulation Based Calibration will be used to identify biases and errors arising from the results of Markov Chain Monte Carlo (MCMC) algorithms.
-The source of any miscalibration include, but are not limited to, estimation strategies, choice of sampler, model specifications and coding mistakes.
-This study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
-Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches.
--->
-
-<!-- The objective of this research is to validate and compare the computational methods used to estimate Bayesian stochastic volatility models.
-Specifically, Simulation Based Calibration will be used to identify biases and errors arising from the results of Markov Chain Monte Carlo (MCMC) algorithms.
-The source of any miscalibration include, but are not limited to, estimation strategies, choice of sampler, model specifications and coding mistakes.
-This study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
-Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches.
--->
-
-<!-- Specifically, they propose a method of sampling by using an offset mixture normal approximation with a reweighting procedure to sample from the posterior distribution.
-By linearing the model and using a mixture of Gaussians, they were able to apply a Kalman Filter to estimate the posterior density over the latent volatilities and static parameters.
--->
-
-<!-- The univariate stochastic volatility model described in their paper outlines the state space representation of the the model which is described below.
- -->
-# 2. Background and Literature Review
-
-## 2.1 Stochastic Volatility
+\section{Background and Literature Review}
     Stochastic volatility models are used in financial econometrics to model the variance or volatility of returns in a financial instrument.
@@
 These models can be estimated using frequent frameworks but this research will focus on Bayesian estimation strategies.
+\\
 
@@
 
-### Estimation strategy
-Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother[^1] (de Jong and Shephard, 1995).
-
-[^1]: In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.
+    \subsection{Estimation strategy}
+        Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.} (de Jong and Shephard, 1995).
 
@@
 
-## 2.2 Hamiltoninan Monte Carlo
+    \subsection{Hamiltonian Monte Carlo}
         Since the paper was written, new MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
@@
         The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
-Stan's default algorithm, the No-U-Turn Sampler (Hoffman & Gelman 2014), allows for direct sampling of the specified stochastic volatility model.
+Stan's default algorithm, the No-U-Turn Sampler (Hoffman \& Gelman 2014), allows for direct sampling of the specified stochastic volatility model.
 Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
@@
 
-## 2.3 Research contribution
-The goal of this research is to determine which algorithm is better at estimating the discrete time stochastic volatility model (conditional on the same priors).
-My main contribution is to design a simulation study that compares these algorithms and determines which sampling strategy is better based on the calibration of posterior estimates and the efficiency of the MCMC sampler.
+\section{Methodology}
 
-<!-- New developments in MCMC algorithms and improved computing resources have enabled estimation of richer and more complicated models.
-However, as the development of tools enables us to fit more complicated models, so does the need to validate the output of these tools.
-Cook, Gelman and Rubin (2006) introduce simulation techniques to evaluate the "correctness" of the software used to sample Bayesian models.
-Their contribution identified that cumulative density quantiles from the posterior samples can be tested for uniformity.
-Talts et al (2018).
-build upon this work by developing Simulation Based Calibration to provide more robust validation of output from a MCMC sampler through the use of rank statistics.
--->
-
- <!-- through the use of the CDF quantiles approximated from the posterior distribution. -->
-
-<!-- , through open source programming languages such as Stan and PyMC, have become
-
-As new and complicated
-
-There have been many developments in MCMC algorithms since the introduction of classical methods such as Gibbs and Metropolis Hastings.
-
-Hamiltoninn Monte Carlo (algorithm introduced when, widely available since 2010...?) is a relatively new im.
-
-- Since then, there have been many developments in MCMC algorithms. In particular, HMC provides a flexible alternative to fitting and sampling from complicated high dimensional Bayesian models.
-- It's key innovation is the use of gradients - enabling the flexible specification of any likelihood function. 
-- Constraints- can only sample from continuous 
-- HMC also provides new diagnostics for evaluating the effectiveness of the sampling procedure and how well the sampler is targeting the joint dist.
-
-- How do we compare models? SBC. ESS in the tails, rhats.
-
-- Given this context, my research looks to compare the efficacy of KSCs proposed MCMC sampling strategy with Hamiltonina Monte Carlo (NUTs).  -->
-
-# 3. Methodology
-
-## 3.1 Design
+    \subsection{Design}
         Simulation Based Calibration (SBC) is a technique that checks the calibration of posterior estimates generated by MCMC algorithms.
@@
         To illustrate this procedure, let $\theta$ be a parameter and $y$ represent the dataset. Start with a single draw from the prior distribution:
+
         $$
@@
 
-<!-- n error - either in computation, code or overall analysis.  -->
-
-<!-- The simulation if done correctly should produce samples from the prior distribution.  -->
-
-If the posterior samples follows the prior distribution, the rank statistic for a given parameter follows a discrete uniform distribution[^2]. The rank statistic is defined as:
-
-[^2]: Proof of this result in Talts et al (2018).
+        If the posterior samples follows the prior distribution, the rank statistic for a given parameter follows a discrete uniform distribution\footnote{Proof of this result in Talts et al (2018).}.
+The rank statistic is defined as:
 
@@
         If the computation is well calibrated and the rank statistics follow a discrete uniform distribution, then the posterior credible intervals have sufficient coverage.
-That is, one way to describe calibration is: for any percentage interval selected over the posterior samples (for example 90%) then there is a 90\% chance that $\theta^{sim}$ falls within this interval.
+That is, one way to describe calibration is: for any percentage interval selected over the posterior samples (for example 90\%) then there is a 90\% chance that $\theta^{sim}$ falls within this interval.
 Another way of saying this is a Bayesian analysis is well calibrated if a 90\% credible interval contains the true parameter in 90\% of the SBC iterations.
 
-## 3.2 Implementation
-SBC for Hamiltonian Monte Carlo and the Gaussian approximation are performed with 1000 simulation iterations, 1000 observations in each dataset, and post warmup/burnin samples of 999 (which gives 1000 possible rank statistics).
-This algorithm is summarised below:
-
-| **1) for sim in 1000 iterations:**
-|   2) Draw from prior: $\theta^{sim}\sim\pi (\theta)$
-|   3) Simulate dataset with 1000 observations: $y^{sim} \sim p(y|\theta^{sim})$
-|   4) Draw 999 posterior samples (post warmup) $\{\theta_1,\dots , \theta_{L}\} \sim p(\theta | y^{sim})$
-|   5) Compute rank statistics $r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim})$
-
-## 3.3 Metrics
+    \subsection{Metrics}
         The key metrics and diagnostics to compare the performance of these methods are the effective sample size (ESS), rank statistics, and chi-squared test statistics.
@@
 
-<!-- 
-The aim of this research is to use Simulation Based Calibration (SBC) to validate the inference of posterior samples generated by different MCMC algorithms in the context of stochastic volatility models.
-This will help identify any bias in computation and inconsistencies in model implementation and estimation strategies.
-Comparison of the SBC results between different computational approaches may help identify any strengths or shortcomings as well as any errors in model specification.
--->
+\section{Discussion}
 
-<!-- ## 3.1 Stochastic Volatility Model
-The univariate stochastic volatility model described by Kim, Shephard and Chib (1998) models the variance as a random variable following some latent stochastic process.
-$y_t$ is the mean corrected returns of some asset for equally spaced intervals t.
-$\beta$ is a constant scaling factor which is also defined as $exp(\mu / 2)$ representing instantaneous volatility.
-$h_t$ is log volatility, where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the autoregressive parameter $\phi$ such that $|\phi|<1$.
-This autoregressive parameter represents the persistence or "stickiness" of log volatility and the dispersion parameter $\sigma_{\eta}$ is its variance.
-$\epsilon_t$ and $\eta_t$ are standard normal white noise shocks.
+\subsection{Limitations}
 
-$$
-\begin{aligned}
-y_t =& \space \beta exp(h_t/2) \epsilon_t \\
-h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
-h_1 \sim& \space normal\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
-\end{aligned}
-$$
+\section{Conclusion}
 
-$$
-\begin{aligned}
-\epsilon_t \sim& \space normal(0,1) \\
-\delta_t \sim& \space normal(0,1)
-\end{aligned}
-$$
-
-Setting $\beta=1$, the model can be expressed more succinctly as:
-
-$$
-\begin{aligned}
-y_t \sim& \space normal(0, exp(h_t/2)) \\ 
-h_1 \sim& \space normal \left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
-h_{t+1} \sim& \space normal(\mu +\phi(h_t - \mu) , \sigma_{\eta}^2), \space\space t\neq 1\\ 
-\end{aligned}
-$$
-
-Priors for the static parameters are defined below with conjugate priors on $\mu$ and $\sigma^2$:
-
-$$
-\begin{aligned}
-\mu \sim& \space normal(0, 10^2) \\
-\sigma_{\eta}^2 \sim& \space IG(5/2, (0.01\times 5) / 2) \\
-\phi^{\ast} \sim& \space beta(20, 1.5) \\
-\phi &=  2\phi^{\ast} - 1
-\end{aligned}
-$$
-
-The prior on $\phi$ is a "stretched" beta distribution. This is a beta distribution (as defined on the parameter $\phi^*$) which has been transformed to have support (-1, 1).
-
-## 3.2 Estimation Strategy
-
-### Gaussian Mixture Approximation
-Kim, Shephard and Chib estimate the stochastic volatility model using a Kalman Filter and smoother[^1] (de Jong and Shephard, 1995) to compute the posterior distribution over the latent states.
-This requires the state space model to be linear and conditionally Gaussian.
-
-[^1]: In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.
-
-Since the model is not linear, a transformation is applied by squaring and taking the log of $y_t$.
-
-$$
-\begin{aligned}
-y_t^{*} &= log(y_t^2) \\ 
-&= log((\epsilon_t exp(h_t/2))^2) \\
-&=  log(exp(h_t)) + log(\epsilon_t^2) \\
-&= h_t + log(\epsilon_t^2)  \\
-&= h_t + z_t \\
-\end{aligned}
-$$
-
-Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
-However, it is not simple to sample from this parameterisation of the model.
-Kim, Sherpherd and Chib use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution.
-This is defined by:
-
-$$
-\begin{aligned}
-f(z_t) = \sum_{i=1}^{K} q_if_N(z_i|m_i-1.2704, \nu_i^2)
-\end{aligned}
-$$
-
-Where K is the mixture of normal densities $f_N$, component probabilities $q_i$, mean $m_i-1.2704$ and variance $\nu_i^2$.
-These parameters were selected using moment matching where they found 7 normal densities with varying mean and variance parameters best approximated the log chi squared moments.
-These parameters and weights can be found in Appendix A.
-
-The model can be sampled via the simulation smoother since the model is now linear and conditionally gaussian.
-The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions where as $\phi$ is sampled via a metropolis hastings accept/reject procedure.
-The details can be be around in Appendix B.
-
-### Hamiltonian Monte Carlo (No-U-Turn Sampler)
-The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
-Stan's default algorithm, the No-U-Turn Sampler (Hoffman & Gelman 2014), allows for direct sampling of the specified stochastic volatility model.
-Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
-This approach will use the same priors as specified in the Gaussian mixture approximation.
-However, Hamiltonian Monte Carlo does not sample directly from the analytical form of conjugate or conditionally conjugate distribution, so different priors can be chosen if necessary.
- -->
-
-<!-- SBC validates posterior samples and inferences generated by MCMC algorithms. It identifies inconsistencies and bias in computation and estimation of a generative model. This procedure  -->
-
-<!-- ## Performance diagnostics
-
-The goal of this simulation is to determine the "calibration" of MCMC when estimating SV models.
-
-ADEMP:
-- Aims
-- Data generating mechanism
-    + Write out mathematical form of the model here. Discrete time. 
-- Estimand /target of analysis (maybe after explaining SBC?)
-- Methods
-    + SBC. Evaluating "calibration" of MCMC. What does calibration mean?
-- Performance measures
-    + Rhats, ESS, histogram
-
-- Convergence diagnostic onto target joint posterior distribution? rhats/ess. This assess "efficiency" of MCMC.
-- Tail ESS calculates ESS in the tail quantiles. Such  -->
-
-# 4. Preliminary Results
-Figure 1 shows the distribution of rank statistics for Hamiltonian Monte Carlo, with the horizontal black line at 50 representing the discrete uniform distribution.
-The preliminary figures focus on the static parameters and arbitrarily chosen states.
-Phi and sigma_sqd look relatively uniform.
-However mu and h.995 (the 995th state parameter) have some lumpiness which may suggest a lack of calibration.
-
-![Distribution of parameter ranks for HMC algorithm (1000 iterations)](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/static_state_hist.png){width=75%}
-
-A drawback of SBC is that any deviation in uniformity from a calibrated analysis may be due to noisy estimates from small samples.
-Figure 2 shows more consistent uniform behaviour from all parameters after increasing the number of iterations from 1000 to 5000.
-This suggests that Hamiltonian Monte Carlo and the overall analysis is well calibrated.
-
-![Distribution of parameter ranks for HMC algorithm (5000 iterations)](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r6_5000_iterations/static_state_hist.png){width=75%}
-
-Figure 3 shows results for the Gaussian approximation.
-The distribution of the rank statistics deviates heavily from the expected discrete uniform distribution, with various peaks at each end of the histogram as well as peaks in the middle.
-A reason for this miscalibration may be due to the Gaussian mixture only approximating the log chi squared distribution.
-Hamiltonian Monte Carlo on the other hand, samples directly from the target model which more closely represents the data generating process.
-
-This result may also be due to the relatively small number of posterior samples drawn from the model.
-The MCMC sampler may not have converged yet which may explain the deviations from uniformity.
-A follow up step is to compare the ESS of the two methods and to increase the length of the MCMC chain to ensure the ESS is approximately the same before comparing histograms.
-
-![Distribution of parameter ranks for Gaussian mixture model (1000 iterations)](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png){width=75%}
-
-<!-- An alternative to checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each bin.
-Let $b_j$ be the number of counts and $e_j$ the expected count in bin $j$.
-Then the chi squared statistic is given by:
-
-$$
-\begin{aligned}
-\chi^2 = \sum_{j=1}^J \frac{(b_{j} - e_{j})^2}{e_j}
-\end{aligned}
-$$
-
-These chi squared statistics can be considered draws from a chi squared distribution with 19 degrees of freedom (since there are 20 bins).
-Visualising this should give a chi squared distribution (if performing a hypothesis test, this would be the sampling distribution under a null of uniformity).
-One caveat about this is the chi squared statistics calculated in this way may not be independent.
-That is, each iteration of SBC may be independent however, the calculation of the rank statistics for a given iteration may be dependent between some parameters (since each state parameter is a function of the previous state parameter).
-
-Figure 4 and 5 are the chi squared statistic histograms results of Hamiltonian Monte Carlo and Gaussian approximation respectively.
-The vertical black line marks the 95th quantile which represents the critical value for hypothesis testing.
-The Hamiltonian Monte Carlo chi squared statistics look like they are sampled from the correct distribution, with 6.2\% of the samples greater than the 95\% critical value (doing this calculation for 5000 iterations leads to 4.8\% of samples in the critical area).
-Figure 5 shows the majority of chi squared statistics for Gaussian approximation within the critical area, suggesting the estimation strategy is not well calibrated.
-
-![Chi squared statistics for Hamiltonian Monte Carlo](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/chi_sq_hist.png){width=70%}
-
-![Chi squared statistics for Gaussian approximation](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/chi_sq_hist.png){width=70%} -->
-
-# 5. Conclusion and next steps
-There are two key extensions to the simulation results worth exploring.
-
-**1) Metropolis Hastings reweighting for KSC model**
-
-Kim, Shephard and Chib apply importance sampling weights to correct for the approximation error of the Gaussian mixture model.
-According to the authors, the correction ensures that samples are from the exact posterior density.
-The next step is to perform this correction and evaluate the SBC results.
-I'd expect the Gaussian mixture approximation diagnostics to improve after this correction.
-
-**2) Centered and non centered parameterisation**
-
-The parameterisation of a model can affect the performance of a sampler.
-This can cause significant problems in getting unbiased MCMC samples for complex models.
-An example of this is Neal's funnel (Neal 2003) which arises in hierarchical models which causes performance issues in the Hamiltonian Monte Carlo sampler.
-Rewriting the model with non centered parameterisation significantly improves the performance of MCMC.
-Another consideration for this research is to consider both centered and non centered parameterisations of the stochastic volatility model to see if there are any improvements to the simulation diagnostics.
-
-The models and parameterisations for this research are summarised in the below table:
-
-![Model matrix](model_table.jpg)
-
-## Plan
-Week 7: Reflect and integrate feedback from presentation into plan for the rest of the semester
-
-Week 8: Finish off any simulation or modelling not yet complete (listed in above matrix)
-
-Week 9: Write up results and have a first draft ready for feedback
-
-Week 10-11: Action any feedback, prepare for final submission and presentation
-
-Week 12: Submit
-
-{{< pagebreak >}}
-
-# References
-
-Abril-Pla O, Andreani V, Carroll C, Dong L, Fonnesbeck CJ, Kochurov M, Kumar R, Lao J, Luhmann CC, Martin OA, Osthege M, Vieira R, Wiecki T, Zinkov R.
-(2023) PyMC: a modern, and comprehensive probabilistic programming framework in Python.
-
-Duane, Simon; Kennedy, Anthony D.; Pendleton, Brian J.; Roweth, Duncan (1987). "Hybrid Monte Carlo". Physics Letters B. 195 (2): 216–222.
-
-Hoffman, M. D., & Gelman, A. (2014). The No-U-Turn sampler: adaptively setting path lengths in Hamiltonian Monte Carlo. J. Mach. Learn. Res., 15(1), 1593-1623.
-
-Kim, Sangjoon, Neil Shephard, and Siddhartha Chib. 1998. “Stochastic Volatility: Likelihood Inference and Comparison with ARCH Models.” Review of Economic Studies 65: 361–93.
-
-Neal, R.M. (1995) Bayesian Learning for Neural Networks. Ph.D. Thesis, Graduate Department of Computer Science, University of Toronto, Toronto.
-
-Neal, R. M. (2003). Slice Sampling. The Annals of Statistics, 31(3), 705–741. http://www.jstor.org/stable/3448413
-
-Neal, R.M. (2011) MCMC Using Hamiltonian Dynamics. In Handbook of Markov Chain Mone Carlo, CRC Press, New York
-
-Stan Development Team. 2023. Stan Modeling Language Users Guide and Reference Manual, 2.32. https://mc-stan.org
-
-Talts, S., Betancourt, M., Simpson, D., Vehtari, A., & Gelman, A. (2018). Validating Bayesian inference algorithms with simulation-based calibration. arXiv preprint arXiv:1804.06788.
-
-{{< pagebreak >}}
-
-# Appendix A: Mixture Gaussian weights
-
-| $\omega$ | $Pr(\omega = i)$ | $m_i$ | $\nu^2_i$ |
-|------|------|------|------|
-| 1    | 0.00730    | -10.12999   | 5.79596 | 
-| 2    | 0.10556    | -3.97281    | 2.61369 | 
-| 3    | 0.00002    | -8.56686    | 5.17950 | 
-| 4    | 0.04395    | 2.77786    | 0.16735 | 
-| 5    | 0.34001    | 0.61942    | 0.64009 | 
-| 6    | 0.24566    | 1.79518    | 0.34023 | 
-| 7    | 0.25750    | -1.08819    | 1.26261 |
-
-{{< pagebreak >}}
-
-# Appendix B: Sampling from mixture of Gaussians
-
-## Conjugate posterior distributions
-
-**Sampling** $\boldsymbol{\sigma_{\eta}^2}$
-
-Inverse gamma conjugate posterior distribution:
-
-$$
-\sigma^2_{\eta} | y,h,\phi,\mu \sim IG \Bigl\{\frac{n+\sigma_r}{2}, \frac{0.05 + (h_1 - \mu)^2 (1 - \phi^2) + \sum_{t=1}^{n-1}((h_{t+1} - \mu) - \phi(h_t - \mu))^2}{2}\Bigr\}
-$$
-
-**Sampling** $\boldsymbol{\mu}$
-
-Gaussian conjugate posterior distribution:
-
-$$
-\mu | h,\phi,\sigma^2_{\eta}  \sim N(\hat{\mu}, \sigma^2_{\mu})
-$$
-
-Where
-
-$$
-\begin{aligned}
-\hat{\mu} &= \sigma^2_{\mu} \Bigl\{\frac{(1-\phi^2)}{\sigma_{\eta}^2}h_1 +\frac{(1-\phi^2)}{\sigma_{\eta}^2} \sum_{t=1}^{n-1} (h_{t+1} - \phi h_t)\Bigr\} \\
-\sigma^2_{\mu} &= \sigma^2_{\eta} \{(n-1)(1-\phi)^2 + (1-\phi^2)\}^{-1}
-\end{aligned}
-$$
-
-**Sampling** $\boldsymbol{\phi}$
-
-Metropolis Hastings accept/reject procedure:
-
-1) Generate proposal $\phi^\ast$ from $N(\hat{\phi}, V_{\phi})$ where $\hat{\phi} = \frac{\sum_{t=1}^{n-1} (h_{t+1} - \mu)(h_t - \mu)}{\sum_{t=1}^{n-1} (h_t - \mu)^2}$ and $V_{\phi} = \sigma^2_{\eta} \{\sum_{t=1}^{n-1} (h_t - \mu)^2\}^{-1}$
-
-2) Accept proposal as $\phi^{(i)}$ with probability $e^{\{g(\phi^\ast) - g(\phi^{(i-1)}\}}$ such that $g(\phi) = log (\pi (\phi)) - \frac {(h_t - \mu)^2 (1-\phi^2)}{2 \sigma_{\eta}^2} + \frac{1}{2} log (1-\phi^2)$
-
-**Sampling mixture density:**
-
-Rewrite mixture density with respect to a indicator variable $s_t$
-
-$$
-\begin{aligned}
-&z_t | s_t = i \sim N(m_i - 1.2704, \nu^2) \\
-&Pr(s_t = i) = q_i
-\end{aligned}
-$$
-
-Sample $s_t$ from probability mass function:
-
-$$
-\begin{aligned}
-Pr(s_t = i | y_t^{\ast}, h_t) \propto q_i f_N(y_t^{\ast} | h_t + m_t - 1.2704, \nu^2)
-\end{aligned}
-$$
-
-**Sampling steps**
-
-1. Set initial values $s$, $\phi$, $\sigma^2_{\eta}$ and $\mu$
-
-2. Sample $h$ from $h|y^{\ast}, s, \phi, \sigma^2_{\eta}, \mu$
-
-3. Sample s from $s|y^{\ast}, h$
-
-4. Sample $\mu|y^{\ast}, s, \phi, \sigma^2_{\eta}, h$ according to conjugate posterior distribution
-
-4. Sample $\sigma_{\eta}^2|y^{\ast}, s, \phi, \mu, h$ according to conjugate posterior distribution
-
-4. Sample $\phi|y^{\ast}, s, \mu, \sigma^2_{\eta}, h$ using metropolis hastings
+\end{document}

### 2023-09-16 manuscript/main.tex
added references

@@
 \usepackage{amsmath, amssymb, amsfonts}
+\usepackage[colorlinks = true,
+			citecolor  = blue]{hyperref}
+
+\usepackage[longnamesfirst,authoryear, round]{natbib}
+\bibliographystyle{unsrtnat}
+
 \usepackage[
@@
 
+- Should make reference to workflow, and reparameterisation of SV.
+
 \section{Background and Literature Review}
     Stochastic volatility models are used in financial econometrics to model the variance or volatility of returns in a financial instrument.
-The return of a financial asset often exhibits patterns in the variance which can be modeled to improve prediction and estimates of risk or tail events.
+The return of a financial asset often exhibits patterns in the variance which can be modelled to improve prediction and estimates of risk or tail events.
 Unlike other volatility models such as ARCH or GARCH, stochastic volatility models treat the variance as a random variable.
@@
 These models can be estimated using frequent frameworks but this research will focus on Bayesian estimation strategies.
-\\
 
-    Kim, Shephard and Chib (1998) estimate a discrete time, univariate stochastic volatility model using Bayesian methods which is outlined below.
+    \citet{kim1998stochastic} estimate a discrete time, univariate stochastic volatility model using Bayesian methods which is outlined below.
 $y_t$ is the mean corrected returns of some asset for equally spaced intervals t.
@@
     \subsection{Estimation strategy}
-        Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.} (de Jong and Shephard, 1995).
+        Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.} \citep{dejong1995}.
 
@@
 Specifically, Hamiltonian Monte Carlo is a MCMC algorithm which has become widely available for efficiently sampling from sophisticated models.
-Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature (Duane et al.
-1987) before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks (Neal, 1995) and statistical computing (Neal 2011).
-The algorithm has since become widely available through open source development projects such as Stan (Stan Development Team, 2023) and PyMC (2023).
+Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature \citep{duane1987hybrid} before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks \citep{neal1995bayesian} and statistical computing \citep{neal2011mcmc}.
+The algorithm has since become widely available through open source development projects such as Stan \citep{stan} and PyMC \citep{pymc2023}.
 
@@
         The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
-Stan's default algorithm, the No-U-Turn Sampler (Hoffman \& Gelman 2014), allows for direct sampling of the specified stochastic volatility model.
+Stan's default algorithm, the No-U-Turn Sampler \citep{hoffman2014no}, allows for direct sampling of the specified stochastic volatility model.
 Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
@@
 
-        If the posterior samples follows the prior distribution, the rank statistic for a given parameter follows a discrete uniform distribution\footnote{Proof of this result in Talts et al (2018).}.
+        If the posterior samples follows the prior distribution, the rank statistic for a given parameter follows a discrete uniform distribution\footnote{Proof of this result in \citet{talts2018validating}.}.
 The rank statistic is defined as:
@@
 
+% \bibliographystyle{apa} % Econometrica referencing
+\bibliography{references}
+
 \end{document}

### 2023-09-17 src/gaussianMixture.py
set stationary initialisation - checked correct

@@ -33,7 +33,7 @@ class TVLLDT(sm.tsa.statespace.MLEModel):
 
         # Initialize base model
         super(TVLLDT, self).__init__(endog, k_states=1, k_posdef=1,
-                                     initialization='known', initial_state =[0], initial_state_cov=np.array([[0.02/(1-0.95**2)]]))
+                                     initialization='stationary')
 
         # Setup time-varying arrays for observation equation
         self['obs_intercept'] = np.zeros((1, self.nobs)) # mod.ssm.obs_intercept

### 2023-09-20 manuscript/main.tex
Updates from Overleaf
Added parametiersation section

@@
 \documentclass[12pt, a4paper]{article}
+\setcounter{tocdepth}{2}
 \usepackage{graphicx} % Required for inserting images
@@
 
-\begin{abstract}
+\begin{abstract}.  
     The objective of this research is to compare the computational methods used to estimate Bayesian stochastic volatility models.
@@
 
+\newpage
+
+\tableofcontents{\protect\newpage}
+
 \section{Introduction}
@@
 \section{Background and Literature Review}
+
+    \subsection{Stochastic Volatility}
         Stochastic volatility models are used in financial econometrics to model the variance or volatility of returns in a financial instrument.
@@
 
-    \subsection{Estimation strategy}
+    \subsection{Gaussian Mixture Approximation}
         Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.} \citep{dejong1995}.
@@
 
-    \subsection{Design}
+    \subsection{Simulation Design}
         Simulation Based Calibration (SBC) is a technique that checks the calibration of posterior estimates generated by MCMC algorithms.
@@
         ESS measures the efficiency of the MCMC sampler.
-It calculates the number of (effectively) independent draws from the posterior draws generated by a Markov chain.
+It calculates the number of effectively independent draws from the posterior draws generated by a Markov chain.
 A poor ESS can arise from high autocorrelation in the Markov chain which leads to highly dependent samples.
@@
 
+    \subsection{Model Parameterisation}
+
+        The parameterisation of a model can affect the performance of a MCMC algorithm when sampling from models with complex posterior geometries.
+An example of this is Neal's funnel \citep{neal2003slice} where the Hamiltonian Monte Carlo sampler encounters performance issues in hierarchical models and produces biased samples.
+
+        The stochastic volatility model as described in section 2.1 is follows a "centered" parameterisation.
+This describes the central location of the latent state parameters which are centered around the mean and lag of the log volatility $\mu +\phi(h_t - \mu)$.
+A model can be reparameterised such that the states are sampled on a distribution centered on 0 (i.e "non centered") which are later transformed to have the correct mean.
+It is useful to compare model parameterisations since sampling strategies may be sensitive to the posterior geometry.
+
+        \subsubsection{Non centered Hamiltonian Monte Carlo}
+        First sample from a standard normal distribuition and multiply by the variance of the log volatility.
+
+        $$
+        \begin{aligned}
+        h_{std} \sim& \space normal(0,1) \\
+        h =& h_{std} \times \sigma_{\eta} \\ 
+        h \sim& normal(0, \sigma_{\eta}^2)
+        \end{aligned}
+        $$
+
+        The states are sampled with the mean centered at zero. Then apply the appropriate rescaling to get samples from the log volatility.
+
+        $$
+        \begin{aligned}
+        h_1 =& \space \frac{h_{std, 1}\times \sigma_{\eta}} {\sqrt{1 - \phi^2}} + \mu \\
+        h_{t+1} =& \space h_{std, t+1}\times \sigma_{\eta} + \mu  + \phi(h_{t} - \mu),\space t\neq 1
+        \end{aligned}
+        $$
+
+        This returns log volatility as desired.
+
+        $$
+        \begin{aligned}
+        h_1 \sim& \space normal \left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
+        h_{t+1} \sim& \space normal(\mu +\phi(h_t - \mu) , \sigma_{\eta}^2), \space\space t\neq 1\\ 
+        \end{aligned}
+        $$
+
+        \subsubsection{Non centered Gaussian Mixture}
+        The non centered model for the Gaussian approximation is expressed slightly differently due to the use of the Kalman Filter.
+While HMC can sample from the joint posterior directly with state vectors centered at zero, applying the Kalman Filter requires the state equation to be a function of the state variable.
+
+        Starting with the log chi squared model:
+
+        $$
+        \begin{aligned}
+        y^{\ast}_t =& h_t + z_t \\
+        h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t
+        \end{aligned}
+        $$
+
+        Expand the state equation:
+
+        $$
+        \begin{aligned}
+        h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
+        =& \space \mu + \phi h_t - \phi\mu + \sigma_{\eta} \eta_t  \\
+        =& \space \mu (1- \phi) + \phi h_t + \sigma_{\eta} \eta_t  \\
+        \end{aligned}
+        $$
+
+        Take out the intercept and put it in the measurement equation:
+
+        $$
+        \begin{aligned}
+        y^{\ast}_t =& h_t + \mu (1- \phi) + z_t \\
+        h^{\ast}_{t+1} =& \space \phi h_t + \sigma_{\eta} \eta_t  \\
+        \end{aligned}
+        $$
+
+        The mean of the log volatility is now inside the measurement equation and de-meaned from the state equation.
+
+        \subsubsection{Prior Predictive check}
+        It is worth noting that the centered and non-centered parameteristions are the same stochastic volatility model.
+Different parameteristions express the same mathematical models in ways that may make it easier or harder for a given algorithm to sample.
+
+        To demonstrate this, a prior predictive simulation of the second state parameter is performed for both a centered and non centered model (using the paramterisation described in section 3.3.1).
+This is the same as the first two steps of SBC, except instead of generating a dataset, generate the parameters implied by the prior draw.
+Let $\boldsymbol{\theta}$ be a vector of static parameters.
+
+        $$
+        \begin{aligned}
+        \boldsymbol{\theta}^{sim} \sim \pi(\boldsymbol{\theta})
+        \end{aligned}
+        $$
+
+        Generate a the second state parameter implied by the joint prior.
+
+        $$
+        \begin{aligned}
+        h_2^{sim} \sim \pi (h_2|h_1, \boldsymbol{\theta}^{sim})
+        \end{aligned}
+        $$
+
+        Generating 1000 samples of $h_2^{sim}$ from both centered and non-centered models gives samples from the same data generating process as shown in figure x.
+
+        % \begin{figure}
+        %     \centering
+        %     \includegraphics[scale=0.1]{ppc_h2.png}
+        %     \caption{Prior predictive samples of second state variable}
+        % \end{figure}
+
+        % Emphasise here that different paramterisations of the models are the same model
+        % Use a prior predictive check from both paramterisations show that the results are
+        % the same
+
+\section{Results}
+
+    \subsection{Hamiltonian Monte Carlo}
+
+    \subsection{Gaussian Mixture Approximation}
+
+    \subsection{Reweigthing rank statistics of Gaussian Mixture}
+
 \section{Discussion}
@@
 
+\newpage
+
 % \bibliographystyle{apa} % Econometrica referencing

### 2023-09-21 R/sbc.r
added ess weights

@@ -23,7 +23,7 @@ load_parameters <- function(rds_path) {
 }
 
 facet_hist <- function(data, variables, nbins, expected_bin_count, rank_scales) {
-    rank_bins %>%
+    data %>%
         select(all_of(variables)) %>%
         tidyr::pivot_longer(everything(), names_to = "variable", values_to = "rank") %>%
         ggplot(.) +
@@ -154,4 +154,10 @@ rhat_boxplot <- function(data, variables, plot_title, rhat_type) {
         geom_boxplot(aes(x = parameters, y = {{rhat_type}})) +
         theme_minimal() +
         labs(title = plot_title)
+}
+
+get_ess_weights <- function(rds_path, model_path) {
+    data = readRDS(here::here(paste(model_path, rds_path, sep = "/")))
+    results = data[["weights_ess"]]
+    return(results)
 }

### 2023-09-21 sim_ksc.py
SIR resampling

@@ -102,7 +102,14 @@ def simulation(seed_data):
     burn_draws = np.concatenate((trace_mu, trace_phi, trace_sigma2, trace_states), axis = 1)
     param_draws = burn_draws[config["burn"]+1:burn_draws.shape[0]]
     draws = np.concatenate((param_draws, weights.reshape(-1,1)), axis = 1)
-    samples = pd.DataFrame(draws)
+
+    if config['reweight_samples']:
+        n_samples = weights.shape[0]
+        resampled_indexes = np.random.choice(np.arange(0,n_samples), size=n_samples, replace=True, p=weights)
+        resampled_draws = draws[resampled_indexes]
+        samples = pd.DataFrame(resampled_draws)
+    else:
+        samples = pd.DataFrame(draws)
     
     static_names = ['mu', 'phi', 'sigma2']
     state_names = [f"h[{state}]" for state in np.arange(1, trace_states.shape[1]+1)]

### 2023-09-23 manuscript/main.tex
remove dot
added resarch context and intro.

@@
 
-\begin{abstract}.  
+\begin{abstract}
     The objective of this research is to compare the computational methods used to estimate Bayesian stochastic volatility models.
@@
 
-- Should make reference to workflow, and reparameterisation of SV.
+Stochastic volatility is a type of volatility model which explicitly treats the variance as a latent random variable.
+These are typically expressed as non linear Gaussian state space models and are difficult to estimate using classical methods.
+The likelihood is unavailable in closed form and there are at least as many variables as data points.
+Bayesian approaches to estimating these models rely on computational Markov Chain Monte Carlo (MCMC) techniques to sample from the target joint posterior of these high dimensional latent parameter spaces.
 
-\section{Background and Literature Review}
+Kim, Shephard and Chib (1998) published a method for sampling this model using a combination of Kalman Filters, Gibbs, and Metropolis Hastings over 20 years ago.
+Since then, there have been many advances in Bayesian, and more broadly, statistical computation to estimate complex, high dimensional state space models.
+In particular, the recent availability of Hamiltonian Monte Carlo, a variant of the Metropolis Hastings algorithm, is a advancement of MCMC for efficient sampling of complex probabilistic models.
+The adoption of such new techniques have been made widely available through the delveopment of various open source libraries such as the Stan programming language and the PyMC library.
+Other algorithms have also improved the speed of estimating high dimensional models using approximations of the posterior such as Integrated nested Laplace approximation and Approximate Bayesian Computation.
+
+These conceptually and practically different techniques to estimate complex models did not exist 20 years ago (or perhaps more accurately, were not easily accessible or implemented 20 years ago).
+As the development of new algorithms rapidly increase, so does the need to develop new methods to assess their output.
+How do we develop our workflow to assesss new sampling algorithms?
+And how can we compare the comptuational strategies used to estimate increasingly complex models?
+
+% - SV models and more broadly state space models are complex, high dimensional models which require computational tools to estimate them.
+% - Modelling variance useful and important?
+
+% - Advanced in Bayesian computation
+%     + Conceptually and practically diff techniques to estimate complex models didn't exist not that long ago
+%     + Performance metrics useful for checking fit on single dataset. However not much emphasis on checking performance between MCMC
+
+% - A principled Bayesian workflow which usually requires checking of multiple componenets: prior, posterior, model choice. 
+%     + Lots of different components which could contribute to an incorrect fit or error.
+
+% - This research focuses on evaluating the computational tools used to esitmate a specific type of model.
+
+% % One para
+% - This research </contributes> examine the algos used to estimate SV model 
+%     + Particular focus is on the computational tools used to estimate these models and whether they are returning what we expect.
+%     + conditional on the same prior
+
+%     - Additionally, check the sensitivity of the simulation and algortihms to the parameterisation of the model.
+Taking idea and advice from Catherine's paper about improvements in sampling performance.
+
+% %  2nd 
+% - Investigating two algorithms, HMC and GMA from KSC.
+
+% % 3
+% - Two samplings methods are used and are compared through simulation
+%     + Specifically, the simulation designed which is taken from talts et al.
+
+% % 4
+% - Key metrics of interests are ESS for efficiency and tests of uniformity for calibration.
+
+% - Goal is to show which algo and parameterisation is "better".
+Moreover, want to demonstarte that this part of a Bayesian workflow which can be applied more generally is important for checking that the model is returning the correct posteriros - and if not, where the problem may be occurring and why.
+Give a set of tools to evaluate algos and parametiersations (before real data is even considered).
+**Why is this important.**
+
+% - Sections outlined below:
+
+\section{Resarch Context}
 
 \subsection{Stochastic Volatility}
-        Stochastic volatility models are used in financial econometrics to model the variance or volatility of returns in a financial instrument.
-The return of a financial asset often exhibits patterns in the variance which can be modelled to improve prediction and estimates of risk or tail events.
-Unlike other volatility models such as ARCH or GARCH, stochastic volatility models treat the variance as a random variable.
-Furthermore, these models have a state space representation since the variance is expressed as a latent stochastic process.
-State space models can be difficult to estimate since there is a latent state for every data observation in the time series, which means there are often more parameters than data points.
-These models can be estimated using frequent frameworks but this research will focus on Bayesian estimation strategies.
-
     \citet{kim1998stochastic} estimate a discrete time, univariate stochastic volatility model using Bayesian methods which is outlined below.
@@
 
-    \subsection{Gaussian Mixture Approximation}
-        Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.} \citep{dejong1995}.
+\subsection{Diagnostic limitations on real data}
+The stochastic volatility model can be estimated using a variety of sampling strategies and MCMC algorithms.
+A common set of metrics are used to check the convergence of the MCMC.
+This includes checking the effective sample size and the $\hat{R}$ calculations for estimated chains and parameters.
+These diagnostics check whether an MCMC has failed to converge onto the target posterior distribution (since we can never know with certainty that a sampler has converged).
 
-        The standard Kalman Filter and smoother is used to compute the posterior distribution over the latent states.
-This requires the state and measurement equations to be linear and conditionally Gaussian.
-Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring and taking the log of $y_t$.
+These diagnostics are useful for identifying when Bayesian computation fail on real data.
+However, confounding issues may arrise when trying to diagnose the cause of computational problems.
+Real data is generated from an unknown data generating process.
+That is, the true parameter or model is unobservable.
+Therefore, failed diagnostic checks could arise from either incorrect model specification or issues in the sampling strategy (algorithm) or both.
 
-        $$
-        \begin{aligned}
-        y_t^{*} &= log(y_t^2) \\ 
-        &= log((\epsilon_t exp(h_t/2))^2) \\
-        &=  log(exp(h_t)) + log(\epsilon_t^2) \\
-        &= h_t + log(\epsilon_t^2)  \\
-        &= h_t + z_t \\
-        \end{aligned}
-        $$
+Furthermore, different sampling strategies and MCMC tools could provide different estimates for the same model.
+Assuming no issues in computation, these diagnostic measures do not provide any evidence to which estimate is closer to the truth.
 
-        Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
-The relationship between $y_t$ and $h_t$ is now linear; however, the error is not Gaussian.
-Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution.
-This is defined by:
+To illustrate this problem, suppose we estimate this model with two different sampling strategies on real data.
+The data are the demeaned log returns of the 2019-2023 S\&P 500 Index.
+Below are the marginal posterior estimates for the static parameters $\mu$, $\phi$ and $\sigma^2$ and corresponding summary statistics.
 
-        $$
-        \begin{aligned}
-        f(z_t) = \sum_{i=1}^{K} q_if_N(z_i|m_i-1.2704, \nu_i^2)
-        \end{aligned}
-        $$
+**Insert results here**
 
-        Where K is the mixture of normal densities $f_N$, component probabilities $q_i$, mean $m_i-1.2704$ and variance $\nu_i^2$.
-These parameters were selected using moment matching where they found 7 normal densities with varying mean and variance parameters best approximated the log chi squared moments.
-These parameters and weights can be found in Appendix A.
+These particular estimates show the marginal posteriors of each parameter sharing similar shapes.
+$\phi$ and $\sigma^2$ possess slightly heavier right tails for SM1 compared to SM2.
+While in this particular example there is no major difference in the estimates, there is no certainty in whether SM1 or SM2 is consistently closer to the truth.
+There are other ways of model or algorithmic selection in this context (for example, out of sample predictive performance), but for parameter estimation it is unclear which one should be selected.
 
-        The model can be sampled via the Kalman Filter and simulation smoother since the model is now linear and conditionally gaussian.
-The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions whereas $\phi$ is sampled via a Metropolis Hastings accept/reject procedure.
-The details can be around in Appendix B.
+\subsection{Diagnostic limitations on a single simulation}
 
-    \subsection{Hamiltonian Monte Carlo}
-        Since the paper was written, new MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
-Specifically, Hamiltonian Monte Carlo is a MCMC algorithm which has become widely available for efficiently sampling from sophisticated models.
-Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature \citep{duane1987hybrid} before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks \citep{neal1995bayesian} and statistical computing \citep{neal2011mcmc}.
-The algorithm has since become widely available through open source development projects such as Stan \citep{stan} and PyMC \citep{pymc2023}.
+Diagnosing computational problems on real data is difficult due to confounding factors.
+A strategy around this is to evaluate a model and sampling strategy on simulated data.
+One approach is to simulate data from the proposed model using known parameters.
+Then, fit the model on the simulated data and see if the model can recover the true parameters.
+This gives us the benefit of knowing or defining the data generating process to be estimated.
+If the model and algorithm cannot adequately capture the true parameter, then we cannot be confident that it will provide reliable estimates on real data.
 
-        The key innovation of Hamiltonian Monte Carlo is using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
-Unlike Random Walk Metropolis Hastings, it takes advantage of the geometry of the posterior to determine its next proposal step.
-A comprehensive explanation of the sampler is beyond the scope of this research and can be found in the above references.
+Furthermore, as discussed in cite"Bayesian workflow", fitting models on simulated data is the only way we can check inference on latent variables.
+This is critical in Stochastic Volatility since the underlying framework is a state space model with latent log volatility parameters.
+Latent variables are unobserved in real data and are only estimated in the context of the model.
+Simulation gives control over the data generation which governs the latent process and reveals what the model can infer about the latent variables.
 
-        The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
-Stan's default algorithm, the No-U-Turn Sampler \citep{hoffman2014no}, allows for direct sampling of the specified stochastic volatility model.
-Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
-This approach will also use the same priors as specified in the Gaussian mixture approximation.
+While simulations enable us to check whether our models and algorithms can capture the true generating process, there are limitations to what can be learned from a single simulation.
+There is always a small probability that the true parameter is in the tails of any estimated posterior distribution.
+Or put differently, there is a 1\% chance the true parameter or any random draw exists outside a 99\% credible interval.
+cite"Talts et al" make the point that a single simulation does not provide sufficient information about the inference made by an algorithm.
+As discussed in their paper, a single simulation may conclude "that an incorrectly coded analysis worked as desired, while a correctly coded analysis failed".
+
+The below results are from a single simulation where data is generated and sampled from the stochastic volatility model with known parameters.
+
+**Insert results here**
+
+The posteriors for x and y sufficiently cover the true parameter, however, the true parameter for z is deep in the tails.
+Such an analysis may incorrectly conclude that the model fails to adequately estimate the z parameter.
+However, it may be the case that the posterior distribution is correctly calculated using this algorithm and the results may be due to the nuances of this specific simulated dataset.
+
+\subsection{Research Goal}
+The objective of this research is to design a simulation study to evaluate the calibration of algorithms used to sample Stochastic Volatility models.
+Indeed, there are limitations to evaluating MCMC algorithms based on fits to real data and single simulations.
+To check the calibration of an algorithm, simulations over multiple "true parameters" are required.
+This methodology is discussed in the next session.
+
+Additionally, two sampling strategies will be assessed using the proposed simulation strategy as well as the performance of the samplers under different parameterisations of the stochastic volatility model.
+Not only will the calibration of any one algorithm be assessed, but also compared across sampling strategies to determine which MCMC approach is most suitable for estimating this model.
 
@@
         Chi squared statistics can be used to test for uniformity and thus calibration of all the parameters in the model.
+
+\section{MCMC Strategies}
+
+    \subsection{Sampling method 1}
+        Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.} \citep{dejong1995}.
+
+        The standard Kalman Filter and smoother is used to compute the posterior distribution over the latent states.
+This requires the state and measurement equations to be linear and conditionally Gaussian.
+Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring and taking the log of $y_t$.
+
+        $$
+        \begin{aligned}
+        y_t^{*} &= log(y_t^2) \\ 
+        &= log((\epsilon_t exp(h_t/2))^2) \\
+        &=  log(exp(h_t)) + log(\epsilon_t^2) \\
+        &= h_t + log(\epsilon_t^2)  \\
+        &= h_t + z_t \\
+        \end{aligned}
+        $$
+
+        Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
+The relationship between $y_t$ and $h_t$ is now linear; however, the error is not Gaussian.
+Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution.
+This is defined by:
+
+        $$
+        \begin{aligned}
+        f(z_t) = \sum_{i=1}^{K} q_if_N(z_i|m_i-1.2704, \nu_i^2)
+        \end{aligned}
+        $$
+
+        Where K is the mixture of normal densities $f_N$, component probabilities $q_i$, mean $m_i-1.2704$ and variance $\nu_i^2$.
+These parameters were selected using moment matching where they found 7 normal densities with varying mean and variance parameters best approximated the log chi squared moments.
+These parameters and weights can be found in Appendix A.
+
+        The model can be sampled via the Kalman Filter and simulation smoother since the model is now linear and conditionally gaussian.
+The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions whereas $\phi$ is sampled via a Metropolis Hastings accept/reject procedure.
+The details can be around in Appendix B.
+
+    \subsection{Sampling method 2}
+    % Need much more detail here about how it works
+        Since the paper was written, new MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
+Specifically, Hamiltonian Monte Carlo is a MCMC algorithm which has become widely available for efficiently sampling from sophisticated models.
+Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature \citep{duane1987hybrid} before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks \citep{neal1995bayesian} and statistical computing \citep{neal2011mcmc}.
+The algorithm has since become widely available through open source development projects such as Stan \citep{stan} and PyMC \citep{pymc2023}.
+
+        The key innovation of Hamiltonian Monte Carlo is using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
+Unlike Random Walk Metropolis Hastings, it takes advantage of the geometry of the posterior to determine its next proposal step.
+A comprehensive explanation of the sampler is beyond the scope of this research and can be found in the above references.
+
+        The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
+Stan's default algorithm, the No-U-Turn Sampler \citep{hoffman2014no}, allows for direct sampling of the specified stochastic volatility model.
+Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
+This approach will also use the same priors as specified in the Gaussian mixture approximation.
 

### 2023-09-23 manuscript/structure/draft.qmd
added resarch context and intro.

--- 2023-09-23 manuscript/main.tex
@@
-\documentclass[12pt, a4paper]{article}
-\setcounter{tocdepth}{2}
-\usepackage{graphicx} % Required for inserting images
-\usepackage[parfill]{parskip} % Removes indent from new paragraphs
-\usepackage{amsmath, amssymb, amsfonts}
-\usepackage[colorlinks = true,
-			citecolor  = blue]{hyperref}
-
-\usepackage[longnamesfirst,authoryear, round]{natbib}
-\bibliographystyle{unsrtnat}
-
-\usepackage[
-top    = 1.0in,
-bottom = 1.2in,
-left   = 1.0in,
-right  = 1.0in]{geometry}
-\linespread{1.5}
-
-\title{Comparison of MCMC algorithms in Stochastic Volatility Models}
-\author{Benjamin Wee}
-\date{October 2023}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
+# Abstract
 The objective of this research is to compare the computational methods used to estimate Bayesian stochastic volatility models.
@@
 This will determine which algorithm is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
-\end{abstract}
 
-\newpage
-
-\tableofcontents{\protect\newpage}
-
-\section{Introduction}
-
+# Introduction
 Stochastic volatility is a type of volatility model which explicitly treats the variance as a latent random variable.
@@
 As the development of new algorithms rapidly increase, so does the need to develop new methods to assess their output.
-How do we develop our workflow to assesss new sampling algorithms?
-And how can we compare the comptuational strategies used to estimate increasingly complex models?
+Developments in our statistical workfow are required to assess new sampling algorithms as well as compare computational strategies used to estimate increasingly complex models.
 
-% - SV models and more broadly state space models are complex, high dimensional models which require computational tools to estimate them.
-% - Modelling variance useful and important?
+This research implements a simulation study to assess the algorithms used to estimate stochastic volatility models.
+Simulation Based Calibration outlined in Talts et al.
+is the primary study design to determine whether the MCMC strategy are returning efficient and well calibrated posterior estimates.
 
-% - Advanced in Bayesian computation
-%     + Conceptually and practically diff techniques to estimate complex models didn't exist not that long ago
-%     + Performance metrics useful for checking fit on single dataset. However not much emphasis on checking performance between MCMC
+Two sampling approaches are considered in this study - Kalman Filter and Metropolis Hastings estimation by Kim Sherphard and Chib, and the Hamiltoninan Monte Carlo algorithim.
 
-% - A principled Bayesian workflow which usually requires checking of multiple componenets: prior, posterior, model choice. 
-%     + Lots of different components which could contribute to an incorrect fit or error.
+Key metrics used to evaluate the model are the effective sample size and visual tests of uniform rank statistics.
 
-% - This research focuses on evaluating the computational tools used to esitmate a specific type of model.
+<!-- 
+ How do we develop our workflow to assesss new sampling algorithms? And how can we compare the comptuational strategies used to estimate increasingly complex models? -->
 
-% % One para
-% - This research </contributes> examine the algos used to estimate SV model 
-%     + Particular focus is on the computational tools used to estimate these models and whether they are returning what we expect.
-%     + conditional on the same prior
+<!-- How can we check whether these algorithms are returning the correct posterior estimates?  -->
+are complex, high dimensional models that can be expressed in the state space framework.
 
-%     - Additionally, check the sensitivity of the simulation and algortihms to the parameterisation of the model.
+A variety of MCMC strategies exist in the literature for estimating these models such as particle filters, Kalman Filters and Metropolis Hastings algorithms.
+
+and more broadly state space models are complex, high dimensional models which require computational tools to estimate them.
+- Modelling variance useful and important?
+
+- Advanced in Bayesian computation
+    + Conceptually and practically diff techniques to estimate complex models didn't exist not that long ago
+    + Performance metrics useful for checking fit on single dataset. However not much emphasis on checking performance between MCMC
+
+- A principled Bayesian workflow which usually requires checking of multiple componenets: prior, posterior, model choice. 
+    + Lots of different components which could contribute to an incorrect fit or error.
+
+- This research focuses on evaluating the computational tools used to esitmate a specific type of model.
+
+% One para
+- This research </contributes> examine the algos used to estimate SV model 
+    + Particular focus is on the computational tools used to estimate these models and whether they are returning what we expect.
+    + conditional on the same prior
+
+    - Additionally, check the sensitivity of the simulation and algortihms to the parameterisation of the model.
 Taking idea and advice from Catherine's paper about improvements in sampling performance.
 
-% %  2nd 
-% - Investigating two algorithms, HMC and GMA from KSC.
+%  2nd 
+- Investigating two algorithms, HMC and GMA from KSC.
 
-% % 3
-% - Two samplings methods are used and are compared through simulation
-%     + Specifically, the simulation designed which is taken from talts et al.
+% 3
+- Two samplings methods are used and are compared through simulation
+    + Specifically, the simulation designed which is taken from talts et al.
 
-% % 4
-% - Key metrics of interests are ESS for efficiency and tests of uniformity for calibration.
+% 4
+- Key metrics of interests are ESS for efficiency and tests of uniformity for calibration.
 
-% - Goal is to show which algo and parameterisation is "better".
+- Goal is to show which algo and parameterisation is "better".
 Moreover, want to demonstarte that this part of a Bayesian workflow which can be applied more generally is important for checking that the model is returning the correct posteriros - and if not, where the problem may be occurring and why.
@@
 
-% - Sections outlined below:
+- Sections outlined below:
 
-\section{Resarch Context}
-
-\subsection{Stochastic Volatility}
-    \citet{kim1998stochastic} estimate a discrete time, univariate stochastic volatility model using Bayesian methods which is outlined below.
+# Stochastic Volatility
+The model of interest in this research is the discrete time, univariate stochastic volatility model outlined by \citet{kim1998stochastic} who use Bayesian estimation methods.
 $y_t$ is the mean corrected returns of some asset for equally spaced intervals t.
@@
 
-\subsection{Diagnostic limitations on real data}
+## Diagnostic limitations on real data
 The stochastic volatility model can be estimated using a variety of sampling strategies and MCMC algorithms.
@@
 
-These diagnostics are useful for identifying when Bayesian computation fail on real data.
-However, confounding issues may arrise when trying to diagnose the cause of computational problems.
+Convergence diagnostics are useful for identifying when Bayesian computation fail on real data.
+However, confounding issues may arise when trying to diagnose the _cause_ of computational problems.
 Real data is generated from an unknown data generating process.
 That is, the true parameter or model is unobservable.
-Therefore, failed diagnostic checks could arise from either incorrect model specification or issues in the sampling strategy (algorithm) or both.
+Therefore, failed diagnostic checks could arise from either inocrrect model specification or issues in the sampling strategy (algorithm) or both.
 
@@
 To illustrate this problem, suppose we estimate this model with two different sampling strategies on real data.
-The data are the demeaned log returns of the 2019-2023 S\&P 500 Index.
+The data are the demaned log returns of the 2019-2023 S&P 500 Index.
 Below are the marginal posterior estimates for the static parameters $\mu$, $\phi$ and $\sigma^2$ and corresponding summary statistics.
@@
 $\phi$ and $\sigma^2$ possess slightly heavier right tails for SM1 compared to SM2.
-While in this particular example there is no major difference in the estimates, there is no certainty in whether SM1 or SM2 is consistently closer to the truth.
-There are other ways of model or algorithmic selection in this context (for example, out of sample predictive performance), but for parameter estimation it is unclear which one should be selected.
+While in this particular example there is no major difference in the estimates, there is no certainty in whether SM1 or SM2 is consistenty closer to the truth.
+There are other ways of model or algorithmic selection in this context (for example, out of sample predictive performance), but for parameter _estimation_ it is unclear which one should be selected.
 
-\subsection{Diagnostic limitations on a single simulation}
+<!-- Could simply just pick the one which has better out of sample performance. Although in the case when computation fails, we wouldn't -->
+<!-- 
+- Workflwo pg 16: "Modeling issues can collide with computational issues" -  can't tell if problem is computation or the model (and whether it's "correct" or capturing the correct DGP). -->
 
+## Diagnostic limitations on a single simulation
 Diagnosing computational problems on real data is difficult due to confounding factors.
@@
 
-Furthermore, as discussed in cite"Bayesian workflow", fitting models on simulated data is the only way we can check inference on latent variables.
+Furthermore, as discussed in \cite{Bayesian workflow}, fitting models on simulated data is the only way we can check inference on latent variables.
 This is critical in Stochastic Volatility since the underlying framework is a state space model with latent log volatility parameters.
@@
 Or put differently, there is a 1\% chance the true parameter or any random draw exists outside a 99\% credible interval.
-cite"Talts et al" make the point that a single simulation does not provide sufficient information about the inference made by an algorithm.
+\cite{Talts et al} make the point that a single simulation does not provide sufficient information about the inference made by an algorithm.
 As discussed in their paper, a single simulation may conclude "that an incorrectly coded analysis worked as desired, while a correctly coded analysis failed".
@@
 The posteriors for x and y sufficiently cover the true parameter, however, the true parameter for z is deep in the tails.
-Such an analysis may incorrectly conclude that the model fails to adequately estimate the z parameter.
+Such an analysis may incorrectly conclude that the model fails to adequatly estimate the z parameter.
 However, it may be the case that the posterior distribution is correctly calculated using this algorithm and the results may be due to the nuances of this specific simulated dataset.
 
-\subsection{Research Goal}
+<!-- Workflow:
+- page 17: simulate data from model with fixed known parameers and see whether method comes close to reproducing known truth.
+If the model cannot sufficiently capture the truth when the truth is known, then we cannot rely on estimates made on real data.
+- pg 18 "Fake-data simulations provide an upper bound of what can be learned about a latent process" [since this is unobserved] 
+Multiple ground truths needed.
+-->
+<!-- 
+- pg 18: Bayes inferencconly calibrated when averaging over the prior - not for any one parameter alue (single simulation anything can happen, i.e a 5\% chance a random draw outside the 95\% interval)
+SBC:
+- pg 2:The behaviour of any individual siulation will not characterise the bility of the inference to fit the model in any meaningful way.
+May prematurely conclude that an incorrectly coded analysis works as desired or a correctly coded analysis failed.
+Need to consider multiple ground truths  -->
+
+## Research Goal
 The objective of this research is to design a simulation study to evaluate the calibration of algorithms used to sample Stochastic Volatility models.
@@
 
-\section{Methodology}
+<!-- 
+- Need to consider multiple ground truths
+- Apply SBC to these 2 algorithms to assess their calibration.
+Tool allows us to not only assess calibration for a single algorithm but enables comparison between strategies (Not only get to validate any one algo but get to compare across algo.) -->
 
-    \subsection{Simulation Design}
-        Simulation Based Calibration (SBC) is a technique that checks the calibration of posterior estimates generated by MCMC algorithms.
+# Methodology
+
+## Simulation Design
+Simulation Based Calibration (SBC) checks the calibration of posterior estimates generated by MCMC algorithms.
 SBC is conducted by comparing the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.
@@
 
-    \subsection{Metrics}
+## Metrics
 The key metrics and diagnostics to compare the performance of these methods are the effective sample size (ESS), rank statistics, and chi-squared test statistics.
@@
 
-\section{MCMC Strategies}
+# MCMC Strategies
 
-    \subsection{Sampling method 1}
+## Sampling method 1
 Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.} \citep{dejong1995}.
@@
 
-    \subsection{Sampling method 2}
+## Sampling method 2
     % Need much more detail here about how it works
@@
 
-    \subsection{Model Parameterisation}
-
+## Model Parameterisation
 The parameterisation of a model can affect the performance of a MCMC algorithm when sampling from models with complex posterior geometries.
@@
 
-        \subsubsection{Non centered Hamiltonian Monte Carlo}
+### Non centered Hamiltonian Monte Carlo
 First sample from a standard normal distribuition and multiply by the variance of the log volatility.
@@
 
-        \subsubsection{Non centered Gaussian Mixture}
+### Non centered Gaussian Mixture
 The non centered model for the Gaussian approximation is expressed slightly differently due to the use of the Kalman Filter.
@@
 
-        \subsubsection{Prior Predictive check}
+### Prior Predictive check
 It is worth noting that the centered and non-centered parameteristions are the same stochastic volatility model.
@@
 
-\section{Results}
+# Results
 
-    \subsection{Hamiltonian Monte Carlo}
+## Hamiltonian Monte Carlo
 
-    \subsection{Gaussian Mixture Approximation}
+## Gaussian Mixture Approximation
 
-    \subsection{Reweigthing rank statistics of Gaussian Mixture}
+## Reweigthing rank statistics of Gaussian Mixture
 
-\section{Discussion}
+# Discussion
 
-\subsection{Limitations}
+## Limitations
+- Weakness of SBC: 
+    + Model uses very uninofmrative priors. Extremely unrealistic datasets which cause SBC to fail, which would otherwise work well on real data. 
+    + Computationally expensive
 
-\section{Conclusion}
-
-\newpage
-
-% \bibliographystyle{apa} % Econometrica referencing
-\bibliography{references}
-
-\end{document}
+# Conclusion

### 2023-09-23 manuscript/motivating_example/single_sim_example.r
single simulation example

@@ -1,0 +1,81 @@
+library(tidyr)
+library(dplyr)
+library(cmdstanr)
+library(ggplot2)
+set.seed(630)
+
+n_time_points <- 1000
+
+gen_ncp_sv_dataset <- function(T){
+    sim_results <- list()
+    # Priors
+    mu <- rnorm(1, 0, sqrt(10))
+    sigma_sqd <- 1 / rgamma(1, 5/2, (0.01*5)/2)
+    sigma <- sqrt(sigma_sqd)
+    p <- rbeta(1, 20, 1.5)
+    phi <- 2*p - 1
+
+    # Sample from standardised normal dist and multiply by sigma
+    h_std <- rnorm(T, mean = 0, sd = sigma)
+
+    # Generate log volatilities
+    h <- rep(0, T)
+    h[1] <- rnorm(1, mean=mu, sd=sigma/sqrt(1-phi^2))
+    for(i in 2:T){
+        h[i] <- h_std[i] + mu + phi*(h[i-1] - mu)
+      }
+
+    # Generate data from prior
+    y_sim <- exp(h/2)*rnorm(T, 0, 1)
+    sim_results[["y_sim"]] <- y_sim
+    sim_results[["sigma_sqd"]] <- sigma_sqd
+    sim_results[["sigma"]] <- sigma
+    sim_results[["mu"]] <- mu
+    sim_results[["phi"]] <- phi
+    
+
+    return(sim_results)
+}
+
+data <- gen_ncp_sv_dataset(T = 1000)
+reuturns <- data$y_sim
+
+model_path <- "models/sv_user_guide_reparameterised_ksc_priors.stan"
+mod <- cmdstan_model(model_path, include_paths = here::here("models", "functions"))
+
+data_list <- list(T = length(returns), y = returns, sample_prior = 0)
+
+model_fit <- mod$sample(
+    data = data_list,
+    seed = 123,
+    chains = 4,
+    parallel_chains = 4,
+    adapt_delta = 0.999
+  )
+
+draws <- model_fit$draws(variables = c("mu", "phi", "sigma"), format = "df") %>% 
+    select(-.chain, -.iteration, -.draw) %>% 
+     pivot_longer(everything())
+
+draws$name <- replace(draws$name, draws$name=="sigma", "sigma_sqd")
+
+means = data.frame(value = rbind(data[['sigma_sqd']], data[['phi']], data[['mu']]),
+           name = rbind('sigma_sqd', 'phi', 'mu'))
+
[…]

### 2023-09-23 manuscript/motivating_example/single_sim_example.r
updated font

@@ -72,10 +72,11 @@ quantiles$name <- replace(quantiles$name, quantiles$name=="sigma", "sigma_sqd")
 
 ggplot(draws, aes(x = value)) +
   geom_histogram(aes(fill = name), alpha = 0.3) +
-  theme_minimal() +
+  theme_minimal(base_size = 18) +
   geom_vline(data = means, aes(xintercept=value, color=name), size = 1.5) +
   geom_vline(data = quantiles, aes(xintercept=value), linetype="dotted", size = 1) +
     facet_wrap(~name, scale = "free") + 
-    theme(legend.position="none")
+    theme(legend.position="none") +
+    labs(x = "Parameter value", y= "Count")
 
 ggsave("manuscript/motivating_example/single_sim.png", bg = "white")

### 2023-09-23 scripts/create_yahoo_data.r
remove config

@@ -1,12 +1,13 @@
 # Load data functions
 source(here::here("R", "data.R"))
-source(here::here("config.r"))
+# source(here::here("configs", "config.r"))
+config <- FALSE
 
 # (Download) and load data
 if (!(config)) {
     stock_index <- "GSPC"
-    start_date <- "2019-01-04"
-    end_date <- "2023-02-01"
+    start_date <- "2023-01-01"
+    end_date <- "2023-09-10"
 }
 
 raw_data_loc <- fin_data_location(stock_index, "raw", start_date, end_date)

### 2023-09-23 manuscript/motivating_example/gma_mcmc_gen.py
real data example

@@ -1,0 +1,98 @@
+#Multiple chains
+#Seeds + datasets
+from __future__ import division
+import math
+
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import pyarrow as pa
+import pyarrow.parquet as pq
+np.set_printoptions(suppress=True, precision=5)
+
+from src.gaussianMixture import *
+
+seed=1349
+np.random.seed(seed)
+
+# ex = pd.read_excel("data/raw/xrates.xls", engine = "xlrd")
+# dta = np.log(ex).diff().iloc[1:]
+# endog = (dta['USXUK'] - dta['USXUK'].mean()) * 100
+
+# df = pd.read_csv("data/preprocessed/GSPC/20190104_20230201.csv")
+df = pd.read_csv("data/preprocessed/GSPC/20230101_20230010.csv")
+dta = df[["log_return"]].iloc[1:]
+endog = (dta - dta.mean()) * 100
+
+# Setup the model and simulation smoother
+mod = TVLLDT(endog)
+mod.set_smoother_output(0, smoother_state=True) # Calculate and return the smoothed states. Other options like cov matrices available
+sim = mod.simulation_smoother()
+
+# Simulation parameters
+n_iterations = 11000
+burn = 10000
+
+seeds = [4324, 5435, 8574, 7675]
+
+results = list()
+
+for idx, val in enumerate(seeds):
+    np.random.seed(val)
+    # Storage for traces
+    trace_smoothed = np.zeros((n_iterations + 1, mod.nobs))
+    trace_states = np.zeros((n_iterations + 1, mod.nobs))
+    trace_mixing = np.zeros((n_iterations + 1, mod.nobs), int)
+    trace_mu = np.zeros((n_iterations + 1, 1))
+    trace_phi = np.zeros((n_iterations + 1, 1))
+    trace_sigma2 = np.zeros((n_iterations + 1, 1))
+    
+    # Initial values (p. 367)
+    trace_mixing[0] = 0 # Initialised all with the 4th gaussian. So all draws from a specific normal distribution. Not specified in the paper how this is done
+    trace_mu[0] = 0
+    trace_phi[0] = 0.95
+    trace_sigma2[0] = 0.02 #0.5 in the code. says 0.02 in the paper
+    
+    # Iterations
+    for s in range(1, n_iterations + 1):
+        # Update the parameters of the model
[…]

### 2023-09-23 manuscript/motivating_example/hmc_mcmc_gen.r

@@ -1,0 +1,38 @@
+library(dplyr)
+library(tidyr)
+library(cmdstanr)
+library(arrow)
+library(posterior)
+library(shinystan)
+
+set.seed(1349)
+
+# data_path <- here::here("manuscript/motivating_example/fin_data.parquet")
+# raw <- arrow::read_parquet(data_path)
+# first_diff <- diff(log(raw$USXUK))
+# returns <- (first_diff - mean(first_diff)) * 100
+
+# data_path <- here::here("data/preprocessed/GSPC/20190104_20230201.csv")
+data_path <- here::here("data/preprocessed/GSPC/20230101_20230910.csv")
+raw <- read.csv(data_path)
+log_return <- raw$log_return[complete.cases(raw$log_return)]
+returns <- (log_return - mean(log_return)) * 100
+
+model_path <- "models/sv_user_guide_reparameterised_ksc_priors.stan"
+mod <- cmdstan_model(model_path, include_paths = here::here("models", "functions"))
+
+data_list <- list(T = length(returns), y = returns, sample_prior = 0)
+
+model_fit <- mod$sample(
+    data = data_list,
+    seed = 123,
+    chains = 4,
+    parallel_chains = 4,
+    adapt_delta = 0.999
+  )
+
+draws <- model_fit$draws(variables = c("mu", "phi", "sigma"), format = "df") %>% 
+    select(-.chain, -.iteration, -.draw) %>% 
+     pivot_longer(everything())
+
+write.csv(draws, file = "manuscript/motivating_example/hmc_draws_2023.csv")

### 2023-09-23 manuscript/motivating_example/real_data_example.r

@@ -1,0 +1,43 @@
+library(dplyr)
+library(ggplot2)
+
+# Maybe consider latent states
+gma <- read.csv("manuscript/motivating_example/gma_draws_2023.csv")
+gma$model <- "gma"
+
+hmc <- read.csv("manuscript/motivating_example/hmc_draws_2023.csv")
+hmc$model <- "hmc"
+
+names(hmc) <- c("index", "name", "value", "model")
+names(gma) <- c("index", "name", "value", "chain", "model")
+
+hmc$name <- replace(hmc$name, hmc$name=="sigma", "sigma2")
+
+df <- rbind(gma %>% select(-chain), hmc)
+means <- df %>% group_by(model, name) %>% summarise(avg = mean(value))
+medians <- df %>% group_by(model, name) %>% summarise(med = median(value))
+
+ggplot(df, aes(x = value)) +
+    geom_histogram(position="identity", aes(fill = model), alpha =0.3) +
+    facet_wrap(~name, scale="free") +
+    # geom_vline(data =means, aes(xintercept=avg, colour=model), size = 1.2) +
+    theme_minimal(base_size = 18) +
+    theme(legend.position="none") +
+    labs(x = "Parameter value", y= "Count")
+
+    # geom_vline(data = medians, aes(xintercept=med, colour=model))
+
+df %>% group_by(model, name) %>% 
+  summarise(as_tibble(rbind(summary(value)))) %>% 
+  ungroup() %>% arrange(name)
+
+ggplot(df, aes(x = value)) +
+    geom_histogram(aes(fill = model), alpha =0.4) +
+    facet_grid(model~name, scale="free")
+
+
+dim(hmc)
+
+gma %>% filter(name=="mu") %>% dim
+hmc %>% filter(name=="mu") %>% dim
+

### 2023-09-24 manuscript/main.tex
Updated examples and added research context

--- 2023-09-20 manuscript/main.tex
@@
 
-\begin{abstract}.  
-    The objective of this research is to compare the computational methods used to estimate Bayesian stochastic volatility models.
-A simulation study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo and their ability to sample from the model's posterior distribution.
-Specifically, Simulation Based Calibration (SBC) is used to check whether these sampling strategies are returning efficient and well calibrated posterior estimates.
+\begin{abstract}
+    This research compares the computational methods used to estimate Bayesian stochastic volatility models.
+The estimation strategies outlined in the stochastic volatility literature and Hamiltoninan Monte Carlo are assesssed in their ability to sample from the model's posterior distribution.
+Specifically, Simulation Based Calibration (SBC) is used to check whether these MCMC algorithm are returning efficient and well calibrated posterior estimates.
 Key metrics of interest are the effective sample size to check the efficiency of the algorithm and tests of uniformity to assess the calibration of the posteriors.
-This will determine which algorithm is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
+This will determine which method is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
+Results reveal that Hamiltonian Monte Carlo provides more efficient and calibrated posterior estimates conditional on the non centered paramterisation of the stochastic volatility model.
 \end{abstract}
@@
 \section{Introduction}
+    The stochastic volatility model is used in financial econometrics to model the behaviour of financial instruments.
+It explicitly treats the variance as a latent random variable.
+These are typically expressed as non linear Gaussian state space models and are difficult to estimate using classical methods.
+The likelihood is unavailable in closed form and there are at least as many variables as data points.
+Bayesian approaches to estimating these models rely on  Markov Chain Monte Carlo (MCMC) techniques to sample from the target joint posterior of these high dimensional latent parameter spaces.
 
-- Should make reference to workflow, and reparameterisation of SV.
+    \citet{kim1998stochastic} propose a method for sampling this model using a combination of Kalman Filters, Gibbs sampler, and Metropolis Hastings over 20 years ago.
+Since then, there have been many advances in statistical computing to estimate complex, high dimensional Bayesian models.
+In particular, the recent availability of Hamiltonian Monte Carlo, a variant of the Metropolis Hastings algorithm, is an advancement of MCMC for efficient sampling of complex probabilistic models.
+The adoption of such new techniques have been made widely available through the development of various open source libraries such as the Stan programming language \citep{stan} and the PyMC library \citep{pymc2023}.
+Other algorithms have also improved the speed of estimating high dimensional models using approximations of the posterior such as integrated nested Laplace approximation \citep{rue2009approximate}.
 
-\section{Background and Literature Review}
+    These conceptually and practically different techniques to estimate complex models did not exist 20 years ago (or perhaps more accurately, were not easily accessible or implemented 20 years ago).
+As the development of new algorithms rapidly increase, so does the need to develop new methods to assess their output.
+Developments in our statistical workfow are required to test new algorithms as well as compare computational strategies used to estimate increasingly complex models.
+
+    This research conducts a simulation study to assess the algorithms used to estimate stochastic volatility models.
+The first algorithm replicates Kim Sherphard and Chib's (KSC) Gaussian mixture approximation which is applied on a transformed stochastic volatility model with a log chi squared error distribution.
+Their sampling method uses the aforementioned Kalman Filter estimating the latent states, conjugate posterior distributions and the Metropolis Hastings algorithm.
+The second algorithm is the Hamiltonian Monte Carlo (HMC) algorithm as implemented in the Stan programming language.
+The objective is to analyse the efficiency of the sampling approach and their ability to return calibrated posterior estimates.
+
+    Comparisons of these algorithms are conducted by Simulation Based Calibration \citep{talts2018validating}.
+Parameters are drawn from the prior distribution and used to create datasets from the generative (stochastic volatility) model which are sampled using a MCMC method.
+Repeating this process multiple times gives insight to how well the sampling approach can estimate the true parameters governing the data generating process on average.
+The key diagnostic metrics are the effective sample size to measure the efficiency of the algorithm and visual tests of uniform rank statistics to assess calibration of the posterior estimates.
+
+    Additionally, two different parametiersations of the stochastic volatility model are considered for each sampling method.
+These are described as either 'centred' or 'non-centred' in location.
+An algorithm may be sensitive to different parameterisations of the same model which may impact MCMC efficiency as described in \citet{strickland2008parameterisation}.
+The key contribution of this research is to use simulations to assess whether a MCMC algorithm can return calibrated posterior estimates as well as the efficiency under the different parametiersations.
+
+    This paper is structured as follows.
+Section 2 provides the context around this research, namely the stochastic volatility model examined and the limitations of single simulation studies and MCMC convergence diagnostics.
+Section 3 describes the simulation design and diagnostic metrics.
+Section 4 details the sampling approaches.
+Section 5 discusses the results and limitations.
+Finally, section 6 concludes and provides points for further reserach.
+
+\section{Resarch Context}
 
 \subsection{Stochastic Volatility}
-        Stochastic volatility models are used in financial econometrics to model the variance or volatility of returns in a financial instrument.
-The return of a financial asset often exhibits patterns in the variance which can be modelled to improve prediction and estimates of risk or tail events.
-Unlike other volatility models such as ARCH or GARCH, stochastic volatility models treat the variance as a random variable.
-Furthermore, these models have a state space representation since the variance is expressed as a latent stochastic process.
-State space models can be difficult to estimate since there is a latent state for every data observation in the time series, which means there are often more parameters than data points.
-These models can be estimated using frequent frameworks but this research will focus on Bayesian estimation strategies.
-
-        \citet{kim1998stochastic} estimate a discrete time, univariate stochastic volatility model using Bayesian methods which is outlined below.
+    The model of interest is the discrete time, univariate stochastic volatility model estimated by \citet{kim1998stochastic} using Bayesian methods.
 $y_t$ is the mean corrected returns of some asset for equally spaced intervals t.
@@
 
-    \subsection{Gaussian Mixture Approximation}
-        Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.} \citep{dejong1995}.
+\subsection{Diagnostic limitations on real data}
+    The stochastic volatility model can be estimated using a variety of sampling strategies and MCMC algorithms.
+Convergence metrics are often used to check the performance of the MCMC chains, such as the effective sample size and the $\hat{R}$.
+These diagnostics provide evidence on whether an MCMC has failed to converge onto the target posterior distribution.
 
-        The standard Kalman Filter and smoother is used to compute the posterior distribution over the latent states.
-This requires the state and measurement equations to be linear and conditionally Gaussian.
-Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring and taking the log of $y_t$.
+    Convergence diagnostics are useful for identifying when Bayesian computation fail on real data.
+However, confounding issues may arise when attempting to diagnose the cause of computational problems.
+Real data is generated from an unknown data generating process.
+That is, the true parameter or model is unobservable.
+Therefore, failed diagnostic checks could arise from either inocrrect model specification or issues in the sampling algorithm or both.
 
-        $$
-        \begin{aligned}
-        y_t^{*} &= log(y_t^2) \\ 
-        &= log((\epsilon_t exp(h_t/2))^2) \\
-        &=  log(exp(h_t)) + log(\epsilon_t^2) \\
-        &= h_t + log(\epsilon_t^2)  \\
-        &= h_t + z_t \\
-        \end{aligned}
-        $$
+    Furthermore, different sampling strategies and MCMC tools could provide different posterior estimates for the same model.
+Assuming no issues in computation, these diagnostic measures do not provide any evidence to which estimate is closer to the truth.
 
-        Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
-The relationship between $y_t$ and $h_t$ is now linear; however, the error is not Gaussian.
-Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution.
-This is defined by:
+    To illustrate this problem, the stochastic volatility model is estimated with two different MCMC algorithms on real data.
+The data are the de-meaned log returns of the 2023 January-September S\&P 500 Index.
+Below are the marginal posterior estimates for the static parameters $\mu$, $\phi$ and $\sigma^2$ and corresponding summary statistics.
 
-        $$
-        \begin{aligned}
-        f(z_t) = \sum_{i=1}^{K} q_if_N(z_i|m_i-1.2704, \nu_i^2)
-        \end{aligned}
-        $$
+    \begin{figure}[h]
+        \centering
+        \includegraphics[scale=0.1]{motivating_example/real_data_ex.png}
+        \caption{Posterior samples from two MCMC algorithms using S\&P 500 data}
+    \end{figure}
 
-        Where K is the mixture of normal densities $f_N$, component probabilities $q_i$, mean $m_i-1.2704$ and variance $\nu_i^2$.
-These parameters were selected using moment matching where they found 7 normal densities with varying mean and variance parameters best approximated the log chi squared moments.
-These parameters and weights can be found in Appendix A.
+    These estimates show the marginal posteriors of each parameter sharing similar shapes.
+$\phi$  has a fatter left tail for and $\sigma^2$ has heavier right tails for sampling method 1 (SM1).
+Furthermore, SM2 has higher peaks at their mode relative to SM1 with a tighter spread.
+Based on this information alone, there is no certainty in whether SM1 or SM2 is has a more correct estimate conditional on data and model.
+There are other methods for model or algorithmic selection in this context (for example, out of sample predictive performance), but for parameter estimation it is unclear which one should be selected.
 
-        The model can be sampled via the Kalman Filter and simulation smoother since the model is now linear and conditionally gaussian.
-The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions whereas $\phi$ is sampled via a Metropolis Hastings accept/reject procedure.
-The details can be around in Appendix B.
+\subsection{Diagnostic limitations on a single simulation}
+    Diagnosing computational problems on real data is difficult due to confounding factors.
+A strategy around this is to evaluate a model and algorithm on simulated data.
+One approach is to simulate data from the proposed model using known parameters.
+Then, fit the model on the simulated data and see if the model can recover the true parameters.
+This gives us the benefit of defining the true parameters of the data generating process to be estimated.
+If the model and algorithm cannot adequately capture the true parameter, then we cannot be confident that it will provide reliable estimates on real data.
 
-    \subsection{Hamiltonian Monte Carlo}
-        Since the paper was written, new MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
-Specifically, Hamiltonian Monte Carlo is a MCMC algorithm which has become widely available for efficiently sampling from sophisticated models.
-Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature \citep{duane1987hybrid} before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks \citep{neal1995bayesian} and statistical computing \citep{neal2011mcmc}.
-The algorithm has since become widely available through open source development projects such as Stan \citep{stan} and PyMC \citep{pymc2023}.
+    Furthermore, as discussed in \citet{gelman2020bayesian}, fitting models on simulated data is the only way we can check inference on latent variables.
+This is critical in Stochastic Volatility since the underlying framework is a state space model with latent log volatility parameters.
+Latent variables are unobserved in real data and are only estimated in the context of the model.
+Simulation gives control over the data generating process which reveals what the model can infer about the latent variables.
 
-        The key innovation of Hamiltonian Monte Carlo is using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
-Unlike Random Walk Metropolis Hastings, it takes advantage of the geometry of the posterior to determine its next proposal step.
-A comprehensive explanation of the sampler is beyond the scope of this research and can be found in the above references.
+    Simulations enable us to check whether our models and algorithms can estimate the true data generating process, however, there are limitations to what can be learned from a single simulation.
+There is always a small probability that the true parameter is in the tails of any estimated posterior distribution.
+Or put differently, there is a 1\% chance the true parameter or any random draw exists outside a 99\% credible interval.
+\cite{talts2018validating} make the point that a single simulation does not provide sufficient information about the inference made by an algorithm.
+As discussed in their paper, a single simulation may conclude "that an incorrectly coded analysis worked as desired, while a correctly coded analysis failed".
 
-        The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
-Stan's default algorithm, the No-U-Turn Sampler \citep{hoffman2014no}, allows for direct sampling of the specified stochastic volatility model.
-Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
-This approach will also use the same priors as specified in the Gaussian mixture approximation.
+    The below results are from a single simulation where data is generated and sampled from the stochastic volatility model with known parameters.
+
+    \begin{figure}[h]
+        \centering
+        \includegraphics[scale=0.1]{motivating_example/single_sim.png}
+        \caption{Posterior samples from known data genearting process. Vertical solid line represents true parameter and dotted lines are 95\% credible intervals around the mean}
+    \end{figure}
+
+    The 95\% credible intervals for $\sigma$ and $\phi$ cover the true parameter.
+The true parameter for $\mu$ however, is deep in the tails and outside the interval.
+Such an analysis may incorrectly conclude that the model fails to adequatly estimate the $\mu$ parameter.
+However, it may be the case that the posterior distribution is correctly calculated using this algorithm and the results may be due to the features of this specific simulated dataset.
+
+\subsection{Research Goal}
+    The objective of this research is to design a simulation study to evaluate the calibration of algorithms used to sample Stochastic Volatility models.
+Indeed, there are limitations to evaluating MCMC algorithms based on fits to real data and single simulations.
+To check the calibration of an algorithm, simulations over multiple "true parameters" are required.
+This methodology is discussed in the next section.
+
+    Additionally, two sampling strategies will be assessed using the proposed simulation design as well as the performance of the samplers under different parameterisations of the stochastic volatility model.
+Not only will the calibration of any one algorithm be assessed, but also compared across sampling strategies to determine which MCMC approach is most suitable for estimating this model.
 
@@
     \subsection{Simulation Design}
-        Simulation Based Calibration (SBC) is a technique that checks the calibration of posterior estimates generated by MCMC algorithms.
+        Simulation Based Calibration (SBC) checks the calibration of posterior estimates generated by MCMC algorithms.
 SBC is conducted by comparing the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.
@@
 
-    \subsection{Metrics}
+    \subsection{Evaluation Metrics}
         The key metrics and diagnostics to compare the performance of these methods are the effective sample size (ESS), rank statistics, and chi-squared test statistics.
 
+        \subsubsection{Effective sample size}
             ESS measures the efficiency of the MCMC sampler.
@@
 An efficient MCMC algorithm takes less resources (for example, time and number of draws) to get a representative sample of the target distribution.
-If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), then we may conclude that this method is a more efficient sampler (conditional on the model).
+If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), then there is evidence that this method is a more efficient sampler.
 
+        \subsubsection{Rank statistics and chi squared statistics}
             Rank statistics as described in the SBC section are used to evaluate the calibration of a posterior.
@@
 
-        Chi squared statistics can be used to test for uniformity and thus calibration of all the parameters in the model.
+            A perfectly discrete uniform distribution will return a $\chi^2$ statistic of zero.
+That is, the number of rank statistics in each bin is equal to the expected number conditional on the number of bins and observations.
+Instead of performing multiple hypothesis tests for each parameter, the distribution of $\chi^2$ statistics is compared across all parameters and simulations.
+This will give a high level summary of calibration and overall performance of the algorithms.
+
+\section{MCMC Strategies}
+
+    \subsection{Sampling method 1}
+        (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.} \citep{dejong1995}.
+
+        The standard Kalman Filter and smoother is used to compute the posterior distribution over the latent states.
+This requires the state and measurement equations to be linear and conditionally Gaussian.
+Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring and taking the log of $y_t$.
+
+        $$
+        \begin{aligned}
+        y_t^{*} &= log(y_t^2) \\ 
+        &= log((\epsilon_t exp(h_t/2))^2) \\
+        &=  log(exp(h_t)) + log(\epsilon_t^2) \\
+        &= h_t + log(\epsilon_t^2)  \\
+        &= h_t + z_t \\
+        \end{aligned}
+        $$
+
+        Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
+The relationship between $y_t$ and $h_t$ is now linear; however, the error is not Gaussian.
+Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution.
+This is defined by:
+
+        $$
+        \begin{aligned}
+        f(z_t) = \sum_{i=1}^{K} q_if_N(z_i|m_i-1.2704, \nu_i^2)
+        \end{aligned}
+        $$
+
+        Where K is the mixture of normal densities $f_N$, component probabilities $q_i$, mean $m_i-1.2704$ and variance $\nu_i^2$.
+These parameters were selected using moment matching where they found 7 normal densities with varying mean and variance parameters best approximated the log chi squared moments.
+These parameters and weights can be found in Appendix A.
+
+        The model can be sampled via the Kalman Filter and simulation smoother since the model is now linear and conditionally gaussian.
+The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions whereas $\phi$ is sampled via a Metropolis Hastings accept/reject procedure.
+The details can be around in Appendix B.
+
+    \subsection{Sampling method 2}
+    % Need much more detail here about how it works
+        Since the paper was written, new MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
+Specifically, Hamiltonian Monte Carlo is a MCMC algorithm which has become widely available for efficiently sampling from sophisticated models.
+Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature \citep{duane1987hybrid} before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks \citep{neal1995bayesian} and statistical computing \citep{neal2011mcmc}.
+The algorithm has since become widely available through open source development projects such as Stan \citep{stan} and PyMC \citep{pymc2023}.
+
+        The key innovation of Hamiltonian Monte Carlo is using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
+Unlike Random Walk Metropolis Hastings, it takes advantage of the geometry of the posterior to determine its next proposal step.
+A comprehensive explanation of the sampler is beyond the scope of this research and can be found in the above references.
+
+        The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
+Stan's default algorithm, the No-U-Turn Sampler \citep{hoffman2014no}, allows for direct sampling of the specified stochastic volatility model.
+Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
+This approach will also use the same priors as specified in the Gaussian mixture approximation.
 
@@
 An example of this is Neal's funnel \citep{neal2003slice} where the Hamiltonian Monte Carlo sampler encounters performance issues in hierarchical models and produces biased samples.
+Efficiency improvements of MCMC algorithms for state space models under different parameterisations are also explored \citet{strickland2008parameterisation}.
 
@@
 A model can be reparameterised such that the states are sampled on a distribution centered on 0 (i.e "non centered") which are later transformed to have the correct mean.
-It is useful to compare model parameterisations since sampling strategies may be sensitive to the posterior geometry.
+The following re-parameterisations are explored as part of the simulation study.
 
@@
         First sample from a standard normal distribuition and multiply by the variance of the log volatility.
+The state variable is sampled with mean centered on zero and variance equal to $\sigma_{\eta}^2$
 
@@
 
-        The states are sampled with the mean centered at zero. Then apply the appropriate rescaling to get samples from the log volatility.
+        Then apply the appropriate rescaling to get samples from log volatility.
 
@@
         The non centered model for the Gaussian approximation is expressed slightly differently due to the use of the Kalman Filter.
-While HMC can sample from the joint posterior directly with state vectors centered at zero, applying the Kalman Filter requires the state equation to be a function of the state variable.
+Non centered HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates.
+Applying the Kalman Filter requires the state equation to be a function of the state variable with the average log volatility parameter $\mu$ entering the measurement equation.
 
@@
 
-        Expand the state equation:
+        Let $g$ be the demeaned state variable. Rewrite the demeaned state equation $g_{t+1}$ to be non centered in location by subtracting average volatility $\mu$.
 
@@
         \begin{aligned}
-        h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
-        =& \space \mu + \phi h_t - \phi\mu + \sigma_{\eta} \eta_t  \\
-        =& \space \mu (1- \phi) + \phi h_t + \sigma_{\eta} \eta_t  \\
+        g_t =& h_t - \mu \\
+        g_{t+1} =& \phi g_t + \sigma_{\eta}\eta_{t}
         \end{aligned}
@@
 
-        Take out the intercept and put it in the measurement equation:
+        Return average volatility into measurement equation and rewrite as a function of the non centered state equation:
 
@@
         \begin{aligned}
-        y^{\ast}_t =& h_t + \mu (1- \phi) + z_t \\
-        h^{\ast}_{t+1} =& \space \phi h_t + \sigma_{\eta} \eta_t  \\
+        y^{\ast}_t =& g_t + \mu + z_t \\
+        g_{t+1} =& \phi g_t + \sigma_{\eta}\eta_{t}
         \end{aligned}
@@
         It is worth noting that the centered and non-centered parameteristions are the same stochastic volatility model.
-Different parameteristions express the same mathematical models in ways that may make it easier or harder for a given algorithm to sample.
+Different parameteristions express the same mathematical models in ways that make it easier or harder for a given algorithm to sample.
 
@@
 
-        % \begin{figure}
-        %     \centering
-        %     \includegraphics[scale=0.1]{ppc_h2.png}
-        %     \caption{Prior predictive samples of second state variable}
-        % \end{figure}
+        \begin{figure}[h]
+            \centering
+            \includegraphics[scale=0.1]{ppc_h2.png}
+            \caption{Prior predictive samples of second state variable}
+        \end{figure}
 

### 2023-09-24 manuscript/structure/draft.qmd
Updated examples and added research context

@@
 # Abstract
+This research compares the computational methods used to estimate Bayesian stochastic volatility models.
+The estimation strategies outlined in the stochastic volatility literature and Hamiltoninan Monte Carlo are assesssed in their ability to sample from the model's posterior distribution.
+Specifically, Simulation Based Calibration (SBC) is used to check whether these MCMC algorithm are returning efficient and well calibrated posterior estimates.
+Key metrics of interest are the effective sample size to check the efficiency of the algorithm and tests of uniformity to assess the calibration of the posteriors.
+This will determine which method is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
+Results reveal that Hamiltonian Monte Carlo provides more efficient and calibrated posterior estimates conditional on the non centered paramterisation of the stochastic volatility model.
+
+<!-- 
 The objective of this research is to compare the computational methods used to estimate Bayesian stochastic volatility models.
@@
 This will determine which algorithm is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
-
+ -->
 # Introduction
-Stochastic volatility is a type of volatility model which explicitly treats the variance as a latent random variable.
+The stochastic volatility model is used in financial econometrics to model the behaviour of financial instruments.
+It explicitly treats the variance as a latent random variable.
 These are typically expressed as non linear Gaussian state space models and are difficult to estimate using classical methods.
 The likelihood is unavailable in closed form and there are at least as many variables as data points.
-Bayesian approaches to estimating these models rely on computational Markov Chain Monte Carlo (MCMC) techniques to sample from the target joint posterior of these high dimensional latent parameter spaces.
+Bayesian approaches to estimating these models rely on  Markov Chain Monte Carlo (MCMC) techniques to sample from the target joint posterior of these high dimensional latent parameter spaces.
 
-Kim, Shephard and Chib (1998) published a method for sampling this model using a combination of Kalman Filters, Gibbs, and Metropolis Hastings over 20 years ago.
-Since then, there have been many advances in Bayesian, and more broadly, statistical computation to estimate complex, high dimensional state space models.
-In particular, the recent availability of Hamiltonian Monte Carlo, a variant of the Metropolis Hastings algorithm, is a advancement of MCMC for efficient sampling of complex probabilistic models.
+Kim, Shephard and Chib (1998) propose a method for sampling this model using a combination of Kalman Filters, Gibbs sampler, and Metropolis Hastings over 20 years ago.
+Since then, there have been many advances in statistical computing to estimate complex, high dimensional Bayesian models.
+In particular, the recent availability of Hamiltonian Monte Carlo, a variant of the Metropolis Hastings algorithm, is an advancement of MCMC for efficient sampling of complex probabilistic models.
 The adoption of such new techniques have been made widely available through the delveopment of various open source libraries such as the Stan programming language and the PyMC library.
-Other algorithms have also improved the speed of estimating high dimensional models using approximations of the posterior such as Integrated nested Laplace approximation and Approximate Bayesian Computation.
+Other algorithms have also improved the speed of estimating high dimensional models using approximations of the posterior such as integrated nested Laplace approximation (INLA).
 
@@
 As the development of new algorithms rapidly increase, so does the need to develop new methods to assess their output.
-Developments in our statistical workfow are required to assess new sampling algorithms as well as compare computational strategies used to estimate increasingly complex models.
+Developments in our statistical workfow are required to test new algorithms as well as compare computational strategies used to estimate increasingly complex models.
 
-This research implements a simulation study to assess the algorithms used to estimate stochastic volatility models.
-Simulation Based Calibration outlined in Talts et al.
-is the primary study design to determine whether the MCMC strategy are returning efficient and well calibrated posterior estimates.
+This research conducts a simulation study to assess the algorithms used to estimate stochastic volatility models.
+The first algorithm replicates Kim Sherphard and Chib's (KSC) Gaussian mixture approximation which is applied on a transformed stochastic volatility model with a log chi squared error distribution.
+Their sampling method uses the aforementioned Kalman Filter estimating the latent states, conjugate posterior distributions and the Metropolis Hastings algorithm.
+The second algorithm is the Hamiltonian Monte Carlo (HMC) algorithm as implemented in the Stan programming language.
+The objective is to analyse the efficiency of the sampling approach and their ability to return calibrated posterior estimates.
 
-Two sampling approaches are considered in this study - Kalman Filter and Metropolis Hastings estimation by Kim Sherphard and Chib, and the Hamiltoninan Monte Carlo algorithim.
+Comparisons of these algorithms are conducted by Simulation Based Calibration (Talts et al.).
+Parameters are drawn from the prior distribution and used to create datasets from the generative (stochastic volatility) model which are sampled using a MCMC method.
+Repeating this process multiple times gives insight to how well the sampling approach can estimate the true parameters governing the data generating process on average.
+The key diagnostic metrics are the effective sample size to measure the efficiency of the algorithm and visual tests of uniform rank statistics to assess calibration of the posterior estimates.
 
-Key metrics used to evaluate the model are the effective sample size and visual tests of uniform rank statistics.
+Additionally, two different parametiersations of the stochastic volatility model are considered for each sampling method.
+These are described as either 'centred' or 'non-centred' in location.
+An algorithm may be sensitive to different parameterisations of the same model which may impact MCMC efficiency as described in \citet{strickland2008parameterisation}.
+The key contribution of this research is to use simulations to assess whether a MCMC algorithm can return calibrated posterior estimates as well as the efficiency under the different parametiersations.
+
+This paper is structured as follows.
+Section 2 provides the context around this research, namely the stochastic volatility model examined and the limitations of single simulation studies and MCMC convergence diagnostics.
+Section 3 describes the simulation design and diagnostic metrics.
+Section 4 details the sampling approaches.
+Section 5 discusses the results and limitations.
+Finally, section 6 concludes and provides points for further reserach.
+
+ <!-- is the primary study design to determine whether the MCMC strategy are returning efficient and well calibrated posterior estimates.
+
+Two sampling approaches are considered in this study - Kalman Filter and Metropolis Hastings estimation by Kim Sherphard and Chib, and the Hamiltoninan Monte Carlo algorithim.  -->
+
+<!-- HMC is able to able to directly sample from the specified stochastic volatility model without any approximations of the error term.
+Its key innovation is uses the gradients of the target joint posterior distribution to generate efficient proposals for the sampler to explore.
+-->
 
@@
 # Stochastic Volatility
-The model of interest in this research is the discrete time, univariate stochastic volatility model outlined by \citet{kim1998stochastic} who use Bayesian estimation methods.
+The model of interest is the discrete time, univariate stochastic volatility model estimated by \citet{kim1998stochastic} using Bayesian methods.
 $y_t$ is the mean corrected returns of some asset for equally spaced intervals t.
@@
 The stochastic volatility model can be estimated using a variety of sampling strategies and MCMC algorithms.
-A common set of metrics are used to check the convergence of the MCMC.
-This includes checking the effective sample size and the $\hat{R}$ calculations for estimated chains and parameters.
-These diagnostics check whether an MCMC has failed to converge onto the target posterior distribution (since we can never know with certainty that a sampler has converged).
+Convergence metrics are often used to check the performance of the MCMC chains, such as the effective sample size and the $\hat{R}$.
+These diagnostics provide evidence on whether an MCMC has failed to converge onto the target posterior distribution.
 
 Convergence diagnostics are useful for identifying when Bayesian computation fail on real data.
-However, confounding issues may arise when trying to diagnose the _cause_ of computational problems.
+However, confounding issues may arise when attempting to diagnose the cause of computational problems.
 Real data is generated from an unknown data generating process.
 That is, the true parameter or model is unobservable.
-Therefore, failed diagnostic checks could arise from either inocrrect model specification or issues in the sampling strategy (algorithm) or both.
+Therefore, failed diagnostic checks could arise from either inocrrect model specification or issues in the sampling algorithm or both.
 
-Furthermore, different sampling strategies and MCMC tools could provide different estimates for the same model.
+Furthermore, different sampling strategies and MCMC tools could provide different posterior estimates for the same model.
 Assuming no issues in computation, these diagnostic measures do not provide any evidence to which estimate is closer to the truth.
 
-To illustrate this problem, suppose we estimate this model with two different sampling strategies on real data.
-The data are the demaned log returns of the 2019-2023 S&P 500 Index.
+To illustrate this problem, suppose we estimate this model with two different MCMC algorithms on real data.
+The data are the de-meaned log returns of the 2023 January-September S&P 500 Index.
 Below are the marginal posterior estimates for the static parameters $\mu$, $\phi$ and $\sigma^2$ and corresponding summary statistics.
@@
 
-These particular estimates show the marginal posteriors of each parameter sharing similar shapes.
-$\phi$ and $\sigma^2$ possess slightly heavier right tails for SM1 compared to SM2.
-While in this particular example there is no major difference in the estimates, there is no certainty in whether SM1 or SM2 is consistenty closer to the truth.
-There are other ways of model or algorithmic selection in this context (for example, out of sample predictive performance), but for parameter _estimation_ it is unclear which one should be selected.
+These estimates show the marginal posteriors of each parameter sharing similar shapes.
+$\phi$  has a fatter left tail for and $\sigma^2$ has heavier right tails for sampling method 1 (SM1).
+Furthermore, SM2 has higher peaks at their mode relative to SM1 with a tighter spread.
+Based on this information alone, there is no certainty in whether SM1 or SM2 is has a more correct estimate conditional on data and model.
+There are other methods for model or algorithmic selection in this context (for example, out of sample predictive performance), but for parameter estimation it is unclear which one should be selected.
 
@@
 Diagnosing computational problems on real data is difficult due to confounding factors.
-A strategy around this is to evaluate a model and sampling strategy on simulated data.
+A strategy around this is to evaluate a model and algorithm on simulated data.
 One approach is to simulate data from the proposed model using known parameters.
 Then, fit the model on the simulated data and see if the model can recover the true parameters.
-This gives us the benefit of knowing or defining the data generating process to be estimated.
+This gives us the benefit of defining the true parameters of the data generating process to be estimated.
 If the model and algorithm cannot adequately capture the true parameter, then we cannot be confident that it will provide reliable estimates on real data.
 
-Furthermore, as discussed in \cite{Bayesian workflow}, fitting models on simulated data is the only way we can check inference on latent variables.
+Furthermore, as discussed in \citet{gelman2020bayesian}, fitting models on simulated data is the only way we can check inference on latent variables.
 This is critical in Stochastic Volatility since the underlying framework is a state space model with latent log volatility parameters.
 Latent variables are unobserved in real data and are only estimated in the context of the model.
-Simulation gives control over the data generation which governs the latent process and reveals what the model can infer about the latent variables.
+Simulation gives control over the data generating process which reveals what the model can infer about the latent variables.
 
-While simulations enable us to check whether our models and algorithms can capture the true generating process, there are limitations to what can be learned from a single simulation.
+Simulations enable us to check whether our models and algorithms can estimate the true data generating process, however, there are limitations to what can be learned from a single simulation.
 There is always a small probability that the true parameter is in the tails of any estimated posterior distribution.
@@
 
-The posteriors for x and y sufficiently cover the true parameter, however, the true parameter for z is deep in the tails.
-Such an analysis may incorrectly conclude that the model fails to adequatly estimate the z parameter.
-However, it may be the case that the posterior distribution is correctly calculated using this algorithm and the results may be due to the nuances of this specific simulated dataset.
+The 95\% credible intervals for $\sigma$ and $\phi$ cover the true parameter.
+The true parameter for $\mu$ however, is deep in the tails and outside the interval.
+Such an analysis may incorrectly conclude that the model fails to adequatly estimate the $\mu$ parameter.
+However, it may be the case that the posterior distribution is correctly calculated using this algorithm and the results may be due to the features of this specific simulated dataset.
 
@@
 To check the calibration of an algorithm, simulations over multiple "true parameters" are required.
-This methodology is discussed in the next session.
+This methodology is discussed in the next section.
 
-Additionally, two sampling strategies will be assessed using the proposed simulation strategy as well as the performance of the samplers under different parameterisations of the stochastic volatility model.
+Additionally, two sampling strategies will be assessed using the proposed simulation design as well as the performance of the samplers under different parameterisations of the stochastic volatility model.
 Not only will the calibration of any one algorithm be assessed, but also compared across sampling strategies to determine which MCMC approach is most suitable for estimating this model.
@@
 
-## Metrics
+## Evaluation Metrics
 The key metrics and diagnostics to compare the performance of these methods are the effective sample size (ESS), rank statistics, and chi-squared test statistics.
 
+### Effective sample size
 ESS measures the efficiency of the MCMC sampler.
@@
 An efficient MCMC algorithm takes less resources (for example, time and number of draws) to get a representative sample of the target distribution.
-If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), then we may conclude that this method is a more efficient sampler (conditional on the model).
+If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), then there is evidence that this method is a more efficient sampler.
 
+### Rank statistics and chi squared statistics
 Rank statistics as described in the SBC section are used to evaluate the calibration of a posterior.
@@
 
-Chi squared statistics can be used to test for uniformity and thus calibration of all the parameters in the model.
+A perfectly discrete uniform distribution will return a $\chi^2$ statistic of zero.
+That is, the number of rank statistics in each bin is equal to the expected number conditional on the number of bins and observations.
+Instead of performing multiple hypothesis tests for each parameter, the distribution of $\chi^2$ statistics is compared across all parameters and simulations.
+This will give a high level summary of calibration and overall performance of the algorithms.
+
+<!-- % Chi squared statistics can be used to test for uniformity and thus calibration of all the parameters in the model. -->
 
@@
 ## Sampling method 1
-Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.} \citep{dejong1995}.
+KSC sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.} \citep{dejong1995}.
 
@@
 
-## Model Parameterisation
+## Model reparameterisation
 The parameterisation of a model can affect the performance of a MCMC algorithm when sampling from models with complex posterior geometries.
 An example of this is Neal's funnel \citep{neal2003slice} where the Hamiltonian Monte Carlo sampler encounters performance issues in hierarchical models and produces biased samples.
+Efficiency improvements of MCMC algorithms for state space models under different parameterisations are also explored in Forbes et al.
 
@@
 A model can be reparameterised such that the states are sampled on a distribution centered on 0 (i.e "non centered") which are later transformed to have the correct mean.
-It is useful to compare model parameterisations since sampling strategies may be sensitive to the posterior geometry.
+The following re-parameterisations are explored as part of the simulation study.
 
 ### Non centered Hamiltonian Monte Carlo
-First sample from a standard normal distribuition and multiply by the variance of the log volatility.
+First sample from a standard normal distribuition and multiply by the variance of the log volatility. The state variable is sampled with mean centered on zero and variance equal to $\sigma_{\eta}^2$
 
@@
 
-The states are sampled with the mean centered at zero. Then apply the appropriate rescaling to get samples from the log volatility.
+Then apply the appropriate rescaling to get samples from log volatility.
 
@@
 The non centered model for the Gaussian approximation is expressed slightly differently due to the use of the Kalman Filter.
-While HMC can sample from the joint posterior directly with state vectors centered at zero, applying the Kalman Filter requires the state equation to be a function of the state variable.
+Non centered HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates.
+Applying the Kalman Filter requires the state equation to be a function of the state variable with the average log volatility parameter $\mu$ entering the measurement equation.
 
@@
 
-Expand the state equation:
+Let $g$ be the demeaned state variable. Rewrite the demeaned state equation $g_{t+1}$ to be non centered in location by subtracting average volatility $\mu$.
 
@@
 \begin{aligned}
-        h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
-        =& \space \mu + \phi h_t - \phi\mu + \sigma_{\eta} \eta_t  \\
-        =& \space \mu (1- \phi) + \phi h_t + \sigma_{\eta} \eta_t  \\
+g_t =& h_t - \mu \\
+g_{t+1} =& \phi g_t + \sigma_{\eta}\eta_{t}
 \end{aligned}
@@
 
-Take out the intercept and put it in the measurement equation:
+Return average volatility into measurement equation and rewrite as a function of the non centered state equation:
 
@@
 \begin{aligned}
-        y^{\ast}_t =& h_t + \mu (1- \phi) + z_t \\
-        h^{\ast}_{t+1} =& \space \phi h_t + \sigma_{\eta} \eta_t  \\
+y^{\ast}_t =& g_t + \mu + z_t \\
+g_{t+1} =& \phi g_t + \sigma_{\eta}\eta_{t}
 \end{aligned}
@@
 It is worth noting that the centered and non-centered parameteristions are the same stochastic volatility model.
-Different parameteristions express the same mathematical models in ways that may make it easier or harder for a given algorithm to sample.
+Different parameteristions express the same mathematical models in ways that make it easier or harder for a given algorithm to sample.
 
@@
 # Conclusion
+<!--
+
+% - SV models and more broadly state space models are complex, high dimensional models which require computational tools to estimate them.
+% - Modelling variance useful and important?
+
+% - Advanced in Bayesian computation
+%     + Conceptually and practically diff techniques to estimate complex models didn't exist not that long ago
+%     + Performance metrics useful for checking fit on single dataset. However not much emphasis on checking performance between MCMC
+
+% - A principled Bayesian workflow which usually requires checking of multiple componenets: prior, posterior, model choice. 
+%     + Lots of different components which could contribute to an incorrect fit or error.
+
+% - This research focuses on evaluating the computational tools used to esitmate a specific type of model.
+
+% % One para
+% - This research </contributes> examine the algos used to estimate SV model 
+%     + Particular focus is on the computational tools used to estimate these models and whether they are returning what we expect.
+%     + conditional on the same prior
+
+%     - Additionally, check the sensitivity of the simulation and algortihms to the parameterisation of the model.
+Taking idea and advice from Catherine's paper about improvements in sampling performance.
+
+% %  2nd 
+% - Investigating two algorithms, HMC and GMA from KSC.
+
+% % 3
+% - Two samplings methods are used and are compared through simulation
+%     + Specifically, the simulation designed which is taken from talts et al.
+
+% % 4
+% - Key metrics of interests are ESS for efficiency and tests of uniformity for calibration.
+
+% - Goal is to show which algo and parameterisation is "better".
+Moreover, want to demonstarte that this part of a Bayesian workflow which can be applied more generally is important for checking that the model is returning the correct posteriros - and if not, where the problem may be occurring and why.
+Give a set of tools to evaluate algos and parametiersations (before real data is even considered).
+**Why is this important.**
+
+% - Sections outlined below: -->

### 2023-09-24 manuscript/motivating_example/gma_mcmc_gen.py

@@ -20,7 +20,7 @@ np.random.seed(seed)
 # endog = (dta['USXUK'] - dta['USXUK'].mean()) * 100
 
 # df = pd.read_csv("data/preprocessed/GSPC/20190104_20230201.csv")
-df = pd.read_csv("data/preprocessed/GSPC/20230101_20230010.csv")
+df = pd.read_csv("data/preprocessed/GSPC/20230101_20230910.csv")
 dta = df[["log_return"]].iloc[1:]
 endog = (dta - dta.mean()) * 100
 

### 2023-09-24 manuscript/motivating_example/real_data_example.r

@@ -25,6 +25,8 @@ ggplot(df, aes(x = value)) +
     theme(legend.position="none") +
     labs(x = "Parameter value", y= "Count")
 
+ggsave("manuscript/motivating_example/real_data_ex.png", bg = "white")
+    
     # geom_vline(data = medians, aes(xintercept=med, colour=model))
 
 df %>% group_by(model, name) %>% 

### 2023-09-24 manuscript/main.tex
updated location
fixed references

@@
 \usepackage[colorlinks = true,
-			citecolor  = blue]{hyperref}
+			citecolor  = blue,
+            linkcolor = black]{hyperref}
 
-\usepackage[longnamesfirst,authoryear, round]{natbib}
-\bibliographystyle{unsrtnat}
+\usepackage[natbibapa]{apacite}
+\bibliographystyle{apacite}
+% \usepackage[longnamesfirst,authoryear, round]{natbib}
+% \bibliographystyle{unsrtnat}
 
@@
 
-        Generating 1000 samples of $h_2^{sim}$ from both centered and non-centered models gives samples from the same data generating process as shown in figure x.
+        Generating 1000 samples of $h_2^{sim}$ from both centered and non-centered models gives samples from the same data generating process as shown in figure 3.
 
@@
             \centering
-            \includegraphics[scale=0.1]{ppc_h2.png}
+            \includegraphics[scale=0.1]{figures/ppc_h2.png}
             \caption{Prior predictive samples of second state variable}
@@
 
-% \bibliographystyle{apa} % Econometrica referencing
 \bibliography{references}

### 2023-09-25 manuscript/main.tex
updated plots

@@
             linkcolor = black]{hyperref}
+\usepackage[font={small,it}]{caption}
+\usepackage[labelfont=bf]{caption}
 
@@
 
-    The 95\% credible intervals for $\sigma$ and $\phi$ cover the true parameter.
-The true parameter for $\mu$ however, is deep in the tails and outside the interval.
-Such an analysis may incorrectly conclude that the model fails to adequatly estimate the $\mu$ parameter.
+    The 95\% credible intervals for $\mu$ and $\sigma^2$ cover the true parameter.
+The true parameter for $\phi$ however, is in the tails and outside the interval.
+Such an analysis may incorrectly conclude that the model fails to adequatly estimate the $\phi$ parameter.
 However, it may be the case that the posterior distribution is correctly calculated using this algorithm and the results may be due to the features of this specific simulated dataset.

### 2023-09-25 manuscript/motivating_example/real_data_example.r

@@ -11,7 +11,13 @@ hmc$model <- "hmc"
 names(hmc) <- c("index", "name", "value", "model")
 names(gma) <- c("index", "name", "value", "chain", "model")
 
-hmc$name <- replace(hmc$name, hmc$name=="sigma", "sigma2")
+hmc$name <- replace(hmc$name, hmc$name=="sigma", "~sigma^2")
+hmc$name <- replace(hmc$name, hmc$name=="mu", "~mu")
+hmc$name <- replace(hmc$name, hmc$name=="phi", "~phi")
+
+gma$name <- replace(gma$name, gma$name=="sigma2", "~sigma^2")
+gma$name <- replace(gma$name, gma$name=="mu", "~mu")
+gma$name <- replace(gma$name, gma$name=="phi", "~phi")
 
 df <- rbind(gma %>% select(-chain), hmc)
 means <- df %>% group_by(model, name) %>% summarise(avg = mean(value))
@@ -19,13 +25,17 @@ medians <- df %>% group_by(model, name) %>% summarise(med = median(value))
 
 ggplot(df, aes(x = value)) +
     geom_histogram(position="identity", aes(fill = model), alpha =0.3) +
-    facet_wrap(~name, scale="free") +
+    facet_wrap(~name, scale="free", labeller = label_parsed) +
     # geom_vline(data =means, aes(xintercept=avg, colour=model), size = 1.2) +
-    theme_minimal(base_size = 18) +
-    theme(legend.position="none") +
-    labs(x = "Parameter value", y= "Count")
+    theme_minimal(base_size = 20) +
+    scale_fill_discrete(labels = c("Method 1", "Method 2")) +
+    labs(title = "Posterior distribution - S&P 500",
+         fill = "MCMC\nSamplers",
+         x = "Parameter value", 
+         y= "Count") + 
+       theme(strip.text.x = element_text(size = 22))
 
-ggsave("manuscript/motivating_example/real_data_ex.png", bg = "white")
+ggsave("manuscript/motivating_example/real_data_ex.png", bg = "white", width = 14, height = 9.42)
     
     # geom_vline(data = medians, aes(xintercept=med, colour=model))
 
@@ -36,10 +46,3 @@ df %>% group_by(model, name) %>%
 ggplot(df, aes(x = value)) +
     geom_histogram(aes(fill = model), alpha =0.4) +
     facet_grid(model~name, scale="free")
-
-
[…]

### 2023-09-25 manuscript/motivating_example/single_sim_example.r

@@ -3,6 +3,7 @@ library(dplyr)
 library(cmdstanr)
 library(ggplot2)
 set.seed(630)
+set.seed(634)
 
 n_time_points <- 1000
 
@@ -38,7 +39,7 @@ gen_ncp_sv_dataset <- function(T){
 }
 
 data <- gen_ncp_sv_dataset(T = 1000)
-reuturns <- data$y_sim
+returns <- data$y_sim
 
 model_path <- "models/sv_user_guide_reparameterised_ksc_priors.stan"
 mod <- cmdstan_model(model_path, include_paths = here::here("models", "functions"))
@@ -57,8 +58,6 @@ draws <- model_fit$draws(variables = c("mu", "phi", "sigma"), format = "df") %>%
     select(-.chain, -.iteration, -.draw) %>% 
      pivot_longer(everything())
 
-draws$name <- replace(draws$name, draws$name=="sigma", "sigma_sqd")
-
 means = data.frame(value = rbind(data[['sigma_sqd']], data[['phi']], data[['mu']]),
            name = rbind('sigma_sqd', 'phi', 'mu'))
 
@@ -68,15 +67,37 @@ quantiles = model_fit$summary(variables = c("mu", "phi", "sigma")) %>%
               select(-name)
 
 names(quantiles) <- c("name", "value")
-quantiles$name <- replace(quantiles$name, quantiles$name=="sigma", "sigma_sqd")
+
+latex_variables <- function(df){
+  df$name <- replace(df$name, df$name=="sigma", "~sigma^2")
+  df$name <- replace(df$name, df$name=="sigma_sqd", "~sigma^2")
+  df$name <- replace(df$name, df$name=="mu", "~mu")
+  df$name <- replace(df$name, df$name=="phi", "~phi")
+
+  return(df)
+}
+
+draws <- latex_variables(draws)
+quantiles <- latex_variables(quantiles)
+means <- latex_variables(means)
+
+quantiles$name <- replace(quantiles$name, quantiles$name=="sigma", "~sigma^2")
+quantiles$name <- replace(quantiles$name, quantiles$name=="mu", "~mu")
+quantiles$name <- replace(quantiles$name, quantiles$name=="phi", "~phi")
+draws$name <- replace(draws$name, draws$name=="sigma_sqd", "~sigma^2")
+draws$name <- replace(draws$name, draws$name=="mu", "~mu")
+draws$name <- replace(draws$name, draws$name=="phi", "~phi")
 
 ggplot(draws, aes(x = value)) +
[…]

### 2023-09-25 manuscript/main.tex
Updated captions

@@
             linkcolor = black]{hyperref}
-\usepackage[font={small,it}]{caption}
-\usepackage[labelfont=bf]{caption}
+\usepackage[font={small,it}]{caption} % Figure caption font
+\usepackage[labelfont=bf]{caption} % Bold figure x.x
 
@@
         \includegraphics[scale=0.1]{motivating_example/real_data_ex.png}
-        \caption{Posterior samples from two MCMC algorithms using S\&P 500 data}
+        \caption{Posterior samples from two MCMC algorithms using S\&P 500 data.
+While the shape of the distributions are similar, the modes of method 2 is much higher and the tails of method 1 are fatter for $\phi$ and $\sigma^2$.
+Convergence diagnostics do not tell us which estimate is "more correct" since we don't observe the true model or true data generating process.}
     \end{figure}
@@
     These estimates show the marginal posteriors of each parameter sharing similar shapes.
-$\phi$  has a fatter left tail for and $\sigma^2$ has heavier right tails for sampling method 1 (SM1).
-Furthermore, SM2 has higher peaks at their mode relative to SM1 with a tighter spread.
-Based on this information alone, there is no certainty in whether SM1 or SM2 is has a more correct estimate conditional on data and model.
+$\phi$  has a fatter left tail for and $\sigma^2$ has heavier right tails for sampling method 1.
+Furthermore, sampling method 2 has higher peaks at their mode relative to method 1 with a tighter spread.
+Based on this information alone, there is no certainty in whether method 1 or method 2 is has a more correct estimate conditional on data and model.
 There are other methods for model or algorithmic selection in this context (for example, out of sample predictive performance), but for parameter estimation it is unclear which one should be selected.
@@
 
-    The below results are from a single simulation where data is generated and sampled from the stochastic volatility model with known parameters.
+    Figure 2 show the posterior distributions of a single simulation where data is generated and sampled from the stochastic volatility model with known parameters.
 
@@
         \includegraphics[scale=0.1]{motivating_example/single_sim.png}
-        \caption{Posterior samples from known data genearting process. Vertical solid line represents true parameter and dotted lines are 95\% credible intervals around the mean}
+        \caption{Posterior samples from known data genearting process.
+Vertical solid line represents true parameter and dotted lines are 95\% credible intervals around the mean.
+The parameter $\phi$ falls outside the credible interval whereas $\mu$ and $\sigma^2$ stay inside.
+In any one simulation conditional on the correct model, there is a 5\% probability that the true parameter falls outside this interval by chance.}
     \end{figure}
@@
 
-    \subsection{Hamiltonian Monte Carlo}
+    \subsection{Centered paramterisation}
 
-    \subsection{Gaussian Mixture Approximation}
+    \subsubsection{Hamiltonian Monte Carlo}
+
+    \subsubsection{Gaussian Mixture Approximation}
+
+    \subsection{Reparameterised models}
+
+    \subsubsection{Hamiltonian Monte Carlo}
+
+    \subsubsection{Gaussian Mixture Approximation}
 
@@
 \subsection{Limitations}
+- SBC is computationally expensive.
+
+- Datasets generated dependent on the priors. If priors are diffuse then can get a wide set of possible datasets which make it harder to get calibrated estimates of the true parameter.
 

### 2023-09-25 R/sbc.r
updated labelling functions

@@ -29,9 +29,10 @@ facet_hist <- function(data, variables, nbins, expected_bin_count, rank_scales) 
         ggplot(.) +
         geom_histogram(aes(rank), bins = nbins, fill = "light blue", alpha = 0.7) +
         scale_x_continuous(labels = function(x) x * rank_scales/nbins) +
-        facet_wrap(~variable) +
-        theme_minimal(base_size = 22) +
-        geom_hline(yintercept = expected_bin_count, size = 0.5, alpha = 0.3) +
+        facet_wrap(~variable, labeller = label_parsed) +
+        theme_minimal(base_size = 20) +
+        theme(strip.text.x = element_text(size = 22)) +
+        geom_hline(yintercept = expected_bin_count, linewidth = 0.5, alpha = 0.3) +
         labs(title = "Distribution of Rank Statistics",
                 y = "Count",
                 x = "Ranks")
@@ -139,25 +140,59 @@ get_time <- function(rds_path, model_path) {
 }
 
 ess_boxplot <- function(data, variables, plot_title, ess_type) {
-    data %>%
+    plot <- data %>%
         filter(parameters %in% variables) %>% 
         ggplot(.) +
         geom_boxplot(aes(x = parameters, y = {{ess_type}})) +
-        theme_minimal() +
-        labs(title = plot_title)
+        theme_minimal(base_size = 22) +
+        labs(title = plot_title,
+             x = "Parameters",
+             y = "Count")
+
+    x_labs = ggplot_build(plot)$layout$panel_params[[1]]$x$get_labels()
+    plot <- plot + scale_x_discrete(labels = parse(text = x_labs))
+
+    return(plot)
 }
 
 rhat_boxplot <- function(data, variables, plot_title, rhat_type) {
-    data %>%
+    plot <- data %>%
         filter(parameters %in% variables) %>% 
         ggplot(.) +
         geom_boxplot(aes(x = parameters, y = {{rhat_type}})) +
-        theme_minimal() +
-        labs(title = plot_title)
+        theme_minimal(base_size = 22) +
+        labs(title = plot_title,
+             x = "Parameters",
+             y = "Count")
+
+     x_labs = ggplot_build(plot)$layout$panel_params[[1]]$x$get_labels()
+     plot <- plot + scale_x_discrete(labels = parse(text = x_labs))
+     
+     return(plot)
+
 }
 
[…]

### 2023-09-25 analysis/sbc/sbc_results.r

@@ -7,22 +7,18 @@ source("R/sbc.r")
 plan(multicore, workers = 9)
 
 # Get all parameter data
-path <- "simulation_output/stan/ncp/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r9_5000_iterations"
+path <- "simulation_output/ksc/cp/sir_sbc_cp_ksc_model_cp_dgf_10kmcmc_1000iter_r3"
 rds_list <- list.files(path = paste(path, "output", sep = "/"), pattern = "*.RDS")
 rds_path <- paste("output", rds_list, sep = "/")
 
 # Rank type
-# rank_type = "weighted_ranks"
 rank_type = "agg_ranks"
+rank_type = "weighted_ranks"
 
-# get_ranks <- function(rds_path, model_path, rank_type) {
-#     data = readRDS(here::here(paste(model_path, rds_path, sep = "/")))
-#     return(data$all_chains[[rank_type]])
-# }
 
 # Parameters
 n_iterations <- length(rds_list)
-posterior_samples <- 999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+posterior_samples <- 9999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
 n_bins <- 20 # Bins
 expected_count <- n_iterations / n_bins # Chi sqd stat
 
@@ -31,21 +27,28 @@ results <- lapply(rds_path, get_ranks, model_path = path, rank_type = rank_type)
 
 if (rank_type == "agg_ranks"){
     df <- as.data.frame(do.call("rbind", results))
+    rank_bins <- as.data.frame(lapply(df, binning, posterior_draws = posterior_samples, bins = n_bins))
 } else if(rank_type == "weighted_ranks"){
     df <- as.data.frame(do.call("rbind", results)) * (posterior_samples+1)
+    rank_bins <- as.data.frame(lapply(df, binning, posterior_draws = posterior_samples+1, bins = n_bins))
 } else {
     print("error")
 }
 
+
+names(df) <- clean_variable_names(df)
+names(rank_bins) <- clean_variable_names(rank_bins)
+
+
 # Create bins and chi sq stats
-rank_bins <- as.data.frame(lapply(df, binning, posterior_draws = posterior_samples, bins = n_bins))
 rank_bins <- rank_bins[!(names(rank_bins) %in% c("y_sim", "sigma"))]
[…]

### 2023-09-26 manuscript/results/rank_histograms.r
preparing results

@@ -1,0 +1,113 @@
+library(tidyverse)
+source("R/sbc.r")
+
+posterior_samples <- 999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+n_bins <- 20 # Bins
+expected_count <- n_iterations / n_bins # Chi sqd stat
+
+
+# HMC results
+# 1000 iterations
+rank_type = "agg_ranks"
+path <- "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_1000_iterations"
+
+rank_bins_f <- function(path, rank_type, posterior_samples, n_bins){
+    rds_list <- list.files(path = paste(path, "output", sep = "/"), pattern = "*.RDS")
+    rds_path <- paste("output", rds_list, sep = "/")
+
+    n_iterations <- length(rds_path)
+    posterior_samples <- 999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+    n_bins <- 20 # Bins
+    expected_count <- n_iterations / n_bins # Chi sqd stat
+
+    results <- lapply(rds_path, get_ranks, model_path = path, rank_type = rank_type)
+
+    if (rank_type == "agg_ranks"){
+        df <- as.data.frame(do.call("rbind", results))
+        rank_bins <- as.data.frame(lapply(df, binning, posterior_draws = posterior_samples, bins = n_bins))
+    } else if(rank_type == "weighted_ranks"){
+        df <- as.data.frame(do.call("rbind", results)) * (posterior_samples+1)
+        rank_bins <- as.data.frame(lapply(df, binning, posterior_draws = posterior_samples+1, bins = n_bins))
+    } else {
+        print("error")
+    }
+
+    names(df) <- clean_variable_names(df)
+    names(rank_bins) <- clean_variable_names(rank_bins)
+
+    rank_bins <- rank_bins[!(names(rank_bins) %in% c("y_sim", "sigma"))]
+
+    return(rank_bins)
+}
+
+static_rank_bins <- function(data, variables, type){
+    result <- data %>%
+        select(all_of(variables)) %>%
+        tidyr::pivot_longer(everything(), names_to = "variable", values_to = "rank") %>% 
+        mutate(type = type)
+
+    return(result)
+}
+
+
[…]

### 2023-09-27 manuscript/main.tex
updated plots

@@
 % \bibliographystyle{unsrtnat}
-
+\usepackage{float}
 \usepackage[
@@
 
-    \subsection{Centered paramterisation}
+    \subsection{Hamiltonian Monte Carlo}
 
-    \subsubsection{Hamiltonian Monte Carlo}
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/hmc_1k.png}
+        \caption{Distribution of rank statistics for centered and non centered Hamiltoninan Monte Carlo (static parameters).
+Non centered histograms look approximately uniform.
+The centered parameterisation struggles to estimate $\sigma^2$}
+    \end{figure}
 
-    \subsubsection{Gaussian Mixture Approximation}
+    \subsection{Gaussian Mixture Approximation}
 
-    \subsection{Reparameterised models}
-
-    \subsubsection{Hamiltonian Monte Carlo}
-
-    \subsubsection{Gaussian Mixture Approximation}
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/ksc_1k.png}
+        \caption{Distribution of rank statistics for centered and non centered Gaussian mixture model(static parameters). All parameters display a lack of calibration}
+    \end{figure}
 
     \subsection{Reweigthing rank statistics of Gaussian Mixture}
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/sir_1k.png}
+        \caption{Distribution of rank statistics for weighted resampled model (static parameters) with 1000 SBC iterations. All parameters display a lack of calibration}
+    \end{figure}
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/sir_5k.png}
+        \caption{Distribution of rank statistics for weighted resampled model (static parameters) with 5000 SBC iterations. Increasing number of SBC parameters does not improve results}
+    \end{figure}
+
+    \subsection{All models and parameters}
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/dist_chisq.png}
+        \caption{Distribution of chi squared statistics for all models.}
+    \end{figure}
 

### 2023-09-27 manuscript/motivating_example/single_sim_example.r

@@ -81,13 +81,6 @@ draws <- latex_variables(draws)
 quantiles <- latex_variables(quantiles)
 means <- latex_variables(means)
 
-quantiles$name <- replace(quantiles$name, quantiles$name=="sigma", "~sigma^2")
-quantiles$name <- replace(quantiles$name, quantiles$name=="mu", "~mu")
-quantiles$name <- replace(quantiles$name, quantiles$name=="phi", "~phi")
-draws$name <- replace(draws$name, draws$name=="sigma_sqd", "~sigma^2")
-draws$name <- replace(draws$name, draws$name=="mu", "~mu")
-draws$name <- replace(draws$name, draws$name=="phi", "~phi")
-
 ggplot(draws, aes(x = value)) +
   geom_histogram(aes(fill = name), alpha = 0.3) +
   theme_minimal(base_size = 20) +

### 2023-09-27 manuscript/results/compare_histograms.r

@@ -1,0 +1,81 @@
+library(tidyverse)
+source("R/sbc.r")
+
+n_iterations <- 1000
+posterior_samples <- 999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+n_bins <- 20 # Bins
+expected_count <- n_iterations / n_bins # Chi sqd stat
+
+
+# HMC CENTERED results
+# 1000 iterations
+rank_type = "agg_ranks"
+path <- "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_1000_iterations"
+
+hmc_cp_onek_bins <- rank_bins_f(path = "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_1000_iterations", 
+                                rank_type = "agg_ranks", 
+                                posterior_samples = 999, 
+                                n_bins= 20)
+
+hmc_cprank_stats <- as.data.frame(lapply(hmc_cp_onek_bins, rank_chi_sqd, expected = expected_count)) %>% 
+            pivot_longer(everything())
+hmc_cprank_stats$name <- names(hmc_cp_onek_bins)
+hmc_cprank_stats$type <- "hmc_cp"
+
+# KSC CENTERED results
+n_iterations <- 1000
+posterior_samples <- 9999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+n_bins <- 20 # Bins
+expected_count <- n_iterations / n_bins # Chi sqd stat
+
+ksc_cp_onek_bins <- rank_bins_f(path = "simulation_output/ksc/cp/sbc_cp_ksc_model_cp_dgf_10kmcmc_1000iter_r2", 
+                                    rank_type = "agg_ranks", 
+                                    posterior_samples = posterior_samples, 
+                                    n_bins= n_bins)
+ksc_cp_rank_stats <- as.data.frame(lapply(ksc_cp_onek_bins, rank_chi_sqd, expected = expected_count)) %>% 
+            pivot_longer(everything())
+
+ksc_cp_rank_stats$name <- names(ksc_cp_onek_bins)
+ksc_cp_rank_stats$type <- "ksc_cp"
+
+
+
+# SIR 1k
+n_iterations <- 1000
+posterior_samples <- 9999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+n_bins <- 20 # Bins
[…]

### 2023-09-28 manuscript/main.tex
Extra histograms and ess plots

@@
         \includegraphics[scale=0.1]{results/hmc_1k.png}
-        \caption{Distribution of rank statistics for centered and non centered Hamiltoninan Monte Carlo (static parameters).
+        \caption{Distribution of 1000 rank statistics for centered and non centered Hamiltoninan Monte Carlo (static parameters).
 Non centered histograms look approximately uniform.
 The centered parameterisation struggles to estimate $\sigma^2$}
+    \end{figure}
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/hmc_5k.png}
+        \caption{Distribution of 5000 rank statistics for centered and non centered Hamiltoninan Monte Carlo (static parameters).
+Non centered histograms look approximately uniform.
+The centered parameterisation struggles to estimate $\sigma^2$}
+    \end{figure}
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/hmc_ess.png}
+        \caption{Boxplots of effective sample size after 5000 iterations of SBC}
     \end{figure}
@@
         \includegraphics[scale=0.1]{results/ksc_1k.png}
-        \caption{Distribution of rank statistics for centered and non centered Gaussian mixture model(static parameters). All parameters display a lack of calibration}
+        \caption{Distribution of 1000 rank statistics for centered and non centered Gaussian mixture model(static parameters). All parameters display a lack of calibration}
     \end{figure}
@@
         \includegraphics[scale=0.1]{results/sir_1k.png}
-        \caption{Distribution of rank statistics for weighted resampled model (static parameters) with 1000 SBC iterations. All parameters display a lack of calibration}
+        \caption{Distribution of 1000 rank statistics for weighted resampled model (static parameters) with 1000 SBC iterations. All parameters display a lack of calibration}
     \end{figure}
@@
         \includegraphics[scale=0.1]{results/sir_5k.png}
-        \caption{Distribution of rank statistics for weighted resampled model (static parameters) with 5000 SBC iterations. Increasing number of SBC parameters does not improve results}
+        \caption{Distribution of 5000 rank statistics for weighted resampled model (static parameters) with 5000 SBC iterations. Increasing number of SBC parameters does not improve results}
     \end{figure}

### 2023-09-28 manuscript/results/ess.r

@@ -1,0 +1,69 @@
+library(tidyverse)
+source("R/sbc.r")
+static_parameters <- c("~mu", "~sigma^2", "~phi")
+
+# HMC results
+# 5000 iterations CP
+path <- "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_5000_iterations"
+rds_list <- list.files(path = paste(path, "output", sep = "/"), pattern = "*.RDS")
+rds_path <- paste("output", rds_list, sep = "/")
+
+hmc_cp_ess_basic_5k <- future_lapply(rds_path,
+                           get_ess, 
+                           model_path = path, 
+                           ess_type = "ess_basic", 
+                           chains = "one_chain", 
+                           future.seed = NULL)
+
+hmc_cp_ess_basic_5k_df <- as.data.frame(do.call("rbind", hmc_cp_ess_basic_5k))
+hmc_cp_ess_basic_5k_df <- latex_variables(hmc_cp_ess_basic_5k_df)
+hmc_cp_ess_basic_5k_df$Type <- "Centered"
+
+
+# 5000 iterations NCP
+path <- "simulation_output/stan/ncp/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r9_5000_iterations"
+rds_list <- list.files(path = paste(path, "output", sep = "/"), pattern = "*.RDS")
+rds_path <- paste("output", rds_list, sep = "/")
+
+hmc_ncp_ess_basic_5k <- future_lapply(rds_path,
+                           get_ess, 
+                           model_path = path, 
+                           ess_type = "ess_basic", 
+                           chains = "one_chain", 
+                           future.seed = NULL)
+
+hmc_ncp_ess_basic_5k_df <- as.data.frame(do.call("rbind", hmc_ncp_ess_basic_5k))
+hmc_ncp_ess_basic_5k_df <- latex_variables(hmc_ncp_ess_basic_5k_df)
+hmc_ncp_ess_basic_5k_df$Type <- "Non Centered"
+
+hmc_ess_df <- rbind(hmc_cp_ess_basic_5k_df, hmc_ncp_ess_basic_5k_df) %>% 
+    filter(parameters %in% static_parameters)
+
+hmc_ess_df <- latex_variables(hmc_ess_df)
+
+plot <- ggplot(hmc_ess_df, aes(x = parameters, y = ess_basic, fill = Type)) +
[…]

### 2023-09-28 manuscript/results/rank_histograms.r

@@ -1,6 +1,7 @@
 library(tidyverse)
 source("R/sbc.r")
 
+n_iterations <- 1000
 posterior_samples <- 999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
 n_bins <- 20 # Bins
 expected_count <- n_iterations / n_bins # Chi sqd stat
@@ -49,7 +50,7 @@ static_rank_bins <- function(data, variables, type){
     return(result)
 }
 
-
+# HMC 1k
 hmc_cp_onek_bins <- rank_bins_f(path = "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_1000_iterations", 
                                 rank_type = "agg_ranks", 
                                 posterior_samples = 999, 
@@ -76,13 +77,129 @@ ggplot(hmc_1k) +
         theme_minimal(base_size = 20) +
         theme(strip.text.x = element_text(size = 22)) +
         geom_hline(yintercept = expected_count, linewidth = 0.5, alpha = 0.3) +
-        labs(title = "Distribution of Rank Statistics",
-                y = "Count",
-                x = "Ranks")
+        labs(y = "Count",
+             x = "Ranks")
+
+ggsave("manuscript/results/hmc_1k.png", bg = "white", width = 14, height = 9.42)
+
+# HMC 5k
+n_iterations <- 5000
+posterior_samples <- 999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+n_bins <- 20 # Bins
+expected_count <- n_iterations / n_bins # Chi sqd stat
+
+hmc_cp_fivek_bins <- rank_bins_f(path = "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_5000_iterations", 
+                                rank_type = "agg_ranks", 
+                                posterior_samples = 999, 
+                                n_bins= 20)
+
+hmc_ncp_fivek_bins <- rank_bins_f(path = "simulation_output/stan/ncp/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r9_5000_iterations", 
+                                rank_type = "agg_ranks", 
+                                posterior_samples = 999, 
+                                n_bins= 20)
+
+hmc_cp_fivek_bin_long <- static_rank_bins(data = hmc_cp_fivek_bins, 
[…]

### 2023-09-28 manuscript/results/ess.r
updated plots

@@ -51,6 +51,55 @@ x_labs = ggplot_build(plot)$layout$panel_params[[1]]$x$get_labels()
 plot <- plot + scale_x_discrete(labels = parse(text = x_labs))
 ggsave("manuscript/results/hmc_ess.png", bg = "white", width = 14, height = 9.42)
 
+# KSC results
+# 5000 iterations CP
+path <- "simulation_output/ksc/cp/sbc_cp_ksc_model_cp_dgf_10kmcmc_5000iter_r1"
+rds_list <- list.files(path = paste(path, "output", sep = "/"), pattern = "*.RDS")
+rds_path <- paste("output", rds_list, sep = "/")
+
+ksc_cp_ess_basic_5k <- future_lapply(rds_path,
+                           get_ess, 
+                           model_path = path, 
+                           ess_type = "ess_basic", 
+                           chains = "one_chain", 
+                           future.seed = NULL)
+
+ksc_cp_ess_basic_5k_df <- as.data.frame(do.call("rbind", ksc_cp_ess_basic_5k))
+ksc_cp_ess_basic_5k_df <- latex_variables(ksc_cp_ess_basic_5k_df)
+ksc_cp_ess_basic_5k_df$Type <- "Gaussian Mixture"
+
+
+# 5000 iterations NCP
+path <- "simulation_output/ksc/cp/sir_sbc_cp_ksc_model_cp_dgf_10kmcmc_5000iter_r1"
+rds_list <- list.files(path = paste(path, "output", sep = "/"), pattern = "*.RDS")
+rds_path <- paste("output", rds_list, sep = "/")
+
+sir_ncp_ess_basic_5k <- future_lapply(rds_path,
+                           get_ess, 
+                           model_path = path, 
+                           ess_type = "ess_basic", 
+                           chains = "one_chain", 
+                           future.seed = NULL)
+
+sir_ncp_ess_basic_5k_df <- as.data.frame(do.call("rbind", sir_ncp_ess_basic_5k))
+sir_ncp_ess_basic_5k_df <- latex_variables(sir_ncp_ess_basic_5k_df)
+sir_ncp_ess_basic_5k_df$Type <- "Importance Resampling"
+
+ksc_sir_df <- rbind(ksc_cp_ess_basic_5k_df, sir_ncp_ess_basic_5k_df) %>% 
+    filter(parameters %in% static_parameters)
+
+ksc_sir_df <- latex_variables(ksc_sir_df)
[…]

### 2023-09-28 manuscript/results/rank_histograms_states.r

@@ -1,0 +1,251 @@
+library(tidyverse)
+source("R/sbc.r")
+
+parameters <- c("h[1]", "h[500]", "h[1000]")
+
+n_iterations <- 1000
+posterior_samples <- 999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+n_bins <- 20 # Bins
+expected_count <- n_iterations / n_bins # Chi sqd stat
+
+
+# HMC results
+# 1000 iterations
+rank_type = "agg_ranks"
+path <- "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_1000_iterations"
+
+# HMC 1k
+hmc_cp_onek_bins <- rank_bins_f(path = "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_1000_iterations", 
+                                rank_type = "agg_ranks", 
+                                posterior_samples = 999, 
+                                n_bins= 20)
+
+hmc_ncp_onek_bins <- rank_bins_f(path = "simulation_output/stan/ncp/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r8_1000_iterations", 
+                                rank_type = "agg_ranks", 
+                                posterior_samples = 999, 
+                                n_bins= 20)
+
+hmc_cp_onek_bin_long <- static_rank_bins(data = hmc_cp_onek_bins, 
+                                         variables = parameters, 
+                                         type = "Centered")
+
+hmc_ncp_onek_bin_long <- static_rank_bins(data = hmc_ncp_onek_bins, 
+                                          variables = parameters, 
+                                          type = "NonCentered")
+
+hmc_1k = rbind(hmc_cp_onek_bin_long, hmc_ncp_onek_bin_long)
+
+ggplot(hmc_1k) +
+        geom_histogram(aes(rank), bins = 20, fill = "light blue", alpha = 0.7) +
+        scale_x_continuous(labels = function(x) x * (posterior_samples+1)/n_bins) +
+        facet_grid(type~factor(variable, levels = parameters), labeller = label_parsed) +
+        theme_minimal(base_size = 20) +
+        theme(strip.text.x = element_text(size = 22)) +
+        geom_hline(yintercept = expected_count, linewidth = 0.5, alpha = 0.3) +
+        labs(y = "Count",
+             x = "Ranks")
+
[…]

### 2023-09-28 manuscript/results/rank_histograms_static.r

@@ -1,0 +1,220 @@
+library(tidyverse)
+source("R/sbc.r")
+
+n_iterations <- 1000
+posterior_samples <- 999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+n_bins <- 20 # Bins
+expected_count <- n_iterations / n_bins # Chi sqd stat
+
+
+# HMC results
+# 1000 iterations
+rank_type = "agg_ranks"
+path <- "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_1000_iterations"
+
+# HMC 1k
+hmc_cp_onek_bins <- rank_bins_f(path = "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_1000_iterations", 
+                                rank_type = "agg_ranks", 
+                                posterior_samples = 999, 
+                                n_bins= 20)
+
+hmc_ncp_onek_bins <- rank_bins_f(path = "simulation_output/stan/ncp/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r8_1000_iterations", 
+                                rank_type = "agg_ranks", 
+                                posterior_samples = 999, 
+                                n_bins= 20)
+
+hmc_cp_onek_bin_long <- static_rank_bins(data = hmc_cp_onek_bins, 
+                                         variables = c("~mu", "~sigma^2", "~phi"), 
+                                         type = "Centered")
+hmc_ncp_onek_bin_long <- static_rank_bins(data = hmc_ncp_onek_bins, 
+                                          variables = c("~mu", "~sigma^2", "~phi"), 
+                                          type = "NonCentered")
+
+hmc_1k = rbind(hmc_cp_onek_bin_long, hmc_ncp_onek_bin_long)
+
+ggplot(hmc_1k) +
+        geom_histogram(aes(rank), bins = 20, fill = "light blue", alpha = 0.7) +
+        scale_x_continuous(labels = function(x) x * (posterior_samples+1)/n_bins) +
+        facet_grid(type~variable, labeller = label_parsed) +
+        theme_minimal(base_size = 20) +
+        theme(strip.text.x = element_text(size = 22)) +
+        geom_hline(yintercept = expected_count, linewidth = 0.5, alpha = 0.3) +
+        labs(y = "Count",
+             x = "Ranks")
+
[…]

### 2023-09-28 manuscript/main.tex
updated plots
typo 5k not 1k

@@
         \centering
+        \includegraphics[scale=0.1]{results/hmc_states_1k.png}
+        \caption{Distribution of 1000 rank statistics for centered and non centered Hamiltoninan Monte Carlo (state parameters).}
+    \end{figure}
+
+    \begin{figure}[H]
+        \centering
         \includegraphics[scale=0.1]{results/hmc_5k.png}
@@
 The centered parameterisation struggles to estimate $\sigma^2$}
+    \end{figure}
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/hmc_states_5k.png}
+        \caption{Distribution of 5000 rank statistics for centered and non centered Hamiltoninan Monte Carlo (state parameters).}
     \end{figure}
@@
         \caption{Distribution of 1000 rank statistics for centered and non centered Gaussian mixture model(static parameters). All parameters display a lack of calibration}
+    \end{figure}
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/ksc_5k.png}
+        \caption{Distribution of 5000 rank statistics for centered and non centered Gaussian mixture model(static parameters). All parameters display a lack of calibration}
     \end{figure}

### 2023-10-01 manuscript/results/rank_histograms_static_states.r
plot static and states on rank facets

@@ -1,0 +1,188 @@
+library(tidyverse)
+source("R/sbc.r")
+
+n_iterations <- 1000
+posterior_samples <- 999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+n_bins <- 20 # Bins
+expected_count <- n_iterations / n_bins # Chi sqd stat
+
+parameters <- c("~mu", "~sigma^2", "~phi", "h[1]", "h[500]", "h[1000]")
+
+# HMC results
+# 1000 iterations
+rank_type = "agg_ranks"
+path <- "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_1000_iterations"
+
+
+# HMC 1k
+hmc_cp_onek_bins <- rank_bins_f(path = "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_1000_iterations", 
+                                rank_type = "agg_ranks", 
+                                posterior_samples = 999, 
+                                n_bins= 20)
+
+hmc_ncp_onek_bins <- rank_bins_f(path = "simulation_output/stan/ncp/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r8_1000_iterations", 
+                                rank_type = "agg_ranks", 
+                                posterior_samples = 999, 
+                                n_bins= 20)
+
+hmc_cp_onek_bin_long <- static_rank_bins(data = hmc_cp_onek_bins, 
+                                         variables = parameters, 
+                                         type = "Centered")
+hmc_ncp_onek_bin_long <- static_rank_bins(data = hmc_ncp_onek_bins, 
+                                          variables = parameters, 
+                                          type = "NonCentered")
+
+hmc_1k = rbind(hmc_cp_onek_bin_long, hmc_ncp_onek_bin_long)
+
+rank_hist <- function(dataframe){
+        ggplot(dataframe) +
+        geom_histogram(aes(rank), bins = 20, fill = "light blue", alpha = 0.7) +
+        scale_x_continuous(labels = function(x) x * (posterior_samples+1)/n_bins) +
+        facet_wrap(~factor(variable, levels = parameters), labeller = label_parsed) +
+        theme_minimal(base_size = 20) +
+        theme(strip.text.x = element_text(size = 22)) +
[…]

### 2023-10-01 manuscript/methodology/rank_stats_examples.r
rank stat examples

@@ -1,0 +1,13 @@
+library(ggplot2)
+df = as.data.frame(list(v1=rnorm(5e6)))
+
+ggplot(df, aes(x=v1))+ geom_density(linewidth=1.2) + 
+    theme_void() +
+    geom_vline(xintercept = -4, linewidth = 1)
+
+hist_df = as.data.frame(list(v1 = c(rep(1,500), runif(3000, 2,20))))
+ggplot(hist_df, aes(x=v1)) +
+    geom_histogram(bins=20) +
+    theme_void()
+
+

### 2023-10-02 manuscript/main.tex
updated results

--- 2023-09-27 manuscript/main.tex
@@
             linkcolor = black]{hyperref}
-\usepackage[font={small,it}]{caption} % Figure caption font
+\usepackage[font={scriptsize}]{caption} % Figure caption font
 \usepackage[labelfont=bf]{caption} % Bold figure x.x 
+\captionsetup{width=0.9\textwidth}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 \section{Results}
+    1000 and 5000 SBC iterations are run for centered and reparameterised versions of the stocashtic volatility model.
+Rank statistics histograms are presented for the static parameters and the 1st, 500th and 1000th latent state parameters with 20 bins constructed for each parameter.
+The black horizontal line represents uniformity conditional on the number of bins and SBC iterations.
+The $\chi^2$ statistic is used to summarise the deviations from uniformity for all parameters since it is unrealistic to visually inspect all the histograms.
+The SBC algorithm is summarised below:
+
+    1) for sim in 1000 iterations:
+        2) Draw from prior: $\theta^{sim}\sim\pi (\theta)$
+        3) Simulate dataset with 1000 observations: $y^{sim} \sim p(y|\theta^{sim})$
+        4) Draw 999 posterior samples (post warmup) $\{\theta_1,\dots , \theta_{L}\} \sim p(\theta | y^{sim})$
+        5) Compute rank statistics $r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim})$
 
     \subsection{Hamiltonian Monte Carlo}
+    Figure \ref{fig:cphmc1k} and Figure \ref{fig:ncphmc1k} show the SBC results for the centered and reparameterisd model for HMC respectively.
+Both sets of results display relatively noisy rank statistic estimates with variation around the discrete uniform distribution.
+
+    The centered parameterisation in particular displays a lack of uniformity for $\phi$ and $\sigma^2$.
+Rank statistics for $\sigma^2$ show two peaks of either end of the histogram and a left peak for $phi$.
+Posterior estimates of $\sigma^2$ are \textit{underdispersed} relative to the true parameter $\sigma^2_{sim}$ generated by the prior distribution.
+On the other hand, the posterior samples of $\phi$ are \textit{over estimating} the true parameter generated by the prior.
+This suggests the posterior geometry for these parameters (at least) are difficult to sample for the HMC algorithm.
 
@@
         \centering
-        \includegraphics[scale=0.1]{results/hmc_1k.png}
-        \caption{Distribution of rank statistics for centered and non centered Hamiltoninan Monte Carlo (static parameters).
-Non centered histograms look approximately uniform.
-The centered parameterisation struggles to estimate $\sigma^2$}
+        \includegraphics[scale=0.09]{results/hmc_cp_1k.png}
+        \caption{1000 SBC iterations for centered model using Hamiltoninan Monte Carlo.
+Histograms show noisy estimates around uniform distribution.
+In particular, $\sigma^2$ displays peaks on both ends of the histogram and $\phi$ shows a major on the left hand side.
+This suggests that the posterior samples of $\sigma^2$ are underdispersed relative to the prior distribution and posterior samples of $\phi$ are over estimating the true parameter.}
+        \label{fig:cphmc1k}
     \end{figure}
 
-    \subsection{Gaussian Mixture Approximation}
+    Estimates of $\sigma^2$ and $\phi$ are more uniform under the reparameterised model.
+Instead, we observe a peak on the left side for the 500th state parameter $h_{500}$.
+Overall the estimates for non centered parameterisation are more favourable.
 
@@
         \centering
-        \includegraphics[scale=0.1]{results/ksc_1k.png}
-        \caption{Distribution of rank statistics for centered and non centered Gaussian mixture model(static parameters). All parameters display a lack of calibration}
+        \includegraphics[scale=0.09]{results/hmc_ncp_1k.png}
+        \caption{1000 SBC iterations for non centered model using Hamiltonian Monte Carlo.
+The rank statistic distributions for $\sigma^2$ and $\phi$ have improved.
+There is some evidence of over estimation of the 500th state parameter.}
+        \label{fig:ncphmc1k}
     \end{figure}
 
-    \subsection{Reweigthing rank statistics of Gaussian Mixture}
+    The noisy histogram estimates can be improved by increasing the number of SBC iterations.
+Figure \ref{fig:cphmc5k} and Figure \ref{fig:ncphmc5k} show the results of the same parameters with 5000 SBC iterations.
+There is an improvement to the reparameterised model with all histograms looking approximately uniform with much smaller variation around the true uniform value.
+
+    Centered parameteristion posess similar improvement for $\mu$ and the state parameters.
+However, there are still major peaks at both ends of $\sigma^2$ and $\phi$ suggesting underdispersion of the posterior estimates relative to the prior distribution.
+
+    These results suggests that HMC overall struggles to sample from the centered parameteristion of the stochastic volatility model.
+The reparameterised model results in much better and consistent inference for this sampling strategy, conditional on the model.
 
@@
         \centering
-        \includegraphics[scale=0.1]{results/sir_1k.png}
-        \caption{Distribution of rank statistics for weighted resampled model (static parameters) with 1000 SBC iterations. All parameters display a lack of calibration}
+        \includegraphics[scale=0.09]{results/hmc_cp_5k.png}
+        \caption{5000 SBC iterations for centered model using Hamiltoninan Monte Carlo.
+$\mu$ and the state parameters look approximately uniform.
+$\sigma^2$ still has major peaks on both the left and right hand side indicating underdispersion relative to the prior distribution.
+$\phi$ exhibits a similar shape to $\sigma^2$ although the peaks are much smaller.}
+        \label{fig:cphmc5k}
     \end{figure}
@@
         \centering
-        \includegraphics[scale=0.1]{results/sir_5k.png}
-        \caption{Distribution of rank statistics for weighted resampled model (static parameters) with 5000 SBC iterations. Increasing number of SBC parameters does not improve results}
+        \includegraphics[scale=0.09]{results/hmc_ncp_5k.png}
+        \caption{5000 SBC iterations for reparameterised model using Hamiltoninan Monte Carlo.
+Parameters display uniformity with less noise around the true uniform value.
+This suggests that the HMC algorithm is returning the correct posterior estimates for the selected parameters.}
+        \label{fig:ncphmc5k}
     \end{figure}
 
-    \subsection{All models and parameters}
+    Further to calibration, the HMC algorithm struggles to efficiently generate posterior samples from the centered parameterisation of stochastic volatility.
+Figure \ref{fig:hmcess} visualises the distribution of ESS for the static parameters after 5000 SBC iterations.
+The non centered model appears to generate independent samples much more efficiently than the centered model.
+This is consistent with HMC struggliing to sample overall from the posteriors of $\sigma^2$ and $\phi$.
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.09]{results/hmc_ess.png}
+        \caption{Effective sample sizes for static parameters after 5000 iterations of SBC and 1000 post warmup samples for Hamiltoninan Monte Carlo.
+The centered parameterisation struggles to generate independent samples for $\sigma^2$ and $\phi$.
+Non parameterisation demonstrates much more efficient sampling behaviour.}
+        \label{fig:hmcess}
+    \end{figure}
+
+    \subsection{Gaussian Mixture Approximation}
+    The Gaussian Mixture Approximation is run for 1000 and 5000 SBC iterations with the first 10,000 samples discarded as burnin (consistent with \citet{kim1998stochastic}).
+However, unlike KSC who take 750,000 posterior draws from the model, the number of posterior samples taken as part of the SBC procedure is 9,999.
+Taking 750,000 draws from the Gaussian mixture approximation would lead to issues given the disk space and compute time required from run SBC.
+Furthermore, the number of posterior samples is larger than the 999 taken for the SBC experiment run for Hamiltonian Monte Carlo.
+Earlier experiments attempted 999 post burn in draws for the Gaussian approximation, however there are concerns that such a short chain would not have converged onto the target posterior distribution.
+Therefore 9,999 draws was chosen as a compromise under both constraints.
+
+    The centered Gaussian mixture approximation contains issues with calibration.
+Figure \ref{fig:cpksc1k} presents the results from SBC where there is a lack of uniformity across all parameters.
+In particular, there is a large peak on the left hand side for $\mu$ and the right hand hand side of $\phi$.
+This suggests over estimation and under estimation of the true parameter respectively.
+All other parameters do not have any key features worth noting beyond noisy estimates around the true discrete uniform value.
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/ksc_cp_1k.png}
+        \caption{1000 SBC iterations for centered Gaussian mixture approximation model.
+The rank statistics for all the selected parameters display a lack of uniformity.
+$\mu$ has a large left peak suggesting the algorithm is over estimating the true parameter.
+The converse can be said about $\phi$ which possesses a peak on the right hand side which implies underestimation of the true parameter.}
+        \label{fig:cpksc1k}
+    \end{figure}
+
+    Increasing the SBC iterations to 5000 does not improve the results.
+$\sigma^2$ and the state parameters contain arguably less noisy estimates around the uniform distribution.
+However, the skewed shapes of $\mu$ and $\phi$ adds further evidence that the sampling strategy is not returning the correct posterior estimates.
+This can be seen in Figure \ref{fig:cpksc5k}.
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/ksc_cp_5k.png}
+        \caption{5000 SBC iterations for centered Gaussian mixture approximation model.
+Estimates for $\sigma^2$ and the state parameters have marginally improved around the uniformity.
+However, there is no improvement in $\mu$ or $\phi$, suggesting a lack of calibration from the MCMC algorithm.}
+        \label{fig:cpksc5k}
+    \end{figure}
+
+    Furthermore, the ESS estimates for this model indicate a high degree of autocorrelation and difficulty in generating independent samples from the posterior distribution.
+The median ESS for $\mu$, $\phi$ and $\sigma^2$ are 559, 103, 39.5 respectively for 10,000 post burn in draws.
+Despite having 10 times the number of draws, the number of effectively independent samples is much smaller relative to HMC, suggesting that this bespoke sampling strategy is highly inefficient.
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/ksc_ess.png}
+        \caption{Effective sample sizes for static parameters after 5000 iterations of SBC and 1000 post burnin samples for the Gaussian Mixture model.
+The bespoke sampling strategy struggles to generate independent samples for all static parameters, suggesting a high degree of autocorrelation in the Markkov Chain.}
+        \label{fig:kscess}
+    \end{figure}
+
+    The reparameterised model using the bespoke sampling strategy fails to generate sufficient results due to numerical errors and highly nonstationary chains.
+Figure X shows the traceplot from one simulation from the reparamterised model.
+The chain for $\phi$ is stuck to a value very close to 1 (unit root) and $\mu$ and $\sigma^2$ continuously increases in value until a numerical error is reached.
+Note that this occurs during the burnin phase of the markov chain.
+As a result, SBC is unable to be performed on this parameterisation of the model.
+
+    There are a variety of reasons why this may happen.
+Perhaps the underlying software used to perform the sampling and kalman filter is not suitable for this paramterisation, or perhaps more broadly the MCMC strategy does not perform well with this parameterisation of the model.
+More work is required to diagnose the cause of this problem.
+Other avenues to improve the sampling of this model could be to rewrite the sampler from scratch or consider a different sampling method.
+Note that this is only one kind of reparamterisation and further research could explore other reparamterisations, such as non centered in scale.
+
+    \subsection{Reweigthing rank statistics of Gaussian Mixture}
+    \citet{kim1998stochastic} correct for approximation error in their method by using importance sampling weights.
+The reweighting procedure ensures that samples are drawn from the correct posterior density.
+This reweighting step is applied to the calculation of the expected value of the posterior density to get a weighted mean estimate from the samples drawn from the Gaussian mixture approximation.
+The rank statistic can be written as a function of the expected value, so the reweighting step can be applied to the calculation of the rank statistics to see if this correction improves the calibration of the sampling approach.
+
+    Let $v(\theta, h)$ be the log weights defined as the log difference between the posterior densities of the true model with log chi squared errors $log\: g(\theta, h | y^{\ast}_t)$ and the Gaussian mixture model $log\:  k(\theta, h_t | y^{\ast}_t)$.
+
+    $$
+    \begin{aligned}
+        v(\theta, h) = log\: g(\theta, h | y^{\ast}_t) - log\:  k(\theta, h_t | y^{\ast}_t) = const + log\: g(y|h) - log\: k(y^{\ast} | h)
+    \end{aligned}
+    $$
+
+    Take the exponential and normalise the weights for the $l^{th}$ posterior draw (note the constants cancel out):
+
+    $$
+    \begin{aligned}
+    w^l = \frac{exp(v(\theta, h)_l)}{\sum_i exp(v(\theta, h)_i)}
+    \end{aligned}
+    $$
+
+    This gives the importance sampling weight. The expectation for any function of the posterior samples can be written as a function of the weights.
+
+    $$
+    \begin{aligned}
+    S(\theta) &= 1[\theta_l < \theta^{sim}] \\
+    E[S(\theta) | y] &= \int s(\theta) g(\theta | y) d\theta\\ 
+    &= \frac{\int s(\theta)\times exp(v(\theta, h)) * k(\theta, h_t | y^{\ast}_t)d\theta d h}{\int exp(v(\theta, h)) * k(\theta, h_t | y^{\ast}_t)d\theta d h} 
+    \end{aligned}
+    $$
+
+    Therefore, the expectation of the reweighted posterior samples is given by:
+
+    $$
+    \begin{aligned}
+    E[S(\hat{\theta}) | y^{\ast}] = \sum_l^L S(\theta_l)w_l
+    \end{aligned}
+    $$
+
+    The rank statistic can be rewritten as a function of the expectation and weights. This gives us the reweighted rank statistics.
+
+    $$
+    \begin{aligned}
+    r = \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}] \approx  L\times E[S(\hat{\theta})] = L\times \sum_l^L S(\theta_l)w_l
+    \end{aligned}
+    $$
+
+    The results from applying this reweighting step to the rank statistics are given on Figure \ref{fig:reweight1k}.
+There are no major improvements to the uniformity of rank statistics.
+Increasing the SBC iterations to 5000 as seen in Figure \ref{fig:reweight5k} also results in no major improvements.
+The shape of both sets of histograms is consistent with the shape of the unweighted rank statistics.
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/sir_cp_1k.png}
+        \caption{1000 SBC iterations for reweighted rank statistics from the Gaussian mixture approximation model. The shape of the histogram is consistent with the unweighted rank statistics.}
+        \label{fig:reweight1k}
+    \end{figure}
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/sir_cp_5k.png}
+        \caption{5000 SBC iterations for reweighted rank statistics from the Gaussian mixture approximation model. No major improvements are observed to the uniformity of rank statistics.}
+        \label{fig:reweight5k}
+    \end{figure}
+
+    \subsection{Evaluating all models and parameters}
+    As discussed, it is not practical to visually inspect the histograms of all parameters in the stochastic volatility model.
+Instead, the Chi squared statistic can be used to summarise the shape of the distribution and if desired, be used for hypothesis testing.
+However, isntead of performing hypothesis tests, another visual approach is taken to compare the calibration of the algorithms and parameters.
+    % However, performing all these tests suffers from the same overcommunication as presenting all the histograms at once.
+
+    The distribution of chi squared statistics across all simulations and all parameters are plotted as histograms on Figure \ref{fig:allchisq}.
+A chi squared statistic of 0 implies the sample is exactly uniformly distributed.
+Any value away from 0 captures any variation or deviation away from uniformity.
+This provides a high level view of all parammeters - how close they are to uniformity and how close or far away they are relative to the other SBC results for other algorithms and parameterisations.
+
     \begin{figure}[H]
@@
         \includegraphics[scale=0.1]{results/dist_chisq.png}
-        \caption{Distribution of chi squared statistics for all models.}
+        \caption{Distribution of chi squared statistics for all models (5000 SBC iterations).}
+        \label{fig:allchisq}
     \end{figure}
+
+    The distribution for HMC are clustered very closely to zero.
+Both results for the bespoke sampling method are distributed far from zero which suggests strong evidence of a lack of uniformity across all parameters.
+This is consistent with the inspected distriubtion of rank statistics in the previous sections.
 
@@
 
+- Fair comparison between 999 samples and 9,999 samples.
+
 \section{Conclusion}

### 2023-10-03 manuscript/main.tex
discussion section

@@
 The reweighting procedure ensures that samples are drawn from the correct posterior density.
-This reweighting step is applied to the calculation of the expected value of the posterior density to get a weighted mean estimate from the samples drawn from the Gaussian mixture approximation.
+This is applied to the calculation of the expected value of the posterior density to get a weighted mean estimate from the samples drawn from the Gaussian mixture approximation.
 The rank statistic can be written as a function of the expected value, so the reweighting step can be applied to the calculation of the rank statistics to see if this correction improves the calibration of the sampling approach.
@@
 Instead, the Chi squared statistic can be used to summarise the shape of the distribution and if desired, be used for hypothesis testing.
-However, isntead of performing hypothesis tests, another visual approach is taken to compare the calibration of the algorithms and parameters.
+However, instead of performing hypothesis tests, another visual approach is taken to compare the calibration of the algorithms and parameters.
     % However, performing all these tests suffers from the same overcommunication as presenting all the histograms at once.
@@
 \section{Discussion}
+Overall, Hamiltonian Monte Carlo applied on the reparameterised model performed best based on the calibration and efficiency evaluation criteria.
+Centered parameterisation using HMC performed moderately well but struggled to produce calibrated posterior estimates for $\sigma^2$ and $\phi$ as well as generate independent samples for these parameters.
+Gaussian mixture approximation struggled to return calibrated posterior estimates and efficient samples for both parameteristions.
+However, the centered parameteristion results may improve if we increased the length of the Markov chain.
+This is discussed later in the limitations section.
+
+These results demonstrate that the performance of a MCMC algorithm is sensitive to the paramterisation of the model in the context of calibration and efficiency.
+Furthermore, the paramterisation of a model is conditional on the choice of MCMC algorithm.
+Evaluating the effectiveness of a modelling strategy (i.e choice of model, paramterisation and sampler) on simulated data gives a lot of a priori information about how the model will perform before it is fit on real data.
+Performing SBC may help isolate confounding factors if any problems occur.
+
+The priors chosen for this study came directly from \citet{kim1998stochastic}.
+The SBC results for the Gaussian mixture may improve if the priors were more informative - although this is only conjecture.
+Whether or not to use tighter priors depends on the modelling context and the problem at hand.
+If the model is expected to handle datasets generated by a set of priors, then uncalibrated results from SBC raises questions about the suitability about the model choice, MCMC or parameterisation.
+
+Other parameterisaions of the stochastic volatility model are available as outlined in \citet{strickland2008parameterisation}, namely non centered in scale or both non centered in scale and location.
+The failure of non centered parameterisaion in location for the Gaussian mixture model could be due to a variety of reasons.
+This may be due to a particular software implementation when it comes to the Kalman Filter and Simulation Smoother (in which a specific software package was applied) that was not designed to handle the non centered parameterisation.
+Or perhaps this particular parameterisation doesn't suit the sampling strategy applied by KSC and another approach (e.g.
+HMC) might be more successful.
+Lots of different design and implementation details could be explored further.
+Pinpointing the reason why this model failed as well as exploring other configurations and MCMC algorithms in this context is left to future research.
+
+% (other implementations exist as well but left for future research. lots of different design and implementation details could be explored. Let alone choice of algos.)
 

### 2023-10-03 manuscript/main.tex
limitations and conclusion

@@
     \begin{aligned}
-    E[S(\hat{\theta}) | y^{\ast}] = \sum_l^L S(\theta_l)w_l
+    E[S(\hat{\theta}) | y^{\ast}] = \sum_l^L S(\hat{\theta}_l)w_l
     \end{aligned}
@@
 Both results for the bespoke sampling method are distributed far from zero which suggests strong evidence of a lack of uniformity across all parameters.
-This is consistent with the inspected distriubtion of rank statistics in the previous sections.
+This is consistent with the distribution of rank statistics in the previous sections.
 
@@
 Gaussian mixture approximation struggled to return calibrated posterior estimates and efficient samples for both parameteristions.
-However, the centered parameteristion results may improve if we increased the length of the Markov chain.
+However, the centered parameteristion results for the Gaussian mixture model may improve if we increased the length of the Markov chain.
 This is discussed later in the limitations section.
@@
 The failure of non centered parameterisaion in location for the Gaussian mixture model could be due to a variety of reasons.
-This may be due to a particular software implementation when it comes to the Kalman Filter and Simulation Smoother (in which a specific software package was applied) that was not designed to handle the non centered parameterisation.
+This may be due to a particular software implementation when it comes to the Kalman Filter and Simulation Smoother (in which a specific software package was applied) that was not designed to handle the non centered parameterisations.
 Or perhaps this particular parameterisation doesn't suit the sampling strategy applied by KSC and another approach (e.g.
@@
 Lots of different design and implementation details could be explored further.
-Pinpointing the reason why this model failed as well as exploring other configurations and MCMC algorithms in this context is left to future research.
+Pinpointing the reason why this model parameterisation failed as well as exploring other configurations and MCMC algorithms in this context is left to future research.
+
+It is worth noting that there are many ways of "comparing" MCMC approaches.
+Other features that may be of interest is the speed in which a MCMC can converge onto the target distribution and generate samples for inference.
+However, "speed" is diffiuclt to define and is subject to a variety of different factors such as hardware, operating system, etc.
+The purpose of this research is to compare algorithms based on calibration and efficiency.
+Additionally, this simulation study does not provide any advice on the appropriateness of the stochastic volatility model on real data.
+Whether the stochastic volatility model is appropriate for modelling a particular financial time series requires further study into the data generating process, domain expertise and a suite of other diagnostic checks.
+Examples of this are posterior predictive checking and out of sample predictive performance.
+Rather, ths SBC approach gives insight into how well calibrated a sampling strategy is conditional on a known data generating process which is important in all applied contexts.
 
@@
 
-\subsection{Limitations}
-- SBC is computationally expensive.
+\section{Limitations}
+A limitation to applying SBC in all modelling contexts is that it is computationally intensive.
+SBC requires fitting multiple models to get reliable estimates.
+In some cases, fitting just one model can take a long time, let alone multiple.
+Producing results for this research within a reasonable timeframe required use of a computer cluster.
+Computing infrastructure may limit the opportunity for SBC to be performed,although it is probably not necessary to do 5000 SBC iterations in other research contexts.
+The choice of SBC iterations for this research was arbitrary and was used to demonstrate that noisy estimates could be reduced by increasing the number of iterations for calibrated models.
+A more principlied approach to picking the number of SBC iterations could be explored in other research.
 
-- Datasets generated dependent on the priors. If priors are diffuse then can get a wide set of possible datasets which make it harder to get calibrated estimates of the true parameter.
+Developing diagnostic tools for evaluating SBC results and models with many parameters is an area for improvement.
+As discussed, inspecting many histograms for highly parameterised models is unrealistic.
+Other summary statistics and visualisations can be applied for a high level comparison, such as the chi squared statistics applied in this research.
+Furthermore, it is not always straightforward to understand why any any one parameter may produce poor calibration results, for example, an arbitrary state parameter.
+SBC may tell us some information about the miscalibration based on the histogram shape (e.g.
+under or over dispersed relative to the prior distribution), but understanding why this occurs may not be straight forward.
+Although, diagnosing sampling problems with complicated posterior geometries speaks more generally to the difficulty of understanding complex models in the first place as opposed to a limitation of SBC.
 
-- Fair comparison between 999 samples and 9,999 samples.
+Lastly, it is unclear what constitutes a "fair comparison" between algorithms when comparing SBC results.
+In particular, the number of post burnin or warmup samples differed between algorithms (999 post burn-in for HMC, 9999 for the bespoke algorithm).
+An argument could be made that the poor results for the bespoke algorithm may be due to the chain not converging onto the target distribution.
+Indeed, the authors of this approach extended their Markov chain to 750,000 posterior draws.
+A farier comparison may be to have roughly similar number of ESS for most of the estimated parameters, although this is difficult to get right with multiple parameters.
+However, given how well Hamiltoninan Monte Carlo performed with the reparamterised model, any improvement to the Gaussian mixture model by increasing the length of the Markov chain would make calibration just as good, but not better due to the inefficiency of the sampler.
 
 \section{Conclusion}
+This research evaluated and compared different MCMC algorithms for fitting stochastic volatility models.
+As our models and algorithms for estimating these models become more complicated, so does the need for principled ways of checking that they are returning correct posterior estimates.
+SBC provides a general simulation design for generative models to check whether a modelling strategy is producing the correct posterior estimates.
+
+In the context of stochastic volatility, applying Hamiltonian Monte Carlo on a reparameterised stochastic volatility model provided the most calibrated and efficient estimates when compared to the bespoke sampling method using a Gaussian mixture model.
+This also outperformed SBC and ESS results when using importance sampling weights to correct for approximation error.
+Other parameterisations of the stochastic volatility model can be considered to see if it improves sampling performance based on these metrics.
+
+This study only considered a handful of potential algorithms, sampling strategies and parameterisations for stochastic volatility.
+Future research could use the same SBC design to compare and evaluate more complicated volatility models (or any model outside this domain) and MCMC approaches.
+Overall, SBC can be applied as part of the development of various probabilistic and Bayesian models and is a useful tool as part of a statistical workflow.
 

### 2023-10-03 manuscript/results/rank_histograms_static_states.r
figs

@@ -34,18 +34,6 @@ hmc_ncp_onek_bin_long <- static_rank_bins(data = hmc_ncp_onek_bins,
 
 hmc_1k = rbind(hmc_cp_onek_bin_long, hmc_ncp_onek_bin_long)
 
-rank_hist <- function(dataframe){
-        ggplot(dataframe) +
-        geom_histogram(aes(rank), bins = 20, fill = "light blue", alpha = 0.7) +
-        scale_x_continuous(labels = function(x) x * (posterior_samples+1)/n_bins) +
-        facet_wrap(~factor(variable, levels = parameters), labeller = label_parsed) +
-        theme_minimal(base_size = 20) +
-        theme(strip.text.x = element_text(size = 22)) +
-        geom_hline(yintercept = expected_count, linewidth = 0.5, alpha = 0.3) +
-        labs(y = "Count",
-             x = "Ranks")
-}
-
 rank_hist(hmc_cp_onek_bin_long) # HMC CP 1k
 ggsave("manuscript/results/hmc_cp_1k.png", bg = "white", width = 14, height = 9.42)
 
@@ -76,10 +64,10 @@ hmc_ncp_fivek_bin_long <- static_rank_bins(data = hmc_ncp_fivek_bins,
                                           type = "NonCentered")
 
 # hmc_5k = rbind(hmc_cp_fivek_bin_long, hmc_ncp_fivek_bin_long)
-rank_hist(hmc_cp_fivek_bin_long) # HMC CP 1k
+rank_hist(hmc_cp_fivek_bin_long) # HMC NCP 1k
 ggsave("manuscript/results/hmc_cp_5k.png", bg = "white", width = 14, height = 9.42)
 
-rank_hist(hmc_ncp_fivek_bin_long) # HMC CP 1k
+rank_hist(hmc_ncp_fivek_bin_long) + ylim(0, 600) # HMC NCP 5k
 ggsave("manuscript/results/hmc_ncp_5k.png", bg = "white", width = 14, height = 9.42)
 
 # KSC 1k

### 2023-10-04 manuscript/main.tex
Rank hist examples

@@
 
-        \subsubsection{Rank statistics and chi squared statistics}
-            Rank statistics as described in the SBC section are used to evaluate the calibration of a posterior.
-Histograms will be used to evaluate the distribution of rank statistics.
-If a posterior is well calibrated then it is expected that the histogram is uniform.
+        \subsubsection{Rank statistics}
+            As discussed in the SBC section, rank statistics are used to evaluate the calibration of the MCMC algorithm.
+If a posterior is well calibrated then it is expected that the distribution of rank statistics is uniform.
 
-            A drawback of this approach is there are more parameters than data points in this model.
+            The shape of the rank statistic distribution gives insight into how a MCMC may be miscalibrated.
+Specifically, it gives information about the type of bias in the posterior estimates for a given parameter.
+Figures \ref{fig:overestimate}, \ref{fig:underestimation} and \ref{fig:underdispersed} gives some examples of the posterior estimates which result in bias of the rank statistics.
+
+            Figure \ref{fig:overestimate} shows a distribution of rank statistics with a large left peak.
+This is consistent with posterior samples overestimating the true parameter.
+The sum of indicator random variables is small since the majority of posterior draws are larger than the true value.
+This results in a small rank statistic.
+Hence, the shape of the histogram is consistent with multiple posterior estimates as part of the SBC procedure being biased towards overestimation of the true parameter.
+
+            Figure \ref{fig:underestimation} displays the converse behaviour. The large right peak is consistent with a bias towards under estimation of the true parameter.
+
+            Figure \ref{fig:underdispersed} has large peaks on both ends of the histogram.
+The sampler is disproportionately over and under estimating the true parameter resulting in biased rank statistics.
+\citet{talts2018validating} describe this bias as underdispersion of the posterior relative to the prior distribuion (since the posterior draws should follow the same distribution as the prior in this simulation design).
+The estimated posterior is too narrow relative to the spread of the prior resulting in bias on both ends of the rank statistics distribution.
+
+            \begin{figure}[H]
+                \centering
+                \includegraphics[scale=0.07]{methodology/lhs.png}
+                \caption{Left: Non uniform rank statistics with peak on the left side.
+Right: Posterior distribution vertical line representing true parameter.
+Bias in rank statistics distribution comes from disproportionate number of SBC iterations overestimating the true parameter.}
+                \label{fig:overestimate}
+            \end{figure}
+
+            \begin{figure}[H]
+                \centering
+                \includegraphics[scale=0.07]{methodology/rhs.png}
+                \caption{Left: Non uniform rank statistics with peak on the right side.
+Right: Posterior distribution vertical line representing true parameter.
+Bias in rank statistics distribution comes from disproportionate number of SBC iterations underestimating the true parameter.}
+                \label{fig:underestimation}
+            \end{figure}
+
+            \begin{figure}[H]
+                \centering
+                \includegraphics[scale=0.07]{methodology/underdispersed.png}
+                \caption{Left: Non uniform rank statistics with peaks on both ends.
+Right: Overlayed posterior and prior distributions for the same arbitrary parameter (taken from \citet{talts2018validating}).
+Bias in ranks comes from underdispersed posterior distribution relative to prior.}
+                \label{fig:underdispersed}
+            \end{figure}
+
+            \subsubsection{Chi squared statistics}
+            A drawback from visualising the rank statistic distributions is there are more parameters (1003 parameters) than data points in this model.
 An alternative to visually checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each histogram bin.

### 2023-10-04 manuscript/methodology/rank_stats_examples.r

@@ -1,13 +1,69 @@
 library(ggplot2)
+library(cowplot)
+
 df = as.data.frame(list(v1=rnorm(5e6)))
 
 ggplot(df, aes(x=v1))+ geom_density(linewidth=1.2) + 
     theme_void() +
     geom_vline(xintercept = -4, linewidth = 1)
 
+
+
+x <- seq(-5,5, 0.01)
+dens <- dnorm(x)
+df <- as.data.frame(cbind(x, dens))
+
+# LHS
+dens_lhs <- ggplot(df, aes(x = x, y = dens)) + 
+    geom_line() +
+    theme_void() +
+    geom_vline(xintercept = -4, linewidth = 1, colour = "red")
+
+
 hist_df = as.data.frame(list(v1 = c(rep(1,500), runif(3000, 2,20))))
-ggplot(hist_df, aes(x=v1)) +
-    geom_histogram(bins=20) +
+hist_lhs <- ggplot(hist_df, aes(x=v1)) +
+    geom_histogram(fill = "red", bins=20, alpha = 0.4) +
+    theme_void()
+
+plot_grid(hist_lhs, dens_lhs)
+ggsave("manuscript/methodology/lhs.png", bg = "white", width = 14, height = 8.42)
+
+# RHS
+dens_rhs <- ggplot(df, aes(x = x, y = dens)) + 
+    geom_line() +
+    theme_void() +
+    geom_vline(xintercept = 4, linewidth = 1, colour = "red")
+
+
+hist_df = as.data.frame(list(v1 = c(rep(20,500), runif(3000, 1,19))))
+hist_rhs <- ggplot(hist_df, aes(x=v1)) +
+    geom_histogram(fill = "red", bins=20, alpha = 0.4) +
+    theme_void()
+
+plot_grid(hist_rhs, dens_rhs)
+ggsave("manuscript/methodology/rhs.png", bg = "white", width = 14, height = 8.42)
+
+# Overdispersion
+x <- seq(-5,5, 0.01)
+post_draws <- dnorm(x, 0, 0.5)
+prior_draws <- dnorm(x,0, 2)
+
+df <- as.data.frame(cbind(x, post_draws, prior_draws))
+names(df) <- c("x", "density", "type")
+
+hist_df = as.data.frame(list(v1 = c(rep(1,500), rep(20,500), runif(3000, 1,19))))
+hist_lhs_rhs <- ggplot(hist_df, aes(x=v1)) +
+    geom_histogram(fill = "red", bins=20, alpha = 0.4) +
[…]

### 2023-10-05 manuscript/final_presentation/presentation.qmd
updated plots and presentation guideline

# Introduce stochastic volatility model
# Research question
# Motivating example 1 (two posterior estimates from two MCMC algorithms)
# Motivating example 2 (Limitations of single simulation)
# Methodology (SBC)
# What is HMC
# HMC results
# What was KSC approach 
# KSC results
# Reparameterisation (HMC)
# Reparam results 
# Reparameterisation (KSC)
# Reparam results (KSC)
# Importance sampling weights
# Discussion/limitations
# Future research

- Stochastic Volatility model (Do i need to motivate this like last time?)
- 2 Motivating examples. If two methods and they're different what do we do. 
- Research Question
- Methodology (how do we answer this. What iss a fair comparison? This is what we're going to do.)
- HMC and KSC algorithms
- Reparameterisation
- Importance sampling reweighting (cut this? Yes)
- Discussion and learnings
- Limitations
- Conclusion

2 motivating examples: Don't have to get into nuts and bolts of diff algos. High level MCMC summary. General structure + bespoke. Stan can flexibly fit any model. It's a programming language. Bespoke only for SV model and how to best sample SV.

In a simpler

 Start withmotivating example 1.
 Then go to movtivating exmaple 2. In this example.

 When talking about SBC: remember motivating example, link it back. Thread the example through the presentation. Could even do distributions of posteriors vs true posteriors. Show results and link it back to the motivating example

In presentation use talts copy the plots. Mkae sure cites.
Reproduce Talts data viz. Replicating examples from this paper are shown in these figures.

Presentation: show results -> then show fig 4-6. Just say for convenience.

Include a summary slide of extensions without going into detail. Give a sense of what else is in the paper. We explored different parameteirsations.

IS: Then they do an Importance Sampling. Wish i could talk about it but no time.

Audience: Give them the impression that you done more than talked about. Made careful decision about what was more important to present. Lot's of content - but you need to understand what's int he presetation first.Done more than minimum.

Reparam: Don't go through steps or why they exist. There is a literature about MCMC efficiency if you reparam these models you can gain MCMC efficiency. Put the citations. There's lit but param is an issue. We look at a couple and these are results but don't go into detail.

Make sure i explicit: I'm really interested in finding out about xyz in context abc. Rsearch is comparing ABC and XyZ. State it early on. Say close to methods? So they can mark you.

In motivating example: say a little bit about why people are interested in this model. This is a difficult model. Ppl are interested GARCH are easy to estimate but don't capture properties of data observwd.

definitely harder on complicated model (not as intresting on simple model). Why we're looking at SV model in the first place. How to fit this model fast. **Check if its correct**. For this SV model in parrticular. Motivate this point. It's a hard model to fit.

It could've gone either way. Might not have gone well stan. **Think about what a fair comparison is**

Particle filtering. There are other approaches. PF within MCMC.

Jump straight into motivating examples. If you're Bayes, -> 2 different approx solutions. So we run both of them. Kind of similar but also different. pointness and fantess. How do we know which one to use.

Then go into SV. And we use it to do this. This is the model. Trying to work out the difference.

Get to movating example as soon as i can.

Start: What the example is. In title slide: prelim, title supervisors, slide gives a rough outline. Main sections of preso to set everyone. 
Suppose i'm an analyst and I need to fit this model and estimate variance. Realise there's more than one way to do that. Check both and this is what i get. Then build onto SBC. Then confirm method.

At a higher level talk about the model (not all priors etc). Just cite the literature on the slide. This is from KSC (cite earlier shephard paper). People have worked with it. Estimate, lots of parameters hard to estimate. Some methods exist here are two of them. Talk in detail later. Reassure them what the thing is. THis is the output.

Side by side histograms then full histgoram

### 2023-10-05 manuscript/main.tex
updated plots and presentation guideline

@@
     \subsection{Evaluation Metrics}
-        The key metrics and diagnostics to compare the performance of these methods are the effective sample size (ESS), rank statistics, and chi-squared test statistics.
-
-        \subsubsection{Effective sample size}
-            ESS measures the efficiency of the MCMC sampler.
-It calculates the number of effectively independent draws from the posterior draws generated by a Markov chain.
-A poor ESS can arise from high autocorrelation in the Markov chain which leads to highly dependent samples.
-An efficient MCMC algorithm takes less resources (for example, time and number of draws) to get a representative sample of the target distribution.
-If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), then there is evidence that this method is a more efficient sampler.
+        The key metrics and diagnostics to compare the performance of these methods are rank statistics, and chi-squared test statistics to measure calibration and the effective sample size (ESS) to measure effciency.
 
@@
 Specifically, it gives information about the type of bias in the posterior estimates for a given parameter.
-Figures \ref{fig:overestimate}, \ref{fig:underestimation} and \ref{fig:underdispersed} gives some examples of the posterior estimates which result in bias of the rank statistics.
+Figures \ref{fig:underestimation} and \ref{fig:underdispersed} are recreated from \citet{talts2018validating} for conveience which gives some examples of the posterior estimates which result in bias of the rank statistics.
 
-            Figure \ref{fig:overestimate} shows a distribution of rank statistics with a large left peak.
-This is consistent with posterior samples overestimating the true parameter.
-The sum of indicator random variables is small since the majority of posterior draws are larger than the true value.
-This results in a small rank statistic.
-Hence, the shape of the histogram is consistent with multiple posterior estimates as part of the SBC procedure being biased towards overestimation of the true parameter.
-
-            Figure \ref{fig:underestimation} displays the converse behaviour. The large right peak is consistent with a bias towards under estimation of the true parameter.
+            Figure \ref{fig:underestimation} shows a distribution of rank statistics with a large right peak.
+This is consistent with posterior samples underestimating the true parameter, where the estimated posterior distrubtion is biased to the left of the prior.
+The sum of indicator random variables is large since the majority of posterior draws are smaller than the true value.
+This results in a large rank statistic and a bias on the right side of the histogram.
+The coverse is true if the peak of the histogram is on the left hand side (i.e the posterior is biased towards the right hand side of the prior).
 
             Figure \ref{fig:underdispersed} has large peaks on both ends of the histogram.
-The sampler is disproportionately over and under estimating the true parameter resulting in biased rank statistics.
-\citet{talts2018validating} describe this bias as underdispersion of the posterior relative to the prior distribuion (since the posterior draws should follow the same distribution as the prior in this simulation design).
+The sampler is disproportionately over and under estimating the true parameter.
+\citet{talts2018validating} describe this bias as underdispersion of the posterior relative to the prior distribuion.
 The estimated posterior is too narrow relative to the spread of the prior resulting in bias on both ends of the rank statistics distribution.
 
-            \begin{figure}[H]
-                \centering
-                \includegraphics[scale=0.07]{methodology/lhs.png}
-                \caption{Left: Non uniform rank statistics with peak on the left side.
+            % \begin{figure}[H]
+            %     \centering
+            %     \includegraphics[scale=0.07]{methodology/lhs.png}
+            %     \caption{Left: Non uniform rank statistics with peak on the left side.
 Right: Posterior distribution vertical line representing true parameter.
 Bias in rank statistics distribution comes from disproportionate number of SBC iterations overestimating the true parameter.}
-                \label{fig:overestimate}
-            \end{figure}
+            %     \label{fig:overestimate}
+            % \end{figure}
 
@@
                 \caption{Left: Non uniform rank statistics with peaks on both ends.
-Right: Overlayed posterior and prior distributions for the same arbitrary parameter (taken from \citet{talts2018validating}).
+Right: Overlayed posterior and prior distributions for the same arbitrary parameter.
 Bias in ranks comes from underdispersed posterior distribution relative to prior.}
@@
 This will give a high level summary of calibration and overall performance of the algorithms.
+
+            \subsubsection{Effective sample size}
+            ESS measures the efficiency of the MCMC sampler.
+It calculates the number of effectively independent draws generated by a MCMC to estimate a target parameter.
+A poor ESS can arise from high autocorrelation in the Markov chain which leads to highly dependent samples.
+An efficient MCMC algorithm takes less resources (for example, time and number of draws) to get a representative sample of the target distribution.
+If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), then there is evidence that this method is a more efficient sampler.
+
+            The ESS is defined as $n_{eff}$ and is a function of the number of chains $M$, the number of samples per chain $N$ and the autocorrelation at lag t $\rho_t$.
+$T$ is a trunction lag chosen such that the sum of any two consecutive autocorrelation lags is negative.
+This truncation (or some kind of downweighting) is required because a large number of lags leads to noiser autocorrelation estimates.
+More details about this and the calculation of the autocorrelation can be found in \citep{vehtari2021rank} and \citet{geyer1992practical}.
+
+            $$
+            \begin{aligned}
+                n_{eff} = \frac{NM}{1+2 \sum_{t=1}^T \hat{\rho}_t}
+            \end{aligned}
+            $$
 

### 2023-10-05 manuscript/methodology/rank_stats_examples.r

@@ -8,16 +8,46 @@ ggplot(df, aes(x=v1))+ geom_density(linewidth=1.2) +
     geom_vline(xintercept = -4, linewidth = 1)
 
 
-
+# Uniform
 x <- seq(-5,5, 0.01)
-dens <- dnorm(x)
-df <- as.data.frame(cbind(x, dens))
+prior_draws <- dnorm(x,0, 2)
+
+df <- as.data.frame(cbind(x, prior_draws))
+names(df) <- c("x", "density")
+
+hist_df = as.data.frame(list(v1 = c(rep(seq(1,20),200))))
+hist_unif <- ggplot(hist_df, aes(x=v1)) +
+    geom_histogram(fill = "red", bins=20, alpha = 0.4, colour = "white") +
+    theme_void() +
+    scale_y_continuous(expand = c(0,0.5))
+
+
+dens_both <- ggplot(df, aes(x = x)) +
+    geom_line(aes(y = post_draws), colour = "blue") + 
+    geom_line(aes(y = prior_draws), colour = 'red') +
+    theme_void() +
+    annotate("text", x = 4.5, y = 0.05, label = "Prior", colour = "red", size = 8) +
+    annotate("text", x = 2.5, y = 0.3, label = "Posterior", colour = "blue", size = 8)
+
+plot_grid(hist_lhs_rhs, dens_both)
+
+ggsave("manuscript/methodology/underdispersed.png", bg = "white", width = 12, height = 8.42)
+
 
 # LHS
-dens_lhs <- ggplot(df, aes(x = x, y = dens)) + 
-    geom_line() +
+x <- seq(-8,4, 0.01)
+post_draws <- dnorm(x, -5, 1)
+prior_draws <- dnorm(x,0, 1)
+
+df <- as.data.frame(cbind(x, post_draws, prior_draws))
+names(df) <- c("x", "density", "type")
+
+dens_lhs<- ggplot(df, aes(x = x)) +
+    geom_line(aes(y = post_draws), colour = "blue") + 
+    geom_line(aes(y = prior_draws), colour = 'red') +
     theme_void() +
-    geom_vline(xintercept = -4, linewidth = 1, colour = "red")
+    annotate("text", x = 0, y = 0.15, label = "Posterior", colour = "red", size = 8) +
+    annotate("text", x = -5, y = 0.15, label = "Prior", colour = "blue", size = 8)
 
 
[…]

### 2023-10-06 manuscript/main.tex
hmc

@@
 
-        The key innovation of Hamiltonian Monte Carlo is using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
-Unlike Random Walk Metropolis Hastings, it takes advantage of the geometry of the posterior to determine its next proposal step.
-A comprehensive explanation of the sampler is beyond the scope of this research and can be found in the above references.
+        The key innovation of Hamiltonian Monte Carlo (HMC) is using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
+HMC is able to reach more distant points with higher acceptance probabilites since proposals are made with information about the target distribution.
+Random walk beahviour is more inefficient in higher dimensions since it is more difficult to randomly select a point with a higher joint density and therefore higher acceptance probability.
+The increasing complexity of high dimensions make it harder for a random walk sampler to explore regions over long distances, resulting in the sampler to become stuck and requiring a lot of iterations to converge onto the target distribution.
+
+        A thorough conceptual and theoretical explanation of HMC can be found in BDA3 and Betancourt with further details found in the aforementioned literature.
+In summary, HMC builds upon the Metropolis Hastings algorithm through the introduction of Hamiltonian equations, the momentum variable $\phi$ and the gradients of the target log posterior distribtuion.
+The vector of model parameters $\theta$ is jointly sampled with a vector of momentum variables $\phi$.
+The momentum variables is also called auxiliary variables since they are required as part of the Hamiltonian equations to generate an efficient proposal but are not a quantity of interest for inference.
+
+        The sampled momentum variable sets out the proposal path for $\theta$ governed by the Hamiltonian dynamics (represented by a set of differential equations).
+These differential equations are solved using a discrete approximation algorithm, such as a leapfrog integrator, and is a function of the log posterior gradients for a given point $\theta$.
+The leapfrog integrator has a set of tuning parameters which determines the size and number of steps with the final step being the proposal value $(\theta^{\ast}, \phi^{\ast})$.
+Finally, the Metropolis Hastings accept/reject step is applied using the parameters at the beginning of the leapfrog process $(\theta^{t-1}, \phi^{t-1})$ and the proposal values $(\theta^{\ast}, \phi^{\ast})$.
 
@@
 This approach will also use the same priors as specified in the Gaussian mixture approximation.
+
+        \subsubsection{Algorithmic steps}
+        Let $\theta$ be a target parameter and $\phi$ be the auxiliary momemtum variable. The Hamiltonian is defined as:
+
+        $$
+        \begin{aligned}
+        H(\theta, \phi) \equiv - log \: \pi(\theta, \phi)
+        \end{aligned}
+        $$
+
+        This can be broken down into kinetic energy $K(\theta, \phi)$ (the density over the auxiliary momentum) and potential energy $V(\theta)$ (the density of the target distribution).
+
+        $$
+        \begin{aligned}
+        H(\theta, \phi) &= - log \: \pi(\phi | \theta) - log \: \pi(\theta) \\ 
+        &\equiv  K(\theta, \phi) + V(\theta)
+        \end{aligned}
+        $$
+
+         Starting with an initial draw of $\theta$ (user defined or randomly generated), the HMC iteration (or proposal path) is determined determined by 3 steps.
+
+        1) Randomly sample a momentum value
+
+        $$
+        \begin{aligned}
+        \phi\sim Multinormal(0, M)
+        \end{aligned}
+        $$
+
+        Where M is some diagnoal mass matrix (assuming independence between momentum variables).
+
+        2) Solve the set of Hamiltonian (differential) equations which generates a proposal $(\theta^{\ast}, \phi^{\ast})$
+
+        $$
+        \begin{aligned}
+        \frac{\partial \theta}{\partial t} &= \frac{\partial H}{\partial \phi} = \frac{\partial K}{\partial \phi} \\
+        \frac{\partial \phi}{\partial t} &= \frac{\partial H}{\partial \theta} = - \frac{\partial K}{\partial \theta} - \frac{\partial V}        {\partial \theta}
+        \end{aligned}
+        $$
+
+        Where $\frac{\partial V}{\partial \theta}$ is the gradient of the target density.
+Solving these differential equations using a Leapfrog integrator which gives a discrete approximate solution.
+The parameters which are tuned in this process are the number of steps $L$ and the size of the steps $\epsilon$.
+The Leapfrog algorithm performs a half step of $\phi$ and then a full step of the parameter $\theta$, and another half step of the momemtum variable $\phi$ and repeats this $L$ times.
+The final step is the proposal $(\theta^{\ast}, \phi^{\ast})$.
+
+        $$
+        \begin{aligned}
+        \phi &\leftarrow \phi - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta} \\
+        \theta &\leftarrow \theta + \epsilon M^{-1} \phi \\
+        \phi &\leftarrow \phi - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta}
+        \end{aligned}
+        $$
+
+        3) Metrpolis Hastings Accept/Reject
+        Let $(\theta^{t-1}, \phi^{t-1})$ be the values before the Leapfrog integrator.
+
+        $$
+        \begin{aligned}
+        r = \frac{\pi(\theta^{\ast} | y) \pi(\phi^{\ast})}{\pi(\theta^{t-1} | y) \pi(\phi^{t-1})}
+        \end{aligned}
+        $$
+
+        With sampled value
+
+        $$
+        \theta^t = \begin{cases}
+            \theta^{\ast},& \text{with probability } min(r,1)\\
+            \theta^{t-1}, & \text{otherwise}
+        \end{cases}
+        $$
 

### 2023-10-06 manuscript/main.tex
updated reparam and mcmc

@@
 
-        A thorough conceptual and theoretical explanation of HMC can be found in BDA3 and Betancourt with further details found in the aforementioned literature.
+        A thorough conceptual and theoretical explanation of HMC can be found in \citet{gelman2013bayesian} and \citet{betancourt2017conceptual} with further details found in the aforementioned literature.
 In summary, HMC builds upon the Metropolis Hastings algorithm through the introduction of Hamiltonian equations, the momentum variable $\phi$ and the gradients of the target log posterior distribtuion.
@@
 
-        This can be broken down into kinetic energy $K(\theta, \phi)$ (the density over the auxiliary momentum) and potential energy $V(\theta)$ (the density of the target distribution).
+        This can be broken down into kinetic energy $K(\theta, \phi)$, the density over the auxiliary momentum, and potential energy $V(\theta)$, the density of the target posterior distribution.
 
@@
 
-         Starting with an initial draw of $\theta$ (user defined or randomly generated), the HMC iteration (or proposal path) is determined determined by 3 steps.
+         Starting with an initial draw of $\theta$ (user defined or randomly generated), the HMC iteration or proposal is determined by 3 steps.
 
@@
 
-        Where $\frac{\partial V}{\partial \theta}$ is the gradient of the target density.
-Solving these differential equations using a Leapfrog integrator which gives a discrete approximate solution.
+        Where $\frac{\partial V}{\partial \theta}$ is the gradient of the target posterior density.
+The solution to these differential equations is approximated using a Leapfrog integrator which gives a discrete approximate solution.
 The parameters which are tuned in this process are the number of steps $L$ and the size of the steps $\epsilon$.
@@
 
+    \subsection{MCMC methods comparison}
+        The KSC sampling method and HMC are both MCMC strategies to genereate samples from the target joint posterior distribution of the stochastic volatility model.
+However, the two sampling approaches are conceptually different in their design and implementation, despite sharing the same objective.
+
+        The KSC sampling method is a bespoke algorithm designed specifically for the sampling of the stochastic volatility model.
+It is a strategy built around how to sample the parameters from the state space model in this specific context.
+The implementation of this MCMC requires careful attention to the sampling of each parameter using different tools and requires these components to be programmed manually, such as the Metropolis Hastings step and conjugate posterior distributions.
+
+        HMC on the other hand is a more general approach to applying MCMC.
+The implementation of HMC is using the Stan programming language which allows for the specfication other generative or Bayesian models, not just stochastic volatility.
+The implementation details of HMC is abstracted away from the user, who only needs to focus on specifying the model using the Stan syntax.
+Any hyperparameter tuning and initialisation is automated away.
+Furthermore, Stan and HMC allows for the specification of different priors, whereas KSC's bespoke algorithm is dependent on the use of conjugate priors.
+
+        The conceptual difference in both methods result in different tradeoffs.
+The abstraction of the MCMC implementation using the Stan programming language means users can focus more on the modelling and the applied research question.
+The algorithm is maintained by other developers and different models can be compared with each other without worrying about differences in MCMC implementations.
+However, the abstraction away from the details means there is less flexibility in MCMC design in the scenario where the current instance of Stan or HMC is insufficient for a specific model.
+
     \subsection{Model Parameterisation}
@@
 An example of this is Neal's funnel \citep{neal2003slice} where the Hamiltonian Monte Carlo sampler encounters performance issues in hierarchical models and produces biased samples.
-Efficiency improvements of MCMC algorithms for state space models under different parameterisations are also explored \citet{strickland2008parameterisation}.
+Efficiency improvements of MCMC algorithms for state space models under different parameterisations are also explored in \citet{strickland2008parameterisation}.
 
@@
 This describes the central location of the latent state parameters which are centered around the mean and lag of the log volatility $\mu +\phi(h_t - \mu)$.
-A model can be reparameterised such that the states are sampled on a distribution centered on 0 (i.e "non centered") which are later transformed to have the correct mean.
-The following re-parameterisations are explored as part of the simulation study.
+A model can be reparameterised in a way that improves MCMC performance.
+Reparameterisation applies a transformation to the model which may change the interpretation of the parameters or the way they are sampled but remains mathematically equivalent to the original model.
+The reparametiersations available for a model is dependent on the MCMC approach.
+Some reparamterisations are available only to specific sampling strategies and may improve or impair a MCMC performance.
 
-        \subsubsection{Non centered Hamiltonian Monte Carlo}
+        SBC is applied to reparameteirsied versions of the stochastic volatility model available to see if there is any improvement in calibration and efficiency.
+
+        \subsubsection{Reparameterised stochastic volatility for Hamiltonian Monte Carlo}
+        The reparameterisation for Hamiltonian Monte Carlo changes how the latent states are sampled.
+Specifically, the states are first sampled from a standard normal distribution and are transformed to be samples from the state equation.
+As noted above, the centered parametersation samples the states directly on a non zero mean.
+
         First sample from a standard normal distribuition and multiply by the variance of the log volatility.
-The state variable is sampled with mean centered on zero and variance equal to $\sigma_{\eta}^2$
+The state vector is sampled with mean centered on zero and variance equal to $\sigma_{\eta}^2$
 
@@
 
-        \subsubsection{Non centered Gaussian Mixture}
-        The non centered model for the Gaussian approximation is expressed slightly differently due to the use of the Kalman Filter.
-Non centered HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates.
+        \subsubsection{Non centered Gaussian Mixture model}
+        The reparameterised model for the Gaussian approximation is expressed differently due to the use of the Kalman Filter.
+Reparameterised HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates.
 Applying the Kalman Filter requires the state equation to be a function of the state variable with the average log volatility parameter $\mu$ entering the measurement equation.
+
+        The following parameterisation is described as non centered in location. This follows the methodology outlined in \citet{strickland2008parameterisation}.
 
@@
     \subsubsection{Prior Predictive check}
-        It is worth noting that the centered and non-centered parameteristions are the same stochastic volatility model.
-Different parameteristions express the same mathematical models in ways that make it easier or harder for a given algorithm to sample.
+        % Different parameteristions express the same mathematical models in ways that make it easier or harder for a given algorithm to sample.
 
-        To demonstrate this, a prior predictive simulation of the second state parameter is performed for both a centered and non centered model (using the paramterisation described in section 3.3.1).
+        A prior predictive simulation of the second state parameter is performed for both a centered and reparameterised model using the approach described in section 3.3.1.
+This is to demonstrate that different parameteristions express the same underlying mathematical model.
 This is the same as the first two steps of SBC, except instead of generating a dataset, generate the parameters implied by the prior draw.
@@
 
-        Generating 1000 samples of $h_2^{sim}$ from both centered and non-centered models gives samples from the same data generating process as shown in figure 3.
+        Generating 1000 samples of $h_2^{sim}$ from both centered and non-centered models gives samples from the same data generating process as shown in Figure \ref{fig:priorpred}.
 
@@
             \caption{Prior predictive samples of second state variable}
+            \label{fig:priorpred}
         \end{figure}

### 2023-10-07 manuscript/results/rank_histograms_static_states.r
ksc ncp plots

@@ -81,13 +81,25 @@ ksc_cp_onek_bins <- rank_bins_f(path = "simulation_output/ksc/cp/sbc_cp_ksc_mode
                                     posterior_samples = posterior_samples, 
                                     n_bins= n_bins)
 
+
+ksc_ncp_onek_bins <- rank_bins_f(path = "simulation_output/ksc/ncp/sbc_ncp_ksc_model_ncp_dgf_10kmcmc_1000iter_r1", 
+                                    rank_type = "agg_ranks", 
+                                    posterior_samples = posterior_samples, 
+                                    n_bins= n_bins)
+
 ksc_cp_onek_bins_long <- static_rank_bins(data = ksc_cp_onek_bins, 
                                          variables = parameters, 
                                          type = "Centered")
 
+ksc_ncp_onek_bins_long <- static_rank_bins(data = ksc_ncp_onek_bins, 
+                                         variables = parameters, 
+                                         type = "Centered")
 
 rank_hist(ksc_cp_onek_bins_long) # KSC CP 1k
 ggsave("manuscript/results/ksc_cp_1k.png", bg = "white", width = 14, height = 9.42)
+
+rank_hist(ksc_ncp_onek_bins_long) # KSC NCP 1k
+ggsave("manuscript/results/ksc_ncp_1k.png", bg = "white", width = 14, height = 9.42)
 
 # KSC 5k
 n_iterations <- 5000
@@ -100,13 +112,24 @@ ksc_cp_fivek_bins <- rank_bins_f(path = "simulation_output/ksc/cp/sbc_cp_ksc_mod
                                     posterior_samples = posterior_samples, 
                                     n_bins= n_bins)
 
+ksc_ncp_fivek_bins <- rank_bins_f(path = "simulation_output/ksc/ncp/sbc_ncp_ksc_model_ncp_dgf_10kmcmc_5000iter_r1", 
+                                    rank_type = "agg_ranks", 
+                                    posterior_samples = posterior_samples, 
+                                    n_bins= n_bins)
+
 ksc_cp_fivek_bins_long <- static_rank_bins(data = ksc_cp_fivek_bins, 
                                          variables = parameters, 
                                          type = "Centered")
 
+ksc_ncp_fivek_bins_long <- static_rank_bins(data = ksc_ncp_fivek_bins, 
+                                         variables = parameters, 
[…]

### 2023-10-07 manuscript/main.tex
updated ncp results

@@
 \section{Results}
-    1000 and 5000 SBC iterations are run for centered and reparameterised versions of the stocashtic volatility model.
-Rank statistics histograms are presented for the static parameters and the 1st, 500th and 1000th latent state parameters with 20 bins constructed for each parameter.
+    1000 and 5000 SBC iterations are run for centered and reparameterised versions of the stochastic volatility model.
+Rank statistic histograms are constructed with 20 bins for the static parameters and the 1st, 500th and 1000th latent state variables.
 The black horizontal line represents uniformity conditional on the number of bins and SBC iterations.
@@
     The centered parameterisation in particular displays a lack of uniformity for $\phi$ and $\sigma^2$.
-Rank statistics for $\sigma^2$ show two peaks of either end of the histogram and a left peak for $phi$.
-Posterior estimates of $\sigma^2$ are \textit{underdispersed} relative to the true parameter $\sigma^2_{sim}$ generated by the prior distribution.
-On the other hand, the posterior samples of $\phi$ are \textit{over estimating} the true parameter generated by the prior.
-This suggests the posterior geometry for these parameters (at least) are difficult to sample for the HMC algorithm.
+Rank statistics for $\sigma^2$ and $\phi$ show two peaks of either end of the histogram.
+The posterior estimates of $\sigma^2$ and $\phi$ are \textit{underdispersed} relative to the prior distribution.
+This implies the posterior geometry of these parameters may be difficult to sample for the HMC algorithm.
 
@@
     Estimates of $\sigma^2$ and $\phi$ are more uniform under the reparameterised model.
-Instead, we observe a peak on the left side for the 500th state parameter $h_{500}$.
+Instead, we observe a peak on the left side for the 500th state variable $h_{500}$.
 Overall the estimates for non centered parameterisation are more favourable.
@@
 The rank statistic distributions for $\sigma^2$ and $\phi$ have improved.
-There is some evidence of over estimation of the 500th state parameter.}
+There is some evidence that the posterior of the 500th state variable is overestimating the distribution of its true value.}
         \label{fig:ncphmc1k}
@@
 
-    Centered parameteristion posess similar improvement for $\mu$ and the state parameters.
+    Centered parameteristion posess similar improvement for $\mu$ and the state variables.
 However, there are still major peaks at both ends of $\sigma^2$ and $\phi$ suggesting underdispersion of the posterior estimates relative to the prior distribution.
 
-    These results suggests that HMC overall struggles to sample from the centered parameteristion of the stochastic volatility model.
-The reparameterised model results in much better and consistent inference for this sampling strategy, conditional on the model.
+    These results suggests that HMC struggles to sample some parameters from the centered parameteristion of the stochastic volatility model.
+The reparameterised model possess stronger evidence of calibration for the parameters examined, implying more reliable and consistent inference.
 
@@
 
-    Further to calibration, the HMC algorithm struggles to efficiently generate posterior samples from the centered parameterisation of stochastic volatility.
+    Additionally, HMC struggles to generate efficient posterior samples from the centered parameterisation of the stochastic volatility model.
 Figure \ref{fig:hmcess} visualises the distribution of ESS for the static parameters after 5000 SBC iterations.
@@
     \subsection{Gaussian Mixture Approximation}
-    The Gaussian Mixture Approximation is run for 1000 and 5000 SBC iterations with the first 10,000 samples discarded as burnin (consistent with \citet{kim1998stochastic}).
+    The Gaussian Mixture Approximation is run for 1000 and 5000 SBC iterations with the first 10,000 samples discarded as burnin (following the method of \citet{kim1998stochastic}).
 However, unlike KSC who take 750,000 posterior draws from the model, the number of posterior samples taken as part of the SBC procedure is 9,999.
-Taking 750,000 draws from the Gaussian mixture approximation would lead to issues given the disk space and compute time required from run SBC.
+Taking 750,000 draws from the Gaussian mixture approximation would lead to issues due to disk space limitations and computational resources required to run SBC.
 Furthermore, the number of posterior samples is larger than the 999 taken for the SBC experiment run for Hamiltonian Monte Carlo.
@@
 
-    The centered Gaussian mixture approximation contains issues with calibration.
-Figure \ref{fig:cpksc1k} presents the results from SBC where there is a lack of uniformity across all parameters.
+    The both the centered and reparameterised Gaussian mixture model contains issues with calibration.
+Figure \ref{fig:cpksc1k} presents the results from the centered model.
+There is a lack of uniformity across all parameters.
 In particular, there is a large peak on the left hand side for $\mu$ and the right hand hand side of $\phi$.
-This suggests over estimation and under estimation of the true parameter respectively.
-All other parameters do not have any key features worth noting beyond noisy estimates around the true discrete uniform value.
+This suggests over estimation and under estimation of the prior distribution respectively.
+All other parameters do not have any key features worth noting beyond noisy estimates around the discrete uniform distribution.
 
@@
 
-    Increasing the SBC iterations to 5000 does not improve the results.
-$\sigma^2$ and the state parameters contain arguably less noisy estimates around the uniform distribution.
+    The non centered model in location exhibits larger issues in calibration as shown in Figure \ref{fig:ncpksc1k}.
+Each histogram exhibits major peaks at one or both ends of the histgoram suggesting major issues in the sampler returning correct posterior estimates.
+The lack of uniformity for this model is distinctly due to a different shape.
+Whereas the lack of uniformity for previous SBC results with 1000 iterations may be due to noisy estimates.
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/ksc_ncp_1k.png}
+        \caption{1000 SBC iterations for non centered in location Gaussian mixture approximation model.
+The rank statistics for all selected parameters are non uniform in shape.
+The bespoke MCMC struggles to return the correct posteriors for this parametersation of the model.}
+        \label{fig:ncpksc1k}
+    \end{figure}
+
+    Increasing the SBC iterations to 5000 does not improve the results for either parametersation.
+$\sigma^2$ and the state parameters contain arguably less noisy estimates around the uniform distribution in the centered parametersation.
 However, the skewed shapes of $\mu$ and $\phi$ adds further evidence that the sampling strategy is not returning the correct posterior estimates.
@@
 
+    Figure \ref{fig:ncpksc5k} shows no improvement to calibration results.
+The issues in calibration observed in Figure \ref{fig:ncpksc1k} are not due to noisy estimates and are likely the result of problems in the bespoke MCMC algorithm and model specification.
+Applying this parametersation and MCMC on real data will on balance produce incorrect posterior estimates.
+
     \begin{figure}[H]
         \centering
-        \includegraphics[scale=0.1]{results/ksc_cp_5k.png}
+        \includegraphics[scale=0.09]{results/ksc_cp_5k.png}
         \caption{5000 SBC iterations for centered Gaussian mixture approximation model.
 Estimates for $\sigma^2$ and the state parameters have marginally improved around the uniformity.
-However, there is no improvement in $\mu$ or $\phi$, suggesting a lack of calibration from the MCMC algorithm.}
+However, there is no improvement in $\mu$ or $\phi$.}
         \label{fig:cpksc5k}
@@
 
-    Furthermore, the ESS estimates for this model indicate a high degree of autocorrelation and difficulty in generating independent samples from the posterior distribution.
-The median ESS for $\mu$, $\phi$ and $\sigma^2$ are 559, 103, 39.5 respectively for 10,000 post burn in draws.
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.09]{results/ksc_ncp_5k.png}
+        \caption{5000 SBC iterations for non centered Gaussian mixture approximation model.
+No improvement is observed in the distribution of rank statistics.
+The MCMC returns uncalibrated posterior estimates for this parametersation of the model.}
+        \label{fig:ncpksc5k}
+    \end{figure}
+
+    The ESS estimates for this model indicate a high degree of autocorrelation and difficulty in generating independent samples from both parameterisations.
+The numerical results are summarised in table x.
+The estimate for $\mu$ in the reparameterised model exhibits a long right tail (longer than the 10,000 posterior draws suggesting the presence of negatively autocorrelated draws).
+Overall, the bespoke MCMC is quite insufficient at sampling the static parameters.
+
+    % The median ESS for $\mu$, $\phi$ and $\sigma^2$ are 559, 103, 39.5 respectively for 10,000 post burn in draws.
 Despite having 10 times the number of draws, the number of effectively independent samples is much smaller relative to HMC, suggesting that this bespoke sampling strategy is highly inefficient.
+
+    % - reparam has massive long tail for mu
 
@@
 
-    The reparameterised model using the bespoke sampling strategy fails to generate sufficient results due to numerical errors and highly nonstationary chains.
-Figure X shows the traceplot from one simulation from the reparamterised model.
-The chain for $\phi$ is stuck to a value very close to 1 (unit root) and $\mu$ and $\sigma^2$ continuously increases in value until a numerical error is reached.
-Note that this occurs during the burnin phase of the markov chain.
-As a result, SBC is unable to be performed on this parameterisation of the model.
+    Overall the results from the KSC algorithm indicates some issues with calibration.
+The centered parameterastion appears to be more favourable for this sampling approach.
+Other avenues to improve the sampling of this model could be to consider a different bespoke sampling method or another parameterisation such as non centered in scale.
 
-    There are a variety of reasons why this may happen.
-Perhaps the underlying software used to perform the sampling and kalman filter is not suitable for this paramterisation, or perhaps more broadly the MCMC strategy does not perform well with this parameterisation of the model.
-More work is required to diagnose the cause of this problem.
-Other avenues to improve the sampling of this model could be to rewrite the sampler from scratch or consider a different sampling method.
-Note that this is only one kind of reparamterisation and further research could explore other reparamterisations, such as non centered in scale.
+    There are a few potential reasons for the poor calibration results.
+It may be due to the inefficiency of the MCMC to generate the correct posteriors.
+Additionally, it could be due to the approximation of the actual stochastic volatility model not producing accurate parameter estimates.
+This is addressed using importance sampling weights by KSC and explored in the next section.
 
-    \subsection{Reweigthing rank statistics of Gaussian Mixture}
-    \citet{kim1998stochastic} correct for approximation error in their method by using importance sampling weights.
+    \subsection{Correcting approximation error using importance weights}
+    \citet{kim1998stochastic} correct for approximation error in their method by using importance weights.
 The reweighting procedure ensures that samples are drawn from the correct posterior density.
-This is applied to the calculation of the expected value of the posterior density to get a weighted mean estimate from the samples drawn from the Gaussian mixture approximation.
-The rank statistic can be written as a function of the expected value, so the reweighting step can be applied to the calculation of the rank statistics to see if this correction improves the calibration of the sampling approach.
+This is applied to the calculation of the expected value of the stochastic volatility posterior density using samples drawn from the Gaussian mixture approximation.
 
+    Importance weights are defined as the ratio of the joint posterior from the stochastic volatility model and the posterior distribution of the approximation.
+A weight is produced for each posterior sample generated by MCMC.
+These weights correct the samples from the approximate distribution by increasing or decreasing the contribution of that sample in the calculation of the expectation.
+
+    These weights can be applied in an additional sampling step using Metropolis Hastings or Importance Resampling (also known as sampling-importance resampling or SIR) to produce samples from the target posterior.
+However, the rank statistic can also be written as a function of the weighted expectation of the indicator random variable.
+This reweighting step is applied to the calculation of the rank statistics to see if the correction improves the calibration of the MCMC sampler.
+
+    \subsubsection{Reweigthing rank statistics of the Gaussian Mixture Model}
     Let $v(\theta, h)$ be the log weights defined as the log difference between the posterior densities of the true model with log chi squared errors $log\: g(\theta, h | y^{\ast}_t)$ and the Gaussian mixture model $log\:  k(\theta, h_t | y^{\ast}_t)$.
@@
 
-    This gives the importance sampling weight. The expectation for any function of the posterior samples can be written as a function of the weights.
+    This gives the normalised importance weight.
+The expectation for any function of the posterior samples can be written as a function of these weights.
+Let $S(\theta)$ be an indicator random variable that is a function of the posterior samples, the expectation can be written as:
 
@@
 
-    The results from applying this reweighting step to the rank statistics are given on Figure \ref{fig:reweight1k}.
+    The results from applying this reweighting step to the rank statistics of the centered model are given on Figure \ref{fig:reweight1k}.
 There are no major improvements to the uniformity of rank statistics.
@@
 The shape of both sets of histograms is consistent with the shape of the unweighted rank statistics.
+The reweighted rank statistics for the reparameterised model exhibit the same behaviour and can be found in Appendix X.
 

### 2023-10-07 manuscript/main.tex
Updated ess and added algo summaries

@@
 
-\title{Comparison of MCMC algorithms in Stochastic Volatility Models}
+\title{Comparing MCMC algorithms in Stochastic Volatility Models using Simulation Based Calibration}
 \author{Benjamin Wee}
@@
 
-        The standard Kalman Filter and smoother is used to compute the posterior distribution over the latent states.
+        The standard Kalman Filter and simulation smoother is used to compute the posterior distribution over the latent states.
 This requires the state and measurement equations to be linear and conditionally Gaussian.
@@
 
+        \subsubsection*{Implementation}
+        The code used to replicate the KSC algorithm was written by \citet{chad2018} and ammended for the purposes of this research.
+Additional code was developed to produce rank statistics, quantities of interest as well as integrated into the SBC framework.
+
+        The steps in the MCMC process is desribed in Algorithm \ref{alg:ksc}.
+The initial values are set to the values outlined in \citet{kim1998stochastic}, except for the mixing indicator which is unspecified.
+This is arbitrarily set at the 4th Gaussian density found in Appendix A.
+
+        \begin{algorithm}[H]
+            \caption{KSC MCMC Algorithm}\label{alg:ksc}
+            \begin{algorithmic}
+            \Require $s_0 = 4$, $\mu_0 = 0$, $\phi_0 = 0.95$, $\sigma^{2}_{\eta,0} = 0.02$
+            \For{\texttt{i in} $1:n_{draws}$}
+                    \State \text{Sample states (Kalman Filter and Smoother): } $\boldsymbol{h}_i \sim h|y^{\ast}, s_{i-1}, \phi_{i-1}, \sigma^{2}_{\eta,i-1}, \mu^{i-1}$ 
+                    \State \text{Sample mixture density: } $s_i \sim s|y^{\ast}, \boldsymbol{h}_{i-1}$
+                    \State \text{Sample conjugate density $\mu$: } $\mu_i \sim \mu|y_{\ast}, s_{i-1}, \phi_{i-1}, \sigma^{2}_{\eta, i-1}, \boldsymbol{h}_{i-1}$
+                    \State \text{Sample conjugate density $\sigma^2_{\eta}$: } $\mu_i \sim \mu|y^{\ast}, s_{i-1}, \phi_{i-1}, \mu_{i-1}, \boldsymbol{h}_{i-1}$
+                    \State \text{Metrpolis Hastings step $\phi$: } $\phi_i \sim \phi|y^{\ast}, s_{i-1}, \mu_{i-1}, \sigma^{2}_{\eta, i-1}, \boldsymbol{h}_{i-1}$
+                  \EndFor
+            \end{algorithmic}
+            \end{algorithm}
+
     \subsection{Sampling method 2}
-    % Need much more detail here about how it works
         Since the paper was written, new MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
@@
 
+        \subsubsection{Implementation}
         The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
@@
 This approach will also use the same priors as specified in the Gaussian mixture approximation.
-
-        \subsubsection{Algorithmic steps}
+A general description of the Hamltonian Monte Carlo algorithm is given below.
+
         Let $\theta$ be a target parameter and $\phi$ be the auxiliary momemtum variable. The Hamiltonian is defined as:
@@
 \section{Results}
-    1000 and 5000 SBC iterations are run for centered and reparameterised versions of the stochastic volatility model.
+    1000 and 5000 SBC iterations are run for centered and reparameterised versions of the stochastic volatility model using simulated datasets of 1000 observations.
 Rank statistic histograms are constructed with 20 bins for the static parameters and the 1st, 500th and 1000th latent state variables.
@@
 The $\chi^2$ statistic is used to summarise the deviations from uniformity for all parameters since it is unrealistic to visually inspect all the histograms.
-The SBC algorithm is summarised below:
-
-    1) for sim in 1000 iterations:
-        2) Draw from prior: $\theta^{sim}\sim\pi (\theta)$
-        3) Simulate dataset with 1000 observations: $y^{sim} \sim p(y|\theta^{sim})$
-        4) Draw 999 posterior samples (post warmup) $\{\theta_1,\dots , \theta_{L}\} \sim p(\theta | y^{sim})$
-        5) Compute rank statistics $r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim})$
+The SBC algorithm is summarised Algorithm \ref{alg:sbc}:
+
+    \begin{algorithm}
+        \caption{SBC}\label{alg:sbc}
+        \begin{algorithmic}
+        \For{\texttt{i in} $1:1000$}
+                \State \text{Draw from joint prior: } $\boldsymbol{\theta}^{sim}_i \sim\pi (\boldsymbol{\theta})$
+                \State \text{Simulate dataset with 1000 observations: } $\boldsymbol{y}^{sim}_i \sim p(y|\boldsymbol{\theta}^{sim}_i)$
+                \State \text{Draw 999 posterior samples post burn in:} $\{\boldsymbol{\theta}_1,\dots , \boldsymbol{\theta}_{999}\}_i \sim p(\theta | \boldsymbol{y}^{sim}_i)$
+                \State \text{Compute rank statistics:} $\boldsymbol{r} = rank(\{\boldsymbol{\theta}_1,\dots , \boldsymbol{\theta}_{999}\}_i, \boldsymbol{\theta}^{sim}_i)$
+              \EndFor
+        \end{algorithmic}
+        \end{algorithm}
 
@@
 
+\newpage
+
+\section{Appendix A: Mixture Gaussian weights}
+
+\newpage
+
+\section{Appendix B: Sampling from mixture of Gaussians}
+
+\subsection*{Conjugate posterior distributions}
+
+\textbf{Sampling} $\boldsymbol{\sigma_{\eta}^2}$
+
+Inverse gamma conjugate posterior distribution:
+
+$$
+\sigma^2_{\eta} | y,h,\phi,\mu \sim IG \Bigl\{\frac{n+\sigma_r}{2}, \frac{0.05 + (h_1 - \mu)^2 (1 - \phi^2) + \sum_{t=1}^{n-1}((h_{t+1} - \mu) - \phi(h_t - \mu))^2}{2}\Bigr\}
+$$
+
+\textbf{Sampling}  $\boldsymbol{\mu}$
+
+Gaussian conjugate posterior distribution:
+
+$$
+\mu | h,\phi,\sigma^2_{\eta}  \sim N(\hat{\mu}, \sigma^2_{\mu})
+$$
+
+Where
+
+$$
+\begin{aligned}
+\hat{\mu} &= \sigma^2_{\mu} \Bigl\{\frac{(1-\phi^2)}{\sigma_{\eta}^2}h_1 +\frac{(1-\phi^2)}{\sigma_{\eta}^2} \sum_{t=1}^{n-1} (h_{t+1} - \phi h_t)\Bigr\} \\
+\sigma^2_{\mu} &= \sigma^2_{\eta} \{(n-1)(1-\phi)^2 + (1-\phi^2)\}^{-1}
+\end{aligned}
+$$
+
+\subsection*{Metropolis Hastings Step}
+\textbf{Sampling}  $\boldsymbol{\phi}$
+
+Metropolis Hastings accept/reject procedure:
+
+1) Generate proposal $\phi^\ast$ from $N(\hat{\phi}, V_{\phi})$ where $\hat{\phi} = \frac{\sum_{t=1}^{n-1} (h_{t+1} - \mu)(h_t - \mu)}{\sum_{t=1}^{n-1} (h_t - \mu)^2}$ and $V_{\phi} = \sigma^2_{\eta} \{\sum_{t=1}^{n-1} (h_t - \mu)^2\}^{-1}$
+
+2) Accept proposal as $\phi^{(i)}$ with probability $e^{\{g(\phi^\ast) - g(\phi^{(i-1)}\}}$ such that $g(\phi) = log (\pi (\phi)) - \frac {(h_t - \mu)^2 (1-\phi^2)}{2 \sigma_{\eta}^2} + \frac{1}{2} log (1-\phi^2)$
+
+\subsection*{Sampling mixture density}
+
+Rewrite mixture density with respect to a indicator variable $s_t$
+
+$$
+\begin{aligned}
+&z_t | s_t = i \sim N(m_i - 1.2704, \nu^2) \\
+&Pr(s_t = i) = q_i
+\end{aligned}
+$$
+
+Sample $s_t$ from probability mass function:
+
+$$
+\begin{aligned}
+Pr(s_t = i | y_t^{\ast}, h_t) \propto q_i f_N(y_t^{\ast} | h_t + m_t - 1.2704, \nu^2)
+\end{aligned}
+$$
+
 \end{document}

### 2023-10-07 manuscript/results/ess.r

@@ -1,4 +1,6 @@
 library(tidyverse)
+library(future.apply)
+plan(multicore, workers = 9)
 source("R/sbc.r")
 static_parameters <- c("~mu", "~sigma^2", "~phi")
 
@@ -34,7 +36,7 @@ hmc_ncp_ess_basic_5k <- future_lapply(rds_path,
 
 hmc_ncp_ess_basic_5k_df <- as.data.frame(do.call("rbind", hmc_ncp_ess_basic_5k))
 hmc_ncp_ess_basic_5k_df <- latex_variables(hmc_ncp_ess_basic_5k_df)
-hmc_ncp_ess_basic_5k_df$Type <- "Non Centered"
+hmc_ncp_ess_basic_5k_df$Type <- "Reparameterised"
 
 hmc_ess_df <- rbind(hmc_cp_ess_basic_5k_df, hmc_ncp_ess_basic_5k_df) %>% 
     filter(parameters %in% static_parameters)
@@ -66,31 +68,30 @@ ksc_cp_ess_basic_5k <- future_lapply(rds_path,
 
 ksc_cp_ess_basic_5k_df <- as.data.frame(do.call("rbind", ksc_cp_ess_basic_5k))
 ksc_cp_ess_basic_5k_df <- latex_variables(ksc_cp_ess_basic_5k_df)
-ksc_cp_ess_basic_5k_df$Type <- "Gaussian Mixture"
-
+ksc_cp_ess_basic_5k_df$Type <- "Centered"
 
 # 5000 iterations NCP
-path <- "simulation_output/ksc/cp/sir_sbc_cp_ksc_model_cp_dgf_10kmcmc_5000iter_r1"
+path <- "simulation_output/ksc/ncp/sbc_ncp_ksc_model_ncp_dgf_10kmcmc_5000iter_r1"
 rds_list <- list.files(path = paste(path, "output", sep = "/"), pattern = "*.RDS")
 rds_path <- paste("output", rds_list, sep = "/")
 
-sir_ncp_ess_basic_5k <- future_lapply(rds_path,
+ksc_ncp_ess_basic_5k <- future_lapply(rds_path,
                            get_ess, 
                            model_path = path, 
                            ess_type = "ess_basic", 
                            chains = "one_chain", 
                            future.seed = NULL)
 
-sir_ncp_ess_basic_5k_df <- as.data.frame(do.call("rbind", sir_ncp_ess_basic_5k))
[…]

### 2023-10-07 manuscript/motivating_example/real_data_example.r
updated tables

@@ -28,7 +28,7 @@ ggplot(df, aes(x = value)) +
     facet_wrap(~name, scale="free", labeller = label_parsed) +
     # geom_vline(data =means, aes(xintercept=avg, colour=model), size = 1.2) +
     theme_minimal(base_size = 20) +
-    scale_fill_discrete(labels = c("Method 1", "Method 2")) +
+    scale_fill_discrete(labels = c("KSC", "HMC")) +
     labs(title = "Posterior distribution - S&P 500",
          fill = "MCMC\nSamplers",
          x = "Parameter value", 
@@ -41,7 +41,7 @@ ggsave("manuscript/motivating_example/real_data_ex.png", bg = "white", width = 1
 
 df %>% group_by(model, name) %>% 
   summarise(as_tibble(rbind(summary(value)))) %>% 
-  ungroup() %>% arrange(name)
+  ungroup() %>% arrange(model, name)
 
 ggplot(df, aes(x = value)) +
     geom_histogram(aes(fill = model), alpha =0.4) +

### 2023-10-07 manuscript/results/ess.r

@@ -53,6 +53,10 @@ x_labs = ggplot_build(plot)$layout$panel_params[[1]]$x$get_labels()
 plot <- plot + scale_x_discrete(labels = parse(text = x_labs))
 ggsave("manuscript/results/hmc_ess.png", bg = "white", width = 14, height = 9.42)
 
+hmc_ess_df %>% group_by(Type, parameters) %>% 
+      summarise(as_tibble(rbind(summary(ess_basic)))) %>% 
+      ungroup() %>% arrange(Type, parameters)
+
 # KSC results
 # 5000 iterations CP
 path <- "simulation_output/ksc/cp/sbc_cp_ksc_model_cp_dgf_10kmcmc_5000iter_r1"
@@ -101,6 +105,9 @@ x_labs = ggplot_build(plot)$layout$panel_params[[1]]$x$get_labels()
 plot <- plot + scale_x_discrete(labels = parse(text = x_labs))
 ggsave("manuscript/results/ksc_ess.png", bg = "white", width = 14, height = 9.42)
 
+ksc_ess_df %>% group_by(Type, parameters) %>% 
+      summarise(as_tibble(rbind(summary(ess_basic)))) %>% 
+      ungroup() %>% arrange(parameters, Type)
 
 # 5000 iterations NCP
 # path <- "simulation_output/ksc/cp/sir_sbc_cp_ksc_model_cp_dgf_10kmcmc_5000iter_r1"

### 2023-10-07 manuscript/main.tex
Updates from Overleaf

@@
 The data are the de-meaned log returns of the 2023 January-September S\&P 500 Index.
-Below are the marginal posterior estimates for the static parameters $\mu$, $\phi$ and $\sigma^2$ and corresponding summary statistics.
+Figure \ref{fig:realdataex} visualises the marginal posterior estimates for the static parameters $\mu$, $\phi$ and $\sigma^2$ with corresponding summary statistics in Tables \ref{tab:realdata}.
 
@@
         \includegraphics[scale=0.1]{motivating_example/real_data_ex.png}
-        \caption{Posterior samples from two MCMC algorithms using S\&P 500 data.
+        \caption{Posterior samples from the Hamiltonian Monte Carlo (HMC) and Kim Sherphard and Chib (KSC) MCMC samplers using S\&P 500 data.
 While the shape of the distributions are similar, the modes of method 2 is much higher and the tails of method 1 are fatter for $\phi$ and $\sigma^2$.
 Convergence diagnostics do not tell us which estimate is "more correct" since we don't observe the true model or true data generating process.}
-    \end{figure}
+        \label{fig:realdataex}
+    \end{figure}
+
+    \begin{table}[H]
+        \centering
+        \begin{tabular}{|c|c|c|c|c|c|c|c|} \hline 
+        Parameter&  MCMC&Min& q25&  Median& Mean & q75&Max\\ \hline 
+        $\mu$&  KSC&-2.03 & -0.472 & -0.354 & -0.355 & -0.231 & 1.31 \\
+     $\mu$&  HMC&-4.55 & -0.508 & -0.373 & -0.370 & -0.229 &2.86  \\\hline 
+     $\phi$&  KSC&0.384 & 0.875 & 0.928 & 0.902 & 0.958 & 0.997 \\
+     $\phi$&  HMC&0.438 & 0.910 & 0.953 & 0.929 & 0.977 &1.00  \\ \hline 
+     $\sigma^2$&  KSC&0.00237 & 0.00891 & 0.0138 & 0.0178 & 0.0210 & 0.130 \\ 
+     $\sigma^2$&  HMC&0.00209 & 0.00731 & 0.0105 & 0.0131 & 0.0159 &0.107 \\ \hline
+        \end{tabular}
+        \caption{Summary statistics for KSC sampler}
+        \label{tab:realdata}
+    \end{table}
 
     These estimates show the marginal posteriors of each parameter sharing similar shapes.
-$\phi$  has a fatter left tail for and $\sigma^2$ has heavier right tails for sampling method 1.
+$\phi$  has a fatter left tail for the KSC sampler and $\sigma^2$ has heavier right tails for HMC.
+This is evident in the quantiles with the 25th quantile for $\phi$ being smaller at 0.875 for KSC compared to 0.910 in HMC.
 Furthermore, sampling method 2 has higher peaks at their mode relative to method 1 with a tighter spread.
@@
     Additionally, HMC struggles to generate efficient posterior samples from the centered parameterisation of the stochastic volatility model.
-Figure \ref{fig:hmcess} visualises the distribution of ESS for the static parameters after 5000 SBC iterations.
+Figure \ref{fig:hmcess} and Table \ref{tab:hmcess} summarises the distribution of ESS for the static parameters after 5000 SBC iterations.
 The non centered model appears to generate independent samples much more efficiently than the centered model.
@@
     \end{figure}
+
+    \begin{table}[H]
+        \centering
+        \begin{tabular}{|c|c|c|c|c|c|c|c|} \hline 
+        Parameter &  Type&Min& q25&  Median& Mean & q75&Max\\ \hline 
+        $\mu$&  Centered&3.79 & 99.7 & 184. & 305. & 457. & 1997.  \\
+     $\mu$&  Reparam&14.5 & 738. & 1099. & 1143. & 1500. & 2993.  \\\hline 
+     $\phi$&  Centered&1.44 & 11.0 & 21.7 & 26.3 & 34.1 & 635.  \\
+     $\phi$&  Reparam&48.8 & 355. & 476. & 521. & 651. & 1445.   \\ \hline 
+     $\sigma^2$&  Centered& 1.17 & 2.99 & 7.04 & 8.05 & 11.3 & 90.7 \\ 
+     $\sigma^2$&  Reparam&46.1 & 403. & 553. & 585. & 734. & 1486. \\ \hline
+        \end{tabular}
+        \caption{HMC: ESS for centered and reparameterised stochastic volatility model.}
+        \label{tab:hmcess}
+    \end{table}
 
@@
     The ESS estimates for this model indicate a high degree of autocorrelation and difficulty in generating independent samples from both parameterisations.
-The numerical results are summarised in table x.
+The numerical results are summarised in Figure \ref{kscess} and Table \ref{tab:kscess}.
 The estimate for $\mu$ in the reparameterised model exhibits a long right tail (longer than the 10,000 posterior draws suggesting the presence of negatively autocorrelated draws).
@@
     \end{figure}
+
+    \begin{table}[H]
+        \centering
+        \begin{tabular}{|c|c|c|c|c|c|c|c|} \hline 
+        Parameter &  Type&Min& q25&  Median& Mean & q75&Max\\ \hline 
+        $\mu$&  Centered&9.57 & 352. & 559. & 1482. & 1552. & 9907.\\
+     $\mu$&  Reparam&1.99 & 205. & 569. & 4174. & 6384. & 34055.\\\hline 
+     $\phi$&  Centered&2.47 & 74.0 & 103. & 112. & 130. & 1222.\\
+     $\phi$&  Reparam&1.03 & 83.4 & 132. & 325. & 419. & 1907. \\ \hline 
+     $\sigma^2$&  Centered&2.33 & 29.0 & 39.5 & 44.2 & 52.8 & 384. \\ 
+     $\sigma^2$&  Reparam&1.25 & 47.4 & 73.4 & 399. & 591. & 6618. \\ \hline
+        \end{tabular}
+        \caption{KSC: ESS for centered and reparameterised stochastic volatility model.}
+        \label{tab:kscess}
+    \end{table}
 
@@
 
+\begin{table}[H]
+    \centering
+    \begin{tabular}{lccc} 
+          $\omega$ &$Pr(\omega = i)$&  $m_i$&  $\nu^2_i$\\ 
+          1&0.00730  &  -10.12999&  5.79596\\ 
+          2&0.10556  &   -3.97281 &  2.61369\\ 
+          3&0.00002 &  -8.56686 &   5.17950\\ 
+          4&0.04395 &  2.77786  &   0.16735 \\ 
+          5&0.34001&   0.61942    &  0.64009\\ 
+          6&0.24566 &  1.79518    &  0.34023 \\ 
+          7&0.25750 &  -1.08819    &  1.26261\\ 
+    \end{tabular} 
+\end{table}
+
 \newpage

### 2023-10-08 manuscript/main.tex
fixed ref
updated most plots

@@
     This research compares the computational methods used to estimate Bayesian stochastic volatility models.
-The estimation strategies outlined in the stochastic volatility literature and Hamiltoninan Monte Carlo are assesssed in their ability to sample from the model's posterior distribution.
-Specifically, Simulation Based Calibration (SBC) is used to check whether these MCMC algorithm are returning efficient and well calibrated posterior estimates.
+The MCMC approaches outlined in the stochastic volatility literature and Hamiltoninan Monte Carlo are assesssed in their ability to sample from the model's posterior distribution.
+Specifically, Simulation Based Calibration (SBC) is used to check whether these MCMC algorithms are returning efficient and calibrated posterior estimates.
 Key metrics of interest are the effective sample size to check the efficiency of the algorithm and tests of uniformity to assess the calibration of the posteriors.
 This will determine which method is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
-Results reveal that Hamiltonian Monte Carlo provides more efficient and calibrated posterior estimates conditional on the non centered paramterisation of the stochastic volatility model.
+Results reveal that Hamiltonian Monte Carlo provides more efficient and calibrated posterior estimates of the stochastic volatility model conditional on model paramterisation.
 \end{abstract}
@@
 \section{Introduction}
-    The stochastic volatility model is used in financial econometrics to model the behaviour of financial instruments.
-It explicitly treats the variance as a latent random variable.
-These are typically expressed as non linear Gaussian state space models and are difficult to estimate using classical methods.
-The likelihood is unavailable in closed form and there are at least as many variables as data points.
-Bayesian approaches to estimating these models rely on  Markov Chain Monte Carlo (MCMC) techniques to sample from the target joint posterior of these high dimensional latent parameter spaces.
-
-    \citet{kim1998stochastic} propose a method for sampling this model using a combination of Kalman Filters, Gibbs sampler, and Metropolis Hastings over 20 years ago.
+    The stochastic volatility model is used in financial econometrics to model the behaviour of financial assets.
+They are typically expressed as non linear Gaussian state space models where the variance is treated as a latent random variable.
+Stochastic volatility is difficult to estimate since the likelihood is unavailable in closed form and there are at least as many variables as data points.
+Bayesian methods can be used to estimate these models which rely on Markov Chain Monte Carlo (MCMC) techniques to sample from the target joint posterior of these high dimensional parameter spaces.
+
+    \citet{kim1998stochastic} proposed a sampling strategy for stochastic volatility through a combination of conjugate posterior distributions, a Kalman filter, and Metropolis Hastings over 20 years ago.
 Since then, there have been many advances in statistical computing to estimate complex, high dimensional Bayesian models.
@@
 $h_t$ is log volatility, where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the autoregressive parameter $\phi$ such that $|\phi|<1$.
-This autoregressive parameter represents the persistence or "stickiness" of log volatility and the dispersion parameter $\sigma_{\eta}$ is the constant variance of the states.
+This autoregressive parameter represents the persistence or stickiness of log volatility and the dispersion parameter $\sigma_{\eta}$ is the constant variance of the states.
 $\epsilon_t$ and $\eta_t$ are standard normal white noise shocks and are uncorrelated with each other.
@@
 
-    The prior on $\phi$ is a "stretched" beta distribution. This is a beta distribution (as defined on the parameter $\phi^*$) which has been transformed to have support (-1, 1).
+    The prior on $\phi$ is a stretched beta distribution. This is a beta distribution (as defined on the parameter $\phi^*$) which has been transformed to have support (-1, 1).
 
@@
 While the shape of the distributions are similar, the modes of method 2 is much higher and the tails of method 1 are fatter for $\phi$ and $\sigma^2$.
-Convergence diagnostics do not tell us which estimate is "more correct" since we don't observe the true model or true data generating process.}
+Convergence diagnostics do not tell us which estimate is ``more correct'' since we don't observe the true model or true data generating process.}
         \label{fig:realdataex}
@@
 \cite{talts2018validating} make the point that a single simulation does not provide sufficient information about the inference made by an algorithm.
-As discussed in their paper, a single simulation may conclude "that an incorrectly coded analysis worked as desired, while a correctly coded analysis failed".
+As discussed in their paper, a single simulation may conclude ``that an incorrectly coded analysis worked as desired, while a correctly coded analysis failed''.
 
@@
 Indeed, there are limitations to evaluating MCMC algorithms based on fits to real data and single simulations.
-To check the calibration of an algorithm, simulations over multiple "true parameters" are required.
+To check the calibration of an algorithm, simulations over multiple true parameters are required.
 This methodology is discussed in the next section.
@@
             A perfectly discrete uniform distribution will return a $\chi^2$ statistic of zero.
-That is, the number of rank statistics in each bin is equal to the expected number conditional on the number of bins and observations.
+That is, the number of rank statistics in each bin is equal to the expected count under a discrete uniform distribution conditional on the number of bins and observations.
 Instead of performing multiple hypothesis tests for each parameter, the distribution of $\chi^2$ statistics is compared across all parameters and simulations.
@@
     \subsection{Sampling method 1}
-        (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.} \citep{dejong1995}.
+        (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable.
+So a more recent simulation smoother is used which is written by the same author of the original software.} \citep{dejong1995}.
 
@@
 
-        The stochastic volatility model as described in section 2.1 is follows a "centered" parameterisation.
+        The stochastic volatility model as described in section 2.1 is follows a centered parameterisation.
 This describes the central location of the latent state parameters which are centered around the mean and lag of the log volatility $\mu +\phi(h_t - \mu)$.
@@
 The black horizontal line represents uniformity conditional on the number of bins and SBC iterations.
-The $\chi^2$ statistic is used to summarise the deviations from uniformity for all parameters since it is unrealistic to visually inspect all the histograms.
+The $\chi^2$ statistic is used to summarise the deviations from uniformity for all state variables since it is unrealistic to visually inspect all the histograms.
 The SBC algorithm is summarised Algorithm \ref{alg:sbc}:
@@
     The ESS estimates for this model indicate a high degree of autocorrelation and difficulty in generating independent samples from both parameterisations.
-The numerical results are summarised in Figure \ref{kscess} and Table \ref{tab:kscess}.
+The numerical results are summarised in Figure \ref{fig:kscess} and Table \ref{tab:kscess}.
 The estimate for $\mu$ in the reparameterised model exhibits a long right tail (longer than the 10,000 posterior draws suggesting the presence of negatively autocorrelated draws).
@@
 However, the rank statistic can also be written as a function of the weighted expectation of the indicator random variable.
-This reweighting step is applied to the calculation of the rank statistics to see if the correction improves the calibration of the MCMC sampler.
+This reweighting step is applied to the calculation of the rank statistics to see if the correction improves the calibration of the MCMC sampler.\footnote{Efficiency estimates of the reweighted posterior samples are omitted.
+It was unclear at the time of writing whether the ESS could be written as a function of the expectation or whether the ESS from resampled posterior samples using the importance weights were valid.}
 
@@
     \begin{aligned}
-    r = \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}] \approx  L\times E[S(\hat{\theta})] = L\times \sum_l^L S(\theta_l)w_l
+    r = \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}] \approx  L\times E[S(\hat{\theta})] = L\times \sum_l^L S(\hat{\theta}_l)w_l
     \end{aligned}
@@
 There are no major improvements to the uniformity of rank statistics.
-Increasing the SBC iterations to 5000 as seen in Figure \ref{fig:reweight5k} also results in no major improvements.
-The shape of both sets of histograms is consistent with the shape of the unweighted rank statistics.
-The reweighted rank statistics for the reparameterised model exhibit the same behaviour and can be found in Appendix X.
-
-    \begin{figure}[H]
-        \centering
-        \includegraphics[scale=0.1]{results/sir_cp_1k.png}
+In fact, there are now large peaks on the left side of the 1st and 500th state variables and a right peak for $\sigma^2$.
+Increasing the SBC iterations to 5000 as seen in Figure \ref{fig:reweight5k} also does not improve the results, with the left peak appearing to be larger for all state variables.
+Applying the reweighting step to the rank statistics appears to have weighted the posterior estimates to the right of the prior distribution.
+Overall the reweighting of the posterior samples from the approximate model did not improve the calibration results.
+The reweighted rank statistics for the reparameterised model also did not improve and can be found in Appendix C.
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/weighted_ksc_cp_1k.png}
         \caption{1000 SBC iterations for reweighted rank statistics from the Gaussian mixture approximation model. The shape of the histogram is consistent with the unweighted rank statistics.}
@@
         \centering
-        \includegraphics[scale=0.1]{results/sir_cp_5k.png}
+        \includegraphics[scale=0.1]{results/weighted_ksc_cp_5k.png}
         \caption{5000 SBC iterations for reweighted rank statistics from the Gaussian mixture approximation model. No major improvements are observed to the uniformity of rank statistics.}
@@
 
-    \subsection{Evaluating all models and parameters}
-    As discussed, it is not practical to visually inspect the histograms of all parameters in the stochastic volatility model.
-Instead, the Chi squared statistic can be used to summarise the shape of the distribution and if desired, be used for hypothesis testing.
-However, instead of performing hypothesis tests, another visual approach is taken to compare the calibration of the algorithms and parameters.
-    % However, performing all these tests suffers from the same overcommunication as presenting all the histograms at once.
-
-    The distribution of chi squared statistics across all simulations and all parameters are plotted as histograms on Figure \ref{fig:allchisq}.
+    \subsection{Evaluating SBC for state parameters}
+    As discussed, it is not practical to inspect the histograms of all parameters and states in the stochastic volatility model.
+Instead, the Chi squared statistic can is used to summarise the shape of the rank statistic distribution.
+This will be used to summarise the state variables only since the static parameters can be compared across simulations individually.
+
+    The state variable chi squared statistics for centered and reparamterised HMC, centered KSC and centered importance weighted KSC are reported as histograms in Figure \ref{fig:allchisq} (using results from 5000 SBC iterations).
+Reparameterised KSC is omitted as its poor SBC performance simply adds noise to the results.
 A chi squared statistic of 0 implies the sample is exactly uniformly distributed.
 Any value away from 0 captures any variation or deviation away from uniformity.
-This provides a high level view of all parammeters - how close they are to uniformity and how close or far away they are relative to the other SBC results for other algorithms and parameterisations.
-
-    \begin{figure}[H]
-        \centering
-        \includegraphics[scale=0.1]{results/dist_chisq.png}
-        \caption{Distribution of chi squared statistics for all models (5000 SBC iterations).}
+This provides a high level summary of the state variables - how close their rank statistics are to being uniformly distributed and their distance relative to the other algorithms and parameterisations.
+
+    The HMC results are relatively close to zero.
+Both KSC and the importance weighted ranks on the other hand are far from zero suggesting major deviations from uniformity.
+There is a large gap between the HMC and KSC chi squared statistics implying the distribution of state variable ranks are much less uniform for KSC relative to the HMC.
+Combined with the results in the previous section, there is a strong evidence that HMC outperforms KSC where it comes to calibration across all parmeters and latent state variables.
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/dist_chisq_all.png}
+        \caption{Distribution of chi squared statistics for latent state variables.
+Results from 5000 SBC iterations are used.
+Values closer to 0 are more consistent with a discrete uniform distribution.
+Both HMC simulations are much closer to zero than KSC.
+This suggests that the HMC algorithm overall produces more calibrated posterior estimates for the state variables}
         \label{fig:allchisq}
     \end{figure}
-
-    The distribution for HMC are clustered very closely to zero.
-Both results for the bespoke sampling method are distributed far from zero which suggests strong evidence of a lack of uniformity across all parameters.
-This is consistent with the distribution of rank statistics in the previous sections.
 
@@
 
-It is worth noting that there are many ways of "comparing" MCMC approaches.
+It is worth noting that there are many ways of comparing MCMC approaches.
 Other features that may be of interest is the speed in which a MCMC can converge onto the target distribution and generate samples for inference.
-However, "speed" is diffiuclt to define and is subject to a variety of different factors such as hardware, operating system, etc.
+However, speed is diffiuclt to define and is subject to a variety of different factors such as hardware, operating system, etc.
 The purpose of this research is to compare algorithms based on calibration and efficiency.
@@
 
-Lastly, it is unclear what constitutes a "fair comparison" between algorithms when comparing SBC results.
+Lastly, it is unclear what constitutes a fair comparison between algorithms when comparing SBC results.
 In particular, the number of post burnin or warmup samples differed between algorithms (999 post burn-in for HMC, 9999 for the bespoke algorithm).
@@
 
+\newpage
+
+\section{Appendix C: Importance weighted rank statistics for reparameterised model}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale=0.09]{results/weighted_ksc_ncp_1k.png}
+    \caption{1000 SBC iterations for reweighted rank statistics from the reparameterised Gaussian mixture approximation model.
+The shape of the histogram is consistent with the unweighted rank statistics.}
+    \label{fig:reweight1k}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[scale=0.09]{results/weighted_ksc_ncp_5k.png}
+    \caption{5000 SBC iterations for reweighted rank statistics from the reparameterised Gaussian mixture approximation model. No major improvements are observed to the uniformity of rank statistics.}
+    \label{fig:reweight5k}
+\end{figure}
+
 \end{document}

### 2023-10-08 manuscript/results/compare_histograms.r

@@ -1,74 +1,172 @@
 library(tidyverse)
+library(cowplot)
 source("R/sbc.r")
 
-n_iterations <- 1000
+n_iterations <- 5000
 posterior_samples <- 999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
 n_bins <- 20 # Bins
 expected_count <- n_iterations / n_bins # Chi sqd stat
 
 
 # HMC CENTERED results
-# 1000 iterations
+# 5000 iterations
 rank_type = "agg_ranks"
-path <- "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_1000_iterations"
+path <- "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_5000_iterations"
 
-hmc_cp_onek_bins <- rank_bins_f(path = "simulation_output/stan/cp/sbc_cp_ksc_priors_0.999_adapt_delta_premade_datasets_r1_1000_iterations", 
+hmc_cp_fivek_bins <- rank_bins_f(path = path, 
                                 rank_type = "agg_ranks", 
                                 posterior_samples = 999, 
                                 n_bins= 20)
 
-hmc_cprank_stats <- as.data.frame(lapply(hmc_cp_onek_bins, rank_chi_sqd, expected = expected_count)) %>% 
+hmc_cprank_stats <- as.data.frame(lapply(hmc_cp_fivek_bins, rank_chi_sqd, expected = expected_count)) %>% 
             pivot_longer(everything())
-hmc_cprank_stats$name <- names(hmc_cp_onek_bins)
-hmc_cprank_stats$type <- "hmc_cp"
+hmc_cprank_stats$name <- names(hmc_cp_fivek_bins)
+hmc_cprank_stats$type <- "Centered HMC"
+hmc_cprank_stats$group <- "HMC"
+
+# HMC NONCENTERED results
+# 5000 iterations
+rank_type = "agg_ranks"
+path <- "simulation_output/stan/ncp/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r9_5000_iterations"
+
+hmc_ncp_fivek_bins <- rank_bins_f(path = path, 
+                                rank_type = "agg_ranks", 
+                                posterior_samples = 999, 
+                                n_bins= 20)
+
[…]

### 2023-10-08 manuscript/results/rank_histograms_static_states.r

@@ -131,6 +131,67 @@ ggsave("manuscript/results/ksc_cp_5k.png", bg = "white", width = 14, height = 9.
 rank_hist(ksc_ncp_fivek_bins_long) # KSC NCP 5k
 ggsave("manuscript/results/ksc_ncp_5k.png", bg = "white", width = 14, height = 9.42)
 
+# REWEIGHTED KSC 1k
+n_iterations <- 1000
+posterior_samples <- 9999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+n_bins <- 20 # Bins
+expected_count <- n_iterations / n_bins # Chi sqd stat
+
+ksc_cp_onek_bins <- rank_bins_f(path = "simulation_output/ksc/cp/sbc_cp_ksc_model_cp_dgf_10kmcmc_1000iter_r2", 
+                                    rank_type = "weighted_ranks", 
+                                    posterior_samples = posterior_samples, 
+                                    n_bins= n_bins)
+
+
+ksc_ncp_onek_bins <- rank_bins_f(path = "simulation_output/ksc/ncp/sbc_ncp_ksc_model_ncp_dgf_10kmcmc_1000iter_r1", 
+                                    rank_type = "weighted_ranks", 
+                                    posterior_samples = posterior_samples, 
+                                    n_bins= n_bins)
+
+ksc_cp_onek_bins_long <- static_rank_bins(data = ksc_cp_onek_bins, 
+                                         variables = parameters, 
+                                         type = "Centered")
+
+ksc_ncp_onek_bins_long <- static_rank_bins(data = ksc_ncp_onek_bins, 
+                                         variables = parameters, 
+                                         type = "Centered")
+
+rank_hist(ksc_cp_onek_bins_long) # Reweighted KSC CP 1k
+ggsave("manuscript/results/weighted_ksc_cp_1k.png", bg = "white", width = 14, height = 9.42)
+
+rank_hist(ksc_ncp_onek_bins_long) # Reweighted KSC NCP 1k
+ggsave("manuscript/results/weighted_ksc_ncp_1k.png", bg = "white", width = 14, height = 9.42)
+
+# REWEIGHTED KSC 5k
+n_iterations <- 5000
+posterior_samples <- 9999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+n_bins <- 20 # Bins
[…]

### 2023-10-08 manuscript/main.tex
First proof read and edits

@@
 
-    \citet{kim1998stochastic} proposed a sampling strategy for stochastic volatility through a combination of conjugate posterior distributions, a Kalman filter, and Metropolis Hastings over 20 years ago.
-Since then, there have been many advances in statistical computing to estimate complex, high dimensional Bayesian models.
-In particular, the recent availability of Hamiltonian Monte Carlo, a variant of the Metropolis Hastings algorithm, is an advancement of MCMC for efficient sampling of complex probabilistic models.
+    \citet{kim1998stochastic} proposed a sampling strategy for stochastic volatility through a combination of conjugate posterior distributions, a Kalman filter, and the Metropolis Hastings algorithm.
+Since then, there have been many advances in statistical computing and algorithm design.
+In particular, Hamiltonian Monte Carlo - a variant of the Metropolis Hastings algorithm - is an advancement in MCMC for efficient sampling of complex, high dimensional probabilistic models.
 The adoption of such new techniques have been made widely available through the development of various open source libraries such as the Stan programming language \citep{stan} and the PyMC library \citep{pymc2023}.
@@
 As the development of new algorithms rapidly increase, so does the need to develop new methods to assess their output.
-Developments in our statistical workfow are required to test new algorithms as well as compare computational strategies used to estimate increasingly complex models.
-
-    This research conducts a simulation study to assess the algorithms used to estimate stochastic volatility models.
-The first algorithm replicates Kim Sherphard and Chib's (KSC) Gaussian mixture approximation which is applied on a transformed stochastic volatility model with a log chi squared error distribution.
-Their sampling method uses the aforementioned Kalman Filter estimating the latent states, conjugate posterior distributions and the Metropolis Hastings algorithm.
-The second algorithm is the Hamiltonian Monte Carlo (HMC) algorithm as implemented in the Stan programming language.
-The objective is to analyse the efficiency of the sampling approach and their ability to return calibrated posterior estimates.
+Developments in statistical workfow are required to test new algorithms as well as compare computational strategies used to estimate increasingly complex models.
+
+    This research conducts a simulation study to analyse the calibration and efficiency of MCMC used to estimate stochastic volatility models.
+The first algorithm replicates Kim Sherphard and Chib's (KSC) Gaussian mixture approximation which is sampled using the aforementioned Kalman Filter and Metropolis Hastings algorithm.
+The second algorithm is Hamiltonian Monte Carlo (HMC) as implemented in the Stan programming language.
+The objective is to analyse the efficiency of the MCMC and their ability to return correct posterior estimates.
 
     Comparisons of these algorithms are conducted by Simulation Based Calibration \citep{talts2018validating}.
-Parameters are drawn from the prior distribution and used to create datasets from the generative (stochastic volatility) model which are sampled using a MCMC method.
-Repeating this process multiple times gives insight to how well the sampling approach can estimate the true parameters governing the data generating process on average.
-The key diagnostic metrics are the effective sample size to measure the efficiency of the algorithm and visual tests of uniform rank statistics to assess calibration of the posterior estimates.
+Parameters are drawn from the prior distribution and used to create datasets from the generative stochastic volatility model which are sampled using one of the mentioned MCMC methods.
+Repeating this process multiple times gives insight to how well the MCMC algorithm can estimate the true parameters governing the data generating process.
+The key diagnostic metrics are the effective sample size to measure the efficiency of the algorithm and visual assessment of uniform rank statistics to determine calibration of the posterior estimates.
 
     Additionally, two different parametiersations of the stochastic volatility model are considered for each sampling method.
-These are described as either 'centred' or 'non-centred' in location.
-An algorithm may be sensitive to different parameterisations of the same model which may impact MCMC efficiency as described in \citet{strickland2008parameterisation}.
+An algorithm may be sensitive to different parameterisations of the same model which may impact MCMC performance as described in \citet{strickland2008parameterisation}.
 The key contribution of this research is to use simulations to assess whether a MCMC algorithm can return calibrated posterior estimates as well as the efficiency under the different parametiersations.
@@
     This paper is structured as follows.
-Section 2 provides the context around this research, namely the stochastic volatility model examined and the limitations of single simulation studies and MCMC convergence diagnostics.
+Section 2 provides the context around this research, namely the stochastic volatility model and the limitations of single simulation studies and MCMC convergence diagnostics.
 Section 3 describes the simulation design and diagnostic metrics.
@@
 Convergence metrics are often used to check the performance of the MCMC chains, such as the effective sample size and the $\hat{R}$.
-These diagnostics provide evidence on whether an MCMC has failed to converge onto the target posterior distribution.
+These diagnostics provide evidence on whether a MCMC has failed to converge onto the target posterior distribution.
 
@@
 That is, the true parameter or model is unobservable.
-Therefore, failed diagnostic checks could arise from either inocrrect model specification or issues in the sampling algorithm or both.
-
-    Furthermore, different sampling strategies and MCMC tools could provide different posterior estimates for the same model.
-Assuming no issues in computation, these diagnostic measures do not provide any evidence to which estimate is closer to the truth.
+Therefore, failed diagnostic checks could arise from either inocrrect model specification or issues in the MCMC algorithm or both.
+Furthermore, different sampling strategies could provide different posterior estimates for the same model.
+Assuming no issues in computation, convergence diagnostics do not provide any information to which estimate is more correct or closer to the true parameter.
 
     To illustrate this problem, the stochastic volatility model is estimated with two different MCMC algorithms on real data.
-The data are the de-meaned log returns of the 2023 January-September S\&P 500 Index.
-Figure \ref{fig:realdataex} visualises the marginal posterior estimates for the static parameters $\mu$, $\phi$ and $\sigma^2$ with corresponding summary statistics in Tables \ref{tab:realdata}.
-
-    \begin{figure}[h]
+The data are the de-meaned daily log returns of the 2023 January-September S\&P 500 Index.
+Figure \ref{fig:realdataex} visualises the marginal posterior estimates for the static parameters $\mu$, $\phi$ and $\sigma^2$ with corresponding summary statistics in Table \ref{tab:realdata}.
+
+    These estimates show the marginal posteriors of each parameter sharing similar shapes.
+$\phi$ has a fatter left tail for the KSC sampler and $\sigma^2$ is also heavier in the right tails relative to HMC.
+This is evident in the quantiles with the 25th quantile for $\phi$ being smaller at 0.875 for KSC compared to 0.910 in HMC.
+Furthermore, HMC has higher peaks at their mode relative to KSC with a tighter spread.
+Based on this information alone, there is no certainty in whether method 1 or method 2 is has a more correct estimate conditional on data and model.
+
+    % Other methods for model selection in this context (for example, out of sample predictive performance), but for parameter estimation it is unclear which one should be selected.
+
+    \begin{figure}[H]
         \centering
@@
         \caption{Posterior samples from the Hamiltonian Monte Carlo (HMC) and Kim Sherphard and Chib (KSC) MCMC samplers using S\&P 500 data.
-While the shape of the distributions are similar, the modes of method 2 is much higher and the tails of method 1 are fatter for $\phi$ and $\sigma^2$.
-Convergence diagnostics do not tell us which estimate is ``more correct'' since we don't observe the true model or true data generating process.}
+While the shape of the distributions are similar, the modes of HMC is much higher and the tails of KSC are fatter for $\phi$ and $\sigma^2$.
+Convergence diagnostics do not tell us which estimate is more correct since we don't observe the true model or true data generating process.}
         \label{fig:realdataex}
@@
         \end{tabular}
-        \caption{Summary statistics for KSC sampler}
+        \caption{Summary statistics for HMC and KSC algorithms.
+KSC's 25th quantile $\phi$ is smaller than HMC which is consistent with the fatter left tail.
+Similarly, the KSC's 75th qunatile for $\sigma^2$ is larger consistent with a larger right tail spread.
+Both sets of estimate are reasonable but it is unclear from estimates on real data which estimate is closer the the truth.}
         \label{tab:realdata}
     \end{table}
-
-    These estimates show the marginal posteriors of each parameter sharing similar shapes.
-$\phi$  has a fatter left tail for the KSC sampler and $\sigma^2$ has heavier right tails for HMC.
-This is evident in the quantiles with the 25th quantile for $\phi$ being smaller at 0.875 for KSC compared to 0.910 in HMC.
-Furthermore, sampling method 2 has higher peaks at their mode relative to method 1 with a tighter spread.
-Based on this information alone, there is no certainty in whether method 1 or method 2 is has a more correct estimate conditional on data and model.
-There are other methods for model or algorithmic selection in this context (for example, out of sample predictive performance), but for parameter estimation it is unclear which one should be selected.
 
@@
 A strategy around this is to evaluate a model and algorithm on simulated data.
-One approach is to simulate data from the proposed model using known parameters.
-Then, fit the model on the simulated data and see if the model can recover the true parameters.
+One approach is to simulate data from a generative model using known parameters.
+Then fit the same model on the simulated data and see if the true parameters can be recovered.
 This gives us the benefit of defining the true parameters of the data generating process to be estimated.
@@
     Furthermore, as discussed in \citet{gelman2020bayesian}, fitting models on simulated data is the only way we can check inference on latent variables.
-This is critical in Stochastic Volatility since the underlying framework is a state space model with latent log volatility parameters.
+This is critical in stochastic volatility since the underlying framework is a state space model with latent log volatility parameters.
 Latent variables are unobserved in real data and are only estimated in the context of the model.
@@
 
-    Simulations enable us to check whether our models and algorithms can estimate the true data generating process, however, there are limitations to what can be learned from a single simulation.
+    Simulations enable us to check whether our models and algorithms can estimate the true data generating process.
+However, there are limitations to what can be learned from a single simulation.
 There is always a small probability that the true parameter is in the tails of any estimated posterior distribution.
@@
 
-    Figure 2 show the posterior distributions of a single simulation where data is generated and sampled from the stochastic volatility model with known parameters.
+    Figure 2 show the posterior distributions of a single simulation where data is generated and posteriors are sampled from the stochastic volatility model with known parameters.
 
@@
 \subsection{Research Goal}
-    The objective of this research is to design a simulation study to evaluate the calibration of algorithms used to sample Stochastic Volatility models.
-Indeed, there are limitations to evaluating MCMC algorithms based on fits to real data and single simulations.
+    The objective of this research is to design a simulation study to evaluate the calibration of algorithms used to estimate stochastic volatility models.
+As discussed, there are limitations to evaluating MCMC algorithms based on fits to real data and single simulations.
 To check the calibration of an algorithm, simulations over multiple true parameters are required.
@@
 
-    Additionally, two sampling strategies will be assessed using the proposed simulation design as well as the performance of the samplers under different parameterisations of the stochastic volatility model.
-Not only will the calibration of any one algorithm be assessed, but also compared across sampling strategies to determine which MCMC approach is most suitable for estimating this model.
+    Additionally, both the HMC and KSC algorithms will be assessed using the proposed simulation design as well as the performance of the samplers under different parameterisations of the stochastic volatility model.
+Results from the study will also be used to compare both sampling strategies to determine which MCMC approach is most suitable for estimating this model.
 
@@
 
-        To illustrate this procedure, let $\theta$ be a parameter and $y$ represent the dataset. Start with a single draw from the prior distribution:
+        To illustrate this procedure, let $\theta$ be an arbitrary parameter from a model and $y$ represent a dataset. Start with a single draw from the prior distribution:
 
@@
 
-        Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (Hamiltonian Monte Carlo or KSC) conditional on this dataset.
+        Then take draws from the posterior distribution generated by a MCMC algorithm (HMC or KSC) conditional on this dataset.
 
@@
 The procedure of generating posterior samples implicitly performs this integral since the expression on the right of the integral is proportional to the prior density.
-Therefore, any deviation of the posterior samples from the prior distribution suggests that the sampling methodology is not producing the correct posteriors.
-
-        If the posterior samples follows the prior distribution, the rank statistic for a given parameter follows a discrete uniform distribution\footnote{Proof of this result in \citet{talts2018validating}.}.
+Therefore, any deviation of the posterior distribution from the prior distribution means that the sampling methodology is not producing the correct posteriors.
+
+        \citet{talts2018validating} prove that the rank statistics for a given parameter follows a discrete uniform distribution if the posterior samples follow the same distribution as the prior.
 The rank statistic is defined as:
@@
 To complete the algorithm, multiple iterations are run and the rank statistics are calculated for each parameter.
-The resulting rank statistics are compared to the discrete uniform distribution to determine if any problematic features exist.
-
-        If the computation is well calibrated and the rank statistics follow a discrete uniform distribution, then the posterior credible intervals have sufficient coverage.
-That is, one way to describe calibration is: for any percentage interval selected over the posterior samples (for example 90\%) then there is a 90\% chance that $\theta^{sim}$ falls within this interval.
+The resulting rank statistics are compared to the discrete uniform distribution to determine if the algorithm is calibrated and returning the correct posteriors.
+
+        Posterior credible intervals are said to have sufficient coverage if Bayesian computation is well calibrated and the rank statistics follow a discrete uniform distribution.
+That said, there are two ways to interpret calibration.
+
+        \begin{enumerate}
+            \item For any percentage interval selected over the posterior samples (for example 90\%) then there is a 90\% chance that $\theta^{sim}$ falls within this interval.
+
+            \item Bayesian analysis is well calibrated if a 90\% credible interval contains the true parameter in 90\% of the SBC iterations. 
+        \end{enumerate}
+
+        % That is, one way to describe calibration is: for any percentage interval selected over the posterior samples (for example 90\%) then there is a 90\% chance that $\theta^{sim}$ falls within this interval.
 Another way of saying this is a Bayesian analysis is well calibrated if a 90\% credible interval contains the true parameter in 90\% of the SBC iterations.
@@
     \subsection{Evaluation Metrics}
-        The key metrics and diagnostics to compare the performance of these methods are rank statistics, and chi-squared test statistics to measure calibration and the effective sample size (ESS) to measure effciency.
+        The key metrics to compare the performance of these methods are rank statistics and chi-squared test statistics to measure calibration, and the effective sample size (ESS) to measure effciency.
 
         \subsubsection{Rank statistics}
-            As discussed in the SBC section, rank statistics are used to evaluate the calibration of the MCMC algorithm.
-If a posterior is well calibrated then it is expected that the distribution of rank statistics is uniform.
+            Rank statistics are used to evaluate the calibration of the MCMC algorithm. If a posterior is well calibrated then it is expected that the distribution of rank statistics is uniform.
 
@@
 Specifically, it gives information about the type of bias in the posterior estimates for a given parameter.
-Figures \ref{fig:underestimation} and \ref{fig:underdispersed} are recreated from \citet{talts2018validating} for conveience which gives some examples of the posterior estimates which result in bias of the rank statistics.
+Figures \ref{fig:underestimation} and \ref{fig:underdispersed} are recreated from \citet{talts2018validating} for convenience which display examples of posterior estimates that give biased rank statistics.
 
@@
 \citet{talts2018validating} describe this bias as underdispersion of the posterior relative to the prior distribuion.
-The estimated posterior is too narrow relative to the spread of the prior resulting in bias on both ends of the rank statistics distribution.
+The estimated posterior is too narrow relative to the spread of the prior resulting in bias on both ends of the rank statistic distribution.
 
@@
                 \caption{Left: Non uniform rank statistics with peak on the right side.
-Right: Posterior distribution vertical line representing true parameter.
-Bias in rank statistics distribution comes from disproportionate number of SBC iterations underestimating the true parameter.}
+Right: Posterior distribution is biased to the left side of the prior distribution.
+Bias in rank statistics distribution comes from disproportionate number of SBC iterations underestimating the true parameter.
+Note that the two plots are independently produced for illustration only.}
                 \label{fig:underestimation}
@@
             \subsubsection{Chi squared statistics}
-            A drawback from visualising the rank statistic distributions is there are more parameters (1003 parameters) than data points in this model.
+            A drawback from visualising the rank statistic distributions is there are more parameters and latent states to estimate than data points in this model.
 An alternative to visually checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each histogram bin.
-Let $b_j$ be the number of counts and $e_j$ the expected count in bin $j$.
+Let $b_j$ be the number of counts and $e_j$ the expected count in bin $j$ (where the expected count is a function of the number of bins and data points under a discrete uniform distribution).
 Then the chi squared statistic is given by:
@@
 
-            A perfectly discrete uniform distribution will return a $\chi^2$ statistic of zero.
-That is, the number of rank statistics in each bin is equal to the expected count under a discrete uniform distribution conditional on the number of bins and observations.
-Instead of performing multiple hypothesis tests for each parameter, the distribution of $\chi^2$ statistics is compared across all parameters and simulations.
+            A perfectly discrete uniform distribution will return a chi squared statistic of zero.
+That is, the number of rank statistics in each bin is equal to the expected count under a discrete uniform distribution.
+The distribution of chi squared statistics is visualised to compare results across multiple simulations.
 This will give a high level summary of calibration and overall performance of the algorithms.
@@
 $T$ is a trunction lag chosen such that the sum of any two consecutive autocorrelation lags is negative.
-This truncation (or some kind of downweighting) is required because a large number of lags leads to noiser autocorrelation estimates.
-More details about this and the calculation of the autocorrelation can be found in \citep{vehtari2021rank} and \citet{geyer1992practical}.
+This truncation (or some kind of downweighting) is required because a large number of lags leads to noisier autocorrelation estimates.
+More details about this ESS formula and the calculation of the autocorrelation can be found in \citep{vehtari2021rank} and \citet{geyer1992practical}.
 
@@
     \subsection{Sampling method 1}
-        (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable.
+        KSC sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable.
 So a more recent simulation smoother is used which is written by the same author of the original software.} \citep{dejong1995}.
@@
 The relationship between $y_t$ and $h_t$ is now linear; however, the error is not Gaussian.
-Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution.
+Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of Gaussians to approximate the first 4 moments of the log chi squared distribution.
 This is defined by:
@@
         The code used to replicate the KSC algorithm was written by \citet{chad2018} and ammended for the purposes of this research.
-Additional code was developed to produce rank statistics, quantities of interest as well as integrated into the SBC framework.
+Additional code was developed to produce rank statistics, quantities of interest as well as to integrate the MCMC into the SBC framework.
 
@@
 This is arbitrarily set at the 4th Gaussian density found in Appendix A.
+
+        For sampling the mixture model the Gaussian mixture density is rewritten with respect to a indicator variable $s_t$
+
+        $$
+        \begin{aligned}
+        &z_t | s_t = i \sim N(m_i - 1.2704, \nu^2) \\
+        &Pr(s_t = i) = q_i
+        \end{aligned}
+        $$
+
+        $s_t$ is then sambled from probability mass function:
+
+        $$
+        \begin{aligned}
+        Pr(s_t = i | y_t^{\ast}, h_t) \propto q_i f_N(y_t^{\ast} | h_t + m_t - 1.2704, \nu^2)
+        \end{aligned}
+        $$
 
@@
     \subsection{Sampling method 2}
-        Since the paper was written, new MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
-Specifically, Hamiltonian Monte Carlo is a MCMC algorithm which has become widely available for efficiently sampling from sophisticated models.
+        Hamiltonian Monte Carlo (HMC) is a MCMC algorithm which has become widely available for efficiently sampling sophisticated models.
 Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature \citep{duane1987hybrid} before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks \citep{neal1995bayesian} and statistical computing \citep{neal2011mcmc}.
@@
 
-        The key innovation of Hamiltonian Monte Carlo (HMC) is using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
+        The key innovation of HMC is using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
 HMC is able to reach more distant points with higher acceptance probabilites since proposals are made with information about the target distribution.
-Random walk beahviour is more inefficient in higher dimensions since it is more difficult to randomly select a point with a higher joint density and therefore higher acceptance probability.
+Random walk samplers are become inefficient in higher dimensions since it is more difficult to randomly select a point with a high acceptance probability.
 The increasing complexity of high dimensions make it harder for a random walk sampler to explore regions over long distances, resulting in the sampler to become stuck and requiring a lot of iterations to converge onto the target distribution.
@@
         A thorough conceptual and theoretical explanation of HMC can be found in \citet{gelman2013bayesian} and \citet{betancourt2017conceptual} with further details found in the aforementioned literature.
-In summary, HMC builds upon the Metropolis Hastings algorithm through the introduction of Hamiltonian equations, the momentum variable $\phi$ and the gradients of the target log posterior distribtuion.
+HMC builds upon the Metropolis Hastings algorithm through the introduction of Hamiltonian equations, auxiliary momentum variables $\phi$, and the gradients of the target log posterior distribtuion.
 The vector of model parameters $\theta$ is jointly sampled with a vector of momentum variables $\phi$.
-The momentum variables is also called auxiliary variables since they are required as part of the Hamiltonian equations to generate an efficient proposal but are not a quantity of interest for inference.
+The momentum variables are also called auxiliary variables since they are required as part of the Hamiltonian equations to generate an efficient proposal but are not a quantity of interest for inference.
 
@@
 These differential equations are solved using a discrete approximation algorithm, such as a leapfrog integrator, and is a function of the log posterior gradients for a given point $\theta$.
-The leapfrog integrator has a set of tuning parameters which determines the size and number of steps with the final step being the proposal value $(\theta^{\ast}, \phi^{\ast})$.
+The leapfrog integrator has a set of tuning parameters which determines the size and number of steps to be taken along the Hamiltonian path with the final step being the proposal value $(\theta^{\ast}, \phi^{\ast})$.
 Finally, the Metropolis Hastings accept/reject step is applied using the parameters at the beginning of the leapfrog process $(\theta^{t-1}, \phi^{t-1})$ and the proposal values $(\theta^{\ast}, \phi^{\ast})$.
+A formal description of the HMC algorithm is provided in Appendix C.
 
@@
 Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
-This approach will also use the same priors as specified in the Gaussian mixture approximation.
-A general description of the Hamltonian Monte Carlo algorithm is given below.
-
-        Let $\theta$ be a target parameter and $\phi$ be the auxiliary momemtum variable. The Hamiltonian is defined as:
-
-        $$
-        \begin{aligned}
-        H(\theta, \phi) \equiv - log \: \pi(\theta, \phi)
-        \end{aligned}
-        $$
-
-        This can be broken down into kinetic energy $K(\theta, \phi)$, the density over the auxiliary momentum, and potential energy $V(\theta)$, the density of the target posterior distribution.
-
-        $$
-        \begin{aligned}
-        H(\theta, \phi) &= - log \: \pi(\phi | \theta) - log \: \pi(\theta) \\ 
-        &\equiv  K(\theta, \phi) + V(\theta)
-        \end{aligned}
-        $$
-
-         Starting with an initial draw of $\theta$ (user defined or randomly generated), the HMC iteration or proposal is determined by 3 steps.
-
-        1) Randomly sample a momentum value
-
-        $$
-        \begin{aligned}
-        \phi\sim Multinormal(0, M)
-        \end{aligned}
-        $$
-
-        Where M is some diagnoal mass matrix (assuming independence between momentum variables).
-
-        2) Solve the set of Hamiltonian (differential) equations which generates a proposal $(\theta^{\ast}, \phi^{\ast})$
-
-        $$
-        \begin{aligned}
-        \frac{\partial \theta}{\partial t} &= \frac{\partial H}{\partial \phi} = \frac{\partial K}{\partial \phi} \\
-        \frac{\partial \phi}{\partial t} &= \frac{\partial H}{\partial \theta} = - \frac{\partial K}{\partial \theta} - \frac{\partial V}        {\partial \theta}
-        \end{aligned}
-        $$
-
-        Where $\frac{\partial V}{\partial \theta}$ is the gradient of the target posterior density.
-The solution to these differential equations is approximated using a Leapfrog integrator which gives a discrete approximate solution.
-The parameters which are tuned in this process are the number of steps $L$ and the size of the steps $\epsilon$.
-The Leapfrog algorithm performs a half step of $\phi$ and then a full step of the parameter $\theta$, and another half step of the momemtum variable $\phi$ and repeats this $L$ times.
-The final step is the proposal $(\theta^{\ast}, \phi^{\ast})$.
-
-        $$
-        \begin{aligned}
-        \phi &\leftarrow \phi - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta} \\
-        \theta &\leftarrow \theta + \epsilon M^{-1} \phi \\
-        \phi &\leftarrow \phi - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta}
-        \end{aligned}
-        $$
-
-        3) Metrpolis Hastings Accept/Reject
-        Let $(\theta^{t-1}, \phi^{t-1})$ be the values before the Leapfrog integrator.
-
-        $$
-        \begin{aligned}
-        r = \frac{\pi(\theta^{\ast} | y) \pi(\phi^{\ast})}{\pi(\theta^{t-1} | y) \pi(\phi^{t-1})}
-        \end{aligned}
-        $$
-
-        With sampled value
-
-        $$
-        \theta^t = \begin{cases}
-            \theta^{\ast},& \text{with probability } min(r,1)\\
-            \theta^{t-1}, & \text{otherwise}
-        \end{cases}
-        $$
+This approach will also use the same priors as specified in the KSC's Gaussian mixture approximation so that HMC is sampling from the same model (although other priors could be chosen).
 
@@
         HMC on the other hand is a more general approach to applying MCMC.
-The implementation of HMC is using the Stan programming language which allows for the specfication other generative or Bayesian models, not just stochastic volatility.
+The implementation of HMC is through the Stan programming language which allows for the specfication other generative or Bayesian models, not just stochastic volatility.
 The implementation details of HMC is abstracted away from the user, who only needs to focus on specifying the model using the Stan syntax.
@@
         The stochastic volatility model as described in section 2.1 is follows a centered parameterisation.
-This describes the central location of the latent state parameters which are centered around the mean and lag of the log volatility $\mu +\phi(h_t - \mu)$.
+This describes the central location of the latent state distribution which is centered on the mean of log volatility $\mu$.
 A model can be reparameterised in a way that improves MCMC performance.
@@
 
-        SBC is applied to reparameteirsied versions of the stochastic volatility model available to see if there is any improvement in calibration and efficiency.
+        SBC is applied to reparameteirsied versions of the stochastic volatility model to see if there is any improvement in calibration and efficiency.
 
@@
 Specifically, the states are first sampled from a standard normal distribution and are transformed to be samples from the state equation.
-As noted above, the centered parametersation samples the states directly on a non zero mean.
+As noted above, the centered parametersation samples the states centered around $\mu$.
 
@@
         \subsubsection{Non centered Gaussian Mixture model}
-        The reparameterised model for the Gaussian approximation is expressed differently due to the use of the Kalman Filter.
+        The reparameterised model for the Gaussian approximation is expressed differently due to the Kalman Filter.
 Reparameterised HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates.
@@
         % Different parameteristions express the same mathematical models in ways that make it easier or harder for a given algorithm to sample. 
-
-        A prior predictive simulation of the second state parameter is performed for both a centered and reparameterised model using the approach described in section 3.3.1.
+        A prior predictive simulation of the second state parameter is performed for both a centered and reparameterised model using the approach described in Section 4.4.1.
 This is to demonstrate that different parameteristions express the same underlying mathematical model.
-This is the same as the first two steps of SBC, except instead of generating a dataset, generate the parameters implied by the prior draw.
+This is similar the first two steps of SBC, except instead of generating a dataset, generate the state variable implied by the prior draw.
 Let $\boldsymbol{\theta}$ be a vector of static parameters.
+Simulate a draw from the joint prior:
 
@@
 
-        Generate a the second state parameter implied by the joint prior.
+        Generate a the second state parameter conditional on prior simulation and first state variable.
 
@@
 
-        Generating 1000 samples of $h_2^{sim}$ from both centered and non-centered models gives samples from the same data generating process as shown in Figure \ref{fig:priorpred}.
+        Simulating 1000 samples of $h_2^{sim}$ from both centered and non-centered models gives samples from the same data generating process as shown in Figure \ref{fig:priorpred}.
 
@@
             \includegraphics[scale=0.1]{figures/ppc_h2.png}
-            \caption{Prior predictive samples of second state variable}
+            \caption{Prior predictive simulation of second state variable from both centered and reparameterised stochastic volatility model.
+Both models are mathematically equivalent and generate samples from the same data generating process.}
             \label{fig:priorpred}
@@
 Figure \ref{fig:cphmc5k} and Figure \ref{fig:ncphmc5k} show the results of the same parameters with 5000 SBC iterations.
-There is an improvement to the reparameterised model with all histograms looking approximately uniform with much smaller variation around the true uniform value.
+There is an improvement to the reparameterised model with all histograms looking approximately uniform with much smaller variation around the uniform distribution.
 
@@
     \subsection{Gaussian Mixture Approximation}
-    The Gaussian Mixture Approximation is run for 1000 and 5000 SBC iterations with the first 10,000 samples discarded as burnin (following the method of \citet{kim1998stochastic}).
+    The Gaussian Mixture Approximation is run with the first 10,000 samples discarded as burnin (following the method of \citet{kim1998stochastic}).
 However, unlike KSC who take 750,000 posterior draws from the model, the number of posterior samples taken as part of the SBC procedure is 9,999.
@@
 There is a lack of uniformity across all parameters.
-In particular, there is a large peak on the left hand side for $\mu$ and the right hand hand side of $\phi$.
-This suggests over estimation and under estimation of the prior distribution respectively.
-All other parameters do not have any key features worth noting beyond noisy estimates around the discrete uniform distribution.
+In particular, there is a large bias on the left hand side for $\mu$ and the 1st and 500th latent state variables.
+There is also a bias on the right hand hand side of $\phi$.
+This suggests overestimation and underestimation of the prior distribution respectively.
+$\sigma^2$ and the 1000th latent state don't appear to be uniform although this may be due to noisy estimates.
 
@@
 $\mu$ has a large left peak suggesting the algorithm is over estimating the true parameter.
-The converse can be said about $\phi$ which possesses a peak on the right hand side which implies underestimation of the true parameter.}
+The converse can be said about $\phi$ which possesses a peak on the right hand side which implies underestimation of the true parameter.
+1st and 500th latent state variables also appear to be over estimating the true value.}
         \label{fig:cpksc1k}
@@
 The lack of uniformity for this model is distinctly due to a different shape.
-Whereas the lack of uniformity for previous SBC results with 1000 iterations may be due to noisy estimates.
+Whereas the lack of uniformity for some of the centered SBC results (for KSC) with 1000 iterations may be due to noisy estimates.
 
@@
 The rank statistics for all selected parameters are non uniform in shape.
-The bespoke MCMC struggles to return the correct posteriors for this parametersation of the model.}
+The KSC bespoke MCMC struggles to return the correct posteriors for this parametersation of the model.}
         \label{fig:ncpksc1k}
@@
     Increasing the SBC iterations to 5000 does not improve the results for either parametersation.
-$\sigma^2$ and the state parameters contain arguably less noisy estimates around the uniform distribution in the centered parametersation.
-However, the skewed shapes of $\mu$ and $\phi$ adds further evidence that the sampling strategy is not returning the correct posterior estimates.
+$\sigma^2$ contain arguably less noisy estimates around the uniform distribution in the centered parametersation (although there may be evidence of a slight right bias).
+The skewed shapes of $\mu$ and $\phi$ adds further evidence that the sampling strategy is not returning the correct posterior estimates.
+All latent state variables appear to have a left side bias.
 This can be seen in Figure \ref{fig:cpksc5k}.
@@
     Figure \ref{fig:ncpksc5k} shows no improvement to calibration results.
-The issues in calibration observed in Figure \ref{fig:ncpksc1k} are not due to noisy estimates and are likely the result of problems in the bespoke MCMC algorithm and model specification.
+The issues in calibration observed in Figure \ref{fig:ncpksc1k} are not due to noisy estimates and are likely the result of problems in the MCMC algorithm and model specification.
 Applying this parametersation and MCMC on real data will on balance produce incorrect posterior estimates.
@@
         \caption{5000 SBC iterations for centered Gaussian mixture approximation model.
-Estimates for $\sigma^2$ and the state parameters have marginally improved around the uniformity.
-However, there is no improvement in $\mu$ or $\phi$.}
+Estimates for $\sigma^2$ may have marginally improved around the uniformity there is still evidence of a ride side bias.
+There is no improvement in $\mu$ or $\phi$ and all latent state variables appear to have a left side bias.}
         \label{fig:cpksc5k}
@@
 The estimate for $\mu$ in the reparameterised model exhibits a long right tail (longer than the 10,000 posterior draws suggesting the presence of negatively autocorrelated draws).
-Overall, the bespoke MCMC is quite insufficient at sampling the static parameters.
+There is some improvement in efficiency from the reparameterised model with higher ESS values across most of the static parameter quantiles.
+Overall, the KSC bespoke MCMC is inefficient at sampling the static parameters.
 
@@
 
-    Overall the results from the KSC algorithm indicates some issues with calibration.
+    The results from the KSC algorithm indicates issues with calibration.
 The centered parameterastion appears to be more favourable for this sampling approach.
-Other avenues to improve the sampling of this model could be to consider a different bespoke sampling method or another parameterisation such as non centered in scale.
+Other ways of improving the sampling of this model using this MCMC strategy could be to use another parameterisation such as non centered in scale.
 
     There are a few potential reasons for the poor calibration results.
-It may be due to the inefficiency of the MCMC to generate the correct posteriors.
-Additionally, it could be due to the approximation of the actual stochastic volatility model not producing accurate parameter estimates.
-This is addressed using importance sampling weights by KSC and explored in the next section.
+It may be due to the ineffectiveness of the MCMC strategy to generate the correct posteriors.
+Additionally, it could be due to the approximation of the actual stochastic volatility model not producing accurate posterior estimates.
+Correcting potential approximation error of the model is explored in the next section.
 
@@
 The reweighting procedure ensures that samples are drawn from the correct posterior density.
-This is applied to the calculation of the expected value of the stochastic volatility posterior density using samples drawn from the Gaussian mixture approximation.
-
-    Importance weights are defined as the ratio of the joint posterior from the stochastic volatility model and the posterior distribution of the approximation.
+This is applied to the calculation of the expected value of the stochastic volatility posterior density using samples drawn from the Gaussian mixture model.
+
+    Importance weights are defined as the ratio of the joint posterior from the stochastic volatility model and the posterior distribution of the approximate model.
 A weight is produced for each posterior sample generated by MCMC.
@@
 This reweighting step is applied to the calculation of the rank statistics to see if the correction improves the calibration of the MCMC sampler.\footnote{Efficiency estimates of the reweighted posterior samples are omitted.
-It was unclear at the time of writing whether the ESS could be written as a function of the expectation or whether the ESS from resampled posterior samples using the importance weights were valid.}
+It was unclear at the time of writing whether the ESS could be written as a function of the expectation or whether the ESS from resampled posterior samples using the importance weights are valid.}
 
@@
 The expectation for any function of the posterior samples can be written as a function of these weights.
-Let $S(\theta)$ be an indicator random variable that is a function of the posterior samples, the expectation can be written as:
+Let $S(\theta)$ be an indicator random variable that is a function of the posterior samples.
+The expectation can be written as:
 
@@
     The results from applying this reweighting step to the rank statistics of the centered model are given on Figure \ref{fig:reweight1k}.
-There are no major improvements to the uniformity of rank statistics.
-In fact, there are now large peaks on the left side of the 1st and 500th state variables and a right peak for $\sigma^2$.
-Increasing the SBC iterations to 5000 as seen in Figure \ref{fig:reweight5k} also does not improve the results, with the left peak appearing to be larger for all state variables.
-Applying the reweighting step to the rank statistics appears to have weighted the posterior estimates to the right of the prior distribution.
-Overall the reweighting of the posterior samples from the approximate model did not improve the calibration results.
-The reweighted rank statistics for the reparameterised model also did not improve and can be found in Appendix C.
+There are no improvements in the distribution of rank statistics.
+Increasing the SBC iterations to 5000 as seen in Figure \ref{fig:reweight5k} also show no major improvements.
+The shape of both sets of histograms is consistent with the shape of the unweighted rank statistics.
+Overall the reweighting of the posterior samples from the approximate model does not improve the calibration results.
+The reweighted rank statistics for the reparameterised model also did not improve and can be found in Appendix D.
 
@@
     As discussed, it is not practical to inspect the histograms of all parameters and states in the stochastic volatility model.
-Instead, the Chi squared statistic can is used to summarise the shape of the rank statistic distribution.
+Instead, the chi squared statistic can is used to summarise the shape of the rank statistic distribution.
 This will be used to summarise the state variables only since the static parameters can be compared across simulations individually.
@@
     The state variable chi squared statistics for centered and reparamterised HMC, centered KSC and centered importance weighted KSC are reported as histograms in Figure \ref{fig:allchisq} (using results from 5000 SBC iterations).
-Reparameterised KSC is omitted as its poor SBC performance simply adds noise to the results.
+Reparameterised KSC is omitted as its poor SBC performance just adds noise to the results.
 A chi squared statistic of 0 implies the sample is exactly uniformly distributed.
-Any value away from 0 captures any variation or deviation away from uniformity.
+Any value away from 0 captures variation or deviations from uniformity.
 This provides a high level summary of the state variables - how close their rank statistics are to being uniformly distributed and their distance relative to the other algorithms and parameterisations.
@@
 \section{Discussion}
-Overall, Hamiltonian Monte Carlo applied on the reparameterised model performed best based on the calibration and efficiency evaluation criteria.
+Overall, Hamiltonian Monte Carlo applied on the reparameterised model performs best based on calibration and efficiency results.
 Centered parameterisation using HMC performed moderately well but struggled to produce calibrated posterior estimates for $\sigma^2$ and $\phi$ as well as generate independent samples for these parameters.
@@
 Whether or not to use tighter priors depends on the modelling context and the problem at hand.
-If the model is expected to handle datasets generated by a set of priors, then uncalibrated results from SBC raises questions about the suitability about the model choice, MCMC or parameterisation.
+If a model is expected to handle data generated by the priors, then uncalibrated results from SBC raises questions about the suitability about the model choice, MCMC or parameterisation.
 
@@
 The failure of non centered parameterisaion in location for the Gaussian mixture model could be due to a variety of reasons.
-This may be due to a particular software implementation when it comes to the Kalman Filter and Simulation Smoother (in which a specific software package was applied) that was not designed to handle the non centered parameterisations.
-Or perhaps this particular parameterisation doesn't suit the sampling strategy applied by KSC and another approach (e.g.
+This may be due to a particular software implementation when it comes to the Kalman Filter and Simulation Smoother (where a specific software package was applied) that was not designed to handle the non centered parameterisations.
+Or perhaps this particular parameterisation does not suit the sampling strategy applied by KSC and another approach (e.g.
 HMC) might be more successful.
@@
 Other features that may be of interest is the speed in which a MCMC can converge onto the target distribution and generate samples for inference.
-However, speed is diffiuclt to define and is subject to a variety of different factors such as hardware, operating system, etc.
+However, speed is diffiuclt to define and is subject to a variety of different factors such as hardware, operating system, and software versions.
 The purpose of this research is to compare algorithms based on calibration and efficiency.
@@
 Examples of this are posterior predictive checking and out of sample predictive performance.
-Rather, ths SBC approach gives insight into how well calibrated a sampling strategy is conditional on a known data generating process which is important in all applied contexts.
+Rather, the SBC approach gives insight into how well calibrated a sampling strategy is conditional on a known data generating process which is important in all applied contexts.
 
@@
 Producing results for this research within a reasonable timeframe required use of a computer cluster.
-Computing infrastructure may limit the opportunity for SBC to be performed,although it is probably not necessary to do 5000 SBC iterations in other research contexts.
+Computing infrastructure may limit the opportunity for SBC to be performed, although it is probably not necessary to do 5000 SBC iterations in other research contexts.
 The choice of SBC iterations for this research was arbitrary and was used to demonstrate that noisy estimates could be reduced by increasing the number of iterations for calibrated models.
-A more principlied approach to picking the number of SBC iterations could be explored in other research.
-
-Developing diagnostic tools for evaluating SBC results and models with many parameters is an area for improvement.
+A more principled approach to picking the number of SBC iterations could be explored in other research.
+
+Diagnostic tools for evaluating SBC results and models with many parameters is an area for improvement.
 As discussed, inspecting many histograms for highly parameterised models is unrealistic.
@@
 SBC may tell us some information about the miscalibration based on the histogram shape (e.g.
-under or over dispersed relative to the prior distribution), but understanding why this occurs may not be straight forward.
+under or over dispersed relative to the prior distribution), but understanding why this occurs may not be immediately clear.
 Although, diagnosing sampling problems with complicated posterior geometries speaks more generally to the difficulty of understanding complex models in the first place as opposed to a limitation of SBC.
@@
 In particular, the number of post burnin or warmup samples differed between algorithms (999 post burn-in for HMC, 9999 for the bespoke algorithm).
-An argument could be made that the poor results for the bespoke algorithm may be due to the chain not converging onto the target distribution.
+An argument could be made that the poor results for the bespoke KSC algorithm may be due to the chain not converging onto the target distribution.
 Indeed, the authors of this approach extended their Markov chain to 750,000 posterior draws.
-A farier comparison may be to have roughly similar number of ESS for most of the estimated parameters, although this is difficult to get right with multiple parameters.
-However, given how well Hamiltoninan Monte Carlo performed with the reparamterised model, any improvement to the Gaussian mixture model by increasing the length of the Markov chain would make calibration just as good, but not better due to the inefficiency of the sampler.
+A farier comparison may be to have roughly similar number of ESS for most of the estimated parameters, although this is difficult to get right with a large number of parameters and variables to estimate.
+Given how well Hamiltoninan Monte Carlo performed with the reparamterised model, any improvement to the Gaussian mixture model by increasing the length of the Markov chain would make calibration just as good, but not better due to the inefficiency of the sampler.
 
@@
 As our models and algorithms for estimating these models become more complicated, so does the need for principled ways of checking that they are returning correct posterior estimates.
-SBC provides a general simulation design for generative models to check whether a modelling strategy is producing the correct posterior estimates.
-
-In the context of stochastic volatility, applying Hamiltonian Monte Carlo on a reparameterised stochastic volatility model provided the most calibrated and efficient estimates when compared to the bespoke sampling method using a Gaussian mixture model.
-This also outperformed SBC and ESS results when using importance sampling weights to correct for approximation error.
-Other parameterisations of the stochastic volatility model can be considered to see if it improves sampling performance based on these metrics.
+SBC provides a general simulation design to check whether a modelling strategy is producing calibrated results for generative models.
+
+In the context of stochastic volatility, applying Hamiltonian Monte Carlo on a reparameterised stochastic volatility model provided the most calibrated and efficient estimates when compared with KSC's bespoke MCMC method.
+This also outperformed the KSC model using importance sampling weights to correct for approximation error.
+Other parameterisations of the stochastic volatility model can be considered to see if it improves sampling performance based on these diagnostics.
 
 This study only considered a handful of potential algorithms, sampling strategies and parameterisations for stochastic volatility.
-Future research could use the same SBC design to compare and evaluate more complicated volatility models (or any model outside this domain) and MCMC approaches.
-Overall, SBC can be applied as part of the development of various probabilistic and Bayesian models and is a useful tool as part of a statistical workflow.
+Future research could use the same SBC design to compare and evaluate more complicated volatility models and MCMC algorithms.
+Overall, SBC is a valuable tool in model development and is a key part of a statistical workflow.
 
@@
 
-\section{Appendix C: Importance weighted rank statistics for reparameterised model}
+\section{Appendix C: Hamiltonian Monte Carlo algorithm description}
+Let $\theta$ be a target parameter and $\phi$ be the auxiliary momemtum variable. The Hamiltonian is defined as:
+
+$$
+\begin{aligned}
+H(\theta, \phi) \equiv - log \: \pi(\theta, \phi)
+\end{aligned}
+$$
+
+This can be broken down into kinetic energy $K(\theta, \phi)$, the density over the auxiliary momentum, and potential energy $V(\theta)$, the density of the target posterior distribution.
+
+$$
+\begin{aligned}
+H(\theta, \phi) &= - log \: \pi(\phi | \theta) - log \: \pi(\theta) \\ 
+&\equiv  K(\theta, \phi) + V(\theta)
+\end{aligned}
+$$
+
+ Starting with an initial draw of $\theta$ (user defined or randomly generated), the HMC iteration or proposal is determined by 3 steps.
+
+1) Randomly sample a momentum value
+
+$$
+\begin{aligned}
+\phi\sim Multinormal(0, M)
+\end{aligned}
+$$
+
+Where M is some diagnoal mass matrix (assuming independence between momentum variables).
+
+2) Solve the set of Hamiltonian (differential) equations which generates a proposal $(\theta^{\ast}, \phi^{\ast})$
+
+$$
+\begin{aligned}
+\frac{\partial \theta}{\partial t} &= \frac{\partial H}{\partial \phi} = \frac{\partial K}{\partial \phi} \\
+\frac{\partial \phi}{\partial t} &= \frac{\partial H}{\partial \theta} = - \frac{\partial K}{\partial \theta} - \frac{\partial V}        {\partial \theta}
+\end{aligned}
+$$
+
+Where $\frac{\partial V}{\partial \theta}$ is the gradient of the target posterior density.
+The solution to these differential equations is approximated using a Leapfrog integrator which gives a discrete approximate solution.
+The parameters which are tuned in this process are the number of steps $L$ and the size of the steps $\epsilon$.
+The Leapfrog algorithm performs a half step of $\phi$ and then a full step of the parameter $\theta$, and another half step of the momemtum variable $\phi$ and repeats this $L$ times.
+The final step is the proposal $(\theta^{\ast}, \phi^{\ast})$.
+
+$$
+\begin{aligned}
+\phi &\leftarrow \phi - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta} \\
+\theta &\leftarrow \theta + \epsilon M^{-1} \phi \\
+\phi &\leftarrow \phi - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta}
+\end{aligned}
+$$
+
+3) Metrpolis Hastings Accept/Reject
+Let $(\theta^{t-1}, \phi^{t-1})$ be the values before the Leapfrog integrator.
+
+$$
+\begin{aligned}
+r = \frac{\pi(\theta^{\ast} | y) \pi(\phi^{\ast})}{\pi(\theta^{t-1} | y) \pi(\phi^{t-1})}
+\end{aligned}
+$$
+
+With sampled value
+
+$$
+\theta^t = \begin{cases}
+    \theta^{\ast},& \text{with probability } min(r,1)\\
+    \theta^{t-1}, & \text{otherwise}
+\end{cases}
+$$
+
+\section{Appendix D: Importance weighted rank statistics for reparameterised model}
 

### 2023-10-09 manuscript/final_presentation/presentation.qmd
final presentation draft

@@
-# Introduce stochastic volatility model
-# Research question
-# Motivating example 1 (two posterior estimates from two MCMC algorithms)
-# Motivating example 2 (Limitations of single simulation)
-# Methodology (SBC)
-# What is HMC
-# HMC results
-# What was KSC approach 
-# KSC results
-# Reparameterisation (HMC)
-# Reparam results 
-# Reparameterisation (KSC)
-# Reparam results (KSC)
-# Importance sampling weights
-# Discussion/limitations
-# Future research
+---
+title: "Comparison of MCMC algorithms in Stochastic Volatility Models using Simulation Based Calibration"
+format: 
+    beamer:
+        classoption:
+            - '`xcolor={dvipsnames}`{=latex}' 
+        header-includes: |
+            \setbeamertemplate{itemize items}[circle]
+author: "By Benjamin Wee"
+institute: "Supervised by Catherine Forbes and Lauren Kennedy"
+---
 
-- Stochastic Volatility model (Do i need to motivate this like last time?)
-- 2 Motivating examples. If two methods and they're different what do we do. 
-- Research Question
-- Methodology (how do we answer this. What iss a fair comparison? This is what we're going to do.)
-- HMC and KSC algorithms
-- Reparameterisation
-- Importance sampling reweighting (cut this? Yes)
-- Discussion and learnings
-- Limitations
-- Conclusion
+## Research Outline
+- Stochastic Volatility model
+- MCMC
 
-2 motivating examples: Don't have to get into nuts and bolts of diff algos.
-High level MCMC summary.
-General structure + bespoke.
-Stan can flexibly fit any model.
-It's a programming language.
-Bespoke only for SV model and how to best sample SV.
+    + Hamiltonian Monte Carlo
 
-In a simpler
+    + Kim Shephard & Chib (1998) bespoke MCMC strategy
 
- Start withmotivating example 1.
- Then go to movtivating exmaple 2. In this example.
+- Compare performance of MCMC
 
- When talking about SBC: remember motivating example, link it back.
-Thread the example through the presentation.
-Could even do distributions of posteriors vs true posteriors.
-Show results and link it back to the motivating example
+    + Simulation based Calibration: check correct posterior estimates from known data generating process
 
-In presentation use talts copy the plots. Mkae sure cites.
-Reproduce Talts data viz. Replicating examples from this paper are shown in these figures.
+    + \color{gray}{Efficiency: Effective sample size}
 
-Presentation: show results -> then show fig 4-6. Just say for convenience.
+- Model reparameterisation - improving MCMC performance
 
-Include a summary slide of extensions without going into detail. Give a sense of what else is in the paper. We explored different parameteirsations.
+- \color{gray}{Kim Shephard \& Chib Importance Weights - correcting approximation error}
 
-IS: Then they do an Importance Sampling. Wish i could talk about it but no time.
+## Log returns S&P 500
+![]("returns.png")
 
-Audience: Give them the impression that you done more than talked about.
-Made careful decision about what was more important to present.
-Lot's of content - but you need to understand what's int he presetation first.Done more than minimum.
+<!-- Suppose you're an analyst looking at this time series. Notice hetero behaviour/volatility clustering. Could try to model with SV -->
 
-Reparam: Don't go through steps or why they exist.
-There is a literature about MCMC efficiency if you reparam these models you can gain MCMC efficiency.
-Put the citations.
-There's lit but param is an issue.
-We look at a couple and these are results but don't go into detail.
+## Stochastic Volatility model
+KSC (1998) estimate a univariate discrete time SV model which models the variance as a latent stochastic process.
 
-Make sure i explicit: I'm really interested in finding out about xyz in context abc. Rsearch is comparing ABC and XyZ. State it early on. Say close to methods? So they can mark you.
+$$
+\begin{aligned}
+y_t =& \space \beta e^{h_t/2} \epsilon_t \\
+h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
+h_1 \sim&  N\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
+\epsilon_t \sim& N(0,1) \quad \eta_t \sim N(0,1) \\
+\end{aligned}
+$$
 
-In motivating example: say a little bit about why people are interested in this model.
-This is a difficult model.
-Ppl are interested GARCH are easy to estimate but don't capture properties of data observwd.
+- State space model
+- Complex! More parameters/unknowns then data points
+- Estimate Using Bayesian methods
 
-definitely harder on complicated model (not as intresting on simple model).
-Why we're looking at SV model in the first place.
-How to fit this model fast.
-**Check if its correct**.
-For this SV model in parrticular.
-Motivate this point.
-It's a hard model to fit.
+## Estimating the model on real data
 
-It could've gone either way. Might not have gone well stan. **Think about what a fair comparison is**
+. . .
 
-Particle filtering. There are other approaches. PF within MCMC.
+<!-- Keen analyst wants to estimate the model straight away. 2 MCMC methods.-->
+<!-- These are your posterior estimates. Which one is correct?-->
+<!-- Don't care about prediction yet. Care about accurate parameter estimates -->
 
-Jump straight into motivating examples.
-If you're Bayes, -> 2 different approx solutions.
-So we run both of them.
-Kind of similar but also different.
-pointness and fantess.
-How do we know which one to use.
+![]("../motivating_example/real_data_ex.png")
 
-Then go into SV. And we use it to do this. This is the model. Trying to work out the difference.
+## Estimate model on simulated data
 
-Get to movating example as soon as i can.
+. . .
 
-Start: What the example is. In title slide: prelim, title supervisors, slide gives a rough outline. Main sections of preso to set everyone. 
-Suppose i'm an analyst and I need to fit this model and estimate variance. Realise there's more than one way to do that. Check both and this is what i get. Then build onto SBC. Then confirm method.
+<!--Benefit of Bayesian model is that they're generative. Model can simulate data based on known parameter values.-->
+<!-- Pick a value for parameteres, simulate the data, estimate the model on simulated data. See if you can capture true parameters -->
+<!-- Phi not good! The model/algorithm doesn't estimate phi very well. -->
+<!-- Not true. 5% chance that any random draw from the true probabiltiy dist will fall outside the 95% interval.-->
 
-At a higher level talk about the model (not all priors etc).
-Just cite the literature on the slide.
-This is from KSC (cite earlier shephard paper).
-People have worked with it.
-Estimate, lots of parameters hard to estimate.
-Some methods exist here are two of them.
-Talk in detail later.
-Reassure them what the thing is.
-THis is the output.
+![]("../motivating_example/single_sim.png")
 
-Side by side histograms then full histgoram
+## Research goal
+- Real data
+    + Don't know much posterior estimate is correct
+    + Confounding causes if convergence diagnostics fail (incorrect model, poor MCMC choice)
+
+. . .
+
+- Single simulation
+    + Incorrect analysis may produce correct result by chance
+    + Correct analysis may produce poor result by chance
+
+. . .
+
+**Research objective**
+
+- Validate the computational methods used to estimate Bayesian stochastic volatility (SV) models
+
+- Design a simulation study to check of correct posterior estimates are returned
+
+- Compare Hamiltonian Monte Carlo with KSC (1998), which MCMC returns calibrated estimates?
+
+# Simulation Based Calibration
+
+## Simulation Based Calibration (SBC)
+Let $\theta$ be a parameter and $y$ represent the dataset. Start with draw a sample from the prior distribution:
+$$
+\begin{aligned}
+\theta^{sim} \sim \pi(\theta)
+\end{aligned}
+$$
+
+. . .
+
+Generate a dataset given by the prior draws.
+
+$$
+\begin{aligned}
+y^{sim} \sim \pi (y|\theta^{sim})
+\end{aligned}
+$$
+
+. . .
+
+Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (Hamiltonian Monte Carlo or Gaussian approximation in our case).
+
+$$
+\begin{aligned}
+\{\theta_1,\dots , \theta_{L}\} \sim \pi (\theta | y^{sim})
+\end{aligned}
+$$
+
+## Simulation Based Calibration (SBC)
+The rank statistic for a given parameter and simulation follows a discrete uniform distribution (proven in Talts et al. (2018)).
+
+$$
+\begin{aligned}
+r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim}) = \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}]
+\end{aligned}
+$$
+
+. . .
+
+Therefore, if the distribution of rank statistics deviates from a discrete uniform distribution, there may be evidence that our analysis or computation is not returning the correct posteriors (not calibrated).
+
+. . .
+
+If Bayesian computation is well calibrated, then the posterior (credible) intervals will have sufficient coverage.
+That is, for any percentage interval selected (for example 90%) then 90\% of these constructed intervals will contain $\theta^{sim}$.
+
+# Markov Chain Monte Carlo (MCMC) algorithms
+
+## Kim, Shephard, Chib (1998)
+
+$$
+\begin{aligned}
+y_t^{*} &= log(y_t^2) \\ 
+&= log((\epsilon_t exp(h_t/2))^2) \\
+&=  log(exp(h_t)) + log(\epsilon_t^2) \\
+&= h_t + log(\epsilon_t^2)  \\
+&= h_t + z_t \\
+\end{aligned}
+$$
+
+Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
+
+- Gaussian Mixture model
+- Kalman Filter
+- Conjugate posteriors
+- Metropolis Hastings
+
+## Hamiltonian Monte Carlo
+- The Stan programming language's implementation of Hamiltonian Monte Carlo, the No-U-Turn Sampler (Hoffman & Gelman 2014), will be compared with KSC's strategy.
+
+- Key innovations:
+
+    + Uses the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
+
+    + Allows for direct sampling of the specified stochastic volatility model
+
+    + Own programming language. Flexibly estimate any model
+
+# Results
+
+## HMC (1000 Iterations)
+
+## HMC (5000 Iterations)
+
+## KSC (1000 Iterations)
+
+## KSC (5000 Iterations)
+
+# Model reparameterisation
+
+## Reparameterised HMC (5000 Iterations)
+
+## Reparameterised KSC (5000 Iterations)
+
+# Discussion
+
+# Conclusion

### 2023-10-10 manuscript/final_presentation/presentation.qmd
updated images

@@
 <!-- Not true. 5% chance that any random draw from the true probabiltiy dist will fall outside the 95% interval.-->
+<!-- Gen models are great because we can simulate data from the process implied by the model conditional on parameters. -->
 
@@
 
-## Kim, Shephard, Chib (1998)
+## Markov Chain Monte Carlo (MCMC) algorithms
+
+**Kim, Shephard, Chib (1998)**
 
@@
 \begin{aligned}
-y_t^{*} &= log(y_t^2) \\ 
-&= log((\epsilon_t exp(h_t/2))^2) \\
-&=  log(exp(h_t)) + log(\epsilon_t^2) \\
-&= h_t + log(\epsilon_t^2)  \\
-&= h_t + z_t \\
+y_t &= \space e^{h_t/2} \epsilon_t \\
+y_t^{*} &= log(y_t^2) = h_t + z_t\\ 
 \end{aligned}
@@
 
-Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
-
-- Gaussian Mixture model
+- Approximate log chi squared error using Gaussian mixture model
 - Kalman Filter
@@
 
-## Hamiltonian Monte Carlo
-- The Stan programming language's implementation of Hamiltonian Monte Carlo, the No-U-Turn Sampler (Hoffman & Gelman 2014), will be compared with KSC's strategy.
+. . .
 
-- Key innovations:
+**Hamiltonian Monte Carlo**
 
-    + Uses the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
+- Own programming language. Flexibly estimate any model
 
-    + Allows for direct sampling of the specified stochastic volatility model
+- Uses the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
 
-    + Own programming language. Flexibly estimate any model
+- Allows for direct sampling of the specified stochastic volatility model
 
@@
 
-## HMC (1000 Iterations)
+## Results
+- 5000 SBC iterations
+- Simulated datasets of 1000 (which means 1000 latent states)
+- 999 posterior draws for HMC (1000 rank statistics per parameter)
+- 9,999 posterior draws for KSC (10,000 rank statistics per parameter)
 
 ## HMC (5000 Iterations)
-
-## KSC (1000 Iterations)
+![]("../results/hmc_cp_5k.png")
 
 ## KSC (5000 Iterations)
+![]("../results/ksc_cp_5k.png")
 
@@
 ## Reparameterised HMC (5000 Iterations)
+![]("../results/hmc_ncp_5k.png")
 
 ## Reparameterised KSC (5000 Iterations)
+![]("../results/ksc_ncp_5k.png")
 
-# Discussion
+## Interpretation of miscalibration
 
-# Conclusion
+## Discussion & Limitations
+- Reparameterised model with HMC performs best based on calibration
+
+- SBC is computationally intensive (required use of Monash HPC Cluster)
+
+- What constitutes a fair comparison? Reparameterised HMC calibrated after 999 draws.
+
+# Questions

### 2023-10-16 manuscript/main.tex

@@
 
-\title{Comparing MCMC algorithms in Stochastic Volatility Models using Simulation Based Calibration}
+\title{Comparing MCMC algorithms for Stochastic Volatility Model Bayesian Posteriors using Simulation Based Calibration}
 \author{Benjamin Wee}
@@
     This research compares the computational methods used to estimate Bayesian stochastic volatility models.
-The MCMC approaches outlined in the stochastic volatility literature and Hamiltoninan Monte Carlo are assesssed in their ability to sample from the model's posterior distribution.
+The MCMC approaches outlined in the stochastic volatility literature and Hamiltonian Monte Carlo are assessed in their ability to sample from the model's posterior distribution.
 Specifically, Simulation Based Calibration (SBC) is used to check whether these MCMC algorithms are returning efficient and calibrated posterior estimates.
@@
 This will determine which method is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
-Results reveal that Hamiltonian Monte Carlo provides more efficient and calibrated posterior estimates of the stochastic volatility model conditional on model paramterisation.
+Results reveal that Hamiltonian Monte Carlo provides more efficient and calibrated posterior estimates of the stochastic volatility model conditional on model parameterisation.
 \end{abstract}
@@
 \section{Introduction}
-    The stochastic volatility model is used in financial econometrics to model the behaviour of financial assets.
-They are typically expressed as non linear Gaussian state space models where the variance is treated as a latent random variable.
-Stochastic volatility is difficult to estimate since the likelihood is unavailable in closed form and there are at least as many variables as data points.
-Bayesian methods can be used to estimate these models which rely on Markov Chain Monte Carlo (MCMC) techniques to sample from the target joint posterior of these high dimensional parameter spaces.
-
-    \citet{kim1998stochastic} proposed a sampling strategy for stochastic volatility through a combination of conjugate posterior distributions, a Kalman filter, and the Metropolis Hastings algorithm.
+    Stochastic volatility (SV) models are used in financial econometrics to model and predict the behaviour of financial asset returns.
+Typically expressed as a non-Gaussian state space model where the conditional variance of the observed return is treated as a latent random variable \citep{hull1987pricing, chesney1989pricing}, such models are difficult to estimate since the likelihood is unavailable in closed form.
+With more parameters than observations,\footnote{We refer to both static unknowns and time-varying latent state variables as 'parameters'.} SV models are commonly analysed using Bayesian methods obtained via a Markov Chain Monte Carlo (MCMC) algorithm designed to sample from the high dimensional joint posterior distribution.
+The focus of this paper is to investigate the performance of two competing MCMC algorithms, each applied to a SV model, in terms of the validity and statistical efficiency of the their resulting inference.
+Performance of these algorithms will also be compared with different parameterisations of the SV model to determine the sensitivity of each MCMC to model specification.
+
+    \citet{kim1998stochastic} propose a sampling strategy for the so-called 'vanilla' stochastic volatility through an off-set mixture approximation.
+Conditional on the mixture allocations and static parameters, the forward filter backward sampling algorithm (Carter and Kohn, 1995; Fr{\"u}wirth-Schnatter, 1995) is used to jointly sample the latent states while conditionally  conjugate prior distributions are also suggested.
 Since then, there have been many advances in statistical computing and algorithm design.
-In particular, Hamiltonian Monte Carlo - a variant of the Metropolis Hastings algorithm - is an advancement in MCMC for efficient sampling of complex, high dimensional probabilistic models.
-The adoption of such new techniques have been made widely available through the development of various open source libraries such as the Stan programming language \citep{stan} and the PyMC library \citep{pymc2023}.
-Other algorithms have also improved the speed of estimating high dimensional models using approximations of the posterior such as integrated nested Laplace approximation \citep{rue2009approximate}.
-
-    These conceptually and practically different techniques to estimate complex models did not exist 20 years ago (or perhaps more accurately, were not easily accessible or implemented 20 years ago).
+In particular, Hamiltonian Monte Carlo (HMC) has be proposed as a generic and efficient sampling method for MCMC.
+Notably, HMC has been shown to be particularly well-suited to complex, high dimensional probabilistic models.
+Importantly, the adoption of such new techniques have been made widely available through the development of various open source libraries such as the Stan programming language \citep{stan} and the PyMC library \citep{pymc2023}.
 As the development of new algorithms rapidly increase, so does the need to develop new methods to assess their output.
-Developments in statistical workfow are required to test new algorithms as well as compare computational strategies used to estimate increasingly complex models.
-
-    This research conducts a simulation study to analyse the calibration and efficiency of MCMC used to estimate stochastic volatility models.
-The first algorithm replicates Kim Sherphard and Chib's (KSC) Gaussian mixture approximation which is sampled using the aforementioned Kalman Filter and Metropolis Hastings algorithm.
-The second algorithm is Hamiltonian Monte Carlo (HMC) as implemented in the Stan programming language.
+Developments in statistical workflow are required to test new algorithms as well as compare computational strategies used to estimate increasingly complex models.
+
+    This research analyses and compares the calibration and efficiency of MCMC used to estimate stochastic volatility models using a simulation study.
+By calibration we refer to a MCMC algorithm that returns correct posterior, on average, conditional on the correct model specification.
+The first algorithm replicates Kim Shephard and Chib's (KSC) Gaussian mixture approximation which is sampled using the aforementioned Kalman Filter and Metropolis Hastings algorithm.
+The second algorithm is the No-U-turn sampler (NUTS) variant of Hamiltonian Monte Carlo (HMC) as implemented in the Stan programming language \citep{hoffman2014no, betancourt2017conceptual, stan}.
 The objective is to analyse the efficiency of the MCMC and their ability to return correct posterior estimates.
 
-    Comparisons of these algorithms are conducted by Simulation Based Calibration \citep{talts2018validating}.
-Parameters are drawn from the prior distribution and used to create datasets from the generative stochastic volatility model which are sampled using one of the mentioned MCMC methods.
-Repeating this process multiple times gives insight to how well the MCMC algorithm can estimate the true parameters governing the data generating process.
-The key diagnostic metrics are the effective sample size to measure the efficiency of the algorithm and visual assessment of uniform rank statistics to determine calibration of the posterior estimates.
-
-    Additionally, two different parametiersations of the stochastic volatility model are considered for each sampling method.
+    Comparisons of these algorithms are conducted using a methodology known as Simulation Based Calibration (SBC) \citep{talts2018validating}.
+Parameters are drawn from the prior distribution and used to create data sets from the generative stochastic volatility model.
+We then obtain posterior samples from each of the MCMC methods.
+Repeating this process multiple times gives insight to how well the MCMC algorithm can estimate the true parameters given the data generating process.
+The key diagnostic metrics are the effective sample size, to measure the efficiency of the algorithm, and uniform rank statistics to determine calibration of the posterior estimates.
+
+    One particular challenge to comparing models is the parameterisation.
+Different parameterisations of a model, whilst the same mathematically, can be very different in terms of sampling difficulty \citep{neal2003slice}.
+In this work two different parameterisations of the stochastic volatility model are compared for each sampling method.
 An algorithm may be sensitive to different parameterisations of the same model which may impact MCMC performance as described in \citet{strickland2008parameterisation}.
-The key contribution of this research is to use simulations to assess whether a MCMC algorithm can return calibrated posterior estimates as well as the efficiency under the different parametiersations.
+The key findings of this research are HMC applied on a re-parameterised SV model produces the most calibrated and efficient posterior estimates and that the MCMC algorithms may perform better or worse based on calibration and efficiency depending on the parameterisation of the model.
 
     This paper is structured as follows.
-Section 2 provides the context around this research, namely the stochastic volatility model and the limitations of single simulation studies and MCMC convergence diagnostics.
-Section 3 describes the simulation design and diagnostic metrics.
-Section 4 details the sampling approaches.
-Section 5 discusses the results and limitations.
-Finally, section 6 concludes and provides points for further reserach.
-
-\section{Resarch Context}
+Section 2 provides the context around this research through the introduction of the stochastic volatility model and a discussion of the limitations of using ad-hoc simulations and MCMC convergence diagnostics to determine posterior calibration (or validity).
+Section 3 describes the SBC methodology, along with the simulation design and diagnostic metrics used for the SV model.
+Section 4 details key aspects of the two MCMC sampling approaches under consideration, while Section 5 discusses the results and limitations.
+Finally, Section 6 concludes and provides points for further research.
+
+\section{Research Context}
 
 \subsection{Stochastic Volatility}
-    The model of interest is the discrete time, univariate stochastic volatility model estimated by \citet{kim1998stochastic} using Bayesian methods.
-$y_t$ is the mean corrected returns of some asset for equally spaced intervals t.
-$\beta$ is a constant scaling factor which is also defined as $exp(\mu / 2)$ representing instantaneous volatility.
-$h_t$ is log volatility, where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the autoregressive parameter $\phi$ such that $|\phi|<1$.
-This autoregressive parameter represents the persistence or stickiness of log volatility and the dispersion parameter $\sigma_{\eta}$ is the constant variance of the states.
+    The model of interest is the discrete time, uni-variate stochastic volatility model estimated by \citet{kim1998stochastic} using Bayesian framework.
+SV is expressed as a state space model which links the observed mean corrected (logarithmic) return of a given asset to the latent conditional variance.
+    \begin{align}
+    y_t =& \space \exp(h_t/2) \varepsilon_t, \\
+    h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t,  \\
+    h_1 \sim& \space \,\mathrm{N}\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right), \\
+    \varepsilon_t \sim& \space \,\mathrm{N}(0,1), \\
+    \eta_t \sim& \space \,\mathrm{N}(0,1).
+    \end{align}
+    $y_t$ is the mean corrected (logarithmic) return of a given asset, observed over an interval with fixed length and associated with period t.
+The measurement equation (1) describes the conditional distribution of $y_t$ given $h_t$, where $h_t$ is the conditional log variance of $y_t$.
+$h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the auto-regressive parameter $\phi$ such that $|\phi|<1$.
+This auto-regressive parameter represents the persistence or stickiness of log volatility and the dispersion parameter $\sigma_{\eta}$ is the constant variance of the states.
 $\epsilon_t$ and $\eta_t$ are standard normal white noise shocks and are uncorrelated with each other.
 
-    $$
-    \begin{aligned}
-    y_t =& \space \beta exp(h_t/2) \epsilon_t \\
-    h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
-    h_1 \sim& \space normal\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
-    \end{aligned}
-    $$
-
-    $$
-    \begin{aligned}
-    \epsilon_t \sim& \space normal(0,1) \\
-    \delta_t \sim& \space normal(0,1)
-    \end{aligned}
-    $$
-
-    Setting $\beta=1$, the model can be expressed more succinctly as:
-
-    $$
-    \begin{aligned}
-    y_t \sim& \space normal(0, exp(h_t/2)) \\ 
-    h_1 \sim& \space normal \left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
-    h_{t+1} \sim& \space normal(\mu +\phi(h_t - \mu) , \sigma_{\eta}^2), \space\space t\neq 1\\ 
-    \end{aligned}
-    $$
-
-    Priors for the static parameters are defined below with conjugate priors on $\mu$ and $\sigma^2$:
-
-    $$
-    \begin{aligned}
-    \mu \sim& \space normal(0, 10^2) \\
-    \sigma_{\eta}^2 \sim& \space IG(5/2, (0.01\times 5) / 2) \\
-    \phi^{\ast} \sim& \space beta(20, 1.5) \\
-    \phi &=  2\phi^{\ast} - 1
-    \end{aligned}
-    $$
-
-    The prior on $\phi$ is a stretched beta distribution. This is a beta distribution (as defined on the parameter $\phi^*$) which has been transformed to have support (-1, 1).
-
-\subsection{Diagnostic limitations on real data}
+    KSC assign certain independent prior distributions and hyperparameters to the static parameters, with conjugate priors for $\mu$ and $\sigma^2$, given by
+    \begin{align}
+    \mu \sim& \space \,\mathrm{N}(0, 10^2) \\
+    \sigma_{\eta}^2 \sim& \space \, \mathrm{IG}(shape=5/2, scale=(0.01\times 5) / 2) \\
+    \phi^{\ast} \sim& \space \,\mathrm{Beta}(20, 1.5) \\
+    \phi =&  2\phi^{\ast} - 1
+    \end{align}
+    The prior distribution for $\phi$ is referred to as a stretched beta distribution, as it corresponds to a transformed beta random variable, $\phi^*$, with support (-1, 1).
+
+\subsection{Challenge 1: Comparing posteriors with no ground truth}
     The stochastic volatility model can be estimated using a variety of sampling strategies and MCMC algorithms.
-Convergence metrics are often used to check the performance of the MCMC chains, such as the effective sample size and the $\hat{R}$.
-These diagnostics provide evidence on whether a MCMC has failed to converge onto the target posterior distribution.
+Convergence metrics assessed from the same MCMC output to be used for posterior inference are commonly used to check the performance of the MCMC chains.
+Most often, the effective sample for each parameter is produced, as it indicates the number of approximately independent draws obtained from the dependent Markov chain for a target parameter \citep{gelman2013bayesian}, and $\hat{R}$, the potential scale reduction factor or measure of variability between different chains \citep{gelman1992inference}, is computed, to gauge whether the MCMC algorithm has converged to the target posterior distribution.
 
@@
 Real data is generated from an unknown data generating process.
-That is, the true parameter or model is unobservable.
-Therefore, failed diagnostic checks could arise from either inocrrect model specification or issues in the MCMC algorithm or both.
-Furthermore, different sampling strategies could provide different posterior estimates for the same model.
-Assuming no issues in computation, convergence diagnostics do not provide any information to which estimate is more correct or closer to the true parameter.
-
-    To illustrate this problem, the stochastic volatility model is estimated with two different MCMC algorithms on real data.
-The data are the de-meaned daily log returns of the 2023 January-September S\&P 500 Index.
-Figure \ref{fig:realdataex} visualises the marginal posterior estimates for the static parameters $\mu$, $\phi$ and $\sigma^2$ with corresponding summary statistics in Table \ref{tab:realdata}.
-
-    These estimates show the marginal posteriors of each parameter sharing similar shapes.
-$\phi$ has a fatter left tail for the KSC sampler and $\sigma^2$ is also heavier in the right tails relative to HMC.
+That is, the true parameter and model are unobservable, and it is very likely that the model being fit is not the true model.
+Therefore, failed diagnostic checks could be the result of attempting to fit a misspecified model, issues with the MCMC algorithm, or both.
+Furthermore, different sampling strategies will result in different posterior estimates for the same model.
+Assuming no issues with the computation or the model specification, convergence diagnostics do not provide any guidance to the analyst regarding which posterior estimate is to be preferred since there is no ground truth.
+
+    To illustrate the concerning issues, the posterior distribution associated with the stochastic volatility model in (1)-(5) and a set of real financial returns is estimated using two different MCMC algorithms.
+The data are demeaned daily (logarithmic) returns on the S\&P 500 Index over the period from 1 January 2023 - 10 September 2023 with sample size $T=171$.
+Figure \ref{fig:realdataex} displays the marginal posterior estimates for the static parameters $\mu$, $\phi$ and $\sigma^2$, with corresponding summary statistics in Table \ref{tab:realdata}.
+Overall the two methods are similar in skew and shape.
+However, closer inspections suggests that $\phi$ has a fatter left tail for the KSC sampler and $\sigma^2$ is also heavier in the right tails relative to HMC.
 This is evident in the quantiles with the 25th quantile for $\phi$ being smaller at 0.875 for KSC compared to 0.910 in HMC.
-Furthermore, HMC has higher peaks at their mode relative to KSC with a tighter spread.
-Based on this information alone, there is no certainty in whether method 1 or method 2 is has a more correct estimate conditional on data and model.
+Furthermore, HMC has taller peaks at their mode relative to KSC with a tighter spread.
+The uncertainty intervals created from each posterior sample would also be quite different and lead to different results.
+However, we can only conclude that there are differences between these methods, not which is correct.
 
@@
         \includegraphics[scale=0.1]{motivating_example/real_data_ex.png}
-        \caption{Posterior samples from the Hamiltonian Monte Carlo (HMC) and Kim Sherphard and Chib (KSC) MCMC samplers using S\&P 500 data.
-While the shape of the distributions are similar, the modes of HMC is much higher and the tails of KSC are fatter for $\phi$ and $\sigma^2$.
-Convergence diagnostics do not tell us which estimate is more correct since we don't observe the true model or true data generating process.}
+        \caption{\textbf{Posterior distribution for static parameters fit on the S\&P 500 Index.} Posterior samples from the HMC (blue) and KSC (red) MCMC samplers.}
         \label{fig:realdataex}
@@
         \end{tabular}
-        \caption{Summary statistics for HMC and KSC algorithms.
-KSC's 25th quantile $\phi$ is smaller than HMC which is consistent with the fatter left tail.
-Similarly, the KSC's 75th qunatile for $\sigma^2$ is larger consistent with a larger right tail spread.
-Both sets of estimate are reasonable but it is unclear from estimates on real data which estimate is closer the the truth.}
+        \caption{\textbf{Summary statistics for HMC and KSC algorithms}. Both sets of estimates appear reasonable but it is unclear from estimates on real data which estimate is closer the the truth.}
         \label{tab:realdata}
@@
 
-\subsection{Diagnostic limitations on a single simulation}
-    Diagnosing computational problems on real data is difficult due to confounding factors.
+\subsection{Challenge 2: Limitations of a single simulation}
+    Diagnosing problems with the approximate posterior from models fit to real data is difficult because we do not have a ground truth to compare with.
 A strategy around this is to evaluate a model and algorithm on simulated data.
@@
 
-    Furthermore, as discussed in \citet{gelman2020bayesian}, fitting models on simulated data is the only way we can check inference on latent variables.
-This is critical in stochastic volatility since the underlying framework is a state space model with latent log volatility parameters.
+    Simulation is particularly important in the SV model as it has latent parameters.
+\citet{gelman2020bayesian} discuss that simulation is the only way we can check inference on latent variables.
+This is critical for the SV model since the underlying framework is a state space model with latent log volatility parameters.
 Latent variables are unobserved in real data and are only estimated in the context of the model.
@@
 
-    Simulations enable us to check whether our models and algorithms can estimate the true data generating process.
+    Simulations enable the analyst to check whether a model and corresponding simulation algorithms appropriately estimate the `true' data generating process.
 However, there are limitations to what can be learned from a single simulation.
-There is always a small probability that the true parameter is in the tails of any estimated posterior distribution.
-Or put differently, there is a 1\% chance the true parameter or any random draw exists outside a 99\% credible interval.
+Even for a correctly specified model, there is always a small probability that the true parameter is in the tails of the corresponding (exact) marginal posterior distribution.
 \cite{talts2018validating} make the point that a single simulation does not provide sufficient information about the inference made by an algorithm.
@@
 
-    Figure 2 show the posterior distributions of a single simulation where data is generated and posteriors are sampled from the stochastic volatility model with known parameters.
-
-    \begin{figure}[h]
-        \centering
-        \includegraphics[scale=0.1]{motivating_example/single_sim.png}
-        \caption{Posterior samples from known data genearting process.
-Vertical solid line represents true parameter and dotted lines are 95\% credible intervals around the mean.
-The parameter $\phi$ falls outside the credible interval whereas $\mu$ and $\sigma^2$ stay inside.
-In any one simulation conditional on the correct model, there is a 5\% probability that the true parameter falls outside this interval by chance.}
-    \end{figure}
-
+    Figure 2 shows marginal posterior distributions given a single set of 1000 returns generated by the SV model.
+The true parameters are $\mu=-0.389$, $\sigma^2_{\nu}=0.0115$ and $\phi=0.547$ and the model is fit using Stan's implementation of HMC, the No-U-Turn sampler.
 The 95\% credible intervals for $\mu$ and $\sigma^2$ cover the true parameter.
 The true parameter for $\phi$ however, is in the tails and outside the interval.
-Such an analysis may incorrectly conclude that the model fails to adequatly estimate the $\phi$ parameter.
-However, it may be the case that the posterior distribution is correctly calculated using this algorithm and the results may be due to the features of this specific simulated dataset.
+Such an analysis may incorrectly conclude that the model fails to adequately estimate the $\phi$ parameter.
+However, it may be the case that the posterior distribution is correctly calculated using this algorithm and the results may be due to the features of this specific simulated data set.
+
+    \begin{figure}[h]
+        \centering
+        \includegraphics[scale=0.1]{motivating_example/single_sim.png}
+        \caption{\textbf{Posterior samples from SV model fit on known data generating process}.
+Vertical solid line represents true parameter and dotted lines are 95\% credible intervals around the mean.
+The parameter $\phi$ falls outside the credible interval whereas $\mu$ and $\sigma^2$ stay inside.}
+    \end{figure}
 
 \subsection{Research Goal}
-    The objective of this research is to design a simulation study to evaluate the calibration of algorithms used to estimate stochastic volatility models.
+    The objective of this research is to design a simulation study to evaluate the calibration of the HMC and KSC offset-mixture algorithms used to estimate SV models.
 As discussed, there are limitations to evaluating MCMC algorithms based on fits to real data and single simulations.
-To check the calibration of an algorithm, simulations over multiple true parameters are required.
+To check the calibration of an algorithm, repeated independent simulations are required.
 This methodology is discussed in the next section.
 
-    Additionally, both the HMC and KSC algorithms will be assessed using the proposed simulation design as well as the performance of the samplers under different parameterisations of the stochastic volatility model.
+    Two parameterisations of the SV model for each sampler will be implemented using this simulation design.
 Results from the study will also be used to compare both sampling strategies to determine which MCMC approach is most suitable for estimating this model.
@@
         Simulation Based Calibration (SBC) checks the calibration of posterior estimates generated by MCMC algorithms.
-SBC is conducted by comparing the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.
-The procedure starts by taking draws from the prior distribution and creating datasets implied by each draw.
-Rank statistics are then calculated on the posterior samples conditional on the simulated data.
-
-        To illustrate this procedure, let $\theta$ be an arbitrary parameter from a model and $y$ represent a dataset. Start with a single draw from the prior distribution:
-
-        $$
-        \begin{aligned}
-        \theta^{sim} \sim \pi(\theta)
-        \end{aligned}
-        $$
-
-        Generate a dataset given by the prior draw.
-
-        $$
-        \begin{aligned}
-        y^{sim} \sim \pi (y|\theta^{sim})
-        \end{aligned}
-        $$
-
-        Then take draws from the posterior distribution generated by a MCMC algorithm (HMC or KSC) conditional on this dataset.
-
-        $$
-        \begin{aligned}
-        \{\theta_1,\dots , \theta_{L}\} \sim \pi (\theta | y^{sim})
-        \end{aligned}
-        $$
-
-        A key result is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples $\theta^{sim}$. This is implied by the following expression:
-
-        $$
-        \begin{aligned}
-        \pi(\theta) &= \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})dy^{sim} d\theta^{sim} \\
-        &= \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) dy^{sim} d\theta^{sim}
-        \end{aligned}
-        $$
-
-        That is, the posterior averaged over the joint distribution follows the same distribution as the prior.
-The procedure of generating posterior samples implicitly performs this integral since the expression on the right of the integral is proportional to the prior density.
-Therefore, any deviation of the posterior distribution from the prior distribution means that the sampling methodology is not producing the correct posteriors.
-
-        \citet{talts2018validating} prove that the rank statistics for a given parameter follows a discrete uniform distribution if the posterior samples follow the same distribution as the prior.
-The rank statistic is defined as:
-
-        $$
-        \begin{aligned}
-        r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim}) = \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}]
-        \end{aligned}
-        $$
+SBC is conducted by comparing the distribution of rank statistics to the uniform distribution which arises when an algorithm is correctly calibrated.
+For each $k$ of $K$ SBC iterations, one value for each parameter is drawn from the prior.
+For this parameter a data set is generated from the model.
+Then a sampling method (originally using Stan in \citet{talts2018validating}) is used to obtain posterior samples given this data and the prior.
+Then the estimated posteriors are compared to the true values by calculating the rank.
+Under a calibrated model, the $K$ rank statistics for each parameter should be uniformly distributed.
+
+        To illustrate the SBC procedure, let $\theta$ be an arbitrary scalar parameter from a model and $y$ represent a data set.
+Start with a single parameter $\theta^{sim}$ drawn from the prior distribution:
+        \begin{align}
+        \theta^{sim} \sim \pi(\theta),
+        \end{align}
+        generate a data set $y^{sim}$ conditionally given $\theta^{sim}$ and the model of interest\footnote{Since all computation presumes a correctly specified model, nottation to condition on the model is suppressed.}:
+        \begin{align}
+        y^{sim} \sim \pi (y|\theta^{sim}).
+        \end{align}
+        Then take $L$ draws from the estimated posterior distribution, conditional on this data set, and generated by the MCMC algorithm under investigation (here, HMC or KSC):
+        \begin{align}
+        \{\theta_1,\dots , \theta_{B}\} \sim \pi (\theta | y^{sim}).
+        \end{align}
+
+        \citet{talts2018validating} prove that the sample of rank statistics for a given parameter, produced from replicate posteriors, follow a uniform distribution if the posterior samples follow the same distribution as the prior on average.
+The rank statistic for $\theta$ produced by the $k^{\mathrm{th}}$ replicate posterior sample, is defined as
+        \begin{align}
+        r^{(k)} = rank(\{\theta_1^{(k)},\dots , \theta_{L}^{(k)}\}, \theta^{sim}) = \sum_{b=1}^{B}1[\theta_{b}^{(k)} < \theta^{sim}].
+        \end{align}
 
@@
 To complete the algorithm, multiple iterations are run and the rank statistics are calculated for each parameter.
-The resulting rank statistics are compared to the discrete uniform distribution to determine if the algorithm is calibrated and returning the correct posteriors.
-
-        Posterior credible intervals are said to have sufficient coverage if Bayesian computation is well calibrated and the rank statistics follow a discrete uniform distribution.
-That said, there are two ways to interpret calibration.
-
-        \begin{enumerate}
-            \item For any percentage interval selected over the posterior samples (for example 90\%) then there is a 90\% chance that $\theta^{sim}$ falls within this interval.
-
-            \item Bayesian analysis is well calibrated if a 90\% credible interval contains the true parameter in 90\% of the SBC iterations. 
-        \end{enumerate}
+The resulting rank statistics are compared to the uniform distribution to determine if the algorithm is calibrated and returning the correct posteriors.
+
+        A key result is that the posterior averaged over the data and true parameters will equal the prior distribution. This is shown in the following expression:
+        \begin{align}
+        \pi(\theta) &= \int \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})d\theta^{sim} dy^{sim}, \\
+        &= \int \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) d\theta^{sim} dy^{sim}.
+        \end{align}
+        That is, the average posterior for $\theta$ over the SBC iterations should approximately equal to prior distribution, if the algorithm is calibrated.
+Therefore, any deviation from the prior distribution and the average posterior draws produced from replicated posterior distributions means that the sampling methodology is not producing the correct posterior, on average.
+
+        Posterior credible intervals are said to have sufficient coverage if Bayesian computation is well calibrated and the rank statistics follow a uniform distribution.
+That is, one way to describe calibration is: for any percentage interval selected over the posterior samples (for example 90\%) then there is a 90\% chance that $\theta^{sim}$ falls within this interval \citep{dawid1982well}.
 
@@
     \subsection{Evaluation Metrics}
-        The key metrics to compare the performance of these methods are rank statistics and chi-squared test statistics to measure calibration, and the effective sample size (ESS) to measure effciency.
+        In this research the following metrics are chosen:
+
+        \begin{enumerate}
+            \item Rank statistics to measure calibration for a given parameter and algorithm.
+
+            \item Chi squared statistics calculated over the $K$ rank statistics to summarise the shape of the distribution and deviation from uniformity.
+
+            \item Effective sample size (ESS) to measure MCMC efficiency
+        \end{enumerate}
 
         \subsubsection{Rank statistics}
-            Rank statistics are used to evaluate the calibration of the MCMC algorithm. If a posterior is well calibrated then it is expected that the distribution of rank statistics is uniform.
-
-            The shape of the rank statistic distribution gives insight into how a MCMC may be miscalibrated.
-Specifically, it gives information about the type of bias in the posterior estimates for a given parameter.
-Figures \ref{fig:underestimation} and \ref{fig:underdispersed} are recreated from \citet{talts2018validating} for convenience which display examples of posterior estimates that give biased rank statistics.
+            Rank statistics are used to evaluate the calibration of the MCMC algorithm.
+If a posterior is well calibrated then it is expected that the distribution of rank statistics over $B$ SBC iterations for each parameter in the model is uniform.
+
+            The shape of the distribution of rank statistics gives insight into how an MCMC may be miscalibrated.
+Specifically, it gives information about the type of bias in a marginal posterior estimate of a given parameter.
+Figures \ref{fig:underestimation} and \ref{fig:underdispersed} are recreated from \citet{talts2018validating} for convenience.
+In these we display examples of posterior sampling bias and the corresponding non-uniform rank statistics.
 
             Figure \ref{fig:underestimation} shows a distribution of rank statistics with a large right peak.
-This is consistent with posterior samples underestimating the true parameter, where the estimated posterior distrubtion is biased to the left of the prior.
-The sum of indicator random variables is large since the majority of posterior draws are smaller than the true value.
-This results in a large rank statistic and a bias on the right side of the histogram.
-The coverse is true if the peak of the histogram is on the left hand side (i.e the posterior is biased towards the right hand side of the prior).
+This distribution is consistent with posterior samples that tend to under estimate the true parameter on average.
+The sum of indicator random variables (see equation (13)) is large since a large proportion of posterior draws are smaller than the true value, resulting in a large rank statistic and hence a larger relative frequency on the right side of the histogram.
+The converse is true if the peak of the histogram is on the left hand side (i.e the average posterior over the SBC iterations is biased towards the right hand side of the prior).
 
             Figure \ref{fig:underdispersed} has large peaks on both ends of the histogram.
-The sampler is disproportionately over and under estimating the true parameter.
-\citet{talts2018validating} describe this bias as underdispersion of the posterior relative to the prior distribuion.
-The estimated posterior is too narrow relative to the spread of the prior resulting in bias on both ends of the rank statistic distribution.
+\citet{talts2018validating} describe this bias as under-dispersion of the posterior relative to the prior distribution on average.
+The estimated posterior is too narrow relative to the spread of the prior resulting in oversampling on both ends of the rank statistic distribution.
 
@@
 Right: Posterior distribution is biased to the left side of the prior distribution.
-Bias in rank statistics distribution comes from disproportionate number of SBC iterations underestimating the true parameter.
+Bias in rank statistics distribution comes from posterior samples underestimating the true parameter on average.
 Note that the two plots are independently produced for illustration only.}
@@
 Right: Overlayed posterior and prior distributions for the same arbitrary parameter.
-Bias in ranks comes from underdispersed posterior distribution relative to prior.}
+Non-Uniformity in ranks comes from under-dispersed posterior distribution relative to prior.}
                 \label{fig:underdispersed}
@@
             \subsubsection{Chi squared statistics}
-            A drawback from visualising the rank statistic distributions is there are more parameters and latent states to estimate than data points in this model.
-An alternative to visually checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each histogram bin.
-Let $b_j$ be the number of counts and $e_j$ the expected count in bin $j$ (where the expected count is a function of the number of bins and data points under a discrete uniform distribution).
-Then the chi squared statistic is given by:
-
-            $$
-            \begin{aligned}
-            \chi^2 = \sum_{j=1}^J \frac{(b_{j} - e_{j})^2}{e_j}
-            \end{aligned}
-            $$
-
-            A perfectly discrete uniform distribution will return a chi squared statistic of zero.
-That is, the number of rank statistics in each bin is equal to the expected count under a discrete uniform distribution.
+            One difficulty of relying on the visualisation of rank statistic distributions for SV models, is there are a very large number of unknowns in the model to consider.
+An alternative to visually checking for uniformity of a given parameter is to calculate a chi-squared statistic based on the counts in each histogram bin.
+Suppose the histogram has $J$ bins, where for each $j^{th}$ bin, $b_j$ is the number of counts and $e_j$ the expected count in bin $j$, where the expected count is the ratio of the number of data points and bins $e_j\frac{T}{J}$.
+The corresponding chi- squared statistic is given by:
+            \begin{align}
+            \chi^2 = \sum_{j=1}^J \frac{(b_{j} - e_{j})^2}{e_j}.\end{align}
+
+            A discrete empirical distribution with the same number of realisations in each of $J$ bins will return a chi squared statistic of zero, since the number realisations in each bin is equal to the expected count under the relevant  uniform distribution.
 The distribution of chi squared statistics is visualised to compare results across multiple simulations.
-This will give a high level summary of calibration and overall performance of the algorithms.
+Thus, the distribution of chi-squared statistics produced from the histograms of SBC rank statistics for a large number of univariate model parameters will give a high level summary of the overall model calibration, and hence for the overall performance of the algorithm.
+By comparing the distributions of chi-squared statistic produced by competing algorithms, one algorithm's degree of calibration, relative to that of the other algorithm, may be assessed.
 
             \subsubsection{Effective sample size}
-            ESS measures the efficiency of the MCMC sampler.
-It calculates the number of effectively independent draws generated by a MCMC to estimate a target parameter.
-A poor ESS can arise from high autocorrelation in the Markov chain which leads to highly dependent samples.
-An efficient MCMC algorithm takes less resources (for example, time and number of draws) to get a representative sample of the target distribution.
-If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), then there is evidence that this method is a more efficient sampler.
-
-            The ESS is defined as $n_{eff}$ and is a function of the number of chains $M$, the number of samples per chain $N$ and the autocorrelation at lag t $\rho_t$.
-$T$ is a trunction lag chosen such that the sum of any two consecutive autocorrelation lags is negative.
-This truncation (or some kind of downweighting) is required because a large number of lags leads to noisier autocorrelation estimates.
-More details about this ESS formula and the calculation of the autocorrelation can be found in \citep{vehtari2021rank} and \citet{geyer1992practical}.
-
-            $$
-            \begin{aligned}
-                n_{eff} = \frac{NM}{1+2 \sum_{t=1}^T \hat{\rho}_t}
+            Another univariate summary of a marginal posterior distribution is its effective sample size (ESS), which provides a measure of the efficiency of the MCMC sampler.
+ESS estimates the number of effectively independent draws generated by a MCMC to estimate a target parameter.
+A poor ESS generally arises from high auto-correlation in the Markov chain, and thus is an indication of a highly dependent posterior sample.
+A more efficient MCMC algorithm will take a relatively lower level of resources (for example, time and number of draws) to get a representative sample from the target distribution when compared to a less efficient algorithm.
+If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), this is seen as evidence of greater efficiency.
+
+            The ESS is denoted as $n_{\text{eff}}$.
+When more than one chain is generated from the target posterior, then the ESS quantity is a function of the number of chains $M$, the number of samples per chain $N$ and the auto-correlation at lag t $\rho_t$ \citep{vehtari2021rank}.
+$T$ is a truncation lag chosen such that the sum of any two consecutive auto-correlation lags is negative \citep{geyer1992practical}.
+This truncation (or some kind of down-weighting) is required because a large number of lags leads to noisier auto-correlation estimates.
+Note that this is only one approach to calculating the ESS.
+Details about this ESS formula, the calculation of the auto-correlation and other ESS calculations can be found in \citet{vehtari2021rank} and \citet{geyer1992practical}.
+
+            $$
+            \begin{aligned}
+                n_{\text{eff}} = \frac{NM}{1+2 \sum_{t=1}^T \hat{\rho}_t}
             \end{aligned}
@@
 
-    \subsection{Sampling method 1}
-        KSC sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother\footnote{In this research the exact software to apply the simulation smoother is unavailable.
-So a more recent simulation smoother is used which is written by the same author of the original software.} \citep{dejong1995}.
-
-        The standard Kalman Filter and simulation smoother is used to compute the posterior distribution over the latent states.
-This requires the state and measurement equations to be linear and conditionally Gaussian.
-Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring and taking the log of $y_t$.
-
-        $$
-        \begin{aligned}
-        y_t^{*} &= log(y_t^2) \\ 
-        &= log((\epsilon_t exp(h_t/2))^2) \\
-        &=  log(exp(h_t)) + log(\epsilon_t^2) \\
-        &= h_t + log(\epsilon_t^2)  \\
-        &= h_t + z_t \\
-        \end{aligned}
-        $$
-
-        Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
-The relationship between $y_t$ and $h_t$ is now linear; however, the error is not Gaussian.
-Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of Gaussians to approximate the first 4 moments of the log chi squared distribution.
-This is defined by:
-
-        $$
-        \begin{aligned}
-        f(z_t) = \sum_{i=1}^{K} q_if_N(z_i|m_i-1.2704, \nu_i^2)
-        \end{aligned}
-        $$
-
-        Where K is the mixture of normal densities $f_N$, component probabilities $q_i$, mean $m_i-1.2704$ and variance $\nu_i^2$.
-These parameters were selected using moment matching where they found 7 normal densities with varying mean and variance parameters best approximated the log chi squared moments.
-These parameters and weights can be found in Appendix A.
-
-        The model can be sampled via the Kalman Filter and simulation smoother since the model is now linear and conditionally gaussian.
+    \subsection{Sampling method 1: Mixture off-set MCMC}
+        KSC sample the conditional posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings within Gibbs algorithms and an implementation of the Kalman Filter and smoother used to sample from the conditional posterior distribution of the latent states \citep{dejong1995}.\footnote{In this research the exact software to apply the simulation smoother is unavailable.
+So a more recent simulation smoother is used which is written by the same author of the original software.} Note that for the simulation smoother to produce an appropriate posterior sample draw of the latent state vector, the state and measurement equations are required to be linear and conditionally Gaussian.
+Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring and taking the log of $y_t$, resulting in
+        \begin{align}
+        y_t^{*} &= log(y_t^2) \notag\\ 
+        &= log((\epsilon_t exp(h_t/2))^2) \notag\\
+        &=  log(exp(h_t)) + log(\epsilon_t^2) \notag\\
+        &= h_t + log(\epsilon_t^2) \notag  \\
+        &= h_t + z_t,  \label{eqn:transform}
+        \end{align}
+        Where $z_t = log(\epsilon_t^2)$ has a distribution that is like that of the logarithm of a chi-squared random variable.
+It is known that the error $z_t$ has mean -1.2704 and variance 4.93, however its distribution is not Gaussian.
+However, by transforming the entire measurement equation, as in (\ref{eqn:transform}), the relationship between $y_t$ and $h_t$ is now linear; however, the error is not Gaussian.
+Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of seven Gaussian distributions to approximate the distribution of $z_t$, with the mixture selected to match the first 4 moments of the log chi squared distribution.
+The mixture distribution is defined by:
+        \begin{align}
+        f(z_t) = \sum_{i=1}^{K} q_if_N(z_i|m_i-1.2704, \nu_i^2),
+        \end{align}
+        where $K=7$, $f_N(z|m_i-1.2704, \nu_i^2)$ denotes the density of the normal distribution with mean $m_i-1.2704$ and variance $v_i^2$, and  component probabilities $q_i$, for $k=1,2,\ldots, K=7$.
+The values of these hyperparameters and component weights are reported in Appendix A.
+
+        Given a set of $T$ mixture component indicator variables, which are added to the MCMC scheme in order to use the approximating Gaussian mixture error distribution, the latent state vector may be sampled via the Kalman Filter and simulation smoother, since the model is now conditionally linear and Gaussian.
 The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions whereas $\phi$ is sampled via a Metropolis Hastings accept/reject procedure.
-The details can be around in Appendix B.
+Details of the full sampling algorithm are found in Appendix B.
 
         \subsubsection*{Implementation}
-        The code used to replicate the KSC algorithm was written by \citet{chad2018} and ammended for the purposes of this research.
-Additional code was developed to produce rank statistics, quantities of interest as well as to integrate the MCMC into the SBC framework.
-
-        The steps in the MCMC process is desribed in Algorithm \ref{alg:ksc}.
+        The code used to replicate the KSC algorithm was written by \citet{chad2018} and amended for the purposes of this research.
+Additional code was developed to produce rank statistics, quantities of interest as well as to integrate the sampling method into the SBC framework.
+
+        The steps in the MCMC process is described in Algorithm \ref{alg:ksc}.
 The initial values are set to the values outlined in \citet{kim1998stochastic}, except for the mixing indicator which is unspecified.
@@
         For sampling the mixture model the Gaussian mixture density is rewritten with respect to a indicator variable $s_t$
-
-        $$
-        \begin{aligned}
+        \begin{align}
         &z_t | s_t = i \sim N(m_i - 1.2704, \nu^2) \\
-        &Pr(s_t = i) = q_i
-        \end{aligned}
-        $$
-
-        $s_t$ is then sambled from probability mass function:
-
-        $$
-        \begin{aligned}
-        Pr(s_t = i | y_t^{\ast}, h_t) \propto q_i f_N(y_t^{\ast} | h_t + m_t - 1.2704, \nu^2)
-        \end{aligned}
-        $$
+        &Pr(s_t = i) = q_i.
+        \end{align}
+        $s_t$ is then sampled from probability mass function: 
+        \begin{align}
+        Pr(s_t = i | y_t^{\ast}, h_t) \propto q_i f_N(y_t^{\ast} | h_t + m_t - 1.2704, \nu^2).
+        \end{align}
 
@@
                     \State \text{Sample states (Kalman Filter and Smoother): } $\boldsymbol{h}_i \sim h|y^{\ast}, s_{i-1}, \phi_{i-1}, \sigma^{2}_{\eta,i-1}, \mu^{i-1}$ 
-                    \State \text{Sample mixture density: } $s_i \sim s|y^{\ast}, \boldsymbol{h}_{i-1}$
-                    \State \text{Sample conjugate density $\mu$: } $\mu_i \sim \mu|y_{\ast}, s_{i-1}, \phi_{i-1}, \sigma^{2}_{\eta, i-1}, \boldsymbol{h}_{i-1}$
-                    \State \text{Sample conjugate density $\sigma^2_{\eta}$: } $\mu_i \sim \mu|y^{\ast}, s_{i-1}, \phi_{i-1}, \mu_{i-1}, \boldsymbol{h}_{i-1}$
-                    \State \text{Metrpolis Hastings step $\phi$: } $\phi_i \sim \phi|y^{\ast}, s_{i-1}, \mu_{i-1}, \sigma^{2}_{\eta, i-1}, \boldsymbol{h}_{i-1}$
+                    \State \text{Sample mixture indicators: } $s_i \sim s|y^{\ast}, \boldsymbol{h}_{i-1}$
+                    \State \text{Sample from conjugate density $\mu$: } $\mu_i \sim \mu|y_{\ast}, s_{i-1}, \phi_{i-1}, \sigma^{2}_{\eta, i-1}, \boldsymbol{h}_{i-1}$
+                    \State \text{Sample from conjugate density $\sigma^2_{\eta}$: } $\mu_i \sim \mu|y^{\ast}, s_{i-1}, \phi_{i-1}, \mu_{i-1}, \boldsymbol{h}_{i-1}$
+                    \State \text{Sample via Metropolis-Hastings $\phi$: } $\phi_i \sim \phi|y^{\ast}, s_{i-1}, \mu_{i-1}, \sigma^{2}_{\eta, i-1}, \boldsymbol{h}_{i-1}$
                   \EndFor
@@
 
-    \subsection{Sampling method 2}
-        Hamiltonian Monte Carlo (HMC) is a MCMC algorithm which has become widely available for efficiently sampling sophisticated models.
+    \subsection{Sampling method 2: Stan's HMC with No-U-Turn Sampler}
+        Hamiltonian Monte Carlo (HMC) is a  general MCMC algorithm used to efficiently sample posterior draws  high associated with high-dimensional and complex stochastic models.
 Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature \citep{duane1987hybrid} before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks \citep{neal1995bayesian} and statistical computing \citep{neal2011mcmc}.
-The algorithm has since become widely available through open source development projects such as Stan \citep{stan} and PyMC \citep{pymc2023}.
-
-        The key innovation of HMC is using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
-HMC is able to reach more distant points with higher acceptance probabilites since proposals are made with information about the target distribution.
-Random walk samplers are become inefficient in higher dimensions since it is more difficult to randomly select a point with a high acceptance probability.
-The increasing complexity of high dimensions make it harder for a random walk sampler to explore regions over long distances, resulting in the sampler to become stuck and requiring a lot of iterations to converge onto the target distribution.
-
-        A thorough conceptual and theoretical explanation of HMC can be found in \citet{gelman2013bayesian} and \citet{betancourt2017conceptual} with further details found in the aforementioned literature.
-HMC builds upon the Metropolis Hastings algorithm through the introduction of Hamiltonian equations, auxiliary momentum variables $\phi$, and the gradients of the target log posterior distribtuion.
+Versions of this algorithm have since become widely available through open source development projects, including Stan \citep{stan} and PyMC \citep{pymc2023}.
+
+        The key innovation of HMC is that it uses the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
+HMC is able to reach more distant points with higher acceptance probabilities than many bespoke MCMC algorithms since its  proposal draws are made using information about the target distribution.
+Random walk samplers (such as Random Walk Metropolis Hastings) tend to be inefficient in higher dimensions since it becomes more difficult to randomly select a point with a high acceptance probability.
+The increasing complexity of high dimensions make it harder for a random walk sampler to explore regions over long distances, and thus the sampler tends to become stuck and will require many iterations to converge to the target distribution.
+
+        A thorough conceptual and theoretical explanation of HMC can be found in \citet{gelman2013bayesian} and \citet{betancourt2017conceptual}.
+HMC builds upon the Metropolis Hastings algorithm through the introduction of Hamiltonian equations, auxiliary momentum variables $\phi$, and the gradients of the target log posterior distribution.
 The vector of model parameters $\theta$ is jointly sampled with a vector of momentum variables $\phi$.
@@
 Stan's default algorithm, the No-U-Turn Sampler \citep{hoffman2014no}, allows for direct sampling of the specified stochastic volatility model.
-Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
+The No-U-Turn Sampler (NUTS) proposes the final parameter using multinomial sampling biased towards the second half of the trajectory steps instead of applying the Metrpolis-Hastings step \citep{betancourt2016identifying}.
+Stan allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
+
         This approach will also use the same priors as specified in the KSC's Gaussian mixture approximation so that HMC is sampling from the same model (although other priors could be chosen).
@@
     \subsection{MCMC methods comparison}
-        The KSC sampling method and HMC are both MCMC strategies to genereate samples from the target joint posterior distribution of the stochastic volatility model.
+        The KSC sampling method and HMC are both MCMC strategies to generate samples from the target joint posterior distribution of the stochastic volatility model.
 However, the two sampling approaches are conceptually different in their design and implementation, despite sharing the same objective.
@@
         HMC on the other hand is a more general approach to applying MCMC.
-The implementation of HMC is through the Stan programming language which allows for the specfication other generative or Bayesian models, not just stochastic volatility.
+The implementation of HMC is through the Stan programming language which allows for the specification other generative or Bayesian models, not just stochastic volatility.
 The implementation details of HMC is abstracted away from the user, who only needs to focus on specifying the model using the Stan syntax.
-Any hyperparameter tuning and initialisation is automated away.
+Any hyper-parameter tuning and initialisation is automated.
 Furthermore, Stan and HMC allows for the specification of different priors, whereas KSC's bespoke algorithm is dependent on the use of conjugate priors.
 
-        The conceptual difference in both methods result in different tradeoffs.
+        The conceptual difference in both methods result in different trade-offs.
 The abstraction of the MCMC implementation using the Stan programming language means users can focus more on the modelling and the applied research question.
-The algorithm is maintained by other developers and different models can be compared with each other without worrying about differences in MCMC implementations.
+The algorithm is maintained and thoroughly tested by developers and different models can be compared with each other without worrying about differences in MCMC implementations.
 However, the abstraction away from the details means there is less flexibility in MCMC design in the scenario where the current instance of Stan or HMC is insufficient for a specific model.
@@
 
-        The stochastic volatility model as described in section 2.1 is follows a centered parameterisation.
+        The stochastic volatility model as described in section 2.1 follows a centered parameterisation.
 This describes the central location of the latent state distribution which is centered on the mean of log volatility $\mu$.
-A model can be reparameterised in a way that improves MCMC performance.
-Reparameterisation applies a transformation to the model which may change the interpretation of the parameters or the way they are sampled but remains mathematically equivalent to the original model.
-The reparametiersations available for a model is dependent on the MCMC approach.
-Some reparamterisations are available only to specific sampling strategies and may improve or impair a MCMC performance.
-
-        SBC is applied to reparameteirsied versions of the stochastic volatility model to see if there is any improvement in calibration and efficiency.
-
-        \subsubsection{Reparameterised stochastic volatility for Hamiltonian Monte Carlo}
-        The reparameterisation for Hamiltonian Monte Carlo changes how the latent states are sampled.
+A model can be re-parameterised in a way that might improve MCMC performance.
+Reparamaterisation applies a transformation to the model which may change the interpretation of the parameters or the way they are sampled but remains mathematically equivalent to the original model.
+The reparamaterisation available for a model is dependent on the MCMC approach.
+Some reparamaterisation are available only to specific sampling strategies and may improve or impair a MCMC performance.
+
+        SBC is applied to reparamaterised versions of the stochastic volatility model to see if there is any improvement in calibration and efficiency.
+
+        \subsubsection{Reparamaterised stochastic volatility in Stan}
+        The reparamaterisation for Hamiltonian Monte Carlo changes how the latent states are sampled.
 Specifically, the states are first sampled from a standard normal distribution and are transformed to be samples from the state equation.
-As noted above, the centered parametersation samples the states centered around $\mu$.
-
-        First sample from a standard normal distribuition and multiply by the variance of the log volatility.
+As noted above, the centered paramaterisation samples the states centered around $\mu$.
+
+        First sample from a standard normal distribution and multiply by the variance of the log volatility.
 The state vector is sampled with mean centered on zero and variance equal to $\sigma_{\eta}^2$
@@
         \begin{aligned}
-        h_{std} \sim& \space normal(0,1) \\
-        h =& h_{std} \times \sigma_{\eta} \\ 
-        h \sim& normal(0, \sigma_{\eta}^2)
-        \end{aligned}
-        $$
-
-        Then apply the appropriate rescaling to get samples from log volatility.
+        h_{std} \sim & \space normal(0,1) \\
+        h =& h_{std} \times \sigma_{\eta}
+        \end{aligned}
+        $$
+
+        Then apply the appropriate re-scaling to get samples from log volatility.
 
@@
 
-        $$
-        \begin{aligned}
-        h_1 \sim& \space normal \left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
-        h_{t+1} \sim& \space normal(\mu +\phi(h_t - \mu) , \sigma_{\eta}^2), \space\space t\neq 1\\ 
-        \end{aligned}
-        $$
-
         \subsubsection{Non centered Gaussian Mixture model}
-        The reparameterised model for the Gaussian approximation is expressed differently due to the Kalman Filter.
-Reparameterised HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates.
+        The re-parameterised model for the Gaussian approximation is expressed differently due to the Kalman Filter.
+Re-parameterised HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates.
 Applying the Kalman Filter requires the state equation to be a function of the state variable with the average log volatility parameter $\mu$ entering the measurement equation.
@@
         % Different parameteristions express the same mathematical models in ways that make it easier or harder for a given algorithm to sample. 
-        A prior predictive simulation of the second state parameter is performed for both a centered and reparameterised model using the approach described in Section 4.4.1.
-This is to demonstrate that different parameteristions express the same underlying mathematical model.
-This is similar the first two steps of SBC, except instead of generating a dataset, generate the state variable implied by the prior draw.
+        A prior predictive simulation of the second state parameter is performed for both a centered and re-parameterised model using the approach described in Section 4.4.1.
+This is to demonstrate that different parameterisations express the same underlying mathematical model.
+This is similar the first two steps of SBC, except instead of generating a data set, generate the state variable implied by the prior draw.
 Let $\boldsymbol{\theta}$ be a vector of static parameters.
@@
 
-        Simulating 1000 samples of $h_2^{sim}$ from both centered and non-centered models gives samples from the same data generating process as shown in Figure \ref{fig:priorpred}.
+        Simulating 5000 samples of $h_2^{sim}$ from both centered and non-centered models gives samples from the same data generating process as shown in Figure \ref{fig:priorpred}.
 
@@
             \includegraphics[scale=0.1]{figures/ppc_h2.png}
-            \caption{Prior predictive simulation of second state variable from both centered and reparameterised stochastic volatility model.
+            \caption{Prior predictive simulation of second state variable from both centered and re-parameterised stochastic volatility model.
 Both models are mathematically equivalent and generate samples from the same data generating process.}
@@
 \section{Results}
-    1000 and 5000 SBC iterations are run for centered and reparameterised versions of the stochastic volatility model using simulated datasets of 1000 observations.
+    1000 and 5000 SBC iterations are run for centered and re parameterised versions of the stochastic volatility model using simulated data sets of 1000 observations.
 Rank statistic histograms are constructed with 20 bins for the static parameters and the 1st, 500th and 1000th latent state variables.
@@
     \subsection{Hamiltonian Monte Carlo}
-    Figure \ref{fig:cphmc1k} and Figure \ref{fig:ncphmc1k} show the SBC results for the centered and reparameterisd model for HMC respectively.
+    Figure \ref{fig:cphmc1k} and Figure \ref{fig:ncphmc1k} show the SBC results for the centered and re-parameterised model for HMC respectively.
 Both sets of results display relatively noisy rank statistic estimates with variation around the discrete uniform distribution.
@@
 Rank statistics for $\sigma^2$ and $\phi$ show two peaks of either end of the histogram.
-The posterior estimates of $\sigma^2$ and $\phi$ are \textit{underdispersed} relative to the prior distribution.
+The posterior estimates of $\sigma^2$ and $\phi$ are \textit{under-dispersed} relative to the prior distribution.
 This implies the posterior geometry of these parameters may be difficult to sample for the HMC algorithm.
@@
         \includegraphics[scale=0.09]{results/hmc_cp_1k.png}
-        \caption{1000 SBC iterations for centered model using Hamiltoninan Monte Carlo.
+        \caption{1000 SBC iterations for centered model using Hamiltonian Monte Carlo.
 Histograms show noisy estimates around uniform distribution.
 In particular, $\sigma^2$ displays peaks on both ends of the histogram and $\phi$ shows a major on the left hand side.
-This suggests that the posterior samples of $\sigma^2$ are underdispersed relative to the prior distribution and posterior samples of $\phi$ are over estimating the true parameter.}
+This suggests that the posterior samples of $\sigma^2$ are under-dispersed relative to the prior distribution and posterior samples of $\phi$ are over estimating the true parameter.}
         \label{fig:cphmc1k}
@@
 
-    Estimates of $\sigma^2$ and $\phi$ are more uniform under the reparameterised model.
-Instead, we observe a peak on the left side for the 500th state variable $h_{500}$.
+    Estimates of $\sigma^2$ and $\phi$ are more uniform under the re-parameterised model.
+Instead, we observe a peak on the left side for the 500th state variable $h_{500}$, although the estimates are noisy.
 Overall the estimates for non centered parameterisation are more favourable.
@@
 Figure \ref{fig:cphmc5k} and Figure \ref{fig:ncphmc5k} show the results of the same parameters with 5000 SBC iterations.
-There is an improvement to the reparameterised model with all histograms looking approximately uniform with much smaller variation around the uniform distribution.
-
-    Centered parameteristion posess similar improvement for $\mu$ and the state variables.
-However, there are still major peaks at both ends of $\sigma^2$ and $\phi$ suggesting underdispersion of the posterior estimates relative to the prior distribution.
-
-    These results suggests that HMC struggles to sample some parameters from the centered parameteristion of the stochastic volatility model.
-The reparameterised model possess stronger evidence of calibration for the parameters examined, implying more reliable and consistent inference.
+There is an improvement to the re-parameterised model with all histograms looking approximately uniform with much smaller variation around the uniform distribution.
+
+    Centered parameterisation posses similar improvement for $\mu$ and the state variables.
+However, there are still major peaks at both ends of $\sigma^2$ and $\phi$ suggesting under-dispersion of the posterior estimates relative to the prior distribution.
+
+    These results suggests that HMC struggles to sample some parameters from the centered parameterisation of the stochastic volatility model.
+The re-parameterised model possess stronger evidence of calibration for the parameters examined, implying more reliable and consistent inference.
 
@@
         \includegraphics[scale=0.09]{results/hmc_cp_5k.png}
-        \caption{5000 SBC iterations for centered model using Hamiltoninan Monte Carlo.
+        \caption{5000 SBC iterations for centered model using Hamiltonian Monte Carlo.
 $\mu$ and the state parameters look approximately uniform.
-$\sigma^2$ still has major peaks on both the left and right hand side indicating underdispersion relative to the prior distribution.
+$\sigma^2$ still has major peaks on both the left and right hand side indicating under-dispersion relative to the prior distribution.
 $\phi$ exhibits a similar shape to $\sigma^2$ although the peaks are much smaller.}
@@
         \includegraphics[scale=0.09]{results/hmc_ncp_5k.png}
-        \caption{5000 SBC iterations for reparameterised model using Hamiltoninan Monte Carlo.
+        \caption{5000 SBC iterations for re-parameterised model using Hamiltonian Monte Carlo.
 Parameters display uniformity with less noise around the true uniform value.
@@
 
-    Additionally, HMC struggles to generate efficient posterior samples from the centered parameterisation of the stochastic volatility model.
+    This computational difficulty in the centered parameterisation is reflected in the effective sample sizes.
 Figure \ref{fig:hmcess} and Table \ref{tab:hmcess} summarises the distribution of ESS for the static parameters after 5000 SBC iterations.
 The non centered model appears to generate independent samples much more efficiently than the centered model.
-This is consistent with HMC struggliing to sample overall from the posteriors of $\sigma^2$ and $\phi$.
+This is consistent with HMC struggling to sample overall from the posteriors of $\sigma^2$ and $\phi$.
 
@@
         \includegraphics[scale=0.09]{results/hmc_ess.png}
-        \caption{Effective sample sizes for static parameters after 5000 iterations of SBC and 1000 post warmup samples for Hamiltoninan Monte Carlo.
+        \caption{Effective sample sizes for static parameters after 5000 iterations of SBC and 1000 post warm-up samples for Hamiltonian Monte Carlo.
 The centered parameterisation struggles to generate independent samples for $\sigma^2$ and $\phi$.
@@
         \end{tabular}
-        \caption{HMC: ESS for centered and reparameterised stochastic volatility model.}
+        \caption{ESS for centered and re-parameterised SV model across 5000 SBC iterations for HMC.}
         \label{tab:hmcess}
@@
     \subsection{Gaussian Mixture Approximation}
-    The Gaussian Mixture Approximation is run with the first 10,000 samples discarded as burnin (following the method of \citet{kim1998stochastic}).
+    The Gaussian Mixture Approximation is run with the first 10,000 samples discarded as burn-in (following the method of \citet{kim1998stochastic}).
 However, unlike KSC who take 750,000 posterior draws from the model, the number of posterior samples taken as part of the SBC procedure is 9,999.
@@
 
-    The both the centered and reparameterised Gaussian mixture model contains issues with calibration.
+    The both the centered and re-parameterised Gaussian mixture model contains issues with calibration.
 Figure \ref{fig:cpksc1k} presents the results from the centered model.
@@
 The rank statistics for all selected parameters are non uniform in shape.
-The KSC bespoke MCMC struggles to return the correct posteriors for this parametersation of the model.}
+The KSC bespoke MCMC struggles to return the correct posteriors for this parameterisation of the model.}
         \label{fig:ncpksc1k}
@@
 
-    Increasing the SBC iterations to 5000 does not improve the results for either parametersation.
-$\sigma^2$ contain arguably less noisy estimates around the uniform distribution in the centered parametersation (although there may be evidence of a slight right bias).
+    Increasing the SBC iterations to 5000 does not improve the results for either parameterisation.
+$\sigma^2$ contain arguably less noisy estimates around the uniform distribution in the centered parameterisation (although there may be evidence of a slight right bias).
 The skewed shapes of $\mu$ and $\phi$ adds further evidence that the sampling strategy is not returning the correct posterior estimates.
@@
 The issues in calibration observed in Figure \ref{fig:ncpksc1k} are not due to noisy estimates and are likely the result of problems in the MCMC algorithm and model specification.
-Applying this parametersation and MCMC on real data will on balance produce incorrect posterior estimates.
+Applying this parameterisation and MCMC on real data will on balance produce incorrect posterior estimates on average.
 
@@
 No improvement is observed in the distribution of rank statistics.
-The MCMC returns uncalibrated posterior estimates for this parametersation of the model.}
+The MCMC returns un-calibrated posterior estimates for this parameterisation of the model.}
         \label{fig:ncpksc5k}
@@
 
-    The ESS estimates for this model indicate a high degree of autocorrelation and difficulty in generating independent samples from both parameterisations.
+    The ESS estimates for this model indicate a high degree of auto-correlation and difficulty in generating independent samples from both parameterisations.
 The numerical results are summarised in Figure \ref{fig:kscess} and Table \ref{tab:kscess}.
-The estimate for $\mu$ in the reparameterised model exhibits a long right tail (longer than the 10,000 posterior draws suggesting the presence of negatively autocorrelated draws).
-There is some improvement in efficiency from the reparameterised model with higher ESS values across most of the static parameter quantiles.
+The estimate for $\mu$ in the re-parameterised model exhibits a long right tail (longer than the 10,000 posterior draws suggesting the presence of negatively auto-correlated draws).
+There is some improvement in efficiency from the re-parameterised model with higher ESS values across most of the static parameter quantiles.
 Overall, the KSC bespoke MCMC is inefficient at sampling the static parameters.
@@
         \includegraphics[scale=0.1]{results/ksc_ess.png}
-        \caption{Effective sample sizes for static parameters after 5000 iterations of SBC and 1000 post burnin samples for the Gaussian Mixture model.
-The bespoke sampling strategy struggles to generate independent samples for all static parameters, suggesting a high degree of autocorrelation in the Markkov Chain.}
+        \caption{Effective sample sizes for static parameters after 5000 iterations of SBC and 1000 post burn-in samples for the Gaussian Mixture model.
+The bespoke sampling strategy struggles to generate independent samples for all static parameters, suggesting a high degree of auto-correlation in the Markov Chain.}
         \label{fig:kscess}
@@
         \end{tabular}
-        \caption{KSC: ESS for centered and reparameterised stochastic volatility model.}
+        \caption{KSC: ESS for centered and re-parameterised stochastic volatility model.}
         \label{tab:kscess}
@@
     The results from the KSC algorithm indicates issues with calibration.
-The centered parameterastion appears to be more favourable for this sampling approach.
+The centered parameterisation appears to be more favourable for this sampling approach.
 Other ways of improving the sampling of this model using this MCMC strategy could be to use another parameterisation such as non centered in scale.
@@
     \citet{kim1998stochastic} correct for approximation error in their method by using importance weights.
-The reweighting procedure ensures that samples are drawn from the correct posterior density.
+The re-weighting procedure ensures that samples are drawn from the correct posterior density.
 This is applied to the calculation of the expected value of the stochastic volatility posterior density using samples drawn from the Gaussian mixture model.
@@
 
-    These weights can be applied in an additional sampling step using Metropolis Hastings or Importance Resampling (also known as sampling-importance resampling or SIR) to produce samples from the target posterior.
+    These weights can be applied in an additional sampling step using Metropolis Hastings or Importance Re-sampling (also known as sampling-importance re-sampling or SIR \citep{gelman2013bayesian}) to produce samples from the target posterior.
 However, the rank statistic can also be written as a function of the weighted expectation of the indicator random variable.
-This reweighting step is applied to the calculation of the rank statistics to see if the correction improves the calibration of the MCMC sampler.\footnote{Efficiency estimates of the reweighted posterior samples are omitted.
-It was unclear at the time of writing whether the ESS could be written as a function of the expectation or whether the ESS from resampled posterior samples using the importance weights are valid.}
-
-    \subsubsection{Reweigthing rank statistics of the Gaussian Mixture Model}
+This re-weighting step is applied to the calculation of the rank statistics to see if the correction improves the calibration of the MCMC sampler.\footnote{Efficiency estimates of the re-weighted posterior samples are omitted.
+It was unclear at the time of writing whether the ESS could be written as a function of the expectation or whether the ESS from re-sampled posterior samples using the importance weights are valid.}
+
+    \subsubsection{Re-weighting rank statistics of the Gaussian Mixture Model}
     Let $v(\theta, h)$ be the log weights defined as the log difference between the posterior densities of the true model with log chi squared errors $log\: g(\theta, h | y^{\ast}_t)$ and the Gaussian mixture model $log\:  k(\theta, h_t | y^{\ast}_t)$.
@@
     \begin{aligned}
-        v(\theta, h) = log\: g(\theta, h | y^{\ast}_t) - log\:  k(\theta, h_t | y^{\ast}_t) = const + log\: g(y|h) - log\: k(y^{\ast} | h)
+        v(\theta, h) = log\: g(\theta, h | y^{\ast}_t) - log\:  k(\theta, h_t | y^{\ast}_t) = \text{const} + log\: g(y|h) - log\: k(y^{\ast} | h)
     \end{aligned}
@@
 
-    Therefore, the expectation of the reweighted posterior samples is given by:
+    Therefore, the expectation of the re-weighted posterior samples is given by:
 
@@
 
-    The rank statistic can be rewritten as a function of the expectation and weights. This gives us the reweighted rank statistics.
+    The rank statistic can be rewritten as a function of the expectation and weights. This gives us the re-weighted rank statistics.
 
@@
 
-    The results from applying this reweighting step to the rank statistics of the centered model are given on Figure \ref{fig:reweight1k}.
+    The results from applying this re-weighting step to the rank statistics of the centered model are given on Figure \ref{fig:reweight1k}.
 There are no improvements in the distribution of rank statistics.
@@
 The shape of both sets of histograms is consistent with the shape of the unweighted rank statistics.
-Overall the reweighting of the posterior samples from the approximate model does not improve the calibration results.
-The reweighted rank statistics for the reparameterised model also did not improve and can be found in Appendix D.
+Overall the re-weighting of the posterior samples from the approximate model does not improve the calibration results.
+The re-weighted rank statistics for the re-parameterised model also did not improve and can be found in Appendix D.
 
@@
         \includegraphics[scale=0.1]{results/weighted_ksc_cp_1k.png}
-        \caption{1000 SBC iterations for reweighted rank statistics from the Gaussian mixture approximation model. The shape of the histogram is consistent with the unweighted rank statistics.}
+        \caption{1000 SBC iterations for re-weighted rank statistics from the Gaussian mixture approximation model. The shape of the histogram is consistent with the unweighted rank statistics.}
         \label{fig:reweight1k}
@@
         \includegraphics[scale=0.1]{results/weighted_ksc_cp_5k.png}
-        \caption{5000 SBC iterations for reweighted rank statistics from the Gaussian mixture approximation model. No major improvements are observed to the uniformity of rank statistics.}
+        \caption{5000 SBC iterations for re-weighted rank statistics from the Gaussian mixture approximation model. No major improvements are observed to the uniformity of rank statistics.}
         \label{fig:reweight5k}
@@
 
-    \subsection{Evaluating SBC for state parameters}
-    As discussed, it is not practical to inspect the histograms of all parameters and states in the stochastic volatility model.
-Instead, the chi squared statistic can is used to summarise the shape of the rank statistic distribution.
+    \subsection{Evaluating SBC for state variables}
+    It is not practical to inspect the histograms of all parameters and states in the stochastic volatility model.
+Instead, the chi squared statistic is used to summarise the shape of the rank statistic distribution.
 This will be used to summarise the state variables only since the static parameters can be compared across simulations individually.
 
-    The state variable chi squared statistics for centered and reparamterised HMC, centered KSC and centered importance weighted KSC are reported as histograms in Figure \ref{fig:allchisq} (using results from 5000 SBC iterations).
-Reparameterised KSC is omitted as its poor SBC performance just adds noise to the results.
+    The state variable chi squared statistics for centered and re-parameterised HMC, centered KSC and centered importance weighted KSC are reported as histograms in Figure \ref{fig:allchisq} (using results from 5000 SBC iterations).
+Re-parameterised KSC is omitted as its poor SBC performance just adds noise to the results.
 A chi squared statistic of 0 implies the sample is exactly uniformly distributed.
@@
 There is a large gap between the HMC and KSC chi squared statistics implying the distribution of state variable ranks are much less uniform for KSC relative to the HMC.
-Combined with the results in the previous section, there is a strong evidence that HMC outperforms KSC where it comes to calibration across all parmeters and latent state variables.
+Combined with the results in the previous section, there is a strong evidence that HMC outperforms KSC where it comes to calibration across all parameters and latent state variables.
 
@@
 \section{Discussion}
-Overall, Hamiltonian Monte Carlo applied on the reparameterised model performs best based on calibration and efficiency results.
+Overall, Hamiltonian Monte Carlo applied on the re-parameterised model performs best based on calibration and efficiency results.
 Centered parameterisation using HMC performed moderately well but struggled to produce calibrated posterior estimates for $\sigma^2$ and $\phi$ as well as generate independent samples for these parameters.
-Gaussian mixture approximation struggled to return calibrated posterior estimates and efficient samples for both parameteristions.
-However, the centered parameteristion results for the Gaussian mixture model may improve if we increased the length of the Markov chain.
+Gaussian mixture approximation struggled to return calibrated posterior estimates and efficient samples for both parameterisations.
+However, the centered parameterisation results for the Gaussian mixture model may improve if we increased the length of the Markov chain.
 This is discussed later in the limitations section.
 
-These results demonstrate that the performance of a MCMC algorithm is sensitive to the paramterisation of the model in the context of calibration and efficiency.
-Furthermore, the paramterisation of a model is conditional on the choice of MCMC algorithm.
-Evaluating the effectiveness of a modelling strategy (i.e choice of model, paramterisation and sampler) on simulated data gives a lot of a priori information about how the model will perform before it is fit on real data.
+These results demonstrate that the performance of a MCMC algorithm is sensitive to the parameterisation of the model in the context of calibration and efficiency.
+Furthermore, the parameterisation of a model is conditional on the choice of MCMC algorithm.
+Evaluating the effectiveness of a modelling strategy (i.e choice of model, parameterisation and sampler) on simulated data gives a lot of a priori information about how the model will perform before it is fit on real data.
 Performing SBC may help isolate confounding factors if any problems occur.
@@
 Whether or not to use tighter priors depends on the modelling context and the problem at hand.
-If a model is expected to handle data generated by the priors, then uncalibrated results from SBC raises questions about the suitability about the model choice, MCMC or parameterisation.
-
-Other parameterisaions of the stochastic volatility model are available as outlined in \citet{strickland2008parameterisation}, namely non centered in scale or both non centered in scale and location.
-The failure of non centered parameterisaion in location for the Gaussian mixture model could be due to a variety of reasons.
+If a model is expected to handle data generated by the priors, then un-calibrated results from SBC raises questions about the suitability about the model choice, MCMC or parameterisation.
+
+Other parameterisations of the stochastic volatility model are available as outlined in \citet{strickland2008parameterisation}, namely non centered in scale or both non centered in scale and location.
+The failure of non centered parameterisation in location for the Gaussian mixture model could be due to a variety of reasons.
 This may be due to a particular software implementation when it comes to the Kalman Filter and Simulation Smoother (where a specific software package was applied) that was not designed to handle the non centered parameterisations.
@@
 Other features that may be of interest is the speed in which a MCMC can converge onto the target distribution and generate samples for inference.
-However, speed is diffiuclt to define and is subject to a variety of different factors such as hardware, operating system, and software versions.
+However, speed is difficult to define and is subject to a variety of different factors such as hardware, operating system, and software versions.
 The purpose of this research is to compare algorithms based on calibration and efficiency.
@@
 In some cases, fitting just one model can take a long time, let alone multiple.
-Producing results for this research within a reasonable timeframe required use of a computer cluster.
-Computing infrastructure may limit the opportunity for SBC to be performed, although it is probably not necessary to do 5000 SBC iterations in other research contexts.
+Producing results for this research within a reasonable time-frame required use of a high performance cluster.
+Computing infrastructure may limit the opportunity for SBC to be performed.
 The choice of SBC iterations for this research was arbitrary and was used to demonstrate that noisy estimates could be reduced by increasing the number of iterations for calibrated models.
@@
 Other summary statistics and visualisations can be applied for a high level comparison, such as the chi squared statistics applied in this research.
-Furthermore, it is not always straightforward to understand why any any one parameter may produce poor calibration results, for example, an arbitrary state parameter.
+Furthermore, it is not always straightforward to understand why any any one parameter may produce poor calibration results, for example, an arbitrary latent state varaible.
 SBC may tell us some information about the miscalibration based on the histogram shape (e.g.
-under or over dispersed relative to the prior distribution), but understanding why this occurs may not be immediately clear.
+underdispersed, overdispersed, or biased relative to the prior distribution), but understanding why this occurs may not be immediately clear.
 Although, diagnosing sampling problems with complicated posterior geometries speaks more generally to the difficulty of understanding complex models in the first place as opposed to a limitation of SBC.
@@
 Lastly, it is unclear what constitutes a fair comparison between algorithms when comparing SBC results.
-In particular, the number of post burnin or warmup samples differed between algorithms (999 post burn-in for HMC, 9999 for the bespoke algorithm).
+In particular, the number of post burn-in or warm-up samples differed between algorithms (999 post burn-in for HMC, 9999 for the bespoke algorithm).
 An argument could be made that the poor results for the bespoke KSC algorithm may be due to the chain not converging onto the target distribution.
-Indeed, the authors of this approach extended their Markov chain to 750,000 posterior draws.
-A farier comparison may be to have roughly similar number of ESS for most of the estimated parameters, although this is difficult to get right with a large number of parameters and variables to estimate.
-Given how well Hamiltoninan Monte Carlo performed with the reparamterised model, any improvement to the Gaussian mixture model by increasing the length of the Markov chain would make calibration just as good, but not better due to the inefficiency of the sampler.
+Indeed, the authors of this approach extended their Markov chain to 750,000 posterior draws (discarding the first 10,000 draws as burn-in).
+A fairer comparison may be to have roughly similar number of ESS for most of the estimated parameters, although this is difficult to get right with a large number of parameters and variables to estimate.
+Given how well Hamiltonian Monte Carlo performed with the re-parameterised model, any improvement to the Gaussian mixture model by increasing the length of the Markov chain would make calibration just as good, but not better due to the inefficiency of the sampler.
 
@@
 
-In the context of stochastic volatility, applying Hamiltonian Monte Carlo on a reparameterised stochastic volatility model provided the most calibrated and efficient estimates when compared with KSC's bespoke MCMC method.
+In the context of stochastic volatility, applying Hamiltonian Monte Carlo on a re-parameterised stochastic volatility model provided the most calibrated and efficient estimates when compared with KSC's bespoke MCMC method.
 This also outperformed the KSC model using importance sampling weights to correct for approximation error.
@@
 
-\section{Appendix B: Sampling from mixture of Gaussians}
+\section{Appendix B: Sampling from mixture of Gaussian's}
 
@@
 \section{Appendix C: Hamiltonian Monte Carlo algorithm description}
-Let $\theta$ be a target parameter and $\phi$ be the auxiliary momemtum variable. The Hamiltonian is defined as:
+Let $\theta$ be a target parameter and $\phi$ be the auxiliary momentum variable. The Hamiltonian is defined as:
 
@@
 
-Where M is some diagnoal mass matrix (assuming independence between momentum variables).
+Where M is some diagonal mass matrix (assuming independence between momentum variables).
 
@@
 The parameters which are tuned in this process are the number of steps $L$ and the size of the steps $\epsilon$.
-The Leapfrog algorithm performs a half step of $\phi$ and then a full step of the parameter $\theta$, and another half step of the momemtum variable $\phi$ and repeats this $L$ times.
+The Leapfrog algorithm performs a half step of $\phi$ and then a full step of the parameter $\theta$, and another half step of the momentum variable $\phi$ and repeats this $L$ times.
 The final step is the proposal $(\theta^{\ast}, \phi^{\ast})$.
@@
 
-\section{Appendix D: Importance weighted rank statistics for reparameterised model}
+\section{Appendix D: Importance weighted rank statistics for re-parameterised model}
 
@@
     \includegraphics[scale=0.09]{results/weighted_ksc_ncp_1k.png}
-    \caption{1000 SBC iterations for reweighted rank statistics from the reparameterised Gaussian mixture approximation model.
+    \caption{1000 SBC iterations for re-weighted rank statistics from the re-parameterised Gaussian mixture approximation model.
 The shape of the histogram is consistent with the unweighted rank statistics.}
@@
     \includegraphics[scale=0.09]{results/weighted_ksc_ncp_5k.png}
-    \caption{5000 SBC iterations for reweighted rank statistics from the reparameterised Gaussian mixture approximation model. No major improvements are observed to the uniformity of rank statistics.}
+    \caption{5000 SBC iterations for re-weighted rank statistics from the re-parameterised Gaussian mixture approximation model.
+No major improvements are observed to the uniformity of rank statistics.}
     \label{fig:reweight5k}

### 2023-10-16 manuscript/main.tex
Fixed up maths spacing

@@
     \begin{align}
-    y_t =& \space \exp(h_t/2) \varepsilon_t, \\
-    h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t,  \\
+    y_t =& \space \,\exp(h_t/2) \varepsilon_t, \\
+    h_{t+1} =& \space \, \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t,  \\
     h_1 \sim& \space \,\mathrm{N}\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right), \\
@@
     \phi^{\ast} \sim& \space \,\mathrm{Beta}(20, 1.5) \\
-    \phi =&  2\phi^{\ast} - 1
+    \phi =& \space \, 2\phi^{\ast} - 1
     \end{align}
@@
         \begin{align}
-        r^{(k)} = rank(\{\theta_1^{(k)},\dots , \theta_{L}^{(k)}\}, \theta^{sim}) = \sum_{b=1}^{B}1[\theta_{b}^{(k)} < \theta^{sim}].
+        r^{(k)} = rank(\{\theta_1^{(k)},\dots , \theta_{B}^{(k)}\}, \theta^{sim}) = \sum_{b=1}^{B}1[\theta_{b}^{(k)} < \theta^{sim}].
         \end{align}
@@
             Figure \ref{fig:underestimation} shows a distribution of rank statistics with a large right peak.
-This distribution is consistent with posterior samples that tend to under estimate the true parameter on average.
+This distribution is consistent with posterior samples that tend to underestimate the true parameter on average.
 The sum of indicator random variables (see equation (13)) is large since a large proportion of posterior draws are smaller than the true value, resulting in a large rank statistic and hence a larger relative frequency on the right side of the histogram.
@@
 
-            Figure \ref{fig:underdispersed} has large peaks on both ends of the histogram.
+            Figure \ref{fig:underdispersed} has large peaks on both ends of the histogram resulting in a `U-shape'.
 \citet{talts2018validating} describe this bias as under-dispersion of the posterior relative to the prior distribution on average.
@@
 An alternative to visually checking for uniformity of a given parameter is to calculate a chi-squared statistic based on the counts in each histogram bin.
-Suppose the histogram has $J$ bins, where for each $j^{th}$ bin, $b_j$ is the number of counts and $e_j$ the expected count in bin $j$, where the expected count is the ratio of the number of data points and bins $e_j\frac{T}{J}$.
+Suppose the histogram has $J$ bins, where for each $j^{th}$ bin, $b_j$ is the number of counts and $e_j$ the expected count in bin $j$, where the expected count is the ratio of K rank statistics and J bins, $e_j=\frac{K}{J}$.
 The corresponding chi- squared statistic is given by:
@@
             The ESS is denoted as $n_{\text{eff}}$.
-When more than one chain is generated from the target posterior, then the ESS quantity is a function of the number of chains $M$, the number of samples per chain $N$ and the auto-correlation at lag t $\rho_t$ \citep{vehtari2021rank}.
-$T$ is a truncation lag chosen such that the sum of any two consecutive auto-correlation lags is negative \citep{geyer1992practical}.
+When more than one chain is generated from the target posterior, then the ESS quantity is a function of the number of chains $M$, the number of samples per chain $N$ and the auto-correlation at lag $l$ $\rho_l$ \citep{vehtari2021rank}.
+$L$ is a truncation lag chosen such that the sum of any two consecutive auto-correlation lags is negative \citep{geyer1992practical}.
 This truncation (or some kind of down-weighting) is required because a large number of lags leads to noisier auto-correlation estimates.
@@
             \begin{aligned}
-                n_{\text{eff}} = \frac{NM}{1+2 \sum_{t=1}^T \hat{\rho}_t}
+                n_{\text{eff}} = \frac{NM}{1+2 \sum_{l=1}^L \hat{\rho}_l}
             \end{aligned}
@@
     \subsection{Sampling method 1: Mixture off-set MCMC}
-        KSC sample the conditional posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings within Gibbs algorithms and an implementation of the Kalman Filter and smoother used to sample from the conditional posterior distribution of the latent states \citep{dejong1995}.\footnote{In this research the exact software to apply the simulation smoother is unavailable.
-So a more recent simulation smoother is used which is written by the same author of the original software.} Note that for the simulation smoother to produce an appropriate posterior sample draw of the latent state vector, the state and measurement equations are required to be linear and conditionally Gaussian.
+        \citet{kim1998stochastic} sample the conditional posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings within Gibbs algorithms and an implementation of the Kalman Filter and smoother used to sample from the conditional posterior distribution of the latent states \citep{dejong1995}.\footnote{In this research the exact software to apply the simulation smoother is unavailable.
+So a more recent simulation smoother is used which is written by the same author of the original software.} Note that for the simulation smoother to produce an appropriate posterior sample draw from the latent state vector, the state and measurement equations are required to be linear and conditionally Gaussian.
 Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring and taking the log of $y_t$, resulting in
@@
         \begin{align}
-        f(z_t) = \sum_{i=1}^{K} q_if_N(z_i|m_i-1.2704, \nu_i^2),
+        f(z_t) = \sum_{i=1}^{I} q_if_N(z_t|m_i-1.2704, \nu_i^2),
         \end{align}
-        where $K=7$, $f_N(z|m_i-1.2704, \nu_i^2)$ denotes the density of the normal distribution with mean $m_i-1.2704$ and variance $v_i^2$, and  component probabilities $q_i$, for $k=1,2,\ldots, K=7$.
+        where $K=7$, $f_N(z_t|m_i-1.2704, \nu_i^2)$ denotes the density of the normal distribution with mean $m_i-1.2704$ and variance $v_i^2$, and  component probabilities $q_i$, for $i=1,2,\ldots, I=7$.
 The values of these hyperparameters and component weights are reported in Appendix A.
@@
 The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions whereas $\phi$ is sampled via a Metropolis Hastings accept/reject procedure.
-Details of the full sampling algorithm are found in Appendix B.
+
+        Since the mixture of Gaussians only approximates the distribution of $\log \epsilon_t$ up to its first four moments, \citet{kim1998stochastic} also recommend using a post-sampling importance sampling step to correct for approximation error.
+The importance weights used in this step are also applied in the context of SBC and can be found in Section 5.3.
 
@@
 Additional code was developed to produce rank statistics, quantities of interest as well as to integrate the sampling method into the SBC framework.
-
-        The steps in the MCMC process is described in Algorithm \ref{alg:ksc}.
-The initial values are set to the values outlined in \citet{kim1998stochastic}, except for the mixing indicator which is unspecified.
-This is arbitrarily set at the 4th Gaussian density found in Appendix A.
-
-        For sampling the mixture model the Gaussian mixture density is rewritten with respect to a indicator variable $s_t$
-        \begin{align}
-        &z_t | s_t = i \sim N(m_i - 1.2704, \nu^2) \\
-        &Pr(s_t = i) = q_i.
-        \end{align}
-        $s_t$ is then sampled from probability mass function: 
-        \begin{align}
-        Pr(s_t = i | y_t^{\ast}, h_t) \propto q_i f_N(y_t^{\ast} | h_t + m_t - 1.2704, \nu^2).
-        \end{align}
-
-        \begin{algorithm}[H]
-            \caption{KSC MCMC Algorithm}\label{alg:ksc}
-            \begin{algorithmic}
-            \Require $s_0 = 4$, $\mu_0 = 0$, $\phi_0 = 0.95$, $\sigma^{2}_{\eta,0} = 0.02$
-            \For{\texttt{i in} $1:n_{draws}$}
-                    \State \text{Sample states (Kalman Filter and Smoother): } $\boldsymbol{h}_i \sim h|y^{\ast}, s_{i-1}, \phi_{i-1}, \sigma^{2}_{\eta,i-1}, \mu^{i-1}$ 
-                    \State \text{Sample mixture indicators: } $s_i \sim s|y^{\ast}, \boldsymbol{h}_{i-1}$
-                    \State \text{Sample from conjugate density $\mu$: } $\mu_i \sim \mu|y_{\ast}, s_{i-1}, \phi_{i-1}, \sigma^{2}_{\eta, i-1}, \boldsymbol{h}_{i-1}$
-                    \State \text{Sample from conjugate density $\sigma^2_{\eta}$: } $\mu_i \sim \mu|y^{\ast}, s_{i-1}, \phi_{i-1}, \mu_{i-1}, \boldsymbol{h}_{i-1}$
-                    \State \text{Sample via Metropolis-Hastings $\phi$: } $\phi_i \sim \phi|y^{\ast}, s_{i-1}, \mu_{i-1}, \sigma^{2}_{\eta, i-1}, \boldsymbol{h}_{i-1}$
-                  \EndFor
-            \end{algorithmic}
-            \end{algorithm}
-
-    \subsection{Sampling method 2: Stan's HMC with No-U-Turn Sampler}
+Details of the full sampling algorithm are found in Appendix B.
+
+    \subsection{Sampling method 2: Stan's Hamiltonian Monte Carlo with No-U-Turn Sampler}
         Hamiltonian Monte Carlo (HMC) is a  general MCMC algorithm used to efficiently sample posterior draws  high associated with high-dimensional and complex stochastic models.
@@
         A thorough conceptual and theoretical explanation of HMC can be found in \citet{gelman2013bayesian} and \citet{betancourt2017conceptual}.
-HMC builds upon the Metropolis Hastings algorithm through the introduction of Hamiltonian equations, auxiliary momentum variables $\phi$, and the gradients of the target log posterior distribution.
-The vector of model parameters $\theta$ is jointly sampled with a vector of momentum variables $\phi$.
+HMC builds upon the Metropolis Hastings algorithm through the introduction of Hamiltonian equations, auxiliary momentum variables $\alpha$, and the gradients of the target log posterior distribution.
+The vector of model parameters $\theta$ is jointly sampled with a vector of momentum variables $\alpha$.
 The momentum variables are also called auxiliary variables since they are required as part of the Hamiltonian equations to generate an efficient proposal but are not a quantity of interest for inference.
@@
 These differential equations are solved using a discrete approximation algorithm, such as a leapfrog integrator, and is a function of the log posterior gradients for a given point $\theta$.
-The leapfrog integrator has a set of tuning parameters which determines the size and number of steps to be taken along the Hamiltonian path with the final step being the proposal value $(\theta^{\ast}, \phi^{\ast})$.
-Finally, the Metropolis Hastings accept/reject step is applied using the parameters at the beginning of the leapfrog process $(\theta^{t-1}, \phi^{t-1})$ and the proposal values $(\theta^{\ast}, \phi^{\ast})$.
+The leapfrog integrator has a set of tuning parameters which determines the size and number of steps to be taken along the Hamiltonian path with the final step being the proposal value $(\theta^{\ast}, \alpha^{\ast})$.
+Finally, the Metropolis Hastings accept/reject step is applied using the parameters at the beginning of the leapfrog process $(\theta^{n-1}, \alpha^{n-1})$ and the proposal values $(\theta^{\ast}, \alpha^{\ast})$.
 A formal description of the HMC algorithm is provided in Appendix C.
@@
 Stan's default algorithm, the No-U-Turn Sampler \citep{hoffman2014no}, allows for direct sampling of the specified stochastic volatility model.
-The No-U-Turn Sampler (NUTS) proposes the final parameter using multinomial sampling biased towards the second half of the trajectory steps instead of applying the Metrpolis-Hastings step \citep{betancourt2016identifying}.
+The No-U-Turn Sampler (NUTS) proposes the final parameter using multinomial sampling biased towards the second half of the trajectory steps instead of applying the Metropolis-Hastings step \citep{betancourt2016identifying}.
 Stan allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
@@
         \begin{aligned}
-        h_{std} \sim & \space normal(0,1) \\
-        h =& h_{std} \times \sigma_{\eta}
+        h_{std} \sim&\space \,\mathrm{N}(0,1) \\
+        h =& \space \, h_{std} \times \sigma_{\eta}
         \end{aligned}
@@
         \begin{aligned}
-        h_1 =& \space \frac{h_{std, 1}\times \sigma_{\eta}} {\sqrt{1 - \phi^2}} + \mu \\
-        h_{t+1} =& \space h_{std, t+1}\times \sigma_{\eta} + \mu  + \phi(h_{t} - \mu),\space t\neq 1
+        h_1 =& \space \,\frac{h_{std, 1}\times \sigma_{\eta}} {\sqrt{1 - \phi^2}} + \mu \\
+        h_{t+1} =& \space \,h_{std, t+1}\times \sigma_{\eta} + \mu  + \phi(h_{t} - \mu),\space \, t\neq 1
         \end{aligned}
@@
         \begin{aligned}
-        y^{\ast}_t =& h_t + z_t \\
-        h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t
+        y^{\ast}_t =& \space \, h_t + z_t \\
+        h_{t+1} =& \space \, \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t
         \end{aligned}
@@
         \begin{aligned}
-        g_t =& h_t - \mu \\
-        g_{t+1} =& \phi g_t + \sigma_{\eta}\eta_{t}
+        g_t =& \space \, h_t - \mu \\
+        g_{t+1} =& \space \, \phi g_t + \sigma_{\eta}\eta_{t}
         \end{aligned}
@@
         \begin{aligned}
-        y^{\ast}_t =& g_t + \mu + z_t \\
-        g_{t+1} =& \phi g_t + \sigma_{\eta}\eta_{t}
+        y^{\ast}_t =& \space \, g_t + \mu + z_t \\
+        g_{t+1} =& \space \, \phi g_t + \sigma_{\eta}\eta_{t}
         \end{aligned}
@@
         \begin{algorithmic}
-        \For{\texttt{i in} $1:1000$}
-                \State \text{Draw from joint prior: } $\boldsymbol{\theta}^{sim}_i \sim\pi (\boldsymbol{\theta})$
-                \State \text{Simulate dataset with 1000 observations: } $\boldsymbol{y}^{sim}_i \sim p(y|\boldsymbol{\theta}^{sim}_i)$
-                \State \text{Draw 999 posterior samples post burn in:} $\{\boldsymbol{\theta}_1,\dots , \boldsymbol{\theta}_{999}\}_i \sim p(\theta | \boldsymbol{y}^{sim}_i)$
-                \State \text{Compute rank statistics:} $\boldsymbol{r} = rank(\{\boldsymbol{\theta}_1,\dots , \boldsymbol{\theta}_{999}\}_i, \boldsymbol{\theta}^{sim}_i)$
+        \For{\texttt{k in} $1:1000$}
+                \State \text{Draw from joint prior: } $\boldsymbol{\theta}^{sim}_k \sim\pi (\boldsymbol{\theta})$
+                \State \text{Simulate data set with 1000 observations: } $\boldsymbol{y}^{sim}_k \sim p(\boldsymbol{y}|\boldsymbol{\theta}^{sim}_k)$
+                \State \text{Draw 999 posterior samples post burn in:} $\{\boldsymbol{\theta}_1,\dots , \boldsymbol{\theta}_{999}\}_k \sim p(\boldsymbol{\theta} | \boldsymbol{y}^{sim}_k)$
+                \State \text{Compute rank statistics:} $\boldsymbol{r} = rank(\{\boldsymbol{\theta}_1,\dots , \boldsymbol{\theta}_{999}\}_k, \boldsymbol{\theta}^{sim}_k)$
               \EndFor
@@
 
-    The centered parameterisation in particular displays a lack of uniformity for $\phi$ and $\sigma^2$.
+    The centered parameterisation in particular displays a lack of uniformity for $\phi$ and $\sigma^2$ which indicates the MCMC algorithm is not returning calibrated estimates for these parameters.
 Rank statistics for $\sigma^2$ and $\phi$ show two peaks of either end of the histogram.
@@
     \subsubsection{Re-weighting rank statistics of the Gaussian Mixture Model}
-    Let $v(\theta, h)$ be the log weights defined as the log difference between the posterior densities of the true model with log chi squared errors $log\: g(\theta, h | y^{\ast}_t)$ and the Gaussian mixture model $log\:  k(\theta, h_t | y^{\ast}_t)$.
-
-    $$
-    \begin{aligned}
-        v(\theta, h) = log\: g(\theta, h | y^{\ast}_t) - log\:  k(\theta, h_t | y^{\ast}_t) = \text{const} + log\: g(y|h) - log\: k(y^{\ast} | h)
-    \end{aligned}
-    $$
-
-    Take the exponential and normalise the weights for the $l^{th}$ posterior draw (note the constants cancel out):
-
-    $$
-    \begin{aligned}
-    w^l = \frac{exp(v(\theta, h)_l)}{\sum_i exp(v(\theta, h)_i)}
+    Let $v(\theta, h)$ be the log weights defined as the log difference between the posterior densities of the true model with log chi squared errors $log\: c(\theta, h | y^{\ast})$ and the Gaussian mixture model $log\:  d(\theta, h | y^{\ast})$.
+
+    $$
+    \begin{aligned}
+        v(\theta, h) = log\: c(\theta, h | y^{\ast}) - log\:  d(\theta, h | y^{\ast}) = \text{const} + log\: c(y|h) - log\: d(y^{\ast} | h)
+    \end{aligned}
+    $$
+
+    Take the exponential and normalise the weights for the $b^{th}$ posterior draw (note the constants cancel out):
+
+    $$
+    \begin{aligned}
+    w^b = \frac{exp(v(\theta, h)_b)}{\sum_z exp(v(\theta, h)_z)}
     \end{aligned}
@@
     \begin{aligned}
-    S(\theta) &= 1[\theta_l < \theta^{sim}] \\
-    E[S(\theta) | y] &= \int s(\theta) g(\theta | y) d\theta\\ 
-    &= \frac{\int s(\theta)\times exp(v(\theta, h)) * k(\theta, h_t | y^{\ast}_t)d\theta d h}{\int exp(v(\theta, h)) * k(\theta, h_t | y^{\ast}_t)d\theta d h} 
+    S(\theta) &= 1[\theta_b < \theta^{sim}] \\
+    E[S(\theta) | y] &= \int s(\theta) c(\theta | y) d\theta\\ 
+    &= \frac{\int s(\theta)\times exp(v(\theta, h)) * d(\theta, h | y^{\ast})d\theta d h}{\int exp(v(\theta, h)) * d(\theta, h | y^{\ast})d\theta d h} 
     \end{aligned}
@@
     \begin{aligned}
-    E[S(\hat{\theta}) | y^{\ast}] = \sum_l^L S(\hat{\theta}_l)w_l
+    E[S(\hat{\theta}) | y^{\ast}] = \sum_b^B S(\hat{\theta}_b)w_b
     \end{aligned}
@@
     \begin{aligned}
-    r = \sum_{l=1}^{L}1[\theta_{l} < \theta^{sim}] \approx  L\times E[S(\hat{\theta})] = L\times \sum_l^L S(\hat{\theta}_l)w_l
+    r = \sum_{b=1}^{B}1[\theta_{b} < \theta^{sim}] \approx  B\times E[S(\hat{\theta})] = B\times \sum_b^B S(\hat{\theta}_b)w_b
     \end{aligned}
@@
 \section{Appendix B: Sampling from mixture of Gaussian's}
+    The steps in the MCMC process is described in Algorithm \ref{alg:ksc}.
+The initial values are set to the values outlined in \citet{kim1998stochastic}, except for the mixing indicator which is unspecified.
+This is arbitrarily set at the 4th Gaussian density found in Appendix A.
+
+    \subsection*{Sampling mixture density}
+    For sampling the mixture model the Gaussian mixture density is rewritten with respect to a indicator variable $s_t$
+        \begin{align}
+        &z_t | s_t = i \sim N(m_i - 1.2704, \nu^2) \\
+        &Pr(s_t = i) = q_i.
+        \end{align}
+        The vector $s$ indexes the seven mixture of Gaussians and a component indicator variable for each t, $s_t$, is sampled from probability mass function: 
+        \begin{align}
+        Pr(s_t = i | y_t^{\ast}, h_t) \propto q_i f_N(y_t^{\ast} | h_t + m_t - 1.2704, \nu^2).
+        \end{align}
 
@@
 
-\subsection*{Sampling mixture density}
-
-Rewrite mixture density with respect to a indicator variable $s_t$
-
-$$
-\begin{aligned}
-&z_t | s_t = i \sim N(m_i - 1.2704, \nu^2) \\
-&Pr(s_t = i) = q_i
-\end{aligned}
-$$
-
-Sample $s_t$ from probability mass function:
-
-$$
-\begin{aligned}
-Pr(s_t = i | y_t^{\ast}, h_t) \propto q_i f_N(y_t^{\ast} | h_t + m_t - 1.2704, \nu^2)
-\end{aligned}
-$$
+        \begin{algorithm}[H]
+            \caption{KSC MCMC Algorithm}\label{alg:ksc}
+            \begin{algorithmic}
+            \Require $s_0 = 4$, $\mu_0 = 0$, $\phi_0 = 0.95$, $\sigma^{2}_{\eta,0} = 0.02$
+            \For{\texttt{b in} $1:B_{draws}$}
+                    \State \text{Sample states (Kalman Filter and Smoother): } $\boldsymbol{h}_b \sim h|y^{\ast}, s_{b-1}, \phi_{b-1}, \sigma^{2}_{\eta,b-1}, \mu_{b-1}$ 
+                    \State \text{Sample mixture indicators: } $s_b \sim s|y^{\ast}, \boldsymbol{h}_{b-1}$
+                    \State \text{Sample from conjugate density $\mu$: } $\mu_b \sim \mu|y_{\ast}, s_{b-1}, \phi_{b-1}, \sigma^{2}_{\eta, b-1}, \boldsymbol{h}_{b-1}$
+                    \State \text{Sample from conjugate density $\sigma^2_{\eta}$: } $\mu_b \sim \mu|y^{\ast}, s_{b-1}, \phi_{b-1}, \mu_{b-1}, \boldsymbol{h}_{b-1}$
+                    \State \text{Sample via Metropolis-Hastings $\phi$: } $\phi_b \sim \phi|y^{\ast}, s_{b-1}, \mu_{b-1}, \sigma^{2}_{\eta, b-1}, \boldsymbol{h}_{b-1}$
+                  \EndFor
+            \end{algorithmic}
+            \end{algorithm}
 
@@
 \section{Appendix C: Hamiltonian Monte Carlo algorithm description}
-Let $\theta$ be a target parameter and $\phi$ be the auxiliary momentum variable. The Hamiltonian is defined as:
-
-$$
-\begin{aligned}
-H(\theta, \phi) \equiv - log \: \pi(\theta, \phi)
-\end{aligned}
-$$
-
-This can be broken down into kinetic energy $K(\theta, \phi)$, the density over the auxiliary momentum, and potential energy $V(\theta)$, the density of the target posterior distribution.
-
-$$
-\begin{aligned}
-H(\theta, \phi) &= - log \: \pi(\phi | \theta) - log \: \pi(\theta) \\ 
-&\equiv  K(\theta, \phi) + V(\theta)
+Let $\theta$ be a target parameter and $\alpha$ be the auxiliary momentum variable. The Hamiltonian is defined as:
+
+$$
+\begin{aligned}
+H(\theta, \alpha) \equiv - log \: \pi(\theta, \alpha)
+\end{aligned}
+$$
+
+This can be broken down into kinetic energy $K(\theta, \alpha)$, the density over the auxiliary momentum, and potential energy $V(\theta)$, the density of the target posterior distribution.
+
+$$
+\begin{aligned}
+H(\theta, \alpha) &= - log \: \pi(\alpha | \theta) - log \: \pi(\theta) \\ 
+&\equiv  K(\theta, \alpha) + V(\theta)
 \end{aligned}
@@
 
-1) Randomly sample a momentum value
-
-$$
-\begin{aligned}
-\phi\sim Multinormal(0, M)
+\textbf{1) Randomly sample a momentum value}
+
+$$
+\begin{aligned}
+\alpha\sim Multinormal(0, M)
 \end{aligned}
@@
 
-2) Solve the set of Hamiltonian (differential) equations which generates a proposal $(\theta^{\ast}, \phi^{\ast})$
-
-$$
-\begin{aligned}
-\frac{\partial \theta}{\partial t} &= \frac{\partial H}{\partial \phi} = \frac{\partial K}{\partial \phi} \\
-\frac{\partial \phi}{\partial t} &= \frac{\partial H}{\partial \theta} = - \frac{\partial K}{\partial \theta} - \frac{\partial V}        {\partial \theta}
+\textbf{2) Solve the set of Hamiltonian (differential) equations which generates a proposal} $(\theta^{\ast}, \alpha^{\ast})$
+
+$$
+\begin{aligned}
+\frac{\partial \theta}{\partial t} &= \frac{\partial H}{\partial \alpha} = \frac{\partial K}{\partial \alpha} \\
+\frac{\partial \alpha}{\partial t} &= \frac{\partial H}{\partial \theta} = - \frac{\partial K}{\partial \theta} - \frac{\partial V}        {\partial \theta}
 \end{aligned}
@@
 The solution to these differential equations is approximated using a Leapfrog integrator which gives a discrete approximate solution.
-The parameters which are tuned in this process are the number of steps $L$ and the size of the steps $\epsilon$.
-The Leapfrog algorithm performs a half step of $\phi$ and then a full step of the parameter $\theta$, and another half step of the momentum variable $\phi$ and repeats this $L$ times.
-The final step is the proposal $(\theta^{\ast}, \phi^{\ast})$.
-
-$$
-\begin{aligned}
-\phi &\leftarrow \phi - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta} \\
-\theta &\leftarrow \theta + \epsilon M^{-1} \phi \\
-\phi &\leftarrow \phi - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta}
-\end{aligned}
-$$
-
-3) Metrpolis Hastings Accept/Reject
-Let $(\theta^{t-1}, \phi^{t-1})$ be the values before the Leapfrog integrator.
-
-$$
-\begin{aligned}
-r = \frac{\pi(\theta^{\ast} | y) \pi(\phi^{\ast})}{\pi(\theta^{t-1} | y) \pi(\phi^{t-1})}
+The parameters which are tuned in this process are the number of steps $G$ and the size of the steps $\epsilon$.
+The Leapfrog algorithm performs a half step of $\alpha$ and then a full step of the parameter $\theta$, and another half step of the momentum variable $\alpha$ and repeats this $G$ times.
+The final step is the proposal $(\theta^{\ast}, \alpha^{\ast})$.
+
+$$
+\begin{aligned}
+\alpha &\leftarrow \alpha - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta} \\
+\theta &\leftarrow \theta + \epsilon M^{-1} \alpha \\
+\alpha &\leftarrow \alpha - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta}
+\end{aligned}
+$$
+
+\textbf{3) Metropolis Hastings Accept/Reject}
+
+Let $(\theta^{n-1}, \alpha^{n-1})$ be the values before the Leapfrog integrator.
+
+$$
+\begin{aligned}
+a = \frac{\pi(\theta^{\ast} | y) \pi(\alpha^{\ast})}{\pi(\theta^{n-1} | y) \pi(\alpha^{n-1})}
 \end{aligned}
@@
 $$
-\theta^t = \begin{cases}
-    \theta^{\ast},& \text{with probability } min(r,1)\\
-    \theta^{t-1}, & \text{otherwise}
+\theta^n = \begin{cases}
+    \theta^{\ast},& \text{with probability } min(a,1)\\
+    \theta^{n-1}, & \text{otherwise}
 \end{cases}
@@
 The shape of the histogram is consistent with the unweighted rank statistics.}
-    \label{fig:reweight1k}
+    \label{fig:ncpreweight1k}
 \end{figure}
@@
 No major improvements are observed to the uniformity of rank statistics.}
-    \label{fig:reweight5k}
+    \label{fig:ncpreweight5k}
 \end{figure}

### 2023-10-17 manuscript/main.tex
Draft with 1000 and 5000 SBC iterations in main text.

@@
 
-\title{Comparing MCMC algorithms for Stochastic Volatility Model Bayesian Posteriors using Simulation Based Calibration}
+\title{Comparing MCMC algorithms in Stochastic Volatility Models using Simulation Based Calibration}
 \author{Benjamin Wee}
@@
 The focus of this paper is to investigate the performance of two competing MCMC algorithms, each applied to a SV model, in terms of the validity and statistical efficiency of the their resulting inference.
-Performance of these algorithms will also be compared with different parameterisations of the SV model to determine the sensitivity of each MCMC to model specification.
+Performance of these algorithms will also be compared with different parameterisations of the SV model to determine the sensitivity of MCMC calibration to model specification.
 
     \citet{kim1998stochastic} propose a sampling strategy for the so-called 'vanilla' stochastic volatility through an off-set mixture approximation.
-Conditional on the mixture allocations and static parameters, the forward filter backward sampling algorithm (Carter and Kohn, 1995; Fr{\"u}wirth-Schnatter, 1995) is used to jointly sample the latent states while conditionally  conjugate prior distributions are also suggested.
+Conditional on the mixture allocations and static parameters, the forward filter backward sampling algorithm \citep{carter1994gibbs, fruhwirth1995bayesian} is used to jointly sample the latent states while conditionally  conjugate prior distributions are also suggested.
 Since then, there have been many advances in statistical computing and algorithm design.
@@
 By calibration we refer to a MCMC algorithm that returns correct posterior, on average, conditional on the correct model specification.
-The first algorithm replicates Kim Shephard and Chib's (KSC) Gaussian mixture approximation which is sampled using the aforementioned Kalman Filter and Metropolis Hastings algorithm.
-The second algorithm is the No-U-turn sampler (NUTS) variant of Hamiltonian Monte Carlo (HMC) as implemented in the Stan programming language \citep{hoffman2014no, betancourt2017conceptual, stan}.
+The first algorithm is the No-U-turn sampler (NUTS) variant of Hamiltonian Monte Carlo (HMC) as implemented in the Stan programming language \citep{hoffman2014no, betancourt2017conceptual, stan}.
+The second algorithm replicates Kim Shephard and Chib's (KSC) Gaussian off-set mixture model which is sampled using the aforementioned Kalman Filter and Metropolis Hastings algorithm.
 The objective is to analyse the efficiency of the MCMC and their ability to return correct posterior estimates.
@@
 Section 3 describes the SBC methodology, along with the simulation design and diagnostic metrics used for the SV model.
-Section 4 details key aspects of the two MCMC sampling approaches under consideration, while Section 5 discusses the results and limitations.
-Finally, Section 6 concludes and provides points for further research.
+Section 4 details key aspects of the two MCMC sampling approaches under consideration, while Section 5 provides the simulation results.
+Finally, Section 6 discusses findings, limitations and provides points for further research.
 
@@
 \subsection{Stochastic Volatility}
-    The model of interest is the discrete time, uni-variate stochastic volatility model estimated by \citet{kim1998stochastic} using Bayesian framework.
-SV is expressed as a state space model which links the observed mean corrected (logarithmic) return of a given asset to the latent conditional variance.
+    The model of interest is the discrete time, univariate stochastic volatility model estimated by \citet{kim1998stochastic} using a Bayesian framework.
+SV is expressed as a state space model which links the observed (logarithmic) return of a given asset over period t, denoted $y_t$, to its latent conditional variance, $h_{t+1}$.
+This model is given by the following expressions
     \begin{align}
@@
     \end{align}
-    $y_t$ is the mean corrected (logarithmic) return of a given asset, observed over an interval with fixed length and associated with period t.
     The measurement equation (1) describes the conditional distribution of $y_t$ given $h_t$, where $h_t$ is the conditional log variance of $y_t$.
@@
 
-        Posterior credible intervals are said to have sufficient coverage if Bayesian computation is well calibrated and the rank statistics follow a uniform distribution.
+        % Posterior credible intervals are said to have sufficient coverage if Bayesian computation is well calibrated and the rank statistics follow a uniform distribution.
 That is, one way to describe calibration is: for any percentage interval selected over the posterior samples (for example 90\%) then there is a 90\% chance that $\theta^{sim}$ falls within this interval \citep{dawid1982well}.
@@
 
-            \item Effective sample size (ESS) to measure MCMC efficiency
+            \item Effective sample size (ESS) to measure MCMC efficiency.
         \end{enumerate}
@@
             Rank statistics are used to evaluate the calibration of the MCMC algorithm.
-If a posterior is well calibrated then it is expected that the distribution of rank statistics over $B$ SBC iterations for each parameter in the model is uniform.
+If a posterior is well calibrated, then it is expected that over $K$ SBC iterations for each univariate parameter in the model:
+
+            \begin{enumerate}
+                \item The distribution of rank statistics is uniform.
+                \item The posterior draws on average are equal to the prior distribution.
+            \end{enumerate}
 
             The shape of the distribution of rank statistics gives insight into how an MCMC may be miscalibrated.
-Specifically, it gives information about the type of bias in a marginal posterior estimate of a given parameter.
+Specifically, it gives information about the average posterior deviates from the prior distribution of a given parameter.
 Figures \ref{fig:underestimation} and \ref{fig:underdispersed} are recreated from \citet{talts2018validating} for convenience.
-In these we display examples of posterior sampling bias and the corresponding non-uniform rank statistics.
+These display examples of posterior miscalibration and the corresponding non-uniform rank statistics.
 
@@
 The sum of indicator random variables (see equation (13)) is large since a large proportion of posterior draws are smaller than the true value, resulting in a large rank statistic and hence a larger relative frequency on the right side of the histogram.
-The converse is true if the peak of the histogram is on the left hand side (i.e the average posterior over the SBC iterations is biased towards the right hand side of the prior).
+The converse is true if the peak of the histogram is on the left hand side (i.e the average posterior over the SBC iterations overestimates prior).
 
@@
 \citet{talts2018validating} describe this bias as under-dispersion of the posterior relative to the prior distribution on average.
-The estimated posterior is too narrow relative to the spread of the prior resulting in oversampling on both ends of the rank statistic distribution.
+The average posterior is too narrow (or overconfident) relative to the spread of the prior resulting in inflated frequency on both ends of the rank statistic distribution.
 
@@
                 \includegraphics[scale=0.07]{methodology/underdispersed.png}
-                \caption{Left: Non uniform rank statistics with peaks on both ends.
+                \caption{Left: Non uniform rank statistics with peaks on both ends (`U-shape' distribution).
 Right: Overlayed posterior and prior distributions for the same arbitrary parameter.
@@
 \section{MCMC Strategies}
-
-    \subsection{Sampling method 1: Mixture off-set MCMC}
+    \subsection{Sampling method 1: Stan's Hamiltonian Monte Carlo with No-U-Turn Sampler}
+        Hamiltonian Monte Carlo (HMC) is a  general MCMC algorithm used to efficiently sample posterior draws  high associated with high-dimensional and complex stochastic models.
+Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature \citep{duane1987hybrid} before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks \citep{neal1995bayesian} and statistical computing \citep{neal2011mcmc}.
+Versions of this algorithm have since become widely available through open source development projects, including Stan \citep{stan} and PyMC \citep{pymc2023}.
+
+        The key innovation of HMC is that it uses the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
+HMC is able to reach more distant points with higher acceptance probabilities than many bespoke MCMC algorithms since its  proposal draws are made using information about the target distribution.
+Random walk samplers (such as Random Walk Metropolis Hastings) tend to be inefficient in higher dimensions since it becomes more difficult to randomly select a point with a high acceptance probability.
+The increasing complexity of high dimensions make it harder for a random walk sampler to explore regions over long distances, and thus the sampler tends to become stuck and will require many iterations to converge to the target distribution.
+
+        A thorough conceptual and theoretical explanation of HMC can be found in \citet{gelman2013bayesian} and \citet{betancourt2017conceptual}.
+HMC builds upon the Metropolis Hastings algorithm through the introduction of Hamiltonian equations, auxiliary momentum variables $\alpha$, and the gradients of the target log posterior distribution.
+The vector of model parameters $\theta$ is jointly sampled with a vector of momentum variables $\alpha$.
+The momentum variables are also called auxiliary variables since they are required as part of the Hamiltonian equations to generate an efficient proposal but are not a quantity of interest for inference.
+
+        The sampled momentum variable sets out the proposal path for $\theta$ governed by the Hamiltonian dynamics (represented by a set of differential equations).
+These differential equations are solved using a discrete approximation algorithm, such as a leapfrog integrator, and is a function of the log posterior gradients for a given point $\theta$.
+The leapfrog integrator has a set of tuning parameters which determines the size and number of steps to be taken along the Hamiltonian path with the final step being the proposal value $(\theta^{\ast}, \alpha^{\ast})$.
+Finally, the Metropolis Hastings accept/reject step is applied using the parameters at the beginning of the leapfrog process $(\theta^{n-1}, \alpha^{n-1})$ and the proposal values $(\theta^{\ast}, \alpha^{\ast})$.
+A formal description of the HMC algorithm is provided in Appendix C.
+
+        \subsubsection{Implementation}
+        The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
+Stan's default algorithm, the No-U-Turn Sampler \citep{hoffman2014no}, allows for direct sampling of the specified stochastic volatility model.
+The No-U-Turn Sampler (NUTS) proposes the final parameter using multinomial sampling biased towards the second half of the trajectory steps instead of applying the Metropolis-Hastings step \citep{betancourt2016identifying}.
+Stan allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
+
+        This approach will also use the same priors as specified in the KSC's Gaussian mixture approximation so that HMC is sampling from the same model (although other priors could be chosen).
+
+    \subsection{Sampling method 2: Mixture off-set MCMC}
         \citet{kim1998stochastic} sample the conditional posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings within Gibbs algorithms and an implementation of the Kalman Filter and smoother used to sample from the conditional posterior distribution of the latent states \citep{dejong1995}.\footnote{In this research the exact software to apply the simulation smoother is unavailable.
@@
         where $K=7$, $f_N(z_t|m_i-1.2704, \nu_i^2)$ denotes the density of the normal distribution with mean $m_i-1.2704$ and variance $v_i^2$, and  component probabilities $q_i$, for $i=1,2,\ldots, I=7$.
-The values of these hyperparameters and component weights are reported in Appendix A.
+The values of these hyperparameters and component weights are applied in this research and reported in Appendix A.
 
@@
 
-    \subsection{Sampling method 2: Stan's Hamiltonian Monte Carlo with No-U-Turn Sampler}
-        Hamiltonian Monte Carlo (HMC) is a  general MCMC algorithm used to efficiently sample posterior draws  high associated with high-dimensional and complex stochastic models.
-Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature \citep{duane1987hybrid} before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks \citep{neal1995bayesian} and statistical computing \citep{neal2011mcmc}.
-Versions of this algorithm have since become widely available through open source development projects, including Stan \citep{stan} and PyMC \citep{pymc2023}.
-
-        The key innovation of HMC is that it uses the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
-HMC is able to reach more distant points with higher acceptance probabilities than many bespoke MCMC algorithms since its  proposal draws are made using information about the target distribution.
-Random walk samplers (such as Random Walk Metropolis Hastings) tend to be inefficient in higher dimensions since it becomes more difficult to randomly select a point with a high acceptance probability.
-The increasing complexity of high dimensions make it harder for a random walk sampler to explore regions over long distances, and thus the sampler tends to become stuck and will require many iterations to converge to the target distribution.
-
-        A thorough conceptual and theoretical explanation of HMC can be found in \citet{gelman2013bayesian} and \citet{betancourt2017conceptual}.
-HMC builds upon the Metropolis Hastings algorithm through the introduction of Hamiltonian equations, auxiliary momentum variables $\alpha$, and the gradients of the target log posterior distribution.
-The vector of model parameters $\theta$ is jointly sampled with a vector of momentum variables $\alpha$.
-The momentum variables are also called auxiliary variables since they are required as part of the Hamiltonian equations to generate an efficient proposal but are not a quantity of interest for inference.
-
-        The sampled momentum variable sets out the proposal path for $\theta$ governed by the Hamiltonian dynamics (represented by a set of differential equations).
-These differential equations are solved using a discrete approximation algorithm, such as a leapfrog integrator, and is a function of the log posterior gradients for a given point $\theta$.
-The leapfrog integrator has a set of tuning parameters which determines the size and number of steps to be taken along the Hamiltonian path with the final step being the proposal value $(\theta^{\ast}, \alpha^{\ast})$.
-Finally, the Metropolis Hastings accept/reject step is applied using the parameters at the beginning of the leapfrog process $(\theta^{n-1}, \alpha^{n-1})$ and the proposal values $(\theta^{\ast}, \alpha^{\ast})$.
-A formal description of the HMC algorithm is provided in Appendix C.
-
-        \subsubsection{Implementation}
-        The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
-Stan's default algorithm, the No-U-Turn Sampler \citep{hoffman2014no}, allows for direct sampling of the specified stochastic volatility model.
-The No-U-Turn Sampler (NUTS) proposes the final parameter using multinomial sampling biased towards the second half of the trajectory steps instead of applying the Metropolis-Hastings step \citep{betancourt2016identifying}.
-Stan allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
-
-        This approach will also use the same priors as specified in the KSC's Gaussian mixture approximation so that HMC is sampling from the same model (although other priors could be chosen).
-
 \subsection{MCMC methods comparison}
-        The KSC sampling method and HMC are both MCMC strategies to generate samples from the target joint posterior distribution of the stochastic volatility model.
+        HMC and the KSC sampling method are both MCMC strategies to generate samples from the target joint posterior distribution of the stochastic volatility model.
 However, the two sampling approaches are conceptually different in their design and implementation, despite sharing the same objective.
 
-        The KSC sampling method is a bespoke algorithm designed specifically for the sampling of the stochastic volatility model.
-It is a strategy built around how to sample the parameters from the state space model in this specific context.
-The implementation of this MCMC requires careful attention to the sampling of each parameter using different tools and requires these components to be programmed manually, such as the Metropolis Hastings step and conjugate posterior distributions.
-
-        HMC on the other hand is a more general approach to applying MCMC.
+        HMC is a general approach to applying MCMC.
 The implementation of HMC is through the Stan programming language which allows for the specification other generative or Bayesian models, not just stochastic volatility.
@@
 Furthermore, Stan and HMC allows for the specification of different priors, whereas KSC's bespoke algorithm is dependent on the use of conjugate priors.
+
+        The KSC sampling method on the other hand is a bespoke algorithm designed specifically for the sampling of the stochastic volatility model.
+It is a strategy built around how to sample the parameters from the state space model in this specific context.
+The implementation of this MCMC requires careful attention to the sampling of each parameter using different tools and requires these components to be programmed manually, such as the Metropolis Hastings step and conjugate posterior distributions.
 
@@
 
-        SBC is applied to reparamaterised versions of the stochastic volatility model to see if there is any improvement in calibration and efficiency.
-
+        This research compares two different parameterisations for the two different sampling methods in a 2x2 design using SBC. This is summarised in Table \ref{tab:params}.
+        \begin{table}
+            \centering
+            \begin{tabular}{|c|c|c|} \hline 
+                 &  \textbf{Centered}& \textbf{Reparameterised}\\ \hline 
+                 \textbf{HMC}&  Centered HMC& Reparameterised HMC\\ \hline 
+                 \textbf{Mixture off-set MCMC (KSC)}&  Centered KSC& Reparameterised (Non-Centered) KSC\\ \hline
+            \end{tabular}
+            \caption{MCMC and parameterisation matrix}
+            \label{tab:params}
+        \end{table}
         \subsubsection{Reparamaterised stochastic volatility in Stan}
@@
 Specifically, the states are first sampled from a standard normal distribution and are transformed to be samples from the state equation.
-As noted above, the centered paramaterisation samples the states centered around $\mu$.
-
-        First sample from a standard normal distribution and multiply by the variance of the log volatility.
+There are two steps to this:
+        \begin{enumerate}
+
+        \item Sample from a standard normal distribution and multiply by the variance of the log volatility.
 The state vector is sampled with mean centered on zero and variance equal to $\sigma_{\eta}^2$
-
-        $$
-        \begin{aligned}
+        \begin{align}
         h_{std} \sim&\space \,\mathrm{N}(0,1) \\
         h =& \space \, h_{std} \times \sigma_{\eta}
-        \end{aligned}
-        $$
-
-        Then apply the appropriate re-scaling to get samples from log volatility.
-
-        $$
-        \begin{aligned}
+        \end{align}
+        \item Apply the appropriate re-scaling to get samples from log volatility. 
+        \begin{align}
         h_1 =& \space \,\frac{h_{std, 1}\times \sigma_{\eta}} {\sqrt{1 - \phi^2}} + \mu \\
         h_{t+1} =& \space \,h_{std, t+1}\times \sigma_{\eta} + \mu  + \phi(h_{t} - \mu),\space \, t\neq 1
-        \end{aligned}
-        $$
-
+        \end{align}
+        \end{enumerate}
         This returns log volatility as desired.
 
-        \subsubsection{Non centered Gaussian Mixture model}
-        The re-parameterised model for the Gaussian approximation is expressed differently due to the Kalman Filter.
+        \subsubsection{Non-centered Gaussian mixture off-set model}
+        The re-parameterised model for the Gaussian mixture off-set is expressed differently due to the Kalman Filter.
 Re-parameterised HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates.
@@
         \begin{algorithmic}
-        \For{\texttt{k in} $1:1000$}
+        \For{\texttt{k in} $1:5000$}
                 \State \text{Draw from joint prior: } $\boldsymbol{\theta}^{sim}_k \sim\pi (\boldsymbol{\theta})$
@@
     The state variable chi squared statistics for centered and re-parameterised HMC, centered KSC and centered importance weighted KSC are reported as histograms in Figure \ref{fig:allchisq} (using results from 5000 SBC iterations).
-Re-parameterised KSC is omitted as its poor SBC performance just adds noise to the results.
+Re-parameterised KSC is omitted as its poor SBC performance skews the scale of the plot so the results are hard to see.
 A chi squared statistic of 0 implies the sample is exactly uniformly distributed.
@@
 The SBC results for the Gaussian mixture may improve if the priors were more informative - although this is only conjecture.
-Whether or not to use tighter priors depends on the modelling context and the problem at hand.
+Whether or not to use tighter priors depends on the modelling context and the problem at hand \citep{gelman2017prior}.
 If a model is expected to handle data generated by the priors, then un-calibrated results from SBC raises questions about the suitability about the model choice, MCMC or parameterisation.
@@
 
-\section{Limitations}
+\subsection{Limitations}
 A limitation to applying SBC in all modelling contexts is that it is computationally intensive.
@@
 
-\section{Conclusion}
+\subsection{Conclusion}
 This research evaluated and compared different MCMC algorithms for fitting stochastic volatility models.

### 2023-10-17 manuscript/main.tex
Updates from Overleaf

@@
 \section{Results}
-    1000 and 5000 SBC iterations are run for centered and re parameterised versions of the stochastic volatility model using simulated data sets of 1000 observations.
+    SBC is conducted for each MCMC sampling method and corresponding parameterisation as outlined in Table \ref{tab:params}.
+For a given MCMC and parameterisation pair, 5000 SBC iterations are run using simulated data sets of size T=1000.
+5000 rank statistics are obtained per univariate parameter from this process.
+
     Rank statistic histograms are constructed with 20 bins for the static parameters and the 1st, 500th and 1000th latent state variables.
-The black horizontal line represents uniformity conditional on the number of bins and SBC iterations.
+The black horizontal line represents a uniform distribution at 250 which is the ratio of  5000 SBC iterations and 20 bins.
 The $\chi^2$ statistic is used to summarise the deviations from uniformity for all state variables since it is unrealistic to visually inspect all the histograms.
-The SBC algorithm is summarised Algorithm \ref{alg:sbc}:
+The SBC algorithm is summarised in Algorithm \ref{alg:sbc}.
 
@@
     \subsection{Hamiltonian Monte Carlo}
-    Figure \ref{fig:cphmc1k} and Figure \ref{fig:ncphmc1k} show the SBC results for the centered and re-parameterised model for HMC respectively.
-Both sets of results display relatively noisy rank statistic estimates with variation around the discrete uniform distribution.
-
-    The centered parameterisation in particular displays a lack of uniformity for $\phi$ and $\sigma^2$ which indicates the MCMC algorithm is not returning calibrated estimates for these parameters.
-Rank statistics for $\sigma^2$ and $\phi$ show two peaks of either end of the histogram.
-The posterior estimates of $\sigma^2$ and $\phi$ are \textit{under-dispersed} relative to the prior distribution.
+    SBC results for the centered model using HMC are displayed in Figure \ref{fig:cphmc5k}.
+The centered parameterisation displays a lack of uniformity in rank statistics for $\sigma^2$ and $\phi$.
+The $\sigma^2$ parameter displays a U-shaped histogram with large peaks on both ends while $\phi$ exhibits a U-shape with smaller peaks.
+This indicates the MCMC algorithm is not returning calibrated estimates for these parameters.
+The average posterior samples of $\sigma^2$ and $\phi$ over the SBC iterations are \textit{under-dispersed}, or \textit{overconfident}, relative to the prior distribution.
 This implies the posterior geometry of these parameters may be difficult to sample for the HMC algorithm.
 
-    \begin{figure}[H]
-        \centering
-        \includegraphics[scale=0.09]{results/hmc_cp_1k.png}
-        \caption{1000 SBC iterations for centered model using Hamiltonian Monte Carlo.
-Histograms show noisy estimates around uniform distribution.
-In particular, $\sigma^2$ displays peaks on both ends of the histogram and $\phi$ shows a major on the left hand side.
-This suggests that the posterior samples of $\sigma^2$ are under-dispersed relative to the prior distribution and posterior samples of $\phi$ are over estimating the true parameter.}
-        \label{fig:cphmc1k}
-    \end{figure}
-
-    Estimates of $\sigma^2$ and $\phi$ are more uniform under the re-parameterised model.
-Instead, we observe a peak on the left side for the 500th state variable $h_{500}$, although the estimates are noisy.
-Overall the estimates for non centered parameterisation are more favourable.
-
-    \begin{figure}[H]
-        \centering
-        \includegraphics[scale=0.09]{results/hmc_ncp_1k.png}
-        \caption{1000 SBC iterations for non centered model using Hamiltonian Monte Carlo.
-The rank statistic distributions for $\sigma^2$ and $\phi$ have improved.
-There is some evidence that the posterior of the 500th state variable is overestimating the distribution of its true value.}
-        \label{fig:ncphmc1k}
-    \end{figure}
-
-    The noisy histogram estimates can be improved by increasing the number of SBC iterations.
-Figure \ref{fig:cphmc5k} and Figure \ref{fig:ncphmc5k} show the results of the same parameters with 5000 SBC iterations.
-There is an improvement to the re-parameterised model with all histograms looking approximately uniform with much smaller variation around the uniform distribution.
-
-    Centered parameterisation posses similar improvement for $\mu$ and the state variables.
-However, there are still major peaks at both ends of $\sigma^2$ and $\phi$ suggesting under-dispersion of the posterior estimates relative to the prior distribution.
-
+    Figure \ref{fig:ncphmc5k} show the SBC results for the reparameterised model using HMC.
+The re-parameterised model produces approximately uniform rank statistics across all reported parameters.
+The distributions of $\sigma^2$ and $\phi$ now exhibit a more uniform shape, implying greater calibration and correct posterior estimates returned on average.
 These results suggests that HMC struggles to sample some parameters from the centered parameterisation of the stochastic volatility model.
-The re-parameterised model possess stronger evidence of calibration for the parameters examined, implying more reliable and consistent inference.
+The re-parameterised model possess stronger evidence of calibration for the parameters examined, producing more reliable and consistent inference.
 
@@
         \includegraphics[scale=0.09]{results/hmc_cp_5k.png}
-        \caption{5000 SBC iterations for centered model using Hamiltonian Monte Carlo.
+        \caption{\textbf{5000 SBC iterations for centered model using Hamiltonian Monte Carlo}.
 $\mu$ and the state parameters look approximately uniform.
 $\sigma^2$ still has major peaks on both the left and right hand side indicating under-dispersion relative to the prior distribution.
-$\phi$ exhibits a similar shape to $\sigma^2$ although the peaks are much smaller.}
+$\phi$ exhibits a similar U-shape to $\sigma^2$ although the peaks are much smaller.}
         \label{fig:cphmc5k}
@@
         \includegraphics[scale=0.09]{results/hmc_ncp_5k.png}
-        \caption{5000 SBC iterations for re-parameterised model using Hamiltonian Monte Carlo.
-Parameters display uniformity with less noise around the true uniform value.
-This suggests that the HMC algorithm is returning the correct posterior estimates for the selected parameters.}
+        \caption{\textbf{5000 SBC iterations for re-parameterised model using Hamiltonian Monte Carlo}.
+The rank statistic distributions for $\sigma^2$ and $\phi$ appear to follow a uniform distribution.
+This suggests that the HMC algorithm is returning the correct posterior estimates for the displayed parameters.}
         \label{fig:ncphmc5k}
@@
         \includegraphics[scale=0.09]{results/hmc_ess.png}
-        \caption{Effective sample sizes for static parameters after 5000 iterations of SBC and 1000 post warm-up samples for Hamiltonian Monte Carlo.
+        \caption{\textbf{Effective sample sizes for static parameters after 5000 SBC iterations}.
 The centered parameterisation struggles to generate independent samples for $\sigma^2$ and $\phi$.
@@
 Overall the re-weighting of the posterior samples from the approximate model does not improve the calibration results.
-The re-weighted rank statistics for the re-parameterised model also did not improve and can be found in Appendix D.
+The re-weighted rank statistics for the re-parameterised model also did not improve and can be found in Appendix E.
 
@@
 
-\section{Appendix D: Importance weighted rank statistics for re-parameterised model}
+\section{Appendix D: 1000 SBC iteration results}
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.09]{results/hmc_cp_1k.png}
+        \caption{1000 SBC iterations for centered model using Hamiltonian Monte Carlo.
+Histograms show noisy estimates around uniform distribution.
+In particular, $\sigma^2$ displays peaks on both ends of the histogram and $\phi$ shows a major on the left hand side.
+This suggests that the posterior samples of $\sigma^2$ are under-dispersed relative to the prior distribution and posterior samples of $\phi$ are over estimating the true parameter.}
+        \label{fig:cphmc1k}
+    \end{figure}
+
+        \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.09]{results/hmc_ncp_1k.png}
+        \caption{1000 SBC iterations for non centered model using Hamiltonian Monte Carlo.
+The rank statistic distributions for $\sigma^2$ and $\phi$ have improved.
+There is some evidence that the posterior of the 500th state variable is overestimating the distribution of its true value.}
+        \label{fig:ncphmc1k}
+    \end{figure}
+
+    Instead, we observe a peak on the left side for the 500th state variable $h_{500}$, although the estimates are noisy. Overall the estimates for non centered parameterisation are more favourable.
+
+\section{Appendix E: Importance weighted rank statistics for re-parameterised model}
 

### 2023-10-18 manuscript/main.tex
BW latest changes before adding new plots

@@
 \begin{abstract}
-    This research compares the computational methods used to estimate Bayesian stochastic volatility models.
+
+    This research uses the simulation-based calibration (SBC) method of \citet{talts2020validating} to analyse two commonly used, competing Markov chain Monte Carlo algorithms for the purpose of estimating a posterior distribution in the context of a stochastic volatility model.
+In particular, the bespoke ‘off-set mixture approximation’ algorithm proposed by \citet{kim1998stochastic} is explored together with a generic Hamiltonian Monte Carlo algorithm implemented through Stan \citep{stan}.
+The SBC analysis involves a simulation study to assess whether each sampling algorithm has the capacity to produce valid inference for the correctly specified model, while also characterising statistical efficiency through effective sample size.
+Results show that Stan's No-U-Turn sampler, an implementation of Hamiltonian Monte Carlo, produces a well-calibrated posterior estimate while the celebrated off-set mixture approach is less efficient and poorly calibrated, though model parameterisation also plays a role.
+Limitations and restrictions of generality are discussed.
+
+    % This research compares the computational methods used to estimate Bayesian stochastic volatility models.
 The MCMC approaches outlined in the stochastic volatility literature and Hamiltonian Monte Carlo are assessed in their ability to sample from the model's posterior distribution.
@@
 
-    Comparisons of these algorithms are conducted using a methodology known as Simulation Based Calibration (SBC) \citep{talts2018validating}.
+    Comparisons of these algorithms are conducted using a methodology known as Simulation Based Calibration (SBC) \citep{talts2020validating}.
 Parameters are drawn from the prior distribution and used to create data sets from the generative stochastic volatility model.
@@
 An algorithm may be sensitive to different parameterisations of the same model which may impact MCMC performance as described in \citet{strickland2008parameterisation}.
-The key findings of this research are HMC applied on a re-parameterised SV model produces the most calibrated and efficient posterior estimates and that the MCMC algorithms may perform better or worse based on calibration and efficiency depending on the parameterisation of the model.
+The key findings of this research are HMC applied on a re-parameterised SV model produces the most calibrated and efficient posterior estimates relative to KSC's MCMC algorithm on the approximate off-set mixture model.
+Additionally, MCMC algorithms may perform better or worse based on calibration and efficiency depending on the parameterisation of the model.
 
@@
 Section 4 details key aspects of the two MCMC sampling approaches under consideration, while Section 5 provides the simulation results.
-Finally, Section 6 discusses findings, limitations and provides points for further research.
+Finally, Section 6 discusses findings, limitations and suggests areas for further research.
 
 \section{Research Context}
+This section describes the stochastic volatility model to be analysed in this research and discusses two examples to motivate the research objectives.
+Section 2.1 outlines the model specificaion and priors, Section 2.2 explores challenges with inference and convergence diagnostics with no ground truth and Section 2.3 mentions the limitations and pitfalls of running only a single simulation.
 
@@
 Even for a correctly specified model, there is always a small probability that the true parameter is in the tails of the corresponding (exact) marginal posterior distribution.
-\cite{talts2018validating} make the point that a single simulation does not provide sufficient information about the inference made by an algorithm.
+\cite{talts2020validating} make the point that a single simulation does not provide sufficient information about the inference made by an algorithm.
 As discussed in their paper, a single simulation may conclude ``that an incorrectly coded analysis worked as desired, while a correctly coded analysis failed''.
@@
 \section{Methodology}
-
+The methodology section describes the simulation study used validate the posterior estimates from MCMC algorithms and the metrics used to empirically measure calibration and efficiency.
+Section 3.1 discusses the procedure for running SBC and the key results for defining calibration.
+Section 3.2 defines the diagnostics and metrics for evaluating SBC experiment results.
     \subsection{Simulation Design}
@@
 For this parameter a data set is generated from the model.
-Then a sampling method (originally using Stan in \citet{talts2018validating}) is used to obtain posterior samples given this data and the prior.
+Then a sampling method (originally using Stan in \citet{talts2020validating}) is used to obtain posterior samples given this data and the prior.
 Then the estimated posteriors are compared to the true values by calculating the rank.
@@
 
-        \citet{talts2018validating} prove that the sample of rank statistics for a given parameter, produced from replicate posteriors, follow a uniform distribution if the posterior samples follow the same distribution as the prior on average.
-The rank statistic for $\theta$ produced by the $k^{\mathrm{th}}$ replicate posterior sample, is defined as
-        \begin{align}
-        r^{(k)} = rank(\{\theta_1^{(k)},\dots , \theta_{B}^{(k)}\}, \theta^{sim}) = \sum_{b=1}^{B}1[\theta_{b}^{(k)} < \theta^{sim}].
+        \citet{talts2020validating} prove that the sample of rank statistics for a given parameter, produced from replicate posteriors, follow a uniform distribution over the integers $\{0,1\dots,B\}$ if the posterior samples follow the same distribution as the prior on average.
+The rank statistic for any one-dimensional function of parameters $f:\Theta\rightarrow\ \mathbb{R}$ produced by the $k^{\mathrm{th}}$ replicate posterior sample, is defined as
+        \begin{align}
+        r^{(k)} = rank(\{f(\theta_1^{(k)}),\dots , f(\theta_{B}^{(k)})\}, f(\theta^{sim})) = \sum_{b=1}^{B}1[f(\theta_{b}^{(k)}) < f(\theta^{sim})].
         \end{align}
@@
 Specifically, it gives information about the average posterior deviates from the prior distribution of a given parameter.
-Figures \ref{fig:underestimation} and \ref{fig:underdispersed} are recreated from \citet{talts2018validating} for convenience.
+Figures \ref{fig:underestimation} and \ref{fig:underdispersed} are recreated from \citet{talts2020validating} for convenience.
 These display examples of posterior miscalibration and the corresponding non-uniform rank statistics.
@@
             Figure \ref{fig:underdispersed} has large peaks on both ends of the histogram resulting in a `U-shape'.
-\citet{talts2018validating} describe this bias as under-dispersion of the posterior relative to the prior distribution on average.
+\citet{talts2020validating} describe this bias as under-dispersion of the posterior relative to the prior distribution on average.
 The average posterior is too narrow (or overconfident) relative to the spread of the prior resulting in inflated frequency on both ends of the rank statistic distribution.
@@
 
-            The ESS is denoted as $n_{\text{eff}}$.
+            The ESS is denoted as $n_{\text{eff}}$ as given by
+            \begin{align}
+                n_{\text{eff}} = \frac{NM}{1+2 \sum_{l=1}^L \hat{\rho}_l}.
+            \end{align}
             When more than one chain is generated from the target posterior, then the ESS quantity is a function of the number of chains $M$, the number of samples per chain $N$ and the auto-correlation at lag $l$ $\rho_l$ \citep{vehtari2021rank}.
@@
 This truncation (or some kind of down-weighting) is required because a large number of lags leads to noisier auto-correlation estimates.
-Note that this is only one approach to calculating the ESS.
+Note that other approaches to estimating the true auto-correlation of the Markov chain and the ESS can be found in the literature.
 Details about this ESS formula, the calculation of the auto-correlation and other ESS calculations can be found in \citet{vehtari2021rank} and \citet{geyer1992practical}.
 
-            $$
-            \begin{aligned}
-                n_{\text{eff}} = \frac{NM}{1+2 \sum_{l=1}^L \hat{\rho}_l}
-            \end{aligned}
-            $$
-
 \section{MCMC Strategies}
+This section defines and compares the two MCMC algorithms used to estimate the SV model.
+Section 4.1 and Section 4.2 describes Stan's No-U-Turn sampler and KSC's mixture offset MCMC respectively.
+Section 4.3 compares the similarities and differences between both methods.
+Section 4.4 describes model reparamterisation in the context of each MCMC algorithm.
     \subsection{Sampling method 1: Stan's Hamiltonian Monte Carlo with No-U-Turn Sampler}
@@
 
-    \subsection{Sampling method 2: Mixture off-set MCMC}
+    \subsection{Sampling method 2: Off-set Mixture MCMC}
         \citet{kim1998stochastic} sample the conditional posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings within Gibbs algorithms and an implementation of the Kalman Filter and smoother used to sample from the conditional posterior distribution of the latent states \citep{dejong1995}.\footnote{In this research the exact software to apply the simulation smoother is unavailable.
-So a more recent simulation smoother is used which is written by the same author of the original software.} Note that for the simulation smoother to produce an appropriate posterior sample draw from the latent state vector, the state and measurement equations are required to be linear and conditionally Gaussian.
+So a more recent simulation smoother is used from the statsmodels Python package by \citet{seabold2010statsmodels}.
+The simulation smoother software within the statsmodels package is based on \citet{durbin2012time}.
+Koopman who co-authored \citet{durbin2012time} also co-authored the software used by \citet{kim1998stochastic} which can be found at \citet{koopman1996ssfpack}.} Note that for the simulation smoother to produce an appropriate posterior sample draw from the latent state vector, the state and measurement equations are required to be linear and conditionally Gaussian.
 Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring and taking the log of $y_t$, resulting in
         \begin{align}
-        y_t^{*} &= log(y_t^2) \notag\\ 
+        y_t^{*} &= log(y_t^2 + c) \notag\\ 
         &= log((\epsilon_t exp(h_t/2))^2) \notag\\
@@
         \end{align}
-        Where $z_t = log(\epsilon_t^2)$ has a distribution that is like that of the logarithm of a chi-squared random variable.
+        where $z_t = log(\epsilon_t^2)$ has a distribution that is like that of the logarithm of a chi-squared random variable.
 It is known that the error $z_t$ has mean -1.2704 and variance 4.93, however its distribution is not Gaussian.
-However, by transforming the entire measurement equation, as in (\ref{eqn:transform}), the relationship between $y_t$ and $h_t$ is now linear; however, the error is not Gaussian.
+The offset $c$ is set to $c=0.001$ to ensure the model is robust to small $y_t^2$ values \citep{fuller1996introduction}.
+However, by transforming the entire measurement equation, as in (\ref{eqn:transform}), the relationship between $y_t^{*}$ and $h_t$ is now linear; however, the error is not Gaussian.
 Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of seven Gaussian distributions to approximate the distribution of $z_t$, with the mixture selected to match the first 4 moments of the log chi squared distribution.
-The mixture distribution is defined by:
+The mixture distribution is defined by
         \begin{align}
@@
                  \textbf{HMC}&  Centered HMC& Reparameterised HMC\\ \hline 
-                 \textbf{Mixture off-set MCMC (KSC)}&  Centered KSC& Reparameterised (Non-Centered) KSC\\ \hline
+                 \textbf{Off-set mixture MCMC (KSC)}&  Centered KSC& Reparameterised (Non-Centered) KSC\\ \hline
             \end{tabular}
@@
     SBC is conducted for each MCMC sampling method and corresponding parameterisation as outlined in Table \ref{tab:params}.
-For a given MCMC and parameterisation pair, 5000 SBC iterations are run using simulated data sets of size T=1000.
+The SV model in Section 2.1 is used for all simulations including the same priors.
+For a given MCMC and parameterisation pair, 5000 SBC iterations\footnote{Results for 1000 SBC iterations returned noisy estimates and can be found in Appendix D.} are run using simulated data sets of size T=1000.
 5000 rank statistics are obtained per univariate parameter from this process.
@@
     Rank statistic histograms are constructed with 20 bins for the static parameters and the 1st, 500th and 1000th latent state variables.
-The black horizontal line represents a uniform distribution at 250 which is the ratio of  5000 SBC iterations and 20 bins.
+The black horizontal line represents the expected counts in each bin if the algorithm is returning calibrated estimates and can be viewed as the uniform distribution.
+This is calculated to be 250 which is the ratio of  5000 SBC iterations and 20 bins.
 The $\chi^2$ statistic is used to summarise the deviations from uniformity for all state variables since it is unrealistic to visually inspect all the histograms.
 The SBC algorithm is summarised in Algorithm \ref{alg:sbc}.
+
+    Section 5.1 and Section 5.2 interpret the SBC results for HMC and KSC's MCMC method respectively.
+Section 5.3 describes the importance weights used to correct for approximation error when applying the off-set mixture model and applying these weights to rank statistics.
+Section 5.4 presents a summary of rank statistic distributions for all latent state variables using chi-squared statistics.
 
@@
 
-    \subsection{Hamiltonian Monte Carlo}
+    \subsection{Hamiltonian Monte Carlo (No-U-Turn Sampler)}
     SBC results for the centered model using HMC are displayed in Figure \ref{fig:cphmc5k}.
@@
 The re-parameterised model possess stronger evidence of calibration for the parameters examined, producing more reliable and consistent inference.
-
         \begin{figure}[H]
@@
     \end{figure}
-
     \begin{figure}[H]
@@
 The rank statistic distributions for $\sigma^2$ and $\phi$ appear to follow a uniform distribution.
-This suggests that the HMC algorithm is returning the correct posterior estimates for the displayed parameters.}
+This suggests that the HMC algorithm is returning the correct posterior estimates on average for the displayed parameters.}
         \label{fig:ncphmc5k}
@@
 Figure \ref{fig:hmcess} and Table \ref{tab:hmcess} summarises the distribution of ESS for the static parameters after 5000 SBC iterations.
-The non centered model appears to generate independent samples much more efficiently than the centered model.
-This is consistent with HMC struggling to sample overall from the posteriors of $\sigma^2$ and $\phi$.
+The reparameterised model generates independent samples much more efficiently than the centered model, reflected in the larger ESS values across every summary statistic.
+This is consistent with HMC struggling to sample overall from the posteriors of $\sigma^2$ and $\phi$ in the centered model.
 
@@
         \end{tabular}
-        \caption{ESS for centered and re-parameterised SV model across 5000 SBC iterations for HMC.}
+        \caption{ESS for centered and re-parameterised SV model across 5000 SBC iterations for HMC.
+The reparameterised model has higher WAA counts for each summary statistic relative to the centered parameterisation.}
         \label{tab:hmcess}
@@
 
-    \subsection{Gaussian Mixture Approximation}
-    The Gaussian Mixture Approximation is run with the first 10,000 samples discarded as burn-in (following the method of \citet{kim1998stochastic}).
-However, unlike KSC who take 750,000 posterior draws from the model, the number of posterior samples taken as part of the SBC procedure is 9,999.
-Taking 750,000 draws from the Gaussian mixture approximation would lead to issues due to disk space limitations and computational resources required to run SBC.
+    \subsection{Mixture off-set MCMC}
+    The Gaussian off-set mixture approximation is run with the first 10,000 samples discarded as burn-in (following the method of \citet{kim1998stochastic}).
+However, unlike KSC who take 750,000 posterior draws from the model, the number of post burn-in posterior samples taken as part of the SBC procedure is 9,999.
+Taking 750,000 draws from the Gaussian off-set mixture would lead to issues due to disk space limitations and computational resources required to run SBC.
 Furthermore, the number of posterior samples is larger than the 999 taken for the SBC experiment run for Hamiltonian Monte Carlo.
-Earlier experiments attempted 999 post burn in draws for the Gaussian approximation, however there are concerns that such a short chain would not have converged onto the target posterior distribution.
+Earlier experiments attempted 999 post-burn in draws for the off-set mixture model, however there are concerns that such a short chain would not have converged onto the target posterior distribution.
 Therefore 9,999 draws was chosen as a compromise under both constraints.
 
-    The both the centered and re-parameterised Gaussian mixture model contains issues with calibration.
-Figure \ref{fig:cpksc1k} presents the results from the centered model.
+    Both the centered and the non-centered Gaussian off-set mixture model contains issues with calibration.
+Figure \ref{fig:cpksc5k} presents the results from the centered model.
 There is a lack of uniformity across all parameters.
-In particular, there is a large bias on the left hand side for $\mu$ and the 1st and 500th latent state variables.
-There is also a bias on the right hand hand side of $\phi$.
-This suggests overestimation and underestimation of the prior distribution respectively.
-$\sigma^2$ and the 1000th latent state don't appear to be uniform although this may be due to noisy estimates.
-
-    \begin{figure}[H]
-        \centering
-        \includegraphics[scale=0.1]{results/ksc_cp_1k.png}
-        \caption{1000 SBC iterations for centered Gaussian mixture approximation model.
-The rank statistics for all the selected parameters display a lack of uniformity.
-$\mu$ has a large left peak suggesting the algorithm is over estimating the true parameter.
-The converse can be said about $\phi$ which possesses a peak on the right hand side which implies underestimation of the true parameter.
-1st and 500th latent state variables also appear to be over estimating the true value.}
-        \label{fig:cpksc1k}
-    \end{figure}
-
-    The non centered model in location exhibits larger issues in calibration as shown in Figure \ref{fig:ncpksc1k}.
-Each histogram exhibits major peaks at one or both ends of the histgoram suggesting major issues in the sampler returning correct posterior estimates.
-The lack of uniformity for this model is distinctly due to a different shape.
-Whereas the lack of uniformity for some of the centered SBC results (for KSC) with 1000 iterations may be due to noisy estimates.
-
-    \begin{figure}[H]
-        \centering
-        \includegraphics[scale=0.1]{results/ksc_ncp_1k.png}
-        \caption{1000 SBC iterations for non centered in location Gaussian mixture approximation model.
+In particular, there is a large peak on the left hand side for $\mu$ and the latent state variables, indicating that the posterior samples on average are overestimating the prior distribution.
+There is also a bias on the right hand hand side of $\phi$, implying underestimation of the prior by the average posterior over the SBC iterations.
+
+    The non-centered model in location exhibits larger issues in calibration as shown in Figure \ref{fig:ncpksc5k}.
+Each histogram exhibits large relative frequencies at one or both ends, suggesting major issues in the sampler returning correct posterior estimates.
+The KSC MCMC algorithm applied to this parameterisation of the model is uncalibrated looking at these parameters.
+Based on the results of both parameterisations, applying this MCMC algorithm on real data will on balance produce incorrect and unreliable posterior samples on average (conditional on the model and the number of post burn-in draws).
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.09]{results/ksc_cp_5k.png}
+        \caption{\textbf{5000 SBC iterations for centered Gaussian mixture approximation model}.
+The rank statistics for all the selected parameters display deviations from uniformity.
+Large relative frequencies (or peaks) exist on the left-hand side for $\mu$ and the latent state variables and there is a peak on the right-hand side for $\phi$.}
+        \label{fig:cpksc5k}
+    \end{figure}
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.09]{results/ksc_ncp_5k.png}
+        \caption{\textbf{5000 SBC iterations for non centered Gaussian mixture approximation model}.
 The rank statistics for all selected parameters are non uniform in shape.
-The KSC bespoke MCMC struggles to return the correct posteriors for this parameterisation of the model.}
-        \label{fig:ncpksc1k}
-    \end{figure}
-
-    Increasing the SBC iterations to 5000 does not improve the results for either parameterisation.
-$\sigma^2$ contain arguably less noisy estimates around the uniform distribution in the centered parameterisation (although there may be evidence of a slight right bias).
-The skewed shapes of $\mu$ and $\phi$ adds further evidence that the sampling strategy is not returning the correct posterior estimates.
-All latent state variables appear to have a left side bias.
-This can be seen in Figure \ref{fig:cpksc5k}.
-
-    Figure \ref{fig:ncpksc5k} shows no improvement to calibration results.
-The issues in calibration observed in Figure \ref{fig:ncpksc1k} are not due to noisy estimates and are likely the result of problems in the MCMC algorithm and model specification.
-Applying this parameterisation and MCMC on real data will on balance produce incorrect posterior estimates on average.
-
-    \begin{figure}[H]
-        \centering
-        \includegraphics[scale=0.09]{results/ksc_cp_5k.png}
-        \caption{5000 SBC iterations for centered Gaussian mixture approximation model.
-Estimates for $\sigma^2$ may have marginally improved around the uniformity there is still evidence of a ride side bias.
-There is no improvement in $\mu$ or $\phi$ and all latent state variables appear to have a left side bias.}
-        \label{fig:cpksc5k}
-    \end{figure}
-
-    \begin{figure}[H]
-        \centering
-        \includegraphics[scale=0.09]{results/ksc_ncp_5k.png}
-        \caption{5000 SBC iterations for non centered Gaussian mixture approximation model.
-No improvement is observed in the distribution of rank statistics.
-The MCMC returns un-calibrated posterior estimates for this parameterisation of the model.}
+The KSC MCMC appears uncalibrated for both parameterisations of the model.}
         \label{fig:ncpksc5k}
@@
 The numerical results are summarised in Figure \ref{fig:kscess} and Table \ref{tab:kscess}.
-The estimate for $\mu$ in the re-parameterised model exhibits a long right tail (longer than the 10,000 posterior draws suggesting the presence of negatively auto-correlated draws).
+The estimate for $\mu$ in the re-parameterised model exhibits a long right tail (longer than the 9,999 post burn-in posterior draws suggesting the presence of negatively auto-correlated draws).
 There is some improvement in efficiency from the re-parameterised model with higher ESS values across most of the static parameter quantiles.
-Overall, the KSC bespoke MCMC is inefficient at sampling the static parameters.
+Overall, the off-set mixture MCMC is inefficient at sampling the static parameters.
 
@@
         \caption{Effective sample sizes for static parameters after 5000 iterations of SBC and 1000 post burn-in samples for the Gaussian Mixture model.
-The bespoke sampling strategy struggles to generate independent samples for all static parameters, suggesting a high degree of auto-correlation in the Markov Chain.}
+The KSC sampling strategy struggles to generate independent samples for all static parameters, suggesting a high degree of auto-correlation in the Markov Chain.}
         \label{fig:kscess}
@@
 
-    The results from the KSC algorithm indicates issues with calibration.
+    The results from the off-set mixture algorithm indicates issues with calibration.
 The centered parameterisation appears to be more favourable for this sampling approach.
@@
 The re-weighting procedure ensures that samples are drawn from the correct posterior density.
-This is applied to the calculation of the expected value of the stochastic volatility posterior density using samples drawn from the Gaussian mixture model.
+This is applied to the calculation of the expected value of the stochastic volatility posterior density using samples drawn from the approximating off-set mixture model.
 
@@
 
-    \subsubsection{Re-weighting rank statistics of the Gaussian Mixture Model}
-    Let $v(\theta, h)$ be the log weights defined as the log difference between the posterior densities of the true model with log chi squared errors $log\: c(\theta, h | y^{\ast})$ and the Gaussian mixture model $log\:  d(\theta, h | y^{\ast})$.
-
-    $$
-    \begin{aligned}
-        v(\theta, h) = log\: c(\theta, h | y^{\ast}) - log\:  d(\theta, h | y^{\ast}) = \text{const} + log\: c(y|h) - log\: d(y^{\ast} | h)
-    \end{aligned}
-    $$
-
-    Take the exponential and normalise the weights for the $b^{th}$ posterior draw (note the constants cancel out):
-
-    $$
-    \begin{aligned}
+    \subsubsection{Re-weighting rank statistics of the Gaussian off-set mixture model}
+    Let $v(\theta, h)$ be the log weights defined as the log difference between the posterior densities of the true model with log chi squared errors $log\: c(\theta, h | y^{\ast})$ and the mixture off-set model $log\:  d(\theta, h | y^{\ast})$.
+The is given in the following expression
+    \begin{align}
+        v(\theta, h) = log\: c(\theta, h | y^{\ast}) - log\:  d(\theta, h | y^{\ast}) = \text{const} + log\: c(y|h) - log\: d(y^{\ast} | h).
+    \end{align}
+
+    Taking the exponential and normalising the weights for the $b^{th}$ posterior draw (note the constants cancel out) returns
+    \begin{align}
     w^b = \frac{exp(v(\theta, h)_b)}{\sum_z exp(v(\theta, h)_z)}
-    \end{aligned}
-    $$
-
-    This gives the normalised importance weight.
+    \end{align}
+    which gives the normalised importance weight.
 The expectation for any function of the posterior samples can be written as a function of these weights.
 Let $S(\theta)$ be an indicator random variable that is a function of the posterior samples.
-The expectation can be written as:
-
-    $$
-    \begin{aligned}
-    S(\theta) &= 1[\theta_b < \theta^{sim}] \\
-    E[S(\theta) | y] &= \int s(\theta) c(\theta | y) d\theta\\ 
-    &= \frac{\int s(\theta)\times exp(v(\theta, h)) * d(\theta, h | y^{\ast})d\theta d h}{\int exp(v(\theta, h)) * d(\theta, h | y^{\ast})d\theta d h} 
-    \end{aligned}
-    $$
-
-    Therefore, the expectation of the re-weighted posterior samples is given by:
-
-    $$
-    \begin{aligned}
-    E[S(\hat{\theta}) | y^{\ast}] = \sum_b^B S(\hat{\theta}_b)w_b
-    \end{aligned}
-    $$
-
-    The rank statistic can be rewritten as a function of the expectation and weights. This gives us the re-weighted rank statistics.
-
-    $$
-    \begin{aligned}
-    r = \sum_{b=1}^{B}1[\theta_{b} < \theta^{sim}] \approx  B\times E[S(\hat{\theta})] = B\times \sum_b^B S(\hat{\theta}_b)w_b
-    \end{aligned}
-    $$
-
-    The results from applying this re-weighting step to the rank statistics of the centered model are given on Figure \ref{fig:reweight1k}.
+The expectation is given as
+    \begin{align}
+    S(\theta) &= 1[f(\theta_b) < f(\theta^{sim})], \\
+    E[S(\theta) | y] &= \int S(\theta) c(\theta | y) d\theta\\ 
+    &= \frac{\int S(\theta)\times exp(v(\theta, h)) * d(\theta, h | y^{\ast})d\theta d h}{\int exp(v(\theta, h)) * d(\theta, h | y^{\ast})d\theta d h}.
+    \end{align}
+
+    Therefore, the expectation of the re-weighted posterior samples can be expressed by
+    \begin{align}
+    E[S(\hat{\theta}) | y^{\ast}] = \sum_b^B S(\hat{\theta}_b)w_b.
+    \end{align}
+
+    The rank statistic can be rewritten as a function of the expectation and weights. This gives us the re-weighted rank statistics
+    \begin{align}
+    r = \sum_{b=1}^{B}1[f(\theta_{b}) < f(\theta^{sim})] \approx  B\times E[S(\hat{\theta})] = B\times \sum_b^B S(\hat{\theta}_b)w_b.
+    \end{align}
+
+    The results from applying this re-weighting step to the rank statistics of the centered model are given on Figure \ref{fig:reweight5k}.
 There are no improvements in the distribution of rank statistics.
-Increasing the SBC iterations to 5000 as seen in Figure \ref{fig:reweight5k} also show no major improvements.
-The shape of both sets of histograms is consistent with the shape of the unweighted rank statistics.
-Overall the re-weighting of the posterior samples from the approximate model does not improve the calibration results.
+The shape of the histograms is consistent with the shape of the unweighted rank statistics.
+Overall re-weighting the posterior samples from the approximate model does not improve the calibration results.
 The re-weighted rank statistics for the re-parameterised model also did not improve and can be found in Appendix E.
@@
         \centering
-        \includegraphics[scale=0.1]{results/weighted_ksc_cp_1k.png}
-        \caption{1000 SBC iterations for re-weighted rank statistics from the Gaussian mixture approximation model. The shape of the histogram is consistent with the unweighted rank statistics.}
-        \label{fig:reweight1k}
-    \end{figure}
-
-    \begin{figure}[H]
-        \centering
         \includegraphics[scale=0.1]{results/weighted_ksc_cp_5k.png}
-        \caption{5000 SBC iterations for re-weighted rank statistics from the Gaussian mixture approximation model. No major improvements are observed to the uniformity of rank statistics.}
+        \caption{\textbf{5000 SBC iterations for re-weighted rank statistics from the Gaussian off-set mixture model}. The shape of the histogram is consistent with the unweighted rank statistics.}
         \label{fig:reweight5k}
@@
 
-    The state variable chi squared statistics for centered and re-parameterised HMC, centered KSC and centered importance weighted KSC are reported as histograms in Figure \ref{fig:allchisq} (using results from 5000 SBC iterations).
-Re-parameterised KSC is omitted as its poor SBC performance skews the scale of the plot so the results are hard to see.
+    The latent state variable chi squared statistics for centered and re-parameterised HMC, centered KSC and centered importance weighted KSC are reported as histograms in Figure \ref{fig:allchisq} using results from 5000 SBC iterations.
+Re-parameterised KSC is omitted as its poor SBC performance skews the scale of the plot which makes it difficult to analyse results.
 A chi squared statistic of 0 implies the sample is exactly uniformly distributed.
@@
 Both HMC simulations are much closer to zero than KSC.
-This suggests that the HMC algorithm overall produces more calibrated posterior estimates for the state variables}
+This suggests that the HMC algorithm overall produces more calibrated posterior estimates for the state variables.}
         \label{fig:allchisq}
@@
 \section{Discussion}
-Overall, Hamiltonian Monte Carlo applied on the re-parameterised model performs best based on calibration and efficiency results.
+This research applied a simulation study to evaluate the calibration and efficiency of different MCMC algorithms in the context of the stochastic volatility model.
+Simulation based calibration was used to determine whether these algorithms were returning the correct posterior estimates on average under different SV model parameterisations.
+
+Hamiltonian Monte Carlo applied on the re-parameterised model performs best based on calibration and efficiency results.
 Centered parameterisation using HMC performed moderately well but struggled to produce calibrated posterior estimates for $\sigma^2$ and $\phi$ as well as generate independent samples for these parameters.
@@
 Evaluating the effectiveness of a modelling strategy (i.e choice of model, parameterisation and sampler) on simulated data gives a lot of a priori information about how the model will perform before it is fit on real data.
-Performing SBC may help isolate confounding factors if any problems occur.
+Performing SBC may help isolate confounding factors in computation if any problems occur.
 
 The priors chosen for this study came directly from \citet{kim1998stochastic}.
-The SBC results for the Gaussian mixture may improve if the priors were more informative - although this is only conjecture.
+The SBC results for the Gaussian off-set mixture may improve if the priors were more informative - although this is only conjecture.
 Whether or not to use tighter priors depends on the modelling context and the problem at hand \citep{gelman2017prior}.
@@
 Other parameterisations of the stochastic volatility model are available as outlined in \citet{strickland2008parameterisation}, namely non centered in scale or both non centered in scale and location.
-The failure of non centered parameterisation in location for the Gaussian mixture model could be due to a variety of reasons.
+The failure of non centered parameterisation in location for the off-set mixture model could be due to a variety of reasons.
 This may be due to a particular software implementation when it comes to the Kalman Filter and Simulation Smoother (where a specific software package was applied) that was not designed to handle the non centered parameterisations.
@@
 Additionally, this simulation study does not provide any advice on the appropriateness of the stochastic volatility model on real data.
+SBC implicitly assumes the data generating process is coming from the `correct model' and does not tell us which MCMC algorithm is better or more calibrated when the model is misspecified.
 Whether the stochastic volatility model is appropriate for modelling a particular financial time series requires further study into the data generating process, domain expertise and a suite of other diagnostic checks.
@@
 Computing infrastructure may limit the opportunity for SBC to be performed.
-The choice of SBC iterations for this research was arbitrary and was used to demonstrate that noisy estimates could be reduced by increasing the number of iterations for calibrated models.
-A more principled approach to picking the number of SBC iterations could be explored in other research.
+The choice of SBC iterations for this research was arbitrary and a more principled approach to picking the number of SBC iterations could be explored in other research.
 
@@
 Other summary statistics and visualisations can be applied for a high level comparison, such as the chi squared statistics applied in this research.
-Furthermore, it is not always straightforward to understand why any any one parameter may produce poor calibration results, for example, an arbitrary latent state varaible.
+Furthermore, it is not always straightforward to understand why any any one parameter may produce poor calibration results, for example, an arbitrary latent state variable.
 SBC may tell us some information about the miscalibration based on the histogram shape (e.g.
-underdispersed, overdispersed, or biased relative to the prior distribution), but understanding why this occurs may not be immediately clear.
+average posteriors underestimating or overestimating the prior distribution, underdispersed or overdispersed average posteriors relative to the prior distribution), but understanding why this occurs may not be immediately clear.
 Although, diagnosing sampling problems with complicated posterior geometries speaks more generally to the difficulty of understanding complex models in the first place as opposed to a limitation of SBC.
@@
 Lastly, it is unclear what constitutes a fair comparison between algorithms when comparing SBC results.
-In particular, the number of post burn-in or warm-up samples differed between algorithms (999 post burn-in for HMC, 9999 for the bespoke algorithm).
+In particular, the number of post burn-in or warm-up samples differed between algorithms (999 post burn-in for HMC, 9999 for the bespoke KSC algorithm).
 An argument could be made that the poor results for the bespoke KSC algorithm may be due to the chain not converging onto the target distribution.
 Indeed, the authors of this approach extended their Markov chain to 750,000 posterior draws (discarding the first 10,000 draws as burn-in).
-A fairer comparison may be to have roughly similar number of ESS for most of the estimated parameters, although this is difficult to get right with a large number of parameters and variables to estimate.
-Given how well Hamiltonian Monte Carlo performed with the re-parameterised model, any improvement to the Gaussian mixture model by increasing the length of the Markov chain would make calibration just as good, but not better due to the inefficiency of the sampler.
+A fairer comparison may be to have roughly similar number of ESS for most of the estimated parameters, although this is difficult to get right with a large number of parameters to estimate.
+Given how well Hamiltonian Monte Carlo performed with the re-parameterised model, any improvement to the Gaussian off-set mixture model by increasing the length of the Markov chain would make calibration just as good, but not better due to the inefficiency of the sampler.
 
@@
 
-\section{Appendix B: Sampling from mixture of Gaussian's}
+\section{Appendix B: Sampling from Gaussian off-set mixture model}
     The steps in the MCMC process is described in Algorithm \ref{alg:ksc}.
@@
 
-    Instead, we observe a peak on the left side for the 500th state variable $h_{500}$, although the estimates are noisy. Overall the estimates for non centered parameterisation are more favourable.
+        \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/ksc_cp_1k.png}
+        \caption{\textbf{1000 SBC iterations for centered Gaussian off-set mixture model}.
+The rank statistics for all the selected parameters display a lack of uniformity.
+$\mu$ has a large left peak suggesting the algorithm is over estimating the true parameter.
+The converse can be said about $\phi$ which possesses a peak on the right hand side which implies underestimation of the true parameter.
+1st and 500th latent state variables also appear to be over estimating the true value.}
+        \label{fig:cpksc1k}
+    \end{figure}
+
+    \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{results/ksc_ncp_1k.png}
+        \caption{\textbf{1000 SBC iterations for non centered in location Gaussian off-set mixture model}.
+The rank statistics for all selected parameters are non uniform in shape.
+The KSC bespoke MCMC struggles to return the correct posteriors for this parameterisation of the model.}
+        \label{fig:ncpksc1k}
+    \end{figure}
 
@@
         \centering
+        \includegraphics[scale=0.1]{results/weighted_ksc_cp_1k.png}
+        \caption{1000 SBC iterations for re-weighted rank statistics from the Gaussian mixture approximation model. The shape of the histogram is consistent with the unweighted rank statistics.}
+        \label{fig:reweight1k}
+    \end{figure}
+
+\begin{figure}[H]
+    \centering
     \includegraphics[scale=0.09]{results/weighted_ksc_ncp_5k.png}

### 2023-10-18 manuscript/main.tex
Updates from Overleaf

@@
 However, we can only conclude that there are differences between these methods, not which is correct.
+        \begin{figure}[H]
+        \centering
+        \includegraphics[scale=0.1]{motivating_example/real_data_ex.png}
+        \caption{\textbf{Posterior distribution for static parameters fit on the S\&P 500 Index.} Posterior samples from the HMC (blue) and KSC (red) MCMC samplers.
+Marginal posterior densities are presented for $\mu$, $\phi$ and $\sigma^2$.
+Note that each panel has their own scale for the x and y axis.
+The parameter $\phi$ falls}
+        \label{fig:realdataex}
+    \end{figure}
 
     % Other methods for model selection in this context (for example, out of sample predictive performance), but for parameter estimation it is unclear which one should be selected. 
-
-    \begin{figure}[H]
-        \centering
-        \includegraphics[scale=0.1]{motivating_example/real_data_ex.png}
-        \caption{\textbf{Posterior distribution for static parameters fit on the S\&P 500 Index.} Posterior samples from the HMC (blue) and KSC (red) MCMC samplers.}
-        \label{fig:realdataex}
-    \end{figure}
-
     \begin{table}[H]
@@
         \includegraphics[scale=0.1]{motivating_example/single_sim.png}
-        \caption{\textbf{Posterior samples from SV model fit on known data generating process}.
+        \caption{\textbf{HMC posterior samples from SV model fit on known data generating process}.
+Marginal posterior densities are presented for $\mu$, $\phi$ and $\sigma^2$.
 Vertical solid line represents true parameter and dotted lines are 95\% credible intervals around the mean.
+Note that each panel has their own scale for the x and y axis.
 The parameter $\phi$ falls outside the credible interval whereas $\mu$ and $\sigma^2$ stay inside.}
@@
 
-            Figure \ref{fig:underestimation} shows a distribution of rank statistics with a large right peak.
+            Figure \ref{fig:underestimation} shows a distribution of rank statistics with inflated frequency on the right side of the histogram.
 This distribution is consistent with posterior samples that tend to underestimate the true parameter on average.
 The sum of indicator random variables (see equation (13)) is large since a large proportion of posterior draws are smaller than the true value, resulting in a large rank statistic and hence a larger relative frequency on the right side of the histogram.
-The converse is true if the peak of the histogram is on the left hand side (i.e the average posterior over the SBC iterations overestimates prior).
-
-            Figure \ref{fig:underdispersed} has large peaks on both ends of the histogram resulting in a `U-shape'.
+The converse is true if the inflated frequency of the histogram is on the left hand side (i.e the average posterior over the SBC iterations overestimates prior).
+
+            Figure \ref{fig:underdispersed} has large inflated frequencies on both ends of the histogram resulting in a `U-shape'.
 \citet{talts2020validating} describe this bias as under-dispersion of the posterior relative to the prior distribution on average.
@@
                 \includegraphics[scale=0.07]{methodology/rhs.png}
-                \caption{Left: Non uniform rank statistics with peak on the right side.
-Right: Posterior distribution is biased to the left side of the prior distribution.
+                \caption{Left: Non uniform rank statistics with inflated frequency on the right side.
+Right: Average posterior distribution is biased to the left side of the prior distribution.
 Bias in rank statistics distribution comes from posterior samples underestimating the true parameter on average.
@@
                 \includegraphics[scale=0.07]{methodology/underdispersed.png}
-                \caption{Left: Non uniform rank statistics with peaks on both ends (`U-shape' distribution).
+                \caption{Left: Non uniform rank statistics with inflated frequencies on both ends (`U-shape' distribution).
 Right: Overlayed posterior and prior distributions for the same arbitrary parameter.
@@
 The centered parameterisation displays a lack of uniformity in rank statistics for $\sigma^2$ and $\phi$.
-The $\sigma^2$ parameter displays a U-shaped histogram with large peaks on both ends while $\phi$ exhibits a U-shape with smaller peaks.
+The $\sigma^2$ parameter displays a U-shaped histogram with inflated frequencies on both ends while $\phi$ exhibits a U-shape with relatively smaller inflated frequencies.
 This indicates the MCMC algorithm is not returning calibrated estimates for these parameters.
@@
 $\mu$ and the state parameters look approximately uniform.
-$\sigma^2$ still has major peaks on both the left and right hand side indicating under-dispersion relative to the prior distribution.
-$\phi$ exhibits a similar U-shape to $\sigma^2$ although the peaks are much smaller.}
+$\sigma^2$ still has inflated frequencies on both the left and right hand side indicating under-dispersion relative to the prior distribution.
+$\phi$ exhibits a similar U-shape to $\sigma^2$ although the U-shape is much smaller.}
         \label{fig:cphmc5k}
@@
 There is a lack of uniformity across all parameters.
-In particular, there is a large peak on the left hand side for $\mu$ and the latent state variables, indicating that the posterior samples on average are overestimating the prior distribution.
-There is also a bias on the right hand hand side of $\phi$, implying underestimation of the prior by the average posterior over the SBC iterations.
+In particular, there is an inflated frequency on the left hand side for $\mu$ and the latent state variables, indicating that the posterior samples on average are overestimating the prior distribution.
+There is also an inflated frequency on the right hand hand side of $\phi$, implying underestimation of the prior by the average posterior over the SBC iterations.
 
     The non-centered model in location exhibits larger issues in calibration as shown in Figure \ref{fig:ncpksc5k}.
-Each histogram exhibits large relative frequencies at one or both ends, suggesting major issues in the sampler returning correct posterior estimates.
+Each histogram exhibits large inflated frequencies at one or both ends, suggesting major issues in the sampler returning correct posterior estimates.
 The KSC MCMC algorithm applied to this parameterisation of the model is uncalibrated looking at these parameters.
@@
 The rank statistics for all the selected parameters display deviations from uniformity.
-Large relative frequencies (or peaks) exist on the left-hand side for $\mu$ and the latent state variables and there is a peak on the right-hand side for $\phi$.}
+Large inflated frequencies exist on the left-hand side for $\mu$ and the latent state variables and there is an inflated frequency on the right-hand side for $\phi$.}
         \label{fig:cpksc5k}
@@
         \includegraphics[scale=0.1]{results/dist_chisq_all.png}
-        \caption{Distribution of chi squared statistics for latent state variables.
+        \caption{\textbf{Distribution of chi squared statistics for latent state variables}.
 Results from 5000 SBC iterations are used.

### 2023-10-19 manuscript/main.tex

@@
 \author{Benjamin Wee}
-\date{October 2023}
-
+\date{\small October 2023\\[0.5cm]{\small Supervised by Professor Catherine Forbes and Dr Lauren Kennedy}}
+% \author{It's me\\[1cm]{\small Advisor: A.D.Visor}}
 \begin{document}
@@
 \begin{abstract}
-
-    This research uses the simulation-based calibration (SBC) method of \citet{talts2020validating} to analyse two commonly used, competing Markov chain Monte Carlo algorithms for the purpose of estimating a posterior distribution in the context of a stochastic volatility model.
+    Simulation based calibration (SBC) \citep{talts2020validating} is applied to analyse two commonly used, competing Markov chain Monte Carlo algorithms for estimating the posterior distribution of a stochastic volatility model.
+In particular, the bespoke ‘off-set mixture approximation’ algorithm proposed by \citet{kim1998stochastic} is explored together with a Hamiltonian Monte Carlo algorithm implemented through Stan \citep{stan}.
+The SBC analysis involves a simulation study to assess whether each sampling algorithm has the capacity to produce valid inference for the correctly specified model, while also characterising statistical efficiency through effective sample size.
+Results show that Stan's No-U-Turn sampler, an implementation of Hamiltonian Monte Carlo, produces a well-calibrated posterior estimate while the celebrated off-set mixture approach is less efficient and poorly calibrated, though model parameterisation also plays a role.
+Limitations and restrictions of generality are discussed.
+
+    % This research uses the simulation-based calibration (SBC) method of \citet{talts2020validating} to analyse two commonly used, competing Markov chain Monte Carlo algorithms for the purpose of estimating a posterior distribution in the context of a stochastic volatility model.
 In particular, the bespoke ‘off-set mixture approximation’ algorithm proposed by \citet{kim1998stochastic} is explored together with a generic Hamiltonian Monte Carlo algorithm implemented through Stan \citep{stan}.
@@
 Typically expressed as a non-Gaussian state space model where the conditional variance of the observed return is treated as a latent random variable \citep{hull1987pricing, chesney1989pricing}, such models are difficult to estimate since the likelihood is unavailable in closed form.
-With more parameters than observations,\footnote{We refer to both static unknowns and time-varying latent state variables as 'parameters'.} SV models are commonly analysed using Bayesian methods obtained via a Markov Chain Monte Carlo (MCMC) algorithm designed to sample from the high dimensional joint posterior distribution.
+With more parameters than observations\footnote{We refer to both static unknowns and time-varying latent state variables as 'parameters'.}, SV models are commonly analysed using Bayesian methods obtained via a Markov Chain Monte Carlo (MCMC) algorithm designed to sample from the high dimensional joint posterior distribution.
 The focus of this paper is to investigate the performance of two competing MCMC algorithms, each applied to a SV model, in terms of the validity and statistical efficiency of the their resulting inference.
@@
 
-    \citet{kim1998stochastic} propose a sampling strategy for the so-called 'vanilla' stochastic volatility through an off-set mixture approximation.
+    \citet{kim1998stochastic} propose a sampling strategy for the so-called `vanilla' stochastic volatility through an off-set mixture approximation.
 Conditional on the mixture allocations and static parameters, the forward filter backward sampling algorithm \citep{carter1994gibbs, fruhwirth1995bayesian} is used to jointly sample the latent states while conditionally  conjugate prior distributions are also suggested.
 Since then, there have been many advances in statistical computing and algorithm design.
-In particular, Hamiltonian Monte Carlo (HMC) has be proposed as a generic and efficient sampling method for MCMC.
-Notably, HMC has been shown to be particularly well-suited to complex, high dimensional probabilistic models.
+In particular, Hamiltonian Monte Carlo (HMC) has been proposed as a generic and efficient sampling method for MCMC.
 Importantly, the adoption of such new techniques have been made widely available through the development of various open source libraries such as the Stan programming language \citep{stan} and the PyMC library \citep{pymc2023}.
@@
 The second algorithm replicates Kim Shephard and Chib's (KSC) Gaussian off-set mixture model which is sampled using the aforementioned Kalman Filter and Metropolis Hastings algorithm.
-The objective is to analyse the efficiency of the MCMC and their ability to return correct posterior estimates.
+The objective is to analyse the efficiency of the MCMC and their ability to return correct posterior estimates on average.
 
@@
     One particular challenge to comparing models is the parameterisation.
-Different parameterisations of a model, whilst the same mathematically, can be very different in terms of sampling difficulty \citep{neal2003slice}.
+Different parameterisations of a model, whilst mathematically the same, can be very different in terms of sampling difficulty \citep{neal2003slice}.
 In this work two different parameterisations of the stochastic volatility model are compared for each sampling method.
 An algorithm may be sensitive to different parameterisations of the same model which may impact MCMC performance as described in \citet{strickland2008parameterisation}.
-The key findings of this research are HMC applied on a re-parameterised SV model produces the most calibrated and efficient posterior estimates relative to KSC's MCMC algorithm on the approximate off-set mixture model.
+
+    The key findings of this research are HMC produces more calibrated and efficient posterior estimates relative to KSC's MCMC algorithm on the approximate off-set mixture model.
 Additionally, MCMC algorithms may perform better or worse based on calibration and efficiency depending on the parameterisation of the model.
+HMC applied on a re-parameterised SV model produces the most calibrated and efficient posterior estimates across all SBC experiments.
 
@@
 \section{Research Context}
-This section describes the stochastic volatility model to be analysed in this research and discusses two examples to motivate the research objectives.
-Section 2.1 outlines the model specificaion and priors, Section 2.2 explores challenges with inference and convergence diagnostics with no ground truth and Section 2.3 mentions the limitations and pitfalls of running only a single simulation.
+This section describes the stochastic volatility model and discusses two examples to motivate the research objectives.
+Section 2.1 outlines the model specification and priors, Section 2.2 explores challenges with inference and convergence diagnostics when the true model is unknown (no ground truth) and Section 2.3 discusses the limitations and pitfalls of running only a single simulation.
+Section 2.4 lays out the research objective given these motivating examples.
 
@@
     The model of interest is the discrete time, univariate stochastic volatility model estimated by \citet{kim1998stochastic} using a Bayesian framework.
-SV is expressed as a state space model which links the observed (logarithmic) return of a given asset over period t, denoted $y_t$, to its latent conditional variance, $h_{t+1}$.
+SV is expressed as a state space model which links the observed (logarithmic) return of a given asset over period $t$, denoted $y_t$, to its latent conditional variance, $h_{t+1}$.
 This model is given by the following expressions
     \begin{align}
-    y_t =& \space \,\exp(h_t/2) \varepsilon_t, \\
-    h_{t+1} =& \space \, \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t,  \\
-    h_1 \sim& \space \,\mathrm{N}\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right), \\
-    \varepsilon_t \sim& \space \,\mathrm{N}(0,1), \\
+    y_t =& \space \,\exp(h_t/2) \varepsilon_t \\
+    h_{t+1} =& \space \, \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
+    h_1 \sim& \space \,\mathrm{N}\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
+    \varepsilon_t \sim& \space \,\mathrm{N}(0,1) \\
     \eta_t \sim& \space \,\mathrm{N}(0,1).
@@
     \begin{align}
-    \mu \sim& \space \,\mathrm{N}(0, 10^2) \\
+    \mu \sim& \space \,\mathrm{N}(0, 10) \\
     \sigma_{\eta}^2 \sim& \space \, \mathrm{IG}(shape=5/2, scale=(0.01\times 5) / 2) \\
     \phi^{\ast} \sim& \space \,\mathrm{Beta}(20, 1.5) \\
-    \phi =& \space \, 2\phi^{\ast} - 1
+    \phi =& \space \, 2\phi^{\ast} - 1.
     \end{align}
@@
 Furthermore, different sampling strategies will result in different posterior estimates for the same model.
-Assuming no issues with the computation or the model specification, convergence diagnostics do not provide any guidance to the analyst regarding which posterior estimate is to be preferred since there is no ground truth.
+Assuming no issues with the computation or the model specification, convergence diagnostics do not provide any guidance to the analyst regarding which posterior estimate is to be preferred.
 
@@
         \includegraphics[scale=0.1]{motivating_example/real_data_ex.png}
-        \caption{\textbf{Posterior distribution for static parameters fit on the S\&P 500 Index.} Posterior samples from the HMC (blue) and KSC (red) MCMC samplers.
+        \caption{\textbf{Posterior distribution for static parameters fit on the S\&P 500 Index.} Posterior samples are drawn from the HMC (blue) and KSC (red) MCMC algorithms.
 Marginal posterior densities are presented for $\mu$, $\phi$ and $\sigma^2$.
-Note that each panel has their own scale for the x and y axis.
-The parameter $\phi$ falls}
+Note that each panel has their own scale for the x and y axis.}
         \label{fig:realdataex}
@@
         \end{tabular}
-        \caption{\textbf{Summary statistics for HMC and KSC algorithms}. Both sets of estimates appear reasonable but it is unclear from estimates on real data which estimate is closer the the truth.}
+        \caption{\textbf{Summary statistics for HMC and KSC algorithms}. Both sets of estimates appear reasonable but it is unclear from estimates on real data which estimate is closer to the truth.}
         \label{tab:realdata}
@@
 \subsection{Challenge 2: Limitations of a single simulation}
-    Diagnosing problems with the approximate posterior from models fit to real data is difficult because we do not have a ground truth to compare with.
+    Diagnosing problems with the approximate posterior from models fit to real data is difficult because we do not observe the true data generating process or the true model.
 A strategy around this is to evaluate a model and algorithm on simulated data.
@@
 \citet{gelman2020bayesian} discuss that simulation is the only way we can check inference on latent variables.
-This is critical for the SV model since the underlying framework is a state space model with latent log volatility parameters.
+This is critical for the SV model since the underlying framework is a state space model with latent log variance parameters.
 Latent variables are unobserved in real data and are only estimated in the context of the model.
@@
     Figure 2 shows marginal posterior distributions given a single set of 1000 returns generated by the SV model.
-The true parameters are $\mu=-0.389$, $\sigma^2_{\nu}=0.0115$ and $\phi=0.547$ and the model is fit using Stan's implementation of HMC, the No-U-Turn sampler.
-The 95\% credible intervals for $\mu$ and $\sigma^2$ cover the true parameter.
+The true parameters are $\mu=-0.389$, $\sigma^2_{\eta}=0.0115$ and $\phi=0.547$ and the model is fit using Stan's implementation of HMC, the No-U-Turn sampler.
+The 95\% credible intervals for $\mu$ and $\sigma^2_{\eta}$ cover the true parameter.
 The true parameter for $\phi$ however, is in the tails and outside the interval.
@@
 \section{Methodology}
-The methodology section describes the simulation study used validate the posterior estimates from MCMC algorithms and the metrics used to empirically measure calibration and efficiency.
-Section 3.1 discusses the procedure for running SBC and the key results for defining calibration.
-Section 3.2 defines the diagnostics and metrics for evaluating SBC experiment results.
+This section introduces the simulation design used to evaluate whether a MCMC algorithm is returning the correct posterior estimates on average (that is, whether a MCMC algorithm calibrated).
+Section 3.1 discusses the procedure for running simulation based calibration and Section 3.2 defines the diagnostics and metrics for evaluating SBC experiment results.\footnote{The software used to implement these methods and produce all results and plots in this paper are the R v4.2.3\citep{rlang}, Python v3.8.13 \citep{10.5555/1593511} and Stan v2.31.0 \citep{stan} programming languages, including open source R packages (tidyverse v1.3.2 \citep{tidyverse}, quantmod v0.4.20 \citep{quantmod}, cmdstanr 0.5.3 \citep{cmdstanr}, posterior v1.4.1 \citep{posteriorr}, bayesplot v1.4.10 \citep{bayesplot}, jsonlite v1.8.4 \citep{jsonlite}, arrow v12.0.1.1 \citep{arrow} and future.apply v1.11.0 \citep{RJ-2021-048}) and python packages (numpy v1.24.4 \citep{harris2020array}, scipy v.1.10.1 \citep{2020SciPy-NMeth}, statsmodels v.0.14.0 \citep{seabold2010statsmodels}, pandas v.2.0.3 \citep{mckinney-proc-scipy-2010} and pyarrow v12.0.1 \citep{arrow}).}
+
     \subsection{Simulation Design}
@@
 For each $k$ of $K$ SBC iterations, one value for each parameter is drawn from the prior.
-For this parameter a data set is generated from the model.
+Conditional on the drawn parameter values, a data set is generated from the model.
 Then a sampling method (originally using Stan in \citet{talts2020validating}) is used to obtain posterior samples given this data and the prior.
-Then the estimated posteriors are compared to the true values by calculating the rank.
+The estimated posteriors are then compared to the true values by calculating the rank statistics.
 Under a calibrated model, the $K$ rank statistics for each parameter should be uniformly distributed.
@@
         To illustrate the SBC procedure, let $\theta$ be an arbitrary scalar parameter from a model and $y$ represent a data set.
-Start with a single parameter $\theta^{sim}$ drawn from the prior distribution:
+Start with a single parameter $\theta^{sim}$ drawn from the prior distribution
         \begin{align}
@@
         \end{align}
-        generate a data set $y^{sim}$ conditionally given $\theta^{sim}$ and the model of interest\footnote{Since all computation presumes a correctly specified model, nottation to condition on the model is suppressed.}:
+        then generate a data set $y^{sim}$ conditionally given $\theta^{sim}$ and the model of interest\footnote{Since all computation presumes a correctly specified model, notation to condition on the model is suppressed.}:
         \begin{align}
@@
         \end{align}
-        Then take $L$ draws from the estimated posterior distribution, conditional on this data set, and generated by the MCMC algorithm under investigation (here, HMC or KSC):
+        Take $L$ draws from the estimated posterior distribution, conditional on this data set, and generated by the MCMC algorithm under investigation (here, HMC or KSC):
         \begin{align}
@@
         \begin{align}
-        r^{(k)} = rank(\{f(\theta_1^{(k)}),\dots , f(\theta_{B}^{(k)})\}, f(\theta^{sim})) = \sum_{b=1}^{B}1[f(\theta_{b}^{(k)}) < f(\theta^{sim})].
+        r^{(k)} = \mathrm{rank}(\{f(\theta_1^{(k)}),\dots , f(\theta_{B}^{(k)})\}, f(\theta^{sim})) = \sum_{b=1}^{B}1[f(\theta_{b}^{(k)}) < f(\theta^{sim})].
         \end{align}
@@
 
-        A key result is that the posterior averaged over the data and true parameters will equal the prior distribution. This is shown in the following expression:
-        \begin{align}
-        \pi(\theta) &= \int \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})d\theta^{sim} dy^{sim}, \\
+        A key result is that the posterior averaged over the data and true parameters will equal the prior distribution. This is shown in the following expression
+        \begin{align}
+        \pi(\theta) &= \int \int \pi(\theta|y^{sim}) \pi(y^{sim}|\theta^{sim}) \pi(\theta^{sim})d\theta^{sim} dy^{sim} \\
         &= \int \int \pi(\theta|y^{sim}) \pi(y^{sim},\theta^{sim}) d\theta^{sim} dy^{sim}.
@@
         That is, the average posterior for $\theta$ over the SBC iterations should approximately equal to prior distribution, if the algorithm is calibrated.
-Therefore, any deviation from the prior distribution and the average posterior draws produced from replicated posterior distributions means that the sampling methodology is not producing the correct posterior, on average.
+Therefore, any deviation between the prior distribution and the average posterior draws produced from replicated posterior distributions means that the sampling methodology is not producing the correct posterior, on average.
 
@@
             The shape of the distribution of rank statistics gives insight into how an MCMC may be miscalibrated.
-Specifically, it gives information about the average posterior deviates from the prior distribution of a given parameter.
+Specifically, it gives information about how the average posterior deviates from the prior distribution of a given parameter.
 Figures \ref{fig:underestimation} and \ref{fig:underdispersed} are recreated from \citet{talts2020validating} for convenience.
@@
             Figure \ref{fig:underdispersed} has large inflated frequencies on both ends of the histogram resulting in a `U-shape'.
-\citet{talts2020validating} describe this bias as under-dispersion of the posterior relative to the prior distribution on average.
+\citet{talts2020validating} describe this as under-dispersion of the posterior relative to the prior distribution on average.
 The average posterior is too narrow (or overconfident) relative to the spread of the prior resulting in inflated frequency on both ends of the rank statistic distribution.
@@
                 \includegraphics[scale=0.07]{methodology/rhs.png}
-                \caption{Left: Non uniform rank statistics with inflated frequency on the right side.
-Right: Average posterior distribution is biased to the left side of the prior distribution.
-Bias in rank statistics distribution comes from posterior samples underestimating the true parameter on average.
-Note that the two plots are independently produced for illustration only.}
+                \caption{\textbf{Left:} Non uniform rank statistics with inflated frequency on the right side.
+\textbf{Right:} The average posterior distribution underestimates the prior distribution.
+The miscalibration from the rank statistic distribution comes from posterior samples underestimating the true parameter on average.}
                 \label{fig:underestimation}
@@
                 \includegraphics[scale=0.07]{methodology/underdispersed.png}
-                \caption{Left: Non uniform rank statistics with inflated frequencies on both ends (`U-shape' distribution).
-Right: Overlayed posterior and prior distributions for the same arbitrary parameter.
-Non-Uniformity in ranks comes from under-dispersed posterior distribution relative to prior.}
+                \caption{\textbf{Left}: Non uniform rank statistics with inflated frequencies on both ends (`U-shape' distribution).
+\textbf{Right}: Average posterior is underdispersed relative to the prior distribution.
+Non-uniformity in rank statistics comes from posteriors which are over-confident relative to the prior.}
                 \label{fig:underdispersed}
@@
             \subsubsection{Chi squared statistics}
-            One difficulty of relying on the visualisation of rank statistic distributions for SV models, is there are a very large number of unknowns in the model to consider.
-An alternative to visually checking for uniformity of a given parameter is to calculate a chi-squared statistic based on the counts in each histogram bin.
-Suppose the histogram has $J$ bins, where for each $j^{th}$ bin, $b_j$ is the number of counts and $e_j$ the expected count in bin $j$, where the expected count is the ratio of K rank statistics and J bins, $e_j=\frac{K}{J}$.
-The corresponding chi- squared statistic is given by:
-            \begin{align}
-            \chi^2 = \sum_{j=1}^J \frac{(b_{j} - e_{j})^2}{e_j}.\end{align}
-
-            A discrete empirical distribution with the same number of realisations in each of $J$ bins will return a chi squared statistic of zero, since the number realisations in each bin is equal to the expected count under the relevant  uniform distribution.
+            One difficulty of relying on the visualisation of rank statistic distributions for SV models is there are a very large number of unknowns in the model to consider.
+An alternative to visually checking for uniformity of a given parameter is to calculate a Chi-squared statistic based on the counts in each histogram bin.
+Suppose the histogram has $J$ bins, where for each $j^{th}$ bin, $b_j$ is the number of counts and $e_j$ the expected count in bin $j$.
+The expected count is the ratio of $K$ rank statistics and $J$ bins, $e_j=\frac{K}{J}$.
+The corresponding Chi-squared statistic is denoted by
+            \begin{align}
+            \chi^2 = \sum_{j=1}^J \frac{(b_{j} - e_{j})^2}{e_j}.
+            \end{align}
+
+            A discrete empirical distribution with the same number of realisations in each of $J$ bins will return a chi squared statistic of zero, since the number of realisations in each bin is equal to the expected count under the relevant  uniform distribution.
 The distribution of chi squared statistics is visualised to compare results across multiple simulations.
 Thus, the distribution of chi-squared statistics produced from the histograms of SBC rank statistics for a large number of univariate model parameters will give a high level summary of the overall model calibration, and hence for the overall performance of the algorithm.
-By comparing the distributions of chi-squared statistic produced by competing algorithms, one algorithm's degree of calibration, relative to that of the other algorithm, may be assessed.
+By comparing the distributions of Chi-squared statistics produced by competing algorithms, one algorithm's degree of calibration, relative to that of the other algorithm, may be assessed.
 
             \subsubsection{Effective sample size}
-            Another univariate summary of a marginal posterior distribution is its effective sample size (ESS), which provides a measure of the efficiency of the MCMC sampler.
+            Another univariate summary of a marginal posterior distribution is its effective sample size (ESS), which provides a measure of efficiency of the MCMC sampler.
 ESS estimates the number of effectively independent draws generated by a MCMC to estimate a target parameter.
 A poor ESS generally arises from high auto-correlation in the Markov chain, and thus is an indication of a highly dependent posterior sample.
-A more efficient MCMC algorithm will take a relatively lower level of resources (for example, time and number of draws) to get a representative sample from the target distribution when compared to a less efficient algorithm.
+A more efficient MCMC algorithm will take a relatively lower level of resources (for example, time and number of draws) to get a sample from the target distribution when compared to a less efficient algorithm.
 If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), this is seen as evidence of greater efficiency.
 
-            The ESS is denoted as $n_{\text{eff}}$ as given by
+            The ESS is denoted as $n_{\text{eff}}$ and is given by
             \begin{align}
@@
 Section 4.3 compares the similarities and differences between both methods.
-Section 4.4 describes model reparamterisation in the context of each MCMC algorithm.
+Section 4.4 describes model reparameterisation in the context of each MCMC algorithm.
     \subsection{Sampling method 1: Stan's Hamiltonian Monte Carlo with No-U-Turn Sampler}
-        Hamiltonian Monte Carlo (HMC) is a  general MCMC algorithm used to efficiently sample posterior draws  high associated with high-dimensional and complex stochastic models.
+        Hamiltonian Monte Carlo (HMC) is a general MCMC algorithm used to efficiently sample posterior draws associated with high-dimensional and complex stochastic models.
 Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature \citep{duane1987hybrid} before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks \citep{neal1995bayesian} and statistical computing \citep{neal2011mcmc}.
@@
         A thorough conceptual and theoretical explanation of HMC can be found in \citet{gelman2013bayesian} and \citet{betancourt2017conceptual}.
-HMC builds upon the Metropolis Hastings algorithm through the introduction of Hamiltonian equations, auxiliary momentum variables $\alpha$, and the gradients of the target log posterior distribution.
+HMC builds upon the Metropolis Hastings algorithm through the introduction of Hamiltonian equations, auxiliary momentum variables, and the gradients of the target log posterior distribution.
 The vector of model parameters $\theta$ is jointly sampled with a vector of momentum variables $\alpha$.
@@
 
-        This approach will also use the same priors as specified in the KSC's Gaussian mixture approximation so that HMC is sampling from the same model (although other priors could be chosen).
+        This approach will also use the same priors as specified in the KSC's off-set mixture approximation so that HMC is sampling from the same model (although other priors could be chosen).
 
     \subsection{Sampling method 2: Off-set Mixture MCMC}
-        \citet{kim1998stochastic} sample the conditional posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings within Gibbs algorithms and an implementation of the Kalman Filter and smoother used to sample from the conditional posterior distribution of the latent states \citep{dejong1995}.\footnote{In this research the exact software to apply the simulation smoother is unavailable.
+        \citet{kim1998stochastic} sample the conditional posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings \citep{metropolis1953equation, hastings1970monte} within Gibbs \citep{geman1984stochastic} algorithms and an implementation of the Kalman Filter \citep{kalman1960new} and smoother used to sample from the conditional posterior distribution of the latent states \citep{dejong1995}.\footnote{In this research the exact software to apply the simulation smoother is unavailable.
 So a more recent simulation smoother is used from the statsmodels Python package by \citet{seabold2010statsmodels}.
@@
 Koopman who co-authored \citet{durbin2012time} also co-authored the software used by \citet{kim1998stochastic} which can be found at \citet{koopman1996ssfpack}.} Note that for the simulation smoother to produce an appropriate posterior sample draw from the latent state vector, the state and measurement equations are required to be linear and conditionally Gaussian.
-Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring and taking the log of $y_t$, resulting in
-        \begin{align}
-        y_t^{*} &= log(y_t^2 + c) \notag\\ 
-        &= log((\epsilon_t exp(h_t/2))^2) \notag\\
-        &=  log(exp(h_t)) + log(\epsilon_t^2) \notag\\
-        &= h_t + log(\epsilon_t^2) \notag  \\
-        &= h_t + z_t,  \label{eqn:transform}
+Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring, adding an offset, and taking the log of $y_t$, resulting in
+        \begin{align}
+        y_t^{*} &= log(y_t^2 + c) = h_t + z_t, \label{eqn:transform}
         \end{align}
         where $z_t = log(\epsilon_t^2)$ has a distribution that is like that of the logarithm of a chi-squared random variable.
-It is known that the error $z_t$ has mean -1.2704 and variance 4.93, however its distribution is not Gaussian.
+It is known that the error $z_t$ has mean -1.2704 and variance 4.93.
 The offset $c$ is set to $c=0.001$ to ensure the model is robust to small $y_t^2$ values \citep{fuller1996introduction}.
-However, by transforming the entire measurement equation, as in (\ref{eqn:transform}), the relationship between $y_t^{*}$ and $h_t$ is now linear; however, the error is not Gaussian.
+By transforming the entire measurement equation, as in (\ref{eqn:transform}), the relationship between $y_t^{*}$ and $h_t$ is now linear; however, the error is not Gaussian.
 Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of seven Gaussian distributions to approximate the distribution of $z_t$, with the mixture selected to match the first 4 moments of the log chi squared distribution.
@@
 
-        Given a set of $T$ mixture component indicator variables, which are added to the MCMC scheme in order to use the approximating Gaussian mixture error distribution, the latent state vector may be sampled via the Kalman Filter and simulation smoother, since the model is now conditionally linear and Gaussian.
-The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions whereas $\phi$ is sampled via a Metropolis Hastings accept/reject procedure.
+        Given a set of $T$ mixture component indicator variables, which are added to the MCMC scheme in order to use the approximating Gaussian mixture error distribution, the latent state vector may be sampled jointly via the Kalman Filter and simulation smoother, since the model is now conditionally linear and Gaussian.
+The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions whereas $\phi$ is sampled via a Metropolis Hastings accept/reject procedure.\footnote{The KSC method was a big advance over the one-state-at-a-time sampling method that was used prior to that time.
+There have also been other advances with particle MCMC since then that still exploit the insights gained from this method.}
 
@@
         \subsubsection*{Implementation}
-        The code used to replicate the KSC algorithm was written by \citet{chad2018} and amended for the purposes of this research.
+        The code used to replicate the KSC algorithm\footnote{The KSC algorithm, KSC MCMC and `off-set mixture MCMC' will be used interchangeably.
+The model itself will be referred to as `off-set mixture model'.} was written by \citet{chad2018} and amended for the purposes of this research.
 Additional code was developed to produce rank statistics, quantities of interest as well as to integrate the sampling method into the SBC framework.
@@
         HMC is a general approach to applying MCMC.
-The implementation of HMC is through the Stan programming language which allows for the specification other generative or Bayesian models, not just stochastic volatility.
-The implementation details of HMC is abstracted away from the user, who only needs to focus on specifying the model using the Stan syntax.
+The implementation of HMC is through the Stan programming language which allows for the specification of other generative or Bayesian models, not just stochastic volatility.
+The implementation details of HMC are abstracted away from the user, who only needs to focus on specifying the model using the Stan syntax.
 Any hyper-parameter tuning and initialisation is automated.
@@
 This describes the central location of the latent state distribution which is centered on the mean of log volatility $\mu$.
-A model can be re-parameterised in a way that might improve MCMC performance.
-Reparamaterisation applies a transformation to the model which may change the interpretation of the parameters or the way they are sampled but remains mathematically equivalent to the original model.
-The reparamaterisation available for a model is dependent on the MCMC approach.
-Some reparamaterisation are available only to specific sampling strategies and may improve or impair a MCMC performance.
+A model can be reparameterised in a way that might improve MCMC performance.
+Reparameterisation applies a transformation to the model which may change the interpretation of the parameters or the way they are sampled but remains mathematically equivalent to the original model.
+The reparameterisation available for a model is dependent on the MCMC approach.
+Some reparameterisation are available only to specific sampling strategies and may improve or impair a MCMC performance.
 
@@
         \end{table}
-        \subsubsection{Reparamaterised stochastic volatility in Stan}
-        The reparamaterisation for Hamiltonian Monte Carlo changes how the latent states are sampled.
+
+        \subsubsection{Reparameterised stochastic volatility in Stan}
+        The reparameterisation for Hamiltonian Monte Carlo changes how the latent states are sampled.
 Specifically, the states are first sampled from a standard normal distribution and are transformed to be samples from the state equation.
@@
         \item Sample from a standard normal distribution and multiply by the variance of the log volatility.
-The state vector is sampled with mean centered on zero and variance equal to $\sigma_{\eta}^2$
+The state vector is sampled with mean centered on zero and variance equal to $\sigma_{\eta}^2$ as given by
         \begin{align}
         h_{std} \sim&\space \,\mathrm{N}(0,1) \\
-        h =& \space \, h_{std} \times \sigma_{\eta}
-        \end{align}
-        \item Apply the appropriate re-scaling to get samples from log volatility. 
+        h =& \space \, h_{std} \times \sigma_{\eta}.
+        \end{align}
+        \item Apply the appropriate re-scaling to get samples from the log variance
         \begin{align}
         h_1 =& \space \,\frac{h_{std, 1}\times \sigma_{\eta}} {\sqrt{1 - \phi^2}} + \mu \\
-        h_{t+1} =& \space \,h_{std, t+1}\times \sigma_{\eta} + \mu  + \phi(h_{t} - \mu),\space \, t\neq 1
+        h_{t+1} =& \space \,h_{std, t+1}\times \sigma_{\eta} + \mu  + \phi(h_{t} - \mu),\space \, t\neq 1.
         \end{align}
         \end{enumerate}
-        This returns log volatility as desired.
+        This returns the log variance as desired.
 
@@
 
-        The following parameterisation is described as non centered in location. This follows the methodology outlined in \citet{strickland2008parameterisation}.
-
-        Starting with the log chi squared model:
-
-        $$
-        \begin{aligned}
+        The following parameterisation is described as non centered in location.
+This follows the methodology outlined in \citet{strickland2008parameterisation}.
+Starting with the log chi squared model
+        \begin{align}
         y^{\ast}_t =& \space \, h_t + z_t \\
-        h_{t+1} =& \space \, \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t
-        \end{aligned}
-        $$
-
-        Let $g$ be the demeaned state variable. Rewrite the demeaned state equation $g_{t+1}$ to be non centered in location by subtracting average volatility $\mu$.
-
-        $$
-        \begin{aligned}
-        g_t =& \space \, h_t - \mu \\
-        g_{t+1} =& \space \, \phi g_t + \sigma_{\eta}\eta_{t}
-        \end{aligned}
-        $$
-
-        Return average volatility into measurement equation and rewrite as a function of the non centered state equation:
-
-        $$
-        \begin{aligned}
+        h_{t+1} =& \space \, \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t.
+        \end{align}
+        Reparameterise the state variable $h_t$ by subtracting the average log variance $\mu$. Let $g_t = \space \, h_t - \mu$. This leads to the state equation
+        \begin{align}
+        g_{t+1} =& \space \, \phi g_t + \sigma_{\eta}\eta_{t}.
+        \end{align}
+        Return average volatility into measurement equation and rewrite as a function of the non centered state equation
+        \begin{align}
         y^{\ast}_t =& \space \, g_t + \mu + z_t \\
-        g_{t+1} =& \space \, \phi g_t + \sigma_{\eta}\eta_{t}
-        \end{aligned}
-        $$
-
-        The mean of the log volatility is now inside the measurement equation and de-meaned from the state equation.
-
-    \subsubsection{Prior Predictive check}
+        g_{t+1} =& \space \, \phi g_t + \sigma_{\eta}\eta_{t},
+        \end{align}
+        where the initial state now is drawn from 
+        \begin{align}
+            g_1 \sim& \space \,\mathrm{N}\left(0, \frac{\sigma_{\eta}^2}{1-\phi^2}\right).
+        \end{align}
+        The mean of the log volatility is now inside the measurement equation and demeaned from the state equation.
+
+    \subsubsection{Prior predictive check}
         % Different parameteristions express the same mathematical models in ways that make it easier or harder for a given algorithm to sample. 
@@
 This is to demonstrate that different parameterisations express the same underlying mathematical model.
-This is similar the first two steps of SBC, except instead of generating a data set, generate the state variable implied by the prior draw.
+This is similar to the first two steps of SBC, except instead of generating a data set, it generates the state variable implied by the prior draw.
 Let $\boldsymbol{\theta}$ be a vector of static parameters.
-Simulate a draw from the joint prior:
-
-        $$
-        \begin{aligned}
-        \boldsymbol{\theta}^{sim} \sim \pi(\boldsymbol{\theta})
-        \end{aligned}
-        $$
-
-        Generate a the second state parameter conditional on prior simulation and first state variable.
-
-        $$
-        \begin{aligned}
-        h_2^{sim} \sim \pi (h_2|h_1, \boldsymbol{\theta}^{sim})
-        \end{aligned}
-        $$
+Simulate a draw from the joint prior
+        \begin{align}
+        \boldsymbol{\theta}^{sim} \sim \pi(\boldsymbol{\theta}).
+        \end{align}
+
+        Generate the second state parameter conditional on prior simulation and first state variable
+        \begin{align}
+        h_2^{sim} \sim \pi (h_2|h_1, \boldsymbol{\theta}^{sim}).
+        \end{align}
 
@@
             \includegraphics[scale=0.1]{figures/ppc_h2.png}
-            \caption{Prior predictive simulation of second state variable from both centered and re-parameterised stochastic volatility model.
+            \caption{\textbf{Prior predictive simulation of second state variable from both centered and re-parameterised stochastic volatility model}.
 Both models are mathematically equivalent and generate samples from the same data generating process.}
@@
 The SV model in Section 2.1 is used for all simulations including the same priors.
-For a given MCMC and parameterisation pair, 5000 SBC iterations\footnote{Results for 1000 SBC iterations returned noisy estimates and can be found in Appendix D.} are run using simulated data sets of size T=1000.
+For a given MCMC and parameterisation pair, 5000 SBC iterations\footnote{Results for 1000 SBC iterations returned noisy estimates and can be found in Appendix D.} are run using simulated data sets of size $T=1000$.
 5000 rank statistics are obtained per univariate parameter from this process.
@@
 This is calculated to be 250 which is the ratio of  5000 SBC iterations and 20 bins.
-The $\chi^2$ statistic is used to summarise the deviations from uniformity for all state variables since it is unrealistic to visually inspect all the histograms.
+The Chi-squared statistic is used to summarise the deviations from uniformity for all state variables since it is unrealistic to visually inspect all the histograms.
 The SBC algorithm is summarised in Algorithm \ref{alg:sbc}.
@@
                 \State \text{Draw from joint prior: } $\boldsymbol{\theta}^{sim}_k \sim\pi (\boldsymbol{\theta})$
-                \State \text{Simulate data set with 1000 observations: } $\boldsymbol{y}^{sim}_k \sim p(\boldsymbol{y}|\boldsymbol{\theta}^{sim}_k)$
-                \State \text{Draw 999 posterior samples post burn in:} $\{\boldsymbol{\theta}_1,\dots , \boldsymbol{\theta}_{999}\}_k \sim p(\boldsymbol{\theta} | \boldsymbol{y}^{sim}_k)$
-                \State \text{Compute rank statistics:} $\boldsymbol{r} = rank(\{\boldsymbol{\theta}_1,\dots , \boldsymbol{\theta}_{999}\}_k, \boldsymbol{\theta}^{sim}_k)$
+                \State \text{Simulate data set with 1000 observations: } $\boldsymbol{y}^{sim}_k \sim \pi(\boldsymbol{y}|\boldsymbol{\theta}^{sim}_k)$
+                \State \text{Draw 999 posterior samples post burn in:} $\{\boldsymbol{\theta}_1,\dots , \boldsymbol{\theta}_{999}\}_k \sim \pi(\boldsymbol{\theta} | \boldsymbol{y}^{sim}_k)$
+                \State \text{Compute rank statistics:} $\boldsymbol{r} = \mathrm{rank}(\{\boldsymbol{\theta}_1,\dots , \boldsymbol{\theta}_{999}\}_k, \boldsymbol{\theta}^{sim}_k)$
               \EndFor
@@
 The distributions of $\sigma^2$ and $\phi$ now exhibit a more uniform shape, implying greater calibration and correct posterior estimates returned on average.
-These results suggests that HMC struggles to sample some parameters from the centered parameterisation of the stochastic volatility model.
+These results suggest that HMC struggles to sample some parameters from the centered parameterisation of the stochastic volatility model.
 The re-parameterised model possess stronger evidence of calibration for the parameters examined, producing more reliable and consistent inference.
@@
         \includegraphics[scale=0.09]{results/hmc_cp_5k.png}
-        \caption{\textbf{5000 SBC iterations for centered model using Hamiltonian Monte Carlo}.
+        \caption{\textbf{Distribution of rank statistics with 5000 SBC iterations for centered model using Hamiltonian Monte Carlo}.
 $\mu$ and the state parameters look approximately uniform.
@@
         \includegraphics[scale=0.09]{results/hmc_ncp_5k.png}
-        \caption{\textbf{5000 SBC iterations for re-parameterised model using Hamiltonian Monte Carlo}.
+        \caption{\textbf{Distribution of rank statistics with 5000 SBC iterations for re-parameterised model using Hamiltonian Monte Carlo}.
 The rank statistic distributions for $\sigma^2$ and $\phi$ appear to follow a uniform distribution.
@@
         \end{tabular}
-        \caption{ESS for centered and re-parameterised SV model across 5000 SBC iterations for HMC.
-The reparameterised model has higher WAA counts for each summary statistic relative to the centered parameterisation.}
+        \caption{\textbf{ESS for centered and re-parameterised SV model across 5000 SBC iterations for HMC}.
+The reparameterised model has higher ESS counts for each summary statistic relative to the centered parameterisation.}
         \label{tab:hmcess}
@@
 
-    \subsection{Mixture off-set MCMC}
-    The Gaussian off-set mixture approximation is run with the first 10,000 samples discarded as burn-in (following the method of \citet{kim1998stochastic}).
+    \subsection{Off-set Mixture MCMC}
+    The off-set mixture MCMC is run with the first 10,000 samples discarded as burn-in (following the method of \citet{kim1998stochastic}).
 However, unlike KSC who take 750,000 posterior draws from the model, the number of post burn-in posterior samples taken as part of the SBC procedure is 9,999.
-Taking 750,000 draws from the Gaussian off-set mixture would lead to issues due to disk space limitations and computational resources required to run SBC.
+Taking 750,000 draws from the off-set mixture model would lead to issues due to disk space limitations and computational resources required to run SBC.
 Furthermore, the number of posterior samples is larger than the 999 taken for the SBC experiment run for Hamiltonian Monte Carlo.
-Earlier experiments attempted 999 post-burn in draws for the off-set mixture model, however there are concerns that such a short chain would not have converged onto the target posterior distribution.
+Earlier experiments attempted 999 post burn-in draws for the off-set mixture model, however there are concerns that such a short chain would not have converged onto the target posterior distribution.
 Therefore 9,999 draws was chosen as a compromise under both constraints.
@@
         \includegraphics[scale=0.09]{results/ksc_cp_5k.png}
-        \caption{\textbf{5000 SBC iterations for centered Gaussian mixture approximation model}.
+        \caption{\textbf{Distribution of rank statistics with 5000 SBC iterations for centered Gaussian mixture approximation model}.
 The rank statistics for all the selected parameters display deviations from uniformity.
@@
         \includegraphics[scale=0.09]{results/ksc_ncp_5k.png}
-        \caption{\textbf{5000 SBC iterations for non centered Gaussian mixture approximation model}.
+        \caption{\textbf{Distribution of rank statistics with 5000 SBC iterations for non centered Gaussian mixture off-set model}.
 The rank statistics for all selected parameters are non uniform in shape.
@@
 Overall, the off-set mixture MCMC is inefficient at sampling the static parameters.
+Note that we cannot directly compare the effect of the reparameterisation on the ESS between Figure \ref{fig:kscess} and Figure \ref{fig:hmcess} due to the difference in post burn-in posterior draws.
 
@@
         \includegraphics[scale=0.1]{results/ksc_ess.png}
-        \caption{Effective sample sizes for static parameters after 5000 iterations of SBC and 1000 post burn-in samples for the Gaussian Mixture model.
+        \caption{\textbf{Effective sample sizes for static parameters after 5000 iterations of SBC and 1000 post burn-in samples for the Gaussian Mixture model}.
 The KSC sampling strategy struggles to generate independent samples for all static parameters, suggesting a high degree of auto-correlation in the Markov Chain.}
@@
         \end{tabular}
-        \caption{KSC: ESS for centered and re-parameterised stochastic volatility model.}
+        \caption{\textbf{ESS for centered and reparameterised SV model across 5000 SBC iterations for KSC}.}
         \label{tab:kscess}
@@
 Additionally, it could be due to the approximation of the actual stochastic volatility model not producing accurate posterior estimates.
-Correcting potential approximation error of the model is explored in the next section.
+Correcting approximation error of the model is explored in the next section.
 
@@
 The re-weighting procedure ensures that samples are drawn from the correct posterior density.
-This is applied to the calculation of the expected value of the stochastic volatility posterior density using samples drawn from the approximating off-set mixture model.
+This is applied to the calculation of the expected value of the stochastic volatility posterior density using samples drawn from the off-set mixture model.
 
@@
     \begin{align}
-    S(\theta) &= 1[f(\theta_b) < f(\theta^{sim})], \\
+    S(\theta) &= 1[f(\theta_b) < f(\theta^{sim})] \\
     E[S(\theta) | y] &= \int S(\theta) c(\theta | y) d\theta\\ 
@@
         \includegraphics[scale=0.1]{results/weighted_ksc_cp_5k.png}
-        \caption{\textbf{5000 SBC iterations for re-weighted rank statistics from the Gaussian off-set mixture model}. The shape of the histogram is consistent with the unweighted rank statistics.}
+        \caption{\textbf{Distribution of rank statistics with 5000 SBC iterations for re-weighted rank statistics from the Gaussian off-set mixture model}.
+The shape of the histogram is consistent with the unweighted rank statistics.}
         \label{fig:reweight5k}
@@
     It is not practical to inspect the histograms of all parameters and states in the stochastic volatility model.
-Instead, the chi squared statistic is used to summarise the shape of the rank statistic distribution.
+Instead, the Chi-squared statistic is used to summarise the shape of the rank statistic distribution.
 This will be used to summarise the state variables only since the static parameters can be compared across simulations individually.
 
-    The latent state variable chi squared statistics for centered and re-parameterised HMC, centered KSC and centered importance weighted KSC are reported as histograms in Figure \ref{fig:allchisq} using results from 5000 SBC iterations.
+    The Chi-squared statistics for centered and re-parameterised HMC, centered KSC and centered importance weighted KSC are reported as histograms in Figure \ref{fig:allchisq} using results from 5000 SBC iterations.
 Re-parameterised KSC is omitted as its poor SBC performance skews the scale of the plot which makes it difficult to analyse results.
-A chi squared statistic of 0 implies the sample is exactly uniformly distributed.
-Any value away from 0 captures variation or deviations from uniformity.
-This provides a high level summary of the state variables - how close their rank statistics are to being uniformly distributed and their distance relative to the other algorithms and parameterisations.
+A chi squared statistic of 0 means the rank statistics for a given parameter is exactly uniformly distributed and any value away from 0 captures deviations from uniformity.
+The chi squared statistics for the latent state variable ranks are expected to be some non-zero value since they are calculated from finite samples and subject to sampling variation even if the draws come from a uniform distribution.
+The purpose of this plot is to provide a high level summary of the latent state variable calibration for each simulation - how close their rank statistics are to being uniformly distributed and their distance relative to the other algorithms and parameterisations.
 
     The HMC results are relatively close to zero.
-Both KSC and the importance weighted ranks on the other hand are far from zero suggesting major deviations from uniformity.
-There is a large gap between the HMC and KSC chi squared statistics implying the distribution of state variable ranks are much less uniform for KSC relative to the HMC.
-Combined with the results in the previous section, there is a strong evidence that HMC outperforms KSC where it comes to calibration across all parameters and latent state variables.
+Both KSC and the importance weighted ranks on the other hand are far from zero suggesting larger deviations from uniformity.
+There is a wide gap between the HMC and KSC chi squared statistics implying the distribution of latent state ranks are much more  uniform for HMC relative to KSC.
+Combined with the results in the previous section, there is a strong evidence that the HMC algorithm outperforms the KSC algorithm where it comes to calibration across all parameters and latent state variables.
 
@@
         \includegraphics[scale=0.1]{results/dist_chisq_all.png}
-        \caption{\textbf{Distribution of chi squared statistics for latent state variables}.
-Results from 5000 SBC iterations are used.
+        \caption{\textbf{Distribution of chi squared statistics for latent state variables from 5000 SBC iterations}.
 Values closer to 0 are more consistent with a discrete uniform distribution.
 Both HMC simulations are much closer to zero than KSC.
-This suggests that the HMC algorithm overall produces more calibrated posterior estimates for the state variables.}
+This suggests that the HMC algorithm overall produces more calibrated posterior estimates for the latent state variables.}
         \label{fig:allchisq}
@@
 Centered parameterisation using HMC performed moderately well but struggled to produce calibrated posterior estimates for $\sigma^2$ and $\phi$ as well as generate independent samples for these parameters.
-Gaussian mixture approximation struggled to return calibrated posterior estimates and efficient samples for both parameterisations.
+The Gaussian off-set mixture approximation struggled to return calibrated posterior estimates and efficient samples for both parameterisations.
 However, the centered parameterisation results for the Gaussian mixture model may improve if we increased the length of the Markov chain.
@@
 The priors chosen for this study came directly from \citet{kim1998stochastic}.
-The SBC results for the Gaussian off-set mixture may improve if the priors were more informative - although this is only conjecture.
+The SBC results for the Gaussian off-set mixture model may improve if the priors were more informative - although this is only conjecture.
 Whether or not to use tighter priors depends on the modelling context and the problem at hand \citep{gelman2017prior}.
@@
 The failure of non centered parameterisation in location for the off-set mixture model could be due to a variety of reasons.
-This may be due to a particular software implementation when it comes to the Kalman Filter and Simulation Smoother (where a specific software package was applied) that was not designed to handle the non centered parameterisations.
+This may be due to a particular software implementation in the Kalman Filter and Simulation Smoother (where a specific software package was applied) that was not designed to handle the non-centered parameterisation.
 Or perhaps this particular parameterisation does not suit the sampling strategy applied by KSC and another approach (e.g.
@@
 Additionally, this simulation study does not provide any advice on the appropriateness of the stochastic volatility model on real data.
-SBC implicitly assumes the data generating process is coming from the `correct model' and does not tell us which MCMC algorithm is better or more calibrated when the model is misspecified.
+SBC implicitly assumes the data generating process is coming from the `correct model' and does not say which MCMC algorithm is more calibrated when the model is misspecified.
 Whether the stochastic volatility model is appropriate for modelling a particular financial time series requires further study into the data generating process, domain expertise and a suite of other diagnostic checks.
-Examples of this are posterior predictive checking and out of sample predictive performance.
+Examples of this are posterior predictive checks and out of sample predictive performance.
 Rather, the SBC approach gives insight into how well calibrated a sampling strategy is conditional on a known data generating process which is important in all applied contexts.
@@
 As discussed, inspecting many histograms for highly parameterised models is unrealistic.
-Other summary statistics and visualisations can be applied for a high level comparison, such as the chi squared statistics applied in this research.
-Furthermore, it is not always straightforward to understand why any any one parameter may produce poor calibration results, for example, an arbitrary latent state variable.
+Other summary statistics and visualisations can be applied for a high level comparison, such as the Chi-squared statistics applied in this research or formal hypothesis tests of uniformity such as the Chi-squared test.
+Furthermore, it is not always straightforward to understand why any one parameter may produce poor calibration results, for example, an arbitrary latent state variable.
 SBC may tell us some information about the miscalibration based on the histogram shape (e.g.
@@
 Lastly, it is unclear what constitutes a fair comparison between algorithms when comparing SBC results.
-In particular, the number of post burn-in or warm-up samples differed between algorithms (999 post burn-in for HMC, 9999 for the bespoke KSC algorithm).
-An argument could be made that the poor results for the bespoke KSC algorithm may be due to the chain not converging onto the target distribution.
+In particular, the number of post burn-in or warm-up samples differed between algorithms (999 post burn-in for HMC, 9999 for the KSC algorithm).
+An argument could be made that the poor results for the KSC algorithm may be due to the chain not converging onto the target distribution.
 Indeed, the authors of this approach extended their Markov chain to 750,000 posterior draws (discarding the first 10,000 draws as burn-in).
@@
 This research evaluated and compared different MCMC algorithms for fitting stochastic volatility models.
-As our models and algorithms for estimating these models become more complicated, so does the need for principled ways of checking that they are returning correct posterior estimates.
+Principled ways of checking that our algorithms are returning correct posterior estimates are important, especially as our models and algorithms become more complicated.
 SBC provides a general simulation design to check whether a modelling strategy is producing calibrated results for generative models.
 
-In the context of stochastic volatility, applying Hamiltonian Monte Carlo on a re-parameterised stochastic volatility model provided the most calibrated and efficient estimates when compared with KSC's bespoke MCMC method.
-This also outperformed the KSC model using importance sampling weights to correct for approximation error.
+In the context of stochastic volatility, applying Hamiltonian Monte Carlo on a re-parameterised stochastic volatility model provided the most calibrated and efficient estimates when compared with KSC's algorithm applied to the approximate off-set mixture model.
+This also outperformed the KSC model using importance sampling weights to correct for approximation error as well as the model with non-centered parameterisation.
 Other parameterisations of the stochastic volatility model can be considered to see if it improves sampling performance based on these diagnostics.
@@
 \section{Appendix B: Sampling from Gaussian off-set mixture model}
-    The steps in the MCMC process is described in Algorithm \ref{alg:ksc}.
+    The steps in the MCMC process are described in Algorithm \ref{alg:ksc}.
 The initial values are set to the values outlined in \citet{kim1998stochastic}, except for the mixing indicator which is unspecified.
@@
         \includegraphics[scale=0.09]{results/hmc_cp_1k.png}
-        \caption{1000 SBC iterations for centered model using Hamiltonian Monte Carlo.
-Histograms show noisy estimates around uniform distribution.
-In particular, $\sigma^2$ displays peaks on both ends of the histogram and $\phi$ shows a major on the left hand side.
-This suggests that the posterior samples of $\sigma^2$ are under-dispersed relative to the prior distribution and posterior samples of $\phi$ are over estimating the true parameter.}
+        \caption{\textbf{1000 SBC iterations for centered model using Hamiltonian Monte Carlo}.
+$\sigma^2$ displays peaks on both ends of the histogram and $\phi$ shows a major on the left hand side.
+This suggests that the average posterior samples of $\sigma^2$ are under-dispersed relative to the prior distribution and posterior samples of $\phi$ are over estimating the true parameter.}
         \label{fig:cphmc1k}
@@
         \includegraphics[scale=0.09]{results/hmc_ncp_1k.png}
-        \caption{1000 SBC iterations for non centered model using Hamiltonian Monte Carlo.
+        \caption{\textbf{1000 SBC iterations for non centered model using Hamiltonian Monte Carlo}.
 The rank statistic distributions for $\sigma^2$ and $\phi$ have improved.
@@
         \centering
-        \includegraphics[scale=0.1]{results/ksc_cp_1k.png}
+        \includegraphics[scale=0.09]{results/ksc_cp_1k.png}
         \caption{\textbf{1000 SBC iterations for centered Gaussian off-set mixture model}.
-The rank statistics for all the selected parameters display a lack of uniformity.
-$\mu$ has a large left peak suggesting the algorithm is over estimating the true parameter.
-The converse can be said about $\phi$ which possesses a peak on the right hand side which implies underestimation of the true parameter.
-1st and 500th latent state variables also appear to be over estimating the true value.}
+$\mu$ has a large left peak suggesting the algorithm is over estimating the true parameter on average.
+The converse can be said about $\phi$ which possesses a peak on the right hand side which implies underestimation of the true parameter.}
         \label{fig:cpksc1k}
@@
         \centering
-        \includegraphics[scale=0.1]{results/ksc_ncp_1k.png}
+        \includegraphics[scale=0.09]{results/ksc_ncp_1k.png}
         \caption{\textbf{1000 SBC iterations for non centered in location Gaussian off-set mixture model}.
@@
         \centering
+        \includegraphics[scale=0.09]{results/weighted_ksc_cp_1k.png}
+        \caption{\textbf{1000 SBC iterations for re-weighted rank statistics from the Gaussian mixture approximation model}.
+The shape of the histogram is consistent with the unweighted rank statistics.}
+        \label{fig:reweight1k}
+    \end{figure}
+
+\begin{figure}[H]
+    \centering
     \includegraphics[scale=0.09]{results/weighted_ksc_ncp_1k.png}
-    \caption{1000 SBC iterations for re-weighted rank statistics from the re-parameterised Gaussian mixture approximation model.
+    \caption{\textbf{1000 SBC iterations for re-weighted rank statistics from the re-parameterised Gaussian mixture approximation model}.
 The shape of the histogram is consistent with the unweighted rank statistics.}
@@
     \centering
-        \includegraphics[scale=0.1]{results/weighted_ksc_cp_1k.png}
-        \caption{1000 SBC iterations for re-weighted rank statistics from the Gaussian mixture approximation model. The shape of the histogram is consistent with the unweighted rank statistics.}
-        \label{fig:reweight1k}
-    \end{figure}
-
-\begin{figure}[H]
-    \centering
     \includegraphics[scale=0.09]{results/weighted_ksc_ncp_5k.png}
-    \caption{5000 SBC iterations for re-weighted rank statistics from the re-parameterised Gaussian mixture approximation model.
-No major improvements are observed to the uniformity of rank statistics.}
+    \caption{\textbf{5000 SBC iterations for re-weighted rank statistics from the re-parameterised Gaussian mixture approximation model.}}
     \label{fig:ncpreweight5k}

### 2023-10-19 manuscript/main.tex
Syntax edits

@@
 Additionally, MCMC algorithms may perform better or worse based on calibration and efficiency depending on the parameterisation of the model.
-HMC applied on a re-parameterised SV model produces the most calibrated and efficient posterior estimates across all SBC experiments.
+HMC applied on a reparameterised SV model produces the most calibrated and efficient posterior estimates across all SBC experiments.
 
@@
 $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the auto-regressive parameter $\phi$ such that $|\phi|<1$.
-This auto-regressive parameter represents the persistence or stickiness of log volatility and the dispersion parameter $\sigma_{\eta}$ is the constant variance of the states.
+This auto-regressive parameter represents the persistence or stickiness of the log variance and the dispersion parameter $\sigma_{\eta}$ is the constant variance of the states.
 $\epsilon_t$ and $\eta_t$ are standard normal white noise shocks and are uncorrelated with each other.
@@
 \subsection{Research Goal}
-    The objective of this research is to design a simulation study to evaluate the calibration of the HMC and KSC offset-mixture algorithms used to estimate SV models.
+    The objective of this research is to design a simulation study to evaluate the calibration of the HMC and KSC off-set mixture algorithms used to estimate SV models.
 As discussed, there are limitations to evaluating MCMC algorithms based on fits to real data and single simulations.
@@
 
-            \item Chi squared statistics calculated over the $K$ rank statistics to summarise the shape of the distribution and deviation from uniformity.
+            \item Chi-squared statistics calculated over the $K$ rank statistics to summarise the shape of the distribution and deviation from uniformity.
 
@@
 
-            \subsubsection{Chi squared statistics}
+            \subsubsection{Chi-squared statistics}
             One difficulty of relying on the visualisation of rank statistic distributions for SV models is there are a very large number of unknowns in the model to consider.
-An alternative to visually checking for uniformity of a given parameter is to calculate a Chi-squared statistic based on the counts in each histogram bin.
+An alternative to visually checking for uniformity of a given parameter is to calculate a chi-squared statistic based on the counts in each histogram bin.
 Suppose the histogram has $J$ bins, where for each $j^{th}$ bin, $b_j$ is the number of counts and $e_j$ the expected count in bin $j$.
 The expected count is the ratio of $K$ rank statistics and $J$ bins, $e_j=\frac{K}{J}$.
-The corresponding Chi-squared statistic is denoted by
+The corresponding chi-squared statistic is denoted by
             \begin{align}
@@
 
-            A discrete empirical distribution with the same number of realisations in each of $J$ bins will return a chi squared statistic of zero, since the number of realisations in each bin is equal to the expected count under the relevant  uniform distribution.
-The distribution of chi squared statistics is visualised to compare results across multiple simulations.
+            A discrete empirical distribution with the same number of realisations in each of $J$ bins will return a chi-squared statistic of zero, since the number of realisations in each bin is equal to the expected count under the relevant  uniform distribution.
+The distribution of chi-squared statistics is visualised to compare results across multiple simulations.
 Thus, the distribution of chi-squared statistics produced from the histograms of SBC rank statistics for a large number of univariate model parameters will give a high level summary of the overall model calibration, and hence for the overall performance of the algorithm.
-By comparing the distributions of Chi-squared statistics produced by competing algorithms, one algorithm's degree of calibration, relative to that of the other algorithm, may be assessed.
+By comparing the distributions of chi-squared statistics produced by competing algorithms, one algorithm's degree of calibration, relative to that of the other algorithm, may be assessed.
 
@@
 This section defines and compares the two MCMC algorithms used to estimate the SV model.
-Section 4.1 and Section 4.2 describes Stan's No-U-Turn sampler and KSC's mixture offset MCMC respectively.
+Section 4.1 and Section 4.2 describes Stan's No-U-Turn sampler and KSC's off-set mixture MCMC respectively.
 Section 4.3 compares the similarities and differences between both methods.
@@
 Koopman who co-authored \citet{durbin2012time} also co-authored the software used by \citet{kim1998stochastic} which can be found at \citet{koopman1996ssfpack}.} Note that for the simulation smoother to produce an appropriate posterior sample draw from the latent state vector, the state and measurement equations are required to be linear and conditionally Gaussian.
-Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring, adding an offset, and taking the log of $y_t$, resulting in
-        \begin{align}
-        y_t^{*} &= log(y_t^2 + c) = h_t + z_t, \label{eqn:transform}
-        \end{align}
-        where $z_t = log(\epsilon_t^2)$ has a distribution that is like that of the logarithm of a chi-squared random variable.
+Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring, adding an off-set, and taking the log of $y_t$, resulting in
+        \begin{align}
+        y_t^{*} &= \log(y_t^2 + c) = h_t + z_t, \label{eqn:transform}
+        \end{align}
+        where $z_t = \log(\epsilon_t^2)$ has a distribution that is like that of the logarithm of a chi-squared random variable.
 It is known that the error $z_t$ has mean -1.2704 and variance 4.93.
-The offset $c$ is set to $c=0.001$ to ensure the model is robust to small $y_t^2$ values \citep{fuller1996introduction}.
+The off-set $c$ is set to $c=0.001$ to ensure the model is robust to small $y_t^2$ values \citep{fuller1996introduction}.
 By transforming the entire measurement equation, as in (\ref{eqn:transform}), the relationship between $y_t^{*}$ and $h_t$ is now linear; however, the error is not Gaussian.
-Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of seven Gaussian distributions to approximate the distribution of $z_t$, with the mixture selected to match the first 4 moments of the log chi squared distribution.
+Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of seven Gaussian distributions to approximate the distribution of $z_t$, with the mixture selected to match the first 4 moments of the log chi-squared distribution.
 The mixture distribution is defined by
@@
 
-        \item Sample from a standard normal distribution and multiply by the variance of the log volatility.
+        \item Sample from a standard normal distribution and multiply by the variance of the states.
 The state vector is sampled with mean centered on zero and variance equal to $\sigma_{\eta}^2$ as given by
@@
         \subsubsection{Non-centered Gaussian mixture off-set model}
-        The re-parameterised model for the Gaussian mixture off-set is expressed differently due to the Kalman Filter.
-Re-parameterised HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates.
+        The reparameterised model for the Gaussian mixture off-set is expressed differently due to the Kalman Filter.
+Reparameterised HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates.
 Applying the Kalman Filter requires the state equation to be a function of the state variable with the average log volatility parameter $\mu$ entering the measurement equation.
@@
 This follows the methodology outlined in \citet{strickland2008parameterisation}.
-Starting with the log chi squared model
+Starting with the log chi-squared model
         \begin{align}
@@
         % Different parameteristions express the same mathematical models in ways that make it easier or harder for a given algorithm to sample. 
-        A prior predictive simulation of the second state parameter is performed for both a centered and re-parameterised model using the approach described in Section 4.4.1.
+        A prior predictive simulation of the second state parameter is performed for both a centered and reparameterised model using the approach described in Section 4.4.1.
 This is to demonstrate that different parameterisations express the same underlying mathematical model.
@@
             \includegraphics[scale=0.1]{figures/ppc_h2.png}
-            \caption{\textbf{Prior predictive simulation of second state variable from both centered and re-parameterised stochastic volatility model}.
+            \caption{\textbf{Prior predictive simulation of second state variable from both centered and reparameterised stochastic volatility model}.
 Both models are mathematically equivalent and generate samples from the same data generating process.}
@@
 This is calculated to be 250 which is the ratio of  5000 SBC iterations and 20 bins.
-The Chi-squared statistic is used to summarise the deviations from uniformity for all state variables since it is unrealistic to visually inspect all the histograms.
+The chi-squared statistic is used to summarise the deviations from uniformity for all state variables since it is unrealistic to visually inspect all the histograms.
 The SBC algorithm is summarised in Algorithm \ref{alg:sbc}.
@@
     Figure \ref{fig:ncphmc5k} show the SBC results for the reparameterised model using HMC.
-The re-parameterised model produces approximately uniform rank statistics across all reported parameters.
+The reparameterised model produces approximately uniform rank statistics across all reported parameters.
 The distributions of $\sigma^2$ and $\phi$ now exhibit a more uniform shape, implying greater calibration and correct posterior estimates returned on average.
 These results suggest that HMC struggles to sample some parameters from the centered parameterisation of the stochastic volatility model.
-The re-parameterised model possess stronger evidence of calibration for the parameters examined, producing more reliable and consistent inference.
+The reparameterised model possess stronger evidence of calibration for the parameters examined, producing more reliable and consistent inference.
         \begin{figure}[H]
@@
         \includegraphics[scale=0.09]{results/hmc_ncp_5k.png}
-        \caption{\textbf{Distribution of rank statistics with 5000 SBC iterations for re-parameterised model using Hamiltonian Monte Carlo}.
+        \caption{\textbf{Distribution of rank statistics with 5000 SBC iterations for reparameterised model using Hamiltonian Monte Carlo}.
 The rank statistic distributions for $\sigma^2$ and $\phi$ appear to follow a uniform distribution.
@@
 
-    This computational difficulty in the centered parameterisation is reflected in the effective sample sizes.
-Figure \ref{fig:hmcess} and Table \ref{tab:hmcess} summarises the distribution of ESS for the static parameters after 5000 SBC iterations.
+    Computational difficulty in the centered parameterisation is reflected in the effective sample sizes.
+Figure \ref{fig:hmcess} and Table \ref{tab:hmcess} summarise the ESS distributions for the static parameters in both parameterisations after 5000 SBC iterations.
 The reparameterised model generates independent samples much more efficiently than the centered model, reflected in the larger ESS values across every summary statistic.
-This is consistent with HMC struggling to sample overall from the posteriors of $\sigma^2$ and $\phi$ in the centered model.
+The median ESS for $\mu$, $\phi$ and $\sigma^2$ in the centered parameterisation is much smaller as a proportion of the posterior draws.
+This is consistent with HMC struggling to sample from the posteriors of the static parameters overall in the centered model.
 
@@
         \end{tabular}
-        \caption{\textbf{ESS for centered and re-parameterised SV model across 5000 SBC iterations for HMC}.
+        \caption{\textbf{ESS for centered and reparameterised SV model across 5000 SBC iterations for HMC}.
 The reparameterised model has higher ESS counts for each summary statistic relative to the centered parameterisation.}
@@
 The numerical results are summarised in Figure \ref{fig:kscess} and Table \ref{tab:kscess}.
-The estimate for $\mu$ in the re-parameterised model exhibits a long right tail (longer than the 9,999 post burn-in posterior draws suggesting the presence of negatively auto-correlated draws).
-There is some improvement in efficiency from the re-parameterised model with higher ESS values across most of the static parameter quantiles.
+The estimate for $\mu$ in the reparameterised model exhibits a long right tail, with values larger than the 9,999 post burn-in posterior draws suggesting the presence of negatively auto-correlated draws.
+The median ESS for each static parameter relative to the number of posterior draws is very small.
+There is some improvement from the reparameterised model with higher ESS values across some of the static parameter quantiles, but these improvements are negligible.
 Overall, the off-set mixture MCMC is inefficient at sampling the static parameters.
-Note that we cannot directly compare the effect of the reparameterisation on the ESS between Figure \ref{fig:kscess} and Figure \ref{fig:hmcess} due to the difference in post burn-in posterior draws.
+Note, Figure \ref{fig:hmcess} and Figure \ref{fig:kscess} are not directly compared due to scale difference from varying post burn-in posterior draws.
+
+    % that we cannot directly compare the effect of the reparameterisation on the ESS between Figure \ref{fig:kscess} and Figure \ref{fig:hmcess} due to the difference in post burn-in posterior draws.
 
@@
     \subsubsection{Re-weighting rank statistics of the Gaussian off-set mixture model}
-    Let $v(\theta, h)$ be the log weights defined as the log difference between the posterior densities of the true model with log chi squared errors $log\: c(\theta, h | y^{\ast})$ and the mixture off-set model $log\:  d(\theta, h | y^{\ast})$.
+    Let $v(\theta, h)$ be the log weights defined as the log difference between the posterior densities of the true model with log chi-squared errors $\log c(\theta, h | y^{\ast})$ and the mixture off-set model $\log d(\theta, h | y^{\ast})$.
 The is given in the following expression
     \begin{align}
-        v(\theta, h) = log\: c(\theta, h | y^{\ast}) - log\:  d(\theta, h | y^{\ast}) = \text{const} + log\: c(y|h) - log\: d(y^{\ast} | h).
+        v(\theta, h) = \log c(\theta, h | y^{\ast}) - \log  d(\theta, h | y^{\ast}) = \text{const} + \log c(y|h) - \log d(y^{\ast} | h).
     \end{align}
@@
 Overall re-weighting the posterior samples from the approximate model does not improve the calibration results.
-The re-weighted rank statistics for the re-parameterised model also did not improve and can be found in Appendix E.
+The re-weighted rank statistics for the reparameterised model also did not improve and can be found in Appendix E.
 
@@
     It is not practical to inspect the histograms of all parameters and states in the stochastic volatility model.
-Instead, the Chi-squared statistic is used to summarise the shape of the rank statistic distribution.
+Instead, the chi-squared statistic is used to summarise the shape of the rank statistic distribution.
 This will be used to summarise the state variables only since the static parameters can be compared across simulations individually.
 
-    The Chi-squared statistics for centered and re-parameterised HMC, centered KSC and centered importance weighted KSC are reported as histograms in Figure \ref{fig:allchisq} using results from 5000 SBC iterations.
-Re-parameterised KSC is omitted as its poor SBC performance skews the scale of the plot which makes it difficult to analyse results.
-A chi squared statistic of 0 means the rank statistics for a given parameter is exactly uniformly distributed and any value away from 0 captures deviations from uniformity.
-The chi squared statistics for the latent state variable ranks are expected to be some non-zero value since they are calculated from finite samples and subject to sampling variation even if the draws come from a uniform distribution.
+    The chi-squared statistics for centered and reparameterised HMC, centered KSC and centered importance weighted KSC are reported as histograms in Figure \ref{fig:allchisq} using results from 5000 SBC iterations.
+Reparameterised KSC is omitted as its poor SBC performance skews the scale of the plot which makes it difficult to analyse results.
+A chi-squared statistic of 0 means the rank statistics for a given parameter is exactly uniformly distributed and any value away from 0 captures deviations from uniformity.
+The chi-squared statistics for the latent state variable ranks are expected to be some non-zero value since they are calculated from finite samples and subject to sampling variation even if the draws come from a uniform distribution.
 The purpose of this plot is to provide a high level summary of the latent state variable calibration for each simulation - how close their rank statistics are to being uniformly distributed and their distance relative to the other algorithms and parameterisations.
@@
 Both KSC and the importance weighted ranks on the other hand are far from zero suggesting larger deviations from uniformity.
-There is a wide gap between the HMC and KSC chi squared statistics implying the distribution of latent state ranks are much more  uniform for HMC relative to KSC.
+There is a wide gap between the HMC and KSC chi-squared statistics implying the distribution of latent state ranks are much more  uniform for HMC relative to KSC.
 Combined with the results in the previous section, there is a strong evidence that the HMC algorithm outperforms the KSC algorithm where it comes to calibration across all parameters and latent state variables.
@@
         \includegraphics[scale=0.1]{results/dist_chisq_all.png}
-        \caption{\textbf{Distribution of chi squared statistics for latent state variables from 5000 SBC iterations}.
+        \caption{\textbf{Distribution of chi-squared statistics for latent state variables from 5000 SBC iterations}.
 Values closer to 0 are more consistent with a discrete uniform distribution.
@@
 
-Hamiltonian Monte Carlo applied on the re-parameterised model performs best based on calibration and efficiency results.
+Hamiltonian Monte Carlo applied on the reparameterised model performs best based on calibration and efficiency results.
 Centered parameterisation using HMC performed moderately well but struggled to produce calibrated posterior estimates for $\sigma^2$ and $\phi$ as well as generate independent samples for these parameters.
@@
 As discussed, inspecting many histograms for highly parameterised models is unrealistic.
-Other summary statistics and visualisations can be applied for a high level comparison, such as the Chi-squared statistics applied in this research or formal hypothesis tests of uniformity such as the Chi-squared test.
+Other summary statistics and visualisations can be applied for a high level comparison, such as the chi-squared statistics applied in this research or formal hypothesis tests of uniformity such as the chi-squared test.
 Furthermore, it is not always straightforward to understand why any one parameter may produce poor calibration results, for example, an arbitrary latent state variable.
@@
 A fairer comparison may be to have roughly similar number of ESS for most of the estimated parameters, although this is difficult to get right with a large number of parameters to estimate.
-Given how well Hamiltonian Monte Carlo performed with the re-parameterised model, any improvement to the Gaussian off-set mixture model by increasing the length of the Markov chain would make calibration just as good, but not better due to the inefficiency of the sampler.
+Given how well Hamiltonian Monte Carlo performed with the reparameterised model, any improvement to the Gaussian off-set mixture model by increasing the length of the Markov chain would make calibration just as good, but not better due to the inefficiency of the sampler.
 
@@
 
-In the context of stochastic volatility, applying Hamiltonian Monte Carlo on a re-parameterised stochastic volatility model provided the most calibrated and efficient estimates when compared with KSC's algorithm applied to the approximate off-set mixture model.
+In the context of stochastic volatility, applying Hamiltonian Monte Carlo on a reparameterised stochastic volatility model provided the most calibrated and efficient estimates when compared with KSC's algorithm applied to the approximate off-set mixture model.
 This also outperformed the KSC model using importance sampling weights to correct for approximation error as well as the model with non-centered parameterisation.
@@
 
-2) Accept proposal as $\phi^{(i)}$ with probability $e^{\{g(\phi^\ast) - g(\phi^{(i-1)}\}}$ such that $g(\phi) = log (\pi (\phi)) - \frac {(h_t - \mu)^2 (1-\phi^2)}{2 \sigma_{\eta}^2} + \frac{1}{2} log (1-\phi^2)$
+2) Accept proposal as $\phi^{(i)}$ with probability $e^{\{g(\phi^\ast) - g(\phi^{(i-1)}\}}$ such that $g(\phi) = \log (\pi (\phi)) - \frac {(h_t - \mu)^2 (1-\phi^2)}{2 \sigma_{\eta}^2} + \frac{1}{2} \log (1-\phi^2)$
 
@@
 \section{Appendix C: Hamiltonian Monte Carlo algorithm description}
-Let $\theta$ be a target parameter and $\alpha$ be the auxiliary momentum variable. The Hamiltonian is defined as:
-
-$$
-\begin{aligned}
-H(\theta, \alpha) \equiv - log \: \pi(\theta, \alpha)
-\end{aligned}
-$$
-
-This can be broken down into kinetic energy $K(\theta, \alpha)$, the density over the auxiliary momentum, and potential energy $V(\theta)$, the density of the target posterior distribution.
-
-$$
-\begin{aligned}
-H(\theta, \alpha) &= - log \: \pi(\alpha | \theta) - log \: \pi(\theta) \\ 
-&\equiv  K(\theta, \alpha) + V(\theta)
-\end{aligned}
-$$
+Let $\theta$ be a target parameter and $\alpha$ be the auxiliary momentum variable. The Hamiltonian is defined as
+\begin{align}
+H(\theta, \alpha) \equiv - \log \pi(\theta, \alpha).
+\end{align}
+
+This can be broken down into kinetic energy $K(\theta, \alpha)$, the density over the auxiliary momentum, and potential energy $V(\theta)$, the density of the target posterior distribution
+\begin{align}
+H(\theta, \alpha) &= - \log \pi(\alpha | \theta) - \log \pi(\theta) \\ 
+&\equiv  K(\theta, \alpha) + V(\theta).
+\end{align}
 
@@
 
-$$
-\begin{aligned}
-\alpha\sim Multinormal(0, M)
-\end{aligned}
-$$
-
-Where M is some diagonal mass matrix (assuming independence between momentum variables).
-
-\textbf{2) Solve the set of Hamiltonian (differential) equations which generates a proposal} $(\theta^{\ast}, \alpha^{\ast})$
-
-$$
-\begin{aligned}
+\begin{align}
+\alpha\sim \mathrm{Multinormal}(0, M),
+\end{align}
+
+where M is some diagonal mass matrix (assuming independence between momentum variables).
+
+\textbf{2) Solve the set of Hamiltonian (differential) equations which generates a proposal} $(\theta^{\ast}, \alpha^{\ast})$ given by
+\begin{align}
 \frac{\partial \theta}{\partial t} &= \frac{\partial H}{\partial \alpha} = \frac{\partial K}{\partial \alpha} \\
-\frac{\partial \alpha}{\partial t} &= \frac{\partial H}{\partial \theta} = - \frac{\partial K}{\partial \theta} - \frac{\partial V}        {\partial \theta}
-\end{aligned}
-$$
-
-Where $\frac{\partial V}{\partial \theta}$ is the gradient of the target posterior density.
+\frac{\partial \alpha}{\partial t} &= \frac{\partial H}{\partial \theta} = - \frac{\partial K}{\partial \theta} - \frac{\partial V}        {\partial \theta},
+\end{align}
+
+where $\frac{\partial V}{\partial \theta}$ is the gradient of the target posterior density.
 The solution to these differential equations is approximated using a Leapfrog integrator which gives a discrete approximate solution.
 The parameters which are tuned in this process are the number of steps $G$ and the size of the steps $\epsilon$.
-The Leapfrog algorithm performs a half step of $\alpha$ and then a full step of the parameter $\theta$, and another half step of the momentum variable $\alpha$ and repeats this $G$ times.
-The final step is the proposal $(\theta^{\ast}, \alpha^{\ast})$.
-
-$$
-\begin{aligned}
+The Leapfrog algorithm performs a half step of $\alpha$ and then a full step of the parameter $\theta$, and another half step of the momentum variable $\alpha$ and repeats this $G$ times where
+\begin{align}
 \alpha &\leftarrow \alpha - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta} \\
 \theta &\leftarrow \theta + \epsilon M^{-1} \alpha \\
-\alpha &\leftarrow \alpha - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta}
-\end{aligned}
-$$
+\alpha &\leftarrow \alpha - \frac{\epsilon}{2} \frac{\partial V}{\partial \theta}.
+\end{align}
+
+The final step is the proposal $(\theta^{\ast}, \alpha^{\ast})$.
 
@@
 
-Let $(\theta^{n-1}, \alpha^{n-1})$ be the values before the Leapfrog integrator.
-
-$$
-\begin{aligned}
-a = \frac{\pi(\theta^{\ast} | y) \pi(\alpha^{\ast})}{\pi(\theta^{n-1} | y) \pi(\alpha^{n-1})}
-\end{aligned}
-$$
-
-With sampled value
-
-$$
+Let $(\theta^{n-1}, \alpha^{n-1})$ be the values before the Leapfrog integrator, then
+\begin{align}
+a = \frac{\pi(\theta^{\ast} | y) \pi(\alpha^{\ast})}{\pi(\theta^{n-1} | y) \pi(\alpha^{n-1})},
+\end{align}
+
+with sampled value
+\begin{align}
 \theta^n = \begin{cases}
     \theta^{\ast},& \text{with probability } min(a,1)\\
-    \theta^{n-1}, & \text{otherwise}
+    \theta^{n-1}, & \text{otherwise}.
     \end{cases}
-$$
+\end{align}
 
@@
 
-\section{Appendix E: Importance weighted rank statistics for re-parameterised model}
+\section{Appendix E: Importance weighted rank statistics for reparameterised model}
 
@@
     \includegraphics[scale=0.09]{results/weighted_ksc_ncp_1k.png}
-    \caption{\textbf{1000 SBC iterations for re-weighted rank statistics from the re-parameterised Gaussian mixture approximation model}.
+    \caption{\textbf{1000 SBC iterations for re-weighted rank statistics from the reparameterised Gaussian mixture approximation model}.
 The shape of the histogram is consistent with the unweighted rank statistics.}
@@
     \includegraphics[scale=0.09]{results/weighted_ksc_ncp_5k.png}
-    \caption{\textbf{5000 SBC iterations for re-weighted rank statistics from the re-parameterised Gaussian mixture approximation model.}}
+    \caption{\textbf{5000 SBC iterations for re-weighted rank statistics from the reparameterised Gaussian mixture approximation model.}}
     \label{fig:ncpreweight5k}

### 2023-10-22 manuscript/main.tex
Final Submission

@@
 \begin{abstract}
-    Simulation based calibration (SBC) \citep{talts2020validating} is applied to analyse two commonly used, competing Markov chain Monte Carlo algorithms for estimating the posterior distribution of a stochastic volatility model.
+    Simulation Based Calibration (SBC) \citep{talts2020validating} is applied to analyse two commonly used, competing Markov chain Monte Carlo algorithms for estimating the posterior distribution of a stochastic volatility model.
 In particular, the bespoke ‘off-set mixture approximation’ algorithm proposed by \citet{kim1998stochastic} is explored together with a Hamiltonian Monte Carlo algorithm implemented through Stan \citep{stan}.
-The SBC analysis involves a simulation study to assess whether each sampling algorithm has the capacity to produce valid inference for the correctly specified model, while also characterising statistical efficiency through effective sample size.
+The SBC analysis involves a simulation study to assess whether each sampling algorithm has the capacity to produce valid inference for the correctly specified model, while also characterising statistical efficiency through the effective sample size.
 Results show that Stan's No-U-Turn sampler, an implementation of Hamiltonian Monte Carlo, produces a well-calibrated posterior estimate while the celebrated off-set mixture approach is less efficient and poorly calibrated, though model parameterisation also plays a role.
@@
 
+\section*{Acknowledgements}
+First and foremost I would like to thank my supervisors Professor Catherine Forbes and Dr Lauren Kennedy.
+Your advice, guidance and patience has been invaluable.
+I am eternally grateful for the opportunity to learn from both of you.
+
+I would also like to acknowledge the Masters research coordinator Professor Param Silvapulle for always checking in and keeping me and my fellow students on track.
+
+Special thanks to Swen for her efforts in proofreading, providing feedback and being a reliable friend who I could talk with about my research.
+
+And to my wife Dana who has supported and encouraged me from the very beginning. Thank you always for your love and patience every single day.
+
+\newpage
+
 \tableofcontents{\protect\newpage}
@@
     This research analyses and compares the calibration and efficiency of MCMC used to estimate stochastic volatility models using a simulation study.
-By calibration we refer to a MCMC algorithm that returns correct posterior, on average, conditional on the correct model specification.
+By calibration we refer to a MCMC algorithm that returns correct the posterior, on average, conditional on the correct model specification.
 The first algorithm is the No-U-turn sampler (NUTS) variant of Hamiltonian Monte Carlo (HMC) as implemented in the Stan programming language \citep{hoffman2014no, betancourt2017conceptual, stan}.
-The second algorithm replicates Kim Shephard and Chib's (KSC) Gaussian off-set mixture model which is sampled using the aforementioned Kalman Filter and Metropolis Hastings algorithm.
+The second algorithm replicates Kim Shephard and Chib's (KSC) Gaussian off-set mixture model which is sampled using the Kalman Filter and Metropolis Hastings algorithm.
 The objective is to analyse the efficiency of the MCMC and their ability to return correct posterior estimates on average.
@@
 Parameters are drawn from the prior distribution and used to create data sets from the generative stochastic volatility model.
-We then obtain posterior samples from each of the MCMC methods.
+Posterior samples are then obtained from each of the MCMC methods conditional on the created data set.
 Repeating this process multiple times gives insight to how well the MCMC algorithm can estimate the true parameters given the data generating process.
@@
 Different parameterisations of a model, whilst mathematically the same, can be very different in terms of sampling difficulty \citep{neal2003slice}.
-In this work two different parameterisations of the stochastic volatility model are compared for each sampling method.
-An algorithm may be sensitive to different parameterisations of the same model which may impact MCMC performance as described in \citet{strickland2008parameterisation}.
+In this work, two different parameterisations of the stochastic volatility model are compared for each sampling method.
+An algorithm may be sensitive to different parameterisations of the same SV model which may impact MCMC performance as described in \citet{strickland2008parameterisation}.
 
@@
 Convergence metrics assessed from the same MCMC output to be used for posterior inference are commonly used to check the performance of the MCMC chains.
-Most often, the effective sample for each parameter is produced, as it indicates the number of approximately independent draws obtained from the dependent Markov chain for a target parameter \citep{gelman2013bayesian}, and $\hat{R}$, the potential scale reduction factor or measure of variability between different chains \citep{gelman1992inference}, is computed, to gauge whether the MCMC algorithm has converged to the target posterior distribution.
+Most often, the effective sample size for each parameter is produced, as it indicates the number of approximately independent draws obtained from the dependent Markov chain for a target parameter \citep{gelman2013bayesian}.
+$\hat{R}$, the potential scale reduction factor or measure of variability between different chains \citep{gelman1992inference}, is also computed to gauge whether the MCMC algorithm has converged to the target posterior distribution.
 
@@
 \subsection{Challenge 2: Limitations of a single simulation}
-    Diagnosing problems with the approximate posterior from models fit to real data is difficult because we do not observe the true data generating process or the true model.
+    Diagnosing problems with the approximate posterior from models fit to real data is difficult because we do not observe the true model.
 A strategy around this is to evaluate a model and algorithm on simulated data.
@@
 
-    Figure 2 shows marginal posterior distributions given a single set of 1000 returns generated by the SV model.
-The true parameters are $\mu=-0.389$, $\sigma^2_{\eta}=0.0115$ and $\phi=0.547$ and the model is fit using Stan's implementation of HMC, the No-U-Turn sampler.
+    Figure 2 shows marginal posterior distributions given a simulated data set of 1000 returns generated by the SV model.
+The true parameters are $\mu=-0.389$, $\sigma^2_{\eta}=0.0115$ and $\phi=0.547$, and the model is fit using Stan's implementation of HMC, the No-U-Turn sampler, with 1000 post burn-in draws from the joint posterior distribution.
 The 95\% credible intervals for $\mu$ and $\sigma^2_{\eta}$ cover the true parameter.
@@
 \subsection{Research Goal}
-    The objective of this research is to design a simulation study to evaluate the calibration of the HMC and KSC off-set mixture algorithms used to estimate SV models.
+    The objective of this research is to evaluate the calibration of the HMC and KSC algorithms used to estimate SV models through a simulation study\footnote{The software used to produce all results and plots in this research are the R v4.2.3\citep{rlang}, Python v3.8.13 \citep{10.5555/1593511} and Stan v2.31.0 \citep{stan} programming languages, including open source R packages (tidyverse v1.3.2 \citep{tidyverse}, quantmod v0.4.20 \citep{quantmod}, cmdstanr 0.5.3 \citep{cmdstanr}, posterior v1.4.1 \citep{posteriorr}, bayesplot v1.4.10 \citep{bayesplot}, jsonlite v1.8.4 \citep{jsonlite}, arrow v12.0.1.1 \citep{arrow} and future.apply v1.11.0 \citep{RJ-2021-048}) and python packages (numpy v1.24.4 \citep{harris2020array}, scipy v.1.10.1 \citep{2020SciPy-NMeth}, statsmodels v.0.14.0 \citep{seabold2010statsmodels}, pandas v.2.0.3 \citep{mckinney-proc-scipy-2010} and pyarrow v12.0.1 \citep{arrow})}.
 As discussed, there are limitations to evaluating MCMC algorithms based on fits to real data and single simulations.
@@
 This section introduces the simulation design used to evaluate whether a MCMC algorithm is returning the correct posterior estimates on average (that is, whether a MCMC algorithm calibrated).
-Section 3.1 discusses the procedure for running simulation based calibration and Section 3.2 defines the diagnostics and metrics for evaluating SBC experiment results.\footnote{The software used to implement these methods and produce all results and plots in this paper are the R v4.2.3\citep{rlang}, Python v3.8.13 \citep{10.5555/1593511} and Stan v2.31.0 \citep{stan} programming languages, including open source R packages (tidyverse v1.3.2 \citep{tidyverse}, quantmod v0.4.20 \citep{quantmod}, cmdstanr 0.5.3 \citep{cmdstanr}, posterior v1.4.1 \citep{posteriorr}, bayesplot v1.4.10 \citep{bayesplot}, jsonlite v1.8.4 \citep{jsonlite}, arrow v12.0.1.1 \citep{arrow} and future.apply v1.11.0 \citep{RJ-2021-048}) and python packages (numpy v1.24.4 \citep{harris2020array}, scipy v.1.10.1 \citep{2020SciPy-NMeth}, statsmodels v.0.14.0 \citep{seabold2010statsmodels}, pandas v.2.0.3 \citep{mckinney-proc-scipy-2010} and pyarrow v12.0.1 \citep{arrow}).}
+Section 3.1 discusses the procedure for running Simulation Based Calibration and Section 3.2 defines the diagnostics and metrics for evaluating SBC experiment results.
 
@@
 The estimated posteriors are then compared to the true values by calculating the rank statistics.
-Under a calibrated model, the $K$ rank statistics for each parameter should be uniformly distributed.
+Under a calibrated algorithm, the $K$ rank statistics for each parameter should be uniformly distributed.
 
@@
         \end{align}
-        then generate a data set $y^{sim}$ conditionally given $\theta^{sim}$ and the model of interest\footnote{Since all computation presumes a correctly specified model, notation to condition on the model is suppressed.}:
+        then generate a data set conditional on $\theta^{sim}$ and the model of interest\footnote{Since all computation presumes a correctly specified model, notation to condition on the model is suppressed.},
         \begin{align}
@@
         \end{align}
-        Take $L$ draws from the estimated posterior distribution, conditional on this data set, and generated by the MCMC algorithm under investigation (here, HMC or KSC):
+        Take $B$ draws from the estimated posterior distribution, conditional on this data set, and generated by the MCMC algorithm under investigation (here, HMC or KSC)
         \begin{align}
@@
 
-        \citet{talts2020validating} prove that the sample of rank statistics for a given parameter, produced from replicate posteriors, follow a uniform distribution over the integers $\{0,1\dots,B\}$ if the posterior samples follow the same distribution as the prior on average.
-The rank statistic for any one-dimensional function of parameters $f:\Theta\rightarrow\ \mathbb{R}$ produced by the $k^{\mathrm{th}}$ replicate posterior sample, is defined as
-        \begin{align}
-        r^{(k)} = \mathrm{rank}(\{f(\theta_1^{(k)}),\dots , f(\theta_{B}^{(k)})\}, f(\theta^{sim})) = \sum_{b=1}^{B}1[f(\theta_{b}^{(k)}) < f(\theta^{sim})].
-        \end{align}
-
-        This completes one iteration of SBC.
-To complete the algorithm, multiple iterations are run and the rank statistics are calculated for each parameter.
-The resulting rank statistics are compared to the uniform distribution to determine if the algorithm is calibrated and returning the correct posteriors.
-
-        A key result is that the posterior averaged over the data and true parameters will equal the prior distribution. This is shown in the following expression
+        The last step is to calculate the rank statistic. The rank statistic for any one-dimensional function of parameters $f:\Theta\rightarrow\ \mathbb{R}$ is defined as
+        \begin{align}
+        r = \mathrm{rank}(\{f(\theta_1),\dots , f(\theta_{B})\}, f(\theta^{sim})) = \sum_{b=1}^{B}1[f(\theta_{b}) < f(\theta^{sim})].
+        \end{align}
+
+        Equations (10)-(13) describes one iteration of SBC. To complete the algorithm, multiple iterations are run and the rank statistics are calculated for each parameter.
+
+        \citet{talts2020validating} prove that the sample of rank statistics for a given parameter follow a uniform distribution over the integers $\{0,1\dots,B\}$ if the MCMC algorithm is calibrated.
+Therefore, the rank statistics calculated from the SBC procedure are compared to the uniform distribution to determine if the algorithm is returning the correct posterior estimates on average.
+
+        Another key result for calibration is that the posterior averaged over the data and true parameters is equal to the prior distribution. This is shown in the following expression
         \begin{align}
@@
         \end{align}
-        That is, the average posterior for $\theta$ over the SBC iterations should approximately equal to prior distribution, if the algorithm is calibrated.
-Therefore, any deviation between the prior distribution and the average posterior draws produced from replicated posterior distributions means that the sampling methodology is not producing the correct posterior, on average.
+        That is, the average posterior for $\theta$ over the SBC iterations should approximately equal the prior distribution, if the algorithm is calibrated.
+Therefore, any deviation between the prior distribution and the average posterior draws over the SBC iterations may suggest that the sampling methodology is not producing the correct posterior on average.
 
@@
 
-            The shape of the distribution of rank statistics gives insight into how an MCMC may be miscalibrated.
+            The shape of the distribution of rank statistics gives insight into how the MCMC may be miscalibrated.
 Specifically, it gives information about how the average posterior deviates from the prior distribution of a given parameter.
@@
 
-            Figure \ref{fig:underestimation} shows a distribution of rank statistics with inflated frequency on the right side of the histogram.
-This distribution is consistent with posterior samples that tend to underestimate the true parameter on average.
+            The left panel of Figure \ref{fig:underestimation} shows a distribution of rank statistics with inflated frequency on the right side of the histogram.
+This distribution is consistent with posterior samples that tend to underestimate the prior distribution on average, as shown on the right panel.
 The sum of indicator random variables (see equation (13)) is large since a large proportion of posterior draws are smaller than the true value, resulting in a large rank statistic and hence a larger relative frequency on the right side of the histogram.
@@
 
-            Figure \ref{fig:underdispersed} has large inflated frequencies on both ends of the histogram resulting in a `U-shape'.
+            The left panel of Figure \ref{fig:underdispersed} has large inflated frequencies on both ends of the histogram resulting in a `U-shape'.
 \citet{talts2020validating} describe this as under-dispersion of the posterior relative to the prior distribution on average.
-The average posterior is too narrow (or overconfident) relative to the spread of the prior resulting in inflated frequency on both ends of the rank statistic distribution.
+The average posterior is too narrow (or overconfident) relative to the spread of the prior, as seen on the right panel, resulting in inflated frequency on both ends of the rank statistic distribution.
 
@@
                 \includegraphics[scale=0.07]{methodology/rhs.png}
-                \caption{\textbf{Left:} Non uniform rank statistics with inflated frequency on the right side.
-\textbf{Right:} The average posterior distribution underestimates the prior distribution.
-The miscalibration from the rank statistic distribution comes from posterior samples underestimating the true parameter on average.}
+                \caption{\textbf{Left panel:} Non uniform rank statistics with inflated frequency on the right side.
+\textbf{Right panel:} The average posterior distribution underestimates the prior distribution.
+The miscalibration from the rank statistic distribution comes from posterior samples underestimating the prior distribution on average.}
                 \label{fig:underestimation}
@@
                 \includegraphics[scale=0.07]{methodology/underdispersed.png}
-                \caption{\textbf{Left}: Non uniform rank statistics with inflated frequencies on both ends (`U-shape' distribution).
-\textbf{Right}: Average posterior is underdispersed relative to the prior distribution.
+                \caption{\textbf{Left panel}: Non uniform rank statistics with inflated frequencies on both ends (`U-shape' distribution).
+\textbf{Right panel}: Average posterior is underdispersed relative to the prior distribution.
 Non-uniformity in rank statistics comes from posteriors which are over-confident relative to the prior.}
@@
 ESS estimates the number of effectively independent draws generated by a MCMC to estimate a target parameter.
-A poor ESS generally arises from high auto-correlation in the Markov chain, and thus is an indication of a highly dependent posterior sample.
+A poor ESS generally arises from high autocorrelation in the Markov chain, and thus is an indication of a highly dependent posterior sample.
 A more efficient MCMC algorithm will take a relatively lower level of resources (for example, time and number of draws) to get a sample from the target distribution when compared to a less efficient algorithm.
@@
             \end{align}
-            When more than one chain is generated from the target posterior, then the ESS quantity is a function of the number of chains $M$, the number of samples per chain $N$ and the auto-correlation at lag $l$ $\rho_l$ \citep{vehtari2021rank}.
-$L$ is a truncation lag chosen such that the sum of any two consecutive auto-correlation lags is negative \citep{geyer1992practical}.
-This truncation (or some kind of down-weighting) is required because a large number of lags leads to noisier auto-correlation estimates.
-Note that other approaches to estimating the true auto-correlation of the Markov chain and the ESS can be found in the literature.
-Details about this ESS formula, the calculation of the auto-correlation and other ESS calculations can be found in \citet{vehtari2021rank} and \citet{geyer1992practical}.
+            When more than one chain is generated from the target posterior, then the ESS quantity is a function of the number of chains $M$, the number of samples per chain $N$ and the autocorrelation at lag $l$, $\rho_l$ \citep{vehtari2021rank}.
+$L$ is a truncation lag chosen such that the sum of any two consecutive autocorrelation lags is negative \citep{geyer1992practical}.
+This truncation (or some kind of down-weighting) is required because a large number of lags leads to noisier autocorrelation estimates.
+Note that other approaches to estimating the true autocorrelation of the Markov chain and the ESS can be found in the literature.
+Details about this ESS formula, autocorrelation and other ESS calculations can be found in \citet{vehtari2021rank} and \citet{geyer1992practical}.
 
@@
 The leapfrog integrator has a set of tuning parameters which determines the size and number of steps to be taken along the Hamiltonian path with the final step being the proposal value $(\theta^{\ast}, \alpha^{\ast})$.
-Finally, the Metropolis Hastings accept/reject step is applied using the parameters at the beginning of the leapfrog process $(\theta^{n-1}, \alpha^{n-1})$ and the proposal values $(\theta^{\ast}, \alpha^{\ast})$.
+Finally, the Metropolis Hastings accept/reject step is applied using the parameters at the beginning of the leapfrog process $(\theta^{b-1}, \alpha^{b-1})$ and the proposal values $(\theta^{\ast}, \alpha^{\ast})$.
 A formal description of the HMC algorithm is provided in Appendix C.
@@
 The No-U-Turn Sampler (NUTS) proposes the final parameter using multinomial sampling biased towards the second half of the trajectory steps instead of applying the Metropolis-Hastings step \citep{betancourt2016identifying}.
-Stan allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
+Stan samples the generative model directly and can flexibly handle complicated likelihood functions.
 
@@
     \subsection{Sampling method 2: Off-set Mixture MCMC}
-        \citet{kim1998stochastic} sample the conditional posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings \citep{metropolis1953equation, hastings1970monte} within Gibbs \citep{geman1984stochastic} algorithms and an implementation of the Kalman Filter \citep{kalman1960new} and smoother used to sample from the conditional posterior distribution of the latent states \citep{dejong1995}.\footnote{In this research the exact software to apply the simulation smoother is unavailable.
+        \citet{kim1998stochastic} sample the conditional posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings \citep{metropolis1953equation, hastings1970monte} within Gibbs \citep{geman1984stochastic} algorithms and an implementation of the Kalman Filter \citep{kalman1960new} and simulation smoother used to sample from the conditional posterior distribution of the latent states \citep{dejong1995}.\footnote{In this research the exact software to apply the simulation smoother is unavailable.
 So a more recent simulation smoother is used from the statsmodels Python package by \citet{seabold2010statsmodels}.
@@
         The stochastic volatility model as described in section 2.1 follows a centered parameterisation.
-This describes the central location of the latent state distribution which is centered on the mean of log volatility $\mu$.
+This describes the central location of the latent state distribution which is centered on the mean of the log variance $\mu$.
 A model can be reparameterised in a way that might improve MCMC performance.
@@
 Reparameterised HMC samples from the joint posterior directly with state vectors centered at zero and transformed to return the correct log volatility estimates.
-Applying the Kalman Filter requires the state equation to be a function of the state variable with the average log volatility parameter $\mu$ entering the measurement equation.
+Applying the Kalman Filter requires the state equation to be a function of the state variable with the average log variance parameter $\mu$ entering the measurement equation.
 
@@
         \end{align}
-        Return average volatility into measurement equation and rewrite as a function of the non centered state equation
+        Return average log variance into measurement equation and rewrite as a function of the non centered state equation
         \begin{align}
@@
         \end{align}
-        where the initial state now is drawn from 
+        where the initial state is drawn from 
         \begin{align}
@@
         \end{align}
-        The mean of the log volatility is now inside the measurement equation and demeaned from the state equation.
+        The mean of the log variance is now inside the measurement equation and demeaned from the state equation.
 
@@
     SBC is conducted for each MCMC sampling method and corresponding parameterisation as outlined in Table \ref{tab:params}.
-The SV model in Section 2.1 is used for all simulations including the same priors.
+The priors in Section 2.1 are used for all simulations.
 For a given MCMC and parameterisation pair, 5000 SBC iterations\footnote{Results for 1000 SBC iterations returned noisy estimates and can be found in Appendix D.} are run using simulated data sets of size $T=1000$.
@@
     Section 5.1 and Section 5.2 interpret the SBC results for HMC and KSC's MCMC method respectively.
-Section 5.3 describes the importance weights used to correct for approximation error when applying the off-set mixture model and applying these weights to rank statistics.
+Section 5.3 describes the importance weights used to correct for approximation error when using the off-set mixture model and applying these weights to rank statistics.
 Section 5.4 presents a summary of rank statistic distributions for all latent state variables using chi-squared statistics.
@@
 The reparameterised model generates independent samples much more efficiently than the centered model, reflected in the larger ESS values across every summary statistic.
-The median ESS for $\mu$, $\phi$ and $\sigma^2$ in the centered parameterisation is much smaller as a proportion of the posterior draws.
 This is consistent with HMC struggling to sample from the posteriors of the static parameters overall in the centered model.
@@
 The centered parameterisation struggles to generate independent samples for $\sigma^2$ and $\phi$.
-Non parameterisation demonstrates much more efficient sampling behaviour.}
+The reparameterised model demonstrates much more efficient sampling behaviour.}
         \label{fig:hmcess}
@@
 Furthermore, the number of posterior samples is larger than the 999 taken for the SBC experiment run for Hamiltonian Monte Carlo.
-Earlier experiments attempted 999 post burn-in draws for the off-set mixture model, however there are concerns that such a short chain would not have converged onto the target posterior distribution.
+Earlier experiments attempted 999 post burn-in draws for the off-set mixture model, however there were concerns that such a short chain would not have converged onto the target posterior distribution.
 Therefore 9,999 draws was chosen as a compromise under both constraints.
 
-    Both the centered and the non-centered Gaussian off-set mixture model contains issues with calibration.
+    Both the centered and the non-centered Gaussian off-set mixture model contain issues with calibration.
 Figure \ref{fig:cpksc5k} presents the results from the centered model.
@@
 Each histogram exhibits large inflated frequencies at one or both ends, suggesting major issues in the sampler returning correct posterior estimates.
-The KSC MCMC algorithm applied to this parameterisation of the model is uncalibrated looking at these parameters.
-Based on the results of both parameterisations, applying this MCMC algorithm on real data will on balance produce incorrect and unreliable posterior samples on average (conditional on the model and the number of post burn-in draws).
+The KSC MCMC algorithm applied to this parameterisation of the model looks uncalibrated overall.
+The results of both parameterisations suggest applying this MCMC algorithm on real data will on balance produce incorrect and unreliable posterior samples on average (conditional on the model and the number of post burn-in draws).
     \begin{figure}[H]
@@
 
-    The ESS estimates for this model indicate a high degree of auto-correlation and difficulty in generating independent samples from both parameterisations.
+    The ESS estimates for this model indicate a high degree of autocorrelation and difficulty in generating independent samples from both parameterisations.
 The numerical results are summarised in Figure \ref{fig:kscess} and Table \ref{tab:kscess}.
-The estimate for $\mu$ in the reparameterised model exhibits a long right tail, with values larger than the 9,999 post burn-in posterior draws suggesting the presence of negatively auto-correlated draws.
+The estimate for $\mu$ in the reparameterised model exhibits a long right tail, with values larger than the 9,999 post burn-in posterior draws suggesting the presence of negatively autocorrelated samples.
 The median ESS for each static parameter relative to the number of posterior draws is very small.
-There is some improvement from the reparameterised model with higher ESS values across some of the static parameter quantiles, but these improvements are negligible.
+There is some improvement for the reparameterised model with higher ESS values across some of the static parameter quantiles, but most of these improvements are negligible.
 Overall, the off-set mixture MCMC is inefficient at sampling the static parameters.
-Note, Figure \ref{fig:hmcess} and Figure \ref{fig:kscess} are not directly compared due to scale difference from varying post burn-in posterior draws.
+Note, Figure \ref{fig:hmcess} and Figure \ref{fig:kscess} are not directly compared due to scale difference from varying posterior draws.
 
@@
         \includegraphics[scale=0.1]{results/ksc_ess.png}
-        \caption{\textbf{Effective sample sizes for static parameters after 5000 iterations of SBC and 1000 post burn-in samples for the Gaussian Mixture model}.
-The KSC sampling strategy struggles to generate independent samples for all static parameters, suggesting a high degree of auto-correlation in the Markov Chain.}
+        \caption{\textbf{Effective sample sizes for static parameters after 5000 iterations of SBC and 9999 post burn-in draws for the off-set mixture model}.
+The KSC sampling strategy struggles to generate independent samples for all static parameters and parameterisations, suggesting a high degree of autocorrelation in the Markov chain.}
         \label{fig:kscess}
@@
         \end{tabular}
-        \caption{\textbf{ESS for centered and reparameterised SV model across 5000 SBC iterations for KSC}.}
+        \caption{\textbf{ESS for centered and reparameterised SV model across 5000 SBC iterations and 9999 post burn-in draws for the off-set mixture model}.}
         \label{tab:kscess}
@@
 
-    Importance weights are defined as the ratio of the joint posterior from the stochastic volatility model and the posterior distribution of the approximate model.
+    Importance weights are defined as the ratio of the joint posterior from the SV model and the posterior distribution of the approximate model.
 A weight is produced for each posterior sample generated by MCMC.
@@
     \subsubsection{Re-weighting rank statistics of the Gaussian off-set mixture model}
-    Let $v(\theta, h)$ be the log weights defined as the log difference between the posterior densities of the true model with log chi-squared errors $\log c(\theta, h | y^{\ast})$ and the mixture off-set model $\log d(\theta, h | y^{\ast})$.
+    Let $v(\theta, h)$ be the log weights defined as the log difference between the posterior densities of the true model with log chi-squared errors $\log c(\theta, h | y^{\ast})$ and the off-set mixture model $\log d(\theta, h | y^{\ast})$.
 The is given in the following expression
     \begin{align}
-        v(\theta, h) = \log c(\theta, h | y^{\ast}) - \log  d(\theta, h | y^{\ast}) = \text{const} + \log c(y|h) - \log d(y^{\ast} | h).
+        v(\theta, h) = \log c(\theta, h | y^{\ast}) - \log  d(\theta, h | y^{\ast}) = \text{const} + \log c(y^{\ast}|h) - \log d(y^{\ast} | h).
     \end{align}
@@
     The results from applying this re-weighting step to the rank statistics of the centered model are given on Figure \ref{fig:reweight5k}.
-There are no improvements in the distribution of rank statistics.
+There are no improvements to the uniformity of the rank statistic distributions.
 The shape of the histograms is consistent with the shape of the unweighted rank statistics.
@@
         \includegraphics[scale=0.1]{results/weighted_ksc_cp_5k.png}
-        \caption{\textbf{Distribution of rank statistics with 5000 SBC iterations for re-weighted rank statistics from the Gaussian off-set mixture model}.
-The shape of the histogram is consistent with the unweighted rank statistics.}
+        \caption{\textbf{Distribution of rank statistics for re-weighted rank statistics from the off-set mixture model}. The shape of the histogram is consistent with the unweighted rank statistics.}
         \label{fig:reweight5k}
@@
 Both KSC and the importance weighted ranks on the other hand are far from zero suggesting larger deviations from uniformity.
-There is a wide gap between the HMC and KSC chi-squared statistics implying the distribution of latent state ranks are much more  uniform for HMC relative to KSC.
+There is a wide gap between the HMC and KSC chi-squared statistics implying the distribution of latent state rank statistics are much more  uniform for HMC relative to KSC.
 Combined with the results in the previous section, there is a strong evidence that the HMC algorithm outperforms the KSC algorithm where it comes to calibration across all parameters and latent state variables.
@@
         \caption{\textbf{Distribution of chi-squared statistics for latent state variables from 5000 SBC iterations}.
-Values closer to 0 are more consistent with a discrete uniform distribution.
+Values closer to 0 are more uniformly distributed.
 Both HMC simulations are much closer to zero than KSC.
@@
 This research applied a simulation study to evaluate the calibration and efficiency of different MCMC algorithms in the context of the stochastic volatility model.
-Simulation based calibration was used to determine whether these algorithms were returning the correct posterior estimates on average under different SV model parameterisations.
+Simulation Based Calibration was used to determine whether these algorithms were returning the correct posterior estimates on average under different SV model parameterisations.
 
@@
 Centered parameterisation using HMC performed moderately well but struggled to produce calibrated posterior estimates for $\sigma^2$ and $\phi$ as well as generate independent samples for these parameters.
-The Gaussian off-set mixture approximation struggled to return calibrated posterior estimates and efficient samples for both parameterisations.
+The off-set mixture MCMC struggled to return calibrated posterior estimates and efficient samples for both parameterisations.
 However, the centered parameterisation results for the Gaussian mixture model may improve if we increased the length of the Markov chain.
@@
 The failure of non centered parameterisation in location for the off-set mixture model could be due to a variety of reasons.
-This may be due to a particular software implementation in the Kalman Filter and Simulation Smoother (where a specific software package was applied) that was not designed to handle the non-centered parameterisation.
+This may be due to a particular software implementation of the Kalman Filter and Simulation Smoother (where a specific software package was applied in this research) that was not designed to handle the non-centered parameterisation.
 Or perhaps this particular parameterisation does not suit the sampling strategy applied by KSC and another approach (e.g.
@@
 Lots of different design and implementation details could be explored further.
-Pinpointing the reason why this model parameterisation failed as well as exploring other configurations and MCMC algorithms in this context is left to future research.
+Pinpointing the reason why this model parameterisation failed as well as exploring other configurations and MCMC algorithms in this context is left for future research.
 
@@
 \subsection{Limitations}
-A limitation to applying SBC in all modelling contexts is that it is computationally intensive.
+A limitation to applying SBC in any modelling context is that it is computationally intensive.
 SBC requires fitting multiple models to get reliable estimates.
@@
 A fairer comparison may be to have roughly similar number of ESS for most of the estimated parameters, although this is difficult to get right with a large number of parameters to estimate.
-Given how well Hamiltonian Monte Carlo performed with the reparameterised model, any improvement to the Gaussian off-set mixture model by increasing the length of the Markov chain would make calibration just as good, but not better due to the inefficiency of the sampler.
+Given how well Hamiltonian Monte Carlo performed with the reparameterised model, any improvement to the off-set mixture MCMC by increasing the length of the Markov chain would make calibration just as good, but not better due to the inefficiency of the sampler.
 
@@
 This study only considered a handful of potential algorithms, sampling strategies and parameterisations for stochastic volatility.
-Future research could use the same SBC design to compare and evaluate more complicated volatility models and MCMC algorithms.
+Future research could use the same SBC design to compare and evaluate more complicated stochastic volatility models and MCMC algorithms.
 Overall, SBC is a valuable tool in model development and is a key part of a statistical workflow.
@@
 \textbf{1) Randomly sample a momentum value}
-
 \begin{align}
@@
 
-Let $(\theta^{n-1}, \alpha^{n-1})$ be the values before the Leapfrog integrator, then
-\begin{align}
-a = \frac{\pi(\theta^{\ast} | y) \pi(\alpha^{\ast})}{\pi(\theta^{n-1} | y) \pi(\alpha^{n-1})},
+Let $(\theta^{b-1}, \alpha^{b-1})$ be the values before the Leapfrog integrator, then
+\begin{align}
+a = \frac{\pi(\theta^{\ast} | y) \pi(\alpha^{\ast})}{\pi(\theta^{b-1} | y) \pi(\alpha^{b-1})},
 \end{align}
@@
 \begin{align}
-\theta^n = \begin{cases}
+\theta^b = \begin{cases}
     \theta^{\ast},& \text{with probability } min(a,1)\\
-    \theta^{n-1}, & \text{otherwise}.
+    \theta^{b-1}, & \text{otherwise}.
     \end{cases}
@@
         \caption{\textbf{1000 SBC iterations for centered model using Hamiltonian Monte Carlo}.
-$\sigma^2$ displays peaks on both ends of the histogram and $\phi$ shows a major on the left hand side.
-This suggests that the average posterior samples of $\sigma^2$ are under-dispersed relative to the prior distribution and posterior samples of $\phi$ are over estimating the true parameter.}
+$\sigma^2$ displays inflated frequencies on both ends of the histogram and $\phi$ shows inflated frequency left hand side.
+This suggests that the average posterior samples of $\sigma^2$ are under-dispersed relative to the prior distribution and posterior samples of $\phi$ are overestimating the true parameter.}
         \label{fig:cphmc1k}
@@
         \includegraphics[scale=0.09]{results/hmc_ncp_1k.png}
-        \caption{\textbf{1000 SBC iterations for non centered model using Hamiltonian Monte Carlo}.
+        \caption{\textbf{1000 SBC iterations for reparameterised model using Hamiltonian Monte Carlo}.
 The rank statistic distributions for $\sigma^2$ and $\phi$ have improved.
@@
         \caption{\textbf{1000 SBC iterations for centered Gaussian off-set mixture model}.
-$\mu$ has a large left peak suggesting the algorithm is over estimating the true parameter on average.
-The converse can be said about $\phi$ which possesses a peak on the right hand side which implies underestimation of the true parameter.}
+$\mu$ has an inflated frequency on the left side suggesting the algorithm is over estimating the true parameter on average.
+The converse can be said about $\phi$ which possesses inflated frequency on the right hand side which implies underestimation of the true parameter.}
         \label{fig:cpksc1k}
@@
         \includegraphics[scale=0.09]{results/weighted_ksc_cp_1k.png}
-        \caption{\textbf{1000 SBC iterations for re-weighted rank statistics from the Gaussian mixture approximation model}.
-The shape of the histogram is consistent with the unweighted rank statistics.}
+        \caption{\textbf{1000 SBC iterations for re-weighted rank statistics from the off-set mixture model}.}
         \label{fig:reweight1k}
@@
     \includegraphics[scale=0.09]{results/weighted_ksc_ncp_1k.png}
-    \caption{\textbf{1000 SBC iterations for re-weighted rank statistics from the reparameterised Gaussian mixture approximation model}.
-The shape of the histogram is consistent with the unweighted rank statistics.}
+    \caption{\textbf{1000 SBC iterations for re-weighted rank statistics from the reparameterised off-set mixture model}.}
     \label{fig:ncpreweight1k}
@@
     \includegraphics[scale=0.09]{results/weighted_ksc_ncp_5k.png}
-    \caption{\textbf{5000 SBC iterations for re-weighted rank statistics from the reparameterised Gaussian mixture approximation model.}}
+    \caption{\textbf{5000 SBC iterations for re-weighted rank statistics from the reparameterised off-set mixture model.}}
     \label{fig:ncpreweight5k}

### 2023-10-22 R/sbc.r
final updates

@@ -196,3 +196,54 @@ clean_variable_names <- function(dataframe){
 
     return(result)
 }
+
+
+rank_bins_f <- function(path, rank_type, posterior_samples, n_bins){
+    rds_list <- list.files(path = paste(path, "output", sep = "/"), pattern = "*.RDS")
+    rds_path <- paste("output", rds_list, sep = "/")
+
+    n_iterations <- length(rds_path)
+    posterior_samples <- posterior_samples # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+    n_bins <- n_bins # Bins
+    expected_count <- n_iterations / n_bins # Chi sqd stat
+
+    results <- lapply(rds_path, get_ranks, model_path = path, rank_type = rank_type)
+
+    if (rank_type == "agg_ranks"){
+        df <- as.data.frame(do.call("rbind", results))
+        rank_bins <- as.data.frame(lapply(df, binning, posterior_draws = posterior_samples, bins = n_bins))
+    } else if(rank_type == "weighted_ranks"){
+        df <- as.data.frame(do.call("rbind", results)) * (posterior_samples+1)
+        rank_bins <- as.data.frame(lapply(df, binning, posterior_draws = posterior_samples+1, bins = n_bins))
+    } else {
+        print("error")
+    }
+
+    names(df) <- clean_variable_names(df)
+    names(rank_bins) <- clean_variable_names(rank_bins)
+
+    rank_bins <- rank_bins[!(names(rank_bins) %in% c("y_sim", "sigma"))]
+
+    return(rank_bins)
+}
+
+static_rank_bins <- function(data, variables, type){
+    result <- data %>%
+        select(all_of(variables)) %>%
+        tidyr::pivot_longer(everything(), names_to = "variable", values_to = "rank") %>% 
+        mutate(type = type)
+
+    return(result)
+}
+
+rank_hist <- function(dataframe){
+        ggplot(dataframe) +
+        geom_histogram(aes(rank), bins = 20, fill = "light blue", alpha = 0.7) +
+        scale_x_continuous(labels = function(x) x * (posterior_samples+1)/n_bins) +
+        facet_wrap(~factor(variable, levels = parameters), labeller = label_parsed) +
+        theme_minimal(base_size = 20) +
+        theme(strip.text.x = element_text(size = 22)) +
+        geom_hline(yintercept = expected_count, linewidth = 1, alpha = 0.6) +
+        labs(y = "Count",
+             x = "Ranks")
+}

### 2023-10-22 analysis/prior_predictive_checks/h2.r

@@ -1,3 +1,5 @@
+library(ggplot2)
+
 gen_ncp_sv_dataset <- function(iter, seed_list, T) {
     set.seed(seed_list[iter])
 
@@ -53,9 +55,12 @@ ncp <- as.data.frame(do.call("rbind", ncp_list))
 cp <- as.data.frame(do.call("rbind", cp_list))
 
 
-ggplot(ncp) +
+hist <- ggplot(ncp) +
     geom_histogram(aes(V1), fill = 'blue', alpha = 0.2) +
     geom_histogram(data = cp, aes(V1), fill = 'red', alpha = 0.2) +
     theme_minimal() +
     labs(title = "Prior simulations of CP (red) and NCP (blue) h[2] parameters",
         x = "h[2]")
+
+ggsave(paste(here::here(), "/analysis/prior_predictive_checks/ppc_h2.png", sep = ""), hist, bg = "white", width = 11)
+

### 2023-10-22 analysis/sbc/sandbox.r

@@ -1,0 +1,204 @@
+library(dplyr)
+library(ggplot2)
+library(future.apply)
+library(posterior)
+library(car)
+source("R/sbc.r")
+plan(multicore, workers = 9)
+
+# Get all parameter data
+path <- "simulation_output/ksc/cp/sbc_cp_ksc_model_cp_dgf_10kmcmc_r2"
+rds_list <- list.files(path = paste(path, "output", sep = "/"), pattern = "*.RDS")
+rds_path <- paste("output", rds_list, sep = "/")
+
+# Rank type
+rank_type = "weighted_ranks"
+# rank_type = "agg_ranks"
+
+# Parameters
+n_iterations <- length(rds_list)
+posterior_samples <- 9999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+n_bins <- 20 # Bins
+expected_count <- n_iterations / n_bins # Chi sqd stat
+
+# Load ranks
+results <- lapply(rds_path, get_ranks, model_path = path, rank_type = rank_type)
+
+if (rank_type == "agg_ranks"){
+    df <- as.data.frame(do.call("rbind", results))
+} else if(rank_type == "weighted_ranks"){
+    df <- as.data.frame(do.call("rbind", results)) * (posterior_samples+1)
+} else {
+    print("error")
+}
+
+# Create bins and chi sq stats
+rank_bins <- as.data.frame(lapply(df, binning, posterior_draws = posterior_samples+1, bins = n_bins))
+rank_bins <- rank_bins[!(names(rank_bins) %in% c("y_sim", "sigma"))]
+
+rank_bins %>% select(mu) %>% distinct() %>% count()
+
+rank_stats <- as.data.frame(lapply(rank_bins, rank_chi_sqd, expected = expected_count))
+
+# Define static and state parameters
+static_parameters <- c("mu", "sigma_sqd", "phi")
+state_parameters <- names(rank_bins)[!(names(rank_bins) %in% static_parameters)]
+additional_parameters <- c("h.1.", "h.10.", "h.100.", "h.500.", "h.1000.", "h.995.")
+
+t = rank_bins %>%
+        select(mu) %>%
+        tidyr::pivot_longer(everything(), names_to = "variable", values_to = "rank") 
+
+t %>% select(rank) %>% distinct() %>% count()
+
+hist(t$rank, breaks = 20)
+        
+    rank_bins %>%
+        select(mu) %>%
[…]

### 2023-10-22 manuscript/methodology/rank_stats_examples.r

@@ -29,7 +29,7 @@ dens_both <- ggplot(df, aes(x = x)) +
     annotate("text", x = 4.5, y = 0.05, label = "Prior", colour = "red", size = 8) +
     annotate("text", x = 2.5, y = 0.3, label = "Posterior", colour = "blue", size = 8)
 
-plot_grid(hist_lhs_rhs, dens_both)
+plot_grid(hist_unif, dens_both)
 
 ggsave("manuscript/methodology/underdispersed.png", bg = "white", width = 12, height = 8.42)
 
@@ -46,7 +46,7 @@ dens_lhs<- ggplot(df, aes(x = x)) +
     geom_line(aes(y = post_draws), colour = "blue") + 
     geom_line(aes(y = prior_draws), colour = 'red') +
     theme_void() +
-    annotate("text", x = 0, y = 0.15, label = "Posterior", colour = "red", size = 8) +
+    annotate("text", x = 0, y = 0.15, label = "Average\nPosterior", colour = "red", size = 8) +
     annotate("text", x = -5, y = 0.15, label = "Prior", colour = "blue", size = 8)
 
 
@@ -55,8 +55,12 @@ hist_lhs <- ggplot(hist_df, aes(x=v1)) +
     geom_histogram(fill = "red", bins=20, alpha = 0.4) +
     theme_void()
 
+plot_grid(hist_lhs, ggplot(df) + theme_void())
+ggsave("manuscript/methodology/half_lhs.png", bg = "white", width = 14, height = 8.42)
+
 plot_grid(hist_lhs, dens_lhs)
 ggsave("manuscript/methodology/lhs.png", bg = "white", width = 14, height = 8.42)
+
 
 # RHS
 x <- seq(-8,4, 0.01)
@@ -71,7 +75,7 @@ dens_rhs <- ggplot(df, aes(x = x)) +
     geom_line(aes(y = prior_draws), colour = 'red') +
     theme_void() +
     annotate("text", x = 0, y = 0.15, label = "Prior", colour = "red", size = 8) +
-    annotate("text", x = -5, y = 0.15, label = "Posterior", colour = "blue", size = 8)
+    annotate("text", x = -5, y = 0.15, label = "Average\nPosterior", colour = "blue", size = 8)
 
 hist_df = as.data.frame(list(v1 = c(rep(20,500), runif(3000, 1,19))))
 hist_rhs <- ggplot(hist_df, aes(x=v1)) +
[…]

### 2023-10-22 manuscript/motivating_example/real_data_example.r

@@ -24,15 +24,15 @@ means <- df %>% group_by(model, name) %>% summarise(avg = mean(value))
 medians <- df %>% group_by(model, name) %>% summarise(med = median(value))
 
 ggplot(df, aes(x = value)) +
-    geom_histogram(position="identity", aes(fill = model), alpha =0.3) +
+    geom_histogram(position="identity", aes(y= after_stat(density), fill = model), alpha =0.3) +
     facet_wrap(~name, scale="free", labeller = label_parsed) +
     # geom_vline(data =means, aes(xintercept=avg, colour=model), size = 1.2) +
-    theme_minimal(base_size = 20) +
+    theme_bw(base_size = 20) +
     scale_fill_discrete(labels = c("KSC", "HMC")) +
-    labs(title = "Posterior distribution - S&P 500",
+    labs(title = "",
          fill = "MCMC\nSamplers",
          x = "Parameter value", 
-         y= "Count") + 
+         y= "Density") + 
        theme(strip.text.x = element_text(size = 22))
 
 ggsave("manuscript/motivating_example/real_data_ex.png", bg = "white", width = 14, height = 9.42)
@@ -41,7 +41,7 @@ ggsave("manuscript/motivating_example/real_data_ex.png", bg = "white", width = 1
 
 df %>% group_by(model, name) %>% 
   summarise(as_tibble(rbind(summary(value)))) %>% 
-  ungroup() %>% arrange(model, name)
+  ungroup() %>% arrange(name)
 
 ggplot(df, aes(x = value)) +
     geom_histogram(aes(fill = model), alpha =0.4) +

### 2023-10-22 manuscript/results/compare_histograms.r

@@ -57,61 +57,136 @@ ksc_cp_rank_stats$name <- names(ksc_cp_fivek_bins)
 ksc_cp_rank_stats$type <- "Centered KSC"
 ksc_cp_rank_stats$group <- "KSC"
 
-# KSC NONCENTERED results
+# KSC CENTERED ISW results
 n_iterations <- 5000
 posterior_samples <- 9999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
 n_bins <- 20 # Bins
 expected_count <- n_iterations / n_bins # Chi sqd stat
 
-ksc_ncp_fivek_bins <- rank_bins_f(path = "simulation_output/ksc/ncp/sbc_ncp_ksc_model_ncp_dgf_10kmcmc_5000iter_r1", 
-                                    rank_type = "agg_ranks", 
+ISW_fivek_bins <- rank_bins_f(path = "simulation_output/ksc/cp/sbc_cp_ksc_model_cp_dgf_10kmcmc_5000iter_r1", 
+                                    rank_type = "weighted_ranks", 
                                     posterior_samples = posterior_samples, 
                                     n_bins= n_bins)
-ksc_ncp_rank_stats <- as.data.frame(lapply(ksc_ncp_fivek_bins, rank_chi_sqd, expected = expected_count)) %>% 
+ISW_rank_stats <- as.data.frame(lapply(ISW_fivek_bins, rank_chi_sqd, expected = expected_count)) %>% 
             pivot_longer(everything())
 
-ksc_ncp_rank_stats$name <- names(ksc_ncp_fivek_bins)
-ksc_ncp_rank_stats$type <- "Reparameterised KSC"
-ksc_ncp_rank_stats$group <- "KSC"
+ISW_rank_stats$name <- names(ISW_fivek_bins)
+ISW_rank_stats$type <- "Importance Weighted Ranks"
+ISW_rank_stats$group <- "KSC"
+
+# # KSC NONCENTERED results
+# n_iterations <- 5000
+# posterior_samples <- 9999 # Posterior samples. m+1 possible ranks. So want (m+1)/J = expected samples for uniform dist.
+# n_bins <- 20 # Bins
+# expected_count <- n_iterations / n_bins # Chi sqd stat
+
+# ksc_ncp_fivek_bins <- rank_bins_f(path = "simulation_output/ksc/ncp/sbc_ncp_ksc_model_ncp_dgf_10kmcmc_5000iter_r1", 
+#                                     rank_type = "agg_ranks", 
[…]

### 2023-10-22 manuscript/results/rank_histograms_static_states.r

@@ -64,10 +64,11 @@ hmc_ncp_fivek_bin_long <- static_rank_bins(data = hmc_ncp_fivek_bins,
                                           type = "NonCentered")
 
 # hmc_5k = rbind(hmc_cp_fivek_bin_long, hmc_ncp_fivek_bin_long)
-rank_hist(hmc_cp_fivek_bin_long) # HMC NCP 1k
+rank_hist(hmc_cp_fivek_bin_long)  # HMC NCP 1k
 ggsave("manuscript/results/hmc_cp_5k.png", bg = "white", width = 14, height = 9.42)
 
-rank_hist(hmc_ncp_fivek_bin_long) + ylim(0, 600) # HMC NCP 5k
+rank_hist(hmc_ncp_fivek_bin_long) + 
+    ylim(0, 600) # HMC NCP 5k
 ggsave("manuscript/results/hmc_ncp_5k.png", bg = "white", width = 14, height = 9.42)
 
 # KSC 1k
@@ -95,7 +96,7 @@ ksc_ncp_onek_bins_long <- static_rank_bins(data = ksc_ncp_onek_bins,
                                          variables = parameters, 
                                          type = "Centered")
 
-rank_hist(ksc_cp_onek_bins_long) # KSC CP 1k
+rank_hist(ksc_cp_onek_bins_long)  # KSC CP 1k
 ggsave("manuscript/results/ksc_cp_1k.png", bg = "white", width = 14, height = 9.42)
 
 rank_hist(ksc_ncp_onek_bins_long) # KSC NCP 1k
@@ -125,10 +126,10 @@ ksc_ncp_fivek_bins_long <- static_rank_bins(data = ksc_ncp_fivek_bins,
                                          variables = parameters, 
                                          type = "Centered")
 
-rank_hist(ksc_cp_fivek_bins_long) # KSC CP 5k
+rank_hist(ksc_cp_fivek_bins_long)  # KSC CP 5k
 ggsave("manuscript/results/ksc_cp_5k.png", bg = "white", width = 14, height = 9.42)
 
-rank_hist(ksc_ncp_fivek_bins_long) # KSC NCP 5k
+rank_hist(ksc_ncp_fivek_bins_long)  # KSC NCP 5k
 ggsave("manuscript/results/ksc_ncp_5k.png", bg = "white", width = 14, height = 9.42)
 
 # REWEIGHTED KSC 1k

### 2023-10-22 src/gaussianMixture.py

@@ -50,7 +50,7 @@ class TVLLDT(sm.tsa.statespace.MLEModel):
         if parameterisation == "centered":
             self['obs_intercept', 0] = ksc_params[indicators, 1] - 1.27036
         elif parameterisation == "noncentered":
-            self['obs_intercept', 0] = ksc_params[indicators, 1] - 1.27036 + params[0] * (1 - params[1])
+            self['obs_intercept', 0] = ksc_params[indicators, 1] - 1.27036 + params[0]
         else:
             raise ValueError("Incorrect parameterisation")
 
@@ -62,7 +62,7 @@ class TVLLDT(sm.tsa.statespace.MLEModel):
         if parameterisation == "centered":
             self['state_intercept', 0, 0] = params[0] * (1 - params[1])
         elif parameterisation == "noncentered":
-            self['state_intercept', 0, 0] = 0
+            self['state_intercept', 0, 0] = params[0] * (-params[1])
         else:
             raise ValueError("Incorrect parameterisation")
         

### 2023-10-22 manuscript/motivating_example/single_sim_example.r
Print true params

@@ -2,7 +2,7 @@ library(tidyr)
 library(dplyr)
 library(cmdstanr)
 library(ggplot2)
-set.seed(630)
+# set.seed(630)
 set.seed(634)
 
 n_time_points <- 1000
@@ -11,10 +11,13 @@ gen_ncp_sv_dataset <- function(T){
     sim_results <- list()
     # Priors
     mu <- rnorm(1, 0, sqrt(10))
+    print(mu)
     sigma_sqd <- 1 / rgamma(1, 5/2, (0.01*5)/2)
+    print(sigma_sqd)
     sigma <- sqrt(sigma_sqd)
     p <- rbeta(1, 20, 1.5)
     phi <- 2*p - 1
+    print(phi)
 
     # Sample from standardised normal dist and multiply by sigma
     h_std <- rnorm(T, mean = 0, sd = sigma)
@@ -82,15 +85,15 @@ quantiles <- latex_variables(quantiles)
 means <- latex_variables(means)
 
 ggplot(draws, aes(x = value)) +
-  geom_histogram(aes(fill = name), alpha = 0.3) +
-  theme_minimal(base_size = 20) +
+  geom_histogram(aes(y= after_stat(density), fill = name), alpha = 0.3) +
+  theme_bw(base_size = 20) +
   geom_vline(data = means, aes(xintercept=value, color=name), size = 1.5) +
   geom_vline(data = quantiles, aes(xintercept=value), linetype="dotted", size = 1) +
     facet_wrap(~name, scale = "free", labeller = label_parsed) +
     theme(legend.position="none") +
-    labs(title = "Posterior distribution - simulated dataset",
+    labs(title = "",
          x = "Parameter value", 
-         y= "Count") + 
+         y= "Density") + 
        theme(strip.text.x = element_text(size = 22))
 
 ggsave("manuscript/motivating_example/single_sim.png", bg = "white", width = 14, height = 9.42)

### 2023-10-22 writeups/research_proposal/propsal.tex
presentations

--- 2023-09-16 writeups/research_proposal/propsal.qmd
@@
----
-title: "Comparison of MCMC algorithms in Stochastic Volatility Models"
-author: 
-    - "By Benjamin Wee (32921020)" 
-    - "Supervised by Catherine Forbes and Lauren Kennedy"
-# abstract: "The objective of this research is to validate and compare the computational methods used to estimate Bayesian stochastic volatility models.
-Specifically, Simulation Based Calibration will be used to identify biases and errors arising from the results of Markov Chain Monte Carlo (MCMC) algorithms.
-The source of any miscalibration include, but are not limited to, estimation strategies, choice of sampler, model specifications and coding mistakes.
-This study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
-Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches."
-format:
-  pdf:
-    fig-pos: H
-    include-in-header:
-        text: |
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  12pt,
+  letterpaper,
+  DIV=11,
+  numbers=noendperiod]{scrartcl}
+
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \ifXeTeX
+    \usepackage{mathspec} % this also loads fontspec
+  \else
+    \usepackage{unicode-math} % this also loads fontspec
+  \fi
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+
 \usepackage{lipsum}
@@
 \linespread{1.5}
-fontsize: 12pt
-geometry: margin=1in
-mathspec: true
----
+\KOMAoption{captions}{tableheading}
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Comparison of MCMC algorithms in Stochastic Volatility Models},
+  pdfauthor={By Benjamin Wee (32921020); Supervised by Catherine Forbes and Lauren Kennedy},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
 
-<!-- # Background and Motivation
-Maybe focus on the MCMC as opposed to SV upfront.
+\title{Comparison of MCMC algorithms in Stochastic Volatility Models}
+\author{By Benjamin Wee (32921020) \and Supervised by Catherine Forbes
+and Lauren Kennedy}
+\date{}
 
-## SV
-Why model volatility. Quantifying risks, tail probabilities, worst case scenarios. These are all inputs in the pricing of assets and derivatives. (pg92 FE slides)
+\begin{document}
+\maketitle
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, boxrule=0pt, frame hidden, breakable, enhanced]}{\end{tcolorbox}}\fi
 
-SV models is one way to model volatility.
-Unlike the classic Black Scholes option pricing formula, it assumes that the underlying variance of an asset's return is not constant.
-That is, the variance of the underlying asset return is treated as a latent random variable.
-This deals with some of the empirical short comings of the BS model such as the volatility smile and skew in the realised variance in financial assets.
+\hypertarget{research-goal}{%
+\subsection{1. Research Goal}\label{research-goal}}
 
-SV models are typically expressed as non linear Gaussian state space models.
-This can make it difficult to estimate using classical methods.
-The likelihood is unavailable in closed form and there are at least as many variables as data points.
-There are however, a variety of Frequentist and Bayesian strategies for estimating SV models.
-Frequentist approaches include quasi likelihood estimation and GMM which rely on asymptotic sampling distributions.
-Bayesian approaches rely on computational Markov Chain Monte Carlo techniques to sample from the target joint posterior of these high dimensional state space models.
+The objective of this research is to compare the computational methods
+used to estimate Bayesian stochastic volatility models. A simulation
+study will compare the estimation strategies detailed in the stochastic
+volatility literature with Hamiltonian Monte Carlo and their ability to
+sample from the model's posterior distribution. Specifically, Simulation
+Based Calibration (SBC) is used to check whether these sampling
+strategies are returning efficient and well calibrated posterior
+estimates. Key metrics of interest are the effective sample size to
+check the efficiency of the algorithm and tests of uniformity to assess
+the calibration of the posteriors. This will determine which algorithm
+is better at estimating stochastic volatility models based on the
+efficiency and accuracy of the sampling strategy.
 
-This research will focus on Bayesian estimation strategies for Stochastic Volatility models.
-Particular attention will be given to the computational tools and approaches used to sample from high dimensional posterior distributions.
-Early approaches for estimating these models relied on classic Gibbs and Metropolis Hastings algorithms to apply the MCMC routines.
-Since then, there have been many developments in the availability of new kinds of MCMC samplers.
+\hypertarget{background-and-literature-review}{%
+\section{2. Background and Literature
+Review}\label{background-and-literature-review}}
 
-Hamiltonian Monte Carlo is a .
-It has been popularised through its implementation in a variety of open source programming languages.
-It presents a new way to sample from increasingly complicated generative models .
+\hypertarget{stochastic-volatility}{%
+\subsection{2.1 Stochastic Volatility}\label{stochastic-volatility}}
 
-However, as the developments of these tools enable the estimation of more complicated models (with relative ease and accessibility through the development of open source software), so do the risk of mistakes and the need for procedures diagnose problems or mistakes when implementing these methods.
-How do we know that these algorithms are doing what we expect them to be doing, and how can we effectively differentiate the strengths and weaknesses of different tools.
+Stochastic volatility models are used in financial econometrics to model
+the variance or volatility of returns in a financial instrument. The
+return of a financial asset often exhibits patterns in the variance
+which can be modeled to improve prediction and estimates of risk or tail
+events. Unlike other volatility models such as ARCH or GARCH, stochastic
+volatility models treat the variance as a random variable. Furthermore,
+these models have a state space representation since the variance is
+expressed as a latent stochastic process. State space models can be
+difficult to estimate since there is a latent state for every data
+observation in the time series, which means there are often more
+parameters than data points. These models can be estimated using
+frequent frameworks but this research will focus on Bayesian estimation
+strategies.
 
-How do we validate the computation and implementation of our sampling strategies and algorithms in increasingly complex models?
-And how do these form part of the computational modelling workflow?
--->
+Kim, Shephard and Chib (1998) estimate a discrete time, univariate
+stochastic volatility model using Bayesian methods which is outlined
+below. \(y_t\) is the mean corrected returns of some asset for equally
+spaced intervals t. \(\beta\) is a constant scaling factor which is also
+defined as \(exp(\mu / 2)\) representing instantaneous volatility.
+\(h_t\) is log volatility, where \(h_1\) is a draw from a stationary
+distribution and the state equation \(h_{t+1}\) follows a stationary
+process governed by the autoregressive parameter \(\phi\) such that
+\(|\phi|<1\). This autoregressive parameter represents the persistence
+or ``stickiness'' of log volatility and the dispersion parameter
+\(\sigma_{\eta}\) is the constant variance of the states. \(\epsilon_t\)
+and \(\eta_t\) are standard normal white noise shocks and are
+uncorrelated with each other.
 
-<!-- ## HMC
-- There have been many developemnts over the past few decades on estimating Bayesian models through Markov Chain Monte Carlo methods.
-
-The original BUGs software implemented Gibbs and Metropolis Hastings MCMC algorithms for estimating this model.
-
-In particular, the implementation of Hamiltonian Monte Carlo, a variant of the Metropolis Hastings algorithm, represents a recent advancement MCMC.
-
-- The development of HMC is another advancement in estimating complicated, high dimensional models. 
-- Development of diagnostics for different types of HMC and tools for evaluating the convergence and efficiency of the MCMC sampler.
-
-- SV models are a primary candidate for using MCMC methods? Difficult to
-
-## SBC -->
-
-<!-- and provide tools for evaluating other types models and algorithms.  -->
-
-<!-- The objective of this research is to evaluate the performance of MCMC algorithms for estimating SV models through a simulation study.
-Simulations are useful since there is control over the data generating process.
-This will help us understand the behaviour of MCMC and evaluate its performance against known parameters.
-Particular attention will be given to the calibration of the MCMC algorithms in the stochastic volatility context.
-Further diagnostics to evaluate performance will be explored to determine the efficiency of the MCMC sampler.
-
-The first step is to understand what "performance" of MCMC algorithms means and what exactly we are measuring.
-At a high level, the objective of MCMC is to sample from the target posterior distribution and to provide reliable inference.
-Convergence to the target posterior can be measured through efficiency measures such as rhat and effective sample size.
-
-- Explcitly outline what is compared. KSC and HMC algorithms and estimation strategies.  -->
-
-<!-- 
-Frequentist methods
-
-- Variety of estimation strategies
-
-- There are a variety of ways of estimating this model using a Bayesian frameworks.
-
-Over the past few decades there have been many developments in computing hardware -->
-<!-- 
-RESEARCH GOAL: To compare the performance of different MCMC algorithms and estimation strategies when fitting stochastic volatility models. The study will use a discrete formulation is described  -->
-
-<!-- The use of simulation and sampling are used as part of model diagnostics and evaluation. -->
-
-## 1. Research Goal
-
-The objective of this research is to compare the computational methods used to estimate Bayesian stochastic volatility models.
-A simulation study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo and their ability to sample from the model's posterior distribution.
-Specifically, Simulation Based Calibration (SBC) is used to check whether these sampling strategies are returning efficient and well calibrated posterior estimates.
-Key metrics of interest are the effective sample size to check the efficiency of the algorithm and tests of uniformity to assess the calibration of the posteriors.
-This will determine which algorithm is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
-
-<!-- Stochastic volatility models are used to explicitly model the variance of returns in financial instruments and can be estimated using frequentist or Bayesian frameworks.  -->
-
-<!-- 
-The objective of this research is to validate the computational methods used to estimate Bayesian stochastic volatility models.
-This study will assess the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
-A simulation study will compare the ability of these methods to sample from the model's posterior distribution.
-Specifically, Simulation Based Calibration (SBC) is used to check whether these sampling strategies are returning efficient and well calibrated posterior estimates.
-Key metrics to check these features are the effective sample size to check efficiency of the algorithm and tests of uniformity to determine calibration of the posteriors.
-This will determine which algorithm is better at estimating stochastic volatility models based on the efficiency and accuracy of the sampling strategy.
--->
-
-<!-- Simulation Based Calibration will be used to identify biases and errors arising from the results of Markov Chain Monte Carlo (MCMC) algorithms.
-The source of any miscalibration include, but are not limited to, estimation strategies, choice of sampler, model specifications and coding mistakes.
-This study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
-Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches.
--->
-
-<!-- The objective of this research is to validate and compare the computational methods used to estimate Bayesian stochastic volatility models.
-Specifically, Simulation Based Calibration will be used to identify biases and errors arising from the results of Markov Chain Monte Carlo (MCMC) algorithms.
-The source of any miscalibration include, but are not limited to, estimation strategies, choice of sampler, model specifications and coding mistakes.
-This study will compare the estimation strategies detailed in the stochastic volatility literature with Hamiltonian Monte Carlo for estimating these class of models.
-Results and diagnostics from these simulations will help identify strengths and limitations that arise from these approaches.
--->
-
-<!-- Specifically, they propose a method of sampling by using an offset mixture normal approximation with a reweighting procedure to sample from the posterior distribution.
-By linearing the model and using a mixture of Gaussians, they were able to apply a Kalman Filter to estimate the posterior density over the latent volatilities and static parameters.
--->
-
-<!-- The univariate stochastic volatility model described in their paper outlines the state space representation of the the model which is described below.
- -->
-# 2. Background and Literature Review
-
-## 2.1 Stochastic Volatility
-Stochastic volatility models are used in financial econometrics to model the variance or volatility of returns in a financial instrument.
-The return of a financial asset often exhibits patterns in the variance which can be modeled to improve prediction and estimates of risk or tail events.
-Unlike other volatility models such as ARCH or GARCH, stochastic volatility models treat the variance as a random variable.
-Furthermore, these models have a state space representation since the variance is expressed as a latent stochastic process.
-State space models can be difficult to estimate since there is a latent state for every data observation in the time series, which means there are often more parameters than data points.
-These models can be estimated using frequent frameworks but this research will focus on Bayesian estimation strategies.
-
-Kim, Shephard and Chib (1998) estimate a discrete time, univariate stochastic volatility model using Bayesian methods which is outlined below.
-$y_t$ is the mean corrected returns of some asset for equally spaced intervals t.
-$\beta$ is a constant scaling factor which is also defined as $exp(\mu / 2)$ representing instantaneous volatility.
-$h_t$ is log volatility, where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the autoregressive parameter $\phi$ such that $|\phi|<1$.
-This autoregressive parameter represents the persistence or "stickiness" of log volatility and the dispersion parameter $\sigma_{\eta}$ is the constant variance of the states.
-$\epsilon_t$ and $\eta_t$ are standard normal white noise shocks and are uncorrelated with each other.
-
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-Setting $\beta=1$, the model can be expressed more succinctly as:
+Setting \(\beta=1\), the model can be expressed more succinctly as:
 
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-Priors for the static parameters are defined below with conjugate priors on $\mu$ and $\sigma^2$:
+Priors for the static parameters are defined below with conjugate priors
+on \(\mu\) and \(\sigma^2\):
 
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-The prior on $\phi$ is a "stretched" beta distribution. This is a beta distribution (as defined on the parameter $\phi^*$) which has been transformed to have support (-1, 1).
+The prior on \(\phi\) is a ``stretched'' beta distribution. This is a
+beta distribution (as defined on the parameter \(\phi^*\)) which has
+been transformed to have support (-1, 1).
 
-### Estimation strategy
-Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic volatility model using a mix of conjugate posterior distributions, Metropolis Hastings and the Kalman Filter and smoother[^1] (de Jong and Shephard, 1995).
+\hypertarget{estimation-strategy}{%
+\subsubsection{Estimation strategy}\label{estimation-strategy}}
 
-[^1]: In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.
+Kim, Shephard and Chib (KSC) sample the posteriors of the stochastic
+volatility model using a mix of conjugate posterior distributions,
+Metropolis Hastings and the Kalman Filter and smoother\footnote{In this
+  research the exact software to apply the simulation smoother is
+  unavailable, so a more recent simulation smoother is used which is
+  based on the software written by the same author.} (de Jong and
+Shephard, 1995).
 
-The standard Kalman Filter and smoother is used to compute the posterior distribution over the latent states.
-This requires the state and measurement equations to be linear and conditionally Gaussian.
-Since the relationship between $y_t$ and $h_t$ in the measurement equation is not linear, a transformation is applied by squaring and taking the log of $y_t$.
+The standard Kalman Filter and smoother is used to compute the posterior
+distribution over the latent states. This requires the state and
+measurement equations to be linear and conditionally Gaussian. Since the
+relationship between \(y_t\) and \(h_t\) in the measurement equation is
+not linear, a transformation is applied by squaring and taking the log
+of \(y_t\).
 
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
-The relationship between $y_t$ and $h_t$ is now linear; however, the error is not Gaussian.
-Since it is not simple to sample from this parameterisation of the model, KSC use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution.
-This is defined by:
+Where \(z_t = log(\epsilon_t^2)\) follows a log chi-squared distribution
+with mean -1.2704 and variance 4.93. The relationship between \(y_t\)
+and \(h_t\) is now linear; however, the error is not Gaussian. Since it
+is not simple to sample from this parameterisation of the model, KSC use
+a mixture of Gaussians to \textbf{approximate} the first 4 moments of
+the log chi squared distribution. This is defined by:
 
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-Where K is the mixture of normal densities $f_N$, component probabilities $q_i$, mean $m_i-1.2704$ and variance $\nu_i^2$.
-These parameters were selected using moment matching where they found 7 normal densities with varying mean and variance parameters best approximated the log chi squared moments.
-These parameters and weights can be found in Appendix A.
+Where K is the mixture of normal densities \(f_N\), component
+probabilities \(q_i\), mean \(m_i-1.2704\) and variance \(\nu_i^2\).
+These parameters were selected using moment matching where they found 7
+normal densities with varying mean and variance parameters best
+approximated the log chi squared moments. These parameters and weights
+can be found in Appendix A.
 
-The model can be sampled via the Kalman Filter and simulation smoother since the model is now linear and conditionally gaussian.
-The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions whereas $\phi$ is sampled via a Metropolis Hastings accept/reject procedure.
-The details can be around in Appendix B.
+The model can be sampled via the Kalman Filter and simulation smoother
+since the model is now linear and conditionally gaussian. The static
+parameters \(\mu\) and \(\sigma^2\) are sampled directly from their
+conjugate posterior distributions whereas \(\phi\) is sampled via a
+Metropolis Hastings accept/reject procedure. The details can be around
+in Appendix B.
 
-## 2.2 Hamiltoninan Monte Carlo
-Since the paper was written, new MCMC algorithms have been developed and enabled the estimation of richer and more complicated models.
-Specifically, Hamiltonian Monte Carlo is a MCMC algorithm which has become widely available for efficiently sampling from sophisticated models.
-Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo, was developed in the physics literature (Duane et al.
-1987) before being applied in the statistics literature by Radford Neal through his works in Bayesian Neural Networks (Neal, 1995) and statistical computing (Neal 2011).
-The algorithm has since become widely available through open source development projects such as Stan (Stan Development Team, 2023) and PyMC (2023).
+\hypertarget{hamiltoninan-monte-carlo}{%
+\subsection{2.2 Hamiltoninan Monte
+Carlo}\label{hamiltoninan-monte-carlo}}
 
-The key innovation of Hamiltonian Monte Carlo is using the gradients of the target posterior distribution to generate an efficient path for the sampler to explore.
-Unlike Random Walk Metropolis Hastings, it takes advantage of the geometry of the posterior to determine its next proposal step.
-A comprehensive explanation of the sampler is beyond the scope of this research and can be found in the above references.
+Since the paper was written, new MCMC algorithms have been developed and
+enabled the estimation of richer and more complicated models.
+Specifically, Hamiltonian Monte Carlo is a MCMC algorithm which has
+become widely available for efficiently sampling from sophisticated
+models. Hamiltonian Monte Carlo, originally called Hybrid Monte Carlo,
+was developed in the physics literature (Duane et al.~1987) before being
+applied in the statistics literature by Radford Neal through his works
+in Bayesian Neural Networks (Neal, 1995) and statistical computing (Neal
+2011). The algorithm has since become widely available through open
+source development projects such as Stan (Stan Development Team, 2023)
+and PyMC (2023).
 
-The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
-Stan's default algorithm, the No-U-Turn Sampler (Hoffman & Gelman 2014), allows for direct sampling of the specified stochastic volatility model.
-Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
-This approach will also use the same priors as specified in the Gaussian mixture approximation.
+The key innovation of Hamiltonian Monte Carlo is using the gradients of
+the target posterior distribution to generate an efficient path for the
+sampler to explore. Unlike Random Walk Metropolis Hastings, it takes
+advantage of the geometry of the posterior to determine its next
+proposal step. A comprehensive explanation of the sampler is beyond the
+scope of this research and can be found in the above references.
 
-## 2.3 Research contribution
-The goal of this research is to determine which algorithm is better at estimating the discrete time stochastic volatility model (conditional on the same priors).
-My main contribution is to design a simulation study that compares these algorithms and determines which sampling strategy is better based on the calibration of posterior estimates and the efficiency of the MCMC sampler.
+The Stan programming language's implementation of Hamiltonian Monte
+Carlo will be used for this study. Stan's default algorithm, the
+No-U-Turn Sampler (Hoffman \& Gelman 2014), allows for direct sampling
+of the specified stochastic volatility model. Hamiltonian Monte Carlo
+allows for sampling of the generative model and can flexibly handle
+complicated likelihood functions. This approach will also use the same
+priors as specified in the Gaussian mixture approximation.
 
-<!-- New developments in MCMC algorithms and improved computing resources have enabled estimation of richer and more complicated models.
-However, as the development of tools enables us to fit more complicated models, so does the need to validate the output of these tools.
-Cook, Gelman and Rubin (2006) introduce simulation techniques to evaluate the "correctness" of the software used to sample Bayesian models.
-Their contribution identified that cumulative density quantiles from the posterior samples can be tested for uniformity.
-Talts et al (2018).
-build upon this work by developing Simulation Based Calibration to provide more robust validation of output from a MCMC sampler through the use of rank statistics.
--->
+\hypertarget{research-contribution}{%
+\subsection{2.3 Research contribution}\label{research-contribution}}
 
- <!-- through the use of the CDF quantiles approximated from the posterior distribution. -->
+The goal of this research is to determine which algorithm is better at
+estimating the discrete time stochastic volatility model (conditional on
+the same priors). My main contribution is to design a simulation study
+that compares these algorithms and determines which sampling strategy is
+better based on the calibration of posterior estimates and the
+efficiency of the MCMC sampler.
 
-<!-- , through open source programming languages such as Stan and PyMC, have become
+\hypertarget{methodology}{%
+\section{3. Methodology}\label{methodology}}
 
-As new and complicated
+\hypertarget{design}{%
+\subsection{3.1 Design}\label{design}}
 
-There have been many developments in MCMC algorithms since the introduction of classical methods such as Gibbs and Metropolis Hastings.
+Simulation Based Calibration (SBC) is a technique that checks the
+calibration of posterior estimates generated by MCMC algorithms. SBC is
+conducted by comparing the distribution of rank statistics to the
+discrete uniform distribution which arises when an algorithm is
+correctly calibrated. The procedure starts by taking draws from the
+prior distribution and creating datasets implied by each draw. Rank
+statistics are then calculated on the posterior samples conditional on
+the simulated data.
 
-Hamiltoninn Monte Carlo (algorithm introduced when, widely available since 2010...?) is a relatively new im.
-
-- Since then, there have been many developments in MCMC algorithms. In particular, HMC provides a flexible alternative to fitting and sampling from complicated high dimensional Bayesian models.
-- It's key innovation is the use of gradients - enabling the flexible specification of any likelihood function. 
-- Constraints- can only sample from continuous 
-- HMC also provides new diagnostics for evaluating the effectiveness of the sampling procedure and how well the sampler is targeting the joint dist.
-
-- How do we compare models? SBC. ESS in the tails, rhats.
-
-- Given this context, my research looks to compare the efficacy of KSCs proposed MCMC sampling strategy with Hamiltonina Monte Carlo (NUTs).  -->
-
-# 3. Methodology
-
-## 3.1 Design
-Simulation Based Calibration (SBC) is a technique that checks the calibration of posterior estimates generated by MCMC algorithms.
-SBC is conducted by comparing the distribution of rank statistics to the discrete uniform distribution which arises when an algorithm is correctly calibrated.
-The procedure starts by taking draws from the prior distribution and creating datasets implied by each draw.
-Rank statistics are then calculated on the posterior samples conditional on the simulated data.
-
-To illustrate this procedure, let $\theta$ be a parameter and $y$ represent the dataset. Start with a single draw from the prior distribution:
-$$
+To illustrate this procedure, let \(\theta\) be a parameter and \(y\)
+represent the dataset. Start with a single draw from the prior
+distribution: \[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
@@
 
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-Then take draws from the posterior distribution generated by a MCMC algorithm or estimation strategy (Hamiltonian Monte Carlo or KSC) conditional on this dataset.
+Then take draws from the posterior distribution generated by a MCMC
+algorithm or estimation strategy (Hamiltonian Monte Carlo or KSC)
+conditional on this dataset.
 
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-A key result is that the posterior sample $\{\theta_1,\dots , \theta_{L}\}$ will share the same distribution as the prior samples $\theta^{sim}$. This is implied by the following expression:
+A key result is that the posterior sample
+\(\{\theta_1,\dots , \theta_{L}\}\) will share the same distribution as
+the prior samples \(\theta^{sim}\). This is implied by the following
+expression:
 
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-That is, the posterior averaged over the joint distribution follows the same distribution as the prior.
-The procedure of generating posterior samples implicitly performs this integral since the expression on the right of the integral is proportional to the prior density.
-Therefore, any deviation of the posterior samples from the prior distribution suggests that the sampling methodology is not producing the correct posteriors.
+That is, the posterior averaged over the joint distribution follows the
+same distribution as the prior. The procedure of generating posterior
+samples implicitly performs this integral since the expression on the
+right of the integral is proportional to the prior density. Therefore,
+any deviation of the posterior samples from the prior distribution
+suggests that the sampling methodology is not producing the correct
+posteriors.
 
-<!-- n error - either in computation, code or overall analysis.  -->
+If the posterior samples follows the prior distribution, the rank
+statistic for a given parameter follows a discrete uniform
+distribution\footnote{Proof of this result in Talts et al (2018).}. The
+rank statistic is defined as:
 
-<!-- The simulation if done correctly should produce samples from the prior distribution.  -->
-
-If the posterior samples follows the prior distribution, the rank statistic for a given parameter follows a discrete uniform distribution[^2]. The rank statistic is defined as:
-
-[^2]: Proof of this result in Talts et al (2018).
-
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-This completes one iteration of SBC.
-To complete the algorithm, multiple iterations are run and the rank statistics are calculated for each parameter.
-The resulting rank statistics are compared to the discrete uniform distribution to determine if any problematic features exist.
+This completes one iteration of SBC. To complete the algorithm, multiple
+iterations are run and the rank statistics are calculated for each
+parameter. The resulting rank statistics are compared to the discrete
+uniform distribution to determine if any problematic features exist.
 
-If the computation is well calibrated and the rank statistics follow a discrete uniform distribution, then the posterior credible intervals have sufficient coverage.
-That is, one way to describe calibration is: for any percentage interval selected over the posterior samples (for example 90%) then there is a 90\% chance that $\theta^{sim}$ falls within this interval.
-Another way of saying this is a Bayesian analysis is well calibrated if a 90\% credible interval contains the true parameter in 90\% of the SBC iterations.
+If the computation is well calibrated and the rank statistics follow a
+discrete uniform distribution, then the posterior credible intervals
+have sufficient coverage. That is, one way to describe calibration is:
+for any percentage interval selected over the posterior samples (for
+example 90\%) then there is a 90\% chance that \(\theta^{sim}\) falls
+within this interval. Another way of saying this is a Bayesian analysis
+is well calibrated if a 90\% credible interval contains the true
+parameter in 90\% of the SBC iterations.
 
-## 3.2 Implementation
-SBC for Hamiltonian Monte Carlo and the Gaussian approximation are performed with 1000 simulation iterations, 1000 observations in each dataset, and post warmup/burnin samples of 999 (which gives 1000 possible rank statistics).
-This algorithm is summarised below:
+\hypertarget{implementation}{%
+\subsection{3.2 Implementation}\label{implementation}}
 
-| **1) for sim in 1000 iterations:**
-|   2) Draw from prior: $\theta^{sim}\sim\pi (\theta)$
-|   3) Simulate dataset with 1000 observations: $y^{sim} \sim p(y|\theta^{sim})$
-|   4) Draw 999 posterior samples (post warmup) $\{\theta_1,\dots , \theta_{L}\} \sim p(\theta | y^{sim})$
-|   5) Compute rank statistics $r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim})$
+SBC for Hamiltonian Monte Carlo and the Gaussian approximation are
+performed with 1000 simulation iterations, 1000 observations in each
+dataset, and post warmup/burnin samples of 999 (which gives 1000
+possible rank statistics). This algorithm is summarised below:
 
-## 3.3 Metrics
-The key metrics and diagnostics to compare the performance of these methods are the effective sample size (ESS), rank statistics, and chi-squared test statistics.
+\textbf{1) for sim in 1000 iterations:}\\
+\hspace*{0.333em}\hspace*{0.333em}2) Draw from prior:
+\(\theta^{sim}\sim\pi (\theta)\)\\
+\hspace*{0.333em}\hspace*{0.333em}3) Simulate dataset with 1000
+observations: \(y^{sim} \sim p(y|\theta^{sim})\)\\
+\hspace*{0.333em}\hspace*{0.333em}4) Draw 999 posterior samples (post
+warmup) \(\{\theta_1,\dots , \theta_{L}\} \sim p(\theta | y^{sim})\)\\
+\hspace*{0.333em}\hspace*{0.333em}5) Compute rank statistics
+\(r = rank(\{\theta_1,\dots , \theta_{L}\}, \theta^{sim})\)
 
-ESS measures the efficiency of the MCMC sampler.
-It calculates the number of (effectively) independent draws from the posterior draws generated by a Markov chain.
-A poor ESS can arise from high autocorrelation in the Markov chain which leads to highly dependent samples.
-An efficient MCMC algorithm takes less resources (for example, time and number of draws) to get a representative sample of the target distribution.
-If a MCMC algorithm possesses higher ESS for the majority of its parameters (relative to another strategy), then we may conclude that this method is a more efficient sampler (conditional on the model).
+\hypertarget{metrics}{%
+\subsection{3.3 Metrics}\label{metrics}}
 
-Rank statistics as described in the SBC section are used to evaluate the calibration of a posterior.
-Histograms will be used to evaluate the distribution of rank statistics.
-If a posterior is well calibrated then it is expected that the histogram is uniform.
+The key metrics and diagnostics to compare the performance of these
+methods are the effective sample size (ESS), rank statistics, and
+chi-squared test statistics.
 
-A drawback of this approach is there are more parameters than data points in this model.
-An alternative to visually checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each histogram bin.
-Let $b_j$ be the number of counts and $e_j$ the expected count in bin $j$.
-Then the chi squared statistic is given by:
+ESS measures the efficiency of the MCMC sampler. It calculates the
+number of (effectively) independent draws from the posterior draws
+generated by a Markov chain. A poor ESS can arise from high
+autocorrelation in the Markov chain which leads to highly dependent
+samples. An efficient MCMC algorithm takes less resources (for example,
+time and number of draws) to get a representative sample of the target
+distribution. If a MCMC algorithm possesses higher ESS for the majority
+of its parameters (relative to another strategy), then we may conclude
+that this method is a more efficient sampler (conditional on the model).
 
-$$
+Rank statistics as described in the SBC section are used to evaluate the
+calibration of a posterior. Histograms will be used to evaluate the
+distribution of rank statistics. If a posterior is well calibrated then
+it is expected that the histogram is uniform.
+
+A drawback of this approach is there are more parameters than data
+points in this model. An alternative to visually checking for uniformity
+of all the parameters is to calculate the chi squared statistics for the
+counts in each histogram bin. Let \(b_j\) be the number of counts and
+\(e_j\) the expected count in bin \(j\). Then the chi squared statistic
+is given by:
+
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-Chi squared statistics can be used to test for uniformity and thus calibration of all the parameters in the model.
+Chi squared statistics can be used to test for uniformity and thus
+calibration of all the parameters in the model.
 
-<!-- 
-The aim of this research is to use Simulation Based Calibration (SBC) to validate the inference of posterior samples generated by different MCMC algorithms in the context of stochastic volatility models.
-This will help identify any bias in computation and inconsistencies in model implementation and estimation strategies.
-Comparison of the SBC results between different computational approaches may help identify any strengths or shortcomings as well as any errors in model specification.
--->
+\hypertarget{preliminary-results}{%
+\section{4. Preliminary Results}\label{preliminary-results}}
 
-<!-- ## 3.1 Stochastic Volatility Model
-The univariate stochastic volatility model described by Kim, Shephard and Chib (1998) models the variance as a random variable following some latent stochastic process.
-$y_t$ is the mean corrected returns of some asset for equally spaced intervals t.
-$\beta$ is a constant scaling factor which is also defined as $exp(\mu / 2)$ representing instantaneous volatility.
-$h_t$ is log volatility, where $h_1$ is a draw from a stationary distribution and the state equation $h_{t+1}$ follows a stationary process governed by the autoregressive parameter $\phi$ such that $|\phi|<1$.
-This autoregressive parameter represents the persistence or "stickiness" of log volatility and the dispersion parameter $\sigma_{\eta}$ is its variance.
-$\epsilon_t$ and $\eta_t$ are standard normal white noise shocks.
+Figure 1 shows the distribution of rank statistics for Hamiltonian Monte
+Carlo, with the horizontal black line at 50 representing the discrete
+uniform distribution. The preliminary figures focus on the static
+parameters and arbitrarily chosen states. Phi and sigma\_sqd look
+relatively uniform. However mu and h.995 (the 995th state parameter)
+have some lumpiness which may suggest a lack of calibration.
 
-$$
-\begin{aligned}
-y_t =& \space \beta exp(h_t/2) \epsilon_t \\
-h_{t+1} =& \space \mu +\phi(h_t - \mu) + \sigma_{\eta} \eta_t  \\
-h_1 \sim& \space normal\left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
-\end{aligned}
-$$
+\begin{figure}[H]
 
-$$
-\begin{aligned}
-\epsilon_t \sim& \space normal(0,1) \\
-\delta_t \sim& \space normal(0,1)
-\end{aligned}
-$$
+{\centering \includegraphics[width=0.75\textwidth,height=\textheight]{../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/static_state_hist.png}
 
-Setting $\beta=1$, the model can be expressed more succinctly as:
+}
 
-$$
-\begin{aligned}
-y_t \sim& \space normal(0, exp(h_t/2)) \\ 
-h_1 \sim& \space normal \left(\mu, \frac{\sigma_{\eta}^2}{1-\phi^2}\right) \\
-h_{t+1} \sim& \space normal(\mu +\phi(h_t - \mu) , \sigma_{\eta}^2), \space\space t\neq 1\\ 
-\end{aligned}
-$$
+\caption{Distribution of parameter ranks for HMC algorithm (1000
+iterations)}
 
-Priors for the static parameters are defined below with conjugate priors on $\mu$ and $\sigma^2$:
+\end{figure}
 
-$$
-\begin{aligned}
-\mu \sim& \space normal(0, 10^2) \\
-\sigma_{\eta}^2 \sim& \space IG(5/2, (0.01\times 5) / 2) \\
-\phi^{\ast} \sim& \space beta(20, 1.5) \\
-\phi &=  2\phi^{\ast} - 1
-\end{aligned}
-$$
+A drawback of SBC is that any deviation in uniformity from a calibrated
+analysis may be due to noisy estimates from small samples. Figure 2
+shows more consistent uniform behaviour from all parameters after
+increasing the number of iterations from 1000 to 5000. This suggests
+that Hamiltonian Monte Carlo and the overall analysis is well
+calibrated.
 
-The prior on $\phi$ is a "stretched" beta distribution. This is a beta distribution (as defined on the parameter $\phi^*$) which has been transformed to have support (-1, 1).
+\begin{figure}[H]
 
-## 3.2 Estimation Strategy
+{\centering \includegraphics[width=0.75\textwidth,height=\textheight]{../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r6_5000_iterations/static_state_hist.png}
 
-### Gaussian Mixture Approximation
-Kim, Shephard and Chib estimate the stochastic volatility model using a Kalman Filter and smoother[^1] (de Jong and Shephard, 1995) to compute the posterior distribution over the latent states.
-This requires the state space model to be linear and conditionally Gaussian.
+}
 
-[^1]: In this research the exact software to apply the simulation smoother is unavailable, so a more recent simulation smoother is used which is based on the software written by the same author.
+\caption{Distribution of parameter ranks for HMC algorithm (5000
+iterations)}
 
-Since the model is not linear, a transformation is applied by squaring and taking the log of $y_t$.
+\end{figure}
 
-$$
-\begin{aligned}
-y_t^{*} &= log(y_t^2) \\ 
-&= log((\epsilon_t exp(h_t/2))^2) \\
-&=  log(exp(h_t)) + log(\epsilon_t^2) \\
-&= h_t + log(\epsilon_t^2)  \\
-&= h_t + z_t \\
-\end{aligned}
-$$
+Figure 3 shows results for the Gaussian approximation. The distribution
+of the rank statistics deviates heavily from the expected discrete
+uniform distribution, with various peaks at each end of the histogram as
+well as peaks in the middle. A reason for this miscalibration may be due
+to the Gaussian mixture only approximating the log chi squared
+distribution. Hamiltonian Monte Carlo on the other hand, samples
+directly from the target model which more closely represents the data
+generating process.
 
-Where $z_t = log(\epsilon_t^2)$ follows a log chi-squared distribution with mean -1.2704 and variance 4.93.
-However, it is not simple to sample from this parameterisation of the model.
-Kim, Sherpherd and Chib use a mixture of Gaussians to **approximate** the first 4 moments of the log chi squared distribution.
-This is defined by:
+This result may also be due to the relatively small number of posterior
+samples drawn from the model. The MCMC sampler may not have converged
+yet which may explain the deviations from uniformity. A follow up step
+is to compare the ESS of the two methods and to increase the length of
+the MCMC chain to ensure the ESS is approximately the same before
+comparing histograms.
 
-$$
-\begin{aligned}
-f(z_t) = \sum_{i=1}^{K} q_if_N(z_i|m_i-1.2704, \nu_i^2)
-\end{aligned}
-$$
+\begin{figure}[H]
 
-Where K is the mixture of normal densities $f_N$, component probabilities $q_i$, mean $m_i-1.2704$ and variance $\nu_i^2$.
-These parameters were selected using moment matching where they found 7 normal densities with varying mean and variance parameters best approximated the log chi squared moments.
-These parameters and weights can be found in Appendix A.
+{\centering \includegraphics[width=0.75\textwidth,height=\textheight]{../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png}
 
-The model can be sampled via the simulation smoother since the model is now linear and conditionally gaussian.
-The static parameters $\mu$ and $\sigma^2$ are sampled directly from their conjugate posterior distributions where as $\phi$ is sampled via a metropolis hastings accept/reject procedure.
-The details can be be around in Appendix B.
+}
 
-### Hamiltonian Monte Carlo (No-U-Turn Sampler)
-The Stan programming language's implementation of Hamiltonian Monte Carlo will be used for this study.
-Stan's default algorithm, the No-U-Turn Sampler (Hoffman & Gelman 2014), allows for direct sampling of the specified stochastic volatility model.
-Hamiltonian Monte Carlo allows for sampling of the generative model and can flexibly handle complicated likelihood functions.
-This approach will use the same priors as specified in the Gaussian mixture approximation.
-However, Hamiltonian Monte Carlo does not sample directly from the analytical form of conjugate or conditionally conjugate distribution, so different priors can be chosen if necessary.
- -->
+\caption{Distribution of parameter ranks for Gaussian mixture model
+(1000 iterations)}
 
-<!-- SBC validates posterior samples and inferences generated by MCMC algorithms. It identifies inconsistencies and bias in computation and estimation of a generative model. This procedure  -->
+\end{figure}
 
-<!-- ## Performance diagnostics
+\hypertarget{conclusion-and-next-steps}{%
+\section{5. Conclusion and next steps}\label{conclusion-and-next-steps}}
 
-The goal of this simulation is to determine the "calibration" of MCMC when estimating SV models.
-
-ADEMP:
-- Aims
-- Data generating mechanism
-    + Write out mathematical form of the model here. Discrete time. 
-- Estimand /target of analysis (maybe after explaining SBC?)
-- Methods
-    + SBC. Evaluating "calibration" of MCMC. What does calibration mean?
-- Performance measures
-    + Rhats, ESS, histogram
-
-- Convergence diagnostic onto target joint posterior distribution? rhats/ess. This assess "efficiency" of MCMC.
-- Tail ESS calculates ESS in the tail quantiles. Such  -->
-
-# 4. Preliminary Results
-Figure 1 shows the distribution of rank statistics for Hamiltonian Monte Carlo, with the horizontal black line at 50 representing the discrete uniform distribution.
-The preliminary figures focus on the static parameters and arbitrarily chosen states.
-Phi and sigma_sqd look relatively uniform.
-However mu and h.995 (the 995th state parameter) have some lumpiness which may suggest a lack of calibration.
-
-![Distribution of parameter ranks for HMC algorithm (1000 iterations)](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/static_state_hist.png){width=75%}
-
-A drawback of SBC is that any deviation in uniformity from a calibrated analysis may be due to noisy estimates from small samples.
-Figure 2 shows more consistent uniform behaviour from all parameters after increasing the number of iterations from 1000 to 5000.
-This suggests that Hamiltonian Monte Carlo and the overall analysis is well calibrated.
-
-![Distribution of parameter ranks for HMC algorithm (5000 iterations)](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r6_5000_iterations/static_state_hist.png){width=75%}
-
-Figure 3 shows results for the Gaussian approximation.
-The distribution of the rank statistics deviates heavily from the expected discrete uniform distribution, with various peaks at each end of the histogram as well as peaks in the middle.
-A reason for this miscalibration may be due to the Gaussian mixture only approximating the log chi squared distribution.
-Hamiltonian Monte Carlo on the other hand, samples directly from the target model which more closely represents the data generating process.
-
-This result may also be due to the relatively small number of posterior samples drawn from the model.
-The MCMC sampler may not have converged yet which may explain the deviations from uniformity.
-A follow up step is to compare the ESS of the two methods and to increase the length of the MCMC chain to ensure the ESS is approximately the same before comparing histograms.
-
-![Distribution of parameter ranks for Gaussian mixture model (1000 iterations)](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/static_state_hist.png){width=75%}
-
-<!-- An alternative to checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each bin.
-Let $b_j$ be the number of counts and $e_j$ the expected count in bin $j$.
-Then the chi squared statistic is given by:
-
-$$
-\begin{aligned}
-\chi^2 = \sum_{j=1}^J \frac{(b_{j} - e_{j})^2}{e_j}
-\end{aligned}
-$$
-
-These chi squared statistics can be considered draws from a chi squared distribution with 19 degrees of freedom (since there are 20 bins).
-Visualising this should give a chi squared distribution (if performing a hypothesis test, this would be the sampling distribution under a null of uniformity).
-One caveat about this is the chi squared statistics calculated in this way may not be independent.
-That is, each iteration of SBC may be independent however, the calculation of the rank statistics for a given iteration may be dependent between some parameters (since each state parameter is a function of the previous state parameter).
-
-Figure 4 and 5 are the chi squared statistic histograms results of Hamiltonian Monte Carlo and Gaussian approximation respectively.
-The vertical black line marks the 95th quantile which represents the critical value for hypothesis testing.
-The Hamiltonian Monte Carlo chi squared statistics look like they are sampled from the correct distribution, with 6.2\% of the samples greater than the 95\% critical value (doing this calculation for 5000 iterations leads to 4.8\% of samples in the critical area).
-Figure 5 shows the majority of chi squared statistics for Gaussian approximation within the critical area, suggesting the estimation strategy is not well calibrated.
-
-![Chi squared statistics for Hamiltonian Monte Carlo](../../simulation_output/sbc_ncp_ksc_priors_0.999_adapt_delta_premade_datasets_r7_1000_iterations/chi_sq_hist.png){width=70%}
-
-![Chi squared statistics for Gaussian approximation](../../simulation_output/sbc_cp_ksc_model_cp_dgf_r1/chi_sq_hist.png){width=70%} -->
-
-# 5. Conclusion and next steps
 There are two key extensions to the simulation results worth exploring.
 
-**1) Metropolis Hastings reweighting for KSC model**
+\textbf{1) Metropolis Hastings reweighting for KSC model}
 
-Kim, Shephard and Chib apply importance sampling weights to correct for the approximation error of the Gaussian mixture model.
-According to the authors, the correction ensures that samples are from the exact posterior density.
-The next step is to perform this correction and evaluate the SBC results.
-I'd expect the Gaussian mixture approximation diagnostics to improve after this correction.
+Kim, Shephard and Chib apply importance sampling weights to correct for
+the approximation error of the Gaussian mixture model. According to the
+authors, the correction ensures that samples are from the exact
+posterior density. The next step is to perform this correction and
+evaluate the SBC results. I'd expect the Gaussian mixture approximation
+diagnostics to improve after this correction.
 
-**2) Centered and non centered parameterisation**
+\textbf{2) Centered and non centered parameterisation}
 
 The parameterisation of a model can affect the performance of a sampler.
-This can cause significant problems in getting unbiased MCMC samples for complex models.
-An example of this is Neal's funnel (Neal 2003) which arises in hierarchical models which causes performance issues in the Hamiltonian Monte Carlo sampler.
-Rewriting the model with non centered parameterisation significantly improves the performance of MCMC.
-Another consideration for this research is to consider both centered and non centered parameterisations of the stochastic volatility model to see if there are any improvements to the simulation diagnostics.
+This can cause significant problems in getting unbiased MCMC samples for
+complex models. An example of this is Neal's funnel (Neal 2003) which
+arises in hierarchical models which causes performance issues in the
+Hamiltonian Monte Carlo sampler. Rewriting the model with non centered
+parameterisation significantly improves the performance of MCMC. Another
+consideration for this research is to consider both centered and non
+centered parameterisations of the stochastic volatility model to see if
+there are any improvements to the simulation diagnostics.
 
-The models and parameterisations for this research are summarised in the below table:
+The models and parameterisations for this research are summarised in the
+below table:
 
-![Model matrix](model_table.jpg)
+\begin{figure}[H]
 
-## Plan
-Week 7: Reflect and integrate feedback from presentation into plan for the rest of the semester
+{\centering \includegraphics{model_table.jpg}
 
-Week 8: Finish off any simulation or modelling not yet complete (listed in above matrix)
+}
+
+\caption{Model matrix}
+
+\end{figure}
+
+\hypertarget{plan}{%
+\subsection{Plan}\label{plan}}
+
+Week 7: Reflect and integrate feedback from presentation into plan for
+the rest of the semester
+
+Week 8: Finish off any simulation or modelling not yet complete (listed
+in above matrix)
 
@@
 
-Week 10-11: Action any feedback, prepare for final submission and presentation
+Week 10-11: Action any feedback, prepare for final submission and
+presentation
 
@@
 
-{{< pagebreak >}}
+\newpage{}
 
-# References
+\hypertarget{references}{%
+\section{References}\label{references}}
 
-Abril-Pla O, Andreani V, Carroll C, Dong L, Fonnesbeck CJ, Kochurov M, Kumar R, Lao J, Luhmann CC, Martin OA, Osthege M, Vieira R, Wiecki T, Zinkov R.
-(2023) PyMC: a modern, and comprehensive probabilistic programming framework in Python.
+Abril-Pla O, Andreani V, Carroll C, Dong L, Fonnesbeck CJ, Kochurov M,
+Kumar R, Lao J, Luhmann CC, Martin OA, Osthege M, Vieira R, Wiecki T,
+Zinkov R. (2023) PyMC: a modern, and comprehensive probabilistic
+programming framework in Python.
 
-Duane, Simon; Kennedy, Anthony D.; Pendleton, Brian J.; Roweth, Duncan (1987). "Hybrid Monte Carlo". Physics Letters B. 195 (2): 216–222.
+Duane, Simon; Kennedy, Anthony D.; Pendleton, Brian J.; Roweth, Duncan
+(1987). ``Hybrid Monte Carlo''. Physics Letters B. 195 (2): 216--222.
 
-Hoffman, M. D., & Gelman, A. (2014). The No-U-Turn sampler: adaptively setting path lengths in Hamiltonian Monte Carlo. J. Mach. Learn. Res., 15(1), 1593-1623.
+Hoffman, M. D., \& Gelman, A. (2014). The No-U-Turn sampler: adaptively
+setting path lengths in Hamiltonian Monte Carlo. J. Mach. Learn. Res.,
+15(1), 1593-1623.
 
-Kim, Sangjoon, Neil Shephard, and Siddhartha Chib. 1998. “Stochastic Volatility: Likelihood Inference and Comparison with ARCH Models.” Review of Economic Studies 65: 361–93.
+Kim, Sangjoon, Neil Shephard, and Siddhartha Chib. 1998. ``Stochastic
+Volatility: Likelihood Inference and Comparison with ARCH Models.''
+Review of Economic Studies 65: 361--93.
 
-Neal, R.M. (1995) Bayesian Learning for Neural Networks. Ph.D. Thesis, Graduate Department of Computer Science, University of Toronto, Toronto.
+Neal, R.M. (1995) Bayesian Learning for Neural Networks. Ph.D.~Thesis,
+Graduate Department of Computer Science, University of Toronto, Toronto.
 
-Neal, R. M. (2003). Slice Sampling. The Annals of Statistics, 31(3), 705–741. http://www.jstor.org/stable/3448413
+Neal, R. M. (2003). Slice Sampling. The Annals of Statistics, 31(3),
+705--741. http://www.jstor.org/stable/3448413
 
-Neal, R.M. (2011) MCMC Using Hamiltonian Dynamics. In Handbook of Markov Chain Mone Carlo, CRC Press, New York
+Neal, R.M. (2011) MCMC Using Hamiltonian Dynamics. In Handbook of Markov
+Chain Mone Carlo, CRC Press, New York
 
-Stan Development Team. 2023. Stan Modeling Language Users Guide and Reference Manual, 2.32. https://mc-stan.org
+Stan Development Team. 2023. Stan Modeling Language Users Guide and
+Reference Manual, 2.32. https://mc-stan.org
 
-Talts, S., Betancourt, M., Simpson, D., Vehtari, A., & Gelman, A. (2018). Validating Bayesian inference algorithms with simulation-based calibration. arXiv preprint arXiv:1804.06788.
+Talts, S., Betancourt, M., Simpson, D., Vehtari, A., \& Gelman, A.
+(2018). Validating Bayesian inference algorithms with simulation-based
+calibration. arXiv preprint arXiv:1804.06788.
 
-{{< pagebreak >}}
+\newpage{}
 
-# Appendix A: Mixture Gaussian weights
+\hypertarget{appendix-a-mixture-gaussian-weights}{%
+\section{Appendix A: Mixture Gaussian
+weights}\label{appendix-a-mixture-gaussian-weights}}
 
-| $\omega$ | $Pr(\omega = i)$ | $m_i$ | $\nu^2_i$ |
-|------|------|------|------|
-| 1    | 0.00730    | -10.12999   | 5.79596 | 
-| 2    | 0.10556    | -3.97281    | 2.61369 | 
-| 3    | 0.00002    | -8.56686    | 5.17950 | 
-| 4    | 0.04395    | 2.77786    | 0.16735 | 
-| 5    | 0.34001    | 0.61942    | 0.64009 | 
-| 6    | 0.24566    | 1.79518    | 0.34023 | 
-| 7    | 0.25750    | -1.08819    | 1.26261 |
+\begin{longtable}[]{@{}llll@{}}
+\toprule\noalign{}
+\(\omega\) & \(Pr(\omega = i)\) & \(m_i\) & \(\nu^2_i\) \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+1 & 0.00730 & -10.12999 & 5.79596 \\
+2 & 0.10556 & -3.97281 & 2.61369 \\
+3 & 0.00002 & -8.56686 & 5.17950 \\
+4 & 0.04395 & 2.77786 & 0.16735 \\
+5 & 0.34001 & 0.61942 & 0.64009 \\
+6 & 0.24566 & 1.79518 & 0.34023 \\
+7 & 0.25750 & -1.08819 & 1.26261 \\
+\end{longtable}
 
-{{< pagebreak >}}
+\newpage{}
 
-# Appendix B: Sampling from mixture of Gaussians
+\hypertarget{appendix-b-sampling-from-mixture-of-gaussians}{%
+\section{Appendix B: Sampling from mixture of
+Gaussians}\label{appendix-b-sampling-from-mixture-of-gaussians}}
 
-## Conjugate posterior distributions
+\hypertarget{conjugate-posterior-distributions}{%
+\subsection{Conjugate posterior
+distributions}\label{conjugate-posterior-distributions}}
 
-**Sampling** $\boldsymbol{\sigma_{\eta}^2}$
+\textbf{Sampling} \(\boldsymbol{\sigma_{\eta}^2}\)
 
@@
 
-$$
+\[
 \sigma^2_{\eta} | y,h,\phi,\mu \sim IG \Bigl\{\frac{n+\sigma_r}{2}, \frac{0.05 + (h_1 - \mu)^2 (1 - \phi^2) + \sum_{t=1}^{n-1}((h_{t+1} - \mu) - \phi(h_t - \mu))^2}{2}\Bigr\}
-$$
+\]
 
-**Sampling** $\boldsymbol{\mu}$
+\textbf{Sampling} \(\boldsymbol{\mu}\)
 
@@
 
-$$
+\[
 \mu | h,\phi,\sigma^2_{\eta}  \sim N(\hat{\mu}, \sigma^2_{\mu})
-$$
+\]
 
@@
 \end{aligned}
-$$
+\]
 
-**Sampling** $\boldsymbol{\phi}$
+\textbf{Sampling} \(\boldsymbol{\phi}\)
 
@@
 
-1) Generate proposal $\phi^\ast$ from $N(\hat{\phi}, V_{\phi})$ where $\hat{\phi} = \frac{\sum_{t=1}^{n-1} (h_{t+1} - \mu)(h_t - \mu)}{\sum_{t=1}^{n-1} (h_t - \mu)^2}$ and $V_{\phi} = \sigma^2_{\eta} \{\sum_{t=1}^{n-1} (h_t - \mu)^2\}^{-1}$
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi})}
+\item
+  Generate proposal \(\phi^\ast\) from \(N(\hat{\phi}, V_{\phi})\) where
+  \(\hat{\phi} = \frac{\sum_{t=1}^{n-1} (h_{t+1} - \mu)(h_t - \mu)}{\sum_{t=1}^{n-1} (h_t - \mu)^2}\)
+  and
+  \(V_{\phi} = \sigma^2_{\eta} \{\sum_{t=1}^{n-1} (h_t - \mu)^2\}^{-1}\)
+\item
+  Accept proposal as \(\phi^{(i)}\) with probability
+  \(e^{\{g(\phi^\ast) - g(\phi^{(i-1)}\}}\) such that
+  \(g(\phi) = log (\pi (\phi)) - \frac {(h_t - \mu)^2 (1-\phi^2)}{2 \sigma_{\eta}^2} + \frac{1}{2} log (1-\phi^2)\)
+\end{enumerate}
 
-2) Accept proposal as $\phi^{(i)}$ with probability $e^{\{g(\phi^\ast) - g(\phi^{(i-1)}\}}$ such that $g(\phi) = log (\pi (\phi)) - \frac {(h_t - \mu)^2 (1-\phi^2)}{2 \sigma_{\eta}^2} + \frac{1}{2} log (1-\phi^2)$
+\textbf{Sampling mixture density:}
 
-**Sampling mixture density:**
+Rewrite mixture density with respect to a indicator variable \(s_t\)
 
-Rewrite mixture density with respect to a indicator variable $s_t$
-
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-Sample $s_t$ from probability mass function:
+Sample \(s_t\) from probability mass function:
 
-$$
+\[
 \begin{aligned}
@@
 \end{aligned}
-$$
+\]
 
-**Sampling steps**
+\textbf{Sampling steps}
 
-1. Set initial values $s$, $\phi$, $\sigma^2_{\eta}$ and $\mu$
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Set initial values \(s\), \(\phi\), \(\sigma^2_{\eta}\) and \(\mu\)
+\item
+  Sample \(h\) from \(h|y^{\ast}, s, \phi, \sigma^2_{\eta}, \mu\)
+\item
+  Sample s from \(s|y^{\ast}, h\)
+\item
+  Sample \(\mu|y^{\ast}, s, \phi, \sigma^2_{\eta}, h\) according to
+  conjugate posterior distribution
+\item
+  Sample \(\sigma_{\eta}^2|y^{\ast}, s, \phi, \mu, h\) according to
+  conjugate posterior distribution
+\item
+  Sample \(\phi|y^{\ast}, s, \mu, \sigma^2_{\eta}, h\) using metropolis
+  hastings
+\end{enumerate}
 
-2. Sample $h$ from $h|y^{\ast}, s, \phi, \sigma^2_{\eta}, \mu$
-
-3. Sample s from $s|y^{\ast}, h$
-
-4. Sample $\mu|y^{\ast}, s, \phi, \sigma^2_{\eta}, h$ according to conjugate posterior distribution
-
-4. Sample $\sigma_{\eta}^2|y^{\ast}, s, \phi, \mu, h$ according to conjugate posterior distribution
-
-4. Sample $\phi|y^{\ast}, s, \mu, \sigma^2_{\eta}, h$ using metropolis hastings
+\end{document}
